\begin{description}

\item[Original Concept and Design] Rob Boyle, Brian Cross

\item[Writing and Design] Lars Blumenstein, Rob Boyle, Brian Cross, Jack
  Graham, John Snead

\item[Additional Writing] Bruce Baugh, Randall N. Bills, Davidson Cole, Tobias
  Wolter

\item[Editing] Rob Boyle, Jason Hardy

\item[Development] Rob Boyle

\item[Line Developer] Rob Boyle

\item[Art Direction] Randall N. Bills, Rob Boyle, Brent Evans, Mike
  Vaillancourt

\item[Cover Art] Stephan Martiniere

\item[Interior Art] Justin Albers, Rich Anderson, Davi Blight, Leanne Buckley,
  Robin Chyo, Daniel Clarke, Paul Davies, Nathan Geppert, Zachary Graves, Tariq
  Hassan, Thomas Jung, Sergey Kondratovich, Sean McMurchy, Dug Nation, Ben
  Newman, Justin Oaksford, Efrem Palacios, Sacha-Mikhail Roberts, Silver
  Saaramael, Daniel Stultz, Viktor Titov, Alexandre Tuis, Bruno Werneck, and
  Dr. CM Wong (Opus Artz Studio)

\item[Graphic Design and Layout] Adam Jury, Mike Vaillancourt

\item[Faction Logos] Michaela Eaves, Jack Graham, Hal Mangold, Adam Jury

\item[Indexing] Rita Tatum

\item[Additional Advice and Input] Robert Derie, Adam Jury, Sally Kats,
  Christian Lonsing, Aaron Pavao, Andrew Peregrine, Kelly Ramsey, Malcolm
  Shepard, Marc Szodruch

\item[Science Advice] Brian Graham, Matthew Hare, Ben Hyink, Mike Miller

\item[Playtesting and Proofreading] Chris Adkins, Sean Beeb Laura Bienz, Echo
  Boyle, Berianne Bramman, Chuck Burhanna, C. Byrne, Nathaniel Dean, Joe
  Firrantello, Nik Gianozakos, Sven Gorny, Björn Grammatke, Aaron Grossman,
  Neil Hamre, Matthew Hare, Kristen Hartmann, Ken Horner, Dominique Immora,
  Stephen Jarjoura, Lorien Jasny, Jan-Hendrik Kalusche, Austin Karpola, Robert
  Kyle, Tony Lee, Heather Lozier, Jürgen Mayer, Darlene Morgan, Trey Palmer,
  Matt Phillips, Aaron Pollyea, Melissa Rapp, Jan Rüther, Björn Schmidt,
  Michael Schulz, Brandie Tarvin, Kevin Tyska, Liam Ward, Charles Wilson, Kevin
  Wortman, plus everyone who participated in a game at Gen Con 2008

\item[Musical Inspiration] Geomatic (Blue Beam), Memmaker (How to Enlist in a
  Robot Uprising), Monstrum Sepsis (Movement)
  
\end{description}

\pagebreak

\paragraph{Dedication}

This book is dedicated first and foremost to the people who made Eclipse Phase
happen, from everyone who contributed time, sweat, ideas, and money into it to
everyone who picks it up, reads it, and plays it. This game is by you and for
you.  Secondly it's dedicated to my grandmother and to Andrea, both important
people in my life who died while I was working on this book and its themes of
defeating death. I sincerely hope that one day such tragic losses are
avoided. Third, this book is dedicated to my son Echo, my entertaining working
companion on this project. Finally this book is dedicated to those visionaries,
especially the anarchists and transhumanists, who are working to bring about a
fantastic future, starting now. —Rob Boyle

\paragraph{Our Resources}

\begin{itemize}
\item \url{http://eclipsephase.com}{Eclipse Phase website and blog}
\item \url{http://del.icio.us/infomorph}{Eclipse Phase-related news and links}
\item \url{http://posthumanstudios.com}{Posthuman Studios website}
\item \url{http://catalystgamelabs.com}{Catalyst Game Labs website}
\item \url{http://eclipsephase.com/store}{Online Ordering and PDFs}
\end{itemize}

\paragraph{Outside Resources}
\label{sec:outside-resources}

\begin{itemize}
\item \url{http://www.humanityplus.org}{Humanity Plus}
\end{itemize}

First Printing by Catalyst Game Labs, an imprint of InMediaRes Productions, LLC
PMB 202 - 303 - 91st Ave. NE, G-701 Lake Stevens, WA 98258.  Printed in China

Creative Commons License; Some Rights Reserved.

This work is licensed under the Creative Commons Attribution-
Noncommercial-Share Alike 3.0 Unported License.

To view a copy of this license, visit:
\url{http://creativecommons.org/licenses/by-nc-sa/3.0/}{} or send a letter to:

\begin{center}
  Creative Commons, 171 Second Street, Suite 300, San Francisco, California,
  94105, USA.
\end{center}

(What this means is that you are free to copy, share, and remix the text and
artwork within this book under the following conditions:

\begin{enumerate}
\item you do so only for noncommercial purposes;
\item you attribute Posthuman Studios;
\item you license any derivatives under the same license. 
\end{enumerate}

For specific details, appropriate credits, and updates/changes to this license,
please see:

\begin{center}
  \url{http://eclipsephase.com/cclicense}{})
\end{center}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ep"
%%% End: 
