\chapter{Lack}
\label{chap:lack}

``What's the number?''

The word dig their claws into my new vocal cords and yank themselves up and out of my parched throat. My diction is predictably poor, as it always is during the first few minutes following a resleeve. The pitch of the voice is apparent despite the mumbled, sandpaper slur of the words. Definitely a biomorph and my latest sex is female. This much I know in the first few seconds. The model eludes me right now, but I'll know for certain soon enough, once motion is mine again. Another fury morph is my first guess.

The slab is hard. Nothing more than frigid metal with crisp white synth-slick wrapped around it. Typical accommodations for a corp dollhouse. The chill seeps through my skin and wraps around my bones.

A cortical cruncher looms over me, waving his welcome-back beam side to side, occasionally catching my pupils. His smug, bored face speaks: ``Consciousness confirmed.'' The beam clicks off. My question should have made it obvious that I am back, but the guy is a slave to procedure. They all are. Corporate body banks like their employees paralyzed by obedience, unable to think for themselves. I mumble the question again. ``What's the number?''

``March 11.''

``How many after the Fall?''

``Are you for real?''

\begin{quotation} I am paranoid. Yes. I need to know the year every time I return upon a body bank slab. Paranoia is just one of the plagues transhumanity faces these days. \end{quotation}

I try to snag the data from my new sleeve's mesh inserts before opening my mouth again. No luck. Asking a sleeve tech for the year is always humiliating. Makes me feel like an amateur, but the circumstances are definitely extenuating, so I press him. Hard.

``Answer the fucking question.''

The corp sloth gives me the lunatic eye before he answers.

``Uh ... 10 AF. You haven't been gone that long. Your last backup ... ''

He scans his entoptics for the info.

``... 14 days, 7 hours ago.''

It takes a second to sink in, but when it does, it stings. It never ceases to shock when time slips away from me. Two weeks. Gone. Completely wiped from my existence. Two weeks ago, there was another me, sleeved in another morph. There was a mission and it led to my death. That is all I know. Either Firewall failed to retrieve the cortical stack off the corpse so I could retain those two weeks, or the fuckers deliberately chose to swipe that time from me. Honestly, though, both possibilities are preferable to having another self bouncing around out there, doing who knows what-the-fuck. Some t-humies get off on having multiple selves traipsing all over, but my ego is in check. One Sava is enough misery to unleash on the ‘verse.

Shit. My brain is wandering off into morose territory; always does during the first moments following a resleeve. I need a physical context. Something tangible to focus my attention on. I bring my hands in front of my eyes, arms feeling like two-ton sacks of rocks. The fingers are thin and long; the knuckles callused, scarred and misshapen. Obviously the work of many thrown punches, fists connecting with jaws, metal, flesh. Yep. A well-worn fury morph. You get what you pay for, I suppose; or what Firewall is willing to pay for. Why do I do it? As far as the org is concerned, I'm nothing more than a cheap precision instrument, tossed into the recycling bin when I snap in half. There will always be more of me, until the horrors prove too intense, until the files get too corrupt, until I know too much and Firewall decides to wipe me, then some other sap will step in to preserve transhumanity. Preserve transhumanity. Fantastic. Now I'm babbling like a Firewall propaganda tweet! My arms weaken and flop back to my sides. The strength just isn't there yet. A few more minutes with nothing but my thoughts.

On his way out of the recovery bay, the cort cruncher laughs at my feeble attempt at motion. ``What's your hurry?'' he says. ``Just relax, willya? You collapse onto the floor, you're gonna stay there until you get yourself up. They don't pay me enough to babysit newbies.'' His flippancy doesn't help my mood and the melancholy returns.

What experiences are no longer a part of my consciousness? Perhaps the thrill of a lifetime. Did I discover true beauty? Fall in love? Have an epiphany? Save a life? I'll never know. Those memories, that life, that version of me, is gone. The new me, lying on this slab, was never shaped by those experiences. My chest hollows out from the weight of the loss.

I gotta shift the thought process.

Fuck it. Maybe there was no joy, no revelations. It was a shit two weeks. I'm certain of it. I was bored out of my mind. Better yet, broken and suffering from an epic heartache. My demise was meaningless. I OD'd on kick, flopping on the floor in a pathetic speed-addled frenzy till my heart exploded. I was gutted by some low-life scumborn in a back station corridor over some lo-rez black market XP. I'm glad the time was wiped. Ecstatic, in fact. Fuck it. Fuck them. I don't need those two weeks.

But these thoughts are lies. I need those two weeks. I don't feel whole without them. Hell, I feel incomplete if even an hour is sacrificed. I have to know.

Someone knows what occurred. No doubt. Probably a Firewall proxy, Jesper most likely. He was my connect this go round. I remember that much. The wipe would have been his call. And proxies have a quick trigger finger when it comes to wiping us sentinels. Even a hard-earned rep score can't save my memories when Firewall deems the results of a mission too sensitive for an outer circle thug like myself to possess. As long as the fucking job gets done. As long as transhumanity perseveres.

What a shit deal.

How did my life, my lives, come to this? Always in the hands of another.

Again, the dread, the paranoia. I gotta shake this off. I have to give the org the benefit of the doubt. I've been a sentinel for decades. I like to think I've saved millions of lives, but I'm just not sure.

Do I trust the org? No. But there is an understanding, a degree of respect. Though as the years continue to race by, and the gaps grow longer and more frequent, I'm beginning to doubt Firewall's commitment to my preservation.

Suddenly, my muse stirs, breaking my dark reverie. Several entoptic displays appear in my field of vision, cycling through diagnostic routines as my mesh inserts finally come online. Careza's familiar feminine voice enters my mind.

[Welcome back, Sava.]

The sound is soothing; like being cradled by my mother, or embraced by a lover. The harmonic upgrade was a worthwhile investment. Careza has learned to use it well. I rarely think of my muse as an AI. It is my only true friend these days. I wonder if it shares the sentiment. I've never shot the thought its way. I keep it to myself. I'm afraid of what the response might be.

Hey Careza. Glad to be back.

[You could use a drink, I suspect.]

You know me too well, Car. Better than I know myself.

[Hospitality now has the request. Wait time, approximately ten minutes.]

Thanks. Careza enjoys our conversations when my brain has a slight buzz. It is always trying to get me drunk.

[You're welcome, Sava. Before you ask, it's been two weeks. I don't have any information on what happened following our last resleeve. Currently, we are in lunar orbit aboard Selardi IV. We are outfitted in a CoreCorp-brand fury morph with minor enhancements. They will be online shortly. I am pleased to report the Titanians were victorious and won the Cup.]

Damnit. Would have made a killing on that one. What did the odds go off at? But before Careza can dig up the info, I shut down the operation. Wait. No. I don't want to know. It'll only irritate me more. A nervous energy starts to itch my entire system and a thick familiar taste begins to coat my tongue. I need a cigarette.

[Yes. I know. The previous occupant of this morph was a heavy smoker. The habit might be difficult to shake this time.]

This resleeve just keeps getting better by the minute. I hate smoking. Booze, fine. I can handle my alcohol, but smoking always makes me feel like shit. Every time I get sleeved in a morph with the addiction, I struggle to kick it. Careza continues with her report as I try to retain my sanity in the face of an intense nicotine craving.

[@-rep remains intact.]

Finally some good news. At least I didn't piss off any allies in the past 2 weeks.

[Indeed. Are you in the right frame of mind for an update on Rati?]

Rati is my passion. The lover I hold above all others. She disappeared on me two years ago. No explanation. The sting still lingers.

Let's skip the update for now, Careza.

[Understood.]

Run a newsfeed scan. Check for any major incidents in the past 2 weeks. Maybe there's a clue as to what we may have been up to.

As Careza runs the scan and continues her standard sitrep, I shift my attention to the new sleeve. The strength to stand is finally there. I push the morph up and swing the feet onto the floor. Spasms shoot through every muscle. New morphs always take a bit of time in which to acclimate. Luckily, I'm familiar with the CoreCorp fury, sleeved it a few times in the past. This one feels like an old pair of shoes, bit worn and abused, but able to pound the pavement if need be. The left ankle is a bit tender. I hold it up a bit to get a look. Bit swollen. Definitely not new sleeve dysmorphia. Probably a nagging injury. Again, a pain in the ass, but you get what you pay for, I suppose. The nanotat encircling the right bicep is rude and obnoxious, even by scum standards—an entire slitheroid entering the genitals of a female pleasure pod, fully animated. Class act, whoever opted to etch that upon the morph. I hate identifying marks, but again, if you can't afford a clean morph, you take what you can get.

I slide off the table, managing not to fall over in the process, and gingerly test the ankle. Sore, but it isn't going to snap off.

Put in a request for a patch, left ankle. Bute should be fine.

[Phenylbutazone. On it's way. And the cocktail will be here in approximately 30 seconds. Nothing unusual on the newsfeed scan.]

Figures.

\begin{quotation}
I plod over to the full-length mirror, standard issue in resleeve waking chambers, and drop the sheet to take a look at the new me. I spy the cortical cruncher lingering in the doorway, my cocktail in his hand, giving my body an appreciative look. I don't recognize myself.
\end{quotation}

``Hand me my drink please.'' I reach out my hand in his direction without even acknowledging his presence. He steps into the room, too close to me, and slips the drink into my hand. His breath smells like some sort of sour sausage.

``Not too bad under the sheet, are you?'' he says. ``I took a peek earlier, but I must say, the slab didn't do you justice. On your feet, the curves really pop. Your face isn't much to look at, but that rack is ... ''

I cut him off before I vomit bile into my mouth. ``It's exquisite. I know. Now shut up and back off before I rip the skin off your face and slap you silly with it.'' He gets the message and slinks from the room.

It is a nice rack.

[If nice is defined by proportion, then I would say yes.]

AIs, always so formal.

[You're approximately 4 centimeters taller than your usual proprieception allows for, so watch your head.]

Thanks for the heads up.

[That was awful.]

Yeah. Yeah. I know. A smile finds its way onto my face as the banter with my muse lightens my mood. Looking in the mirror, I try to broaden the smile, to get a better sense of my new face. I show some teeth. Nicotine stains all over them. I take a long sip from my cocktail, swish the alcohol around a bit. I can feel my blood respond instantly to the sauce. I close my eyes and let out a sigh. Just a few moments of peace is all I ask.

[We have a guest, Sava.] Damnit. No such luck.

Who?

[Our last Firewall proxy, Jesper, has sent a beta-level fork of himself. It is rather impatient to speak with you.]

Connect him.

They just cannot leave me alone, can they? Officially, Firewall doesn't even exist. It's because of Rati that they got their tentacles wrapped all around me, through me. The whole mess on Mars. That's where it all started. The last time I saw Rati. All that knowledge they allowed me to retain. But why? Until that day, I had never realized just how scary the universe truly was. No, not scary. Horrific. No other word for something so vast, so uncaring. Transhumanity could be wiped out completely and it would all just continue on as before. Horrific. No other way to explain the feeling you have when you come face to face with things truly beyond comprehension. Hell, no other term could encapsulate transhumanity's actions towards each other—much less what other beings lurking in the void have in store for us. Perhaps that was why. To teach me a lesson. To make certain I would never forget, so I would never cease assisting the org, because even the briefest glimpse of what is actually out there is enough.

Jesper's fork materializes in my field of vision.

[Welcome back, Sava.]

Fuck off, Jesper. You know I hate waking up with lack.

[Sorry. Nothing I could do.] His expression is serious and concerned, but his kinesics indicate he is as calm as can be. What an act! Fucking proxies never panic. They hold all the cards and it's never their minds that are on the line.

Yeah. Right. Get to the point. You don't have me sleeved in a combat morph to get some downtime, so you must have something serious lined up. Are Berk, Pivo, and Sarlo here?

[Yes, they have been resleeved in the same facility.]

At least my team is with me. People I could count on. To a certain degree.

All right. What are the details?

\begin{center}
***
\end{center}

Pivo gripped the smooth outer surface of the station with all eight arms. Nano-magnetics at the tips of his vacsuit arms were the only difference between a secure hold and an endless drift into the depths of space. He peered up through his faceplate at the dark orb above him.

Earth.

His eyes locked on an expanse of dead black ocean through the ominous clouds. Pivo longed to swim in those ancient depths. Born and bred for space, he had never once immersed himself in the former ecological niche of his kind. Odds were against his ever taking a plunge into the salty waters of an Earth ocean. The planet was now a plagued death trap. A wasteland of skeletal forms.

He imagined a time before the Fall, when his ancestors thrust through blue waters and slipped effortlessly through mazes of coral, or gently floated along with the current, not bothered by the burden of sapience. Perhaps octopi still survived beneath the black waters of the present, eking out a brief existence, biding their time, keeping the species true and alive until the Earth could be reclaimed, and Pivo would join them on that glorious day, abandoning knowledge altogether, and returning to the ways of instinct.

Vacsuit sensors interrupted Pivo's fantasy, detecting a laser light that bathed his form—contact from Sava by line-of-sight laser link. It was the preferred method of communication when a mission required discretion. Pivo's muse processed the message, and Sava's voice entered his head.

[Something wrong? Why'd you stop moving?]

[Just enjoying the view,] Pivo beamed back.

[Enjoy it on the way down, for hours if you want. Get inside the station before one of the sentry bots finds us.]

Pivo didn't bother to respond. There was no arguing with Sava. No use in defending your actions. Pivo began crawling along the shell of the station again. The station itself was tethered to the end of a long, black, carbon nanotube cable that stretched all the way down to the planet's surface—the sole surviving space elevator.

Pivo located the breach, a thin scar in the station's metallic hull, the result of an internal explosion responsible for the station's demise during the Fall. The breach was exactly where Sava said it would be and the description of its size was dead-on: a gap barely large enough for a human infant to slip through. According to Sava, years ago, the self-repairing nanosystems operating in the hull's metal had malfunctioned before the breach had been fully repaired. The level of mission details Sava managed to extract from Firewall was scary sometimes. Paranoia bloomed for a moment, but he quickly dismissed his suspicion, compressed his cephalopod form, and squeezed his body through the breach.

In blackness, Pivo activated his infrared emitter, casting the room in a light outside the normal visual spectrum. The interior of the lifeless station became visible to his enhanced eyes in the eerie altered colors of infrared. Pivo almost preferred the dark. Ice crystals glittered from every surface, the reult of flash frozen moisture in the long-absent atmosphere. Frigid clumps of human remains floated alongside chunks of hull metal in a macabre zero-gravity ballet. Pivo floated through the wreckage and the gore, lightly tapping aside metal or flesh to clear a path deeper into the room. A female head drifted slowly by, the face frozen in a gaping silent scream. An intact cortical stack dangled from the severed neck. For a second, Pivo considered snatching the stack, but he was not here to retrieve lost souls. Instead, he placed two of his arms upon the top of the head and pushed it beneath him, towards the floor. Like so many others lost during the Fall, this person would remain forgotten here.

Pivo made it to the airlock without incident, but he knew his luck would run dry eventually. A run-in with hypercorp guardians on a derelict station was unavoidable. Sensors may have already detected his presence. It was only a matter of time before bots converged on his position. He just hoped that when it occurred (and it most certainly would), it would happen after he had opened up the airlock and the rest of the team was inside the station.

The airlock had been welded shut from the inside. Pivo was prepared for this eventuality, but it made his detection by guardian bots a certainty. He composed himself for a few seconds, focused on the task at hand, then fired up the plasma torch built into one of his vacsuit arms. A sizzling hiss and a harsh hot blue glare filled the room. Seconds were now his most precious possession.

He was almost through the inner door when his muse pinged him with a warning from the passive teraherz sensor. An object was moving towards Pivo's position rapidly, now only twenty meters away. A sentry bot would be upon him in soon.

[Almost through the first door,] Pivo transmitted calmly, even though it took every ounce of his will to keep the torch steady. [I have company. Be ready.]

[Copy that,] Sava replied.

Finally, Pivo cut through the seal. The octomorph slithered four arms through the still smoldering sliced metal, and with a strained yank, pulled the door from the frame. The door slowly floated away into the chamber, the edges rapidly cooling. The interior airlock door was not welded shut. With a vocal sigh of relief, all eight of Pivo's arms began a frenzied assault upon the airlock door's manual controls.

[Few more seconds. Just a few more seconds.] But the seconds had expired.

In his 360-degree field of vision, Pivo could see the security bot thrust into view behind him. The bot unloaded its weapons immediately, the shots ricocheting off the floating airlock door. The bot advanced on the door, and with a furious swat knocked the obstruction aside. It clanged upon the crystalline surface of the wall. Just as Pivo pulled the last lever to release the airlock door, blazing plasma fire engulfed him.

\begin{center} *** \end{center}

Sava had instructed Careza to surge the neurochem the instant the airlock portal was open. The muse did not fail to deliver. In what seemed like an eternal slow-mo to Sava's charged brain, the airlock door swung open into the station, aided by a thudding steel leg kick courtesy of Berk, the team's muscle. With a flash of thought, Sava's targeting radar snapped up an entoptic display and locked on two targets: Pivo and a sentry bot. The robotic guard dog was already leveling its weapons, but Sava was faster. Retinal-searing plasma fire erupted from Sava's weapon, singeing one of Pivo's arms and slamming the sentry back. A second shot punched through the bot's armored carapace, melting critical components within, rendering the bot a useless pile of fused scrap metal.

Sava moved quickly past the cursing octomorph and unloaded two more shots into the smoking bot.

[We're clear,] Sava transmitted. [One down, but there is always more. Count on it. Pivo, you shiny?]

[You scorched my breeding arm, puta.] Pivo shot back with clear agitation rumbling in the harmonics.

[You rather I leave you to the bot next time?] Sava turned to Sarlo. [Sarlo, get in here and find the console you need. Berk, we're going to need to set up defensive positions, to give hacker boy here time to crunch his bits.]

Pivo cut through his vacsuit and detached his damaged arm, cursing Sava under his breath as the vacsuit rapidly repaired itself and sealed the gap.

[Hey. Don't worry, Pivo. You've got seven more. And besides, you don't really strike me as the breeding type anyway.] Sava relished giving Pivo a hard time. It was one of the true joys in life.

Pushing off from one wall to the next, Sarlo moved along the chamber with ease and grace. His neotenic morph was slighter and even more diminutive than the average human child sleeve, completely augmented and customized to match his ``preferences.'' He had paid a fortune for it. The others never understood Sarlo's penchant for juvenile human sleeves, so much so that he always kicked in his own credits to ensure an augmented neotenic resleeve, even when Firewall was footing the bill. They also didn't know where his seemingly endless supply of personal funds came from, nor did they want to. As long as he got the job done.

Two minidrones followed after Sarlo, lighting the area in infrared and actively scanning on other wavelengths. [This way,] he said, transmitting an entoptic map to each team member's overlay. [It's not far, a hundred meters or so.] A highlighted route appeared on the map.

Sava and Pivo followed closely behind Sarlo, while Berk struggled to keep pace in her armored gynoid shell.

[Keep up, flatlander. We'll be down the gravity well soon enough,] Sava beamed to Berk.

[Not soon enough for me,] Berk replied.

The abandoned station was eerily quiet. Signs of long-forgotten violence and desperation lingered everywhere. Floating debris. Ruptured and frozen bodies. Scorch marks and twisted metal. Death owned this place.

When the team reached the control station, Sava and Berk took up defensive positions in the corridor while Sarlo and Pivo went to work on the station's dormant systems.

[I'll be damned! The mission spec was actually right. The station systems are active but dormant. Whomever's guarding this place didn't wreck the systems, they left open the possibility that the space elevator could be activated again.] Sarlo gleefully began his procedures to hack the system.

[Who the fuck would want to risk going down to that ball of ash?] Berk piped in.

Pivo waved one of his arms in agitation. [Need I remind you that some of us happen to think that reclaiming our home planet is a good idea?]

[Reactionary thinking, if you ask me.] Berk replied. [Shrugging off all of our old nation-state loyalties is one of the best steps transhumanity has ever made. Leave reveling in the glories of the past to the bio-cons. I'll take a future where we step boldly outward into space, thank you much.]

[Let's cut the politics.] Sava pointed at Berk. [You're an anarchist, I get it.] Then Sava pointed at Pivo. [And you're on a reclamation kick. Fine.] But Sava's rant was interrupted by half a dozen fast moving dots upon the team's entoptic radars. [Incoming pings. Sarlo, you in yet?]

[Working on it. Fuck. Fuck. Shit.] Sarlo's childish voice sounded petulant.

[Work faster. If these bots have heavy ordnance, we're screwed.] Sava and Berk both unloaded suppression fire down their respective stretches of corridor before the bots even made it to the corners. The bots halted their approach momentarily, taking cover just around the bend. More bots began to appear on the radar, moving towards the position of the first responders.

[We're running out of time, Sar! More bots gathering!] Sava unloaded another round of suppression at the bend. Berk kept her weapon quiet, waiting for a bot to make a move into the corridor before lighting it up, but the bots remained put. More gathered, and even more appeared on radar, moving to the same position.

[They're gonna be all over us any second now!]

[Consider this a gift, ladies and gents ... ] And with a final operation, Sarlo seized control of the station's entire security system.

Suddenly, one of the bots turned on the others. Another soon joined it. In a matter of seconds, fumes and debris came drifting down the corridor as all-out warfare broke out between the bots. Sava and Berk lowered their weapons and admired the sounds of Sarlo's handiwork.

[Damn, Sar! I guess that is why you are one of the best hackers in the system!]

[Applause, applause, ya waify freak!]

[When you've got cutting edge-exploits courtesy of the leet coding AGIs on Extropia, there's not a whole lot you can't do.] Sarlo delivered the line with a calm harmonic, but Sava was watching his kinesics, and they were off the charts. The neotenic's little heart was beating like a drum roll. Sava opted not to bust his furless balls about it, and instead let Sarlo have his moment in the sun. This had been a ``close one,'' and another close one might not end up in their favor.

Sava allowed a few seconds of relieved silence before getting the team back to business. [Sarlo. How soon till the elevator is active?]

\begin{center} *** \end{center}

Pivo stuck to the portal, watching as they descended below the soot-filled layer of clouds and the Earth below came into view. They were in the atmosphere now, descending on a taut beanstalk stretched between the Earth and station above, a massive feat of engineering built from carbon nanotubes. The shuttle car crawled down the elevator cable, bringing them closer and closer to the ruined planet.

Earth's atmosphere was now choked with a thick dust, the color of rust. The winds whipped over the planet's surface with breakneck velocity, swirling dangerously in certain pockets. The world's weather systems had been irretrievably ravaged by the Fall, when transhumanity had seemingly gone to war with a group of rogue AIs known as the TITANs. Bombs, raging fires, chemical attacks, biowar plagues, voracious nanoswarms—even nukes—had taken their toll. It was now an inhospitable place, gripped by nuclear winter. Some of the clouds were formed into unusual shapes, defying the high winds, even seeming to writhe as they moved—the thriving descendents of self-replicating airborne nanoswarms, Pivo suspected. Who knew what other monstrosities waited for them below, evolved from the remnants of AI war machines?

The Earth was off-limits now. Abandoned to the enemy. Though the TITANs were presumed to be long gone, escaping the solar system via secretly-constructed wormholes gates, taking millions of forcibly uploaded transhuman minds with them—they had left many of their tools and weapons behind. Likewise, some of the weapons transhumanity had unleashed on the AIs—and, quite often, themselves—had taken on a life of their own. So Earth had been abandoned and interdicted, with hypercorp killsats laced into orbit to shoot down anything that attempted to leave or land on the planet's surface.

As a reclaimer, Pivo was part of a small but vocal faction that advocated a return to Earth. There was still hope for the planet, they believed. It had always perservered, and this was no time to give up on it. Earth needed to be cleansed and terraformed, resuscitating transhumanity's home. But the reclamers were a minority. To most survivors of the Fall, the Earth held too many horrible memories. Lives ruined. Loved ones lost. Their own deaths. It was a monument to transhumanity's arrogance and mistakes, a grim reminder that they were not above destroying themselves despite all of their advances and technology, or perhaps because of them.

This didn't prevent some from trying, of course. Scavengers still raided the planet's ruins, retrieving long-lost treasures, cultural artifacts, or even the preserved mind-states of those who failed to escape. Some reclaimers had initiated their own secret missions, intending to establish a basecamp from which they would begin operating their own reclamation projects. Most were never heard from again.

The team of four rested and prepared equipment in the shuttle's large open lounge, Sava and Sarlo in a cramped inflatable survival bubble so the biomorphs could escape the confines of their vacsuits for a while. Pivo elected to remain outside the bubble and in the vacsuit. Close confines with Sava during the descent did not sound pleasant to him. The walls of the lounge were smeared with decadesold blood, now frozen into a crystalline brown in the depressurized cabin. Whoever the last passengers were to ride this shuttle, fleeing the doomed Earth, must have set violently upon each other, fueled by madness or despair.

[I wonder what it was like.] Sarlo tossed the thought out to the group.

[What?] Pivo replied.

Sava quickly jumped in and put an end to the discussion Sarlo was yearning to start. [Quit with the philosophizing and the dramatizing. You know I cannot stand that shit.] Sava tried desperately to maintain order and an air of gruff detachment. It was too easy to let the brain wander off into the past and the fate of the millions who perished during the Fall. To counter this, Sava always resorted to the diatribe. [Listen. We all know the mission specs. We're locating someone. A courier. Most likely a corpse. Last known position while alive was the base station we will drop into when this ride stops. Mount Kilimanjaro. Which, according to quite reliable sources, was once overrun by killbots, which are most likely still in the vicinity.] Sava paused for dramatic effect before continuing. [We retrieve something from the courier. What, we don't fucking know. Only that it is quite valuable to the org. We stick to what we know. I don't want to hear any more bullshit ``what ifs'' and ``I wonders.'' If your thoughts are anywhere other than the mission, keep them to yourself. I don't want to hear them.] And with that declaration, the rest of the journey to the Kilimanjaro station was in silence, each confined to their own thoughts, not a single ping between them.

\begin{center} *** \end{center}

The shuttle rattled to a stop inside the dark cavernous hangar. At one time, the Kilimanjaro hangar was the busiest Earth-to-space station port in the world, servicing millions of customers annually. Now, as Pivo clung to a shuttle window and stared out into the black emptiness of the hangar, it seemed as if the place was a soulless vacuum.

[Ready when you are.] Sarlo pinged Sava, poised to hack open the shuttle door and allow the stale dust-choked air of Earth to waft over the team. Sava nodded to Sarlo and the shuttle door slid open with a rush of decompression. A blinding red-gray dust blasted into the shuttle from the hangar and coated the shuttle interior almost immediately.

Sava's first step into the Kilimanjaro hangar landed firmly onto the brittle ribcage of a child's skeleton. The bones snapped into splinters and powder with a crunch. The floor surrounding the shuttle airlock was carpeted with skeletons entangled in a mass of tattered clothing. There was no way to avoid stepping upon them. One by one, the others stepped from the airlock.

[This place is a tomb,] Berk beamed to the group.

[This whole planet is a tomb,] Sava replied, with an extra echo harmonic allowing the word tomb to continue on well after the phrase was transmitted, added specifically to annoy Pivo, who immediately shut down the echo in his head with a countermeasure from his muse.

Sava took a few more crunching steps forward, then stopped. The rest of the team followed suit.

[Something is not right here.] Sava kicked at one of the skeletons. The bones rattled and cracked. [I don't see any skulls.]

[Forced uploading,] Sarlo transmitted. [TITAN machines harvested the heads of the dead for scanning.] He shrugged. [That's my guess, anyway.]

[Shut up!] Sava signaled the team to silence. [Who else hears that?]

A low mechanical whir reverberated nearby. [I'm picking it up.] Pivo replied. [Up a bit to the north. About 30 meters.] As if in response to Pivo's observation, another whir began, this one behind the team, from the south end of the hangar. Another whir from the east joined in the chorus. The sounds were coming closer, becoming more distinct, more aggressive.

[No visual, yet. This fucking place is so deep and thick with this dust shit, seems to act like chaff too. Infrared is giving me only about twenty feet!] Sava motioned for the team to move to the right. [Stay close, we move slow and keep the triggers itchy. The passenger lounges are just east of us. We start the search there.] The whirs were now all around them, hovering just outside visual range.

[What the fuck is that?] A flying insectoid bot with six articulated arms ending in small buzzsaws lunged from the dusty darkness at Berk, who dropped to the floor and unleashed plasma fire into it. The bot slammed into a pile of bones and rags and set it alight. The fire spread quickly, leaping from dry cloth to dry cloth. The blazing hangar floor now illuminated the area in the hot orange glow of flame. At least a dozen insectoid bots hovered in a perimeter around the team, awaiting an opportunity to strike. Another bot dove at Berk, its buzzsaw arms slashing wildly. Berk fired, but missed. The bot slammed into Berk's head and the buzzsaws ground into her neck. Sparks flew in all directions as metal met metal. She dropped her rifle and pushed against the body of the bot till the saws were off her neck. [Fucking run you idiots! I've got this!]

Sava fired and dropped a bot, then dashed east, leaping over spreading waist-high flames. [Make for the lounge!]

Pivo elevated onto two arms and ran behind Sava, his five remaining arms flopping wildly above his head. [Out of the way, ya poke!] Sarlo outpaced the slower octomorph, running through the flames towards the lounge.

Berk flung the frenzied bot into a flaming pile of bones, scrambled to her feet, and followed after the group, covered in bone bits and dust, the bot swarm in whirring pursuit.

Sava reached the lounge first and the portal was open. Turning with rifle raised, Sava took cover against the door frame. Sarlo and Pivo were past the flames and Berk was closing the gap, as were the bots. Sava unloaded cover fire that sizzled over Sarlo's head, knocking another bot out, but the rest of the swarm remained unphased. They just kept coming. Suddenly, more bots appeared out of the shrinking darkness near the lounge.

[There's more! They're flanking!] Sava blasted at the new bots to try and slow down their gambit. Sarlo was only thirty feet from the portal when he tripped on a tangle of bones. His boyish body collapsed face first into the dust and human remains. Pivo made an awkward leap over him, skidded across the floor, and squished into the outer lounge wall right near the door. Sava reached out, snagged the octomorph by a arm, and dragged him into the safety of the lounge. Berk tried to stop and help Sarlo up, but her momentum was too much and her footing upon the dusty floor too unstable. She tumbled forward in a roll of dust cloud, chipped bone, and tattered rags, finally slamming into Sava in the doorway.

The three team members within the lounge gathered themselves just in time to witness a bot latch onto Sarlo's head from above as he stood up. The machine stretched two arms out to the side, then plunged their spinning blades into Sarlo's neck. Sarlo's eyes went wide and his body tensed as the saw blades ground through flesh and bone, working through his neck in seconds. The instant his head was severed from the torso, the bot swooped around and zipped off over the flames, into the dark oblivion of the far end of the hangar.

Sarlo's headless body wavered for a second, then collapsed, spurting blood in long, lazy arcs.

\begin{center} *** \end{center}

Pivo, Sava, and Berk sat in silence. They had managed to seal the portal into the lounge, locking out the horrors of the hangar. The headhunter bots could still be heard hovering outside the portal, occasionally clanging and grinding their blades against the sealed door.

Berk finally broke the silence. [I'm trying very hard not to think about what they're going to do with him.]

[Try harder. Sarlo knew the odds of survival were slim when he signed on. We all did.] Sava stood up.

[Should we tell him? When he resleeves?] Pivo knew this was going to set Sava off, but he blurted it anyway.

[Would that be kindness or cruelty, Pivo? And besides, there is no guarantee that any of us will survive. So who gives a shit? Whenever your last backup was, I sure hope you're not gonna miss anything since. Let's get moving.]

\begin{center} *** \end{center}

With Sarlo gone, Pivo took over the navigation duties. They were nearing the corporate VIP lounge, the last known location of the courier.

The team moved through dark corridors filled with headless skeletons and mummified remains. Years ago, the corporate forces defending the structure had been overrun by AI war machines, which mercilessly slaughtered everyone inside. The walls were scarred from battle, covered in dried blood. Destroyed remnants of the AI war machines littered the halls as well, haunting monuments to the few victories humanity had in their losing battle. Even as piles of scrap, the machines had a menacing presence.

[Too bad this isn't a salvage op,] Berk commented. [The autonomists could use a look at this tech. At the very least, figure out what the hypercorps might try to do with it.]

As they entered a long concourse, the remains and debris abruptly disappeared, as if cleared out.

[I'm getting some strange thermal readings here. Patterns that don't make sense,] transmitted Pivo.

[What is that supposed to mean?] Sava beamed back.

Before Pivo could give thought to ``I don't know,'' his muse issued a chilling warning: [My nanosensors register the presence of unknown nanobots in large numbers of a highly sophisticated design, suggesting a TITAN manufacture. Countermeasures have been initiated.]

[Nanoswarm. Move! Move!] Pivo broadcasted in a panic as he launched into a full two-armed sprint. Sava and Berk followed Pivo's lead without question. They all knew the dangers of a TITAN nanoswarm. Unlike the nanobots Pivo often made, which were manufactured with particular purposes in mind, and which were neither self-sustaining or intelligent, this particular nanoswarm was autonomous, self-replicating, adaptive, and capable of making almost anything it needed. Even as they fled, individual nanosensors were measuring up the three agents, transmitting details on their morphs and gear to the rest of the swarm.

A junction came into view ahead, the pathway narrowing into a smaller tunnel. Suddenly, Pivo stopped, just a meter before the tunnel. [Do not move forward!] The others crashed to a halt.

[What the fuck Pivo!?] Sava looked back down the hall. [Fucking swarm could be finishing us as we speak!]

[My muse picked up a burst of thermal energy here. The swarm is up to something,] Pivo warned.

[But there's nothing here,] Berk replied, as she waved her hand across the tunnel entrance. Her metal hand suddenly clanged to the floor, separated from her wrist.

[Monomolecular wire.] Even though the situation grew more dire by the minute, Pivo was impressed and fascinated with the inventiveness of the alien nanotech. [It laced the door with it. Cuts through anything. Weak tensile strength though—you probably snapped it.]

[We're fucked. Let's face it.] Berk picked her severed hand off the floor. Down the hall, the nanoswarm began to take a visible shape as the nanobots accreted. The swarm was congealing into a fog, creeping closer. Berk continued, [The entirety of this port is probably filled with this shit. I'm useless at this point. These things are already all over my systems, my diagnostics are going crazy.]

[So what are you saying, Berk? You done?] Sava transmitted.

[Yeah. I'm done.] Berk shook her head in disgust. [Who knows what these little bastards have infected me with. I don't want to risk it. I'd rather resort to a clean back up. Forget this shit ever happened. You keep running if you want. I'll try to buy you some time.] Berk turned and ran directly into the fog. The nanoswarm sucked in around her immediately and the disassembly began. Berk's metal frame began to dissolve as she ran further and further away from Pivo and Sava, leaving a wispy trail of nanoswarm behind her.

[Get fucking moving fools! This isn't for my amusement! I'll see ya the next time around.] A few minutes later, Berk's signal went dead.

\begin{center} *** \end{center}

Sava and Pivo entered the VIP lounge. When the spaceport was overrun so many years ago, this was the site of the humans' last stand. Piles of security personnel skeletons littered the floor just inside the doorway. The charred remnants of a hopeless barricade were scattered beside the mounds of bone. Skeletons draped in torn singed civilian garb were clustered around the walls and corners, sometimes three or four deep, as if they had all scrambled as far as possible from some avatar of death in the middle of the room.

Pivo started an operation to locate the RFID tag the courier was supposedly chipped with in his left shoulder blade. The code triggered a ping within three meters. Pivo pointed a lengthy arm at a small bone pile. [He's in there somewhere.]

Sava stepped over to the pile of three skeletons and began rummaging through the bones, yanking out or snapping off all the femurs. [Goddamnit I want a cigarette. This morph has me so tweaked. Haven't I made it clear I don't smoke? Yet, every time, they sleeve me in a morph nailed with the habit.] Sava handed the bundle of bones to Pivo.

[Must be a fury thing. Should just take a few minutes to scan these for the nanoscale etching.] Pivo got to work. [Enough time for a smoke, if you want.]

[Yeah. Real funny. How about I grind you up into dust and smoke you?] Sava sat down on the floor as Pivo sent out a chuckle.

The deceased courier, whomever he was, had been entrusted with information too sensitive to transmit. No one knew the true capabilities of the TITANs to intercept and decode, so the courier had been injected with nanobots that etched a nanoscopic encoded message directly onto one of his femur bones. However, he had never made it off the planet. His message had never been delivered.

Pivo and Sava had no idea what the information was, but someone at Firewall obviously deemed it worthy of capture. Information on the TITANs perhaps. Or some CEO's secret family recipe for pasta sauce.

[This is the one.] Pivo held out the femur to Sava and tossed the others to the floor.

[What does it say?]

[I don't know. Not sure I want to know.] Pivo continued to hold out the femur.

[Enough with the drama Pivo. Just get your nanos to read it. We need a copy of the data. If you don't want to carry, I will.]

[I'd prefer that. Thank you.] Pivo set his nanobots to work on deciphering the inscription. When they were done, the intel was transmitted directly to Sava. Pivo wanted no part of it.

[So, now what? How do we get out of here? The only way out is the way we came in, and that's suicide.] Pivo's complexion changed from a milky green to an almost royal blue. It always happened when helplessness began to settle in.

Sava did not hesitate to answer, choosing to speak as opposed to transmit. ``We're not gonna leave, Pivo. Not even gonna try.'' Sava raised the plasma rifle and aimed it directly at Pivo's oblong head. ``See you next time, calamari.'' Sava pulled the trigger, and a fiery bolt of plasma reduced Pivo to a twitching mass of bloody scorched cartilage atop writhing arms. The arms continued to flop on the floor in a growing pool of blood as Sava sat down next to a pile of bones and leaned against the wall.

Sava pulled out a cigarette and lit it. The first inhale was virtually orgasmic. Sava loved to smoke.

Upon exhale, Careza pinged. [Shall I contact Project Ozma?]

Yeah. Get our lady on the line.

A woman's voice, cold and harsh, entered Sava's head, so different from the soothe of Careza. [Are you prepared to deliver, Agent Sava?]

[That depends.] Sava took another drag.

[Perhaps I did not make myself clear during our initial negotiations, Agent Sava. Your options are rather limited. You are unlikely to make it off the planet alive, and we cannot afford to lose this information, nor can we afford to have it fall into the hands of your organization. You are going to have to follow through, and trust that we will do the same.]

[Either you give me her location right now, or I take your precious info with me.] There was a long pause before the woman transmitted again. [You realize there will be consequences, Agent Sava. For you and for Rati.]

[Yeah. I suppose so.] The cigarette burned to the filter and Sava flicked it into a bone pile. [So what's it gonna be?]

[We do not bargain, Agent Sava, after a deal has been struck. Do as you will, and we will react accordingly.] The connection with the woman terminated. Sava stood up and walked over to where the courier's femur lay and picked it up. Pivo's gore coated the bone. Sava wiped it off and held it up to take a close look.

Sorry, Careza. Info payload only. Leave the ego behind.

[Understood.]

With the flash of a thought, Sava instructed Careza to activate the cortical stack's emergency farcaster—a one-shot neutrino transmitter, powered by the tiniest amount of antimatter. Sava's head exploded all over the room, taking the courier's femur with it. The information contained on the femur, however, found its way almost instantly through the blackest depths of space, landing safely onto a dedicated Firewall receiver elsewhere in the solar system.

\begin{center} *** \end{center}

``What's the number?''

The words dig their claws into my new vocal cords and yank themselves up and out of my parched throat. My diction is predictably poor, as it always is during the first few minutes following a resleeve. The pitch of the voice is apparent despite the mumbled, sandpaper slur of the words. Definitely a biomorph and my latest sex is female. This much I know in the first few seconds.

%%% Local Variables: %%% mode: latex %%% TeX-master: "ep" %%% End: 
