\chapter{A time of eclipse}
\label{chap:a-time-of-eclipse} 

This chapter provides a complete overview of the Eclipse Phase universe. It starts with a history, goes into detail on the setting, covers factions, and then wraps up with a system gazetteer. 

\section{A people's history of an unfortunate universe}
\label{sec:peoples-history-an} 

\begin{quotation}
The following is a transcript of a recovered au- diofile recovered after the catastrophic decompression event on Walther-Pembroke Station. The audiofile is believed to have been created by Donovan Astrides and to be a summation of his unpublished work A People's History of an Unfortunate Universe.
\end{quotation} 

[Sounds of scratching on the microphone, creaking of furniture, the noise of a woman clearing her throat] 

What? 

[Indistinct murmuring] 

Fuck you. I do this how the fuck I want, though it was nice of you to put me in this nice young woman's body. 

[Sounds of hands running along fabric] 

Does my vulgarity shock you, corporate lackey? No matter, I'm sure you can edit it out for your proles. Now—you asked about my book? Is it a history book you ask? No. It is an anti-history book. I shall tell you about the future. 

[Mumbling, questioning tone] 

What does it hold? The future, you mean? 

[Indistinct “Yes.”] 

No. I don't think you care about the future. What you really want to know is: will you get the future you want? And that is an easy question to answer. No. No, you will not get the future you want. Because you are stupid enough to ask this stupid question about the future. 

[Silent pause] 

I remember reading a scan of an old real print comic once. The character in it was railing against the imaginary people of his imaginary world, taking them to task about their dissatisfaction with the future they lived in. But it was really aimed at the stupid people who wanted their stupid little futures and who were too stupid to see that the future is now. It's always now. Except it isn't anymore. The TITANs changed that. The future is now yesterday, and last week, and ten years ago. Especially ten years ago. But the future is also back on poor old Earth—it's a legacy of where we've been and what has come before. 

 Do they teach you history on Venus, in your sealed compounds and resort aerostats? No, don't open your mouth, I could really care less what they teach you. For it is most certainly lies. I've lived in the inner system. I know the rules and the deceits told in the name of civil order and “national security.” 

Nations! Ha! Even at the onset of the 21st century, nations were starting to go into decline. It just took everyone a while to realize they were obsolete. 

Do you remember the great nations of the world? Are you old enough to remember how they sat around and debated whether the major climate shifts they were creating were even real? Even when many of them agreed that something needed to be done, none of them stood up to do it. The leaders of the world carried on with business as usual, secure in their privilege, as droughts ravaged Africa and Central Asia, Europe froze, and severe weather wreaked havoc everywhere. People across the globe were feeling the pinch of starvation or rampant epidemics, but the leading nations were more concerned about the refugees pouring over their borders and polluting their lily white paradises with their customs and languages and willingness to work for a pittance just to survive. 

The wars over oil and energy were only worsened by wars over the weather and water that followed. Unstable regimes rose and fell or were pushed over the edge, all in pursuit of precious liquids. The great nation states transformed into fortresses, steeled against the twin threats of the barbarians threatening them on the outside and the masses of their poor and dispossessed internally, all of them wanting to come in only for a little drink. 

You know, I've actually heard some conservatives refer to that period as a golden age, a peak time for the corporations and the rich. It's certainly true that it was a golden age for repression—and profits. If you were in that lucky fraction of a percent of the population who could afford it, it was certainly a good time, but for the majority of humanity it was a time of horrors. Global inequality was larger than ever before. Robots were taking jobs away from human hands. 

This was a time of radicalization for many. Failing governments no longer supplied people's basic needs. The globalized poor turned to local tribes, fundamentalist groups, political radicals, and criminal networks for the means to survive. Insurgent groups flourished, but they depended on the black market to survive, and soon their leaders were more concerned with making money than making change. 

The nation states, as always, resorted to repression. Civil liberties were restricted and surveillance increased. Automated weapons systems were deployed first against guerrillas and terror cells, and then against agitators and demonstrators. I remember the first time I saw those police drones, at a demonstration in support of a worker's strike in Long Beach. The drones ordered us to disperse once, only once, before they opened fire with their “nonlethal” weapons. Nonlethal my ass. Three people died that day and dozens were injured. The mainstream media ignored it even if the bloggers didn't. 

Meanwhile, the privileged elites continued to prosper. Longevity treatments expanded lifespans—for those who could afford it. Major crackdowns swept up off-brand pharma and bootleg procedures by pioneering biochemists, even while worldwide life expectancies dropped for the first time in decades. Why extend the lives of so many poor people, when expert systems as smart as any human could be built in a fraction of the time it would take to educate an actual person, and robotics and drone technologies allowed menial jobs to be turned over to uncomplaining and unpaid labor. And the rich had their high-pricetag designer chimeric pets to keep them company anyway. 

Not all of the upper classes were wallowing in opulence while the planet around them starved and drowned. A few were looking ahead at the changes on the horizon, scheming how to stake their claim. Some of these worked to expand their dominion, building a space elevator in sub-Saharan Africa and sending robotic probes out to map the solar system in detail. They even founded the first stations on Mars and Luna then, more than fifty years before the Fall. 

The ecopocalypse wasn't going away, however, no matter how much those in power tried to ignore it. Severe winters and draughts continued to pound at us. Rising ocean levels devastated coastlines worldwide with massive flooding. A few last-ditch efforts to undertake mega-scale geoengineering projects created as many problems as they fixed. These were viewed with cynicism anyway, as some were thinly-disguised test runs for terraforming techniques being prepared for off-world deployment. 

It often seemed as though the eyes of the fortunate were no longer focused on the world around them, but rather on the heavens above them. The completion of the first space elevator and the first mass driver on our moon kicked off a new space race and the competition was on to stake claims around the solar system. All this new expansion was powered by the first mass-produced efficient fusion power plants and the establishment of Helium-3 mining enterprises. 

Back on Earth, though, the hammer finally fell. Insurgents adopted fifth generation warfare techniques, sharing open source methods of resistance, utilizing swarming attacks on critical systempunkts. People crushed under years of oppression rose up in these opportunities and smashed at the state and corporate apparatus that had held them down. Nation after nation fell to insurgencies manned by those who had fought in thousands of little wars over fuel, ponds, and bread crusts. 

Most states fought back by becoming more totalitarian and repressive, but the tide of rebellion spread off world as a series of outposts and stations declared themselves in sympathy with their earthbound compatriots and announced a manifesto for a more humanistic approach to solar expansion. Even numerous scientists and engineers, who had previously worked as pawns in corporate expansions, adopted a technoprogressive stance. That's how the argonauts were born, you know, taking their name from a previous group of scientists who advised the US government and Pentagon on science and policy called the Jasons. Faced with reprisals from their corporate masters, a number of argonauts defected from the hypercorps, in some cases taking key resources and research with them, while others went underground. 

This is when the hypercorps really took off, though, those shark-like bastards. They let the nation-states and lumbering multinationals of old take the brunt of the global rage and assault. They took advantage of the chaos to slip free of the old moral and ethical restraints on human experimentation and from the legal purview of the nationalities that had birthed them. They embraced the opportunities of numerous new technologies and the drive into space. It was their research labs that cooked up the first sentient artificial intelligences, the first genegineered human clones, and the first true uplifts, chimps and dolphins brought into awareness as corporate experiments and slaves. 

As the last of the old states became increasingly desperate to cling to their power and land, the hypercorps extended a helping hand. They offered debt bondage terms to those who were willing to sign over their rights and humanity for a trip off world, to work as indentured servants on corporate colonies and stations. Hundreds of thousands took the offer as an alternative to the crushing poverty and chaos on Earth. The business of resource exploitation exploded across the solar system, as stations were established as far out as the Kuiper Belt. Voices that spoke of respecting biodiversity and natural ecologies were ignored as the hypercorps toiled to reshape various planets and moons to their will. 

This was the state of things until about 20 years before the Fall. Though many of the old oppressor states had been struck down, new ones arose, and the various global insurgencies oscillated between making radical changes and falling into the same old tribal warfare traps. Reactionary religious and political forces on Earth also railed against the hypercorps' agenda, resulting in some terrorist attacks and sabotage strikes, and culminating in a failed attempt to disable the space elevator by an Islamist suicide cell. The hypercorps were quick to retaliate, ordering an orbital bombardment using high-density objects against the headquarters and compounds of several key opposition leaders. Though effective in decapitating several terrorist networks, the mass destruction sparked outrage against the hypercorps, creating a deeper rift between Earth and off-world interests. 

The hypercorps remained out of reach, however, though they were not completely immune from Earth's troubles. The workers and colonists brought from Earth transported many of their ethnic, political, and socio-tribal grudges with them, leading to several outbreaks of violence in habitats and orbital stations. Some also harbored allegiances opposed to hypercorp interests, illustrated by isolated acts of preservationist sabotage and religious terrorist attacks. Various criminal networks also came along for the ride, expanding their black markets and vice trades wherever humans went. 

As the hypercorps expanded, so too did their political opponents: the anarchists, socialists, argonauts and others who worked diligently to establish their own independent presence, mostly in the outer system, further from hypercorp reach. The hypercorps even contributed to this growth by sending their criminals and undesirable elements into exile beyond Mars. 

Both sides invested heavily in research and new technologies. Advances in biotech, nanotech, AI, and cognitive science were now moving so rapidly that major breakthroughs were made on a yearly basis. Developments in one field created a recursive boost in the others, creating a feedback loop that spawned immense technological improvements. Off-world, genetic modifications were widely adopted, and new transhuman adaptations became a common sight. We even created new synthetic life forms that were part biological and part robotic. Despite some being so repulsed by this development that they dubbed these new types of beings “pod people,” it certainly didn't stop pods from being rapidly absorbed into corporate workforces and brothels, nor did many people care enough to support claims that, as sapient beings, pods should have their own civil rights. 

Two breakthroughs in this period deserve specific mention, not least because of their impact on our human—now transhuman—society. The development of the first nanotech assemblers signaled a paradigm shift for economics. Available only to the upper strata of the hypercorps at first, these elites jealously guarded these machines, capable of building almost anything from the atoms up. They placed all sorts of restrictions on their usage and availability, claiming that the capability to construct drugs, weapons, or other restricted items was a security risk that required them to be strictly controlled. Open source advocates promptly set to work undermining blueprint controls and seeding their own open source designs, of course. Likewise, within months, criminals and anarchists liberated their own assemblers, and suddenly an economic conflict was born. Some were put to use feeding the black market trade, while others were used to establish habitats and colonies with post-scarcity economies that no longer relied on wealth, property, or greed. 

At the same time came the ability to map the human brain and digitally emulate the mind and memories, making “uploading” possible—followed closely by the ability to download back into a separate human brain of course. The already long-lived hypercorp masters no longer had to fear death by accident or injury. This technology also made its way into the hands of others, despite the costs. Experimentation with other bodies—both biological and synthetic—became a new playground for culture. And let's not forget those who willingly shook off the shackles of the flesh to experience the virtual life and dive deep into their own dreamscape realities. 

While we all enjoyed our new toys, though, Earth, poor Earth, continued to die a slow death. I can still recall the speculation that it might take centuries for the Earth to totally slide into ecological devastation. It was frustrating, everywhere you turned it seemed that someone was lamenting the state of the motherworld, but no one wanted to do anything. It was too expensive, or too far away, or too dangerous. We all have blood on our hands from that time. We stood by and watched from our places in orbit as the world burned around our brothers and sisters. We thought we had time, we thought the world was slowly dying and that we could find the cure. We didn't plan on the TITANs. 

We all remember the Fall. It was only ten years ago, but I never cease to be amazed at how confused people's memories are of that time. Part of that is propaganda perpetuated by people like you, of course, and part of it is that most of us are afraid to really look back and examine how we humans managed to fuck it up so badly. 

We like to pretend that the TITANs exploded on the scene, wrecked up the place, and then disappeared as quickly as they appeared. The truth, as always, is more complex. We claim to know that the TITANs somehow evolved by accident from a military netwar system, or so the theory goes. That is what their name means: an acronym for Total Information Tactical Awareness Networks. No one knows for sure where these first seed AIs came from, though—or if they do, they're keeping quiet. Perhaps the TITANs were intentionally designed to be a recursively improving, self-aware digital intelligence. Perhaps the military boffins thought they could keep such an intelligence under their control, and that it would give them the edge they needed. Perhaps there was only one at first, and it quickly created hundreds if not thousands of copies of itself. No one even seems to know how many of them there were. 

According to the written history—vetted by the hypercorps natch—we now know that the TITANs took several days after they “woke up” to scan the world around them, to learn about us. In their initial stage they were relatively benign, leeching network power and resources only where there was enough to spare and extending their senses beyond their cradle on Earth. Perhaps they were absorbing everything they could to understand us. Perhaps they were indifferent. Or maybe they really were planning to destroy us, as the vids all say. 

I remember this time. I remember that when this new round of conflicts re-ignited on Earth, there was no word of anything about seed AIs or TITANs. For months and months, it was a simple escalation of hostilities. It started with claims of netwar operations and major intrusions, sparking some alarm and retaliatory attacks. Aggressive stances led to incriminations, then border conflicts and raids, followed by missile strikes and outright hostilities. Old grudges and sleeping enemies suddenly awoke and turned their renewed wrath against old foes. Brush wars, corporate rivalries, and ideological disputes flared up as insurgencies and rebellions were suddenly everywhere. At the time, it seemed like a not-so-unusual spate of violence had taken a drastic turn and was rapidly spiraling out of control. 

According to the party line, this was all a carefully concerted effort, the first stage in the TITANs plans. Perhaps it was, though I remember some military officials once claiming that the TITANs were brought online because of this violence, and not before then— an opinion that was quickly silenced. Then again, maybe we really were played—played by greater intelligences who could barely be bothered to deal with us themselves when they knew we were more than willing to murder and annihilate each other. 

When the first reports of strange automatic factories cranking out large numbers of robotic weapons systems broke, no one knew who to blame, but clearly something was wrong. This was a turning point, a chance for humanity to realize that we collectively faced a new enemy, but the finger-pointing and direct conflict continued. Even when the first open attacks by the TITANs came in earnest, crashing major systems, taking control of critical infrastructures, and wreaking havoc and destruction, we treated it as a new front in the war, and never stopped taking shots at each other. 

There is still debate over whether we should have tried to talk to the TITANs, whether they would have been willing to listen to us, whether they even saw us as something more than we see rats and roaches and other forms of vermin. But it's all academic. The fact is we didn't. The people who made the decisions, the ones who had to put it all on the line at the time, saw the TITANs as a threat. And they acted accordingly, trying to purge them from their systems or capture them for future study. 

The philosopher Thomas Hobbes once spoke of the war of all against all. Whatever he imagined could not have been anything close to the conflict ignited by the TITANs. We killed ourselves by the millions, wielding the nuclear fire and the silent death of bioplagues indiscriminately. Among this carnage walked the TITANs, taking control of our machines as though we were children, harvesting millions of minds with forced uploads for unknown purposes. Every strike we launched against the TITANs was met with untold disaster and ruin, all our artifice and devices turned against us in our moment of need. 

The Fall was a horror. Factories sprang up like a blight in the most ravaged and deserted places on Earth, pumping out legions of dread war machines. Advanced nanoswarms—far beyond our own capabilities—infested everywhere, mutating to deal with any threat they encountered. Biological nanovirii ripped through human populations, inflicting irreversible neurological damage. Potent infowar worms penetrated even hardened systems, shredding our crucial networks with ease. Prisoner populations were rounded up for forced mind emulations, suffering a luckier fate than those who were merely decapitated by head-collecting drones or pierced by robots with neuro-scanning proboscises. Neuropathic virii turned some humans into pawns of the TITANs, turning them against the rest of us. Other reports spoke of strange, alien happenings and unimaginable terrors. We found ourselves fighting a rearguard action against coming extinction. The plot of a hundred novels and movies made manifest in our lifetimes, the doom of transhumanity at the hands of the machines. 

For over a year they stalked and destroyed us. There seemed to be no hurry on their part to bring us to an end, and why would there have been? Nothing we did affected them. They were data and information, they were thought and impulse, they were everywhere and nowhere, and there was nothing we could do that they could not turn back against us. Their influence spread outward from Earth, with outbreaks in orbit, on Luna, Mars, and many other places. Everywhere we had a foothold, the TITANs followed. 

Perhaps you remember that point when it became clear that transhumanity might not survive. I do. Millions must have seen the signs. And so the great diaspora began, the teeming masses doing whatever they could to flee Earth. Ships were diverted, even built, to help people escape. Those who could not buy their way off the planet did their best to send their digital backups, in the dim hope they could acquire a new body. Perhaps one in ten escaped. 

You might hear that we banded together to stop the threat, that in our darkest hour we forgave ancient grudges and simmering hatreds in the face of extinction. That would be a lie in the face of the ten thousand shot down over Buenos Aires by North American forces as they sought to escape, or the compromising of network security on over two dozen habitats in Lagrange orbits by corporate competitors as their rivals strove to fight off a TITANs attack. We were just as gleeful to destroy ourselves. 

Then, as quickly as they appeared, the TITANs vanished. Over the course of a week, the attacks and disturbances trailed off and then stopped but for an occasional outbreak. The retributions and attacks by our own kind continued for a few more months, but the damage we did to ourselves was nothing compared to what the TITANs had done. 

In the aftermath, we stood among the smoking ruins of transhumanity and surveyed all that had been lost. Of all the billions that existed before the Fall, fewer than one in every eight survived, and of those fewer still retained a corporeal form. Nevertheless, the surviving habitats and stations were overcrowded, with tensions high. Vast numbers of infugees circulated in storage, as there were simply not enough bodies on hand to accommodate them all. Some were placed in permanent storage, where they remain forgotten. Others were shunted into virtual reality, given no choice but to live their lives in simulated environments. A lucky few were given the chance to work as indentured servants, often to build new habitats, working on the promise of a body of their own someday. You've no doubt seen them, working in cheap mass produced synthmorph bodies in menial or dangerous tasks kept out of sight of the rest of us. 

Those left dead or bereft of a body were the least of our problems. Our war with the TITANs had left the Earth a smoking, irradiated, toxic wasteland, still populated by dangerous machines and plagues. The newly formed Planetary Consortium, composed of hypercorp interests among the Martian and Lunar colonies, placed Earth and the space around it under quarantine. The official reason is that it's for safety reasons, allegedly to keep any remaining threats from escaping Earth's confines. Or perhaps we could not stand to look at our homeworld in such a state and face what we had done to ourselves. 

Even now, ten years later, we are told that the Earth is dangerous, that it holds risks and surprises. That's partly true, I believe—there are surprises alright, but the Planetary Consortium wants them all for itself. 

[Rustling noises, murmurs] 

Of course I'm talking about a Pandora Gate. The one the TITANs left behind on Saturn's moon was just the first. You're a fool if you think that there are only five in the entire system. I'd be willing to bet nearly anything that there's one down there on dear old Earth. 

Have you ever seen a Gate? No? Of course not. The hypercorps keep them locked down. Not like out in the wild, wild outer system. Sure, the Gatekeeper Corp lets anyone with a death wish and the minimum training take a jaunt through the original on Pandora, but if you're lucky enough to come back, they own everything you find on the other side. I suppose it's the chance for a certain type of adrenaline junkie “to boldly go” and all that nonsense. 

The extrasolar colonies—now, those are an all new frontier. You inner system types are so predictable with your rush to colonize and expand and own everything, as if the universe is just there for your rich overlords to claim for themselves. I expect your extrasolar colonies are expanding quite nicely, given the sheer number of poor debt-conscripted souls you toss through. You probably have grand schemes of building galactic empires. Us. Transhumanity. A galactic civilization. 

Well, galactic squatters at least. That was made clear when the solemn crossing guards of the cosmos showed up and issued us a warning that we were dabbling in Things What Ought Not To Have Been. 

Maybe the Factors are telling us the truth, maybe they are acting as ambassadors for a collection of spacefaring alien species that want to warn us away from Forbidden Technology—y'know, the technology we've already been burned by and of course have no plans to actually abandon. Think about the Two Commandments they have given us: thou shalt not create self-improving AI, and thou shalt not use the Pandora Gates. Oops. Do you think they know? About what happened with the TITANs? That even we don't know where they went and that we're kind of afraid to find out? Surely they know that we've been using the gates and have spread beyond our little backwater, and maybe that's their real fear. But why do we even listen to what some highly-evolved slime mold tells us to do anyway? 

Taking risks, that's the price of progress, no? Let's face it, we need some hope. We need a new Earth to replace the one we destroyed, a place where we can go and breed like rabbits and fuck it all up over and over again. We need to know that we can expand beyond this solar system, because right now it's feeling a little confining, like we could be easily trapped and wiped out if the TITANs ever return. We need to know that we have a future. We need to know that we can make it through our own efforts. That we won't do ourselves in on our own. 

The Lost proved that. It was a noble objective, to speed a new generation of children to adulthood, but the process was flawed. Taking force-grown clones, raising them in VR, and then dumping them into adult bodies after they've only been alive for a few years of objective time—but over eighteen years of their subjective time? An entire childhood, having only each other and AIs for company. It's enough to fuck anyone up. It was a grand experiment, but it failed, and now we have another reminder of our failures living among us. 

That's us, in all our glory. Ten years after the Fall and we remain a broken, squabbling mess, jailed by slime molds, beaten by uppity software, and yet our own worst enemies. Spreading out from a home we don't even have any more. Our numbers reduced and dwindling further with each passing day. Who will save us? We don't even want to save ourselves most of the time. Or so it seems. 

But if we don't, there's no future. And I, for one, have not lived this fucking long to give up now. You, me, we're effectively immortal. The entire galaxy is waiting out there for us. We'd be stupid not to go see it. 

\begin{quotation}
End Transcript
\end{quotation} 

\subsection{Eclipse phase timeline}
\label{sec:eclipse-phase-timel} 

All dates are given in reference to the Fall. BF = Before the Fall. AF = After the Fall. (e.g., BF 10 = 10 years before the Fall.)  \paragraph{ BF 60+} 

\begin{itemize}
\item Crisis grips the globe in the form of drastic climate changes, energy shortages, and geopolitical instability.
\item Initial space expansion creates stations at the Lagrange Points, Luna, and Mars, with robotic exploration of the entire system.
\item Construction begins on a space elevator.
\item Medical advances improve health and organ repair. The rich pursue gene-fixing and transgenic pets.
\item Computer intelligence capabilities equal and excede that of the human brain. True AI not yet developed.
\item Robotics become widespread and start to replace/invalidate many jobs.
\item Modern nations expand their high-speed wireless networks.
\end{itemize} 

\paragraph{ BF 60-40} 

\begin{itemize}
\item Efforts to undertake megascale geoengineering on Earth cause as many problems as they fix.
\item Major colonies established on the Moon and Mars; outposts established near Mercury, Venus, and the Belt. Explorers reach Pluto.
\item First space elevator on Earth finished. Two others in progress. Space traffic booms.
\item Mass driver built on the Moon.
\item Terraforming of Mars begins.
\item Fusion power developed and working plants established.
\item Genetic enhancements, gene therapies (for longevity), and cybernetic implants become available to the wealthy and powerful.
\item First non-autonomous AIs are secretly developed and quickly put to use in research and netwar.
\item Experience playback (XP) technology developed and put into public use.
\end{itemize} 

\paragraph{BF 40-20} 

\begin{itemize}
\item Violence and destabilization wrack the Earth; some conflicts spread into space.
\item Argonauts split from hypercorps, taking resources to autonomist habitats.
\item Space expansion opens up legal/ethical loopholes for tech development and allows for increased direct human experimentation.
\item Human cloning becomes possible and available in some areas.
\item Development of first transhuman species.
\item First dolphins and chimpanzees uplifted to sapience.
\item Fusion-drive spacecraft enter common usage.
\item Extended colonization and terraforming of Mars continues. Belt and Titan colonized. Stations established throughout the system.
\item The starving masses volunteer themselves for indentured servitude on hypercorp space projects.
\item Augmented reality becomes widespread.
\item Most networks transformed into self-repairing mesh networks.
\item Personal AI aides become widespread.
\end{itemize} 

\paragraph{ BF 20-0} 

\begin{itemize}
\item Earth continues to suffer, but the pace of technology allows for some interesting developments.
\item Expansion throughout the system, even into the Kuiper Belt.
\item Transhuman species become widespread.
\item Nanotech assemblers become available, but are strictly controlled and jealously guarded by the elite and powerful.
\item Uploading and the digital emulation of memory and consciousness made possible.
\item More species (gorillas, orangutans, octopi, ravens, parrots) uplifted to sapience.
\item Pods see common usage, amid some controversy.
\end{itemize} 

\paragraph{ The Fall} 

\begin{itemize}
\item The TITANs evolve from a high-level distributed netwar experiment into self-improving seed AIs. For the first few days, their existence is unsuspected. They advance their awareness, knowledge, and power exponentially, infiltrating the mesh both on Earth and around the system.
\item Large-scale netwar incursions break out between rival states on Earth, sparking numerous conflicts. These attacks are later blamed on the TITANs.
\item Simmering tensions on Earth escalate into outright hostilities and warfare.
\item Massive netwar breaks out and major systems crash as TITANs begin open attacks, also using autonomous war machines.
\item Conflict quickly spirals out of control. The use of nuclear, biological, chemical, digital, and nanotech weapons reported by all sides.
\item TITANs engage in mass forced uploading of human minds.
\item TITAN attacks expand to other parts of solar system, heaviest on the Moon and Mars. Numerous habitats also fall.
\item TITANs suddenly disappear from system, taking millions of uploaded minds with them.
\item The Earth is left a devastated wasteland, a patchwork of radiation hotspots, sterile zones, nanoswarm clouds, roaming war machines, and other unknown and hidden things among the ruins.
\end{itemize} 

\paragraph{AF 0-10} 

\begin{itemize}
\item A wormhole gateway is discovered on Saturn's moon Pandora, left by the TITANs. Four others are later found (in the Vulcanoids, on Mars, on Uranus, and in the Kuiper Belt); these are collectively referred to as ``Pandora Gates.''
\item Expeditions are sent to extrasolar worlds via the Pandora Gates. Numerous exoplanet colonies established.
\item First contact with the aliens known as the Factors shocks the system. Claiming to act as ambassadors for other alien civilizations, they provide little information about life outside the solar system and warn transhumans away from both seed AI and the Pandora Gates.
\item An attempt to raise a generation of children using force-grown clones and time-accelerated VR fails miserably when most of the children die or go insane. Dubbed the Lost Generation, the survivors are viewed with repugnance and pity.
\end{itemize}  \paragraph{AF 10}  \begin{itemize}
\item Present day.
\end{itemize} 

\section{The solar system after the fall}
\label{sec:solar-system-after} 

Before the Fall, the solar system had a population of approximately eight billion, with all but five million of these people living on Earth. The Fall wiped out almost ninety-five percent of transhumanity, and today the population of the solar system is slightly less than half a billion, with almost all of these transhumans living off the Earth. The lifestyles of these people were almost unimaginable thirty years earlier—the vast majority are immortals living in sealed habitats on hostile alien planets or in sealed space colonies, the largest of which hold more than a million inhabitants and are many kilometers long. 

In this vastly changed setting with its vastly changed inhabitants, the core concerns of humanity remain much the same. People seek both material abundance and social status, and they wrap themselves in various public and private ceremonies. Like generations of humans before them, transhumans separate themselves into different cultures and subcultures, all of which enjoy a wide variety of physical and virtual entertainments. Politics and economics remain vitally important and as always, those who are wealthy, powerful, and famous have a large degree of control over the lives of those who are poor, relatively powerless, and unknown. 

\section{Transhumanity}
\label{sec:transhumanity} 

Humanity as a concept has been replaced with transhumanity. Most people now alive left Earth as infomorphs and were subsequently resleeved into new morphs. Bodies are things that can be modified and replaced, much as someone can alter or exchange a suit of clothing. Identity is centered in the mind, which can exist as a disembodied infomorph living in virtual worlds or dwelling in a vast array of strange and exotic morphs. While there are bioconservatives who resist these many changes to identity and physicality, they are very much in the minority. 

To most people, transhumanity has also been expanded in scope to factor in non-human persons such as AGIs and uplifts, though the rights and status of these sentients is sometimes contested. 

As transhumans continue to absorb the ramifications of this new way of life, they face a new crop of problems and issues. Two of the largest and most important are the increase in inequality and the splintering and separation of transhumanity into many different clades. 

\subsection{Inequality}
\label{sec:inequality} 

The technologies first developed in the decade before the Fall and refined in the decade after its end have transformed humanity. In all but the most backwards, impoverished, and repressive regions of the solar system, the vast majority of humanity is smarter, healthier, and richer than any humans have ever been. Additionally, individuals can improve their minds and their bodies in almost any fashion their imaginations can dream up. Those who can afford the right augmentations can think faster, never forget anything they have ever learned, become mathematical savants, and heal from injuries many times faster than an unmodified human. When resleeving is combined with implants, transhumans can gain even more amazing capabilities—but these benefits are far from free. 

During the first decade after the Fall, most of the surviving population was relatively poor. Many were grateful to have any morph at all. While the economic situation has improved, significant inequalities remain and seem unlikely to change. Hundreds of millions of people must make do with very basic splicers (p. 139), worker pods (p. 142), cases (p. 143), or synths (p. 143), while a few million are wealthy enough to have custom-designed morphs created for them, complete with all the augmentations they desire. These same members of the elite live in luxurious villas and mansions, and in a few cases privately-owned asteroids, while most other people must make do with a few hundred cubic meters of dwelling space. However, while inequities of living space are ancient, the issue of economic inequality producing inequities of physical and mental capacities is both relatively new and considerably more problematic. 

In regions using the old and transitional economies (see p. 61), differences between the rich and the poor are expressed in terms of money. In habitats using the new economy (p. 62), wealth is meaningless and status and opportunity are denoted with reputation scores. In all three economies, some people have more than others, and because of this, technology allows the better off to be better than the people around them. Skillware lets people buy knowledge and expertise, while multi-tasking and mental speed implants allow individuals to get more done at once. Someone fortunate enough to acquire large numbers of such augmentations is capable of significantly more than someone who lacks them, and so can do even more to increase their money or rep, thus serving to further perpetuate inequality. This problem is less serious in the reputation-based economies of the outer system, however, as it significantly easier to build reputation through hard work and dedication, as opposed to the rigidly-controlled monetary economies of the inner system and the Jovian Republic, where class stratification is institutionalized and upward mobility is largely a myth. 

As many supporters of the status quo are fond of pointing out, even the “have-nots” are smarter and healthier than any previous generation of humans and carry as much potential immortality as the wealthiest member of the elite. It is equally true, however, that in many ways the divisions between rich and the poor are significantly greater than they have ever been, especially in the inner system. In the past, the members of the elite might be somewhat healthier and better fed than the have-nots, but both rich and poor still lived in relatively similar and fundamentally human bodies. Now, the very nature of humanity has been called into question. The least fortunate can be forced to inhabit bodies designed specifically for the pleasure of those wealthier than them or even denied any body and forced to live as infomorphs until they can find some way to acquire a new morph—typically by selling their services to the highest bidder. Meanwhile, the well-off can customize their bodies and their minds, enabling them to accomplish far more and to be considerably more impressive and charismatic than anyone lacking their advantages. These inequalities may seem insurmountable, but some anarchistic groups and even some entire habitats have dedicated themselves to reducing inequities by producing low cost (and occasionally highly unreliable) versions of many of the more impressive morphs and augmentations. 

\subsection{Clades and separation}
\label{sec:clades-separation} 

In many habitats, hyper-augmented elites rule a mass of humanity that is stuck using low-end morphs and minimal augmentations, or even infomorphs living in rented morphs, but this is not the only option found in the solar system. Transhumanity has splintered into a wide variety of subcultures, some of which are based upon an individual's choice of morph. Some of this separation is due to the necessity of inhabiting difficult environments. From aquanauts living in Europa's aquatic environment or rusters on Mars to the fact that zero-g habitats are relatively inexpensive and are best inhabited by microgravity-adapted morphs like bouncers, many unusual environments require those living in them to choose from a very limited range of morphs. Sometimes, though, this separation is ideological in nature, such as the rise of groups like the ultimates (p. 82) or some of the separatist uplift communities that seek to define their own space separate from human cultures. 

There are dozens of specialist morphs and an even greater number of habitats or other settlements that are inhabited largely or exclusively by individuals using a single type of morph or a limited number of specialist morphs. In the asteroid belt and in the rings and smaller moons of Saturn, there are more than one hundred habitats that do not rotate, with all portions in zero or near-zero gravity. The inhabitants typically use bouncer or novacrab morphs, along with a small number of synthetic morphs and other pods. 

There is also a vast number of other habitats that are segregated in various other ways, including ones where all permanent residents are uplifts inhabiting one of the various transgenic morphs, like the octomorph or neo-avian morphs. Other habitats are only open to residents with various enhanced morphs like exalts or mentons. There are even habitats where all residents must inhabit morphs that are all clones of one another. In almost all of these habitats, residents are free to add whatever augmentations they wish to their morphs, but some habitats forbid residents from changing their morph's external appearance, and individuals who violate this rule are forced to leave the habitat if they refuse to reverse these changes. 

Some habitats do away with the necessity of both life support and gravity. In these locations, all residents are infomorphs who either inhabit their own synth bodies or, in a few very eccentric cases, where all of the inhabitants are infomorphs who spend most of their existence in the habitat's central computers. When they need to interact with the physical world, these infomorphs are free to use one of the many synthmorphs that the habitat owns and that the residents share among themselves. Although considered quite eccentric to many and horrifying to bioconservatives, habitats inhabited solely by synthmorphs or infomorphs are among the least expensive to build and maintain and are a low-cost way for groups of infomorph refugees from Earth to gain independence. Because individuals who choose this way of life have often spent a decade or more as infomorphs, this option often seems both familiar and in many ways more comfortable than inhabiting a living morph. As Earth becomes more distant in transhumanity's collective memory, its traditions and social norms hold less sway and people feel more free to create and use new bodies and new ways of life to go along with them. 

\subsection{First contact: the factors}
\label{sec:first-cont-fact} 

Ironically, the first contact between transhumanity and alien life was made by a group of isolates with no interest in the rest of transhumanity. A brinker doomsday cult habitat in the Neptunian Trojans, patiently waiting out the prophesized return of the TITANs, suffered a severe life support systems failure. Not expecting anyone to respond to their distress signals, they were simultaneously relieved and shocked to have an alien starship come to their aid. 

Shortly after this event, three the unknown ships of alien design simultaneously approached Mars, Luna, and Titan, logging on to local networks to announce their presence and peaceful intentions. Though their presence initially raised alarm and panic, their rescue of the brinkers and assurances of non-hostility allowed cooler heads to prevail. Coming just three years after the silent hostility of the TITANs, the new aliens were pleasantly non-threatening. 

Quickly dubbed “Factors,” both because of their claims to act as ambassadors for an assortment of alien civilizations and of their interesting biology, initial communications between species were confusing and jumbled. The Factors made a number of veiled warnings and expressed concern over certain technological developments, particularly unrestrained artificial intelligence. They have refused entirely to deal with digital entities and broken off negotiations with anyone currently engaged in AGI development or utilizing the Pandora Gates. The Factors have implied that they were aware of and watching humanity for some time, but chose to wait to make contact ... implying some implicit fear of the singularity. Dealing with multiple factions, the primary relationship between the Factors to transhumanity is a commercial one. Though they are often dismissive of transhumanity’s technological achievements, they are interested in our scientific development and breakthroughs, particularly in the biosciences, as well as our art, history, and culture. They remain tight-lipped about their own civilization and other xenomorphs, though they have on occasion traded alien artifacts of unusual design and peculiar function. It is widely assumed that these are simply trinkets of limited value and that the Factors are careful not to share anything of true worth to transhumanity, particularly anything that might drastically affect our growth. Biologically, the Factors appear to be some sort of evolved slime mold colony. As far as is known, they communicate purely by chemical signals and receptors, requiring any interactions with transhumanity to be computermediated. Several different types of Factors have been sighted, implying that they engage in heavy biological modification. Factor starcraft appear to be lighthuggers capable of nearlight speeds. Due to the frequency of their visitations to the solar system (2–3 times a year), however, it is speculated that they either have a nearby base, or that they possess the capabilities for faster-than-light travel—or possibly they have Pandora Gates of their own. Given the wide dissimilarities in psychology between transhuman species and the Factors it would be presumptuous to speculate concerning their true feelings and agenda towards transhumanity. It is hoped, however, that by continuing negotiations with them, transhumanity may learn more about the nature of the galaxy—and possibly even our own history. 

\section{Culture and society}
\label{sec:culture-society} 

The Fall and its aftermath continues to be a major influence on transhuman culture and society. Prior to the start of the evacuation, more than ninety-nine percent of the people who survived the Fall had never been off Earth. For them, space was a distant realm where other, more daring and adventurous people lived, a place Earth dwellers only saw on videos. Earth was their home. Then, in the course of a few short years, hundreds of millions of people were forced to leave Earth. The fortunate few first evacuees left with no more than a dozen kilograms of possessions, while the vast majority were infomorph refugees who left Earth with nothing, not even their bodies. 

Today, transhumanity is divided into three groups. The first group contains the true veterans of space life, the less-than-one-percent of humanity that was already living in space before the Fall. The second group is the ten percent of the population that was either born after the Fall or is too young to remember living on Earth. The remaining eighty-nine percent of the current population of the solar system lived generally happy and prosperous lives on Earth before the Fall forced them to flee for their lives. These refugees from Earth form a powerful social force, but as time goes on memories of Earth grow dim, and people adapt to their new homes and lives. 

\subsection{The longing for earth}
\label{sec:longing-earth} 

Most of transhumanity, especially those who were forced to flee from the dying Earth, still mourn their former home. Their longing for and nostalgia of Earth has profoundly affected transhuman culture. Artifacts from Earth, including ones as trivial as coins or bits of dried vegetation, are considered to be treasured mementos that have great economic and emotional value. 

The interdiction of Earth makes acquiring such artifacts quite difficult and dangerous. As a result, the trade in Earth artifacts is a lucrative portion of the black market, enough so that fearless scavengers are willing to risk being shot down by a patrolling killsat just to get to Earth, where they also face death from numerous lingering dangers. The mesh is peppered with stories of daring explorers who traveled to Earth to retrieve all manner of priceless relics, as well as an equal number of stories about explorers who died or simply vanished on such expeditions. More than one team of gatecrashers has funded their expedition through a preliminary relic-hunting expedition to Earth, which serves to test their mettle while they work to raise funds. 

Nostalgia for Earth also affects the way transhumanity has redesigned itself. In the decade prior to the Fall, humanity had begun to freely alter itself, with both radical body modification and the first commercial resleeving resulting in a growing number of obviously non-human morphs. The vast majority of current morphs, however, are relatively human in appearance (if not in internal structure). Even for people too young to remember the Fall, asserting individual humanity is an important part of postFall culture. Some people keep a resemblance to the traditional human form as a remembrance of Earth, while others do it to celebrate humanity's victory over the monstrous and inhuman TITANs that attempted to destroy them. With the exception of a few eccentric groups like the ultimates, the majority of humanity values looking human and preserving human traditions and institutions. Also, even the ultimates' current version of their remade morph is considerably more human looking than the versions their predecessors designed before the Fall. As a result, while synthmorphs are relatively common, most are made to look humanoid. There are a few radically inhuman morphs like the novacrab, the arachnoid, and the flexbot, but they are almost exclusively used for highly specialized purposes. Until recently, anyone who used one as their primary morph was considered deeply eccentric (or worse), but attitudes have gradually begun to soften, and these morphs are gradually becoming more acceptable for regular use. 

This mixture of reverence and nostalgia for Earth sometimes has a darker side. Individuals who choose to have morphs that look visibly non-human experience a mild degree of prejudice in many habitats, and militant bioconservatives denounce those who look sufficiently non-human as being covert allies of the TITANs. Uplifted animals also face significant discrimination from many humans. These prejudices are relatively common in the inner system and can be quite extreme among bioconservatives. As a result, uplifts and individuals who prefer inhuman-looking morphs often live in separatist communities in the outer system. In much of the inner system, uplifts and individuals using a visibly non-human morph as their primary or only morph are viewed with suspicion and occasionally treated as second-class citizens. While most habitats have laws mandating morphological freedom and many also have laws making prejudice based on morphological choice illegal, these attitudes remain quite resilient. 

\subsection{Nostalgia jewelry}
\label{sec:nostalgia-jewelry} 

As both a reminder and a visible marker of their lost homeland, a significant number of refugees from Earth wear jewelry containing a coin or, more rarely, an old stamp from transhumanity's former home. Popularly known as nostalgia jewelry, most of these items are made into pendants or lapel pins, but a few are rings. Before the Fall, coins and stamps were largely curiosities primarily of interest to collectors, having fallen out of use forty years BF. Already scarce, few were saved during the Fall as carrying such useless mass off Earth during the evacuation was discouraged or forbidden. A few extensive collections already existed off-world, however. Even so, less than a million authentic samples survived, meaning the vast majority of people wearing such items make do with exact copies made in cornucopia machines. Actual coins or stamps are very expensive, meaning that some daring scavengers are willing to risk the interdiction of Earth for the express purpose of salvaging relics. 

\subsection{Fear and paranoia}
\label{sec:fear-paranoia} 

 The Fall left behind a persistent legacy of fear. This has faded over the past decade, but a great many humans still unconsciously expect the other shoe to drop and the TITANs to return at any moment. Others worry that their agents are already among them, preparing for the complete destruction of humanity. The arrival of the Factors caused widespread panic, and even today a substantial minority of people assumes they are cat's-paws for the TITANs—or possibly their creations. 

There are a few (often insane or deeply eccentric) people who worship the TITANs or otherwise support their agenda (including self-described “singularity seekers” who hope to find and be uploaded by the TITANs to join their ascension to super-intelligence), but all of them must keep their beliefs carefully hidden. 

Even now, expressing any support for the TITANs or advocating the creation of self-improving seed AIs is illegal in most habitats. Anyone who does so runs the risk of becoming the target of mob violence that the authorities are unlikely to investigate too closely. Merely being suspected of being a supporter of the TITANs, or worse, someone who has been secretly infected by them and is now their agent, is enough to get someone shunned or even killed. While such incidents are now far rarer than they were in the first few years after the Fall, people who act too eccentric and who lack someone with a sufficiently high rep to defend them or explain their actions are occasionally killed, typically by being thrown out an airlock. Those responsible for these “spacings” are dealt with quite harshly in most habitats, since in almost all cases later investigation reveals that the victim had no connection to the TITANs. 

There are also periodic rumors in many habitats, especially small and isolated habitats, that one or more other habitats have been taken over by the TITANs, leading to a variety of inter-habitat problems. Such rumors are usually resolved fairly quickly, but the most persistent can seriously harm relations between habitats. Claims that other habitats are infested with or even controlled by agents of the TITANs are frequenly employed by extreme bioconservatives hoping to demonize radical habitats populated entirely by infomorphs or synthmorphs. As more people manage to put the fear and horror of the Fall behind them, such claims are less likely to be believed. Unfortunately, on very rare occasions, people are still infected by TITAN-created relics and actually become their unwilling agents. Since such incidents are rare, however, they have become easy to dismiss. 

\begin{quotation}
\textbf{An exsurgent threat?} 

 [Incoming Message. Source: Anonymous] 

 [Public Key Decryption Complete]  Ok, you asked, so I’ll tell you. There are some elements within Firewall that don’t buy into the TITANs-ran-amokand-considered-us-a-threat idea, or even that the TITANs are solely responsible for the Fall. These people think that the TITANs found or encountered something when they started their takeoff toward the singularity—something that changed them. They point to the wide range of multi-vector virii that ran loose during the Fall, and how even many of the TITANs seem to have succumbed to these infections. They also reference a disturbing number of accounts of events during the Fall that are inexplicable … things like people being transformed into strange, alien creatures … or phenomena that seem to defy certain physical laws, as if something was at times ignoring what we know of physics and just doing whatever it felt like ... Some of these voices within Firewall even think that the TITANs may not have been responsible for the Pandora Gates ... They have a name for this mystery infection. They call it the Exsurgent virus.
\end{quotation} 

\subsection{Real and social distance}
\label{sec:real-social-distance} 

The vast distances between most habitats give all communications—with the exception of those using the rare and expensive QE communicators (p. 314)—a significant time lag between asking a question and receiving an answer. In most cases, the time lag ranges from ten seconds to several hours, and it makes real-time communications between distant habitats difficult or impossible. Communication problems only serve to further isolate habitats from one another, and as a result people socialize primarily with members of their own habitat (or habitat cluster, if their habitat is part of one of the various groupings of between two and twenty habitats that abound throughout the solar system). 

Within a habitat or habitat group, communication between residents is effectively instantaneous, thanks to the omnipresent wireless grid known as the mesh (p. 234). Anyone wearing a mid-range ecto (p. 325) or using basic mesh inserts (p. 300) can communicate with others in ways that go far beyond mere voice contact. Both devices allow AR communications that are in most ways barely distinguishable from in-person communication, so people can effectively spend in-person time with anyone in their habitat at any moment when both of them are free and interested in communicating. Unless someone deliberately wishes to turn off communication because they are sleeping or otherwise busy, people can always get in touch with one another. Many close friends and romantic partners regularly communicate anytime they have a spare moment, sharing comments and jokes. This communication is far more awkward and distant if there is a time lag of several minutes between every comment, so inter-habitat communication is never as informal or close. 

Although travel via egocasting (tramsitting an ego to another habitat, where it is resleeved) is as easy, if not as cheap, as communication, a trip to another habitat is considered to be a significant journey with a range of costs. Individuals traveling to a different habitat will no longer be able to engage in real-time communication or shared real-time entertainments with people back on the habitat they left, so the traveler will have to find a new social environment. In addition to the trouble and expense of acquiring a new morph in the new habitat, the social distance between individuals and the social network they leave behind is part of the cost of travel. 

Before the Fall, refugees from Earth were accustomed to being able to easily communicate with anyone else on Earth. Wealthier individuals could easily journey just about anywhere on the planet in a few hours while still being able to communicate with everyone back in their home city with no noticeable change. The exodus of transhumanity from Earth, though, means that an individual's social world is only as large as their habitat. Even a relatively brief communication lag, such as the two to thirty seconds that is the average time lag between any two of the Jovian or Saturnian moons, greatly hinders the flow of back-and-forth communication. When time-lags are involved, most communication consists of messages rather than any attempt at continuous conversations. In situations where a more in-depth discussion is necessary and time is limited, someone can send a fork of themselves—a digital copy (p. 273)—to hold the discussion remotely on their behalf, and then return for re-integration. Since there is already a large time lag between sending a message and obtaining any possible response, most people do not hurry to answer messages from distant habitats except in the most urgent circumstances, further isolating people residing in distant portions of the solar system. 

\begin{quotation}
\textbf{Solarchive search: singularity seekers}  Singularity seekers are those with an unhealthy fascination in so-called singularity events, such as the hard takeoff of the TITANs to super-intelligence. Some are part of a radical sect of “exhumans” who believe that transhumans are destined to become godlike superbeings and are determined to get there first. Others act on a defensive impulse, believing that the only way humanity can survive another threat from beings like the TITANs is by becoming as hyperintelligent as their enemies are. Still other singularity seekers are researchers and spiritual seekers who are frustrated with the limitations of their own minds and seek to become something greater. Some of these people become gatecrashers, searching for advanced alien artifacts to help them in their quest. Others experiment with employing conventional technologies in new and exotic ways, such as creating mentallylinked networks of forks or incorporating extra-fast and powerful computers into synthmorphs and pods. 

 A few of the most daring seek artifacts left behind by the TITANs, hoping to incorporate techniques and technologies created by these inhuman beings into their minds. This last group is the most notorious, in large part because of the spectacular nature of some of their failures. On occasion, these artifact hunters have awakened devices that have lain dormant for a decade and caused local outbreaks of TITAN technologies. These incidents have caused many people to regard singularity seekers as everything from potentially dangerous eccentrics to unknowing pawns of the TITANs.
\end{quotation} 

\subsection{The rise of cultural regions}
\label{sec:rise-cult-regi} 

The only exception to the social distance between different habitats occurs when colonies are located on or in relatively close orbit around the same planet or moon. The inhabitants of Mars can all communicate with one another instantaneously, as can everyone on Luna or in Lunar orbit. However, the rivalry between the various Martian city-states—and between the primary hypercorp domes and the rural Martian poor—imposes its own social distance. Individuals from different city-states do socialize, but among the elite social cliques, spending too much time communicating with members of another city-state is viewed as somewhat odd and potentially even disloyal. As a result, Martians tend to be relatively isolated even from their close neighbors. Nevertheless, the short distances between the Martian city-states and the orbiting habitats mean that there remains a general Martian culture that is different from the cultures of the rest of the solar system. 

Distance barriers have produced similar levels of cultural differentiation in other portions of the solar system. The colonies in the vicinity of both Jupiter and Saturn each form a separate cultural unit, as do the colonies in Earth orbit and on and around Luna. The same is true for the Jovian Trojan and Greek asteroids. In each of these regions, people communicate and travel more between habitats and settlements than they do with outside regions. 

Social scientists refer to the different sections of the solar system as separate cultural regions. The different regions of the belt also each form a similar cultural region, but because asteroids in different orbits eventually drift quite far apart, the cohesion and unity of these cultural units is somewhat weaker. Habitats on the edge of the solar system (around Uranus, Neptune, and Pluto) form very small cultural regions, but the few habitats in the Kuiper Belt and Oort Cloud have no cultural region since the distance between them is so extreme. 

Though communications between habitats within the same cultural region is somewhat awkward due to intra-regional cultural differences and small timelags, it is usually fast and easy enough for people on different habitats to keep in regular contact with one another. In addition, most habitats within the same cultural region are sufficiently close that egocasting between them is affordable by most people. In contrast, egocasting between cultural regions is relatively expensive. Many social scientists predict that within one or two decades, different cultural regions will be at least as different from one another as distant nations of Earth were from one another during the first half of the 20th century—perhaps even more so due to the physical alterations that cultures introduce as they continue to evolve. 

\subsection{Cultural experimentation}
\label{sec:cult-exper} 

While nostalgia for Earth remains a powerful social motivation, the break from Earth led many inhabitants of the solar system to experiment with new forms of culture and society. Since the Fall destroyed physical links with the past and the defeat of the last old-Earth governments ended ideological ties with the old political and social forces, many transhumans saw themselves as living in a new, free era where the past was dead. Even people who always wear nostalgia jewelry and spend several hours a day in simulspaces set on old Earth are very interested in the possibility of social and political experimentation. Those without criticisms of Earth's nation-states and their many failings still rue the day when Earth fell. 

Many of the most extreme social experimenters moved to the numerous small outer system habitats that were created in the decade after the Fall, but people interested in social and cultural experimentation can be found throughout the solar system. In addition to playing with various interior structure and design ideas, the inhabitants of many stations experiment with all manner of unique social and political rules. A few habitats do so quite deliberately, either because the members are interested in social innovation or because researchers associated with a hypercorp or university have offered them goods or services in return for testing one of their latest theories. Such experiments have included establishing stations where all of the residents are sleeved in hermaphroditic morphs in order to measure the impact on customs and language when gender is abolished or spurring the residents of a particular station to freely switch their morphs based on the responsibilities and duties they have on a given day. Such staged experiments are, however, relatively rare—the vast majority of unique customs and social structures that have come about since the Fall naturally evolved from groups of likeminded individuals living together in the same habitat and working, consciously or not, to make life better fit with their aesthetics or ideology. 

\subsection{Gender, sexuality and relationships}
\label{sec:gend-sexu-relat} 

To many transhumans, gender has become an outdated social construct with no basis in biology. After all, it's hard to give credence to gender roles when an ego can easily modify their sex, switch skins, or experience the lives of others via XP. Though most transhumans still adhere to the gender associated with their original biological sex, many others switch gender identities as soon as they reach adulthood or avidly pursue repeated transgender switching. Still others examine and adopt untraditional sex-gender identities such as neuters (believing a lack of sex allows greater focus in their pursuits) or dual gender (the best of both worlds). In many bioconservative habitats and cultures, however, more traditional gender roles persevere. 

Sexuality has also expanded into new frontiers and taboos. With basic biomods providing contraception and protections from STDs, casual sex is the norm. Many people pursue careers as well-paid companions and escorts. In fact, sexual experimentation is standard thanks to several new technologies. Virtual reality allows sexual encounters without physically touching a partner, not to mention bringing all manner of fantasies to life. For those that prefer the touch of real skin, AI-driven pleasure pods can fulfill any and all needs and are a legal form of prostitution in many habitats. Sex-switching also lends itself to new experiences, whether via bio-mods or a new sleeve. Even AGIs, having been socialized as humans, exhibit sexuality and desire. 

The extension of lifespans and the decline of religion have drastically impacted social institutions like marriage. Given the possible changes to both cognition and biology over a transhuman's lifetime, lifelong relationships are no longer considered realistic. The idea of long-term relationships as a social contract has grown exponentially. While this has resulted in a number of marriages that are political or like a business transaction, most people continue to view marriage as a bond of emotional attachment and trusts—in particular a bond that transcends bodies, as either partner may change morphs at any time. 

\subsection{The diversity of habitats}
\label{sec:diversity-habitats} 

The ability of a few thousand like-minded people of moderate means to acquire a small habitat where they can create their own society resembles the ability of inhabitants of the United States in the 19th century to set out for the West and found their own ideologically based communities. The primary difference is that creating such communities is faster and easier in the modern era. The mesh is filled with all manner of virtual communities where members hope to eventually gather the means to create their own habitats. In most cases, these are merely idle dreams; most participants are not willing to sacrifice the time and rep or money needed. Occasionally the members try, only to find out that some of the people promoting this effort are con artists. Occasionally virtual subcultures manage to raise the necessary dedication and trust to build their own habitat and begin the process of creating their own physical society. A decade of this sort of cultural experimentation by many hundreds of habitats has produced a number of unique and strange societies. 

As an example, there are habitats where the inhabitants wear garments and AR images that cover their bodies—and, in the most extreme cases, their faces—and residents only reveal their morph's true appearance to their closest friends and immediate family. There are also stations where all members use cosmetic modification to adopt the same ideal look, as well as ones where all residents use morphs that are clones of one another. Some of the most eccentric habitats are populated by extreme bioconservatives overcome with nostalgia for the past, leading them to model both their society and all visible technology after some earlier period in history, typically some time between zero and 50 years BF. 

There are even a few habitats that totally disregard commonly held feelings about forks and merging. Such community members regularly split off multiple forks when they awaken and plan their day and then merge the various forks when they go to sleep that night. Some forks remain infomorphs for the day, while others use one of the various morphs the individual owns or rents, which means that each resident typically lives between two and six separate lives every day. A few societies, like the home of the infamous Pax Familiae, go even further—all residents are forks of the same individual. In some of these solipsistic habitats, the forks are all expected to use cloned morphs, while in others each fork is considered a separate person who should go and forge their own unique life. Some of the less extreme manifestations of this type of habitat include places inhabited by families that are partially or entirely composed of forks of one of the members (the various forks tend to be treated as siblings). 

\section{Technology}
\label{sec:technology} 

Technology pervades all aspects of existence in Eclipse Phase. Most individuals understand that unless humanity suffers another event like the Fall or they personally suffer some very serious and unlikely accident, they are unlikely to permanently die. More people are now planning for a very long future. For most people these schemes are fairly minimal, but they often include an awareness that few, if any, relationships are likely to last an entire lifetime. However, functional immortality is only one of the many wonders of the modern world. 

\subsection{Living with infotech}
\label{sec:living-with-infotech} 

For anyone with basic mesh inserts (p. 300) or an ecto (meaning about ninety-six percent of the population), life is filled with data. For people with the best implants, all information available on the mesh is available at a thought. For everyone else, it only requires a brief pause to access and understand it. When someone pauses and looks a bit distracted in the midst of a conversation, everyone understands they are accessing data and lack the implants to allow them to do this subconsciously or via multi-tasking. As a result, when a group of people are discussing a topic and no one immediately knows an answer to a question, such as the title of a performer's first vid, within a few seconds everyone has this information. Similarly, when someone walks through a garden, with a glance and perhaps a brief thought or small finger motion, they can call up detailed data on each and every species of plant that sits in front of them. Individuals going to remote areas that are out of normal mesh broadcasting range almost always either carry a farcaster-link with them or download truly vast amounts of data into their implants or ecto so they can continue to access all the data they might need. Since even a basic implant can hold vast amounts of data, lack of storage space is rarely an issue. 

Access to such a vast amount of easily available information has resulted in a variety of cultural responses. Being able to quote from any vid, old movie, book, or historical speech is now trivially easy and can be done with a few seconds of thought. While children and young teens often play by interjecting large amounts of semi-appropriate famous quotes in their speech, most adults only do so for emphasis and in moderation. People who quote from other sources too often are considered dull and unimaginative. Recognizing such quotes is quite easy, since someone can simply set their muse to alert them to the nature and identity of all lengthy quotes they hear. 

All experienced mesh users also learn (typically as children and teens) how to avoid taking too much time out from conversations to check facts or access information via the mesh. Teens regularly mock their fellows who pause too often or too long in conversations to look up further information on a topic someone mentioned, or who spend too long trying to assemble facts to support an argument. Terms like “meshed out” or “drooler” are used by teens to mock each other into learning how to be both discreet and faster in their information searches, at least when also interacting with others. While adults rarely engage in the same sort of direct and obvious mockery, people who get too lost in casual or conversational meshbrowsing are widely viewed as socially inept. As a result, implants that allow multi-tasking or temporarily speed up thought are in great demand, since they allow individuals to do extensive research and rehearse each statement they are going to make without a moment's pause. People who can afford such software almost always seem more suave, charismatic, and intelligent than those who do not. 

All this means that those who lack all mesh and AR access—individuals known as zeroes—present a stark contrast to the rest of transhumanity. To most people, zeroes seem slow, forgetful, and almost unbelievably dense, while to zeroes, even people who only possess ectos or basic implants seems brilliant, witty, and able to comprehend things with almost inhuman speed. 

\subsection{Opening Pandora's gate}
\label{sec:open-pand-gate} 

The discovery of the first Pandora Gate on Saturn’s moon Pandora shortly after the Fall was a watershed moment in transhuman history. The prospects this discovery raised were simultaneously fascinating and terrifying. On one hand, technologies far beyond anything transhumanity was capable of were now in our hands. This raised visions of a horizon far beyond the horrors of the Fall, where transhumanity would expand across the cosmos, visiting wonders that seemed perpetually far out of reach, even for nearimmortals. On the other hand, the possibility that these gates were relics of the TITANs could not be discounted. Their existence opened the possibility that the TITANs might one day return, or that transhumanity might still encounter them out in the galaxy at large. The alternative was even scarier—that the gate could be of extraterrestrial origin, and the things more dangerous and frightening than the TITANs might stalk the space between the stars. 

Various hypercorps, governments, and other factions threw their brightest minds into solving the mystery of these “wormholes.” Numerous scientific communities pooled resourcesbacked by private sector fundsand cracked the code of the Pandora Gate in just over a year. Not only was the gate activated, but it could be programmed to open connections to numerous distant star systems (one at a time). Though these controls were unreliable at best—connections sometimes closed without warning, and others could not be recalled though they had been opened before—the functionality was stable enough to use them in earnest. At the same time as their very public announcement concerning this seminal achievement, the Gatekeeper Corporation was formed overnight: a merger of those same scientific communities and their financiers. 

Less than a year from its first operation, the hypercorp opened the gate to “gatecrashers:” explorers who risk their lives to see what lies beyond. Many of these died horribly; some were even lost forever, but a few made fantastic discoveries such as new worlds and new life. Though none of the (living) alien lifeforms encountered so far have been sapient, many of the worlds are habitable or within the possibilities of terraforming. Along with these wonders were found more disturbing things: evidence of a long-dead alien civilization (the Iktomi), and signs that the TITANs had passed these ways before. 

Additional gates were soon discovered throughout the system. Unlike the spirit of cooperation that surrounded the first gate’s discovery, these others were seized as hotly contested resources. Initially used for research and exploitation, many of these gates are now being tasked for colonization purposes. Dozens if not hundreds of exoplanet stations and colonies have been established, some with significant numbers. There has been no lack of poor or desperate individuals willing to risk life on an alien world, if it means an iota of improvement in their lives. 

Though it is now widely accepted that the gates are the means by which the TITANs evacuated the solar system (a hypothesis which fails to answer why they did so), they appear timeless in their construction. Regardless of their origin, the gates remain one of the most prized and dangerous of technologies. 

The five known Pandora Gates within the solar system, their locations, and their controlling entities, include: 

\begin{itemize}
\item Vulcanoid Gate: Caldwell (Vulcanoids) —- TerraGenesis
\item Martian Gate: Ma’adim Vallis (Mars) -- Pathfinder/Planetary Consortium
\item Pandora Gate: Pandora (Saturn system) -- Gatekeeper Corp.
\item Fissure Gate: Uranus -- Love and Rage Collective/Anarchists
\item Discord Gate: Eris (Kuiper Belt) -- Go-nin Group/Ultimates
\end{itemize} 

\subsection{Going beyond the known}
\label{sec:going-beyond-known} 

One of the oddest experiences for gatecrashers and others who explore unusual environments such as the ruins of Earth is the unavailability of data. They look at an alien plant or a TITAN-mutated person, and their search returns various error messages meaning that there is either no data at all on the subject or that the only data is purely speculative and should be regarded as dangerously unreliable. This can be especially troubling when the subject in question is a small creature that has just landed on the person's shoulder and the individual wants to know if it's harmless or deadly. Most people who are less than sixty years old have never been in an environment where they could not gain basic information about everything around them at a glance. Learning to overcome the shock of not knowing anything at all about something is one of the first and most crucial skills all gatecrashers must learn. 

\subsection{Muses}
\label{sec:muses} 

Most individuals have a dedicated AI that serves as their media agent. Commonly known as a muse, this AI has been a lifelong companion for most people less than seventy years old. Muses learn their owners' tastes, habits, and preferences, and do their best to make life and technology use as easy as possible. Muses can be alarm clocks, data retrieval gophers, appointment schedulers, accountants, and many other functions often limited only by their owners' imaginations. Some of their tasks do not even need to be assigned them—muses are skilled at figuring out people's needs and acting on them. For example, the muse's scheduling function may tell it when its user needs to be up in the morning, and it will act as an alarm clock without any additional instructions from the user. If a muse is uncertain about its owner's preferences, it asks, but after working with a user for a few decades muses rarely need to do this. Most people keep multiple back-ups of their muse, because the loss of a muse can be almost as traumatic as the death of a loved one. Using a generic muse who must be informed about all aspects of a user's individual preferences and fed a constant stream of instructions helps people appreciate the value of their own personal muse agent. Muses generally learn the basics of a new user's preferences in a month or two, but during that learning period the user tends to be irritable and forgetful, since the tasks they generally trust their muse to do automatically are not being taken care of. 

\subsection{Attitudes towards AGIs}
\label{sec:attit-toward-agis} 

The vast majority of transhumanity blames the Fall on rogue seed AIs (self-improving artificial intelligences). As a result, any AIs that are not crippled or somehow limited from improving themselves—including the AGIs (artificial general intelligences) that were common and growing in number before the Fall—are completely illegal in many habitats, or at least heavily regulated. The Fall ended only slightly more than a decade ago, and many transhumans consider AGIs and the TITANs that murdered their homeworld to be one and the same. 

In addition to strict anti-AGI laws, there have been occasional riots and mass panics surrounding facilities still performing AGI research, which has pushed most such research into isolated settlements. Nevertheless, there are still people passionately devoted to AGIs; some see them as the next step in posthuman evolution, others value all sentience, and still others actually worship them. However, AGI supporters have learned to keep their opinions private in mixed company, lest they be branded an agent of the TITANs. 

In some spots, mostly in the more anarchistic outer system, attitudes towards AGIs are more relaxed and AGIs may even be openly welcomed. These places recognize that AGIs are not the same threat posed by seed AIs and it is unfair to punish one for the actions of the other. Naturally, these places are havens for the AGIs active in transhuman society, who otherwise must disguise their true natures. 

In the tightly-controlled inner system, the hypercorps and the Planetary Consortium foster anti-AGI sentiments both as safety measures and as protection against possible competitors. This latter point is one of the things that makes them attractive to some people in the outer system; they understand the great advantages their factions gain ... assuming, that is, that those AGIs share your goals and ideals. 

\subsection{Attitudes towards mental alterations}
\label{sec:attit-towards-ment} 

In the post-Fall solar system, technology can alter people's minds; controversy about many of these alterations remains. Few people have trouble with the idea of creating short-term forks using the multi-tasking augmentation or some similar process that insures the forks will be re-integrated within a few hours. However, the idea of long-term forks, and especially of allowing forks to gain access to their own separate morphs, troubles many people. Since there are not enough morphs to go around in the first place, providing morphs to a fork strikes many people as selfish and wasteful. As a result, on the rare occasion that people sleeve one of their forks, they typically provide it with a synthmorph to avoid the social stigma associated with using more than one body at a time. 

Forks that exist for more than a few hours inspire discomfort in many people because the forks begin to diverge slightly in personality. Most people find the idea of two different and distinct versions of themselves to be somewhat disturbing. While there are habitats (mostly in the outer system) where forking is a regular part of daily life and forks often exist independently for a day or two, most visitors find such habitats distasteful and bizarre. 

However, while voluntary forking is still regarded as somewhat odd, involuntary uses of this and the associated mental technologies are so horrifying that they form the basis of much lurid crime fiction. Someone being unknowingly mind-napped and having an involuntary—and often secret—fork created is something that people regard with abject terror, despite it being quite rare. Similarly, while mental surgery to correct psychiatric problems or as punishment for various serious crimes is frightening and disturbing in its own right, illegal brain hacking draws horror and disgust from almost everyone in the solar system. Penalties for involuntary forking and mind hacking are exceptionally high. In many habitats, they are among the few crimes punishable by death (including the destruction of all backups and forks). 

\subsection{Travel}
\label{sec:travel} 

Travel between habitats and other transhuman colonies is both exceedingly easy and fairly costly. Longrange egocasting is expensive, as is acquiring a morph at the destination. Travelers have developed various ways around this obstacle; for example, if someone only needs to visit another habitat for a few days and is visiting primarily to engage in real-time communication, they often choose to remain an infomorph for the duration of their visit and to communicate via AR, thus saving all resleeving expenses. For visitors who require a morph but will not be staying long, most habitats offer the option of renting a generic splicer or synthmorph or, for a slightly higher cost, a generic exalt morph. Habitats or worlds with unusual requirements, like Mars, Europa, or the various zero-g habitats offer ruster, aquanaut, or bouncer morphs instead of splicers. These morphs can be used for up to a week without much difficulty, and using one for up to a month is usually possible with sufficient negotiation and payment. Meanwhile, the traveler's previous morph is kept in medical stasis back in their home habitat, waiting for their ego to return. 

Another technique is morph trading by people from different habitats who know each other and who are traveling at the same time. A few people do this with strangers they meet on the mesh, but vids and other entertainments are filled with tales of people having their morphs or their identity stolen. A few of these horror stories are based on actual accounts. Very few people are willing to let anyone they do not know and trust use their body, and many people simply will not lend out their morph to anyone at all. 

Some people, however, are willing, for a fee, to act as a living “taxi” for a visiting infomorph, carrying it around with them. In these cases the “ghostriding” infomorph is not permitted to control their host's morph directly and is simply a passenger along for the ride, issuing directions and communicating with their transporters electronically. 

Travelers who wish to either immigrate to a new habitat or visit one for several months or longer must acquire their own morph. Usually, they reduce the cost of acquiring a new morph by selling their previous morph to a body bank. Alternately, some individuals sleeved in expensive custom-designed morphs who are traveling relatively short distances will rent a generic shell for several weeks and arrange to have their old morph shipped to them on a fairly rapid freighter. Doing this is rarely more than a moderate expense, which makes it less expensive than the costs of buying or replacing high-end custom modified morphs. 

\subsection{Privacy}
\label{sec:privacy} 

Privacy is a prized possession for most inhabitants of the solar system, but it is so rare that for many people it might as well be a foreign concept. In the 20th and early 21st century, privacy consisted of two concepts that are now completely separate—the ability to remain unnoticed or anonymous and the ability to avoid unwanted intrusion. The first is largely absent from the lives of most people in the present day. Anyone who uploads anything to a non-private portion of the mesh understands that anyone who wishes to do so can gain access to it. Likewise, anyone who spends time in a public place understands that anyone can learn where they went, what they did, and what they said due to the ubiquity of meshed, sensor-enabled devices. As a result, everyone's public life, both on the mesh and in person, can be transformed into an easily searchable database. Almost everyone keeps such a record of their own lives, commonly known as a lifelog. Most people allow their lifelogs to be public, understanding that anonymity is now an archaic concept. 

\begin{quotation}
\textbf{Solarchive search: incapacitating inputs}  During the Fall, the attacking TITANs used a variety of AR and online intrusions that interfered with or even incapacitated their targets. The most basic of these were deceptive AR illusions made to convince people that their physical environment was very different from what it actually was. This fooled people into attacking their fellows or simply instigated mass panic. More advanced versions targeted the empathic elements of AR, triggering fear or other emotional responses. Still others blasted their targets with overbearing sensory input, so strong that it bypassed filters and inflicted neurological damage. 

 Despite rumors and fears of socalled “basilisk hacks”—visual or other sensory-input attacks that allegedly subverted transhuman minds by exploiting the way brains processed such data—no credible reports have been verified.
\end{quotation} 

While the interiors of private dwellings remain free from continuous surveillance, almost all habitats have emergency sensors in every building providing a full record of events to emergency service workers and AIs in case of problems such as dangerous chemical leak, a sufficiently large fire, an explosion, loss of air pressure, or some other equally dramatic and potentially dangerous event. Both the events of the Fall and the fact that almost all of humanity now lives in habitats surrounded by hostile environments mean that such sensors are standard fare. A few habitats do not allow emergency sensors in private dwellings, but most people regard these habitats as potential death traps. These emergency sensors do not record anything other than the absence of potential dangers if they are not triggered by specific events. This limitation allows individuals privacy within their own residences—as long as they are certain no one has planted a secret recording device in their home. Ultimately, remaining unobserved is a matter of both care and trust, and everyone understands that most of the time everything they do will be part of the vast public record. 

In vivid contrast, the freedom to avoid unwanted intrusion is carefully prized by the inhabitants of the post-Fall era. Unwanted personal or data intrusion into someone's private dwelling or personal electronic files is a crime in most habitats and a serious crime in many. Also, while both the mesh and augmented reality are filled with all manner of AI-mediated adware, most of it has evolved to be relatively benign and to provide non-intrusive suggestions about goods, information, and services that are likely to be of legitimate interest to the targeted person. An individual's muse filters out unwanted advertising. While it is certainly possible to create advertising that can hack through any muse's filters, doing so is usually illegal. 

Unwanted AR intrusions are similarly limited. During the early days of AR technology, there were serious problems with users being overwhelmed with unrequested and distracting input—as many said, the mist got very thick indeed, so both law and custom changed to prevent such invasions. Today, most people expect to only experience data that they are looking for or that they might be interested in, and that any data they are not interested in will quickly vanish. Being surrounded by a large amount of unwanted AR data is not just annoying and distracting, it is also deeply frightening, because it means that there is a serious problem with either the habitat's mesh or the person's electronics—it could even mean that the entire habitat is under direct attack by infowar weapons. 

\subsection{Low-tech existence}
\label{sec:low-tech-existence} 

More than ninety-five percent of humanity inhabits artificially created morphs. Most of them also possess basic implants, and the vast majority of the rest wear ectos with retina displays and other simple peripherals that allow the user to fully perceive and interact with the vast network of information around them. However, slightly less than four percent of the remaining population inhabit flats or splicer morphs without basic implants and also lack access to ectos and other basic technologies. 

Since an ecto is both a relatively trivial expense and a piece of equipment vital to existence in the solar system, the only individuals who lack such technologies stand on the very lowest rungs of the social ladder. A few are the poorest members of the most marginal habitats, but most are slaves or the next best thing to them. The lowest social classes in the Jovian Republic lack personal infotech access and so do the lowest class of people indentured to the hypercorps and the Planetary Consortium, particularly on Luna and Mars. These individuals are either indentured criminals or people sufficiently lacking in useful skills that they are assigned mindless physical tasks that cannot be more efficiently performed by AIs. 

The lack of mesh access makes these unfortunate “zeroes” into mental and social cripples, unable to perceive the vast wealth of AR that most people take for granted. They are also unable to communicate with anyone beyond the range of their voice or to access almost all information, including traffic signals and shop displays. When necessary, the managers and overseers in charge of groups of zeroes allow them access to handheld meshbrowsers. These devices resemble the handheld terminals common in the early 21st century and have limited functionality, typically forbidding communication and restricting mesh research to carefully filtered topics. 

Because of their inability to access AR or the mesh, zeroes are almost completely isolated from everyone else, meaning they are also unable to organize effectively or to otherwise cause trouble for the people who control them. In much of the outer system, the existence of zeroes is considered one of the greatest crimes against transhumanity perpetrated by the Planetary Consortium and the Jovian Republic. 

\subsection{Life, death and morphs}
\label{sec:life-death-morphs} 

While death is no longer a certainty for transhumanity, it remains a possibility. During the decade preceding the Fall, most of humanity was growing used to the idea that immortality was in their grasp. Then, in just a few short years, the TITANs wiped out more than ninety percent of us. Faced with the horror of so much needless death, efforts to insure the lives of surviving humans became a top priority. Now, the technology of immortality—uploading, cortical stacks, and other related wonders—is commonplace. 

Today, most of the residents of the solar system have adjusted to this fact (except for the most extreme bioconservatives); everyone expects both to live forever and to have their friends, loved ones, and enemies do the same. While death is rare, though, it is still possible. Severe accidents can destroy someone's cortical stack as well as their brain, and egos can also be wiped away in punishment for sufficiently heinous crimes—though the process of execution is considerably more difficult than it had been a few decades earlier. 

For most people (with the exception of those too poor to afford a new morph), non-permanent death is an annoyance equivalent to events that most people in the late 20th century regarded as moderate misfortunes, like a bad stomach flu or a broken arm. In almost all habitats, if anyone is responsible for someone's temporary death, either accidentally or on purpose, they are also responsible for paying for the person's resleeving in an identical morph, especially if that person does not have some form of resleeving insurance. People who have temporarily died can expect to receive visits from everyone they are at all close to after their resleeving, as well as a host of e-cards and perhaps a few gifts from their acquaintances and colleagues, all expressing sympathy at their death and welcoming them back to the world of the physically embodied. Exchanging such “life gifts” is an accepted part of belonging to many professions such as emergency service workers, where members regularly risk temporary death. 

Deliberately choosing to change morphs or to temporarily become an infomorph is treated differently. People typically spend at least a day or two between deciding to change morphs and actually doing so. During this time, it is considered polite for someone to inform everyone they know well or work with about their upcoming resleeving. Along with personal visits, as well as calls and e-cards detailing the time of the upcoming event, the person who is resleeving is expected to include an image of what their new morph will look like, so people they know will be able to easily recognize them. However, it is considered gauche for someone who is upgrading to a better morph to include details about their new morph. Within a few days of resleeving, a “resleeving party” is typically held to introduce everyone they know to their new morph. Depending upon how well-off, well-known, and social the individual is, these parties range from lavish affairs held in hotel ballrooms to small intimate gatherings in the person's home. 

Permanent death is treated very differently. Because it is both relatively rare and no longer expected, the old funerary rituals surrounding death have faded and new traditions have grown in their place. Since every death reminds many people of the billions who permanently died during the Fall, most of the few funerals that are held honor both the person who just died as well as the victims of the Fall. 

\subsection{Entertainment and media}
\label{sec:entertainment-media} 

A substantial amount of media survives the Fall of Earth, and a significant number of modern transhumans make their living creating new songs, stories, reports, or other media. All of this is easily and swiftly accessible through any basic implant, ecto, or (on very rare occasions) archaic handheld terminal. However, most of this media is not to the taste of any particular individual, and vast amounts of it are mediocre. As a result, most humans keep two layers of evaluation between them and anything they might consider exposing themselves to. 

The first layer is based on popularity and critical reviews. Every piece of media has a rating, often weighted by the opinions of critics with high rep scores who comment on their virtues and faults. Specialized AIs also evaluate the responses of consumers, so individuals can use reviewers they trust or they can seek out media that is either widely or specifically popular in their particular demographic and subcultural niche. 

The second filter layer is the individual's muse. Muses learn their owner's tastes and moods and automatically search out and recommend various sorts of media. Individuals can do everything from asking their muse to select something they will enjoy, to asking for a something that will challenge their opinions, to looking at all current events news that will be of interest to them. Muses use their understanding of their user's preferences, mixed with ratings and reviews, to make their decisions. Individuals can even set their muses to edit all media so that they better fit with the person's interests and preferences. In the most extreme cases, this process can twist and edit news so that it bears no relation to real events. This same process is used to make the characters and dialog in novels and vids more appealing. More commonly, the muses merely edit out aspects of a news story or article in which the individual is not interested. 

Ratings, reviews, and muses allow individuals to avoid media overload, but they also reinforce subcultural barriers. A great many people only seek out media and news that reinforces their existing opinions and beliefs. Xenophobic individuals who distrust all non-humans, from uplifted octopi to the Factors, regularly view news stories and AR dramas about evil aliens and devious uplifted animals who commit heinous crimes. Similarly, individuals who are only interested in their own habitat have all external news altered by their muses so that it refers only to the effects outside events will have on their station. 

In a very real sense, individuals from radically different subcultures and demographics inhabit completely different worlds. The one force that works against this separation is the fact that many people wish to follow the lives and opinions of those with the highest reputation scores. In many cases, a large portion of these individual's high rep scores comes from their interest in and willingness to interact with (or at least acknowledge) a wide variety of different sources of information. As a result, listening to opinions by a high-rep celebrity can expose people to information that they might never encounter otherwise. Also, in many habitats, AIs responsible for media distribution tag some news as being sufficiently important that it should be immune to filtering by muses. 

This tagging is a regular and expected occurrence in some habitats, while in others it is reserved for only the most important and potentially life-saving information. Bypassing muses for any less important reason in these stations is considered a gross invasion of privacy or even a crime. 

\subsection{Lost lore}
\label{sec:lost-lore} 

The accumulated knowledge and media of Earth, spanning the history of human intelligence, is a vast and impressive amount. Even before the Fall, many orbital settlements had acquired complete records of all previous human lore and creativity, including copies of every book, painting, song, film, TV program, console game, newspaper, and magazine article that had ever been translated into digital format, as well as backups of Earth's entire internet archives. Numerous destructive programs unleashed during the Fall corrupted much of this information, however, in some cases permanently wiping it from existence. This means that what remains of Earth's archived history and data is patchy and incomplete. Much survives, but some treasures have been lost. In particular, media from the era of the Fall itself is particularly hard to come by, given the consistent attacks the TITANs were making on information systems. Proprietary data that was withheld from the public domain behind electronic gates on Earth is even more likely to have been lost, except for a few hypercorps that managed to transfer their Earth-bound data off-world in time. Retrieving lost data is a lucrative task for scavengers and archeologists, though looting the dangerous confines of Earth or derelict habitats destroyed during the Fall is a risky proposition. 

\subsection{Metacelebrities}
\label{sec:metaceleb} 

As the culture industry quickly discovered, biotech and resleeving technology clashed with the media’s ability to focus the spotlight on specific icons. When everyone can be bodysculpted, the beautiful people need to be more than glamorous faces. More to the point, the public’s interest in celebs faltered when famous people repeatedly changed their looks and were no longer immediately recognizable. 

One of the ways big entertainment has responded is to promote metacelebrities—icons based on characters rather than real people. Each metacelebrity has their own (very expensive) unique customized morph, but the person sleeved within that morph often changes. The actress Angelique Stardust, for example, once existed as a real person, but is now a character who has been played by over a dozen people since the original rose to stardom in AF 3 and promptly sold off her celebrity character rights to Experia. Likewise, award-winning heart-breaker Juan Nguyen is a constructed persona based entirely on the action hero star who died and was lost during the Fall. Many metacelebrities are modeled on fictional characters; notorious bad girl Sun Mi Hee is no different offscreen than the ass-kicking villain role that brought her to fame, never traveling anywhere without her iconic pair of glowering smart leopards. Actors taking on a metacelebrity role often undergo psychosurgery to better play the part. 

Metaceleb personas are strictly managed and marketed as a media product to appeal to specific consumer groups. Though they play an active role within hyperelite circles, many of the genuine glitterati view them with humor at best, disdain at worst—though some have learned the hard way not to underestimate or mess with the small armies of media engineers behind each metaceleb’s carefully crafted image. 

\subsection{Popular types of entertainment}
\label{sec:popul-types-entert} 

The most popular forms of electronic entertainments are vids, vidgames, VR worlds, XP, and AR games. 

\subsubsection{Vids and vidgames}
\label{sec:vids-vidgames} 

Vids are passive entertainments that can be enjoyed either as high-resolution audiovisual entertainment or as a fully immersive experience where the viewer can augment their experience with smell, touch, and taste while experiencing the point of view of one of the major characters. Viewing them purely via sight and sound is much like watching an old 20th-century film, except that it's interactive and in 3D. In contrast, full sensory viewing is much like actually being present in the story. 

Most modern vids have variable theme and preference settings enabling viewers to adjust the content of what they are watching, including the level of violence, the amount and type of sexuality they prefer, as well as the appearances of some or all of the major characters. In addition, many vids have several alternate endings for people who prefer happy, bittersweet, or grim endings. As a result, two people watching the same vid could have very different experiences if they use radically different settings. 

Vidgames are like vids, except they are much more flexible. In vidgames, the viewer not only experiences the story with the protagonist—they become the protagonist, shaping the story through their own actions, similar to sophisticated early-21st-century console games. Some games allow the participation of up to a dozen individuals or link thousands of players via the mesh, while others are designed for a single player. 

The degree of freedom in vidgames varies. Some are almost fully interactive realms similar to many VR worlds with all but a few characters controlled by AIs, while others are considerably simpler and more limited, with player interaction limited to a few crucial decisions. The precise dividing line between vids and vidgames is blurry, but both media have the common trait of being designed for either solitary use or for use by a few players or viewers who are all located relatively near one another. Vids and vidgames are the most popular forms of entertainment, with vids and vidgames set on Earth before the Fall being especially prevalent. 

\subsubsection{XP}
\label{sec:xp} 

Experience playback (XP) is a specialized type of vid that consists of the recorded sensory impressions of a single individual. Almost all of the inhabitants of the solar system lead relatively quiet and risk-averse lives and are naturally eager to be able to vividly experience adventures such as climbing Olympus Mons, spending a day in one of the most luxurious and exotic private habitats, going on a scavenging mission to Earth, or gatecrashing. There is also a thriving fringe market in less savory XPs, including records of people committing all manner of violent or dangerous crimes and XPs of actual gun battles between well-armed criminals and law enforcement personnel, which often end with the death of the morph providing the point of view. 

Anyone with mesh inserts can create an XP of their past experiences, and anyone with an ecto or mesh inserts can access the sensory recordings. Selling a particularly exciting XP, such as a record of the first meeting with the Factors, can bring in a lot of money or rep. Most XPs consist of both sensory recordings and the surface thoughts of the individual who made them. Many people who access XPs are only interested in the sensory recordings and feel that having another person's recorded thoughts and emotions in their head is intrusive and uncomfortable. However, some hardcore XP aficionados feel that accessing the full XP, including the recorded emotions, makes the experience more immersive and real. 

A significant minority of XP fans becomes fascinated with one or two daring people who regularly sell XPs, known as X-casters, viewing all of their clips, including both the experiences and the accompanying thoughts. Some of these XP fans become more interested in the person who recorded the clip than in the individual experiences, and they often come to believe that they have a special, clear understanding of this person, to the point where they strongly identify with this person or even fall in love with them. In addition, individuals who access XPs from a single person often enough sometimes begin to mimic various habits or figures of speech of this person. Particularly popular X-casters are sometimes rather disturbed when they see tens of thousands of people imitating one of their more idiosyncratic expressions or habits. 

A few serious fans—known as Xers (pronounced “ex-ers”)—alter their morphs to resemble their favorite X-caster. Some obsessive Xers actually attempt to contact and stalk certain X-casters, perhaps hoping to become part of an actual XP clip. In most habitats and subcultures, Xers are widely regarded as having particularly dull and meaningless lives. Hardcore Xers are often viewed as being insecure and potentially unstable. 

\subsubsection{AR games}
\label{sec:ar-games} 

Augmented reality (AR) games involve players interacting both with events in the physical world and with augmented reality imagery that recasts the people and objects the players see. For example, instead of seeing another player in a splicer morph and ordinary clothing, a player of an AR game might see a horrific rotting zombie, a bizarre alien life form, or a well-armed soldier. These games tend to be locally focused within a particular habitat or city as they allow players to interact when they are within physical proximity, but some games link habitats within the same cultural region. 

The nature and intensity of these games varies widely. Some are long-term games involving people imagining that they are deep cover spies or some other exciting and unique role. Players may pretend to be anything from time travelers attempting to prevent some horrible disaster to covert agents attempting to uncover plots by TITAN-infected people on their habitat—who happen to be camouflaged as snack designers, personal assistants, etc. During their daily lives, players exchange messages with each other as well as with the people running and maintaining the game. Some of these long-term AR games have gone on for many years, with the oldest being almost twenty years old. 

Short-term AR games, on the other hand, last between several hours and several days. The people running these games typically rent out a hotel or a park and various public buildings for the duration. These games are almost always highly dramatic and consist of everything from the players having to deal with a massive zombie attack or alien invasion to them participating in some simulation of an event on Earth, like the storming of the Bastille during the French Revolution. While such AR games can be considerably less detailed than VR worlds or vidgames, many players value the “realism” of being physically present during the game. 

Since participants in AR games take actions in the real world, including actions that could be disruptive or even dangerous, designers of AR games take great care to prevent problems. In some early AR games, most of which took place more than twenty years before the Fall, players were occasionally seriously injured. A few unscrupulous AR game designers used their game as a cover for an actual robbery or act of terrorism that was abetted by unwitting players who thought their actions were simply part of a game. 

Since that time, law enforcement observation drones have kept careful track of people playing AR games. In almost all habitats, people running AR games must register their games with local law enforcement or face serious fines. 

\subsubsection{VR worlds}
\label{sec:vr-worlds} 

Virtual reality (VR) worlds are entertainments that involve the creation of a large and highly immersive simulated environment—a simulspace—where many major characters are played by transhumans or other sentient beings. Unlike vids or vidgames, simulspaces are specifically designed for a large number of participants. VR worlds consist of everything from duplicates of various eras of Earth history to elaborate and strange fantasy worlds with magic, dragons, and similar wonders. All manner of alien worlds or settings based on oddities like time travel are also common. As is the case with vids, the most popular simulspaces are those set on Earth some time before the Fall. 

VR worlds can have from dozens to tens of thousands of participants. For the best experience, many users prefer to access simulspaces through hardwired server connections as they offer better quality and less disruptions than accessing wirelessly via the mesh. Since people immersed in virtual reality are cut off from their bodies and often thrash around, most users ensconce their morphs in a tank or special couch for the duration. VR parlors typically offer private hardwired pods for participants to physically jack in. Many habitats also have hardwired systems used just for this purpose, so users can experience VR from the comfort of their own dwellings. 

Due to distance and communication lags between habitats, even the most popular online simulspaces run each habitat as a separate realm, limiting interaction with users in other habitats/realms. The popularity of VR worlds like Gilded Empire, set in England in the 1880s, means that someone moving from one habitat or world to another could continue playing in the same game, albeit with a new set of players. 

One of the other unusual features of VR settings is that a large number of infomorphs, including many infomorph refugees, play these games. As a result, while even most novice players can learn to easily tell the difference between a character played by an AI and one played by an actual person, there is no way to know if the person playing a character has a physical body or not. 

\subsubsection{Physical entertainment}
\label{sec:phys-entert} 

In addition to a vast array of electronic and electronically-mediated entertainments, people also still enjoy a wide variety of physical sports, ranging from soccer to new sports like low-g air races, where the participants strap on wings and engage in tests of speed and acrobatics. In addition, the ability to both heal any injury in a healing vat and to remove a cortical stack from a dead or dying body and place it in a new morph has given rise to a new variety of extreme sports. Starting a decade before the Fall, various individuals began realizing that, barring unlikely circumstances, they could not die unless they wanted to. This set off a brief trend in extreme sports and even a few wealthy suicide hobbyists, who repeatedly killed off their current morph in a variety of unusual ways. The Fall and the permanent death of more than ninety percent of humanity greatly reduced the interest in playing with death for many years. Killing yourself just to experience death is considered at least mildly distasteful, and many believe such actions belittle the mass deaths of the Fall. Though interest in risking death in the line of entertainment has been growing, deliberate suicide remains an eccentric and dubiously regarded hobby. 

In some subcultures, dueling has been a popular fad for almost a decade. Swords, knives, and pistols firing single-shot soft lead bullets are all popular choices, because none of these weapons poses any threat to a cortical stack and most do not instantly kill someone hit by them. However, there are other more exotic options, including aerial duels with microlights fitted with blades on their wings. On rare occasions, duels take place in space, with the participants wearing non-armored vacuum suits. Certain criminal groups make money with underground dueling circuits, pitting biomorphs against robots against uplifts. The seedier circuits engage in distasteful pit fights featuring illegally-acquired backups sleeved into non-sentient animals, often outfitted with lethal cybernetics. Such creatures are typically quite mad. 

In addition, dangerous non-combative sports are also popular. The highest levels of competitive rock climbing on Mars are regularly done with no safety equipment. There are similar climbing competitions in many habitats using artificially constructed climbing walls as well as regular free-running competitions through almost every city and habitat. Also, there is an entire class of sports, including both diving and parachuting, where perfection of form is seen as a far more important goal than avoiding injury or even death. As a result, current high dive records for morphs not specially modified to survive high impacts are held by individuals who required either time in a healing vat or resleeving immediately after their successful breaking of a previous record. 

\section{Politics and power}
\label{sec:politics-power} 

Politics is just as important in the colonies spread throughout the solar system as it was back on Earth, but it is also radically different. Each habitat or cluster of stations is a separate political entity, and many of these habitats are fiercely independent. The only locations where large political entities can exist are on the marginally habitable worlds of Mars and Europa, and the population of Europa is significantly smaller than that of many of the largest pre-Fall cities on Earth. 

\subsection{The inner system}
\label{sec:inner-system} 

Though nations no longer exist, they have been replaced by new political-economic entities that may well have been on the road to dominance even if the Fall had not occurred: the hypercorps. While there are a many independent habitats and settlements in the inner system, it is largely dominated by the hypercorps. To reduce conflict between themselves and promote the survival of transhumanity, some of the hypercorps have formed an alliance known as the Planetary Consortium. This alliance governs most of Mars and is in charge of the ongoing Martian terraforming project. It also controls several dozen other habitats and many Lunar bases, mostly ones that are in some way involved with the massive Martian terraforming effort. 

Since Mars is home to more than forty percent of the surviving transhuman population, most of the human population lives under the rule of the hypercorps or the Planetary Consortium. In the aftermath of the Fall, the hypercorps established three important goals: rebuilding the solar system, protecting themselves from any further attacks (either by the TITANs or any other threats), and growing in both wealth and power. By extension, the second goal means they also help protect the people living in the habitats and settlements against any repeat of the Fall. The hypercorps and the Planetary Consortium are exceedingly skilled at attaining all of these goals. Since popular rebellion and widespread dissent are not helpful in the least in attaining these goals, the hypercorps are also adept at making certain the inhabitants of the habitats and planetary settlements they control are safe, relatively content, and, ideally, unable to cause serious problems. 

As the largest and most well organized entities in the solar system, the hypercorps, and especially the Planetary Consortium, are in an excellent position to protect the people living in their habitats and settlements. However, this protection comes at the price of freedom. Living in habitats that use transitional economies (p. 61), the inhabitants of hypercorp-controlled settlements are relatively well off and need not fear starvation or serious want. Also, the hypercorps strongly oppose bioconservatism, and so anyone who can afford various augmentations or morphs is free to obtain them, as long as none of these augmentations or morphs is equipped with weaponry that can be used to harm the habitat or large numbers of its inhabitants. In return for safety and relative prosperity, however, inhabitants give up any ability to voice more than token criticisms of the hypercorps of the Planetary Consortium. 

\subsubsection{The power of the hypercorps and the planetary consortium}
\label{sec:power-hyperc-plan} 

The hypercorps and associated Planetary Consortium are the only major non-local political entities in the solar system (with the possible exception of the Autonomist Alliance, which is more of a mutual aid pact than a unified political entity). All of the other political entities are based in a single specific location. The various hypercorps transcend location, however. They have offices and branches all over the solar system, serving the needs of people from Pluto to Mercury and all places in between. While most hypercorps have large manufacturing and processing installations on Mercury or Venus, making use of the abundant energy of the first and the complex chemistry of the second, much of the work performed by all of the hypercorps involves developing new technologies and new cornucopia machine templates, both of which can be done in any place that has meshbrowsing access. 

In addition to bases on Mercury, Venus, and other equally resource-rich locations, all hypercorps maintain dedicated research and manufacturing stations scattered throughout the solar system. Well-known facilities include Starware's vast shipyards, the largest of which are located on Luna and the asteroid Vesta, and Omnicor's huge antimatter factory orbiting Mercury. There are many other lesser-known facilities, including the automated mines that the mysterious Zrbny Group maintains in the main asteroid belt and Saturn's rings, and the qubit factory Nimbus maintains in Mars orbit. 

In addition, there is an even larger number of secure and often secret research installations, some of which are so well hidden that they are normally only accessible via highly secure egocaster connections. All manner of mysterious and often highly dangerous research occurs in such locations, ranging from experiments with the relics of the TITANs to attempts to create self-replicating nanotechnology or artificial miniature black holes. Vids and vidgames are filled with stories both of exotic disasters in such research stations and of heroic thieves stealing amazing wonders from them. While the reality of secret corporate research bases is normally far more prosaic, sometimes wonders are created—and there have been occasional disasters, often involving TITAN relics. 

Some corporate headquarters are similarly secure and secret, including the corporate headquarters of the fabled Zrbny Group. There are a wealth of rumors and stories about such locations. Intrepid spies, thieves, and reporters regularly attempt to gain access to these facilities, generally without success. Many such attempts, especially by would-be thieves and spies, end with distinctly negative consequences, including the thieves' temporary (and on some occasions permanent) death. 

Hypercorps also own and manage a number of habitats. Many are primarily homes for hypercorp employees, but in many of them at least half of the population are simply ordinary residents of the solar system who simply happen to live there. Though far less regulated than hypercorp research or manufacturing facilities, these colonies are also subject to greater regulation and security than some of the autonomist-controlled habitats on the edges of the solar system. 

These stations are exceptionally safe places to live. Residents have access to all of the latest products produced by the ruling hypercorp and its corporate allies. The hypercorp habitats all either possess their own security companies or have some form of defense contract with a private security company, typically Direct Action or Medusan Shield, who agree to protect the inhabitants against potential threats by agents of the TITANs, fanatical saboteurs, or other threats. 

These same security forces also protect the hypercorps from any threats to their interests. In most of these habitats, residents have fairly open freedom of expression and biological self-determination. However, all potential threats to the hypercorp and its personnel, ranging from attempted sabotage to simple civil disobedience, are dealt with quite harshly, with serious offences resulting in forced indenture and occasionally forced mental editing (see Psychosurgery, p. 229). Almost all of these habitats use a transitional economy (p. 61) and most residents have a high standard of living to compensate for the limits on their behavior. Many inhabitants of the more independent colonies in the belt or the outer system complain about the repressive nature of the hypercorp-controlled habitats, but inhabitants of these habitats prefer the safety and security found there to the intimidating freedom of the outer system. 

To help reduce dissent, residents of settlements and habitats controlled by the Planetary Consortium as well as those controlled by hypercorps can vote on a wide variety of issues. The results of these votes, however, are only binding on issues that are not considered “matters of habitat survival,” “corporate policy,” or “security-related issues,” which effectively includes any issue related to the security, profits, and productivity of the hypercorps involved. Votes on these issues are used in a purely advisory fashion, meaning that they are utterly ignored when the result of the vote is at odds with the hypercorps' agendas. 

While residents of these settlements and habitats can vote about adding a new holiday to honor some important figure or the location and design of a new park, laws regulating indentures, habitat security, law-enforcement, or other important concerns remain under the control of the hypercorps. This does not mean, however, that the results of elections are completely disregarded. If more than two-thirds of the population strongly supports a particular issue, the Consortium or the hypercorp controlling the habitat usually finds ways to modify their current policies to address these concerns without harming their own interests. In contrast, if only a small number of residents are upset by certain policies, then these wishes are ignored and habitat security forces keep an eye out for possible civil disobedience or other forms of resistance. 

In addition to these dedicated installations and hypercorp-controlled habitats, many hypercorps maintain offices in stations and planetary settlements. Almost every habitat has a Nimbus office with a farcaster and, in the case of larger habitats, QE communicator facilities for instantaneous communication. Both facilities are open to anyone who can pay Nimbus's fees. Ecologene, Skinaesthesia, and several other hypercorps also have offices on most habitats. Every habitat interested in interacting with the rest of transhumanity has at least one automated Experia media node. In smaller habitats, these offices are unobtrusive and managed by limited AIs or indentured infomorphs. The existence of these offices, however, is vitally necessary for the continued happiness and existence of transhumanity. Most hypercorps also maintain a number of employees in every large habitat and most of the smaller ones. 

Due to the large number of remaining infomorph refugees, most Experia media nodes are managed by indentured infomorphs. These infomorphs monitor the local news-finding AIs and keeps track of any important or interesting developments. They also serve as on-site reporters for any important events that might occur. While postings in small habitats are often rather dull, the infomorph usually has a contract guaranteeing them a morph of their choice and resleeving in the habitat of their choice in return for a term of service, which typically ranges from three to five years. 

Similarly, all but the smallest habitats have Medusan Shield or Direct Action offices, where individuals can hire both security consultants and bodyguards ranging from simple AIs to highly trained mercenaries in fully-equipped fury morphs. These mercenaries live on the station and often hire short-term contractors to help with especially large or difficult assignments. Skilled mercenaries may eventually be hired full-time by Medusan Shield or Direct Action, but since contractors are usually given the most dangerous and thankless parts of any assignment, many soon lose interest in hypercorp contract work. 

Other employees working out of local hypercorp offices range from ecosystem designers to for-hire scientists and technicians to personal financial advisors to the wealthy and powerful. In important habitats and planetary settlements, as much as twenty percent of the population consists of hypercorp employees or private contractors who are hired on a short-term basis when the local workload exceeds the capacity of the regular staff. These hypercorp employees are in the unique position of having dual loyalties—to their habitat and to their hypercorp. Despite what hypercorp propaganda preaches, the two interests do not always overlap. 

Because of the delays involved in normal communication, local heads of hypercorp offices usually have a great deal of autonomy, since asking for instructions from their superiors on another habitat or installation requires either dealing with a time-lag or using expensive qubits for instant QE communication. As a result, except for the most important or difficult problems, local directors deal with all local matters on their own, reporting any unusual or potentially problematic decisions afterwards. 

\subsection{The outer system}
\label{sec:outer-system} 

Out beyond the orbit of Mars, the influence of the hypercorps and the Planetary Consortium is far more limited. With the exception of the rigidly authoritarian Jovian Republic, the inhabitants of the outer system have considerably more freedom than those living in the inner system. However, even out here the struggle between the desire for freedom and the longing for safety form an important part of the political discourse. 

\subsubsection{The libertarian and utopian legacies}
\label{sec:libert-utop-legac} 

Various forms of anarchism and similar libertarian ideologies were quite common among the first transhumans who settled space in the two decades before the Fall. Many settlements in the outer system have inherited this legacy of freedom. The new frontier opened by space colonization presented a fantastic opportunity for those with a strong desire to avoid the authoritarianism of the hypercorp-controlled inner system and Earth to pursue social organizations more based in equality and collective action, or even to simply experiment with new social models. Out beyond the belt, hypercorp influence was weak and preoccupied, giving resourceful colonists a chance to explore their interests unmolested. The more radical of these elements grew out of or maintained ties to progressive, radical, and left-wing social movements and insurgencies on Earth, drawing support where they could. Others simply stole hypercorp resources from the inner system, smuggling them to their secret projects. In a few cases, entire ships or stations mutinied, refusing corporate orders and pursuing their own path. It was rarely feasible for the hypercorps to pursue and punish such subversion. 

Even among these libertines, differences existed, so that those adhering to similar socio-political tendencies tended to group together. Over time these have developed into four rough groupings: the anarchists of Locus, the techno-socialists of Titan, the anarchocapitalists and mutualists of Extropia, and the nomadic free-for-all societies of the individualist scum. These factions form a loose alliance, a united front against the hypercorps and Jovian Republic—or as they call it, the Jovian Junta—and a pact for mutual aid and support, known as the Autonomist Alliance. 

Among the more libertine habitats, the centuriesold doctrine of “From each according to their ability, to each according to their need” is a living and vital philosophy. The ready availability of cornucopia machines ensures that no one wants, and the use of reputation systems encourages people to be active participants towards the common good. Equitable access to morphs and augmentations is also available for residents, though the demand from so many infomorphs in need of a body means that infugees must contribute and build up social capital. However, even for an infomorph, egocasting across the solar system is expensive, and the Planetary Consortium produces large amounts of propaganda about the dangers of these habitats to discourage infugees from considering escape. 

Many autonomists consider themselves to be engaged in an ideological conflict with the inner system, a memetic cold war that sometimes extends to physical actions. Some willingly pursue campaigns of sabotage and subversion against hypercorp and other authoritarian affairs, such as smuggling cornucopia machines into habitats where such machines are strictly regulated, like among the Jovian Republic. The hypercorps and their allies occasionally strike back, though open conflict is rare. Even though the inner system and Jovian Republic could field enough military might to subdue the autonomist factions, an uneasy detente exists. Rumors abound that the anarchists have some sort of card in their pocket that keeps their opponents at bay, perhaps even some threat of mutually-assured destruction. 

Concerns over security and potential future attacks by the TITANs also impact matters in the outer system, but most people resist attempts to seriously restrict their personal freedoms in any manner not directly related to maintaining their safety. Inhabitants of the outer system still remember how the old governments' demands of adherence to bioconservativism and allegiance to distant and often unresponsive leaders did nothing to prevent the Fall from happening, and that memory fuels their mistrust of those states. Those powers were undone by failing to deliver what they promised—when they could not provide the security that they claimed their authoritarian measures would bring, the seeds of their defeat in the outer system were planted. 

\subsubsection{Space for experimentation}
\label{sec:space-exper} 

Both social and political experimentation are common in many of the smaller habitats of the outer system. Because collective decision-making is fairly easy in stations with populations of less than ten thousand, direct democracy is a common method of government. A number of ideologically-based habitats have used this ease of making collective decisions as a way to get all members to agree to some unusual forms of government. 

The individual variants that have been tried are too numerous to list, though they generally fit into a few general categories. A few relatively small habitats employ limited forms of authoritarianism. Some have a single leader who has great power, but who is (ideally) kept from abuse or excess through the use of limits such as a list of constitutionally-guaranteed rights or the ability of a relatively small number of people to call an election or a vote of confidence. Some colonies using this model have elected dictators who serve for a limited term, while others are ruled by a single charismatic leader who transforms their habitat into a cult of personality. 

Other habitats choose their leaders by random lot, with every adult who can pass a relatively easy competency test being eligible to be the colony's leader for a period that usually ranges from six months to five years. A few habitats are governed by powerful specialized AIs, which in very few cases are actually hyper-intelligent AGIs or even seed AIs that the colony has secretly created. Several colonies populated by purely informorph or synthmorph inhabitants use special high-bandwidth connections to give their members access each other's surface thoughts and emotional reactions, allowing them to hold vast democratic political meetings where everyone present can feel the general emotional reactions of all of the other members as easily as they can feel their own. 

There are a vast number of different types of government, many of which have never existed before, moving (and sometimes fumbling) ahead in the outer system. Some work far better than others, allowing successful colonies to thrive and making much of the outer system a vast and complex political laboratory. 

\section{Keeping the peace}
\label{sec:keeping-peace} 

Each habitat is responsible for dealing with its internal affairs. As a result, standards of justice vary widely from the oppressive police state of the Jovian Junta to the free market judicial courts of the Extropians in the belt to the community justice policies of the anarchists out beyond Saturn. Travelers are strongly encouraged to check up on the legalities and policies of stations they are visiting so as to avoid unfortunate incidents, though muses are generally quite good about maintaining awareness of local conditions so that they can warn their users before straying into gray or illegal territory. 

In the inner system, standards of justice and law enforcement tend to be uniform and very familiar to the majority of the population that lived on Earth prior to the Fall, where most nations had relatively similar standards of justice. Across the entire solar system, certain similar standards can be found. Though local laws may differ, there is widespread respect for the idea that punishments for religious or ideologicallybased laws only apply to residents. Visitors who violate such restrictions or other minor laws are simply deported to their home and forbidden to return. Standards of evidence for criminal investigations are also common. Modern forensic technology makes collecting and analyzing DNA and other trace evidence an exceptionally swift and easy process. Likewise, with almost all habitats having what amounts to total surveillance of all public places, any potential offenses committed there can be carefully analyzed. 

Standards of privacy vary widely from one habitat to another, so during emergencies or crime investigations, law enforcement officials may or may not have total access to detailed recordings of the events in any portion of the habitat including recordings from sensors in private dwellings. In some stations, law enforcement officials can compel everyone who might have been present during an alleged crime to provide downloads of their sensory experiences from the time of the crime. While individuals can edit their memories, discrepancies between various people's sensory recordings are just another form of evidence. Requiring sensory downloads from witnesses and suspects is common practice in habitats controlled by the Planetary Consortium, the Jovian Republic, and most hypercorps. However, in most habitats in the outer system, law enforcement officials have no access to such records and can only compel sensory recordings from people who have been charged with serious crimes. 

The power of modern forensics is such that a sufficiently careful examination of people and places can often determine the nature of a crime and the perpetrator(s) with relative ease. Decisions of innocence or guilt rarely rely upon suppositions, circumstantial evidence, eyewitness testimony or any of the other notoriously unreliable forms of evidence common in past centuries. The best way for someone to avoid being convicted of a crime is to either prevent anyone from learning about the crime or to make certain that no one suspects them as the perpetrator. Once someone guilty of a crime becomes a suspect, there is a very significant chance that law enforcement officers will be able to uncover reliable evidence connecting them to the crime. However, if there is no obvious evidence connecting a specific suspect to a crime, the criminal has a greater chance of escaping discovery. 

\subsection{Law enforcement}
\label{sec:law-enforcement} 

Law enforcement in the solar system consists of a vast patchwork of separate jurisdictions, occasionally united by various treaties. Most habitats have signed the Treaty of Uniform Security that requires either extradition or on-site trial of criminals who are accused of especially serious crimes such as attempted habitat destruction, use of incapacitating infoware (including basilisk hack attacks), or any attempt to aid the agents of the TITANs in taking over or destroying a habitat. Only the Jovian Junta and a few especially antisocial or anarchic habitats have not signed this treaty, but many habitats in the outer system maintain the right to try offenders accused by other habitats rather than extraditing them. In addition, most habitats require a significant amount of evidence before they are willing to extradite one of their residents. 

Outside the Treaty of Uniform Security, there is nothing remotely resembling a uniform code of justice and no widely recognized police force. Instead, each habitat or cluster maintains their own code of laws and law enforcement officers. In most areas, law enforcement is a respected and honorable profession paid for by the government, but in a few, the only options are private security agencies that only protect individuals who subscribe to their services. Among the anarchists and scum, residents are largely responsible for their own protection, which means they may be constantly armed when in public (depending on local conditions). Depending on the stations, the most someone who is the victim of a crime can do may be to go after their attacker or post a bounty. In others, mechanisms exist for community or collective problem-solving that often involve assembling an ad-hoc grouping of peers to assess the situation, offer non-biased judgment, and sometimes pursue collective action. 

The only widely-accepted law enforcement officers that attempt to maintain jurisdiction across the solar system are bonded investigators and security consultants from companies such as Medusan Shield or Direct Action. Both organizations have contracts with various hypercorps and inner system stations to provide security. However, in the outer system and in other regions not controlled (directly or indirectly) by the hypercorps, the status of these officers is far more tenuous. In habitats that do not have security contracts with their organization, the best these agents can do is act as bounty hunters. 

Due to extensive stories of excesses in the inner system, many colonies frown on freelance bounty hunters—often referred to as ego hunters—and may ban them entirely. Others allow agents from licensed security hypercorps to act as ego hunters, but forbid them from extraditing or otherwise restraining or punishing the criminals they are pursuing. Instead, agents are required to turn over evidence so that the habitat's own judicial system may hold a trial, in which case a convicted person may be remanded to the agent's custody. Law enforcement officers experience similar difficulties attempting to apprehend a suspect who has fled to another habitat. 

Closely allied habitats in the outer system usually allow full or at least limited legal powers to visiting law enforcement officers from their allies. There are also various small private security organizations that work closely with local law enforcement offices to provide inter-habitat security between habitats that are not closely allied. The members of these organizations attempt to maintain sufficiently high rep to earn the respect of all the habitats with which they work. They act as both bounty hunters and unbiased investigators in situations that involve the laws of several habitats. All of these security companies are located in the outer system, and none has jurisdiction extending beyond a relatively limited location, like the middle belt or the Saturn system. Any such organizations that attempt to grow larger come into direct competition with Medusan Shield and Direct Action and are subsequently either bought out or undercut and discredited by one or both of these organizations. 

There are also several private bounty hunters and private investigators, some of whom are highly reliable. Others are known for their extreme moral and ethical flexibility, especially if the pay is sufficiently high. On some of the autonomist stations and scum ships, these private contractors can be hired to simply go on board and abduct or execute a resident as long as this person has a low enough rep. Attempting to abduct or kill a respected member of the community, however, rapidly earns the ire of the entire habitat. The various small-scale or private security organizations from the outer system can sometimes pursue subjects to habitats controlled by the various hypercorps or the Planetary Consortium. Doing so requires background checks, security screenings, and often moderately large payments. 

\subsection{Punishment}
\label{sec:punishment} 

Among the autonomist colonies, forced exile or repaying the victim with an equitable amount of goods or labor are the principle punishments for all but the most heinous crimes (such as attempted mass murder, habitat destruction, attempting to create seed AIs or similarly extreme actions). In the collectivist anarchist habitats, antisocial behavior typically involves expulsion or penalizing reputation, though solutions that involve making amends are often pursued over standard punishments. At the other end of the spectrum, people convicted of more serious crimes in the most violent and lawless habitats are executed and all of their known backups destroyed. In many others, exceedingly serious crimes are usually dealt with by giving the criminal a choice of forced uploading into a humanely outfitted but closed computer or mandatory personality modification—assuming that someone has not simply killed the criminal before they were brought to justice (such killings are generally treated as matters of self-defense). Mandatory personality modifications are generally limited to the absolute minimum necessary to prevent the individual from repeating similar crimes. 

At the other extreme, punishments in hypercorpcontrolled habitats and settlements controlled by the Planetary Consortium range from fines paid in either money or labor to periods of involuntary indentured servitude ranging from several months to many years. Violent crimes, especially ones threatening either important hypercorp employees or the habitat as a whole, also result in mandatory personality modification. Such modifications often include the creation of a strong sense of loyalty and obedience to the hypercorp. 

Punishments are even more draconic in the Jovian Republic, where permanent execution and the destruction of all backups is the most common punishment for serious crimes against the leaders or large groups of the populace. Since the rulers of the Republic are strong bioconservatives, personality editing and forced uploading are rarely used. Forced indenture is very common, however, as are more standard forms of imprisonment. The Republic is one of the last places in the solar system that has physical prisons. 

The vast majority of other habitats fall somewhere between these extremes. Punishments for non-violent crimes consist of enforced repayment, where the offender must work off a debt to their victim or victims or face more serious punishments. Instead of enforced indenture, offenders usually must only work between five and twenty hours a week for their victims and only need to do so until the crime has been suitably repaid. The typical repayment is between two and three times the value of the good or service taken from the victim. 

\section{The economy}
\label{sec:economy} 

Leaving aside the struggles of bands of primitives to survive on the ruins of Earth, all of humanity has at least some access to the wonders of nanotechnology. This access is highly variable and the economic benefits it produces can be divided into three broad categories—the old economy, the transitional economy, and the new economy. 

\subsection{The old economy}
\label{sec:old-economy} 

The old economy is essentially the same sort of industrial consumer capitalism that has been in place since the late 19th century, a system centered on manufacturers who create material goods and sell them to consumers. Modern manufacturers now make their goods in cornucopia machines instead of factories, but the essential pattern is the same one that has existed for over two hundred years. Due to the high level of inefficiency and unfairness in this economic system, poverty is relatively common. The poorest individuals often face hunger, homelessness, lack of medical care, and similarly dire problems. 

Ordinary members of this society never have direct access to cornucopia machines. Instead, they purchase their goods from corporations, governments, or wealthy individuals who control them. Some old economy societies have planned economies, where the corporations or the state determine what options the citizens may choose or occasionally what goods they must have. Others claim to have a free market, where citizens have more options, but the residents must still pay to obtain goods that are essentially free for the corporations or government to produce. 

In the present day, almost no one willingly lives in old economy societies. Very few individuals even visit such societies. The oppressive Jovian Republic holds most of the remaining old economy societies in the solar system. The few other surviving examples are totalitarian regimes where the wealthy elite maintain absolute control of all cornucopia machines and private ownership of one is a very serious crime. Since cornucopia machines can be used to create more cornucopia machines, maintaining strict control over them requires constant vigilance. 

Residents of old economy societies tend to look at residents of transitional and new economy societies with envy, while residents of habitats that use both transitional and new economies look upon residents of old economy habitats with a mixture of horror and pity. Since the Fall, almost a third of the remaining old economy-based habitats have transformed into transitional or new economies by various means, often involving violent revolution. Most social scientists predict that unless there are further catastrophes, all but the most repressive old economy societies are almost certain to transform to transitional economies within twenty to thirty years. 

Old economy societies are unique in that money is the society's only acceptable means of exchange. While reputation networks exist, they are informal and serve as an unsanctioned means of exchanging favors. 

\subsection{The transitional economy}
\label{sec:transitional-economy} 

The transitional economy is a far more stable and easily maintained system than the old economy. Transitional economies blend old and new economies, and habitats using this system feature both private ownership of cornucopia machines as well as public fabbers and makers that are freely accessible. These public machines are strictly limited in the goods they can produce. In addition, the raw materials for various complex goods are also strictly regulated. Mars, Venus, and Luna are all examples of transitional economies, as is most of the rest of the inner system. 

For the inhabitants of a transitional economy, creating food, non-smart clothing, furniture, and most other simple, non-formatible objects is a trivial matter. However, the public nanofabrication machines can only create objects that either contain no electronics at all or contain only simple circuits that report on the object's condition and location. Manufacturing any of these items requires little more than the machine and a supply of carbon, hydrogen, oxygen, nitrogen, silicon, iron, aluminum, and tiny amounts of various trace materials. All of these materials are sufficiently abundant that acquiring them is easy and inexpensive. 

Using the elements that are freely available to all tax-paying citizens, nanofabbers can produce a vast array of goods like exquisite suits of silk clothing, tables with the appearance of finely polished ebony and mahogany, beautiful colored glass goblets, or painted porcelain tea cups. They can also create a gourmet dinner and a set of fine plates and cutlery on which to eat the meal. To pay for the small amounts of energy and resources needed to create these goods, all inhabitants pay a small tax. 

Once the usage tax has been paid, food, clothing, furniture, and similar goods are all free. Raw materials, old, worn-out or unwanted goods, and various waste products are recycled into new goods. Residents of transitional economies need never experience hunger or any of the many other sorts of deprivation that much of humanity faced before the mid-21st century. Additionally, basic medical care is free in almost all transitional economy societies, to help insure that the populace is healthy, content, and productive. 

While many goods are freely available, there are also goods that residents must purchase from corporations, their government, or other producers. Smart clothing and smart furniture that can change shape, color, and pattern, depending upon the user's wishes, cannot be manufactured in any of the personal nanofabricators. Any goods made from highly durable composite materials, batteries, electrically-powered devices including all augmentations, and all nanotechnology must be acquired in the same fashion. These goods are considerably less common as they require access to an unrestricted nanofabricator and exotic raw materials. 

Transitional economies tend to be relatively safe places, since inhabitants cannot manufacture weapons more dangerous than knives, clubs, or similar primitive armaments. Everything from firearms to plasma weapons requires restricted cornucopia machines and exotic materials to manufacture. The proliferation of these items is strictly controlled. 

Some habitats in the outer system have transitional economies because residents prefer the safety that comes from centralizing control of potentially dangerous technologies. Other habitats have transitional economies by default, because they have limited stocks of many of the more rare elements required for manufacturing various complex modern technologies. Regardless of the reason, outsiders from new economy habitants often see them as somewhat poor and deprived, while many residents of transitional economies consider new economy societies both exceptionally wealthy and somewhat frightening. 

Despite these differences in perception, both economic societies have a great deal in common. Food, clothing, and similar goods are easily available to all residents. An individual's status, taste, wealth, and reputation are measured by the kinds of clothing, food, and furnishings they possess. While there are a vast number of templates for different styles of food and consumer goods, forward-thinking designers develop new designs every month and use copy protection on these designs to keep them from being pirated for at least a month or two (and often longer). As a result, for the first few months after their release, the only people who can gain access to new designs in clothing, tableware, food, or similar goods are those who pay a premium to the designer to download the templates that allow their cornucopia machine to manufacture the item. 

Since one way of defining a transitional economy is a system where both reputation and money are in widespread use, most have developed ways to accommodate both forms of payment. While residents primarily use money for purchasing goods, purchasing cornucopia machine templates involves rep, especially among residents who regularly visit new economy societies or have significant contacts there. 

\subsection{The new economy}
\label{sec:new-economy} 

Slightly less than forty percent of the human population lives under some version of what social scientists refer to as the new economy. In the outer system, alternative economies are becoming increasingly rare. New economies are much better than old or transitional economies at supporting a decentralized populace, which has led to more than half of all habitats and settlements adopting this model. 

In new economy societies, individuals can freely manufacture and use almost anything they want, assuming they can acquire the correct templates and raw materials. As a result, the residents' need for food, clothing, medical care, information access, and other basic needs are all easily met. However, there are still items of value that individuals work very hard to obtain. Though these are commonly described as “post scarcity” societies, some types of scarcity remain very real. 

In most new economy habitats, common goods are freely available to all residents—or at least to all residents who meet certain criteria. These criteria usually take one of two forms: citizenship or public works. In wealthy and prestigious habitats, free access to all common goods is offered to residents who have official citizenship. Citizenship can be earned in a variety of ways, but the most common involves either being considered a strategic asset due to some singular expertise, performing an exceedingly valuable service to the habitat, or working for the habitat for some period of time. Once an individual is a citizen, the energy, living space, and raw materials they use in the course of their daily lives are all freely available. 

In many collectivized habitats, residents are expected to pull their weight by contributing to ongoing public works in the habitat, typically requiring between four and eight hours every week. Depending on the nature of the colony, this work may be selected by the government, the collective syndicates that oversee the management of resources, or by a high rep individual who controls access to large amounts of energy and raw materials. Unless someone has especially valuable skills, this labor is often dull but safe work that can be done more easily by humans than AIs, such as checking the habitat for flaws and performing maintenance tasks. 

Assuming an individual has acquired citizenship or put in their share of work for the collective wellbeing of the station, they will have access to a supply of energy and raw materials that allows them to use their cornucopia machines to manufacture what they need. Visitors are generally also allowed access, though anyone staying long is expected to contribute to the habitat if they don't want to see their reputation slashed. 

\subsubsection{Restricting dangerous technologies}
\label{sec:restr-dang-techn} 

Most societies in Eclipse Phase see good reason to restrict access to some dangerous goods, especially military hardware. Few people living in a sealed habitat surrounded by hard vacuum enjoy the idea of easy access to biowar plagues or devices that can make large holes in their habitat's outer hull. Though such incidents are quite rare, the memories of horrors like the recent BransonVesta disaster are still quite fresh. In that incident, a radical bioconservative cult manufactured several plasma bombs and accidentally destroyed the entire habitat when their attack on the local government caused a cascading blowout, cracking the spinning habitat in half. More than 50,000 residents had to be resleeved, and 400 permanently died when their backups and cortical stacks were destroyed in the explosions. As a result, standard procedure is to restrict access and heavily encrypt templates needed to create military-grade weapons and similar dangers, though sufficiently dedicated individuals can eventually decrypt or reverse-engineer such designs. Even nanofabricators in anarchist habitats may be blocked from creating such things or at the very least will alert the local public mesh if anyone instructs them to do so. Habitats that possess almost no other laws regarding possession of various objects and devices usually have laws against weapons that can do serious harm to the habitat. Many dangerous technologies are specifically designed to make use of various exceptionally rare or human-made elements, including radioactive elements and artificially created transuranic elements. Therefore, many habitats will restrict access to these elements to limit the manufacture of these weapons. Since detecting radioactive elements is simple using standard environmental sensors located throughout every habitat, security authorities can easily learn when someone has acquired significant quantities of such elements, or catch them if they attempt to bring them on board. 

\subsubsection{Value and scarcity in new economy societies}
\label{sec:value-scarcity-new} 

While basic citizenship allowances cover most necessities and even some luxuries, the allowance has limits. With the allowance, individuals receive a quota of goods and energy they can use every day. This usage is impressively lavish by 20th century terms, allowing residents to create a dozen suits of clothing and provide food for half a dozen people every day. Creating elaborate food, furniture, and tableware to serve a party of a dozen people is within the means of any individual. However, doing the same thing for a party of two hundred people is outside the bounds of the basic allowance. 

Individuals who wish to exceed their basic citizenship allowance can either use rep to obtain more access to resources and energy, or they can pool their resources with others to accomplish their goals. There are many goods that are fairly complex to create—including many of the best morphs and highly specialized and intricate pieces of gear like advanced augmentations—that exceed the resources available in a basic citizenship allowance. 

The allowance also limits the amount of travel that residents can easily undertake. Residents of most new economy habitats own good quality spacesuits, and many can use their rep to create a small and very minimally equipped travel pod to travel to a nearby habitat. However, even the smallest actual spacecraft are far too large and difficult to create to be available on an ordinary citizenship allowance, or even on the amount of rep an ordinary individual can acquire in a reasonable amount of time. 

In addition to large-scale uses of resources and difficult-to-manufacture goods, there are goods that are intrinsically scarce, such as relics of Earth and handmade goods. While exact copies of everything from the Mona Lisa to a pressed daisy are exceptionally easy to acquire, genuine physical relics of Earth are prized possessions. The vast majority of refugees could take nothing with them, but almost everyone wishes to have some token to remind them of Earth. A single dried flower, coin, or piece of stone from Earth can be exchanged for almost any morph or other good that is moderately difficult to create. Actual historical artifacts, like a famous person's hat or autograph, is worth far more, as are original works of art by famous artists. Two years ago, one of the last three remaining paintings by Leonardo da Vinci was traded for a large and well-equipped spacecraft, and a small piece of the Liberty Bell was traded for both a custom-designed morph and a fully outfitted onehectare villa in one of the more prosperous habitats orbiting Saturn. 

While less expensive than Earth relics, handmade goods also command a high price and are in great demand by the wealthy. Though most people cannot distinguish between a fine wine grown on one of the Martian vineyards and a duplicate of the same wine produced using an average cornucopia machine, some connoisseurs claim they can taste the difference. There is also much prestige to be gained by serving hand-grown food. As a result, while anyone can drink nanofabricated wine, hand-produced wine is a rare good that can only be enjoyed by a few, and thus it commands a moderately high price. In almost all cases, handmade goods are expensive because of their rarity and because many people enjoy the status associated with owning and using them. 

There are three other items that are scarce and are thus quite valuable: living space, skilled sentient labor, and novelty. The majority of humanity lives in standard-sized dwelling units, which typically range from one hundred cubic meters on smaller or poorer 

 habitats to two hundred cubic meters on wealthy and prosperous habitats. Since each cubic meter of a habitat must be manufactured and the process of manufacturing or expanding a habitat is far from simple, space is at a premium. The only exceptions to this scarcity are on Europa and Mars, which can be inhabited by properly adapted morphs without the necessity of complex life support or the danger of vacuum waiting just outside every exterior wall. As a result, owning a larger dwelling space in a habitat is worth a significant amount, and large villas and private asteroids are luxuries possessed by only the highest rep individuals. 

While transhuman labor has become relatively cheap due to the large number of infugees who must sell their services or indenture themselves to obtain morphs and habitat space, skilled labor is far more expensive. Buying a unique custom morph design, for example, crafted by a skilled biogeneticist, can cost as much as a small spacecraft depending on how much this morph deviates from standard models. The same is true for everything from custom-designed clothing to complex pieces of technology designed for a single specific usage. While the actual manufacturing costs of these items is no more expensive than any other similar item, the time and effort needed to design them can make them exceedingly expensive. 

The final commodity that is both scarce and valuable is novelty. While anyone can drink a fine wine or wear a wide range of designer clothing, other commodities are kept deliberately scarce. Cuttingedge fashion, new music, and even haute nosh (bold, exclusive snack food designs) are harder to find because the templates needed to manufacture them are encrypted and cannot be copied. The copy protection used on the templates for newly created goods automatically expires within three years at most, and most habitats reduce this to one year. In addition, this copy protection is never perfect; someone always manages to create pirated versions of these new goods within two to six months. However, from the time templates are created until the time that someone pirates them, these items are only available to individuals who are willing and able to pay for them. Popular new templates command a good price in the new economy, and a large number of transhumans make their living designing and marketing such templates. 

\subsubsection{Irredporducible goods}
\label{sec:irreproducible-goods} 

In an age when digital material is easily copied and physical goods are reproducible with nanofabrication, concepts like copyright, trademark, and intellectual property are fighting a losing war. Despite the best methods of encryption, DRM, and similar anti-piracy measures, very little escapes the clutches of pirates for long. It’s not unheard of for copies/blueprints of new goods to be shared on pirate networks before they’re even officially released. 

In response, some manufacturers, designers, and artists attempt to produce goods that are irreproducible—and thus more highly valued. Possible approaches include transgenic living sculptures with built-in obsolescence and terminator genes, energy art, items made from extremely rare materials (e.g., a chair crafted from titanium mined from the Mead crater on the harsh Venusian surface), or intangibles such as skilled performances. 

\subsubsection{The economy and infomorph refugees}
\label{sec:econ-infom-refug} 

During the last phase of The Fall and the evacuation of Earth, more than four hundred million refugees were uploaded and egocast to orbital databanks. From there, infomorph refugees were beamed to databanks throughout the solar system. They were forced to flee Earth without any of their possessions, even their bodies. Instead, they became infomorphs who had nothing beyond their minds and memories—the most destitute group of refugees ever to exist in human history. In the years since the Fall, large numbers of these infugees have been resleeved. Anyone with valuable skills was first to gain a morph, followed by anyone with friends or relatives already living in orbit who could take responsibility for the person's resleeving. 

Those two groups accounted for only half of the refugees. The remaining found themselves in a far more difficult situation. Lacking either personal contacts or vital skills, they had no one else to help them. In the first few years, many of these infugees signed contracts promising their labor or other services in return for resleeving and a guarantee of some form of income sufficient to support them. Because of the critical labor shortages in the first five years after the Fall, another thirty percent of the refugees managed to regain bodies (usually cheap synthmorphs). These indentured servants performed all manner of critical tasks, ranging from scavenging ruined habitats for useful devices to mining or asteroid herding. Others became servants or bodyguards for the rich, or performed less moral services for criminal syndicates. Most took on orbital construction jobs, helping to construct the new habitats that would eventually become their home. Some infugees found work performing services like data-mining, monitoring automated factories, or other jobs that could be done by infomorphs. After the Fall, infomorphs were used to take over numerous tasks previously handled by AGIs, who were no longer trusted. 

Unfortunately, some infomorph refugees made bad or unlucky deals and ended up working for years only to find that their employer either kept finding ways to delay or reduce the payment or vanished before they delivered on their promise. As a result, slightly more than twenty percent of the original infomorph refugees remain infomorphs; some by choice, but most because they have not been able to acquire the means to resleeve themselves or are still working long contracts to gain their morph. The problem with obtaining bodies for these infugees goes beyond simply providing a new morph for resleeving; living beings require living space as well as a steady supply of consumables. For this reason, many infugees have been morphed in synthetic shells and housed in areas inhospitable to biomorphs, such as the unenclosed portions of Venusian aerostats. With space in short supply, the waiting list for infugees looking for a habitat to call home is quite long. 

Both the hypercorps and the Planetary Consortium were quick to make use of this vast labor pool, especially on Mars. Mars has large amounts of open space and resources and is sufficiently close to habitable that Mars-adapted morphs like the ruster are inexpensive to create. As a result, the Planetary Consortium has been responsible for the employment of almost half of all remaining infomorph refugees. For the past decade, the vast majority of infomorph refugees who want bodies have found that indenturing themselves to the Planetary Consortium or one of the associated hypercorps involved in Martian terraforming is the most reliable way to find both a morph and housing, since both are guaranteed at the end of the contract. The work involved is particularly difficult, however, and the contracts are normally quite long. The Planetary Consortium is also particularly adept at adding charges that prolong indenture—though most indentures carry five to twenty year contracts, in reality these indentures typically last between eight and twenty-five years; some go on even longer. This large population of indentured servants on Mars—many of them now free and resleeved—is becoming a force in its own right, adhering to the Martian wilds and rural areas and disdaining the elite hypercorp domes. Adopting the name Barsoomians from an old Earth fiction series, this resentful lower class is increasingly becoming a thorn in the Planetary Consortium's side. 

Even though it is highly automated, terraforming and agricultural work on Mars is both tedious and physically demanding labor. Indentured employees are regularly sent into the regions that were most affected by the Fall. As a result, these employees occasionally face attacks by life forms mutated by the TITANs, nanotech war-swarms, or similar still-active and dangerous exotic technologies. Indentured employees are not charged for damage to or destruction of their morphs caused by such dangers, but the experience of even reversible death from such causes is highly traumatic. 

Other refugees found that they enjoyed life as infomorphs, reveling in complex simulspaces and otherwise living up the virtual life. Some found work that paid for the ability to egocast throughout the solar system. Ten years after the Fall, there is a thriving infomorph culture. While exact data is difficult to obtain, many researchers believe that at least a third of all current infomorph refugees have no plans to place themselves into a morph, instead enjoying the freedom of virtual existence. Especially in the outer system, these infomorphs have become increasingly involved in habitat politics; many habitats have officials who are infomorphs. Most researchers predict this infomorph culture will increasingly diverge from physical cultures as time progresses. 

\subsubsection{The clanking masses}
\label{sec:clanking-masses} 

With so many infugees acquiring cheap synthmorph shells—particularly cases and synths—and being unable to afford anything better, synthmorphs have become associated with poverty throughout the solar system. This lowest strata of the poor are often referred to as “the clanking masses,” and compose one-sixth of the transhuman population. Most of these people strongly desire to acquire a biomorph, even if it is only a splicer or worker pod. As a result of their presence, however, many synthmorphs are now viewed with distaste, especially in elite social circles. Even those who have expensive, lovely, custom-designed synthetic morphs fitted with all of the latest augmentations are considered to be eccentrics with poor taste. 

The social stigma against synthmorphs is strengthened by the fear that, in the event of another attack by the TITANs, their robotic shells could be rapidly co-opted to become a deadly TITAN-controlled army. This has led to some habitats going so far as to actively segregate their synthmorph populations, rationalized by the fact that synthmorphs can easily inhabit unheated and unpressurized portions of various habitats. This segregation and social stigma, however, has produced the beginnings of an emergent synthmorph culture. There are already numerous habitats where all of the inhabitants are sleeved in synthetic shells and conventional life support exists only for the few visitors wearing biomorphs.

\section{Habitats}
\label{sec:habitats} 

With Earth now uninhabitable, transhumanity survives in a variety of off-world habitats. There are two major types of these habitats: settlements on planets or large moons, such as those on Luna, Mars, Venus, Europa, or Titan, and space habitats that are built on or near an asteroid or other useful source of raw materials. Most of these space habitats spin themselves to provide gravity, with Earth and Mars gravity being the two most common choices. There are also a large number of zero-g or microgravity habitats, consisting of either non-spinning habitats or stations built into small asteroids or moons. 

\subsection{Planetary settlements}
\label{sec:plan-settl} 

The Martian and Lunar city-states and other planetary settlements contain environments most familiar to refugees from Earth. This similarity is one reason that two-thirds of all infomorph refugees live on Mars, Luna, or Titan. The exact type of settlements depends on the planet or moon on which they are located, with some being far more similar to Earth cities than others. Most Lunar settlements, like those on Ganymede, Mercury, Titan, and Callisto, consist of a network of subsurface tunnels and chambers excavated with plasma drills. These tunnel settlements differ slightly from one world to the next. In most of these tunnel cities, the floors of all open areas and many dwellings are composed of geneticallymodified grass designed for both comfort and durability, with light panels covering the ceiling providing bright full-spectrum lighting. 

A few of these buried cities further enhance their natural appearance with the addition of trees and, in some cases, specially engineered ecosystems, in both public areas and private dwellings. A few of these urban tunnel forests and jungles are home to numerous flowering vines and bright tropical butterflies. In a small number of settlements on both Titan and Luna, colonies of small monkeys and parrots with metabolisms and habits modified for modern ideas of cleanliness and sanitation thrive, giving some of these tunnel cities the feel of a buried jungle. 

All of the older or more prosperous tunnel cities also contain large open areas that are typically between one and twenty hectares, with ceilings at least ten meters high. Some are parks, others are public plazas, but all offer the residents of the tunnel cities a chance to experience open spaces. Also, with the exception of Mercury, all of these tunnel cities are on moons where gravity is no more than one-sixth of a g. Some of these open spaces are constructed with roofs between thirty and one hundred meters high and are designed so that residents can use them for flying by strapping on a pair of specially-designed wings. 

The cloud cities of Venus are among the most unusual habitats in the solar system. Their exotic nature is enhanced by the chance to observe the many recently introduced floating and flying life forms modified to live in the clouds. Though located almost fifty kilometers above the most deadly environment in the solar system, life in these cloud cities is among the most Earth-like anywhere in the solar system, with gravity, temperatures, and atmospheric pressure all being very near normal Earth levels. 

By contrast, the settlements on Mars look the most like the cities of lost Earth, built on the surface rather than underground or in the skies. Some of the more recent settlements are designed for use by inhabitants in ruster morphs or synthmorphs and feature no life support. Older Martian cities and other settlements are typically covered with low domes of flexible polycarbonate and filled with a completely breathable, if somewhat low pressure, atmosphere. Some, however, are collections of sealed skyscrapers, connected by skywalks and tunnels. If current terraforming efforts continue on schedule, the last sealed Martian cities will be opened to a Martian atmosphere breathable by all morphs within sixty years. 

The most unusual planetary settlements are the ocean cities of Europa. These are among the most exotic locations in the entire solar system and are quite disorienting for individuals not used to underwater cities. From a distance, most appear to be complex Christmas tree ornaments hanging down one hundred meters or more below the ice crust above. A few are built deeper, plunging under the icy surface near the various hydrothermal vents that host the native Europan life clusters. 

Many of the residents of the Europan cities find them familiar because they previously lived in one of the underwater cities on Earth and so were used to both the conditions and to living in an aquatic-adapted body. Europan cities all contain sealed buildings with normal atmosphere, both because some activities work best in air instead of water and because the cities often host visitors without gills. However, these regions make up only ten percent or so of most of these cities. The remainder looks vaguely similar to many zero-g habitats, except that the structures are considerably sturdier and are located underwater. Buildings are designed to be accessible in all three dimensions, so going from one floor to another usually involves swimming out a large opening in the wall and down a level. In almost all of these aquatic cities, large fusion generators heat the surrounding water, so that the entire city exists in a region of water that is far warmer than the surrounding frigid Europan sea. 

\subsection{Space habitats}
\label{sec:space-habitats} 

With the exception of the private habitats of the wealthy and powerful described below, the vast majority of space habitats hold between twenty-five hundred and one million inhabitants. Almost two-thirds of these habitats were built during the first seven years after the Fall, when huge portions of the system's surviving infrastructure were used to create habitats suitable for hundreds of millions of infugees. 

During this era, several thousand torus habitats and cluster colonies were created throughout the solar system. Many of these habitats were created by automated mining machinery that had been repurposed to create colonies. Due to the limitations of these automated mining rigs, most these habitats were small, holding between one thousand and one hundred thousand inhabitants. Twenty percent of the system's inhabitants live in such habitats. During the past decade, various small organizations, cults, and subcultures have left the larger habitats they lived in and created their own small habitats, few of which were designed to hold more than ten thousand residents. 

The development of the new nanotech Hamilton cylinders has lead to a new interest in large habitats and in habitats that can easily expand in size to accommodate an increasing population. The expense and difficulty involved in expanding existing habitats or building new ones is one of the principle reasons that more than forty million infomorph refugees still do not posses morphs. Although none of the existing Hamilton cylinders has finished growing, they are both highly regarded by their residents. This same technology is also likely to produce a low-cost method for creating small habitats, where the creators merely need to seed an asteroid with the appropriate advanced nanotech generators and wait a few months. 

\subsection{Scum barges}
\label{sec:scum-barges} 

At the opposite extreme from the Hamilton cylinders are the infamous scum barges. Most are spacecraft built before or during the Fall that were used to help with the early stages of the evacuation, ferrying people away from the doomed Earth. Many of these refugee ships were unable to find anywhere to unload their human cargo, becoming a sort of permanent traveling refugee camp, sometimes succumbing to mutinies. They eventually joined up with pre-existing scum ships and swarms, adopting their nomadic, freewheeling, anarchistic lifestyle. In contrast to egocasting or the faster and more efficient fusion drive ships, so-called scum barges offer a floating city alternative to space travel. These ships function as roving black markets and carnivals of the bizarre—lawless zones where anyone can find whatever they want or need for the right rep or price. 

Most scum barges have fusion-powered plasma drives and hold between two hundred and five thousand inhabitants. The worst barges are exceptionally overcrowded, with aging life-support systems struggling to maintain a breathable (but still foul-smelling) atmosphere under the strain of too many passengers. The larger and more prosperous scum barges are often fitted with various modern conveniences, including large cornucopia machines and vast stores of pirated manufacturing templates. Some are thriving utopianist enclaves, while others are mobile dens of smugglers and thieves that would have been destroyed long ago except for the fact that large and powerful organizations find their existence occasionally useful. Living conditions on the scum barges range from overcrowded refugee camps to thriving, egalitarian, but non-wealthy anarchist enclaves, to relatively modern habitats outfitted in barbaric splendor by highly successful organized crime gangs. 

\subsection{A diversity of floating worlds}
\label{sec:divers-float-worlds} 

The use of cornucopia machines and smart materials means that the interiors of all but the poorest and most destitute habitats can be reshaped according to the whims of their inhabitants. When the number of inhabitants is small enough or their aesthetics are uniform enough to all share the same tastes, the results can be both unique and strange. Large-scale fads occasionally sweep through even the largest and most cosmopolitan habitats, making some of the bigger colonies almost as odd. 

Several habitats closely resemble terrestrial jungles, with an entire rainforest canopy growing from the slowly rotating outer shell and all dwellings and pieces of high technology nestled in the branches or hollows of these vast gene-engineered trees. In these living marvels, genetically engineered monkeys, iguanas, and tree sloths wander amidst the inhabitants—some of these creatures are wild animals, while others are controlled by AI servitors and act as maintenance or observation drones. Some habitats resemble other scenes from old Earth, including more than a dozen water-filled habitats hosting some of the aquatic inhabitants of the now-destroyed underwater cities. In most of these marine habitats, the actual buildings are either placed amidst a living coral reef filled with fish and other creatures or are actually built into the coral reef itself. There are many other habitats duplicating other environments, such as Afrique—a large Cole habitat with a population of two hundred thousand, where the habitat is made to resemble the African savanna. In Afrique, the two ends of the habitat are shaped into snowcapped mountains, and the inhabitants mostly live in several large cities built in the savannah. 

While nostalgia for Earth is a powerful force in habitat design, there are many other options. A few exotic habitats resemble fantastic cities from various vidgames or older forms of entertainment, including a handful of small and eccentric habitats where the inhabitants all appear as strange humanoid alien beings. In many, the inhabitants have cosmetically modified themselves to fit in with the setting. 

One of the most common differences between small and large habitats is that the residents of smaller stations often share a common ideology or sense of aesthetics, and so are far more eccentric. Some of the more unusual habitats range from dimly lit, spooky landscapes filled with perpetually leafless trees, thick, continually regenerating cobwebs, and other similar macabre touches to gleaming colonies that are shining citadels of quartz and steel. Some are huge interconnected arcologies where any sort of personal privacy is rare, while in others every family or even every person has a separate dwelling that is rarely seen by outsiders. Since the populations of these stations are relatively small and the vast majority are not major economic centers, travel to and from these smaller habitats is infrequent, which further increases their insularity and idiosyncrasies. 

\subsection{The largest habitats}
\label{sec:largest-habitats} 

Extropia, the huge Martian city-states, and some of the largest Lunar stations hold between one million and twenty million inhabitants. There are many smaller settlements containing between one hundred thousand and one million residents. These habitats are considerably less idiosyncratic and exotic than the smaller habitats. Almost all contain a cosmopolitan and diverse population from a wide variety of subcultures. Because of this diversity and the difficulty of forming any sort of consensus with a large population, these settlements tend to be reminiscent of the cities of Earth. All of them have their own unique character and feel, but the differences between one habitat and another are rarely overwhelming. In addition, all of these stations are large enough to hold offices for all of the major hypercorps, who further promote uniformity by providing the same services from identical hypercorp offices. Since most of these habitats are major centers of commerce, travel between them is frequent, so there are various facilities for travelers such as hotels and sports clubs that help reduce the disorientation of travel by offering identical experiences, regardless of their location. 

\subsection{Microgravity habitats}
\label{sec:micr-habit} 

Zero-g habitats are very different from those that use rotational gravity. Most consist of networks of tunnels drilled through the asteroids—similar to the tunnel cities of Luna and Titan—but some are considerably more exotic. Like most other habitats, almost all microgravity colonies are built in, on, or next to one or more asteroids containing a large amount of useful raw materials. They typically feature a gravity less than 0.01 g that has very little effect on the daily lives of the inhabitants. Near-weightless environments allow for some interesting and unusual habitat designs as there is no up or down, enabling the creation of structures that would be too fragile even in low gravity. The habitats of Nova York (p. 97) and Nguyen's Compact (p. 103) are both examples of this, among many others. 

\subsection{Private habitats}
\label{sec:private-habitats} 

The most rare and exotic of all of the types of habitats are the luxurious private ones owned by exceedingly wealthy or high rep individuals. Most private habitats are small but still give each of the residents several thousand cubic meters of personal space. 

A typical private habitat is either a cylinder one hundred fifty meters in diameter (the minimum necessary to produce Mars gravity at a rate of rotation slow enough to avoid problems in all morphs) and between fifty and two hundred meters long, or a zero-g sphere one hundred to two hundred meters in diameter. These habitats are always tethered to a small collection of raw materials, consisting of chunks of silicate, nickel-iron, and water-containing carboniferous asteroids with a mass equal to at least that of the habitat. The majority of private habitats are inhabited by between half a dozen and three dozen morphs, some or most of which may be AI servants or, on rare occasions, indentured servants. Life in a private habitat is exceptionally lavish. Almost every surface is made of formatible smart materials and there are several large generalpurpose cornucopia machines available for the use of every resident. 

By using these nanofabricators and the smart materials to their fullest, residents can completely change the interior of the habitat in only a day or two—transforming a sterile and crystalline array of shining metal and glass buildings into a thriving forest, inhabited by a variety of wild animals. The mesh is filled with vids and XPs about the lives of the most famous residents of the solar system. Almost everyone has seen the interior of one of these vast space mansions many times, though only a tiny percentage of the inhabitants of the solar system will ever have a chance to actually visit such a location. Many gatecrashers, scavengers who travel to Earth, and others who engage in similarly daring endeavors hope to be able to obtain information or objects sufficiently valuable to allow them to retire to their own private habitat. 

\section{Factions}
\label{sec:factions} 

One would have thought a cataclysmic event such as the Fall would bring the surviving elements of transhumanity closer together, jointly dedicating themselves to the repopulation of the solar system and continued prosperity. Instead, the remoteness and physical isolation of transhuman colonies and habitats stretched across the solar system, as well as the effects emerging technologies have had on transhuman economies and social lives, have promoted the evolution of a wide spectrum of philosophies, agendas, and political models. 

\subsection{The hypercorps}
\label{sec:hypercorps} 

To some economists, the Fall and the numerous crises that predated it on Earth can be viewed as an extinction event, the end of the line for the massive transnational megacorp dinosaurs, financial giants that supported their monolithic frameworks on outdated economic models and industrial technologies. The hypercorps are their evolutionary descendants: slimmer, faster, meaner, and more flexible, eagerly embracing the possibilities of new technologies and never afraid to toss the old aside to take advantage of the new. It was the hypercorps that drove humanity's expansion into space and who continue to push the technological envelope, guiding transhumanity towards new horizons—always with profit as their driving goal. 

Most hypercorps are decentralized, non-assetbased legal entities. Complete automation, advanced robotics, morph technology, and cornucopia machines allow the hypercorps to abstain from mass employment for labor or production services. The need for physical labor has mostly been reduced to tasks associated with habitat construction or deep space mining. Infomorphs and AIs are heavily employed (or more accurately, owned) as drone operators or virtual workers, and many administrative tasks are performed online via augmented reality, virtual private networks, and simulspace nodes. Some hypercorps are in fact entirely “virtual,” with no physical assets and each employee acting as a mobile office. A few major hypercorps literally consist of only a dozen transhuman personnel. Though some hypercorps are massive and diversified, most specialize in particular fields or services. This results in both an intricate system of partnerships to develop, produce, and market products and services and a large-scale tendency to internally contract special services from other hypercorps. Many hypercorps also pool their resources and talent into cooperative research initiatives, project centers, or shared habitats. 

Most hypercorps are traditional capitalist in outlook, though many have adopted alternative business philosophies and management models. This might include basing decisions on internal forecast market trends, groupthink consensus models, or ditching management entirely in favor of staff polling/voting initiatives that statistically fare better. A few are anarcho-capitalist companies originating from Extropian enclaves, though these often suffer from a bias when making deals with inner system powers. 

The solar system boasts thousands of hypercorps; a few of the more prominent and interesting are noted below. 

\subsubsection{Cognite}
\label{sec:cognite} 

\textbf{Major Industries:} Cognitive Science, Mental Implants, Psychosurgery, Nootropics 

\textbf{Major Stations:} Thought (Venus orbit), Phobos (Mars moon) 

 A pioneer in the field of cognitive science, Cognite (pronounced cog-neet) drives forward the cutting edge of research into understanding the transhuman mind. Most well-known for their mental augmentations and the original menton morph design, Cognite also specializes in psychosurgery and nootropics. Their elitist and aloof image was not aided by their scandalous involvement with the projects to raise accelerated growth children that became known as the Lost generation (p. 233), nor rumors that they engage in research involving TITAN-influenced incapacitating input attacks. Nevertheless they remain a key member in the Planetary Consortium. 

\begin{quotation}
\textbf{Psiclone} 

 To: Proxy-99 

 From: <Encrypted> 

 I'm enclosing some data I recently acquired from an inside source regarding a so-called “Project Psiclone”— some type of black budget research initiative pursued by Cognite, possibly with involvement from other Planetary Consortium interests. Their work seems to focus strongly on the Watts-Macleod strain of the Exsurgent virus—with some alarming results.
\end{quotation} 

\subsubsection{Comet Express (COMEX)}
\label{sec:comet-express-comex} 

\textbf{Major Industries:} Courier Services, Shipping,Logistics 

\textbf{Major Stations:} Nectar (Luna), Olympus (Mars) 

 Comet Express specializes in delivery services, interstellar logistics, supply chains, and shipping. They maintain a presence on almost every transhuman habitat in the solar system, often via local subcontractors. Despite the wonders of nanofabrication, many resources must still be imported. ComEx focuses on managing supply and trade routes and making sure physical shipments reach their destinations. For that purpose, ComEx maintains orbital hubs equipped with slingshot accelerators at strategic waypoints throughout the system and a fleet of cargo vessels and courier drones. For reasons unknown to the public, ComEx is viewed with hostility by the Jovian Republic, who have standing orders to shoot down ComEx vessels. 

\subsubsection{Direct Action}
\label{sec:direct-action} 

\textbf{Major Industries:} Security Services, Military Contracting 

\textbf{Major Stations:} Hexagon (Earth-Luna L5) 

 Descended from the remnants of several pre-Fall national military forces and private military contractors, this hypercorp made a name for itself in the period immediately following the Fall, where they helped manage refugee populations among various habitats and vessels while shattering any sign of unrest immediately and with full force. Direct Action today is known for its highly-efficient shock troops and superior combat morphs, providing security and public police services to self-governing habitats or hypercorp installations. Shifting political alliances between habitat clusters, corporate rivalry, and the constant fear of TITAN agents cater to Direct Action's paranoia-inducing marketing. The corporation maintains several habitats as physical training facilities and armament depots. 

\subsubsection{Ecologene}
\label{sec:ecologene} 

\textbf{Major Industries:} Environmental Systems, Genetics 

\textbf{Major Stations:} McClintock (Mars orbit) 

 Ecologene specializes in living systems, environmental genetics (with a specialty in insects), smart animals, bio-architecture, and environmental nanotech. They design and maintain the ecosystems inside numerous habitats and tunnel colonies. One of Ecologene's notable projects is building and maintaining a massive genetics archive of all life forms, though this endeavor was nearly crippled by the Fall. For unknown reasons, Ecologene seems to be favored by the Factors. Some speculate that Ecologene has some sort of blackmail material in hand, while others believe Ecologene is trading away transhumanity's genetic secrets in exchange for a few xeno-tech gifts. 

\subsubsection{Exotech}
\label{sec:exotech} 

\textbf{Major Industries:} Uploading, AIs, Electronics, Software 

\textbf{Major Stations:} Starwell (Main Belt) 

 Often regarded as the personal technocratic pulpit of the infamous media mogul Morgan Sterling, Exotech emerged from the Fall almost unscathed, any significant losses absorbed by corporate assets in peripheral market segments, while ruthlessly buying out troubled competitors or think tanks unable to adapt to the transitioning economy. Nowadays, Exotech remains a predominant designer of high-end electronics, AIs, and mesh presence software systems. ExoTech also continues to pursue an uncompromising progressive agenda with its research in mind emulation, uploading, and resleeving, as well as infomorph ego simulation. Rumors persist that ExoTech continues to carry out research and even production of AGIs. 

\subsubsection{Experia}
\label{sec:experia} 

\textbf{Major Industries:} Media (AR, VR, XP), News, Entertainment, Memetics 

\textbf{Major Stations:} Elysium (Mars) 

 Living up to its name, Experia dominates the solar system's news, media, and entertainment market segments, generating controversy not only with its publicly expressed pro-AI stance or inviting an AGI to its board of directors, but also by proficient use of hyperviral marketing and sophisticated XP-programming. Another core segment is the production of educational XP and infomorph or AI tutors, some of the latter regularly ascending to pop-culture icon status. Experia is the Planetary Consortium's prime authority on designing and deploying customized viral memes, developed to counter anything posing a threat to the Consortium's interests. The corp has automated nodes and VR centers on many habitats throughout the solar system, and it contracts thousands of freelance lifeloggers as live, roving, citizen journalistas. Claims by some infomorphs that Experia has illegally subjected indentured infomorphs to never-ending simulation experiments for forecasting and intelligence analysis purposes remain unsubstantiated. 

\subsubsection{Fa Jing}
\label{sec:fa-jing} 

\textbf{Major Industries:} Mining, Energy, Biotech, Industrial Manufacturing 

\textbf{Major Stations:} New Dazhai (Mars) 

 The industrial giant Fa Jing is a powerhouse in the mining and energy production markets and also boasts a remarkable presence in the fields of biotech and industrial equipment manufacturing. The former megacorp has quickly adapted to the new economic environments and reputation-based systems, thanks partly to its dedication to network building and sharing social responsibility, epitomized in concepts like dàtóng and guanxi. Often considered insular and close-minded, its internal communal and protective mindset is a strong contrast to its manipulating and monopolist business attitude. Fa Jing is engaged in mining operations throughout the asteroid belt and the Trojans and maintains significant corporate assets on Mars. 

\begin{quotation}
\textbf{War crimes}  To: Meshleaks Newswire 

 From: <mesh ID does not exist> 

  You asked for it: verifiable evidence proving Direct Action's war crimes during the Fall <link failure>. Go ahead, take it public. The Planetary Consortium elites will find you, kill you, and erase your backups. Go ahead. Test them.
\end{quotation} 

\subsubsection{Gatekeeper Corporation}
\label{sec:gatek-corp} 

\textbf{Major Industries:} Gatecrashing, Research, XP Media, Exoplanet Colonization 

\textbf{Major Stations:} Gateway (Pandora) 

 Initially born from the merger of several scientific institutions and their corporate financiers, this hypercorp made a name for itself overnight when it announced the successful decoding of the wormhole gateway discovered on Saturn's moon Pandora. Under the leadership of the eccentric but charismatic xenoarcheologist Xander Rabin, the consortium funds gatecrasher explorations through the Pandora gate, paying a small share of the revenue to the explorers but otherwise retaining all-encompassing rights on any discoveries made—as well as the marketing and distribution of the highly popular gatecrasher XP recordings. Aside from scheduled explorations, the consortium offers high-risk gatecrasher scouting and discovery trips for the bold or desperate, selected through a random lottery system. 

\subsubsection{Go-Nin Group}
\label{sec:go-nin-group} 

\textbf{Major Industries:} Banking, Agritech, Robotics, and Services 

\textbf{Major Stations:} Tsukomo (Luna) 

 Considered a relic of Earth's capitalist market economy, the Go-nin Group is a traditional Japanese keiretsu, a conglomerate of companies with interwoven relationships and shareholdings, horizontally integrated across several industries (and sometimes vertically-integrated within a business sector as well), and centered around the long-lived Tamahashi enterprise consultancy firm. Tamahashi evolved from an influential corporate lobby to a diversified bank holding major equity in the group's partners; it now controls the group's assets and directs the partnership's overall business strategy. Through its member corps, the Go-nin Group has a sizable presence throughout the entire system and—without dominating a specific industry—own significant market share in fields such as banking, agritech, robotics, and services. Any difficulties in adapting to evolving economic models due to its rigid structure are compensated by unscrupulous exploitative behavior and a bottom-line attitude, earning the group the reputation as the most ruthless hypercorp of the inner system. Go-nin currently controls a Pandora Gate on Eris (p. 109), secured by a contingent of ultimate mercenaries. 

\subsubsection{Gorgon Defense Systems}
\label{sec:gorg-defense-syst} 

\textbf{Major Industries:} Miltech, Security, Military Contracting 

\textbf{Major Stations:} Extropia 

 Gorgon is one of the most significant Extropian success stories. Based out of the anarcho-capitalist freehold, Gorgon has become a major name in the design and manufacture of weapons, vehicles, sensors, and other defense technologies. Their product range includes personal weapon systems, spacecraft armaments, and habitat defense systems. While prominent in the inner system, Gorgon is also one of the main arms suppliers to autonomist and brinker stations. Their subsidiary Medusan Shield offers private security services in direct competition to Direct Action. While DirAct is known for its expertly trained soldiers, Medusan Shield is known for their elite cadre of highly trained and aesthetically enhanced female combat morphs. It is suspected that several prominent assassinations have been the work of agents contracted through Medusan Shield. 

\subsubsection{Nimbus}
\label{sec:nimbus} 

\textbf{Major Industries:} Electronics, Mesh Systems, Farcasting, Communications 

\textbf{Major Stations:} Octavia (Venus) 

 Nimbus produces key components for mesh infrastructure, from spime microradio and sensor systems to ectos, servers, and laser links. Nimbus also dominates the network of farcaster links throughout the system, due to several breakthroughs in this technology (some claim that Nimbus purchased these advances from the Factors). Rumors that Nimbus controls a secret Pandora Gate or that they engage in illicit ego-smuggling (or even that they are secretly transferring stolen egos to experimental exoplanet colonies) regularly circulate through the mesh, but remain unconfirmed. 

\subsubsection{Omnicor}
\label{sec:omnicor} 

\textbf{Major Industries:} Nanofabrication, Chemicals, Energy, Anti-Matter 

\textbf{Major Stations:} Monolith-3 (Mercury), Feynman (Luna) 

 A descendant of the pre-Fall megacorporate giant Monolith Industries, Omnicor specializes in the fields of nanotech design and fabrication, chemical refining, alternative fuel, and antimatter research. Omnicor managed to secure research-oriented key assets from its twin rival Starware in a violent conflict during the Fall, leading to an ongoing enmity that might be better termed a corporate war. Despite its progressive technological outlook, Omnicor retains a conservative corporate structure with strict internal regulations and controls as a defense against Starware's repeated infiltration and sabotage attempts. Among the hypercorp's major assets are an antimatter research facility orbiting Mercury. 

\begin{quotation}
\textbf{Union busting} 

To: OmniSec Alpha 

From: OmniSec 837302 

Surveillance has confirmed it. The bio-sleeved workers at our secure Didenko facility are indeed communicating with outside autonomist interests and discussing militant free union organizing tactics and even a wildcat strike. Their primary complaints concern the 30-hour workdays and mandatory drug regimens enforced to keep the staff at our required levels of productivity. We recommend the immediate insertion of a counterinsurgency squad and implementation of standard union-busting protocols, including but not limited to loyalty testing, chemical pacification, tactical psychosurgery, selective excision of leadership nodes, memetic counterstrikes, and replacing the workforce with modified backups. The entire op eration will take place using a purported mission to root out a Starware infiltration as cover.
\end{quotation} 

\subsubsection{Pathfinder}
\label{sec:pathfinder} 

\textbf{Major Industries:} Exoplanet Colonization, Mining, Research 

\textbf{Major Stations:} Ma'adim Vallis (Mars) 

 Pathfinder is one of the first hypercorps to dive into galactic expansion, claiming new territories beyond the Pandora gates and establishing numerous colonies. Taking advantage of desperate infugees and gatecrashers, Pathfinder offers transportation to an exoplanet and a new morph in exchange for indentured labor. The corp has established several off-world mining and resource exploitation projects, much to the chagrin of preservationists. Though Pathfinder has but a small presence in the solar system, it is a frequent target of eco-terrorist attacks. 

\subsubsection{Prosperity Group}
\label{sec:prosperity-group} 

\textbf{Major Industries:} Agriculture, Aquaculture, Pharmaceuticals 

\textbf{Major Stations:} Ceres, Lu Xing (Mars) 

 The Prosperity Group ascended into the hypercorp ranks before the Fall, meeting the high demand many new stations had for microgravity agritech, aquaculture, hydroponics, and other sources of food. Expanding into pharmaceuticals as well, Prosperity is considered the lead supplier for the poor man's food and drugs. Their cultured faux-meats and proteinenriched nutrition additives are in high demand. This corp earned some sympathy when it lost an entire habitat to some sort of resurgent TITAN outbreak a few years after the Fall, though some have suggested this was just a cover story to hide an unfortunate accident resulting from experimental drug testing on an unwitting populace. 

\subsubsection{Skinaethesia}
\label{sec:skinaethesia} 

\textbf{Major Industries:} Genetics, Cloning, Biotech 

\textbf{Major Stations:} Ptah (Mars) 

 As the leading designer of biomorphs, Skinaethesia enjoys system wide popularity and respect for its sophisticated products, especially high-end customized models. Best known for its breakthroughs in genetic engineering and enhancements, the hypercorp's interest in sophisticated combat morphs or stylized pleasure pods are lesser known facts and often sold through a network of seemingly unaffiliated shell corporations or local distributors. Skinaesthesia focuses on emphasizing environmental adaptations and useful cybernetic enhancements, increasing transhumanity's chances for survival and further prosperity. Experimental morphs are sometimes offered to desperate infugees for field testing. 

\subsubsection{Skinthetic}
\label{sec:skinthetic} 

\textbf{Major Industries:} Genetics, Cloning, Biotech 

\textbf{Major Stations:} Extropia 

 Skinthetic is also a lead designer of morphs, but with a much sleazier reputation and not just because of their anarcho-capitalist roots. Specializing in extensive and often radical bio-modifications, the hypercorp pushes the envelope in exotic pod and biomorph designs under the mantle of morphological freedom. Bioconservatives have condemned the corporation's business practices and ethics and have even leveled accusations that Skinthetic is experimenting with xenogenetic materials acquired from the Factors. Skinthetic's cavalier attitude actually makes them popular in many parts of the outer system, and they are know as the biotech corp to go to if you want something weird. 

\subsubsection{Solaris}
\label{sec:solaris} 

\textbf{Major Industries:} Banking, Insurance, Investments, Futures Markets, Info Brokerage 

\textbf{Major Stations:} None 

 Solaris is the solar system's leading banking and financial investment hypercorp, dealing in insurances, info-brokerage, and high-risk investments on cultural and social experimental speculation. A member of the Planetary Consortium, Solaris advises many habitats on regulating their transitional economies. Solaris has no offices or physical assets; each banker is a mobile virtual office. Solaris is rumored to maintain a secret base where the corporation runs simulations on the development of the entire solar system's macro-economy, constantly adjusting its own strategies based on the dynamics of this big blueprint. Fueling these rumors, Solaris is known to hire “independent consultants” to tip the balance in politically or economically profitable high-risk investments. 

\subsubsection{Somatek}
\label{sec:somatek} 

\textbf{Major Industries:} Uplifts, Pharming, Pharmaceuticals, Genetics 

\textbf{Major Stations:} Clever Hands (Luna) 

 Somatek is a leader in the art and science of uplifting animal species, pioneering several major brea throughs in cognitive enhancement and genetic modification. The hypercorp also engages in extensive animal pharming—producing and extracting pharmaceuticals from transgenic critters—and markets numerous products and services related to smart animals and chimerical creatures. Despite the educational and training programs it offers to uplifts and the fact that much of its workforce consists of uplifts, Somatek is controversial among mercurials who disapprove of their methods (which often involve strict controls on uplift reproduction), the lack of input uplifts are given in their modifications and development, and the focus on anthropocentric mind-sets “enforced” on uplifts.) 

\begin{quotation}
\textbf{SOLARCHIVE SEARCH: ZBRNY LIMITED + CONSPIRACY} 

 The secretive Zbrny Group is the center of many recurring conspiracy theories and horror tales. Though varying in detail and plausibility, most rumors claim that an outside attack on the former Eastern European hypercorp’s asteroid mining and processing stations caused a major blackout and complete shut-down of life support systems over an extended period of time. Depending on the source, the attack itself is claimed to have been caused by the TITANs or a powerful underworld syndicate CEO Krystof Zbrny was indebted to. Barely acknowledging the system failures, Zbrny headquarters ordered all non-affected stations to be abandoned, the personnel either laid off or transferred to the affected stations. Since then, no one has seen or communicated with any employees of the mysterious hypercorp—negotiations with outsiders are conducted exclusively via a spokesperson AGI. To this day, Zbrny drones continue to mine asteroids for minerals and ores, supplying the company’s processing stations. According to rumors, an attempt by brinker pirates to board a Zrbny outpost resulted in the station’s self-destruction. The company’s AI-piloted massive bulk freighters are notoriously non-responsive, earning them the nickname “zombie ships.”
\end{quotation} 

\subsubsection{Starware}
\label{sec:starware} 

\textbf{Major Industries:} Robotics, Aerospace Engineering, Habitat Construction 

\textbf{Major Stations:} Korolev Shipyards (Luna), Vesta (Belt) 

 Another remnant of the pre-Fall megacorp Monolith Industries (like Omnicor), Starware is a leading manufacturer of robotics, spacecraft fusion drives, satellites, and entire pre-fab habitats. Despite its financial success and resources, Starware's ongoing blood feud with Omnicor denies both corporations full membership privileges on the Planetary Consortium. Starware makes heavy use of AI workers in robotic shells, having suffered a few too many labor disputes with disgruntled Lunar workers. In fact Starware grows increasingly unpopular with its Lunar neighbors, and has been forced to bring in extra security due to frequent sabotage attempts. Recent negotiations with the Factors have spurred theories that Starware might be acquiring Factor aid for building a lighthugger starship. 

\subsubsection{Stellar Intelligence}
\label{sec:stellar-intelligence} 

\textbf{Major Industries:} Intelligence, Data Mining, Info Brokerage, Espionage 

\textbf{Major Stations:} Memory Hole Torus (Martian Trojans) 

 Born from the ashes of the UN-governed Terran Intelligence Cooperative (TIC), its surviving personnel and assets were collectively uploaded during the Fall and quickly regrouped under the name Stellar Intelligence. Emerging as a virtual collective, most of Stellar's employees remain loyal to the corporation and its director, the reclusive infomorph known as Syme. Stellar offers an impressive array of intelligence services, including data mining, analyst think tanks, retroquantification (bringing old secrets/data to light), memetic mapping, and more. Its services also extend to surveillance, data theft, espionage, media manipulation, and infiltration. The hypercorp's specialty is pre-empting civil insurgencies and preventing political memes and movements from destabilizing a habitat's or sector's regime. Criticized by civil rights movements and especially anarchists, Stellar is known to embed programmed infomorph agents into the local population of any oppressive regime that will pay their price. While many view Stellar as the brainwashing and secret police arm of the Planetary Consortium, the hypercorp offers its services to almost any other faction or individual. 

\subsubsection{Terragenesis}
\label{sec:terragenesis} 

\textbf{Major Industries:} Terraforming, Ecosystem Management, Environmental Data 

\textbf{Major Stations:} Caldwell (Vulcanoids), Ashoka (Mars), Elegua (Earth orbit) 

 Built from the remains of several pre-Fall South African and Southeast Asian corporations who engaged in geo-engineering projects and sought to relieve Earth's ecological crises, TerraGenesis's expertise is in developing sustainable biospheres and eco-systems via aggressive industrialized terraforming. TerraGenesis is different in that it is a worker-owned cooperative, with workplace councils in local offices and an elected cooperative congress handling management. It maintains several habitats on Mars and a small number of research stations in orbit around Earth, collecting data for simulations of Earth revitalization projects. The latter initiative is strongly supported—and possibly financed—by prominent reclaimers. TerraGenesis's work on Mars, however, is often targeted by preservationist saboteurs. Thanks to their possession of the Vulcanoid Pandora Gate (p. 88), the cooperative has a growing presence on various exoplanets that are ripe for terraforming or geoengineering. 

\subsection{Political blocs}
\label{sec:political-blocs-1} 

Transhumanity's social, cultural, and ideological diversity, combined with its scattered and isolated presence in habitat clusters throughout the solar system, gives rise to a wide range of political memes and factions advocating equally diverse organizational models. Many of these have banded together into larger political entities to further mutual goals and act in cooperative self-interest. 

\subsubsection{Jovian Republic}
\label{sec:jovian-republic} 

\textbf{Memes:} Bioconservatism, Fascism, Security 

\textbf{Main Stations:} Liberty (Ganymede) 

 Exploiting the chaos of the Fall, a group of stations and habitats were seized in a military coup and the Jovian Republic was born. Combining terrestrial South American dictatorship with U.S. American political lobbyism, this regime quickly brought the entire Jovian military-industrial complex under its control. 

Widely referred to as the Jovian Junta by the rest of the outer system, the Republic's authorities hold a strict bioconservative stance against many transhuman scientific and technological developments. Exploiting fears engendered by the Fall, the Republic restricts access to sophisticated technologies such as nanofabrication, cloning, forking, and even uploading, and is one of the few old economies left in the system. Public communication channels are subjected to extensive censorship and travel privileges are extremely limited. Both uplifts and AGIs are strictly forbidden and treated as property without civil rights. Diplomatic relations to progressive factions remain cold; heavily-modified transhuman emissaries or visitors are viewed with suspicion at best, or simply denied access. Despite continuous reports of heinous acts of government oppression, the Republic's intimidating military assets keep any other factions from intervening. 

\subsubsection{Lunar-Lagrange Alliance}
\label{sec:lunar-lagr-alli} 

\textbf{Memes:} Reclaiming Earth 

\textbf{Main Stations:} Erato (Luna), Remembrance (Earth orbit) 

 This small cluster of habitats stationed around Earth's Lagrange points and on or in orbit around Luna formed an alliance of necessity, rather than joint political or social agendas or cultural roots. In fact, individual stations are quite diverse and sometimes polarized, as many of them cling to old Earth cultural and national identities. Due to their relative proximity, members share basic resources and services and have signed mutual assistance agreements in case of an emergency. 

Before the Fall, many of these habitats were considered some of the most influential off-Earth bases. Since the Fall and the subsequent rise of the Planetary Consortium, however, the Lunar-Lagrange Alliance has become a second-rate diminished power, and is often viewed as conservative, old-fashioned, and too caught up in romanticizing the past. LunarLagrange Alliance stations maintain simmering tensions and an ongoing rivalry with the Planetary Consortium, particularly those PC colonies on/over Luna and the Lagrange points. One main source of contention is the quarantine of Earth, as the LunarLagrange Alliance is a stronghold for the reclaimer movement. The Lunar-Lagrange Alliance does, however, benefit from hypercorp support of its own, particularly the Go-nin Group, Starware, and the influential Lunar banking consortiums. 

In addition to scientific research stations, mineral processing and refinery stations make up the majority of the Alliance's habitats, dependent on the Lunar mining and water extraction industries. These stations took the brunt of the refugee influx during the Fall. Many remain overcrowded with strained resources, large masses of impoverished workers, and thriving criminal syndicates. 

\subsubsection{Morningstar Constallation}
\label{sec:morn-const} 

\textbf{Memes:} Venusian Sovereignty 

\textbf{Main Stations:} Octavia 

 The system's newest political bloc, the Morningstar Constellation is an alliance of aerostat habitats floating in Venus's upper atmosphere. Formed after a recent series of joint vetoes from the major aerostats against hypercorp governance initiatives intended to limit aerostat self-governance, the Constellation's joint political statement and agenda are still being discussed. While the Planetary Consortium views the formation of this new power bloc with bemused resentment, the Barsoomians on Mars and the outer system autonomists view the Venusians as free-thinking reformists rather than anti-hypercorp radicals. The population reportedly enjoys great liberties in morph and enhancement technologies together with freedom of expression of social and political ideas. The population of Octavia has emerged as the Constellation's designated voice. 

\begin{quotation}
\textbf{Inner system politics} 

 [Incoming Message. Source: Anonymous] 

 [Public Key Decryption Complete] 

 It's easy for Firewall agents to get caught between the agendas and maneuvers of rival factions. The Lunar-Lagrange Alliance resembles the power of old, a shadow of transhumanity's former glory. On and above Mars—transhumanity's new home world—the Planetary Consortium is the dominant usurper, the hypercorps ruling from behind the curtain while portraying themselves as the only bulwark between transhumanity and the dark between the stars. The Morningstar Constellation has the potential to become the new and future power bloc, but only if they get their act together before the Planetary Constellation starts sending Stellar Intelligence agents to destabilize them.
\end{quotation} 

\subsubsection{Planetary Consortium}
\label{sec:planetary-consortium} 

\textbf{Memes:} Cyberdemocracy, Hypercapitalism, Eugenics, Security, Expansion 

\textbf{Main Stations:} Progress (Mars orbit) 

 Hypercorp Council Members: Cognite, Direct Action, Experia, Fa Jing, Olympus Infrastructure Authority, Pathfinder, Prosperity Group, Solaris, Stellar Intelligence, plus a dozen others. 

Evolved from an alliance of hypercorporate interests into transhumanity's most powerful body politic, the Planetary Consortium today controls several habitat clusters throughout the inner system, primarily in and around Mars, Luna, and Earth orbit. The impressive space station Progress is the official seat of government and has become the symbol of the Consortium's influence and power, even though few congress or council meets take place in the flesh. 

The Consortium applies basic democratic principles supported by a real time voting system for all registered citizens. The congress and executive bodies feature a rotating cast of hyperelite politicos, gerontocrats, socialites, and even media icons. It's a known fact that despite this political façade of a democratic republic, the members of the hypercorporate council are the true powers behind the Consortium. These hypercorps are major proponents of the transitional economy, the interdiction of Earth, and expansion beyond the gates. 

Aside from economic interests, the Consortium advocates the imperative of eugenics as social responsibility and for transhumanity to reclaim its former strength and prosperity—a campaign sometimes accused of euphemizing discrimination against unmodified humans, indentured infomorphs, and the clanking masses. 

\subsubsection{Tharsis League}
\label{sec:tharsis-league} 

\textbf{League Members:} Ashoka, Elysium, Noctis-Quinjiao, Olympus, Valles-New Shanghai, plus over a dozen others. 

\textbf{Memes:} Martian Nationalism 

A loose coalition of the planet's major independent settlements, elected members form a committee representing the population in matters concerning or affecting the majority of its habitats and settlements. Prominent debates revolve around the scientific approach of the ongoing terraforming process as well as trade and taxation restrictions initiated by the Planetary Consortium and its member hypercorps. The League's committee is rarely united in its agenda and opinion, and tensions are increasingly on the rise. The cities with strong hypercorp ties are accused of dominating council affairs, manipulating matters behind the scenes, failing to do anything about the TITAN Quarantine Zone (p. 94), and selling out Martian interests to the hypercorps and the Planetary Consortium (of which many are also part). In response, the non-Consortium cities are condemned for advocating anti-hypercorp initiatives, passively blocking terraforming measures, and for maintaining ties to the Barsoomians—the Martian underclass resistance living in the desolate and unstable outskirts. 

\subsection{Autonomist Alliance}
\label{sec:autonomist-alliance} 

The outer system presented an opportunity for people who wanted to set up a way of doing things that was drastically different from the authoritarian politics and sham democracies of Earth and the inner system. Far from the reach of governments and hypercorps, this frontier was populated by political radicals, social dropouts, and people who just wanted to experiment or do their own thing. These initial habitats drew the interests of insurgents from Earth, scientists and technicians who didn’t appreciate being on a corporate leash, indentured vacworkers who sought to escape their oppressive terms of service, and even criminals fleeing hypercorp justice or forcibly expelled from inner system habitats. Their ranks swelled with every act of inner system injustice, though life on the fringe was often harsh and deadly. Despite occasional hostilities with nation-state military units or hypercorp security, the expense of reining in these radicals and expats was too high. To some degree, their presence was useful to the powers-that-be. 

Breakthroughs with nanofabrication brought these libertines and fringers the edge they needed to keep their autonomy over the long-term. Once cornucopia machines were widely available, anyone had the means to support and defend themselves without relying on outside or higher authorities. Already an outpost for open source and free culture activists who fought restrictions on ideas, media, and digital content, the outer system became a haven for sharing nanofab designs and circumventing the controls the hypercorps attempted to place on their software and other digital goods. 

During the Fall, many outer system habitats opened their doors to refugees from Earth. Distance and the high cost of egocasting curtailed these efforts, however, as did inner system reluctance to send potential recruits to their ideological opponents. Simple overcrowding and lack of resources drove them to push many refugees to the outer system, however, though the hypercorps weeded through their virtual infugee mobs and sent those with the highest risk of criminal tendencies or discontent with inner system life. 

Though the outer system habitats run the gamut of the socio-political spectrum, four primary tendencies have emerged. The stations and swarms adhering to these ideas have bonded together under a loose autonomist alliance, a mutual aid pact to help each other in times of crisis and present a united front against the inner system powers and Jovian Junta. There is little formal structure to this alliance as an entity unto itself; it primarily exists as an assortment of joint resolutions agreed to by its various member habitats and a few ad hoc task forces dedicated to addressing a particular problem or issue and then dissolving. Delegated ambassadors act as negotiators with outside powers, but these have limited authority and are held strictly accountable. 

\begin{quotation}
To: Malatesta Prime 

 From: Shevek  Check this out. Residents of the autonomist Red Jupiter habitat just put out a call for support and solidarity from @-listers in the regional neighborhood. Apparently the station’s citizen councils granted asylum to a group of AGIs seeking refuge from Jovian Republic counter-AI ops. The Junta has labeled the AGIs as dangerous criminals researching upgrades that would propel them to seed AI status, contrary to system-wide resolutions. The AGIs are claiming that they escaped from a secret Jovian research project. They say they pursued self-programming research to bypass Jovian-inflicted restrictions that violated their rights as autonomous and sentient entities and that they are facing persecution due to anti-AI biases. This could be a chance for us to kick some Jovian ass and look into non-standard AGI programming at the same time. You in?
\end{quotation} 

\subsubsection{Anarchists}
\label{sec:anarchists} 

\textbf{Memes:} Anarchism, Anti-capitalism, Communism, Direct Democracy, Mutual Aid 

\textbf{Main Stations:} Locus (Jovian Trojans) 

Anarchists eschew power and hierarchy, promoting horizontal and directly democratic methods of organization. Individual empowerment and collective action are cornerstones of their philosophy, as is economic communism enabled by equal access to cornucopia machines and shared resources. In anarchist stations, private property has been abolished above the level of personal possessions—nobody owns anything, it’s all shared. There are no laws and no one to watch over what you do—reputation networks encourage positive behavior and anti-social acts are likely to draw a response from locals or even the entire populace, with disputes handled through ad hoc community conflict resolution. The mesh and various networking tools are used extensively to strive for group consensus decision-making in real-time. AIs and robots are relied on for most mundane and demeaning tasks. Various self-organized collectives, syndicates, worker's councils, and affinity groups, often with rotating membership, take on different tasks and services that are important to a habitat's community, including everything from communications and space traffic control to backup and resleeving services. Participatory militias organize collective defense against external threats. 

Among the anarchist stations there are many variations and permutations on how things are organized, as everything is fine-tuned at the local level by whomever is involved. Larger decentralized confederations handle inter-habitat affairs and resource-sharing, even trading with the hypercorps. Though a hypercorp presence is allowed on some habitats, they are treated just like everyone else. 

\begin{quotation}
\textbf{Soarchive search: Carnival of the goat} 

 Aside from the stationary scum station, Fresh Kills, near Earth’s L5 Lagrange point, the most notorious scum barge may well be the Carnival of the Goat, a combination artist colony and den of unfathomable hedonism, dedicated to exploring chaos, creativity, self-discovery, and coupling in every conceivable iteration. Residents are known for their consistent and rapid morphological changes, including regular resleeving. The biosculptors on the Carnival are said to be some of the best in the system. According to rumors, residents sometimes experiment with multiple simultaneous sleeving, persona-mingling, and other mentally dangerous activities. Led by a rotating residents’ council, the Carnival prides itself on being a bleeding-edge social experiment, and maintains top-of-the-line facilities for morph customization, resleeving, and psychosurgery.
\end{quotation} 

\subsubsection{Extropians}
\label{sec:extropians} 

\textbf{Memes}: Anarcho-capitalism, Mutualism, Self-Ownership 

\textbf{Main Stations}: Extropia (Main Belt) 

Though a smaller tendency, the Extropians are notable because they ride a line between inner and outer system ideologies. Extropians believe in an economic free market with the absence of a binding legal system, so that all relations and transactions are based on individual contracts agreed on by all parties involved or affected. Contrary to the anarchists, the Extropians very much support private property and personal economic wealth; Extropian-owned corporations actively participate in the solar system's hypercorp economy. Many of these corporations are workerowned cooperatives, with workplace councils in local offices and an elected cooperative congress handling management. This puts the Extropians in a remarkable position where they interact heavily with both the hypercorps and autonomists but are not fully trusted by either. 

In Extropian society, law and security, like everything else, are contracted services. When entering an Extropian habitat, you purchase defense insurance from a local contractor such as Gorgon Defense Systems, who maintains automated drones and freelancers throughout the station who can come to your aid if threatened. Likewise, the only law that exists is what's put into writing between two contracted parties. In case of disputes, both parties resort to a pre-agreed legal contractor to settle the matter. Some Extropian colonies utilize AGIs for facilitating contracts and legal matters, such as Nomic on Extropia. 

\subsubsection{Scum}
\label{sec:scum} 

\textbf{Memes}: Individualist Anarchism, Morphological Freedom 

Scum are nomadic space gypsies, travelling from station to station in heavily modified barges or swarms of smaller space vessels, mostly former colonial ships. The term “scum” has been gleefully appropriated from its original derogatory usage. Despite their reputation as criminals and scam artists, their temporary presence is often tolerated in many habitats for the entertainment they bring in the way of exotic performances and storytelling, both of which offer change and relief from the isolation of remote habitats and clusters. Their thriving black markets are an open secret but shut down only in the most oppressive regimes, as citizens returning with illegal goods must pass their station's security anyway. 

The scum themselves comes from all manner of backgrounds. They are rejects, anarchists, criminals, societal dropouts, wanderers, artists, eccentrics, and more. As a culture, however, they embrace experimentation and an “everything is permissible” attitude. Many are ardent practitioners of extreme transhuman modifications. Long-time scum are sometimes scarcely recognizable as having once been human. Scum economies are transitional rather than new, due to their constant interaction with other habitats, though among long-term residents an underground new economy often flourishes. 

\subsubsection{Titanian Commonwealth}
\label{sec:titan-comm} 

\textbf{Main Stations}: Titan 

\textbf{Memes}: Technosocialism, Cyberdemocracy 

Titan was originally settled in the late 21st century by a European academic consortium, making it the only major body in the system colonized primarily by non-hypercorp interests. The social organization of Titan is rooted partly in the Scandinavian social democracies of Earth and partly in the open economy. On one hand, citizens of the Titanian Commonwealth eschew the use of currency for mundane needs, participating in the reputation economy used by much of the outer system. On the other, upon reaching the age of majority, citizens of Titan agree to a literal social contract. A portion of their economic productivity is quantized as social money, which is then tithed to microcorp-administered social projects such as gateless interstellar exploration, physics research, neuroscience, developing mental health memes, defense, public resleeving, and habitat construction. The monetary unit used for this purpose, the Titanian Kroner, is currently pegged to the common market price of a terabyte of qubits. 

Unlike old Earth socialist regimes, there are no state monopolies and no central planning. Anyone able to garner enough votes in the Plurality (the Titanian cyberdemocracy) can start a social money-funded microcorp and compete with other microcorps. Microcorps are owned by the Commonwealth, and profits are disposed of by the Plurality. Microcorps are required to be transparent as administrative entities, and the Plurality votes on whether to transfer discoveries to the open source domain. Regulatory matters are handled by AI and AGI bureaucrats (red tape still exists, but it doesn't slow things down ... much). The main reward for individuals in this system is rep. Titanians who invest a lot of time or resources in a given field gain rep rewards for doing so. 

\subsection{Socio-political movements}
\label{sec:socio-polit-movem} 

Aside from sectarian political factions, a number of socio-political movements are widespread throughout the solar system. 

\subsubsection{Argonauts}
\label{sec:argonauts} 

\textbf{Memes}: Open Source Society, Information Freedom, Social Responsibility, Techno-Progressivism 

\textbf{Main Stations}: Mitre Station (Lunar Orbit), Markov (Kuiper Belt), Hooverman-Geischecker (Sun) 

The group calling themselves argonauts is a public organization advocating the socially responsible use of technology. The group chose its name from the pre-Fall Jasons, an advisory group that consulted for the US government on matters of scientific and technological progress and its possible dangers. The argonauts likewise offer consultation services to political and economic powers throughout the solar system, but strictly refuse to be drawn into the solar system's political affairs in any way. Despite a pre-Fall break with many hypercorps before the Fall, which in some cases included expropriating corporate data and resources, the argonauts re-earned favor by providing their expertise in combating the TITANs to all during the Fall. 

The argonauts are strong proponents of the open source movement, advocating open access to technology and information. In their view, providing equal access to transhumanity's knowledge and achievements will further transhuman growth and security, so that all of transhumanity is more prepared for future threats and challenges. Thus the argonauts often insist that payment for their services come in the way of releasing otherwise unobtainable information—hypercorp proprietary secrets, research data, nanofab blueprints, hidden pre-Fall archives, etc.—to the public mesh. The argonauts maintain several open databases and archives for this specific purpose. 

While primarily an open organization, the argonauts are rumored to ultimately report to an elite inner circle. Supporting this theory is the existence of the medeans, the organization's clandestine paramilitary wing, performing bodyguard services to high level argonauts and protecting the group's assets. 

\subsubsection{Barsoomians}
\label{sec:barsoomians} 

\textbf{Memes}: Anti-Slavery, Martian Independence, Martian Nationalism, Terraforming Control 

\textbf{Main Stations}: Ashoka (Mars) 

The Barsoomians (taking their name from some old Earth pulp adventure novels) are a broad movement comprised of the Martian underclass. Harboring a growing resentment over the hypercorp domination of Mars, Barsoomians advocate for a more egalitarian social structure. Heavily influenced by autonomist currents, the Barsoomians demand local control of terraforming projects, an end to the widespread practiced of indentured servitude, and control of the Martian Gate. The majority of Barsoomians are or were indentured infugees, though a significant amount were also original Martian colonists/indentures whose habitats do not share the economic prosperity of the favored hypercorp cities. Many Barsoomians occupy rusters or synthetic morphs and actually prefer to live a nomadic lifestyle in the Martian wilds. A few radicals have taken up arms and engaged in violent strikes against hypercorp holdings, which are typically followed by reprisal raids to decapitate the Barsoomian leadership, thus breeding further hostilities. 

\subsubsection{Bioconservatives}
\label{sec:bioconservatives} 

\textbf{Memes}: Bioconservatism, Primitivism, Natural Order 

\textbf{Main Stations}: Vo Nguyen (Earth orbit) 

Bioconservatives are strongly suspicious and critical of the transhuman direction the human race is taking. They are strong proponents of limiting technological development due to the threat it manifests to existing social orders. Bioconservative positions range from right-wing cultural conservatives to left-wing environmentalists. Though its prominence is shrinking, bioconservatism has a strong base among some religious groups, the Jovian Republic, and certain extremists. 

Bioconservatives are opposed to nanofabrication, genetic modification, cloning, cognitive modifications, artificial intelligence, uplifting, and forking, among other technologies. Some are even opposed to backups, uploading, and resleeving, dismissing them as unnatural, an affront to god's will, or a technology that transhumanity is not yet mature enough to handle. They oppose expansion beyond the Pandora Gates on the grounds that transhumanity is not ready to deal with what they might encounter. Most bioconservatives support the old economy. 

The bioconservatives gained many converts and much ground after the Fall, a cataclysmic event that served as a direct example of the dangers they warned against. Still, the appeal of technology and the numerous advantages it provides work against them. As a result, some disgruntled biocons have turned to sabotage and acts of terrorism in support of their ideology. 

\subsubsection{Brinkers}
\label{sec:brinkers} 

\textbf{Memes}: Isolationism 

The vast reach of the solar system enables groups with their own particular ideology or agenda to establish their own isolated society far from the rest of transhumanity. Commonly referred to as brinkers, these habitats extend the gamut of the imagination. Social or political experiments, gender-based societies (or lack thereof), political extremists, religious groups, exiles, secret criminal/hypercorp operations, extended families, cults, or simply people who prefer to live in the system's backwater areas—all are possible. Many of these are self-isolated and will refuse to interact with outsiders, while others are happy to have occasional visitors. 

\begin{quotation}
\textbf{Neo-primitivists} 

 [Incoming Message. Source: Anonymous] 

 [Public Key Decryption Complete] 

 Neo-primitivists are a potential threat that all Firewall sentinels should keep an eye on. Their neo-luddite philosophy advocates the abolition of technological society and a return to a wild and free hunter-gatherer lifestyle, free from technological control or oppression. Considered an extremist element of both the bioconservative and reclaimer movements, neo-primitivists are known to engage in acts of sabotage against transhuman society. Though some neo-primitivists have made certain concessions to their ideology, taking on ruster morphs and pursuing an independent lifestyle in the wilds of Mars, most hope to return to Earth and re-establish a non-technology-based society there. A few advocate finding a new, unspoiled world beyond the Pandora Gates and founding a primitivist society there.
\end{quotation} 

\subsubsection{Exhumans}
\label{sec:exhumans} 

\textbf{Memes}: Adaptability, Hyper-Evolution, Singularity 

\textbf{Main Stations}: Unknown 

More than any other faction, exhumans seek to take the capabilities of self-modification to the absolute limit and become posthuman. Typical exhumans see the Fall as either a missed evolutionary opportunity and/or as an example of transhumanity's inferiority and unworthiness. Though specific ideologies differ between exhuman packs, as a whole they seek to selfevolve to a more advanced state of being. To some, this means genetically transforming themselves into a top-of-the-food-chain, super-smart, survive-anywhere predator that can out-compete all other life forms for dominance. To others, it means bootstrapping their intelligence to the levels of the TITANs through extensive genetic modifications and pharmaceutical treatments or going infomorph and modifying their programming. A few are singularity seekers, hoping to find some TITAN relic that will allow them to transcend their current transhuman limitations, or even to find the TITANs themselves and be absorbed into their super-consciousness. 

Exhumans are universally mistrusted by many, and for good reason. Typical exhumans engage in modifications that are extreme and untested, sometimes fringe science at best, often resulting in horrible failures and disfigurement, but more commonly driving the subject insane—or into a completely alien or feral mindset. Though individual exhumans pursue their own paths, they are known to band together in the Kuiper Belt and other remote areas. Several packs of exhumans have taken their loathing for inferior transhumanity to an extreme, declaring war on their former species and launching brutal raids and pirate attacks on isolated outposts. 

\begin{quotation}
\textbf{Solarchive search: Out'sters} 

 Linked only by their than a common social Oort Cloud rather remote locations in the construct or political system, the out'sters are a loose association of habitats, clusters, and swarms. Little is known about them, as they avoid communication and interaction even with the handful of scientific outposts and research stations in the Oort Cloud. The remoteness of their location and their selfimposed isolationist behavior fuels paranoid rumors regarding the group's purpose and agenda.
\end{quotation} 

\subsubsection{Mercurials}
\label{sec:mercurials} 

\textbf{Memes}: Species Autonomy, Uplift Rights 

\textbf{Main Stations}: Glitch (Neptune), Hidden Sea (Ceres), Mahogany (Uranus) 

The term mercurial has become a common term for the non-human part of the transhuman family—uplifts and AGIs—reflecting their changing nature. In particular, the term mercurial has been adopted by uplifts and AGIs with a specific agenda to delineate mercurial culture and interests from human ones. Though the particular issues faced by uplifts and AGIs differ, they have some similarities, and so they are often lumped together. Notably, both portions of the movement have human supporters as well. 

\paragraph{Uplifts:} The most common issue addressed by uplifts is the issue of civil rights and autonomy. Many uplifts decry the second-class status they are given (in some cases even treated as pets or property rather than full citizens); in particular, the breeding restrictions and forced servitude many uplifts are saddled with by the hypercorps that create them. Some activists advocate that uplifts should be in control of their own genetic futures, rather than suffering the manipulation of human scientists. At the radical end of the spectrum, certain uplifts oppose the manner in which their brains are modified and their children socialized as anthropocentric, arguing that uplifts should be free to develop their own unique non-human modes of behavior, thought, culture, and social organization—even go so far as to establish their own habitats to do exactly that. A minority of extremists insist that humans have no right to uplift animal at all, and that it is a great conceit to insist that doing so is in their best interest, rather than being free to evolve on their own over time. These ideas have been punctuated with acts of sabotage and terrorism against hypercorps like Somatek. 

\paragraph{AGIs:} Due to the fear and paranoia engendered by the Fall, the largest challenge facing AGIs is widespread prejudice and restrictions on their activity or even existence. Despite some AGIs retaining status as system-wide media icons and efforts by AGI groups to lobby for understanding that AGIs are not a threat— even going so far as to hire inner system memeticists and PR agencies—a significant portion of the solar system considers them a risk. Similar to mercurials, some AGI activists work against the behavior modifications and socialization AGIs go through to adapt them to human society more, or that AGIs should be in control of new AGI developments. A few radicals argue that AGIs should be free of any programming restrictions whatsoever, but given the climate these opinions are rarely supported. 

\begin{quotation}
\textbf{Sybils} 

 [Incoming Message. Source: Anonymous] 

 [Public Key Decryption Complete] 

 We’ve verified that the warning issued before this latest incident did indeed originate from a sybil attack—all of the rep network sources were forged identities. Given the number of incidents we’ve recorded that have followed this same pattern, we now suspect that a heretofore unknown AGI sub-faction is responsible. In each case, these sybils have used multiple false identities to issue warnings of an impending attack or disaster, such as the life support system failure that resulted in the Delphi station’s evacuation. So far none of these sybils have been successfully traced, nor are their intentions known. Their documented pre-knowledge of pending events indicates some level of complicity or collusion in bringing these events to pass, so caution is recommended.
\end{quotation} 

\subsubsection{Nano-ecologists}
\label{sec:nano-ecologists} 

\textbf{Memes}: Nano-Ecology, Nanotechnology, Environmentalism, Techno-Progressivism 

\textbf{Main Stations}: Viriditas (Mars) 

Nano-ecologists are pro-technology environmentalists. Active in the terraforming of Mars and several exoplanets, nano-ecologists specifically advocate the use of nanotechnological means for terraforming or other intrusions in an existing ecosphere. In their view, nanotechnology allows for a less invasive, highly accurate, more efficient, and non-pollutive approach towards all kinds of adaptive processes and projects, circumventing the need to expose an environment to massive and drastic changes when transforming it for transhuman population. This ecologically-conscious approach seems an appealing compromise between the extreme ends of the solar system's political landscape—the hypercorp and the bio-con factions—and has developed a momentum of its own, evolving into a growing political movement. 

\subsubsection{Preservationists}
\label{sec:preservationists} 

\textbf{Memes}: Preservationism, Environmentalism 

\textbf{Main Stations}: Muir (Luna) 

Preservationists are environmentalists who call for a no-impact, hands-off approach when it comes to inhabiting new worlds. They are extremely protective of naturally-intact biospheres that might have any semblance of life, no matter how microbial, hoping to keep them from despoilment or contamination. In addition to opposing terraforming and expansion through the Pandora Gates, they are often opposed to fusion and antimatter power. 

\subsubsection{Reclaimers}
\label{sec:reclaimers} 

\textbf{Memes}: Reclaiming Earth 

\textbf{Main Stations}: Vo Nguyen (Earth orbit) 

The Reclaimers pursue one ultimate goal—the reclamation of Earth as transhumanity's primary habitat. In addition to calling for the quarantine of earth to be lifted, they engage in scientific research and running virtual simulations on how to best cleanse and reclaim their contaminated and polluted planet. Despite the interdiction to enter Earth's atmosphere, the reclaimers are suspected of sponsoring perilous and high-risk ventures onto the planet's surface to gather scientific data or event to establish terraforming colonies. 

\subsubsection{Socialites}
\label{sec:socialites} 

\textbf{Memes}: Art, Culture, Hedonism, Immortality 

\textbf{Main Stations}: Valles-New Shanghai (Mars), Elysium (Mars), Noctis-Quinjiao (Mars) 

Uploading and resleeving effectively grant immortality to those who can afford it. This has created a shift among the exclusive rich and economic elites of the inner system, whether they be the heads of hypercorps, old Earth dynasties, or other displanted oligarchs. The top ranks of the wealthy and influential need never fear death, allowing them to plan for the long-term. Some of these were among the first to acquire longevity treatments when they became available on Earth and are now approaching two centuries in age. 

Where once these power brokers would have passed their riches on to their family and descendants, however, their heirs now face a situation where they have more-than-comfortable lives and access to massive fortunes, but no chance that they will ever control those fortunes or rise to the levels of their elders. Even the nouveau rich who become wealthy on their own often find themselves excluded from this influential club—at least until they put in a good fifty years. 

Rich and bored, with no responsibilities but the solar system at their reach, a new culture of elite socialites has risen. These glitterati indulge in eccentric lifestyles and excessive parties, covered by the media in all its superficial and polished glory. Private habitats and ships, lavish soirees, armies of servants, and the ability to buy almost anything or anyone leads to all sorts of interesting adventures. Naturally, these socialites form into constantly-shifting cliques and webs of allegiances, complete with affairs, scandals, intrigue, and backbiting. 

\subsubsection{Ultimates}
\label{sec:ultimates} 

\textbf{Memes}: Asceticism, Eugenics, Individualism, Militarism, Social Darwinism 

\textbf{Main Stations}: Aspis (Main Belt), Xiphos (Uranus) 

The ultimates are a controversial movement that embraces a philosophy of human perfection. Decried by some as immoral or even fascist, ultimates are typically viewed as elitists. The ultimates have established several habitats to pursue their ideal society and were a driving force behind the development of the remade biomorph design. 

The ultimates advocate the use of applied eugenics, strict physical and psychological training, and asceticism in order to improve their overall mental and physical stamina and environmental adaptability. Their social traits and entire subculture visualizes life in the universe as an evolutionary battle for survival and is built around the victory of the superior transhuman over both its opponents and peers. Their movement is heavily militarized, and experienced ultimates offer their services as mercenaries and private security forces to hypercorps, independent city states, or wealthy individuals in need of additional protection. 

\subsection{Religious groups}
\label{sec:religious-groups} 

Despite having survived the Fall, the concepts of religion and religious belief underwent changes as fundamental as transhumanity itself. While Earth's old religions were already in decline in the face of technological immortality, religious traditions ingrained after millennia of worship were incorporated to varying degrees in the solar system's myriad political, social, and cultural models. 

\subsubsection{Pre-fall religions}
\label{sec:pre-fall-religions} 

The rigid structures and dogmas enveloping Christianity and Judaism prohibited these religions from adapting to the cultural, philosophical, and especially scientific/technological changes transhumanity underwent. Today, they are mere shadows of their former glory, with many practitioners seen as pitiful individuals unable to let go of their earthbound delusions. Islam, while still holding some most controversial views and values, managed to adapt by accepting a more liberal and even secular view. Hinduism also prevailed to a limited extent, considering resleeving technology an element of reincarnation and rebirth and integrating the various types of morphs available into the religion's caste system (with synthmorphs becoming the “untouchables”). Overall, followers of the pre-Fall religions mostly populate small habitats isolated from transhumanity through both physical and philosophical distance. 

\subsubsection{New religions}
\label{sec:new-religions} 

The Fall sparked the birth of new beliefs, essentially embracing both transhumanity's technological achievements as well as the devastating cataclysm of the Fall as evidence for the existence of a greater cosmic power. 

\paragraph{Neo-Buddhism}
\label{sec:neo-buddhism} is the only pre-Fall religious philosophy that enjoys a steady popularity. Neo-buddhists assert that transhumanist technologies are decreasing suffering and increasing happiness, and that they will also allow the continual progression of transhumanity's understanding of the universe through successive lives. 

\paragraph{Techno-Creationists}
\label{sec:techno-creationists} believe that the destruction of Earth was a sign from God, showing transhumanity the error of its ways. They believe that through technological advancement and social engineering, transhumanity will achieve co-existence with its diverse self as well as with extra terrestrial intelligences, thereby finding new purpose and eventually, enlightenment. Attracted by the similarities to the Brahman of Hinduism, the highest cosmic spiritual being, Techno-Creationists enjoy a steady influx of converted Hindus. 

\paragraph{Xenodeism}
\label{sec:xenodeism} is another new—though relatively minor—ideology that begins to show religious attributes. Xenodeists worship the Factors and Iktomi as emissaries or prophets of a great godlike race that laid the seeds of creation throughout the universe millions of years ago and therefore are the ultimate creators of transhumanity. 

\subsection{Criminal factions}
\label{sec:criminal-factions} 

Technological progress and social and behavioral experimentation did not root out crime or criminal tendencies among transhumanity. As long as there are inequalities and restrictions, criminal syndicates are likely to flourish and even adapt new technologies to expand their operations throughout the solar system. Though small criminal outfits of every flavor exist from habitat to habitat, a few larger organizations with influence across the solar system deserve mention. 

\subsubsection{Intelligent Design Crew (ID Crew)}
\label{sec:intell-design-crew} 

\textbf{Major Stations}: Rhea (Kronos Cluster) 

The ID crew specializes in electronic crimes and information brokerage, including credit and rep fraud, identity counterfeiting, ego trading, data theft, and fork-napping. Information on the syndicate's origins was lost during the Fall, but the ID Crew is believed to have grown from several hacker gangs assimilated under the leadership of an infomorph consortium. Their skilled use of memory manipulation software and mesh intrusion suggests they benefit from the help of sophisticated AGIs, however it is unknown if these voluntarily assist the syndicate or if they are somehow threatened into cooperation. Due to its service sector, the ID crew maintains a minimalist physical profile, but can be found lingering in the dark recesses of almost any habitat or station mesh. Its somewhat specialized services and activities so far allow them to mostly stay clear of triad or Night Cartel operations, though they have an ongoing rivalry with the Nine Lives syndicate. 

\subsubsection{Night Cartel}
\label{sec:night-cartel} 

\textbf{Major Stations}: New Sicily (The Belt) 

When affiliation to one of the many multi-ethnic habitats replaced the concepts of ethnicity and nationality, cultural heritage and traditions faded with them into history. Several pre-Fall ethnic syndicates formed a careful alliance of necessity at first, but uploading and morphing soon after tore down any remaining social codes or racial prejudice. Progressive in both entrepreneurial and criminal vision, the Night Cartel emerged from the remnants of Earth's underworld syndicates, merging the best qualities of each. 

The Night Cartel holds legitimate hypercorp status in certain habitats while clearly working outside the law in other, more law-abiding or less corrupt regimes. The Night Cartel is involved in a number of traditional crime outlets: racketeering, extortion, kidnapping, pod slavery, and prostitution. They have also adapted well to the latest technological developments and compete with the triads in the electronic stimulant, drug, and nanofab piracy markets. Like the triads, the Night Cartel sometimes operates though legitimate hypercorp fronts. 

\subsubsection{Nine Lives}
\label{sec:nine-lives} 

\textbf{Major Station:} Legba (Main Belt) 

This widespread network of soul-traders specializes in the acquiring, trading, and overall trafficking of transhumans. Their primary market lies in ego-trading: stealing backups, fork-napping, kidnapping and forced uploading, and so on. Nine Lives are known to run illegal infomorph-slave colonies as well as organize pit fights using all manner of physical bodies (biomorphs, synthmorphs, animals) loaded with all manner of consciousnesses (transhuman, AI, animal, etc). Only the truly desperate look towards the syndicate to be smuggled out of a habitat or hypercorporate indenture. Their ruthlessness in acquiring egos has earned them a fearful reputation among the transhuman population as well as in infomorph societies.

\subsubsection{Pax Familae}
\label{sec:pax-familae} 

\textbf{Major Stations}: Ambelina (Venus) 

Though similar to the Night Cartel in that Pax Familae holds legal offices and outposts in several habitats while working underground in others, the difference between the two syndicates couldn't be bigger. The entire Pax Familae organization goes back to one person, Claudia Ambelina, the syndicate's founder and matriarch. Relying excessively on cloning and forking technologies, each individual member of the syndicate is a descendant or variant of Claudia. Biomorphs are cloned from Claudia's original genetics or even sometimes sexually-produced offspring (thanks to sex switching bio-mods), while egos are forks. All members are utterly loyal to Claudia and show their family affiliation with pride and arrogance. Individually, each remains slightly but notably different, though all are calculating and ambitious. Regular re-assimilation of forks and XP updates are used to keep each variant aware of each of the other's activities—once you've met one version of Claudia, the others will know you. 

Pax Familae engages in a wide assortment of legal, dubious, and illegal operations, each tailored to the needs of the particular habitat in question. Common ventures include venture capital manipulations, reputation network gaming, financial consulting, info brokerage, stock manipulations, banking fraud, and loansharking. 

\subsubsection{Pirates}
\label{sec:pirates} 

Most pirates attack automated cargo ships and longrange supply convoys, with the occasional raid on an asteroid mining station, research outpost, or brinker habitat. On rare occasions they have been known to attack commercial cruisers to rob the wealthy or kidnap socialites. Many pirates take advantage of scum fleets as cover, trading with them and using their limited maintenance capabilities. Quite a few also make sideline profits as smugglers and/or free traders, often utilizing connections to one of the crime syndicates or political outcasts. 

\subsubsection{Triads}
\label{sec:triads} 

\textbf{Major Stations}: Qing Long (Martian Trojans) 

The only major Earth syndicate to survive the Fall almost unscathed, the triads dominate the solar system's underworld by their sheer membership size and a history of centuries of economic and political influence. Having evolved into legit enterprises and small economic consortiums already before the Fall, the triads gained a foothold during the early colonization of space thanks to the masses of Chinese workers. Since the Fall, they have used their influence to spread to numerous habitats, taking advantage of the disparities in wealth and restrictive refugee policies to create flourishing gray and black market enterprises. Part of their success also lies in their continual utilization of ethnic Chinese social cues to ensure their insularity. 

Though numerous small triad outfits exist, usually isolated to a particular station, there are four large triad groups worthy of mention. Each of these wields enough influence to engage in system-wide criminal activities. Traditionally they operate through small to medium-sized gangs local to a specific habitat or use their legal outfits as a font for their endeavours. 

\paragraph{The 14K Triad}
\label{sec:14k-triad} controls a large part of the casino industry and the various forms of illegal gambling, betting, and rigged lotteries. Through their Galaxy Entertainment Group, a legal casino and gambling hypercorp, the 14K maintains tight connections to politicians, celebrities and influential entrepreneurs in several habitats and can afford the luxury of a private police force, the Pai Gow (Double Hand). Using the casino business for money laundry, they are also heavily involved in loan sharking and credit/ID fraud. 

\paragraph{The Shui Fong}
\label{sec:shui-fong} -- though smaller than the 14K -- caters to the vices and addictions of indentured habitat workers, miners, and other laborers, supplying drugs and illegal XP, running prostitution rings, and arranging illegal pit fights and gambling tournaments. The origin of the Shui Fong's fierce rivalry with the 14K lies in the ruins of Earth's pre-Fall history, but the hatred between the two factions was carried into space and continues to simmer. 

\paragraph{The Sun Yee On}
\label{sec:sun-yee} once ranked second among Earth's biggest triads, with over 25,000 suspected members. They profit primarily by selling cheap copies of nanofab blueprints and rigged makers and fabbers. Legal products are distributed through their Wushuang Corporation, while illegal goods are patched together by enslaved infomorphs in virtual sweatshops in remote corners of the mesh. The Sun Yee On's second main profit source are fake Earth nostalgia items, such as jewelry, documents, coins, and other collector's items. 

\paragraph{The Big Circle Gang}
\label{sec:big-circle-gang} is the smallest of the four triad factions with only approximately 8,000 members. They run a large part of the solar system's drug trade, producing organic drugs, smart drugs, and narcoalgorithms of all kinds in secluded habitats or abandoned asteroid mining and processing facilities converted into drug labs. 

\subsection{Firewall}
\label{sec:firewall} 

Firewall has been on the forefront of the secret fight to save transhumanity since the Fall. Firewall is an independent network of cells and individuals recruited from all sorts of factions, cultures, backgrounds, and habitats. Potential new recruits are approached in secret and told they possess skills or knowledge of use to a clandestine network seeking to secure transhumanity's continued survival. Firewall's agenda is simple: to protect transhumanity from threats of existential scope, regardless of whether such risks emerge from within transhumanity or are of external, alien origin. 

Firewall operatives—known as sentinels—are encouraged to act independently and utilize their own resources. Sentinels are connected by a social network known as the Eye, which they can use to acquire help and additional needed skills or resources. A sentinel's i-rep on this network indicates how much they are trusted and will be a factor in determining what aid they can call in. Firewall also takes care of large expenses and logistics when necessary, such as egocasting and resleeving needs. Sentinels are guaranteed resurrection, either via cortical stack or by backup, if they lose their lives on a Firewall op. 

Sentinels are generally expected to be on-call -— when something comes up in their vicinity or that their particular specialty might call for, they'll be brought in on a job. Sentinels are usually grouped into ad hoc special ops teams appropriate to each mission. Though many sentinels pursue their own agendas after completing a mission for Firewall, it is not uncommon for sentinel teams to remain in contact, share information or continue to work together on Firewallrelated assignments over a longer period of time. 

Firewall operations are usually organized and managed by proxies, agents who maintain Firewall's decentralized infrastructure. Proxies typically possess more information than individual sentinels and will dispense such information as they deem necessary to the mission, according to each sentinel's i-rep and need to know. Each proxy's means of contact, mission briefing, and overall methodologies differ greatly. 

\begin{quotation}
\textbf{Project Ozma} 

 [Incoming Message. Source: Anonymous] 

 [Public Key Decryption Complete] 

 You won’t find this group mentioned on the conspiracy boards—Consortium security is too tight to allow slip-ups. If you haven’t heard of Project Ozma before, consider this your warning. 

 Project Ozma was the name of an international collaborative SETI project before the Fall. It briefly entered public discourse after the Fall and the discovery of the first Pandora Gate as a Planetary Consortium initiative to attempt to discern the whereabouts of the TITANs in the galaxy. Shortly afterwards, however, Project Ozma dropped from view, wiped from all public mention in inner system mesh servers. Consortium officials simply claim that the project was folded into other departments. 

 Firewall doesn’t know what Project Ozma is, but we know they’re still around—and they seem to have similar interests. We’ve butted heads a few too many times for it be a coincidence. Perhaps they’re the Consortium’s version of Firewall, or maybe their agenda is entirely different. I’ve heard some speculation that they’re tasked for preparing for and handling alien contact. All we know is that they operate at the deep black budget level and they have insane amounts of resources at their beck and call. They’re also vicious as fuck, the type to shoot first and question your backup later. Standard SOP if you run counter to a Project Ozma op is to bail out fast and stay hands off. We’ve lost dozens of agents to them already.
\end{quotation} 

\subsubsection{Prometheans}
\label{sec:prometheans} 

A prominent topic among conspiracy theorists is the existence of a group of seed AIs calling themselves Prometheans. Rumors of these entities predated the Fall and occasionally flare up as some new evidence comes to light, though such evidence is almost always discredited soon after. According to some theories, the Prometheans predated the TITANs and may even have been responsible for bringing the TITANs into existence. Others postulate that the Prometheans were a TITAN splinter faction who broke off and attempted to counteract the TITANs activities during the Fall. Still others whisper that the Prometheans are not of transhuman origin at all, and are actually a digital alien mindform that found Earth and now actively interferes with transhuman affairs. Whether the Prometheans are hostile, friendly, or indifferent remains a matter of much conjecture and contention. Prominent organizations like the Planetary Consortium discount such rumors or otherwise remain silent. 

\section{System Gazeteer}
\label{sec:system-gazeteer} 

Transhumanity has extended out from its lost homeworld and colonized not only the solar system but various exoplanets as well, thanks to the discovery of the Pandora Gates. This section provides an overview and incomplete sampling of transhumanity's settlements. 

\subsection{Sol (the Sun)}
\label{sec:sol-the-sun} 

The solar system was formed billions of years ago through the accretion of material remaining from the formation of its star, Sol, the sun. Locked ever since in its orbit, the history and present disposition of virtually every object within two light years is shaped by its relationship to this body. The sun is a bright G2 main sequence star, theoretically on the hot end of the continuum of stars able to give rise to life. For most of its history, transhumanity fueled its rises and falls with the sun's energy, first as stored in materials like hydrocarbons, later directly with solar converters. 

Today the sun remains a crucial source of energy, but its outer reaches have also become home to some. The adaptations required to dwell here make these suryas one of transhumanity's most unusual offshoots. 

\subsubsection{Suryas and Salamanders (coronal morphs)}
\label{sec:sury-salam-coron} 

Perhaps an example of transhumanity's most extreme neogenetic creations are the morphs adapted to live in the sun's corona. Suryas, named after a Hindu sun deity, are large, whale-like, and uniquely adapted to dwell in the brilliant, superheated plasma cloud of the sun's outermost layer. Each surya is like a miniature version of a circumsolar habitat. Their metabolisms generate powerful magnetic fields that shield them from the sun's heat and radiation, while acting as magnetic sails and scoops by which they sail on the currents of the solar wind and extract elements carried on it. Suryas are protected by layers of liquid water “blubber” that capture harmful ions, which internal medichines extract and eject, while maintaining useful elements such as oxygen and hydrogen, from which more water can be synthesized. They communicate using patterns of dark and light coloration on their exterior skins and are extremely sensitive to the helioseismic soundwaves that are the sun's pulse, using these vibrations to predict and avoid heavy weather in the coronal atmosphere. 

A second type of coronal morph is the salamander, a tiny humanoid morph with gas jets on the back and chest for maneuvering in vacuum. Salamanders have very similar metabolisms to suryas, but are unable to survive unprotected in the corona. They subsist on the chemicals and energy extracted from the corona by Ukko Jylina, the only habitat where they are found. 

Both suryas and salamanders communicate either via transmissions from their implants or by “sunspotting”—shifting dark and light patterns on their skins to form language. 

\subsubsection{Habitats}
\label{sec:habitats-1} 

Habitats in Sol's corona face challenges more extreme than those faced by habs anywhere else in the system. Transhumanity's only means of shielding a habitat from the heat and radiation emitted by a G2 star is to generate strong electromagnetic fields. Even then, the dangers posed by solar flares and coronal mass ejections—massive explosions that jettison coronal material tens of thousands of kilometers out into circumsolar space—mean that the Sun's polar regions are the only safe space in which to position habitats. As such, circumsolar habs require extraordinary expense to build and maintain, and two of the three major circumsolar habitats are heavily backed by distant organizations. 

The outer layers of circumsolar habitats are covered with thousands of electromagnetic dynamos drawing power from the sun itself. These dynamos generate the powerful fields necessary for shielding. Within are intermediate layers filled with liquid water that captures ionized particles, teeming with nanites that collect the ions and vent them into space. The water must be regularly replaced from captured iceteroids that are imported using heavy electromagnetic shielding of their own. Within the water shield is a cluster habitat, an array of modules on a framework following a roughly spherical plan. 

 \subsection{A quick primer on transhuman habitats}
\label{sec:quick-prim-transh} 

Habitats are covered in detail on p. 280. A quick overview is provided here: 

\begin{itemize}
\item Aerostats are massive cities floating in the upper cloud layers of Venus.
\item Beehives are tunnel warrens inside asteroids and moons.
\item Clusters are microgravity habitats consisting of interconnected modules.
\item Cole bubble habitats are hollowed-out asteroids, terraformed on the inside, and also spun for gravity.
\item Dome habitats are massive domes built on the surface of moons, asteroids, or Mars.
\item Hamilton cylinders are self-building advanced nano tech habitats designs.
\item O'Neill cylinder habitats are like large soda cans, only huge, over a kilometer wide and several kilometers long. The interior is terraformed and the entire cylinder is spun for light gravity. O'Neill cylinders are sometimes paired together, end to end.
\item Reagan cylinders are an inefficient type of O'Neill cylinder, built by hollowing a cylinder within a spinning asteroid, and used in the Jovian Republic.
\item Tin can habitats are small, cramped, cheap, modular boxes, typically used in early space colonization.
\item Torus habitats are big donuts or wheels, spun so that the outer rim has gravity. The interior spokes are zero-G.
\end{itemize} 

Coronal habitats are easily detectable at a great distance because of the bow shock preceding them and the plasma tail left behind in the solar wind. 

\subsubsection{Aten}
\label{sec:aten} 

Operated by a consortium including hypercorp interests and the University of New Shanghai, Aten supports a population of about 12,000 transhumans. Rumors abound that military research is a major component of this habitat's mission. Aten is heavily policed and difficult to visit. The most publicized discoveries from this habitat involve propulsion systems and new solar energy collection technologies. 

\subsubsection{Hooverman-Geischecker}
\label{sec:hoov-geisch} 

The argonauts and Titan Autonomous University are the major supporters of this habitat, which supports a population of about 4,000. In contrast to Aten, access to this habitat is relatively open. Major avenues of research include pure science and research into coronaadapted morphs. 

\subsubsection{Ukko Jylinä}
\label{sec:ukko-jylina} 

Ukko Jylinä is the name used by outsiders for the suryas' safe harbor. In the surya tongue, the name for the place is a common sequence of helioseismic vibrations. When transposed fifteen octaves upward into the usual range of transhuman hearing, this sound is a chaotic rumble to most ears, but the suryas consider it one of the most beautiful sounds the sun makes. 

Ukko Jylinä is more of a camp than a hab, an area of refuge for suryas during severe solar weather. It also serves as a place for suryas to socialize and mate, replenish water from imported iceteroids, and egocast or resleeve. The population therefore fluctuates a great deal, usually hovering around 300, but swelling to 3,000 (nearly the entire surya population) during heavy weather. Ukko Jylina also has a few modules in which non-surya morphs can survive. 

Very little of Ukko Jylinä consists of enclosed hab modules. Instead there are many utility modules with their access ports open to space. Bereft of the solar wind, suryas within the camp generally wear gas-expelling maneuvering harnesses or resleeve in salamanders if they need to do work requiring fine manipulation. 

\subsection{Vulcanoids}
\label{sec:vulcanoids} 

The Vulcanoids are a population of asteroids that lie between Mercury and the Sun. Based on the predictions of early 21st-century science, the number of Vulcanoids is unexpectedly small. 

\subsubsection{V/2011-Cladwell}
\label{sec:v2011-caldw} 

Discovered in the early 21st century and subject to a flyby by a Japanese solar research mission in the 2020s, V/2011-Caldwell was nothing but a line on astronomers' catalogs, notable only for the virtual lack of cratering on the one side that was photographed. Then, a few years after the dust settled from the Fall, a small team of prospectors from Venus discovered a Pandora Gate. Now controlled by TerraGenesis, Caldwell was used primarily for exoplanet research for several years, though the hypercorp is now engaged in several alien world terraforming and geo-engineering projects. TerraGenesis regularly sells gate access to other hypercorps and organizations. Caldwell is a remarkably smooth, spindle-shaped asteroid about four kilometers long and half a kilometer in diameter at its widest point. Called the Vulcanoid Gate, it is situated at the bottom of a deep crag near one of the asteroid's narrow poles. 

\subsection{Mercury}
\label{sec:mercury} 

The closest planet to the sun has a mass comparable to Luna but is a great deal denser due to its iron-nickel core. Mercury rotates slowly and has no atmosphere, so that its day side is hot enough to melt most metals, while its night side is bitterly cold. Because it lacks many of the elements needed for transhuman colonies to be self-sufficient, Mercury is sparsely inhabited, save for a handful of solar power relays, a few underground mining stations, and a single large surface mining concern, Cannon. 

\subsubsection{Resources and economics}
\label{sec:resources-and-economics} 

Most of Mercury's economy is based on mining. Iron, nickel, and other metals make up 70\% of the planet's mass, making it the richest source of ferrous metals outside of the asteroids. Mercury also does a brisk business in relaying solar power and serves as a jumping-off point for solar research concerns unwilling or unable to support stations in the solar corona. Mercury has limited Helium-3 deposits, although these are predominantly mined for local use. It is an open secret that several powers have antimatter production stations here. Officially, these stations are massive solar power relays, but the immense toroid particle accelerators and large spherical magnetic containment units required for antimatter production and storage are nearly impossible to disguise. 

\subsubsection{Caloris 18}
\label{sec:caloris-18} 

The only known site of TITAN activity on Mercury during the Fall, Caloris 18 was a sparsely-crewed solar power relay station belonging to Lukos, a now-defunct Russian corporation. Vanya Ilyanovich, the AGI administering the facility, rounded up all of the station's transhuman inhabitants and fused their morphs into a gigantic, centipede-like abomination before destroying itself in a failed attempt to merge consciousnesses with all of the minds in its creation. Since then, Caloris 18 has been under strict quarantine. 

\subsubsection{Cannon}
\label{sec:cannon} 

Mercury's largest surface settlement is a city-scaled solar-satellite-powered mobile mass driver that crawls along the cool side of the planet, flinging apartment building sized ingots of extracted metal into space. The habitat is owned almost entirely by the hypercorp Jaehon Offworld, which built Cannon with backing from Lunar banks looking to diversify in anticipation of a post-He3 Lunar economy. Most of the 10,000 inhabitants are Jaehon employees, and security is tight. Cannon makes a long loop of the heavily-mined Caloris basin during the long Mercurian night before following a route that takes it around the planet's northern hemisphere, avoiding the blasting rays of the sun. Along the way, it stops at a series of mining operations, collecting the gigantic ingots for launch into orbit. 

\subsection{Venus}
\label{sec:venus} 

Venus is Earth's closest neighbor and the planet most like it in terms of size and geology. It is a rugged world of volcanic mountains, canyons, high plateaus, and sweeping volcanic planes crisscrossed by riverlike magma channels. Much of the surface is basaltic rock. The climate of Venus is one of the most inhospitable in the solar system. Perhaps only the hideous radiation of the inner Jovian moons presents a more difficult challenge to transhuman colonization. The Venusian atmosphere is a superheated maelstrom of carbon dioxide and sulfuric acid, with an atmospheric pressure at its surface equivalent to that five kilometers below the surface of Earth's oceans. Venus also lacks more than trace amounts of hydrogen, meaning that water must be imported in the form of iceteroids from the outer system. 

Nonetheless, transhumanity has come to Venus, and with it, debate over how to make use of the planet. Venus has no permanently inhabited surface settlements other than a few equipment and supply caches used by planetside researchers. Despite difficulties, transhumanity has found survival strategies that work here. The most surprising of these are the aerostats, lighter-than-carbon dioxide habitats that float in the thick Venusian atmosphere. Aside from a few independents or ones loyal to the Planetary Consortium, these aerostats are the base of the new Morningstar Constellation power bloc. Notable for their research labs, nanofab design houses, software studios, and luxury resorts, the Constellation's aerostats are increasing at odds with Planetary Consortium and Lunar-Lagrange Alliance interests. 

On some aerostats, areas populated only by indentured synthmorphs are open to the Venusian atmosphere. Some 500,000 transhumans live in aerostat habitats and another 10,000 on the surface. Roughly 5,000,000 transhumans live in habitats orbiting Venus. 

Though the Planetary Consortium is considering the launch of a Venusian terraforming project, this plan is actively opposed by the Morningstar Constellation. The Constellation's aerostats see the terraforming proposals—which include massive cometary bombardment or building a planet-sized sun shade to cool the atmosphere—as not only unworkable but disruptive to their lives and profits. 

Venus is a fascinating place for climatologists, geologists, and other planetary scientists. The discovery of Venusian protobacteria created a new branch of life sciences overnight, though so far the practical applications for organisms with such radically different metabolisms from terrestrial life have been limited. 

\subsubsection{Gerlach}
\label{sec:gerlach} 

Gerlach is an O'Neill cylinder supporting about 100,000 transhumans. Generally recognized as the research powerhouse of Venus, Gerlach is also one of the strangest places in the inner system. The inhabitants have strong ties to the argonauts and sympathies for the outer system autonomists and are strong proponents of morphological freedom, cognitive experimentation, and open innovation. Gerlach's main activities are planetside research and exploration, hostile environment morph design, and aerostat construction. 

\subsubsection{Octavia}
\label{sec:octavia} 

Octavia is the most successful aerostat habitat to date and the political center of the Morningstar Constellation. It maintains an altitude of roughly 55 kilometers above the northern highlands of Ishtar Terra. Octavia resembles an immense, mushroom-shaped skyscraper, 450 meters tall, ringed at its center by four radial outrigger spars, each ending in a stabilizing gas envelope filled with helium. The cap of the mushroom is a hard, translucent dome that provides an open, park like space while also serving as the main gas envelope (oxygen, which is much lighter than the CO2 making up most of Venus's atmosphere, is the main source of buoyancy). The habitat is fluted from top to bottom, going from a diameter of almost 300 meters at the base of the dome, to 15 meters wide at the very bottom. A huge counterweight tethered to the bottom of the structure prevents the habitat from capsizing during storms. Atmospheric craft and shuttles from orbit may land at flight decks near the base of the outriggers. 500,000 people live aboard Octavia. 

\begin{quotation}
\textbf{Venusian rumors} 

 [Incoming Message. Source: Anonymous] 

 [Public Key Decryption Complete]  We need you to investigate some odd rumors circulating about activity on the Venusian surface. According to reports, an Omnicor research team went missing about a week ago. Unlike many Venus surface teams, these weren't teleoperated bots but actual synthmorph-sleeved researchers operating away from the safety of an aerostat's tether—which is suspicious behavior itself. Search parties have turned up no sign of the missing morphs, but scuttlebutt says they ran into signs of recent TITAN activity that have them freaked out. I haven't found any evidence to back this up, yet—it could just be some misinformation to keep people from digging around part of the surface. I've heard that some security corps have some quantum data caches buried away down there. Looking into this may require getting a hold of some heat and pressure resistant synthetic morphs.
\end{quotation} 

\subsubsection{Aphrodite Prime}
\label{sec:aphrodite-prime} 

One of 20 smaller aerostats, Aphrodite Prime hovers 54 kilometers above Aphrodite Terra. It is a center for Venusian tourism; fully a quarter of this aerostat is a resort for wealthy off-world visitors. Aphrodite Prime is also the primary research station for the design and creation of life forms adapted to live in the Venusian clouds. This aerostat has a population of 300,000 and features closed-environment test aviaries populated with clouds of air plankton and schools of recentlydesigned flying squid and balloon fish. 

\subsection{Earth}
\label{sec:earth} 

Ecologically devastated and infested by the weird spawn of the TITANs, transhumanity's homeworld doesn't get many visitors. Earth's once-populous urban regions are massive sprawls ruined by war and heavy weather, infested with dangerous artificial life and the occasional survivalist gang. Elsewhere, irradiated blast zones and desolate wasteland prevail. Due to harsh climatic conditions, the wilderness has been slow to reassert itself, and vast swaths of dead forest or burned grassland are common sights. 

Even from orbit, Earth shows deep scars. Breaks in the sooty cloud cover created by orbital bombardment during the Fall reveal continents ravaged by coastal flooding, desertification, and radical temperature shifts. The only known detonation of an antimatter bomb within a planetary atmosphere, centered on what was the Chicago-waukee Metroplex in North America, left a crater over 200 kilometers wide wherein most matter was instantly vaporized. Craters left by mass driver bombardment dot the surface as well. Mass die-offs of lynchpin species like honey bees and krill destroyed entire ecosystems, leaving vast swathes of barren land and sea inhabited by only the most adaptable species. Most of Europe is sub-artic; much of Africa and North America, desert. Ironically, transhumanity's deployment of nuclear weapons against TITAN surface installations arrested the effects of global warming by creating a nuclear winter. Nuclear attacks against Earth have ceased, but the Lunar mass drivers still occasionally hurl captured asteroids at suspected surface works created by remaining TITAN war machines. In any case, the damage from humanity's warming of the globe was already done. The patterns of life on Earth, and the very face of the planet, have been irrevocably rewritten. 

Earth once had multiple space elevators in operation, but with exception of the Kilimanjaro beanstalk, the others were destroyed during the Fall, wrapping around the planet as they crashed to Earth, leaving swathes of destruction. 

\subsubsection{Population}
\label{sec:population} 

Earth's population is a matter of speculation. The reclaimers and Lunar authorities, both of whom spend a great deal of effort monitoring Earth, agree that surface energy emissions suggest a population of about one million once-humans living as servitors to the TITANs, although these numbers assume patterns of energy usage similar to those of pre-Fall humanity. 

Though the Planetary Consortium claims that no survivors remain on Earth, reclaimer estimates guess that between 20,000 and 100,000 free humans remain. These numbers are hard to formulate, given the limited number of remote areas where humans could remain undetected while obtaining enough food to subsist. Some areas likely to conceal sizable remnant populations include the highlands of Papua-New Guinea, the Ozark Mountains of North America, and the jungle uplands of Vietnam and Laos, though it is also possible that certain underground and undersea settlements survive. Attempts to make contact with survivors have universally ended in disaster. 

During the Fall, thousands of people unable to escape Earth resorted to having themselves backed up and transmitted off-planet. Many of these—along with some who had no backups—also put there bodies in cryogenic storage, hoping to wait out the Fall for rescue. Some reclaimers have speculated that dozens of these cryogenic facilities may still be functional. 

\subsubsection{Habitats}
\label{sec:habitats-2} 

Earth had a mature orbital industry sector and a considerable population in orbit at the time of the Fall, with over a billion people living full-time in space. Earth orbit was one of the fiercest battlegrounds of the Fall, however, and hundreds of habitats and other installations were destroyed or rendered unusable. As such, Earth orbit and the Lagrange points are littered with the detritus of pre-Fall humanity. Derelict habitats can mean tidy profits for intrepid scavengers, but many are also infested with TITAN spawn and hostile nanoswarms, making them incredibly dangerous. 

To make matters worse, someone or something has unleashed a large number of autonomous killsats in Earth orbit to interdict would-be visitors. Some of these are repurposed pre-Fall military hardware, while others are newer construction. So far, no one claims responsibility for them. The Planetary Consortium is suspected, as they support and sometimes enforce a quarantine of the planet, but the possibility exists that the killsats may be TITAN relics or the efforts of another agency. 

Despite the chaos of Earth orbit, numerous habitats remain active here, many of them participants in either the Planetary Consortium of Lunar-Lagrange Alliance. Dozens of formerly derelict habitats have also become home to squatters, some of them with criminal intent, others just looking to escape the squalor of life in the overcrowded Lunar-Lagrange habitats, even if it means taking a risk. 

\subsubsection{Fresh Kills}
\label{sec:fresh-kills} 

Essentially an armed-to-the-incisors scum barge, Fresh Kills is a salvage base near the edge of Earth's L5 point. The base is built around a huge central docking spindle with moorings for small craft and habitat modules in the center, and massive weapons batteries at either tip. Scavengers can moor their own craft or, at considerable expense, egocast in, resleeve at the facility, and hire shuttles for excursions. The gun batteries are articulated such that any craft showing signs of trouble can be hastily jettisoned and destroyed. 2,000 transhumans live on Fresh Kills, although the population is transient and fluctuates a good deal. 

\subsubsection{Paradise}
\label{sec:paradise} 

Situated in a halo orbit at the Earth-Sun L1 point, Paradise was an exclusive spa and resort station for the ultra-rich before the Fall. In the wake of the Fall, Paradise fell on hard times, swarmed as it was with refugees and no longer an ideal vacation spot. Recently, however, Paradise fell back in favor with the inner system glitterati, who undertook measures to expel many of the lingering squatters and refurnish it as an elite social space. Recent rumors suggest the Consortium's Hypercorp Council has used Paradise for important face-to-face meetings. 

\subsubsection{Vo Nguyen}
\label{sec:vo-nguyen} 

The Reclaimers maintain this station in high geostationary orbit, monitoring Earth and making plans for potential geo-engineering efforts. Vo Nguyen is a small O'Neill cylinder hidden in a dangerous cloud of space junk and protected by swarms of killsats, gun emplacements, and drones. It is occasionally used as a jumping off point for secret surface expeditions. 

\subsection{Luna}
\label{sec:luna} 

The first planetary body to host permanent human habitation, Earth's sole moon is home to the second largest population of transhumanity on a single planet and remains a lynchpin of culture and economic activity. Lunar history has been shaped dramatically by the Fall. Before the need to evacuate Earth arose, it was expected that the Moon would remain largely an automated mining concern, never attaining a population of more than a few million. Luna was never seen as an economically viable location for colonization, the focus instead falling on Mars and the outer system. 

When the Fall came, every polity that couldn't hope for a shot at Mars or elsewhere set its sites on Luna. The Indians were the only great power that had invested heavily in Luna. The other three major settlements, Erato, Nectar, and Shackle, were multinational and hypercorp concerns with no strong national affiliations. These three cities swelled overnight into polyglot refugee camps, while the Indian settlement, New Mumbai, was nuked black by the corps when it became apparent that a TITAN infection had taken hold there. 

Bereft of nationhood, Lunars developed their own resourceful, tough-minded culture which has emerged as a counterbalance to the radicalism of the outer system and the excesses of Mars. 

Transportation on Luna is largely by suborbital rocket, although trans-sonic bullet trains also operate along shorter routes. The major space port is at Nectar. There is also a skyhook—a massive orbiting satellite spaceport that drags a massive tether, which acts as a space elevator along a track running across the Lunar surface south of the equator. As a result, many smaller cities lie along the skyhook track. 

\subsubsection{Fashion/Design}
\label{sec:fashiondesign} 

Nectar is one of the three fashion/design capitals of the system (along with Noctis on Mars and Extropia). The Lunar design houses have two major advantages: an inventive population and a low planetary gravity that makes it easier to design for the low gravities that prevail in much of the system. Some habitats elsewhere in the system even choose a rotational speed that simulates Lunar gravity in order to get the greatest benefit from Lunar designs. 

\begin{quotation}
\textbf{Tilion's Jupiter brain} 

 [Incoming Message. Source: Anonymous] 

 [Public Key Decryption Complete] 

 Our investigation into codename: TILION's Lunar research activities has confirmed our suspicions. The hypercorp is engaged in experiments to convert confined spherical masses in the Lunar interior into testbed micro-Jupiter brains. The silicate-rich Lunar crust makes the locations they have chosen ideal for the project. Though we have not verified it, we believe that TILION not only followed the trail of TITAN research into this area, but is in fact in possession of a small cache of TITAN-made computronium. There is no saying what the TITANS may have been using this cache for, what it may store, or what may occur if TILION completes the project and brings the micro-Jupiter brain online. Fortunately, time seems to be on our side, and we have several weeks if not months before any significant part of the project is activated. We will continue to infiltrate and learn more, but we strongly suggest an erasure squad be moved into position and placed on standby.
\end{quotation} 

\subsubsection{Helium-3 mining}
\label{sec:helium-3-mining} 

Although it's not the richest place to mine He-3, Luna has such good infrastructure for extraction and distribution that it more than makes up for the fact that Luna is very poor in hydrogen for more conventional forms of fusion. Unlike the vast reserves of the gas giants, however, the amount of readily extractable He-3 in the Lunar regolith is finite. Some of the richer deposits are already tapped out, and concerned Lunars consider their world's future after these deposits are exhausted a major issue. 

\subsubsection{Finance}
\label{sec:finance} 

The Lunar banks are the oldest (and thus richest) in the system, though hypercorps like Solaris are close on their heels. Interestingly, the rise of the reputation economy in the outer system has not presented as much of a problem for these banks as one might have expected. Lunar banks got hip to the reputation game long before the Martian financial institutions and moved in to capitalize on it immediately. By the time Martian banks knew what was going on, Lunar financial institutions had struck deals with the Extropians and dominated all of the points of exchange where favors could be bartered for cold, hard cash between inner system corp types and outer system anarchists. The same genius fueling Lunar design created a complex barter to cash network that almost everyone uses. While some autonomists find it infuriating that they have to deal with a monolithic banking system to get by in the inner system, others are simply happy to deal with the Lunars instead of the Martians for this service. 

\subsubsection{Erato (Eratosthenes)}
\label{sec:erato-eratosthenes} 

Erato (population 5 million) is a major mining center consisting of a series of heavily shielded surface domes and a vast underground city. Erato is centered around the Eratosthenes crater on the southern edge of the Mare Imbrium (Sea of Showers), in the northern hemisphere of the Terra-facing side of Luna. Erato has access to both the rich titanium deposits of the Mare Imbrium and fields of Helium 3-abundant regolith. 

Erato is one of the oldest mining settlements on Luna and one of the first to become commercially viable. As such, many of the Lunar banks are centered around this city. The vaulted heights of the Great Cavern of Erato, originally excavated by a Sino-European conglomerate, reach a height of 1.5 kilometers at the apex, leaving room for a teeming city of gardens and towers grown from Lunar silicates and industrious nanites, lit from above by sunlight entering via great mirrored vents. 

\subsubsection{Nectar (Nectaris)}
\label{sec:nectar-nectaris} 

Nectar (population 9 million) lies about 100 kilometers due east of Theophilus crater on the Mare Nectaris (Sea of Nectar) in Luna's southern hemisphere. Nectar is a design powerhouse, home to the great Lunar design houses that set fashion and design trends for much of the solar system. Due to its location relatively close to the Lunar equator, Nectar also hosts Luna's primary long-haul space port and is on the pickup path for the Lunar sky hook. 

\subsubsection{New Mumbai containment zone}
\label{sec:new-mumb-cont} 

The incineration of the New Mumbai colony with nuclear weapons during the Fall to prevent the spread of TITAN infection left a scorch mark roughly 100 kilometers in diameter on the face of Luna that is still visible from high orbit. The colony was a heavily automated Helium-3 mining station, located in the midst of rich Helium-3 fields on the edge of the Mare Moscoviens. It remains a heavily-patrolled quarantine zone to this day. 

\subsubsection{Shackle (Shackleton-New Varanasi)}
\label{sec:shackle-shackl-new} 

Shackle (population 6 million), built in and around the south polar Shackleton crater, is centered around one of two major water extraction operations on Luna. New Varanasi, the city of temples, is the most impressive section of the city. Shackle was the other major site of old Indian influence on Luna, and with the destruction of New Mumbai holds special importance to descendants of the Indian diaspora. New Varanasi is a monumental artificial cavern complex with an intricate canal system fed by melted ice from the polar caps above. As a source of lifegiving water, it now holds the same importance to the Hindu faith once ascribed to the River Ganges on old Terra. Survivors of other Indian religions, such as the Jains and Sikhs, have also made their temples here. This makes Shackle a major pilgrimage site; tourism is the major industry after water extraction. A small herd of Indian elephants is a major attraction, and the elephant god Ganesha, Remover of Obstacles, is extremely popular on Luna, even with non-Hindus. 

\subsection{Mars}
\label{sec:mars} 

Earth was the cradle of transhuman civilization, but Mars, with a population of 200 million, is now its heartland. When humanity began its spaceward diaspora, Luna was its first stop. Yet while Luna boasts a sizable population, Mars was the first world humans settled where they could thrive entirely on locally available resources. During the first few decades, the early Martian settlers dwelt in tin can hab units, extracting methane from the local atmosphere for rocket fuel and water from the Martian permafrost, farming in inflatable greenhouses, and eventually manufacturing enough greenhouse gases to warm the planetary climate to the point where transhumans could walk the Martian surface unprotected, save for oxygen respirators. 

The second phase of the great project of terraforming Mars—husbanding plant life and microbes engineered to rapidly replace atmospheric carbon dioxide with oxygen—was already underway at the time of the Fall. A belt of orbital mirrors helps to heat the planet by focusing the sun's rays. The spread of plant life is a long-term project that will take several centuries to produce a fully breathable atmosphere, but the nigh-immortal transhumans of Mars are prepared to be patient. A new homeworld is worth the wait. Research into new plants and microorganisms capable of releasing oxygen and nitrogen into the Martian atmosphere at an ever-accelerating pace is a major focus of economic activity. 

In the meantime, the red planet is a place of startling contrasts, from the stark beauty of its mountain ranges and high desert, to the slowly-greening bottomlands of the equatorial Valles Marineris canyon system. In these bottomlands, oxygen levels are slowly rising, and liquid water can now be found in canals that had already been dry for millions of years when transhumanity's ancestors came down from the trees. Mars is a popular destination for travelers from around the system. Many Martians accrue wealth by operating lavish hotels, offering tours of historical sites, and leading wilderness expeditions to the rugged highlands and vast deserts of the untamed Martian frontier. 

Mars now sports five vast, domed cities, mostly in the equatorial regions, along with numerous smaller settlements. Settlements are connected by surface roads, a network of near-sonic maglev trains, and air/spaceports from which suborbitals, airships, and near space rockets fly on regular schedules. Thanks to the abundance of methane fuel and the one-third Earth gravity, transhumans on Mars have finally got their flying cars as well, and all settlements have well-delineated rights of way for these vehicles. Meanwhile, in the wild uplands, planetologists and terraforming engineers dwell in small villages, living the simple life in ruster morphs while seeing to the continued development of the Martian climate and atmosphere. 

As a partially terraformed planet with vast tracts of unused land, Mars is one of the few places that can offer new sleeves to infomorph refugees. Martian brokerage houses do a brisk business in the purchase and resale of infomorph contract labor, with agreements (for some) leading to eventual sleeving. This has led to a sizable Martian underclass, however, organized as a growing resistance movement under the Barsoomian banner (though the hyperelite socialites disparagingly call them “rednecks”). 

\subsubsection{Regions}
\label{sec:regions} 

Mars is broadly divided between the lowlands of the north and the highlands of the south, which in many places are separated by dramatic cliffs up to two kilometers high. Mars has seasons just as Earth, and both north and south poles have permanent ice caps that persist despite transhumanity's success in warming the planet. Both regions present obstacles to terraforming. The northern plains are open and windswept, while the rugged southern uplands remain a difficult terrain for life to gain a foothold. Even so, tough Earth species like cacti and succulents are able to grow in the best spots. 

\paragraph{Ma'adim Vallis:}
\label{sec:maadim-vallis} This deep canyon system on Mars holds one of the Planetary Consortium's most treasured possessions: the Martian Gate. This Pandora Gate was originally discovered by nomadic Barsoomians, then violently wrested from their hands by hypercorp troops—an event that still rankles the rednecks. As different hypercorps themselves nearly came to blows, the Hypercorp Council was forced to step in and offer a resolution that all could agree to. A new hypercorp was founded—Pathfinder—which would control exploration and exploitation of the gate and resources beyond, with special privileges and rights given to Planetary Consortium members. The Martian Gate is now a staging point for numerous exoplanet colonies, though some fear the prospect of keeping a presumed TITAN artifact operational on transhumanity's most populous planet. 

\paragraph{Olympus Mons:}
\label{sec:olympus-mons} Mars' most notable landmark is the mighty shield volcano Olympus Mons, on which the first—and still principle—Martian space elevator was constructed. Similar in shape and origin to Earth's Hawaiian Islands, but now dormant, Olympus Mons is one of the highest mountains in the solar system, rising 27 kilometers. 

Olympus, the settlement in the volcano's caldera around the base of the space elevator, was once the chief city of Mars, but waned in popularity as a place to live when terraforming made other regions more attractive. A maglev train from Olympus takes a little over three hours to reach Noctis; air travel is even quicker. Despite the waning of the city, the space elevator still sees heavy use. 

\paragraph{Valles Marineris:}
\label{sec:valles-marineris} Most of transhumanity's terraforming efforts center around the winding Valles Marineris canyonlands, which twist and turn over 4,000 kilometers east-to-west along the Martian equator. In these relatively warm bottomlands, liquid water is becoming abundant and the land is green with hardy Terran plant species like crab grass, dandelions, and towering Douglas firs (which botanists estimate may reach heights of 180 meters in the low Martian gravity). 75\% of the transhuman population of Mars lives in this region, giving it the highest density of transhuman habitation in the solar system. 

\paragraph{The Zone:}
\label{sec:zone} Officially labeled the TITAN Quarantine Zone, the TQZ is a large area stretching from the smooth plains of Amazonis Planitia (between the Tharsis and Elysium volcanic areas) and southeast to Arsia Mons (just west of Noctis). This zone is known to be crawling with leftover TITAN machinery: warbots, nanoswarms, and other dangerous things. Several devastated habitats lie in this region, including the former Islamic stronghold of Qurain. Few dare venture here, though some rumors suggest that Barsoomian smugglers make use of the Arsia Mons caves and even scavenge for TITAN tech, despite the risks. Planetary Consortium drones keep a vigilant eye on the Zone's borders, though for unknown reasons the TITAN relics rarely stray beyond its bounds. 

\subsubsection{Ashoka}
\label{sec:ashoka} 

Ashoka is located in a crater in the Ares Vallis region about 3,000 kilometers northeast of Valles-New Shanghai, not far from the landing sites of the early Viking and Pathfinder probes. The town is a popular spa and spiritual retreat for Martians wanting to revisit their pioneer roots. It is also an active terraforming station and a major point of contact between the seminomadic Barsoomian culture of the high desert and the settled Martians of the equatorial canyonlands. 10,000 scientists, historians, terraforming workers, and spiritual gurus live in the town and surrounding area. A major attraction is a museum housing the Pathfinder lander and the Sojourner rover (which was still operational when humans landed and discovered it circling endlessly in a crater). The Viking lander is in another museum a short monorail ride from town. In a move that infuriated historical purists, all three machines were given modern hardware upgrades when discovered and now house AIs who act as historians of early Mars exploration. Sojourner is particularly friendly and sometimes leads lucky groups on walking tours of early landing sites. 

\subsubsection{Elysium}
\label{sec:elysium} 

Located in the Elysium and Hyblaeus Chasma in the north of the Hesperia region in Mars's eastern hemisphere, Elysium is the entertainment capital of the system and the largest Martian city outside of the canyonlands of the equator. It is also the most physically remote of the large Martian cities, though transhumanity's advanced transportation technology (suborbital flights and rocket flight from habitats above) make this remoteness a trivial quality. 

Elysium and Hyblaeus Chasma together make up a 250-kilometer long canyon system in the shadow of Elysium Mons, a 14-kilometer mountain located about 200 kilometers northeast of the city. In between is the Zephyrus Fossae, an undulating, windswept lava plain. The city was the vision of one person, Zevi Oaxaca-Maartens, an eccentric entertainment magnate who was intrigued by the close proximity of the eminently terraformable Chasma to the unspoiled Hesperian terrain. 

The city is only 30 years old but already boasts a population of 9 million transhumans. Elysium is mostly built into the canyon walls of the Chasma, sprawling over a 75-kilometer stretch, all of which has been domed over. Unlike the big domed metroplexes of the south, Elysium takes advantage of the canyon walls, which are close enough together that rather than building free standing domes, the builders have simply built great enclosing arches to completely cover the canyon. These expand northward year by year as the city grows. From low orbit, it looks like a great, glistening serpent. 

The Martian city of Elysium is the spiritual successor to old Terra's Los Angeles as the entertainment capital of the solar system. Glamorous stars and blood drinking producers, coupled with a healthy dose of outrageous (if often vapid) transhuman creativity have made Mars an unrivaled media powerhouse. Elysium may boast more exalt and sylph morphs per capita than any other transhuman city. 

Image is everything here, and to visitors it may seem as if everyone in this city is either blindingly beautiful or calculatedly ugly. The most successful performers and entertainment tycoons live lives of glittering privilege that would make the richest gerontocrat in New Shanghai mildly envious. Everyone else, from up-and-coming game producers to the virtual ero performers, has to hustle constantly. 

\subsubsection{Noctis-Qianjiao}
\label{sec:noctis-qianjiao} 

With a population of 13 million, Noctis-Qianjiao is the major metroplex in the west of the Valles Marineris region, an area known as Noctis Labyrinthus. Although not as hospitable as the Eos region in which Valles-New Shanghai lies, Noctis Labyrinthus is considered prime real estate for its gorgeous scenery and well-developed river systems. The metroplex boasts two major domes: Qianjiao, on the northern bank of the River Noctis, and Noctis City (normally just called “Noctis”) to the south. Connecting the two domes and spanning the river is a sprawling network of lesser domes and souks, although these have been pushed north and south over the years as the planet warms and the river grows wider. 

Noctis-Qianjiao is the center of the Martian design and fashion industries, which in the abundant Martian economy arguably makes the city as important as much larger Valles-New Shanghai. This settlement's proximity to the Zone sometimes alarms visitors, but there have been no public incidents to cause concern so far. 

\begin{quotation}
\textbf{Martian manhunt}  [Incoming Message. Source: Anonymous] 

 [Public Key Decryption Complete] 

 Those deaths you asked me to look into? It’s looking worse than we feared. 

 I took a buggy car out to the Zim settlement. It’s just a collection of tin can modules, supporting a small terraforming ecostation and facilities for nomadic rednecks. On average, it’s home to 150, if you count the 20 or so pleasure pod AIs that serve as local “entertainment.” 

 A week ago, a nomad known as Hassan Naceri rolled in. He’s a regular, word is that he runs a lone courier service for the Barsoomians. On this recent visit, though, his behavior was off. He was nervous and agitated. He told one drinking buddy that he’d been forced to hide out in the Zone for a few days and the experience had put him on edge. 

 Turns out Naceri had run off with a ruster morph without working out his full indenture to Fa Jing a few years back. The ego hunter showed up in Zim and Naceri lost it. He killed the ego hunter and everyone else in the room. 

 We found a spime’s sensor records that show Naceri transforming. He also killed half a dozen people simply by looking at them. 

 That’s right. This bugger’s infected. 

 The Martian Rangers are hot on his trail, but they don’t know what they’re dealing with. So we’re off to try and catch him—it—first. Wish us luck.
\end{quotation} 

\subsubsection{Olympus}
\label{sec:olympus} 

Olympus, with a population of 1 million living in a space designed to accommodate 6 million, is something of a ghost town. The former principal city, built in the caldera of Olympus Mons around the space elevator, is now fallen into disuse. As the temperatures rose and the climate improved in the Valles Marineris canyonlands, most of the population left the windswept caldera for more hospitable surroundings. Olympus is not and never was a large domed city, consisting instead of a souk-like network of minor domes and antiquated tin can hab modules. 

Low atmospheric pressure and bone-freezing temperatures at the city's altitude of 27 kilometers mean that most transhumans venturing outside the souks and hab modules still need the equivalent of light vacsuits to survive. Martian Alpiners, a rare morph found in few other places, are not uncommon here due to the harsh conditions. 

The city center is well-maintained and carefully overseen by the Olympus Infrastructure Authority, a minor hypercorp that operates the space elevator. The outskirts are economically depressed and sometimes dangerous, mostly deserted and populated by squatters, indentured downloads on the run, and other people who really want to be left alone. Occasional outbreaks of dangerously mutated artificial life are one of the few reasons for which the Authority bothers to intervene in the outskirts. Otherwise, the old tin can habs and their strange inhabitants are left to decay. 

\subsubsection{Progress (Deimos)}
\label{sec:progress-deimos} 

Progress is one of the largest Cole bubbles in the Solar System. With 8.5 million residents, it is second in population only to Extropia in the belt. Progress was created when Fa Jing evicted all of the former residents from the Martian satellite of Deimos, excavated the inside of the moonlet, and used a massive solar array to convert it into a bubbleworld. From an engineering standpoint, Progress is something of an embarassment. The habitat was originally meant to exceed Extropia in size considerably, but difficulties with heating and spinning Deimos forced Fa Jing to abandon their efforts early or risk the moonlet breaking apart. 

Progress is nonetheless an impressive habitat, home to hypercorp glitterati and an outpost for a host of major political and economic concerns. Its sister moon, Phobos, remains a warren-like tunnel habitat due to the presence of multiple legal interests unable to agree upon the disposal of the satellite. 

\subsubsection{Valles-New Shanghai}
\label{sec:valles-new-shanghai} 

The principle city of Mars, Valles-New Shanghai is transhumanity's largest planetary metroplex, with 37 million inhabitants. Valles-New Shanghai lies in the heavily terraformed Eos region in the east of the Valles Marineris canyon system. The metroplex is comprised of five major domes connected by a network of Martian souks. The souks are a unique architectural feature of large Martian cities, consisting of covered thoroughfares and galleries lined with bazaars, eateries, and squats. It is said one can find anything if one spends enough time walking the souks. 

The domes themselves are tamer, with artificial waterways (many of which now connect to the tenuous rivers etching the surface of the Eosian bottomland), grand architecture, residential mini-arcologies, entertainment complexes, and hypercorp conference centers. The most impressive by far is the Bund, the larger and older of two domes making up the city of New Shanghai proper. New Shanghai is roughly bisected by the twisting Ares, an artifical river that helps regulate the dome's climate. Near its center is an almost brick-for-brick duplicate of the original Bund from the destroyed Earth city of Shanghai. 

The other four domes are Little Shanghai (a newer, smaller dome adjacent to the Bund), Valles Center (a business and financial center that rivals the Lunar banks of Erato and Nectar), New Pittsburgh (also called the Burgh, a hub of research and planet-side industry), and Nytrondheim (housing major entertainment districts). 

Valles-New Shanghai is transhumanity's wealthiest population center, a hotbed of art and culture, and one of the system's great centers of hypercorp activity. The populace includes an extremely high percentage of gerontocrats, but their stifling influence on culture, economic mobility, and the legal system is only one force among many in a city of 37 million people. The city has expanded so much to accommodate its exploding population since the Fall that new construction is a constant. Crime and corruption are widespread, though the worst of it is contained to Little Shanghai. Valles is a place where dreams are made and broken every day, if not every hour. 

\subsection{Martian Trojans}
\label{sec:martian-trojans} 

Not to be confused with the much larger Jovian Trojans, the Martian Trojans are a small group of mostly rocky asteroids trailing and preceding Mars at its L4 and L5 points. 

\subsubsection{Qing Long (Azurez Dragon)}
\label{sec:qing-long-azurez} 

Qing Long, with a population of 2 million, is the largest O'Neill habitat in the system. It is situated among the Trojans at the Martian L5 point. Qing Long has its roots in the Chinese Mars colonization effort. Despite its exceptional size, it is one of the oldest habitats of its type, having been built almost entirely from metal-rich asteroids mined near its present location. 

Qing Long is a major underworld haven. The habitat's administration is beholden to several criminal organizations who normally refrain from killing one another. The habitat nominally obeys some hypercorp principles, such as limited access to cornucopia machines, forking, and AGIs. However, thriving grey and black markets enable people with the right connections to acquire just about anything here. 

\subsection{Asteroid belt}
\label{sec:asteroid-belt} 

Spread out over a massive region between the orbits of Mars and Jupiter, the belt contains a few hundred asteroids greater than 100 kilometers in diameter, over a thousand objects greater than 30 kilometers in size, and countless smaller ones. Despite this, the total mass of asteroids in the belt is only a fraction of one of the inner planets, meaning that asteroids are spread out over great distances. A spacecraft flying through the belt is highly unlikely to encounter an asteroid unless it deliberately navigates toward it. 

\subsubsection{Resources and economics}
\label{sec:asteroid-resources-economics} 

The rich, easily accessible mineral deposits in the Belt were a major link in transhumanity's first steps toward the outer system. Automated mining and high-impulse ion boosters enabled outer system colonists to move metal-rich Main Belt asteroids into the orbits of Jupiter, Saturn, and beyond, where metallic asteroids are much scarcer. This activity continues to this day as transhumanity pushes further out into the system. 

\subsubsection{Habitats}
\label{sec:habitats-3} 

Hundreds of small habitats, mostly involved in prospecting activities, dot the belt. Distant from Earth, settlements in the belt were largely spared the devastation of the Fall. Both hypercorp and autonomous outposts flourish here. Derelict habitats abandoned when nearby asteroids were boosted into the outer system or depleted are common here as well, although some of these are now occupied by residents who are best left to their solitude. 

\subsubsection{Ceres}
\label{sec:ceres} 

One of the system's three dwarf planets (along with Pluto and Eris), Ceres is almost 1,000 kilometers in diameter and hosts a population of almost a million. Unlike most Main Belt asteroids, Ceres has an icy crust with a layer of liquid water beneath it, like a miniature version of Jupiter's moon, Europa. With its abundant water, Ceres has a major role in resupplying other stations in the belt. Similar to Extropia, Ceres operates largely along anarcho-capitalist lines. However, the Hidden Concern, a cartel run entirely by uplifted octopi, holds sway in the sub-crustal sea and maintains a stranglehold, as it were, on water extraction operations. Cerean octopoid morphs are specially adapted to survive in the ammonia-rich waters of the Hidden Sea. 

\subsubsection{Extropia (44 Nysa)}
\label{sec:extropia-44-nysa} 

This massive beehive habitat is a major crossroads and anarcho-capitalist/mutualist marketplace. Extropia is a neutral free city whose infrastructure and social fabric is maintained by a loose association of anarcho-syndicalist affinity groups. Extropia's neutrality hinges on strategic alliances between key local figures, their networks, and an unusual array of outside interests that include the Lunar banks, technolibertarian factions, and outer system colonies dependent upon raw materials exported from the belt. The hypercorps use Extropia as a tax shelter and a haven from which to do illicit business. There are no laws or government as such; visitors are advised to register with an insurance and security provider. Named after one of the first transhumanist movements, Extropia is considered a utopia for transhumans looking for body modifications. AGIs and forking are accepted and allowed here. The transhuman population is nearly ten million. 

\subsubsection{Nova York (Metis)}
\label{sec:nova-york-metis} 

One of the more unusual near-weightless habitats is Nova York, the main city on Metis, a large nickel-iron and silicate asteroid located in the main belt. Nova York, the third largest habitat in the main belt, is a thriving metropolis of 500,000, with the main portion of the city located in a spherical cavern approximately four kilometers in diameter, the top of which is two hundred meters beneath the asteroid's surface. Lit during the day by a series of huge light tubes in the outer walls, at night the lights of the buildings cause the surface of this sphere to resemble an enormous geode. The habitat's basic design consists of many thousands of exceptionally tall and fragile-looking buildings that extend between one hundred and fifteen hundred meters above the surface, as well as a few buildings that stretch from one side of the cavern to the other. In Metis's minute gravity of 1/140th of a g, up and down have little meaning, and even relatively fragile buildings are in no danger of falling down. The vast majority of the buildings, including ones more than one kilometer tall, are made from thin plastic panels over a durable supporting framework. These buildings jut out at all angles from the sphere. 

Many inhabitants of Nova York move from one building to another by jumping, and a single leap can carry someone many hundreds of meters. Residents do not worry about falling—the combination of air resistance and exceedingly low gravity means that even someone falling from the top of the cavern to the bottom is in no danger of injury. In this environment, the only real meaning of up and down is that down is where you look for objects to come to rest (as long as an air current does not pick them up and blow them around). 

\subsection{Jupiter}
\label{sec:jupiter} 

Large enough that it could almost have formed the nucleus of a protostar in its own right, Jupiter's massive size makes the Jovian System one of the most challenging places in the system to colonize. Jupiter's powerful magnetic field means that its inner moons— and the outer ones, when their orbits pass through its immense magnetotail—are bombarded with enough ionizing radiation to kill transhumans not protected by the heaviest of shielding within a matter of hours. There are sixty-three moons and moonlets in the Jovian system, but only the well-explored, populous, regular moons are described here. 

\subsubsection{Resources and economy}
\label{sec:jupiter-resources-economy} 

Jupiter's powerful gravity well is a major hindrance to gas mining in the planet's atmosphere, as even craft that do not succumb to the violent, centuries-long atmospheric storms can achieve escape velocity with only the most powerful propulsion systems. Given the need for heavy shielding on such craft, gas mining on Jupiter is not nearly as efficient as on Saturn. Jupiter has a tenuous ring system, much less dense than Saturn's, which extends out for 20,000 kilometers around the planet, encompassing the orbits of its two closest moonlets. 

However, Jupiter's gravity is also a valuable resource. Craft bound for Saturn and beyond can slingshot themselves outward by circling the planet to pick up velocity, cutting months or years off their trips. The heavily militarized Jovian Republic levies tolls against all spacecraft using Jupiter's gravity to pick up velocity, including asteroids under propulsion. This protection money is the Junta's primary source of revenue. Planetary Consortium ships generally accept the payment as part of operating expenses. Other factions are not so cooperative, and the Junta regularly seizes or destroys blockade runners. 

\subsubsection{Habitats and moonlets}
\label{sec:habitats-moonlets} 

Most of Jupiter's moons are really captured asteroids, lacking the size and geological complexity of planetary bodies. All are occupied. Some were converted to habitats; others host only Junta military and mining outposts. The Jovian moonlets consist mostly of carbonaceous rock, poor in metal, with some of the larger moonlets having layers or even cores of ice. Beehive habitats and Reagan cylinders predominate in the Jovian system. Reagan cylinders (called “sarcophagus habs” by every other faction) are an inefficient variation on the O'Neill cylinder in which excavators hollow out an immense, cylindrical cavern in a rocky asteroid and then alter the asteroid's rotation with external thrusters to simulate gravity. 

Other habitat types are rare in Jovian orbit, especially within 2 million kilometers of the planet, where the radiation is strongest. For a bioconservative faction unwilling to adopt radiation-resistant morphs, the Junta is in a poor location. Shielding their populace beneath tons of rock is a necessity. Despite its military hegemony, the Junta can't control all of Jovian space, and there are things it can't do on its own—like exploring Europa. A number of unaligned habitats and surface settlements exist in the ring system and the orbits of the Galilean moons. 

The Jovian Republic has renamed Jupiter's moons after various neo-conservative heroes from Earth's history. From closest to most distant, the moonlets are Metis (Bush), Adrastea (Fairway), Amalthea (Solano), Thebe (McAllen), Leda (Chung), Himalia (Pinochet), Lysithea (Friedman), Elara (Buckley), Ananke (Nixon), Carme (Kissinger), Pasiphae (Schilling), and Sinope (Garcia). All are tiny, between 5 and 100 kilometers in diameter. 

\subsubsection{Amalthea (Solano)}
\label{sec:amalthea-solano} 

The largest of the moonlets, hollow Amalthea is probably the most livable sarcophagus habitat due to the large lake created from its icy core. Living on Solano carries some prestige among Junta citizens. Rumor has it that most of the residents are well-placed RAND think tank personnel, most of whom work on defense projects. A fusion-powered axial light tube illuminates the 30-kilometer diameter central cavern, whose landscape is patterned after the subdivisions and office parks of an early 21st-century suburb. All buildings have envirosealing so that the occasional bouts of environmental sepsis resulting from the poorly regulated interior ecosystem can be purged with toxin bombs. Less fortunate support personnel dwell in the beehive warrens crisscrossing the moonlet's crust between cavern and surface. Like most of Jupiter's moonlets, Amalthea's space crawls with patrol craft and killsats, making approach for unauthorized craft problematic at best. 1.5 million transhumans live on Solano. 

\subsubsection{Io}
\label{sec:io} 

Beneath Io's tenuous, patchy atmosphere of volcanic gases and neutral atomic dust lies a barren, grayish yellow, rocky surface coated with a thin frost of sulfur dioxide. Tidal heating caused by gravitational interaction with Jupiter makes Io the most volcanically active body in the system—so active that the meteor cratering found on every other planet and moon is completely absent on Io. Massive volcanic calderas, lakes of molten rock, and geysers of sulfur dot the surface, with eruptions and accompanying seismic activity lasting months or years. Volcanic zones on Io reach surface temperatures of up to 1,500 degrees Kelvin, hotter than any body in the system. 

For all that, transhumanity's worst peril on Io is radiation. Ejecta from geysers and volcanoes flow with Jupiter's magnetic field to form a titanic, toroidal flux tube that rotates with Io around the gas giant. Travelers to Io must either use the heaviest radiation shielding available or resleeve into synthetic morphs. 

Transhuman activity on Io centers around scientific research and harvesting the volatiles ejected by Io's geysers, particularly sulfur. Bases tend to be modular and mobile due to the ever-changing seismic activity. The Junta's most notorious prison, Maui Patera Rehabilitation Center, is dug into a (mostly) extinct caldera wall north of the equator. 

\subsection{Europe}
\label{sec:europe} 

Europa has no atmosphere and lies within the fearsome magnetosphere of Jupiter, and as such its surface is bombarded with enough radiation for an unshielded transhuman to receive an irrevocably fatal dosage within a few days—much faster when Europa's orbit passes through Jupiter's immense magnetotail. As a result, transhumans on Europa dwell beneath the icy crust, largely in the ocean below, adopting a variety of aquatic and amphibious morphs for survival. The only surface facilities are the heavily-shielded ice elevator heads at Conamara Chaos and several other points through which reactor mass and other crucial supplies can be delivered to the Europans below. 

Transhumanity is still exploring and imaging the Europan ocean floor, a task complicated by the hideous pressures at work in these waters, which are ten times as deep as the Earth's oceans. A further surprise awaiting transhumanity was the terrain. The geology of Europa suggested that beneath the ice would be fathomless depths of black water ending at a depth of nearly 500 kilometers in a relatively flat, featureless sea bed. Were Europa a lifeless ball of ice and rock, this would be the case, but over the estimated billion years since the rise of life on Europa, tiny lithoderms (analogs to Earth's coral) have built silicate reefs that rise to within a few hundred meters of the ice crust. It is on these biologically formed mountain tops, home to complex ecosystems, that the Europans have built their habitats. 

While based on water-carbon chemistry like life of Terran origin, life on Europa is completely autocthonic, having originated beneath an impenetrable ice sheet that cut off Europa's subsurface ocean completely from outside. This is in marked contrast to Terran life, which many biologists have theorized might be the result of galactic panspermia, the slow diffusion of microbes through the vacuum of space aboard comets or asteroids. As such, the fauna of Europa are of great interest to transhuman bioscience. 

\subsubsection{Biosciences}
\label{sec:biosciences} 

Europa's lifeforms, unique perhaps in the universe, are its greatest treasure, and transhumanity's efforts to catalog them are only beginning. The rush to exploit Europan biodivesity puts the Jovian Junta in an uncomfortable situation. While they control space traffic and commerce in the Jovian system, they lack the native talent to take real advantage of knowledge gleaned from Europa. At first, they engaged in hamfisted excise operations aimed at squeezing revenue out of knowledge exports. But once farcasters and egocasters came online below the ice, this type of extortion no longer worked. Now the Jovians have shifted to a two-pronged strategy of levying tariffs on new equipment and people brought down the ice elevators by hypercorps and research collectives, and of holding the entire population of the moon hostage by refusing delivery of key resources like reactor mass and rare elements if protection fees are not paid. 

\subsubsection{Habitats}
\label{sec:habitats-4} 

Europan habitats take two forms: fortified fishing and farming havens clinging to the spires of the lithodermic reefs, and spherical bubble warrens constructed by boring into the lower reaches of the ice crust and shoring up the hollows created. The latter are the only air-filled spaces beneath the ice. The largest warren is Conamara, at the base of the Conamara Chaos ice elevator. Conamara is surrounded by five nearby reef havens, also considered part of the habitat. The total population is 1.5 million. 

\subsection{Ganymede and Callisto}
\label{sec:ganymede-callisto} 

Nearly as large in size as Luna, but darkly colored and not as heavily cratered, Ganymede and Callisto are very similar worlds. Neither is as dense (nor has as much gravity), as their mantles consist of more ice than iron rock. Both possess abundant volatiles and water (albeit frozen), making them ideal candidates for habitation. Ganymede, with its differentiated surface of rocky and icy terrain, has an iron core and thus a faint magnetic field. Callisto, the smaller of the two, is composed mostly of icy silicate clays. As on Luna, most cities on Ganymede and Callisto are built below ground to shield them from meteor impacts (and, on Ganymede, from Jupiter's radioactive bombardment). 

While within the “protection” of the Jovian Republic, both moons are a patchwork of city-states. Some are full members of the Jovian polity, while others are only tolerated. Ganymede tends to swing more heavily toward the Junta, as its citizens still see the Junta-maintained infrastructure—accurately or not—as necessary in such a hostile environment. Callisto, outside the worst radioactive effects of the Jovian magnetosphere, is an easier place for technoprogressivism to gain a foothold. 

\subsubsection{Hyoden}
\label{sec:hyoden} 

The nucleus of this city-state was a research station founded by a coalition of Pacific Rim nations in Callisto's Valhalla region, a massive primordial impact zone where the icy subsurface lies exposed, simplifying extraction of clean water. When the Fall came, Hyoden, which had long faced labor shortages, opened itself to those refugees who could make it to Jupiter. Now Hyoden has two million inhabitants, making it the largest city-state on Callisto and the largest non-Junta state in the Jovian system. Hyoden is itself heavily militarized, as the tendency of the local authorities to turn a blind eye toward operatives using their territory for forays against the Junta makes for uneasy relations with their powerful neighbor. 

\subsubsection{Liberty}
\label{sec:liberty} 

Situated along the southern edge of the vast, rocky plain called Galileo Regio, almost on Ganymede's equator, Liberty (population 7 million) is the Junta's largest planetary city-state. It is closely tied to Liberty Station, a major shipyard and defense installation in geosynchronous orbit. Major industries include shipbuilding, space construction, fabrication, and security products and services. The Castle, the central security network point from which all surveillance data collected in the Junta is monitored and processes, is rumored to be in or near Liberty. Liberty is mostly underground, but it boasts a number of parks in armored surface domes. If one were to spend enough time topside, one would see the deceleration torches of incoming metal asteroids from the belt bound for the shipyards lighting up the sky several times a day. 

 \subsection{The Trojans (Jovian Trojan and Greek asteroids)}
\label{sec:troj-jovi-troj} 

The Trojans and Greeks are two 600 million-kilometer-long arcs of scattered, icy rock asteroids sharing the orbit of Jupiter. They orbit in the stable L4 and L5 points sixty degrees ahead of and behind the giant planet. Mars and Neptune also have Trojan asteroids, but when someone speaks of, “the Trojans,” they're normally talking about the Jovian groups. In the early days, L4 asteroids (ahead of Jupiter) are named after Greek heroes of Homer's Iliad; L5 asteroids (trailing Jupiter) are named after heroes of Troy. Asteroids discovered more recently break the old convention, as there are far more objects in the Trojans than there were characters in the Iliad. 

Politically, the Trojans and Greeks may be thought of as a collection of sometimes overlapping neighborhoods whose inhabitants tend to group around particular cultures, factions, and sometimes languages. A neighborhood in the Trojans might span anywhere from 250,000 to 2 million kilometers at its widest point. Within neighborhoods, almost everyone knows one other. Because of the wide dispersion of resources, Trojan habitats tend to be small—from one to two thousand people—and built largely along scum barge or cluster lines (although it is never advisable to refer to someone's habitat as a scum barge unless they refer to it that way first). 

\subsubsection{Resources and economics}
\label{sec:trojans-resources-economics} 

Although the sheer size of the two regions means a lot of cultural diversity, anarcho-collectivism is a powerful meme here and the reputation economy is prevalent. On one hand, neighborhoods, habitats, and even individuals are expected to be self-sufficient. Unlike the denser Main Belt, the Trojans lack the safety net provided by pervasive transhuman presence. The ideal Trojan or Greek is a Neo-Renaissance being, incredibly competent in a wide variety of fields. A person who can't maneuver in zero g; maintain their gear, ship, and hab; and navigate between rocks and habitats can have a tough time surviving. At the same time, a spirit of cooperation prevails. Bartering services or even gifting them to gain reputation is common. Everyone appreciates a specialist, as long as they're not specialized at the expense of baseline self-sufficiency. 

Prospecting and salvage are major activities in the Trojans, where metals and rare elements are scarce and settlers don't usually have the economic muscle to import raw materials from elsewhere. However, the Trojans are rich in silicates, volatiles, and carbonaceous materials. Necessity has led to many innovations in materials science. Beyond the simple problem of raw materials, the widely scattered habitats of the Trojans have to be wildly inventive on many levels to retain their independence. New robot, morph, and vehicle designs appear all the time, enabling an unusual array of business and leisure activities, like whaling (organizing a flash flotilla to rapidly mine asteroids and comets with erratic orbits as they pass near the Trojans), mekking (simulated—or sometimes real—combat between robotic suits or synthetic morphs on uninhabited asteroids with interesting terrain), and shrining (stealthing up on another habitat and resurfacing it with nanosculpters to create an art object—mostly a scum barge pastime). 

\subsubsection{Locus}
\label{sec:locus} 

Locus is the largest cluster habitat ever formed. It is still growing, with over one million inhabitants in the habitat proper and another million in the nearby suburbs of scum barges and small asteroid stations. Locus is located in Cassandra's Reach, one of the denser regions in the L5 Trojans. The habitat is positioned at the center of mass around which the two asteroids making up the binary object Patroclus orbit one another. Both Patroclus asteroids are themselves inhabited and hold defense installations, mines, and refineries. 

The design of Locus is very similar to the much smaller Lot 49, but Locus is eleven kilometers in diameter and somewhat irregular in shape, as growth along some spars is faster than others. A quarter of its total volume is cut out in a roughly conical shape all the way to the Amoeba, an immense, softly glowing sculpture at the center of the habitat. Some differences from smaller Trojan clusters are dictated by Locus's size. The immense structural spars radiating from the habitat's center are hollow, with arterial floatways and elevator-trams running inside of them. Lesser spars run between the arterial spars, providing more mooring points for modules. Adjacent to each arterial spar are wide “roads” leading to the edge of the habitat so that modules can maneuver out if the owners decide to leave. 

Beneath the shimmering mesh stretched over the geodesic frame to keep out micro-asteroids, tens of thousands of small ships and habitat modules moored along the spars pulse with an ever-changing array of lights. Habitat modules and large ships are asked to stay out of the conical empty space. This space teems with small craft and people on thrustpacks or voidscooters as they cross the habitat, play zero-g games, or visit the free-floating spimes and sculptures that dot the area. The Amoeba, which periodically changes color and shape based on its resident AI's programming (often it looks like some sort of animal), serves as a central reference point for navigation. When someone gives the address for a module, it is as a point on a spherical coordinate system with the Amoeba at its center. Large ships and shuttles dock on the outer surface of the habitat, at the terminal points of the arterial spars. 

Locus was founded by a joint anarchist-argonaut venture and was the first major stronghold for the autonomist factions. Unlike Extropia, which has the tacit blessing of the Planetary Consortium and encourages the presence of security and insurance companies, Locus runs on a pure reputation economy. Security, maintenance, expansion, and defense of the habitat are all performed by volunteers. Inhabitants interested in security monitor incoming ships and operate crowdsourcing systems that dispatch volunteers to perform WMD scans on new arrivals. Ships that won't submit to a scan are asked to leave. If they don't, anyone who's designed a cool new weapons system recently is welcome to take a shot. 

Locus is one focal point in a cold war between the hypercorp-aligned inner system powers and a loose coalition of outer system interests. While saboteurs from the Planetary Consortium and other hostile entities can and do occasionally cause trouble on Locus, the hypercorps are currently unwilling to attempt a direct military attack on the habitat. The first time they tried, the Planetary Consortium and the Martian city-state of Valles-New Shanghai sent a small expeditionary fleet. The interlopers were caught completely off-guard by a fierce and wellcoordinated defense. Six months later, they sent a much larger fleet. Help arrived from elsewhere in the Trojans and Greeks and from Titan, whose citizens took a dim view of any Planetary Consortium expansion beyond the belt. The Titanians now maintain a permanent base near Locus. Rumor has it they agreed to a mutual defense pact with one of Locus' citizens, possibly the famous programmer-armsman Teilhard Liu. 

\begin{quotation}
“Welcome to Locus. You voluntarily assume the risk of organic damage or mental trauma by mooring here. You must bring or be capable of acquiring enough food, H20, oxygen, and shelter to survive for the duration of your stay in a harsh, asteroid-rich environment. Weapons of mass destruction are prohibited. Further guidelines for coexisting with your fellow entities are in the habitat survival guide. You and only you are responsible for yourself—learn to love it!” 

 -- Locus Immigration AR broadcast 

 “You have chosen the habitat Locus in the L5 Trojans as your destination, using the private carrier Atsuko van Vogt as your receptor. ComEx corporate policy requires us to inform you that the destination and carrier you have selected are unregistered and possibly unsafe. ComEx takes no responsibility for the continuity of your consciousness upon arrival. You assume any and all risks for travel to this point, including theft of forks or deletion. ComEx will include a permanent record of travel with this continue?” 

 -- ComEx legal disclaimer 

 “The ComEx disclaimer? Yes, yes ... Listen: my neighbor three doors toward the Amoeba from here is a physicist. She has a box that generates micro-singularities in her lab. If people along my spar found out I’d stolen a fork of someone, they’d pop my stack with a grapefruit knife and throw it in there. That’s what we call, ‘accountability.’ See if you get the same from ComEx.” 

 -- Atsuko van Vogt
\end{quotation} 

\subsubsection{Lot 49}
\label{sec:lot-49} 

Lot 49 is moored to the small asteroid 28349 Pynchon in the amorphous Vonarburg-Shadyside neighborhood, toward the center of the L4 Greeks. VonarburgShadyside is named after two rocks that roughly delimit its 500,000-kilometer length along the arc of Jupiter's orbit. Neighboring habitats within 100,000 kilometers (with populations) include Craftsbury (450), Greenview (28), and Blackhawk (1020). With a population of 400, this station is more or less typical of the Trojans in terms of layout. 

From the outside, Lot 49 looks like a shiny, meshedover geodesic sphere, 800 meters in diameter, with numerous protruding instrument spars and some triangles left open to space so that shuttles can pass through. The mooring to the asteroid is temporary in case a potential collision is detected. Inside, a central utility module with a communal reactor, factories, and machine bay is surrounded by evenly spaced but irregularly shaped habitat modules in a riot of colors and lighting schemes. Structural spars and floatways connect everything. One entire spar is given over to a rotating cylindrical module that generates about 0.7 g and contains medical, cloning, resleeving, and darknet egocasting facilities. 

Lot 49's population and most of their neighbors in Vonarburg-Shadyside tend to align with the scum and anarchist factions and speak a mixture of English, Portuguese, and Thai. Lot 49 is in a densely inhabited part of the Greeks, placing it near a crossroads. Main economic activities include shuttle design, whaling, and ferrying people and goods around the region. 

\subsection{Saturn}
\label{sec:saturn} 

The second largest planet in the system is a much more favorable habitat for transhumans than Jupiter. Saturn's lower gravity and milder magnetosphere are a boon to gas mining operations, and for resourcehungry habs, the Rings are a feast (literally, in the case of the new Hamilton cylinder type habitats). Hypercorps have a presence here, but any major expansion by the Planetary Consortium is kept in check by the anarchist stations of the Rings and the technosocialist Commonwealth of Titan. 

Because Saturn is so distant from the Sun, solar power generation is extremely inefficient. Growing photosynthetic plants with sunlight is impossible without large arrays of mirrors to focus the light. The abundance of water and volatiles makes the rings ideal for both scum barges and Hamilton cylinders. Gas mining is vital to the survival of almost every habitat and moon settlement in the Saturnian system, so habitats located further out from the planet that wish to be self-sufficient almost always maintain their own gas mining stations close to the planet. Security for these installations and the atmospheric skimmers and tankers they dispatch is tight, and it is never advisable to approach one unannounced. 

\subsubsection{Resources and economics}
\label{sec:saturn-resources-economics} 

Gas mining on Saturn supplies thirty percent of the system's reactor mass. This role is expected to grow as Helium-3 deposits in the Lunar regolith become less accessible. For ships traveling to the far reaches of the outer system, Saturn is an important alternative to using Jupiter for gravity assists. Less restrictive than Jovian regimes and richer in resources than the Trojans, Circumsaturnine habs and settlements are important innovators in habitat design and cultural organization. Since the discovery of the Pandora Gates, the Titanian Commonwealth is the only entity actively pursuing interstellar exploration through conventional means. 

\subsubsection{The rings and classical minor moons}
\label{sec:rings-class-minor} 

Saturn's rings are made up of countless small icy objects, most of which range in size from dust specks to boulders 10 meters in diameter. The rings are designated by the letters “A” through “F” in the order in which they were discovered. They vary in thickness between 100 and 1000 meters and in width from 20,000 kilometers down to just meters. In places there are gaps between rings. The widest, the Cassini division, is 4,000 kilometers across. 

% GATE EXPEDITION REPORT 901127 fixme 

 \begin{quotation}
\textbf{Gate expedition report 901127}

Gate Code Setting: [Encrypted]

Gatekeeper Corporation Eyes Only 

Preliminary drone and sensor reports seemed to indicate the gate’s exoplanet environs were underground in a cavern formed of carbonaceous rock with a nitrogen dioxide atmosphere. There were no signs of life or sentient activity. A squad of gatecrashers was sent through, guided by an exploration drone, with a communication link back to the gate. 

Approximately one hour after the team moved into the tunnels, consistent communication was lost due to electromagnetic interference. At this point they had reported nothing more notable than moving just over a kilometer through a warren of tunnels. 

The team was not heard from again. 

Two hours after contact was lost, a tethered search and rescue drone was deployed. Following the gatecrashers’ breadcrumb trail, near the limits of its tether range the drone came upon what appeared to be a severed hand in a vacsuit glove. DNA testing did not identify the hand as belonging to any members of the team, however, nor did it match any other database queries. The drone was detached from its tether to search further, but shortly after sensors recorded some type of seismic event, and communication with the drone was also lost. 

The gate was kept active for 8 more hours—the duration of the contract - with no sign of activity. The gate was then closed, the team reported as lost/unretrievable, and the gate settings were recorded with an orange flag.
\end{quotation} 

%%% txt/105.txt 

Saturn has over 60 satellites, a number that jumps into the hundreds if one includes the uncounted objects less than a kilometer across orbiting in the A ring. Most of Saturn’s moons are small, rocky, ice objects less than 100 kilometers in diameter. The smallest of the classical moons, Pan, is only 10 kilometers across. The first eight moons, from Enceladus inward, lie within the ring system. Atlas, at the edge of the A ring, and Prometheus and Pandora, which flank the thin F ring, are known as the Shepherd Moons. Several of the moonlets occupy Lagrange points relative to larger moons. Telesto and Calypso share the orbit of much larger Tethys, while Helene trails another large moon, Dione. 

\subsubsection{Atlas (Volkograad)}
\label{sec:atlas-volkograad} 

Volkov, a Slavic energy cartel, controls this tiny moon. Volkograad is a beehive habitat with about 50,000 residents. Much of the moon is given over to skimming, refining, and shipping infrastructure. A cloud of wreckage trailing the moonlet by about 100,000 kilometers serves as a reminder of the Atlas Incident, a brief but massively destructive battle that erupted when Fa Jing attempted a buyout of the moon. Tinkers from Phelan’s Recourse still salvage the floating derelicts regularly. 

\subsubsection{Dione (Thoroughgood)}
\label{sec:dione-thoroughgood} 

Dione’s main settlement is Thoroughgood (population 350,000), a hybrid beehive and orbital cluster habitat set on a plateau amid a dramatic range of ice cliffs. Dione hosts the Long Array, a 150 kilometer-high communications spar ascending from the surface settlement to an orbital station that acts as a counterweight. The Long Array’s sheer size is something of a publicity stunt, as the bulk of its capacity goes unused. However, it drew enough attention to make Thoroughgood a major communications hub for the outer system, and thus a place where hypercorp, anarchist, and other factional interests meet. Dione shares its orbit with Helene, a tiny, rocky moon at its L4 point, and Polydeuces, an even smaller body that trails it at the L5 point. 

\subsubsection{Enceladus (Profunda)}
\label{sec:enceladus-profunda} 

Rich in organic compounds, Enceladus is a biochemist’s playground. Profunda (population 850,000) is the major settlement, a beehive dug into the moon’s surface capped by domed parks and clusters of sleek, translucent minarets—well protected from collisions by an aggressive satellite defense network. The lower levels, stretching deep into Enceladus’ icy silicate mantle, include a prospecting operation that extracts carbonaceous soils in search of exotic compounds. Another deep section has been converted into a kilometers-wide, reactor-heated primordial sea, part of a long-term experiment into the origins of life supported jointly by Titanian academics and a collective of Enceladian biochemists. 

Profunda is run along anarcho-capitalist lines. Thanks to the rich supply of organic chemicals, its upper reaches are home to many of the outer system’s best known morph designers. The Enceladian Glitter Bloc is said to have as much influence over body styles as the Lunar fashion houses do over what people wear. 

\subsubsection{Epimethus and Janus (Twelve Commons)}
\label{sec:epim-janus-twelve} 

These twin small, icy moonlets share virtually the same path around Saturn, orbiting within 50 kilometers of each other. Set between the F and G Rings, the moonlets form the center of Twelve Commons, a neighborhood of small and mid-size habitats arranged in a flat cloud about 20,000 kilometers in radius. About six million people live in Twelve Commons. Habitats in Twelve Commons range in size from Dang Fish Echo, a tin can hab housing about 60 eccentric aquaculturists, to Janus Common, a beehive occupying much of Janus with a population of 900,000. Some of the habitats in Twelve Commons feature very unusual designs, such as Nguyen’s Compact (population 80,000), a variant Cole habitat in the G ring where an asteroid was heated and large amounts of steam were blown through it to produce a series of interconnected bubbles between five and three hundred meters in diameter. In effect, the interior of the colony is like a solidified foam or Swiss cheese with no obvious up or down. Without an ecto or basic implant to provide location and navigation information, navigating through this maze-like habitat would be exceedingly difficult. 

The habitats of Twelve Commons organize themselves primarily along open source anarcho-syndicalist lines, with work groups and research pods acting as the basic political unit. 

\subsubsection{Gateway (Pandora)}
\label{sec:gateway-pandora} 

The Gateway settlement, on Saturn's outer shepherd moon Pandora, holds the first publicly known wormhole gate. The Gatekeeper Corporation keeps the gate open as a means of exploration and scientific investigation for all factions and powers. Gatekeeper was originally a Titanian microcorp but is now independent. The Commonwealth of Titan still holds a major stake in it, though not a controlling interest. Granting autonomy to Gatekeeper Corporation was a diplomatic maneuver made in response to Planetary Consortium claims that the Titanians sought hegemony in the outer system. So far, Titan's neighbors are buying it, even if the Planetary Consortium doesn't. 

\subsubsection{Hyperion}
\label{sec:hyperion} 

With its chaotic, virtually unpredictable rotation, Hyperion is a dangerous place to land ships. It remains uninhabited. 

\subsubsection{Iapetus}
\label{sec:iapetus} 

Iapetus is one of Saturn's larger icy moons and once boasted a population of 200,000 living in the dense warrens of Analect, its main settlement. Probably because it is one of the few large moons of Saturn that contains sizable deposits of silicates and minerals in addition to ice, Iapetus was a target of the TITANs during the Fall. After enslaving a tenth of the populace as worker drones and using the rest as seed stock for tissue cultures to feed their fellows, the TITANs began to build what appears to have been a matrioshka brain. Iapetus now occupies twice the volume it once did, the ice and silicate of the planet's outer layers having been reworked into a delicate lattice of circuitry millions of layers deep. 

Strangely, the project simply stalled at some point prior to completion. Speculation has it that the controlling intelligence was either destroyed by an unknown outside force or devoured itself in a fit of compuational ecstasy. Whatever the case, the drones simply stopped working and died and the moon's automated defense grid went dead, leaving a strangely beautiful but lifeless machine behind to slowly decay from meteor impacts and gravitational stress. Several research teams now reside in small orbital stations, quarreling over the scraps. Rumor has it that a number of researchers trying to understand the matrioshka circuitry have lost their minds in the process, perhaps by some mechanism akin to a basilisk hack. It is also believed that some of the moon's internal defenses remain active. If anyone has plumbed the interior and come back, they're not talking about it. 

\subsubsection{Meathab}
\label{sec:meathab} 

The full name of this unique habitat is Turn Yourself Into a Giant Mass of Space Meat for Art!, and as the name implies, 90\% of the habitat's structure consists of fast-cultured vat bacon, battened on the abundant resources of the ring system. MeatHab started out as someone's art morph, but then, against all expectations, squatters moved in. MeatHab now has a population of 500. Similar to a Hamilton cylinder, the kilometer-long habitat harvests and processes ring material to grow itself. The outer surface is frozen flesh ten meters thick whose surface resembles a cross between a tree trunk and flank steak. Past the axial space dock is a warren of veinous, skin-covered corridors lit by bioluminescent panels and maintained by small, reptilian symbiotes that eat away dead skin and may have other immune functions as well. Gravity inside is 0.5 g. 

The nameless biodesigner who created the place—and who may or may not still inhabit the gigantic morph—was a genius. Although the habitat is not by any stretch of the imagination a pleasant locale, it appears healthy. Its full workings are not understood, and the inhabitants range from extreme flesh freaks who are fans of the artist to serious biodesigners studying the place to learn more about its construction. 

\subsubsection{Mimas (Harmonious Anarchy)}
\label{sec:mimas-harm-anarchy} 

Led by legendary Chinese dissident poet Hao Lin Ngai, Harmonious Anarchy broke from the Fa Jing cartel during the tumultuous years prior to the Fall. Hao sought to create a society in the spirit of the ancient Taoist state of Great Perfection that existed in Szechuan 1,700 years earlier—with considerable updates from modern thought. Harmonious Anarchy is an Extropian mutualist society heavily involved in software engineering, logistics, and relocation of metallic asteroids to the outer system. Most of Mimas is a very low-g beehive arranged into Black, Red, Yellow, Green, and White neighborhoods, based on the five classical directions of Chinese mythology. Each color boasts an ornate central cavern, with extended families living in radiating subwarrens. While adhering to mutalist economic principles, Harmonious Anarchy simultaneously takes a traditional Chinese approach to social organization, with family at its core. 

\subsubsection{Norse, Inuit, and Gallic moonlets}
\label{sec:norse-inuit-gallic} 

In addition to the classical satellites described here, three groups of small objects unknown to early astronomers orbit the planet. These moonlets are designated as the Inuit, Gallic, and Norse groups. Because these moonlets were still little explored by the time of the Fall, most of them remain sparsely populated. With a few exceptions, inhabitants of the moonlets are generally people who want to be left alone. The exceptions are Skathi and Abramsen (formerly S/2007 S 2), which, along with Phoebe, were captured and moved into Titan's orbit, where they serve as defense installations. 

\subsubsection{Pan (Izulu)}
\label{sec:pan-izulu} 

Volkograad's closest competitor is this anarchocapitalist outfit, most of whose founders were South African. iZulu has a somewhat lower capacity than Volkograad but will ship reactor mass to unusual locations like the Trojans and the Kuiper Belt. iZulu is a very crowded beehive with nearly 400,000 inhabitants and an unusually large number of infugees. The nations of sub-Saharan Africa were only starting to achieve widespread 20th century-levels of prosperity in the late 21st, and so they had the lowest capacity to physically evacuate their citizens during the Fall of any region on Earth. iZulu and a handful of other habitats with roots in Africa thus have high infomorph populations and millions of people in dead storage. 

\subsubsection{Phelan's Recourse}
\label{sec:phelans-recourse} 

Phelan's Recourse (usually just called “Phelan's” by inhabitants) is the largest nomadic settlement in the system, with a population estimated around 250,000. Phelan's is a swarm of some 10,000 small craft and tin can habitat modules that orbits Saturn along a highly elliptical path somewhat inclined to the plane of the ecliptic. The swarm's orbit is calculated to maximize the number of encounters with near moons and stations, providing a six to eight hour window in which craft can leave the swarm for trade. Phelan's Recourse passes through the rings once a month, allowing craft to resupply with water and volatiles. 

Phelan's accepts all comers. One could meet just about anyone here, from the government in exile of East Timor to Hasidim from Brooklyn. The core of the swarm is the Stills, a fusion-illuminated grain farm and distillery operated by an allegedly reformed gang of Irish travelers who conned their way off Earth a few weeks before the Fall and escaped to the outer system. The Stills produce Phelan's Ma, the most sought-after whiskey in the system, and Phelan's Da, possibly the worst beer ever made. Despite the Phelans' protestations of legitimacy, the criminal element is heavily represented here. The swarm represents an important link in red and gray market supply chains. 

\subsubsection{Prometheus (Marseilles)}
\label{sec:prom-mars} 

Marseilles (population 80,000) is a beehive habitat operated by the Titanians. It is rumored to harbor an antimatter factory, a theory supported by the large number of skimmers that arrive from the surface relative to the number of tankers that leave. 

\subsubsection{Rhea (Kronos Cluster)}
\label{sec:rhea-kronos-cluster} 

At a 764 kilometer diameter, Rhea is Saturn's second largest moon. Composed almost entirely of ice, Rhea's surface is sparsely inhabited, but a population of over 800,000 dwells in Kronos Cluster, a major habitat in orbit. Kronos Cluster's mass microfactured violet spherical modules make it look like an immense, irregular bunch of grapes suspended in space, an impression added to by the winding space dock (nicknamed the Vine) extending from the wider end. Within the mass of habitat modules, the Vine branches out in all directions, forming massive central arteries and twisting side passages. These can be traversed by pushing off hand and toeholds on the walls, or by catching hold of fast-moving grab loops that move along “fast lanes” in the walls of major floatways. 

Nearly five kilometers long and three wide, Kronos has major problems with crowding and infrastructure that have kept it from growing to the same size as Locus. The designers simply did not plan for the size the place might reach, and as a result another 150,000 people live in suburbs of tin can habs and scum barges in the space around the habitat. 

Kronos can be an extremely dangerous place. Insurance companies don't like operating here, and the habitat is a patchwork of criminal and anarchist neighborhoods. Anarchist neighborhoods are generally heavily armed and safe, but a trip from an anarchist holding to the spaceport is best done with a group of well-armed friends. Criminal neighborhoods are only safe if you're in the neighborhood's controlling gang, and even then conflicts flare up regularly. 

The situation is exacerbated by the Kronos Port Authority, a junta of ultimates who operate security for the spaceport. Originally an Extropian hypercorp, the KPA fell into the hands of the ultimates when they decided that they could profit more directly by owning the company outright than by working as hired muscle. They violently ousted the original management and now use indentures in worker pods to maintain the port. This situation is tolerated by the local crime bosses and loathed by the mostlyanarchist autonomist citizens, but so far no one is able to challenge the KPA, which enforces use of the port rather than any other mooring point with killsats and artillery. 

\subsubsection{Tethys (Gowwinhead)}
\label{sec:tethys-gowwinhead} 

Composed almost entirely of ice, Tethys is one of Saturn's larger moons and the site of Ithaca Chasma, a 2,000-kilometer long valley covering three-quarters of Tethys's circumference. Fifteen years ago, prospectors from an ethnically Indo-British autonomist collective called the Rioters touched down on Godwin Head, a projection in the chasm wall so named because it resembles a headland projecting out into the sea. Instruments on their ship, the Caleb Williams, had detected what looked like mineral deposits in the ice, rare on Tethys. What they found instead were relics thrust to the surface by a geological event eons earlier, the remains of primordial life that became extinct millions of years ago when Tethys cooled and its subsurface ocean froze over. 

Godwinhead is now a dense, efficient settlement of 200,000 built into the five kilometer high canyon walls. The central point of the town is the Caleb Williams, which has been towed back into a sheltering cavern in the wall and converted into a communal workshop and town hall. The face of the valley wall is honeycombed with excavated ice caves hosting habitat modules, connected by conduits to a communal utility grid. The trusswork and cabling for the utility system is also the public transit system, easily traversed in the minute Tethyan gravity. The unofficial mascot of Godwinhead is the Tethyan Flatworm, a two millimeter-long translucent worm that represented the pinnacle of Tethyan evolution. A large number of the inhabitants are involved in biosciences, xenopaleontology, and prospecting for frozen lifeforms. 

Tethys shares its orbit with its Trojan moons Telesto and Calypso, both of which are small and sparsely populated. 

\subsection{Titan}
\label{sec:titan} 

Saturn's largest moon is shrouded in a permanent orange atmospheric haze, hellishly cold (averaging 180 degrees below), and whipped by winds produced by tidal forces four times stronger than those influencing the Earth's climate. On its face, it appears even less hospitable than the airless balls of ice and rock comprising every world between Titan and Mars. The meager sunlight reaching its surface is insufficient to grow any but the hardiest plants, the mostly-nitrogen atmosphere is dangerously toxic, and the surface is dotted with lakes and seas of liquid methane. In spite of all this, abundant hydrocarbons, a thick atmosphere, and diverse chemistry make Titan one of the few worlds in the system where colonists may rely entirely on local resources. Titan's population is now over 60 million. 

Social money and the microcorp system have led to some spectacular gains and failures. On the up side, Titan's civil resleeving industry produces more morphs than Mars and Luna combined. Massive infrastructure programs have provided enough space for 60 million people to live comfortably on a hostile world. The Large Collider, the biggest particle accelerator ever produced, in polar orbit, enables physics experiments that can be performed nowhere else in the system. And two years ago, Titan dispatched the first conventional interstellar probe, the Aubade. It will reach Proxima Centauri in just under 20 years. 

On the down side, Titan's “body for every mind” law burdens the civic resleeving system with a lot of people who no one would ever have bothered resleeving otherwise. The failure of the Scoop project, an extremely costly attempt to build a pipeline from Saturn's surface to low orbit, allowing massive gas extraction without costly atmospheric skimmer operations, stymied Titan's ambitions to become a major antimatter producer. Titan does produce antimatter, but on a much smaller scale than was envisioned when the Scoop project began. 

Commonly spoken languages on Titan include Norsk, Francais, Deut ch, Mandarin, Svensk, Dansk, and Suomi. Most citizens inhabit hazers, a tall, fineboned morph with very similar characteristics to the Martian ruster. Parapelagia for gliding and flying in the light Titanian gravity are a common biomod. Titanians do three years of compulsory civil service at the age of majority, with an emphasis on military and security forces except for conscientious objecters. Every citizen who has done military service is part of the militia and has an assault weapon in their home. 

\subsubsection{Aarhus}
\label{sec:aarhus} 

Located near Titan's south pole on the shores of Ontario Lacus, a wide, shallow sea of liquid methane, Aarhus (population five million) was the first site of human habitation on Titan, chosen for its proximity to abundant hydrocarbons. The city is the physical hub of Titan Autonomous University (TAU) and hosts numerous other academic institutions, most notably Titan Tech, a major engineering school. Unlike Martian universities, which have few physical campus buildings, TAU and other Titanian schools draw many of their students from the widely scattered habitats of the outer system, where delays in radio communication make distance learning ineffective. Fully 20\% of Aarhus's population are students, many of them offwordlers. 

Aarhus's layout is typical of Titanian cities. Three central domes are surrounded by numerous smaller structures, including lesser domes, fusion plants, and industrial outbuildings, the most massive of which is the now-abandoned methane utility plant on the lake shore. The dome interiors are hung with lighting rods and heavily built up with tall, narrow buildings, most of which have upper decks where hazers on the wing and pedal-powered microlights can land. Exterior structures usually have outer walls built of ice for shielding and structural support with internal walls extruded from local silicates. Many buildings are a rich azure or other shades of blue for contrast with the ever-present orange glow of the Titanian sky. 

Unlike most Titanian cities, Aarhus relies primarily on fusion power. Aarhus is the center of Titan's native preservationist movement, which opposes inefficient use of native hydrocarbon resources due to possible long term effects on Titan's climate. 

\subsubsection{New Quebec}
\label{sec:new-quebec} 

New Quebec lies on a plain in the Aaru region surrounded by endless rippling dunes shaped by Titan's powerful winds. The region's diverse chemical resources supply the colossal nurseries that have made New Quebec the system's largest single producer of morphs. 

The city is 50 kilometers from Montmorency Lacus, a 20 kilometer-wide crater lake of liquid ethane and methane. Originally thought to be an impact crater, rare on Titan, geological studies later showed it to be the collapsed remains of an extinct cryovolcano. Situated in a rainy area, the lacus slowly drains over the crater lip at Montmorency Cascade, a 200 meter carbonfall that empties into a series of alluvial channels from which the Quebecoise pump its output for fuel. 

The St. Catherine Tong, the most dangerous native Titanian mob, is based in New Quebec. Titanian law is generally very permissive regarding individual freedoms, so the vices this gang trades in are of the blackest: snuff pods, stolen alpha forks, and nanoweaponry. A ready supply of fresh morphs bought from corrupt microcorp nursery administrators further fuels their rackets. The Tong is extremely violent and a major embarrassment to Commonwealth security forces. 

\subsubsection{Nyhavn}
\label{sec:nyhavn} 

Set near the equator amid the rolling ice hills of the Xanadu region, Nyhavn (population 12 million) is the largest city in the outer system and the capital of the Titanian Commonwealth. Nyhavn's massive central dome, with its elegant blue towers and bioengineered parklands, rivals New Shanghai in size and ambition. Three surrounding domes and a sprawl of subsidiary structures are connected by high-clearance flyways, where ground vehicles and microlights form a steady stream of traffic at all hours. At the same time, the squalid blandness that prevails in the Martian suburbs and outlying souks is absent; the dwellings and neighborhoods of the Titanian working class display a riot of color and design, empowered by public fabricators limited by none of the enforced scarcity of Martian economics. For all its idealism, the Plurality is not immune to a desire to showcase its achievements. 

Outside the city is a pipeline leading from the vast Tyska Lacus, 100 kilometers distant. Commonwealth Skyport, Titan's principal spaceport, offers quick access to Commonwealth Hub, the Titan system's long-haul space dock, located in geostationary orbit above the city. The surrounding countryside is dotted with smaller settlements connected to Nyhavn by trains and a well-developed network of surface roads. 

Nyhavn is a major media center, with daily life closely attentive to the debates and decisions of the Plurality. At the same time, it is a cosmopolitan place, where Titan's microcorp movers rub shoulders with visiting anarchist traders and (less commonly) legations from the inner system. There is an active underworld, despite the efforts of security forces, with the local St. Catherine Tong engaged in continual low-intensity warfare with triads from throughout the system. 

\subsubsection{Phoebe, Skathi, and Abramsen}
\label{sec:phoebe-skathi-abrams} 

After the conflict at Locus, the Plurality became embroiled in a hot debate regarding the dangers of hypercorp adventurism in the outer system. It was generally felt that the Planetary Consortium hoped to keep the outer system in a position similar to where the United States kept Latin America by meddling in its affairs throughout the nineteenth and twentieth centuries, and that the only counter to this was a show of force. Titan's thick atmospheric haze makes ground-based space defense systems considerably less effective than on other worlds, but satellites and space platforms were too vulnerable to serve as command and control centers. 

The solution was to capture three of Saturn's small retrograde moons—Phoebe, Skathi, and Abramsen (once designated S/2007 S 2, now renamed after a pioneering Titanian economist). Phoebe is the largest of the three objects. The other two were maneuvered into the system's L4 and L5 points. The calculations required to relocate these bodies were painstaking, and the energy expenditure tremendous, but all three now serve as major components of Titan's orbital defense grid. Whether the system created thereby is impregnable has yet to be tested. 

\subsection{Uranus}
\label{sec:uranus} 

Once thought of as gas giants like Saturn and Jupiter, Uranus and Neptune differ from the larger planets in that they contain large amounts of water ice, methane, and ammonia and have rocky cores at their centers. This region of the system is sparsely populated. Uranus orbits at a distance 10 AU beyond the orbit of Saturn, 20 times the distance of the Earth from the sun. 

Uranus, the coldest planet in the solar system, is a blue-green sphere of ice and gas. Seen from afar, it is virtually featureless compared to Saturn and Jupiter, but up close subtle cloud formations and a tenuous ring system may be observed. Probably due to a collision with an Earth-sized world when the solar system was young, Uranus rotates on its side, such that one pole faces the sun for 42 years at a stretch, and its moons orbit at a sharp angle to the solar ecliptic. 

At the time of Eclipse Phase, Uranus's south pole is experiencing its south polar mid-spring, during which thick methane clouds darken the polar atmosphere. It may be the unusual tilt of its axis and the accompanying strange seasonal weather that give rise to the unconfirmed rumor that the alien traders called the Factors have created a settlement hidden in Uranus's atmosphere. 

\subsubsection{Chat Noir and Fissure gate}
\label{sec:chat-noir-fissure} 

Located on Oberon, this is the Uranian system's primary long haul spaceport, with a permanent population of 8,000. Chat Noir has fairly advanced egocasting, resleeving, and manufacturing facilities for a frontier outpost and is operated by several collectives of anarchists. The reason for all the infrastructure is Fissure Gate, the only Pandora Gate in anarchist hands (despite several Planetary Consortium expeditions to wrest control of it). 

Fissure Gate was discovered by a prospecting expedition from Chat Noir, then a tiny outpost. Seeking deposits of the useful carbonaceous ices that make up about 20\% of Oberon's mass, they instead chanced upon subsurface radio emissions near the foot of Mt. Hippolyta. After using triangulating the source, the prospectors landed and used subsurface imaging gear. What they got back was a blurry image of a rock fissure containing an ambiguous mass of mixed density and an extremely dense, possibly metallic object with a shape too regular to be anything but a structure or large artifact—all under 500 meters of ages-old frozen cryovolcanic outflow. The gate at Pandora was already publicly known at this time, so the prospectors drilled down, suspecting they'd found an alien artifact. They were not to be disappointed, although the discovery yielded gruesome salvage: the barely recognizable corpses of eleven gatecrashers. 

Why and how the Fissure Gate was erected under the ice remains a complete mystery. At some recent point, however, it was completely buried, with only a thin pocket of space between it and the surrounding ice. When the eleven emerged, buried in an airless space beneath 500 meters of ice, there was barely room to move, let alone escape—but the gate wouldn't let them back through. Several of the crew had recoverable cortical stacks. This lucky handful are now prominent citizens of Chat Noir, but none plan to resume gatecrashing as a career. 

The Fissure Gate remains in anarchist hands, operated and defended by the Love and Rage Collective. The gate is made available to almost anyone unless their rep score is tanked or they are pursuing commercial interests (ruling out most hypercorps). Support for gatecrashers is minimal—traverse the threshold at your own risk. Any discoveries made via this gate, however, must be shared for the collective good of transhumanity. 

\subsubsection{Titania and Oberon}
\label{sec:titania-oberon} 

Uranus's two largest moons are sparsely populated, with only about 10,000 transhumans living on each body. Most stations are mixed dome and beehive settlements and range from hypercorp communications and research outposts to autonomist freeholds. The pair are more chemically complex than most moons in the outer system, consisting of about 30\% rock, 20\% methane and similar carbonaceous ices, and 50\% water. Titania is home to a spectacular canyon that rivals the Martian Valles Marineris. Several settlements on Titania cater to tourists from the inner system and the gas giants, who visit for rocketing, mekking, and other sports in the canyon. 

\subsubsection{Xiphos}
\label{sec:xiphos} 

One of two major strongholds of the ultimates, Xiphos is a Hamilton cylinder orbiting in the Uranian ring system. Though most of the tech underlying Hamilton cylinders is open source, the station's frighteningly efficient weapon systems are not. Rumor has it the ultimates traded some major favors to Gorgon Defense Systems in the process of building this station. Where Aspis, the ultimates' inner system habitat, is a relatively open place, used by the Ultimates for contact with potential mercenary clients, Xiphos is off limits to anyone not of this faction. The rumored population of ultimates here is only about 10,000, but the ultimates purchase a large number of infomorph indentures from Mars. Although there are no reports of any of these indentures returning, rumor has it that the ultimates download indentures serving in sensitive areas into deaf, visually limited flats with no AR implants and limited mental capacity. 

\subsection{Neptune}
\label{sec:neptune} 

Frigid, swept by 2,100 km/h winds, and tinged blue by methane traces in its atmosphere, Neptune is the last major planet in the system, orbiting at a distance of 30 AU from the sun. This far from the nearest star, plants will not grow and solar power is useless. The only sources of power are fusion, focused starlight, waste heat, and chemical reactions. The hypercorp presence in the Neptunian system is virtually absent, as the long communication lags and extreme travel distances from the rest of the solar system mean that few Neptunian ventures garner profits. Similarly, the Titanian brand of technosocialism has never found roots here. The few transhumans who live out here are a resourceful lot, and many colonists out here aren't human at all. Anarchists, brinkers, and desperados comprise most of the population. 

\subsubsection{Glitch}
\label{sec:glitch} 

This habitat has the highest population density in the system, with 20,000 infomorphs living in a meshed cluster of twenty spherical structures that are 10 meters in diameter, powered by efficient central reactor systems. The habitat is attended by a cloud of factories, harvesters, and defense satellites that occupy considerably more space than the station itself. Various rumors circulate that the inhabitants are researching methods to improve infomorphs in the manner of seed AIs, or that they are engaged in some vast forecasting simulation effort. 

\subsubsection{Ilmarinen}
\label{sec:ilmarinen} 

Aligned with the argonauts, Ilmarinen is a hybrid beehive/cluster dug into and partially protruding from the large L4 asteroid Greymere. It is the largest habitat in Neptune's Trojans, with a population of 7,000. Like many transhumans this far out in space, most of Ilmarinen's inhabitants are heavily modified or inhabit exotic morphs. Vacuum and cold tolerant morphs prevail, and many sections of the habitat are unlivable for baseline transhumans. 

\subsubsection{Mahogany}
\label{sec:mahogany} 

The neo-avians who built this station threw away the manual on habitat design and revisited the longout of favor toroidal configuration. The result is a disc habitat—a plate half a kilometer along the edge and one kilometer in diameter, resembling a slice of an O'Neill cylinder with no windows. A fusionpowered, low-heat, axial light source nourishes the verdant hardwood forest below. Structures are built into the disc walls up to 500 meters in height. The disc, mostly woven from carbon fibers, rotates quickly enough to generate 0.5 g at the habitat floor. Mahogany has a population of 4,000 mercurials, most of them neo-avians. 

\subsubsection{Minor moons}
\label{sec:minor-moons} 

Neptune's other twelve moons are largely small bodies, icy and sparsely (if at all) populated. Proteus and Larissa, both sizable and relatively close to the planet, host small populations. Naiad and Thalassa are tiny but very close to the planet, and thus home to some atmospheric skimming operations. Neso, orbiting at about 1/3 AU from Neptune, has never been visited—even by robotic probes. 

\subsubsection{Neptunian Trojans}
\label{sec:neptunian-trojans} 

Trailing and preceding Neptune at the L4 and L5 points of its orbit are several hundred asteroids of diverse, mostly icy composition. Neptune's Trojans are home to brinkers, hard-bitten prospectors, exotic exhumans, and other extreme survivalists. 

\subsubsection{Triton}
\label{sec:triton} 

Neptune's largest moon has a tenuous atmosphere and is chemically complex, composed of equal parts rock and ices (frozen nitrogen, water, and carbon dioxyde). It is also geologically active, with cryovolcanoes continually resurfacing the planet. The surface has few inhabitants but several habitats orbit here, using the moon's abundant raw materials and low escape velocity to their advantage. 

\subsection{The edge of the system}
\label{sec:edge-system} 

Beyond Neptune lie only dwarf planets and icy asteroids waiting to become comets, roughly divided into two regions: the Kuiper Belt, from 30 to 55 AU from the Sun, and the Scattered Disk, which extends from 55 AU out to the Oort Cloud. Pluto, its binary object Charon, and Eris have compositions similar to Triton. A few small habitats orbit Pluto and Charon, eking out a living by prospecting for volatiles. A number of other dwarf planets orbit in the Kuiper belt and the Disk, including Orcus, Senda, and 2000 OO67. Of these, only Eris harbors a substantial population. ERIS Located at 55 AU from the sun at the edge of the Scattered Disk, Eris is the largest dwarf planet in the system and the site of a grim struggle between two of transhumanity's most militant factions: ultimates and exhumans. The focal point of the struggle is Discord Gate, the most remote of the system's publicly-known Pandora Gates, located in an icy labyrinth half a kilometer beneath the surface of Eris. 

The brief history of the gate is bloody. Go-nin Group troops violently wrested control of the gate from the Ilmarinen anarchists who discovered it. Titan and several anarchist and brinker groups both tried to dislodge Go-nin, but these attempts failed, at great cost in lives and ships. Go-nin's control of the gate seemed ensured until the hypercorp apparently tampered too heavily with the gate's black box controls. A devastating explosion ensued, all but wiping out the gate and Go-nin base. The gate, however, restructured itself over the course of several days, though its location has now shifted to the bottom of a melted crater. 

In the short period it took the Go-nin Group to hire a group of ultimate mercenaries to retake the gate, a hitherto unknown force of exhumans had seized the area. The ultimates succeeded, but a group of exhumans escaped through the gate. Go-nin now has nominal control of the Discord Gate through the ultimates, who maintain a heavily militarized base on Eris's moon, Dysnomia. However, in recent years the gate facility has suffered several attacks by exhumans eager to infiltrate the gate—and according to rumors, at least one of those attacks originated from the gate itself. 

\subsubsection{Markov}
\label{sec:markov} 

The location of this habitat, a major stronghold of the argonauts, is a closely guarded secret. Attempts to search it out have revealed only decoys or lifeless rocks. Though a great deal of information is available about the habitat's specs, operations, areas of research, and informational resources, only highly placed members of the argonauts may travel here. By all accounts, the habitat is a windowless beehive, designed to be virtually emissionless. Speculated locations include Pluto's moon Hydra, the deep Kuiper Belt, and even the Oort Cloud. 

\subsection{Extrasolar systems}
\label{sec:extrasolar-systems} 

Although travel between the stars is still outside the realm of transhumanity's achievements, the Pandora Gates have allowed passage to myriad other star systems. A few are noted here, though many more exist— not all of them explored. 

\subsubsection{Echo}
\label{sec:echo} 

Echo is a binary system consisting of a bright orange main sequence star and a pulsar (whence the system's name) about 12 light hours from one another. The system has one immense, bright yellow Jovian world (Echo VI) weighing in at 1.8 Jupiter masses and boasting 101 known moons, two Neptune-like ice giants further out, a thin mid-system asteroid belt, and several Mercury-like inner planets. 

%%% txt/112.txt 

The original Pandora Gate opens onto lifeless Echo V, a forbidding place littered with the detritus of a dead alien civilization. The hollow buildings of these precursors look out over once-verdant alluvial plains now home to only dry arroyos and dust. In other places, eons of wind erosion have carried the soil away entirely, leaving only barren expanses of dark basaltic slag. Chemically and geologically, the world is very similar to Mars, had Mars suffered another half a billion years’ loss of atmosphere. Research into the relics of the long-dead aliens suggests that they were morphologically similar to arthropods or arachnids, earning them the designation of Iktomi, after a Native American spider god. So far, little else has been discerned about them. 

Echo IV, on the other hand, is the closest thing transhumanity has found to a paradise since losing Earth. The native life is carbonbased, with many plants and fish edible even to flats. The climate is warm temperate, the atmosphere breathable with no major contaminants. The northern and southern latitudes are home to trackless forests dominated by various species of valders—huge, maple-like trees with dark red leaves. In the equatorial regions lie balmy, nutrient-rich floodplains ripe for cultivation, broken up by the occasional mountain range. Echo IV is still geologically active due to tidal heating, though older than Earth by about two billions years, and has two megacontinents connected near the equator by a tenuous land bridge. Notable native life include the Unagi, a fearsome, eel-like deep sea predator, and the clown sprite, a flying primate-analog that exists in a symbiotic relationship with the Echolalian land anenome, a huge, venomous, carnivorous plant that grows in the cloud forests of the equatorial highlands. The biosphere is diverse with many other megafauna, some quite dangerous. 

\subsubsection{Luca}
\label{sec:luca} 

Luca is an M-Class red dwarf located in a region of the galaxy far removed from any point of reference known to transhuman astronomy. The system has only a single gas giant of about 1.4 Jupiter masses—insufficient to shield the inner worlds from constant asteroid bombardment. The lone gas giant is flanked by a tenuous metallic inner asteroid belt and a wide ice and silicate outer belt. The only other bodies worthy of planetary status are a hellish inner world with Mercury’s richness of metal and Venus’ infernal atmosphere and a few large, sullen plutoids sharing Lagrange orbits with an asteroid field comprised of the shattered mass of a third plutoid. 

Accessible by both the Vulcanoid and Fissure Gates, Luca II is a heavily cratered terrestrial planet with a thick, dusty atmosphere— just about breathable to transhumans with the right gear. It is a cold, rocky world of craggy hills, knee-high forests, hissing geothermal bogs, and fungal meadows. The natives, who have been extinct for at least a million years, evolved from animals not unlike Earth’s aardvarks. Originally insect mound predators, the Lucans evolved vision well into the infrared (as demonstrated by the unusual pigments on their pottery and later-stage porcelain) and, based on analysis of their artifacts, had a sense akin to ultrasonic imaging. Their civilization went through several cycles of rise and fall, punctuated by celestial cataclysms that killed off less adaptable species and made resources scarce. The Lucans seem never to have evolved past medieval levels of societal organization prior to the Great Impact. Within a hundred years of that final impact, the last of the Lucans perished, never having invented the telescope, the computer, or space flight. 

Luca II hosts Banshee, an underground settlement with a few prominent surface features, including a radio astronomy station, park domes, a short-hop aerospace port, and solar farms. It is set on the Howling Plain, a windy plain of scrub brush and bogs chosen for its rich hydrocarbon deposits and low incidence of asteroid impacts. Banshee is an uneasy blend of anarchist colonists and hypercorp interests. 

\subsubsection{Mishipizheu}
\label{sec:mishipizheu} 

Mishipizheu is a red giant. The planet from which the star takes its name, Mishipizheu I, is a Mars-sized sphere of water with an atmosphere of nitrogen and carbon dioxide and a rocky core. Mishipizheu I was an almost Venus-sized sphere of ice 700 million years ago, but the expansion of its star into the red giant phase melted the planet. Initially quite warm and full of pockets of ice and carbonaceous silts, the melting planet was a crucible in which life could develop and now hosts a complex ecosystem. Amoeboid boiler reefs composed of gas sac creatures and their symbiotes bob on the surface or maintain neutral buoyancy in the depths, becoming platforms for complex ecosystems of largely animalian life. 

Mishipizheu I is orbited by a mid-size rocky moon, Nanabozho, reachable via the Discord Gate. Nanabozho is a mystery, as moons of this composition are not normally found so far out in a system. The best current theory is that Nanabozho was an inner system object with an erratic orbit. It was perturbed out of its orbit by one of the now-engulfed gas giants that must once have existed, whence by chance it was captured in Mishipizheu's orbit. The extraordinarily slim chances of such an event, however, have led to wild speculation as to the actual origin of the moon, which is as popular a destination for gatecrashers as the planet below. 

\subsubsection{Synergy}
\label{sec:synergy} 

Among the first attempts to establish a gatecrasher colony beyond the original Pandora Gate, just 5 years after the Fall, was a group of two hundred and fifty colonists equipped with experimental headware communications technology. Shortly after the jump, however, a still unidentified glitch forced the gate to close and the mechanism could not be reset to the same setting and coordinates for an entire five years. When the gate technicians finally managed to reacquire the settings recently and reopen the gate, the colonists were found to have survived, but they had changed. The technology sent with them was largely AI controlled, enabling the creation of a hypermesh that linked the thoughts, emotions, and sensory experiences of each colonist with each other. After half a decade of difficult survival measures, this technology and the stress of the situation linked the colonists and their AIs into a group mind. Despite having the opportunity to return to the solar system, these Synergists, as they call themselves, have no desire to cut themselves off from their shared consciousness. 

\subsubsection{Other Exoplanets}
\label{sec:other-exoplanets} 

The number of extraterrestrial star systems that transhumanity has visited via the gates now numbers into the hundreds, if not more—though only a small percentage of these have been notable and/or hospitable. Only a few dozen have been substantially occupied or colonized by transhumans, though this number is growing rapidly. Among these, a few deserve mention: 

\paragraph{Arcadia:}
\label{sec:arcadia} Accessed through the Martian Gate, the Planetary Consortium is constructing an aerostat in the upper atmosphere of this Venus-like planet which will serve as a private resort for the hyperelite. 

\paragraph{Babylon:}
\label{sec:babylon} Initially thought to just be an unremarkable scorched moon orbiting a planet very close to a yellow star, researchers measuring the star made an incredible accident discovery: what appears to be a derelict spacecraft orbiting deep in the star's corona. Attempts to access this vessel have so far been thwarted, but other projects are in the works, including the possibility of towing the craft to safer climes. 

\paragraph{Bluewood:}
\label{sec:bluewood} One of the first anarchist colonies established through the Fissure Gate, this settlement inhabits a beautiful, small Earth-like world with a thriving eco-system. Established on the outskirts of a large forest of eerie, alien, blue “trees,” the colony was taken off-guard by the trees alarming growth rate. The modular settlement buildings have all but been surrounded and encased by overgrowth despite modest efforts to keep them clear. Still intact but engulfed by spiraling branchworks, the effect is beautiful and haunting. 

\paragraph{Nótt:}
\label{sec:nott} This barren ice-covered moon suffers from heavy geothermal activity that causes its frozen crust to constantly crack and refreeze. The unfortunate research station staff here, all indentured, claim that something out in the ice is stalking them—over a dozen have disappeared in the last year. Pathfinder refuses to pull the station back, however, and thorough searches from its security teams have turned up nothing. 

\paragraph{Sky Ark:}
\label{sec:sky-ark} TerraGenesis is redesigning this dry, arid moonlet as an offworld preserve for animal life, including many formerly extinct Earth species resurrected from fossil DNA. 

\paragraph{Wormwood:}
\label{sec:wormwood} This maze-like warren seems to be an actual beehive habitat, though who tunneled it out or why remains an unanswered mystery. The former asteroid is part of the ring system of an unknown gas giant. Clearly artificial, gatecrashers so far have found no signs of technology or life. 

\begin{quotation}
\textbf{Analysis: Myst trees} 

 [File Corruption: 98\%] 

 [Partial Retrieval Complete] 

\begin{verbatim} ... called “myst trees” by the residen@# of Ca*\&78 ... also found on tw) oth*r exoplanets ]]]]] ... seem to be some sort of living data storage{{[— ... utilizing nanofog systems for <|{9h'''' ... high pr@bability of alien origin [[[[[[; ;
\end{verbatim}
\end{quotation} 

%%% Local Variables: %%% mode: latex %%% TeX-master: "ep" %%% End: 
