\chapter{Skills}
\label{chap:skills}

In a setting where physical looks and capabilities are
easily changed at the push of a button, who you are and
what you know is more important than any inborn ability.
Skills represent the knowledge your character has,
the accumulated set of experience, education, and inherent
know-how possessed by each and every sentient
transhuman in Eclipse Phase. They are what allow you
to sneak into a hypercorp station, disable the security
systems, hack the mesh hub, and then impersonate security
personnel to make your escape. Your skills represent
the one thing you have no matter what you look like or
where you find yourself. When your characters explore
what they can do, their skills, or lack thereof, often determine
the margin between success and failure.
Having a well-rounded set of skills is vital to survival
and success in Eclipse Phase. The skills below encompass
a wide selection of talents, enough so that each character
can be unique in their abilities and knowledge.

\section{Skill overview}
Skills are divided into aptitudes and learned skills (see
Character Skills, p. 123). Most (but not all) learned
skills are built on and linked to an aptitude. If a
character lacks the specific skill needed in a situation,
they may default to the linked aptitude. You may also
choose to specialize in certain skills (see Specializations,
p. 123), reflecting an enhanced knowledge of a
particular aspect of a certain skill.

\subsection{CORE SKILLS: APTITUDES}
Aptitudes represent inherent skills and abilities acquired
at birth or during the course of growing up. Aptitudes
are sometimes used for tests, but their primary use is
determining the starting point at which learned skills
are developed. Aptitudes determine the starting value
of their linked skills. For example, a character with
Somatics aptitude 10 who wishes to purchase points
in the Freerunning skill (which is linked to Somatics)
would start with a Freerunning rank of 10 and then
buy additionally points in that skill.
Aptitudes are also used when a character doesn’t
posses knowledge of a needed skill (see Defaulting, p.
116). Aptitudes represent the basic knowledge that a
character has acquired regarding rudimentary use of that
skill. They may not have ever received any formal training
with the skill, but they can still attempt to use it.
Aptitudes range in value from 1 to 30, with 10 being
the unaugmented human average and 15 representing
the average of most genetically modified transhumans.
Since aptitudes represent untrained ability, they are
capped at a maximum rating of 30.
There are seven different aptitudes that all players
possess. These aptitudes are purchased during character
creation (p. 128), but depending on the morph the
character is currently inhabiting, they may find their aptitudes
capped by the quality of the morph (see p. 124).

\subsubsection{LEARNED SKILLS}
A player’s learned skills are the most important part of
their character, representing the acquired knowledge
they carry with them from morph to morph, knowledge
that plays a fundamental role in helping define
the person’s ego. Learned skills encompass nearly any
skill that you might need to use in Eclipse Phase, and
they range in value from 0 to 99.
All learned skills have a linked aptitude that is
used to calculate their initial value, and which is
also defaulted to if the player does not have that
particular skill.


 \subsubsection{SKILL CATEGORIES}
Each learned skill is classified as either an Active skill
or a Knowledge skill. Active skills represent skills
that typically require physical actions and are used in
action scenes within game play. Knowledge skills are
more knowledge-based and intellectual, representing
ideas and facts. Knowledge skills may play a less dramatic role in certain action-oriented game play moments, but they flesh out the character’s background
and interests and are integral to roleplaying interactions. Active and Knowledge skills are purchased
separately during character creation.

Active skills are further divided into Combat,
Mental, Physical, Psi, Social, Technical, and Vehicle
skills. Certain traits and abilities may apply to specific
categories.


 \subsubsection{FIELD SKILLS}
Some learned skills are field skills, meaning that when
this skill is chosen a particular field of emphasis must
also be selected. For example, the skill of Academics
requires the character to specify a specific academic
discipline in which they are knowledgeable, such as
Biology, Chemistry, or Xenosociology. Field skills are
written as “[skill]: [field];” for example: “Art: Painting.” Field skills can be taken multiple times, choosing
a different area of emphasis each time, reflecting skills
in different fields; that is to say, each field is a separate
skill. Several suggested fields are listed for each field
skill, but gamemasters and players may also cooperate
to create others that fit their games.

Field skills may also have specializations; for example, Professional: Accounting (Money Laundering).


 \subsubsection{PSI SKILLS}
Psi refers to the ability to perceive and manipulate
biological minds via psi waves and/or other inexplicable phenomena. Due to the uniqueness of this ability,
characters that wish to wield psi must acquire the Psi

%%% txt/175.txt
trait (p. 147). Psi use also requires a number of specialized skills (Control, Psi Assault, and Sense) that reflect
special training characters acquire to tap into their psi
powers. Psi skills may not be defaulted on; the only
way to use a psi skill is to possess the trait along with
training in that skill. For more details, see Psi, p. 220.


 \subsubsection{SPECIALIZATIONS}
Any character may opt to specialize in a given skill
(see Specializations, p. 123). This specialization reflects
increased knowledge in one particular aspect of the
skill. Many of the skills offered below include sample
specializations. Gamemasters and players are encouraged to develop other specialization ideas together for
their campaigns.

Specialization provides a +10 modifier when
using that skill in a situation appropriate to that
specialization.




\section{USING SKILLS}
Whenever a character wants to do something using
a skill, they must succeed at a skill test (see Making
Tests, p. 115). The difficulty of the action is applied
as a modifier, as are any other extenuating circumstances that may affect the test (see Difficulty and
Modifiers, p. 115). As with other types of tests, all skill
tests are successful when the character rolls less than
or equal to the test’s target number after any modifiers have been applied. In the case of skill tests, the
target number is the character’s skill rating with that
particular skill. Modifiers representing difficulty and
other factors are applied directly to the target number
(see Difficulty and Modifiers, p. 115). A roll of a 00 is
always a success, regardless of modifiers, and a result
of 99 is always a failure, again despite any modifiers
that may increase a character’s target number over
100. Standard critical success and failure rules apply
to skill tests (see Criticals: Rolling Doubles, p. 116),
so any time a character rolls a double (i.e. 00, 11, 22,
33, etc.) they score a critical success or failure.


 \subsubsection{DEFAULTING}
Sometimes you lack the skill needed in a certain situation. In these instances, characters may default their
skill test to the linked aptitude. This reflects the fact
that most learned skills are developed from some sort
of baseline physical ability. Even though you may not
know how to do something, you’ve likely seen how
it’s done at some point or have some idea of how to
do it, or can at least take a shot at it. Naturally, you’re
not as good as someone who has training in that skill,
but it still allows you to make an attempt.

Not all skills can be defaulted. Some skills are
simply too complex or obscure, or demand special
knowledge or ability, for someone to attempt their
use untrained. For example, brain surgery or most psi
skills are simply beyond anyone who doesn’t have that
ability or the knowledge of what they’re attempting.

 \subsubsection{DEFAULTING TO FIELD SKILLS}
In some cases, a character may not possess the particular field skill that a test calls for, but they may be
skilled in another related field. For example, a test to
conduct an alien autopsy might call for an Academics:
Xenobiology roll, but a character who doesn’t have
that skill may be allowed to default to Academics:
Biology instead. The gamemaster decides if and when
to allow this, perhaps applying a modifier to the test
based on the difference between fields.


 \subsubsection{DEFAULTING TO RELATED SKILLS}
If the gamemaster allows it, characters may default to
a related skill that also has some relevance to the test
at hand. For example, a character skilled in Kinetic
Weapons might not be trained in the use of a laser, but
they know enough to point at the target and pull the
trigger. Likewise, a character might not be skilled in
Investigation, but the gamemaster could still allow
them to use their Perception skill instead in order to
realize that a body had been moved from the place
where it had been shot. In situations like this, when the
gamemaster allows defaulting to a related skill, a –30
modifier should be applied to the test.

\begin{quotation}
\textbf{Example}
\\
Srit is wandering through a black market souk on
Mars, trying to find a particular piece of sensory
equipment. The gamemaster calls for a Scrounging
Test, but Srit does not have that skill. She could
default her INT of 22, but instead she asks the
gamemaster if she can default to the related skill
of Perception, which she has at 82. The gamemaster agrees, and so Srit rolls against a target number
of 52 (82 – 30).
\end{quotation}




\subsubsection{COMPLEMENTARY SKILLS}
 Sometimes more than one skill may apply to a particular test, or knowledge in one area can aid your skill
 in another. In this case, the gamemaster may apply a
 modifier to the skill test based on the strength of the
 complementing skill, as noted on the Complementary
 Skill Bonus table.


\begin{quotation}
\textbf{Example}
\\
Dav is hoping to persuade a brinker pilot to take him
to an isolated habitat that doesn’t welcome visitors.
To impress upon the pilot that he is a friend of these
particular isolates, he calls on his knowledge of their
particular cultural practices (Interests: Religious
Cults skill at 45). The gamemaster allows this and
applies a +20 modifier to Dav’s Persuasion Test.
\end{quotation}

\begin{quotation}
\textbf{COMPLEMENTARY SKILL BONUS}
\\
\\
\begin{tabular}{cc}
SKILL RATING & MODIFIER \\
01-30 & +10 \\
31-60 & +20 \\
61+ & +30 \\
\end{tabular}
\end{quotation}


\begin{quotation}
\textbf{APTITUDE RANGE}
\\
\\
\begin{tabular}{lllll}

RATING & ASSESSMENT & SOMATICS & COORDINATION & REFLEXES \\
1–5 & child average & inept & clumsy & slow \\
6–10 & adult average & weak & able & paced \\
11–15 & transhuman average & fit & coordinated & swift \\
16–20 & enhanced & enhanced & agile & fast \\
21–25 & superhuman & gifted & nimble & lightning \\
26-30 & posthuman & elite & unerring & synaptic \\
%\end{tabular}
%\\
%\begin{tabular}{ccccccccc}
\\
RATING & COGNITION & INTUITION & SAVVY & WILLPOWER \\
1–5 & limited & aware & awkward & distracted \\
6–10 & intelligent & perceptive & personable & controlled \\
11–15 & bright & sharp & charismatic & focused \\
16–20 & learned & uncanny & dazzling & resolute \\
21–25 & brilliant & prescient & mesmerizing & unwavering \\
26-30 & genius & near omniscient & hypnotic & unshakable \\
\end{tabular}
\end{quotation}




\subsection{SKILL RANGES}
What is the difference between being a clumsy neophyte wobbling in zero gravity and being a veteran
gliding effortlessly through space as though you were
dancing? The answer is training and skill. The greater
your skill, the more likely you are to not only succeed
at what you want to do, but succeed well.

 Aptitudes in Eclipse Phase range from 1 to 30,
while learned skills range from 0 to 99. These numbers are an abstraction of the range of transhuman
abilities and traits. The Aptitude Range table provides
a breakdown of different aptitude levels and how they
relate to each other. Likewise, the Learned Skill Range
table provides an interpretation for the capabilities at
different skill levels.


\begin{quotation}
\textbf{LEARNED SKILL RANGES}
\\
\\
\begin{tabular}{cl}
SKILL & EQUIVALENCE \\
00 & No exposure or familiarity, completely unskilled \\
10 & Very rudimentary knowledge \\
20 & Basic operator’s proficiency (driver’s license, gun permit, high school diploma) \\
30 & Hands-on experience, some professional training \\
40 & Basic professional certification (police driving, army rifle certified, college diploma) \\
50 & Experience from professional-level work, some advanced training \\
60 & Expert competence (competitive driver, marksman, PhD) \\
70 & Experience from expert-level work, has had unique innovations or insights \\
80 & Worthy of being a system-renowned authority on the subject \\
90 & Nobel/Olympic/grandmaster \\
99 & Pinnacle of current understanding and innovation \\
\end{tabular}
\end{quotation}
% %%% txt/176.txt


\section{APTITUDES}
There are 7 aptitudes in Eclipse Phase, described on p.
123. Each character has these aptitudes at a minimum
rating of 1.

\subsection{APTITUDE-ONLY TESTS}
In rare cases, a test may call for using an aptitude
only, rather than a learned skill. This should only
occur when no learned skills are appropriate to the
test, and these circumstances are usually noted in
the rules.
Aptitude-only tests must be handled carefully, as
the range of aptitude ratings (1–30) is typically much
smaller than the rating of learned skills (0–99). For
this reason, most aptitude tests should use a target
number equal to the aptitude x 3. In rare cases where
the test is more difficult, the gamemaster may simply
use an aptitude x 2, or just the straight aptitude rating.
In some cases, more than one aptitude may be relevant
to the test, and so they may be added together
to derive the target number.
What follows are a few examples where an aptitudeonly
test might be appropriate. Gamemasters may call
for similar tests in other situations, but learned skills
should be used whenever possible.



\begin{quotation}
\textbf{APTITUDE COMPARISON: FLATS VS. SPLICERS AND EXALTS}
\\
Compared to humans in the early 21st Century, the average transhuman in the world of Eclipse Phase is faster,
smarter, stronger, and healthier than their unaugmented predecessors. Normal unaugmented humans, called
flats (p. 139), most closely approximate the type of person that was born in our time. The majority of people,
however, inhabit bodies that are known as splicers (p. 139) or exalts (p. 139) (well, those with biological bodies
anyway). Splicers are genefixed to avoid genetic defects and optimized for certain characteristics, while exalts
are tweaked to make them superior across the board: they are more attractive, more athletic, have greater
cognitive capacity, and are more attuned to the world around them than their unaugmented kin.
\end{quotation}


\section{COMPLETE SKILL LIST}



This section details all of the learned skills available
in Eclipse Phase. Gamemasters and players
may, of course, agree to add additional skills to
this list as appropriate to their campaign.

\begin{quotation}
\textbf{NECESSARY SKILLS}

While characters will need a mix of skills to
succeed in the varied tasks they encounter in
Eclipse Phase, some skills are crucial for any
character. If a character lacks these, they will
have a difficult time getting by, so it is important
for players and gamemasters to know
these particular skills.

\textbf{Fray}: Fray is the primary skill you use to avoid
getting hit in combat. Even if you plan to avoid
combat, being able to get out of the way when
necessary is a handy survival skill to have.

\textbf{Networking}: Unless you live in total isolation,
you need a Networking skill—preferably several.
Networking is how you interact with people in
a particular social circle to obtain information,
spread rumors, call in favors, and so on.

\textbf{Perception}: Perception Tests get called for
quite often, so if you want your character to
know what’s going on around them, make sure
to get this skill. Investigation and Scrounging
are also good, but Perception is king.
\end{quotation}


\begin{quotation}
\begin{tabular}{lll}
& \textbf{SKILL LIST} &\\
\\
SKILL &  LINKED APTITUDE  & CATEGORY\\
Academics: [Field] & COG & Knowledge\\
Animal Handling & SAV & Active, Social\\
Art: [Field] & INT & Knowledge\\
Beam Weapons & COO & Active, Combat\\
Blades & SOM & Active, Combat\\
Climbing & SOM & Active, Physical\\
Clubs & SOM & Active, Combat\\
Control & WIL (no defaulting) & Active, Mental, Psi\\
Deception & SAV & Active, Social\\
Demolitions & COG (no defaulting) & Active, Technical\\
Disguise & INT & Active, Physical\\
Exotic Melee Weapon: [Field] & SOM & Active, Combat\\
Exotic Ranged Weapon: [Field] & COO & Active, Combat\\
Flight & SOM & Active, Physical\\
Fray & REF & Active, Combat\\
Free Fall & REF & Active, Physical\\
Freerunning & SOM & Active, Physical\\
Gunnery & INT & Active, Combat\\
Hardware: [Field] & COG & Active, Technical\\
Impersonation & SAV & Active, Social\\
Infiltration & COO & Active, Physical\\
Infosec & COG (no defaulting) & Active, Technical\\
Interest: [Field] & COG & Knowledge\\
Interfacing & COG & Active, Technical\\
Intimidation & SAV & Active, Social\\
Investigation & INT & Active, Mental\\
Kinesics & SAV & Active, Social\\
Kinetic Weapons & COO & Active, Combat\\
Language: [Field] & INT & Knowledge\\
Medicine: [Field] & COG & Active, Technical\\
Navigation & INT & Active, Mental\\
Networking: [Field] & SAV & Active, Social\\
Palming & COO & Active, Physical\\
Perception & INT & Active, Mental\\
Persuasion & SAV & Active, Social\\
Pilot: [Field] & REF & Active, Vehicle\\
Profession: [Field] & COG & Knowledge\\
Programming & COG (no defaulting) & Active, Technical\\
Protocol & SAV & Active, Social\\
Psi Assault & WIL (no defaulting) & Active, Mental, Psi\\
Psychosurgery & INT & Active, Technical\\
Research & COG & Active, Technical\\
Scrounging & INT & Active, Mental\\
Seeker Weapons & COO & Active, Combat\\
Sense & INT (no defaulting) & Active, Mental, Psi\\
Spray Weapons & COO & Active, Combat\\
Swimming & SOM & Active, Physical\\
Throwing Weapons & COO & Active, Combat\\
Unarmed Combat & SOM & Active, Combat\\
\end{tabular}
\end{quotation}

\subsubsection{ACADEMICS: [FIELD]}
\textbf{Type:} Field, Knowledge
\\ \textbf{Linked Aptitude:} COG
\\ \textbf{What it is:}
Academics covers any sort of specialized
non-applied knowledge you can only get
through intensive education. Most theoretical and
applied sciences, social sciences, transhumanities,
etc. are covered by this skill. Most of the other
skills listed in this chapter could also be taken as
an Academics field, reflecting a working theoretical
knowledge of the skill—for example, Academics:
Armorer or Academics: Interrogation.
\\ \textbf{When you use it:} Academics is used when a
character wishes to call upon a specific body of
knowledge. For example, Academics: Chemistry
could be used to identify a particular substance,
understand an unusual chemical reaction, or determine
what elements are needed to nanofabricate
something that requires exotic materials. At the
gamemaster’s discretion, some Academics-related
tests might not be defaultable, given that only
someone who has been educated in that subject is
likely to be able to tackle it.
\\ \textbf{Sample Fields:} Archeology, Astrobiology, Astronomy,
Astrophysics, Astrosociology, Biochemistry,
Biology, Botany, Computer Science,
Cryptography, Economics, Engineering, Genetics,
Geology, Linguistics, Mathematics, Memetics,
Nanotechnology, Old Earth History, Physics, Political Science, Psychology, Sociology, Xeno-archeology,
Xenolinguistics, Zoology
\\ \textbf{Specializations:} As appropriate to the field

\subsubsection{ANIMAL HANDLING}
\textbf{Type:} Active, Social
\\ \textbf{Linked Aptitude:} SAV
\\ \textbf{What it is:} Skilled animal handlers are able to
train and control a wide variety of natural and transgenic animals, including partial uplifts. Though many
animal species went extinct during the Fall, a few
“ark” and zoo habitats keep some species alive, and
many others can be resurrected from genetic samples.
Exotic animals are considered a sign of prestige
among the hypercorp elites, and guard animals are
occasionally used to protect high-security installations.
Likewise, many habitats and settlements employ small
armies of partially uplifted, genetically modified, and
behavior-controlled creatures for sanitation or other
purposes. Many new and strange breeds of animal are
created daily to serve a variety of roles.
\\ \textbf{When you use it:} Animal Handling is used whenever
you are trying to manipulate an animal, whether your
intent is to calm it down, keep it from attacking, intimidate it, acquire its trust, or goad it into attacking.
Your Margin of Success determines how effective you
are at convincing the creature. At the gamemaster’s
discretion, modifiers may be applied to the test. Likewise, winning an animal over may sometimes take
time, and so could be handled as a Task Action with a
timeframe of five minutes or more.
\\ \textbf{Specializations:} Per animal species (dogs, horses, smart
rats, etc.)

\begin{quotation}
\textbf{TRAINING ANIMALS}
\\
Training animals is a time-consuming task requiring
repeated efforts and rewards to reinforce the trained
behavior. Treat this as a Task Action with a timeframe
of one day to one month, depending on the complexity of the action. Apply modifiers to this test based on
the relative intelligence of the animal being trained,
how domestic it is, and the complexity of the task.
Once an animal has been trained, commanding
it is treated as a Simple Success Test (p. 118) except
for unusual or stressful situations, in which case the
trainer receives a +30 modifier on their Animal Handling Tests when convincing the animal to complete
the trained action.
\end{quotation}

\subsubsection{ART: [FIELD]}
\textbf{Type:} Field, Knowledge
\\ \textbf{Linked Aptitude:} INT
\\ \textbf{What it is:} Art confers the ability to create and evaluate artistic endeavors. This is a particularly useful
skill in Eclipse Phase, especially in the post-scarcity
economies where creativity and vision can be a key
component to a character’s reputation.
\\ \textbf{When you use it:} The Art skill can be used to either
create a new work of art or to duplicate an existing
piece of art in the hopes of passing it off as your own.
The skill can also determine the approximate value of
a piece of art either on the open market, for monetary
exchange systems, or in terms of reputation for the
artist.
\\ \textbf{Sample Fields:} Architecture, Criticism, Dance, Drama,
Drawing, Painting, Performance, Sculpture, Simulspace Design, Singing, Speech, Writing
\\ \textbf{Specializations:} As appropriate to the field

\subsubsection{BEAM WEAPONS}
\textbf{Type:} Active, Combat
\\ \textbf{Linked Aptitude:} COO
\\ \textbf{What it is:} The Beam Weapons skill covers the usage
and maintenance of standard coherent beam energy
weapons such as lasers, particle beam weapons,
plasma rifles, and microwave weapons (p. 338).
\\ \textbf{When you use it:} A player uses their Beam Weapons skill whenever attacking with a beam weapon in
combat (p. 191). Beam Weapons may also be used for
tests involving maintenance of the weapon, but not
for repairing or modifying the weapon (that would be
Hardware: Armorer skill).
\\ \textbf{Specializations:} Lasers, Microwave Weapons, Particle
Beam Weapons, Plasma Rifles

\subsubsection{BLADES}
\textbf{Type:} Active, Combat
\\ \textbf{Linked Aptitude:} SOM
\\ \textbf{What it is:} The Blades skill covers the usage and
maintenance of standard bladed weapons (p. 334).
\\ \textbf{When you use it:} A player uses their Blades skill
whenever attacking with a blade weapon in melee
combat (p. 191). Blades may also be used for tests
involving maintenance of the weapon, but not for
repairing or modifying the weapon (that would be
Hardware: Armorer skill). This skill is used for blade
weapons implanted in the body at the end of an appendage (hands, forearms, feet, octomorph arms, etc.),
but the Exotic Melee Weapon skill is used for blades
implanted in other parts of the body.
\\ \textbf{Specializations:} Axes, Implant Blades, Knives, Swords

\subsubsection{CLIMBING}
\textbf{Type:} Active, Physical
\\ \textbf{Linked Aptitude:} SOM
\\ \textbf{What it is:} Climbing is the skill of ascending and
descending sheer surfaces with or without the aid of
specialized equipment.
\\ \textbf{When you use it:} This skill is used whenever a character wishes to scale a climbable surface. For heights
greater than one story, climbing is handled as a Task
Action with a timeframe equivalent to one meter per
Action Phase. For rappelling, the timeframe for descent is 50 meters per Action Turn. Climbing gear (p.
332-333) provides appropriate modifiers.
\\ \textbf{Specializations:} Assisted, Freehand, Rappelling
%%% txt/180.txt

\subsubsection{CLUBS}
\textbf{Type:} Active, Combat
\\ \textbf{Linked Aptitude:} SOM
\\ \textbf{What it is:} The Clubs skill covers the usage and
maintenance of standard blunt melee weapons such
as batons or sticks (see p. 334).
\\ \textbf{When you use it:} Players use their Clubs skill whenever they want to attack with a blunt weapon in melee
combat (p. 191). The Clubs skill may also be used for
tests involving maintenance of the weapon, but not
for repairing or modifying the weapon (that would be
Hardware: Armorer skill).
\\ \textbf{Specializations:} Batons, Hammers, Staffs

\subsubsection{CONTROL}
\textbf{Type:} Active, Mental, Psi
\\ \textbf{Linked Aptitude:} WIL
\\ \textbf{What it is:} Control is the use of psi to manipulate
individuals or actively penetrate their mental processes. This skill is only available to characters with
the Psi trait (p. 147).
\\ \textbf{When you use it:} Use Control when taking a psionic
tour through a foreign ego—messing around included.
See Psi, p. 220.
\\ \textbf{Specializations:} By sleight

\subsubsection{DECEPTION}
\textbf{Type:} Active, Social
\\ \textbf{Linked Aptitude:} SAV
\\ \textbf{What it is:} Deception is your ability to act, bluff,
con, fast talk, lie, misrepresent, and pretend. Accomplished users of deception are able to convince anyone
of nearly anything. This skill does not include using a
physical disguise to appear to be another person (the
Impersonate skill covers that area).
\\ \textbf{When you use it:} Use this skill whenever you want
to deceive someone with words or gestures. A successful skill test means that you have passed off your deception convincingly. At the gamemaster’s discretion,
someone who is actively alert for signs of deception
may make an Opposed Test using the Kinesics skill.
\\ \textbf{Specializations:} Acting, Bluffing, Fast Talk

\subsubsection{DEMOLITIONS}
\textbf{Type:} Active, Technical
\\ \textbf{Linked Aptitude:} COG (no defaulting)
\\ \textbf{What it is:} Demolitions covers the use of controlled
explosives.
\\ \textbf{When you use it:} Use it when making, placing, and
disarming explosives and explosive devices. See Demolitions, p. 197.
\\ \textbf{Specializations:} Commercial Explosives, Disarming,
Improvised Explosives

\subsubsection{DISGUISE}
\textbf{Type:} Active, Physical
\\ \textbf{Linked Aptitude:} INT
\\ \textbf{What it is:} Disguise is the art of physically altering
your appearance so that you look like someone else.
This includes both the use of props (wigs, contacts,
skin pigments) and the altering of subtle physical
characteristics (gait, posture, poise).
\\ \textbf{When you use it:} Use Disguise to fool someone into
thinking you’re someone you’re not. This can be used
to hide your identity or to make yourself look like
someone in particular. When used against someone
who knows your true look or the appearance of the
person you are imitating, this is handled as an Opposed Test against Perception or Investigation.
\\ \textbf{Specializations:} Cosmetic, Theatrical

\subsubsection{EXOTIC MELEE WEAPON: [FIELD]}
\textbf{Type:} Field, Active, Combat
\\ \textbf{Linked Aptitude:} SOM
\\ \textbf{What it is:} The Exotic Melee Weapon skill covers
the use and maintenance of all melee weapons not
covered by the Clubs or Blades skills (see p. 334).
\\ \textbf{When you use it:} Use the Exotic Melee Weapon skill
when attacking someone with an exotic melee weapon
in melee combat (p. 191).
\\ \textbf{Sample Fields:} Morning Star, Spear, Whip
\\ \textbf{Specializations:} N/A

\subsubsection{EXOTIC RANGED WEAPON: [FIELD]}
\textbf{Type:} Field, Active, Combat
\\ \textbf{Linked Aptitude:} COO
\\ \textbf{What it is:} Exotic Ranged Weapon skill includes
the use and maintenance of all ranged weapons not
covered by the Beam, Flechette, Kinetic, Sonic, or
Throwing Weapons skills.
\\ \textbf{When you use it:} Use this skill whenever attacking with
an exotic ranged weapon in ranged combat (p. 191).
\\ \textbf{Sample Fields:} Blowgun, Crossbow, Flamethrower,
Slingshot
\\ \textbf{Specializations:} N/A

\subsubsection{FLIGHT}
\textbf{Type:} Active, Physical
\\ \textbf{Linked Aptitude:} SOM
\\ \textbf{What it is:} Flight is the skill of using your body to
fly. This skill is used when sleeved in or jamming a
winged or otherwise flight-capable morph (manual and
remote-control flight are handled using Pilot skill).
\\ \textbf{When you use it:} Use this skill whenever you need to
make an aerial maneuver, land in difficult conditions,
maintain your course in steep winds, or otherwise
keep from crashing or falling.
\\ \textbf{Specializations:} Diving, Landing, Takeoff, specific
maneuvers

\subsubsection{FRAY}
\textbf{Type:} Active, Combat
\\ \textbf{Linked Aptitude:} REF
\\ \textbf{What it is:} Fray is the ability to get out of the way
of incoming attacks, debris, or inconvenient passers-by.
Characters that have a high Fray score are able to react
quicker than others when dodging or maneuvering.
\\ \textbf{When you use it:} Whenever a character is physically
attacked by an opponent in melee combat, roll Fray to
avoid getting hit (see p. 191). Fray may also be used
%%% txt/181.txt
to dodge other events that may harm the character,
such as avoiding a charging vehicle or jumping out of
the way of a collapsing stack of crates.
\\ \textbf{Specializations:} Blades, Clubs, Full Defense, Unarmed

\subsubsection{FREE FALL}
\textbf{Type:} Active, Physical
\\ \textbf{Linked Aptitude:} REF
\\ \textbf{What it is:} Free Fall is about moving in free-fall and
microgravity environments.
\\ \textbf{When you use it:} Use whenever you need to maneuver in a zero-g situation, such as propelling yourself
across a large open space or making sure you don’t
accidentally send yourself spinning off into space.
Free Fall is also used when moving with spacesuit
maneuvering jets and when parachuting.
\\ \textbf{Specializations:} Microgravity, Parachuting, Vacsuits

\subsubsection{FREERUNNING}
\textbf{Type:} Active, Physical
\\ \textbf{Linked Aptitude:} SOM
\\ \textbf{What it is:} Freerunning is part running, part gymnastics. It is about moving fast, maneuvering over/under/
around/through obstacles, and placing your body
where it needs to go. Freerunning/parkour is a popular
pastime in habitats where open space is limited.
\\ \textbf{When you use it:} Use Freerunning whenever you
need to overcome an obstacle via movement, such
as hurdling a railing, rolling across the hood of a
car, jumping across a pit, or swinging around a pole.
Freerunning is also used for sprinting (p. 191) and full
defense against attacks (p. 198).
\\ \textbf{Specializations:} Balance, Gymnastics, Jumping, Running

\subsubsection{GUNNERY}
\textbf{Type:} Active, Combat
\\ \textbf{Linked Aptitude:} INT
\\ \textbf{What it is:} Gunnery skill covers the use and maintenance of large, vehicular, or non-portable weapons
systems. Firing these weapons is more like playing a
video game than firing a gun.
\\ \textbf{When you use it:} Use Gunnery when attacking with
a vehicle-mounted weapon or weapon emplacement
in ranged combat (p. 191).
\\ \textbf{Specializations:} Artillery, Missiles

\subsubsection{HARDWARE: [FIELD]}
\textbf{Type:} Field, Active, Technical
\\ \textbf{Linked Aptitude:} COG
\\ \textbf{What it is:} This skill encompasses the ability to
build, repair, physically hack, and upgrade equipment
of a specific type.
\\ \textbf{When you use it:} Hardware is primarily used to repair
devices, vehicles, habitat systems, or synthetic morphs.
See Building, Repairing, and Modifying below.
\\ \textbf{Sample Fields:} Aerospace (all air and space vehicles),
Armorer (armor and weapons), Electronics (all
computerized devices), Groundcraft, Implants, Industrial (habitat, factory, and life support systems),
Nautical (watercraft and submarines), Robotics
(synthetic morphs)
\\ \textbf{Specializations:} As appropriate to the field


\begin{quotation}
\textbf{Building}
\\
Creating an item from scratch is handled as a Task
Action with a timeframe determined by the gamemaster. The timeframe should be set according to the complexity of the object and could range from an hour
(constructing a set of shelves) to days (assembling a
robot from spare parts) to even months (building a
house). Numerous factors may apply modifiers to
the test, such as the use of entoptic blueprints/help
manuals (+20) or poor working conditions (–10 to
–30). Tools are also a factor, perhaps making the job
easier (superior tools +10 to +30), more difficult (poor
or inadequate tools, –10 to –30), or even impossible
(lack of required tools).
\end{quotation}

\begin{quotation}
\textbf{Repairing}
\\
Damaged items may be repaired in a similar manner. See
the rules for Synthmorph and Object Repair, p. 209.
\end{quotation}


\begin{quotation}
\textbf{Modifying}
\\
Altering an object’s design and function follows the
same basic rules as build and repair, above. The time-frame is determined by the gamemaster as appropriate
to the modification.
\end{quotation}



%%% txt/182.txt

\subsubsection{IMPERSONATION}
\textbf{Type:} Active, Social
\\ \textbf{Linked Aptitude:} SAV
\\ \textbf{What it is:} Impersonation is the skill of trying to
pass yourself off as someone else in social situations,
including virtual ones. This includes copying mannerisms and speech patterns and using accumulated information to convince others that you are that person.
In a universe where appearance is highly variable, the
question of identity is largely one of both trust and
picking up on behavioral quirks and verbal cues to
recognize a given individual.
\\ \textbf{When you use it:} Sometimes it’s fun to pretend
you’re someone else, and sometimes it’s profitable
or lifesaving. Use this skill whenever you attempt
to convince someone that you are actually someone
else through some sort of social or online interaction.
Forks use this skill when passing themselves off as
their alpha ego. Impersonate is handled as an Opposed Test against the Kinesics skill.
\\ \textbf{Specializations:} Avatar, Face-to-Face, Verbal

\subsubsection{INFILTRATION}
\textbf{Type:} Active, Physical
\\ \textbf{Linked Aptitude:} COO
\\ \textbf{What it is:} Infiltration is the art of escaping
detection.
\\ \textbf{When you use it:} Use Infiltration whenever you
need to physically hide or move with stealth to avoid
someone sensing you, whether you are hiding behind
a tree, sneaking past a guard, or blending into a
crowd. Infiltration can also be used to follow people
(shadowing) without them detecting you. Infiltration
is an Opposed Test against the Perception of whomever you are hiding from. The gamemaster may wish
to roll such tests in secret so the player does not know
whether they have succeeded or failed.
\\ \textbf{Specializations:} Blending In, Hiding, Shadowing,
Sneaking

\subsubsection{INFOSEC}
\textbf{Type:} Active, Technical
\\ \textbf{Linked Aptitude:} COG (no defaulting)
\\ \textbf{What it is:} Infosec is short for “information security.” It encompasses training in electronic intrusion
and counterintrusion techniques, as well as encryption
and decryption.
\\ \textbf{When you use it:} Infosec is used both for hacking
into electronic devices and mesh networks and for
protecting them. See the Mesh chapter, p. 234, for
more details.
\\ \textbf{Specializations:} Brute-Force Hacking, Decryption,
Probing, Security, Sniffing, Spoofing

\subsubsection{INTEREST: [FIELD]}
\textbf{Type:} Field, Knowledge
\\ \textbf{Linked Aptitude:} COG
\\ \textbf{What it is:} Interest includes just about any topic
that captures your attention that isn’t covered by
another skill. This includes hobbies, obsessions,
causes, pastimes, and other recreational pursuits.
\\ \textbf{When you use it:} Use the Interest skill whenever
you need to recall or use knowledge related to the
particular interest in question.
Field Examples: Ancient Sports, Celebrity Gossip,
Conspiracies, Factor Trivia, Gambling, Hypercorp
Politics, Lunar Habitats, Martian Beers, Old Earth
Nation-States, Reclaimer Blogs, Science Fiction, Scum
Drug Dealers, Spaceship Models, Triad Economics,
Underground XP
\\ \textbf{Specializations:} As appropriate to the field

\subsubsection{INTERFACING}
\textbf{Type:} Active, Technical
\\ \textbf{Linked Aptitude:} COG
\\ \textbf{What it is:} Interfacing is about using computerized
electronic devices and software.
\\ \textbf{When you use it:} Use Interfacing to understand an
electronic device you are not familiar with, use a program according to its normal operating parameters,
manipulate electronic files of various types (including
images, video, XP, and audio files), scan for wireless
devices, and otherwise interact with and command
your ecto, muse, and other computerized devices.
Some Interfacing actions may be Task Actions, with
a timeframe determined by the gamemaster. For more
detail, see the Mesh chapter, p. 234.
\\ \textbf{Specializations:} Forgery, Scanning, Stealthing, by
program

\subsubsection{INTIMIDATION}
\textbf{Type:} Active, Social
\\ \textbf{Linked Aptitude:} SAV
\\ \textbf{What it is:} Intimidation is convincing someone to
do what you want based on direct threats (implied or
actual) or sheer force of personality.
\\ \textbf{When you use it:} Use Intimidation to scare someone
into submission, browbeat them into getting your way,
command them to follow your orders, or berate them
into giving up information. Influence is handled as an
Opposed Test, pitted against the target’s WIL + WIL
+ SAV.
\\ \textbf{Specializations:} Interrogation, Physical, Verbal

\subsubsection{INVESTIGATION}
\textbf{Type:} Active, Mental
\\ \textbf{Linked Aptitude:} INT
\\ \textbf{What it is:} Investigation is the art of analyzing
evidence, piecing together clues, solving mysteries,
and making logical deductions from groups of facts.
Investigation differs from Perception in that it is the
conscious search for clues or pieces of a puzzle.
\\ \textbf{When you use it:} Use Investigation to draw conclusions from assorted details. For example, Investigation could be used to determine the likely sequence
of events at a crime scene, determine a possible social
connection between two people, or deduce how an
enemy made their escape. Investigation is a great way
%%% txt/183.txt
to provide clues to players, especially when the subject
matter is something their character might know well
but the player does not.
\\ \textbf{Specializations:} Evidence Analysis, Logical Deductions,
Physical Investigation, Physical Tracking

\subsubsection{KINESICS}
\textbf{Type:} Active, Social
\\ \textbf{Linked Aptitude:} SAV
\\ \textbf{What it is:} Kinesics is the art of empathy and non-vocal communication.
\\ \textbf{When you use it:} Use Kinesics to read body language,
tells, social cues, and other subconscious indicators. It
can also be used to emote more effectively. Kinesics
is used defensively whenever someone is trying to deceive you; make an Opposed Test against that person’s
Deception or Impersonation skill.
Though synthetic morphs are also designed to
emote, reading them is not as easy. Apply a –30 modifier when judging a synthetic morph inhabited by a
character or AGI. Likewise, standard AIs are also difficult to read; apply a –60 modifier when judging a
synthetic morph or pod operated by an AI.
\\ \textbf{Specializations:} Judge Intent, Nonvocal Communication

\begin{quotation}
\textbf{JUDGING EMOTIONS AND INTENTIONS}
\\
Transhumans are empathic beings, and so you can
attempt to gauge the demeanor and/or intent of
someone you are dealing with by rolling a Kinesics
Test. This attempt to read someone is far from exact,
however, and it is easy to misjudge. The gamemaster
should make this test in secret and only allow a hint
if successful—it is not possible to read someone with
absolute certainty. If the person being judged is intentionally trying to deceive the character, this should be
an Opposed Test against their Deception skill.
\end{quotation}

\begin{quotation}
\textbf{NONVOCAL COMMUNICATION}
\\
Experts in Kinesics can effectively communicate
with each other simply by posture, stances, gestures,
demeanors, and looks. Such communication is necessarily limited in the amount of information it can
convey, but feelings, attitudes, affirmation/negation,
and simple concepts may be passed. To effectively
communicate complex concepts, the gamemaster may
require successful Kinesics Tests from both parties, applying modifiers as appropriate.
\end{quotation}

\subsubsection{KINETIC WEAPONS}
\textbf{Type:} Active, Combat
\\ \textbf{Linked Aptitude:} COO
\\ \textbf{What it is:} Kinetic Weapons covers the use and
maintenance of standard kinetic projectile weapons
like firearms and railguns (p. 335).
\\ \textbf{When you use it:} Use this skill whenever attacking
with a kinetic weapon in ranged combat (p. 191).
\\ \textbf{Specializations:} Assault Rifles, Machine Guns, Pistols,
Sniper Rifles, Submachine Guns


\subsubsection{LANGUAGE: [FIELD]}
\textbf{Type:} Field, Knowledge
\\ \textbf{Linked Aptitude:} INT
\\ \textbf{What it is:} Language covers the speaking and
reading of languages other than the player’s native
tongue. A speaker is considered fluent at a skill level
of 50; anything above this indicates further refinement in technical vocabulary, accents, and knowledge of dialects.
\\ \textbf{When you use it:} Use the Language skill whenever
you want to speak, understand, or read something in
a language at which you are skilled. Most speaking
and reading comprehension tests can be considered
Simple Success Tests if your skill is over 50, unless the
gamemaster rules the subject is sufficiently complex
that a non-native speaker would have trouble understanding it.
\\ \textbf{Sample Fields:} Arabic, Cantonese, English, French,
Hindi, Japanese, Mandarin, Portuguese, Russian,
Spanish
\\ \textbf{Specializations:} As appropriate to the field, representing dialects, technical jargon, and subcultural slang

\begin{quotation}
\textbf{LANGUAGES IN ECLIPSE PHASE}
\\
With the Fall of Earth, the languages that
remain most prominent in the solar system are
those that were extensively carried into space
by countries and hypercorps with aggressive
space programs or by the large populations
of poor laborers and infomorph refugees that
followed. No single language dominated the
realm of space expansion, and multilingualism was common. Many habitats and (sub)
cultural groupings cling to specific languages
as a method of retaining cultural identity. Despite the availability of instant translation via
the mesh, many people remain versed in two or
more languages.
The ten most common languages in the solar
system by speaking populations are: Arabic,
Cantonese, English, French, Hindi, Japanese,
Mandarin, Portuguese, Russian, and Spanish.
Other languages that remain strong include
Bengali, Dutch, Farsi, German, Italian, Javanese,
Korean, Polish, Punjabi, Swedish, Tamil, Turkish,
Urdu, Vietnamese, and Wu. Some languages
were effectively lost during the Fall, especially
those in some undeveloped regions, as their
speaking populations did not migrate into space
pre-Fall and were not privileged enough to survive in large numbers as infomorph refugees.
\end{quotation}

%%% txt/184.txt

\subsubsection{MEDICINE: [FIELD]}
\textbf{Type:} Field, Active, Technical
\\ \textbf{Linked Aptitude:} COG
\\ \textbf{What it is:} Medicine is the applied care and maintenance of biological beings and life.
\\ \textbf{When you use it:} Use Medicine whenever you need
to apply medical care beyond the immediate help
provided by first responders. This includes conducting physical exams, diagnosing ailments, treating
problems and illnesses, surgery, using biotech and
nanotech medical tools, and long-term care. See Healing and Repair, p. 208.
\\ \textbf{Sample Fields:} Biosculpting, Exotic Biomorphs,
Gene Therapy, General Practice, Implant Surgery,
Nanomedicine, Mercurials (by type), Paramedic,
Pods, Psychiatry, Remote Surgery, Trauma Surgery,
Veterinary
\\ \textbf{Specializations:} As appropriate to the field

\subsubsection{NAVIGATION}
\textbf{Type:} Active, Mental
\\ \textbf{Linked Aptitude:} INT
\\ \textbf{What it is:} Navigation is the art of finding your way,
whether using AR maps, a compass, the stars, or an
astrogation AI.
\\ \textbf{When you use it:} Use Navigation whenever you
need to plot out a course, determine a direction, or
otherwise keep from getting lost.
\\ \textbf{Specializations:} Astrogation, Map Making, Map
Reading

\subsubsection{NETWORKING: [FIELD]}
\textbf{Type:} Active, Social
\\ \textbf{Linked Aptitude:} SAV
\\ \textbf{What it is:} Networking is your skill at working your
contacts, trading favors, and keeping your finger on
the pulse of a particular faction or cultural grouping.
\\ \textbf{When you use it:} Use Networking to gather information or call on services using your Reputation (see
Reputation and Social Networks, p. 285).
\\ \textbf{Sample Fields:} Autonomists (@-rep), Criminals (g-rep),
Ecologists (e-rep), Firewall (i-rep), Hypercorps (c-rep), Media (f-rep), Scientists (r-rep). At the game-master’s discretion, this list can be expanded to
other (sub)cultural groupings.
\\ \textbf{Specializations:} As appropriate to each field

\subsubsection{PALMING}
\textbf{Type:} Active, Physical
\\ \textbf{Linked Aptitude:} COO
\\ \textbf{What it is:} Palming is the skill of handling items quickly and nimbly without others noticing. Palming is not
only about dexterous manipulation of objects but also
relies heavily on obfuscation, timing, and misdirection.
\\ \textbf{When you use it:} Use Palming any time you are
trying to conceal an item on your person, shoplift,
pick a pocket, surreptitiously discard something, or
perform a magic trick. Palming is an Opposed Test
against the Perception of any onlookers. The game-master may wish to make this roll secretly.
\\ \textbf{Specializations:} Pickpocketing, Shoplifting, Tricks

\subsubsection{PERCEPTION}
\textbf{Type:} Active, Mental
\\ \textbf{Linked Aptitude:} INT
\\ \textbf{What it is:} Perception is the use of your physical
senses (including cybernetic) and awareness of the
physical world around you. Perception differs from
Investigation in that it is noticing things by chance,
rather than actively searching for something.
\\ \textbf{When you use it:} Use Perception whenever you
wanted to take a detailed account of your surroundtion) or by lying (covered by Deception).
\\ \textbf{When you use it:} Use Persuasion any time you are
trying to bargain with, convince, or manipulate someone. This can include motivating your subordinates or
peers to take action, seducing a companion, winning
a political debate, or negotiating a contract, among
other things. Persuasion is handled as an Opposed
Test against the target’s WIL + WIL + SAV when one
person is simply trying to win over another. If both
parties are trying to convince each other, make it an
Opposed Test between Persuasion skills.
\\ \textbf{Specializations:} Diplomacy, Morale Boosting, Negotiating, Seduction

\subsubsection{PILOT: [FIELD]}
\textbf{Type:} Field, Active, Vehicle
\\ \textbf{Linked Aptitude:} REF
\\ \textbf{What it is:} Pilot is your skill at driving/flying a vehicle of a particular type.
\\ \textbf{When you use it:} You use Pilot skill whenever you
need to maneuver, control, or avoid crashing a vehicle,
whether you are in the pilot’s seat, remote controlling
a robot, or directly jamming a vehicle with VR. Each
vehicle has a Handling modifier that applies to this
test, along with other situational modifiers (see Bots,
Synthmorphs, and Vehicles, p. 195).
\\ \textbf{Sample Fields:} Aircraft, Anthroform (walkers), Exotic
Vehicle, Groundcraft (wheeled or tracked), Spacecraft, Watercraft
\\ \textbf{Specializations:} As appropriate to the field

\subsubsection{PROFESSION: [FIELD]}
\textbf{Type:} Field, Knowledge
\\ \textbf{Linked Aptitude:} COG
\\ \textbf{What it is:} Profession skills indicate training in a
profession practiced in Eclipse Phase. This can indicate either formal training or informal, on-the-job type
training, and includes both legal and extralegal trades.
\\ \textbf{When you use it:} Use Profession to perform work-related tasks for a specific trade (i.e. mining, balancing
accounts, designing a security system, etc.) or to reference specialized knowledge that someone trained in
that profession might have.
\\ \textbf{Sample Fields:} Accounting, Appraisal, Asteroid Prospecting, Banking, Cool Hunting, Con Schemes,
Distribution, Forensics, Lab Technician, Mining,
Police Procedures, Psychotherapy, Security Ops,
Smuggling Tricks, Social Engineering, Squad Tactics,
Viral Marketing, XP Production
\\ \textbf{Specializations:} As appropriate to the field

\subsubsection{PROGRAMMING}
\textbf{Type:} Active, Technical
\\ \textbf{Linked Aptitude:} COG (no defaulting)
\\ \textbf{What it is:} Programming is your talent at writing
and modifying software code.
\\ \textbf{When you use it:} Use Programming to write new
programs, modify or patch existing software, break
copy protection, find or introduce exploitable flaws,
write virii or worms, design virtual settings, and so
on. See the Mesh chapter, p. 234. Programming is also
applied when using nanofabrication devices.
\\ \textbf{Specializations:} AI Code, Malware, Nanofabrication,
Piracy, Simulspace Code

\subsubsection{NANOFABRICATION}
Nanofabrication is use of Programming skill to create
objects using a cornucopia machine, fabber, or maker
(p. 327). If you have appropriate blueprints and raw
materials, most uses of a nanofabricator can be treated
as a Simple Success Test (p. 118). If you wish to create
an item for which you do not have blueprints or the
proper raw materials, however, or you wish to alter an
item’s design, then a Nanofabrication Test is called for.
See Nanofabrication, p. 284.
\\ \textbf{Specializations:} Art, Clothing, Electronics, Food, Forgery, Weapons

\subsubsection{PROTOCOL}
\textbf{Type:} Active, Social
\\ \textbf{Linked Aptitude:} SAV
\\ \textbf{What it is:} Protocol is the art of making a good
impression in social settings. This includes keeping
up with the latest memes, trends, gossip, interests and
habits of various (sub)cultural group.
\\ \textbf{When you use it:} Use Protocol whenever you need
to choose your words carefully, determine who is the
appropriate person to speak to, impress someone with
your grasp of customs, or otherwise fit into a specific
social/cultural grouping. Part etiquette, part streetwise,
Protocol allows you to navigate treacherous social
waters and put people at ease. If the character is dealing with a suspicious or hostile audience, make this an
Opposed Test against the target’s WIL + WIL + SAV.
\\ \textbf{Specializations:} Anarchist, Brinker, Criminal, Factor,
Hypercorp, Infomorph, Mercurial, Reclaimer, Preservationist, Scum, Ultimate

\subsubsection{NEGATING SOCIAL GAFFES}
Sometimes a player will make a mistake that their
character never would, whether that’s failing to stand
in the presence of hypercorp royalty, confusing a gang
leader for a peon, or accidentally insulting someone’s
heritage. In cases like this, the player may make a Protocol Test for the appropriate field in order to negate
the gaffe. If successful, the character never actually
screwed up, or at least managed to cover their tracks
without ruffling any feathers.
%%% txt/186.txt

\subsubsection{PSI ASSAULT}
\textbf{Type:} Active, Mental, Psi
\\ \textbf{Linked Aptitude:} WIL
\\ \textbf{What it is:} Psi Assault is the skill of damaging another ego’s mind. It can only be purchased by characters with the Psi trait (p. 147).
What it does: Use Psi Assault when attacking another ego’s mind in psi combat.
\\ \textbf{Specializations:} By sleight

\subsubsection{PSYCHOSURGERY}
\textbf{Type:} Active, Technical
\\ \textbf{Linked Aptitude:} INT
\\ \textbf{What it is:} Psychosurgery is the use of machine-aided psychological techniques to repair, damage, or
manipulate the psyche.
\\ \textbf{When you use it:} Use Psychosurgery to attempt the
tricky process of editing someone’s mind (see Psychosurgery, p. 229). Psychosurgery can be used beneficially to help patients who remember their deaths, feel
disconnected after remorphing, or have experienced
other sorts of mental traumas. This skill may also be
used to interrogate, torture, or otherwise mess with
captive minds in a VR environment.
\\ \textbf{Specializations:} Memory Manipulation, Personality
Editing, Psychotherapy

\subsubsection{RESEARCH}
\textbf{Type:} Active, Technical
\\ \textbf{Linked Aptitude:} COG
\\ \textbf{What it is:} Research is the skill for looking up information on the Mesh: searching, sifting, mining, and
interpreting data. This includes knowing where to look,
what links to follow, and how to optimize your queries.
\\ \textbf{When you use it:} Use the Research skill whenever
you need to look up the answer to a question, find
databases, search archives, or track down anything
online. Research is typically a Task Action with the
timeframe and difficulty modifier determined by the
gamemaster. See the Online Research, p. 249.
\\ \textbf{Specializations:} Tracking, by information type

\subsubsection{SCROUNGING}
\textbf{Type:} Active, Mental
\\ \textbf{Linked Aptitude:} INT
\\ \textbf{What it is:} Scrounging is your ability to find things,
particularly things of use or value that are concealed,
buried, or hard to find. This includes knowing where
to look and what to look for. Scrounging differs from
both Perception and Investigation in that it is about
finding items hidden among others, and in most cases
about finding something in particular (food, valuables, etc.).
\\ \textbf{When you use it:} Use Scrounging to dumpster-dive a
meal, search ruins for relics, find bargains at a bazaar,
forage berries in the forest, locate a spacesuit in an
abandoned ship, etc. Scrounging is typically handled
as a Task Action with a timeframe and difficulty
modifier determined by the gamemaster.
\\ \textbf{Specializations:} Bazaars, Forests, Habitats, Ruins

\subsubsection{SEEKER WEAPONS}
\textbf{Type:} Active, Combat
\\ \textbf{Linked Aptitude:} COO
\\ \textbf{What it is:} Seeker Weapons covers the use and
maintenance of seeker launchers (p. 339) and seeker
missiles (p. 340).
\\ \textbf{When you use it:} Use this skill when attacking with
a seeker in ranged combat (p. 191).
\\ \textbf{Specializations:} Armband, Pistol, Rifle, Underbarrel

\subsubsection{SENSE}
\textbf{Type:} Active, Mental, Psi
\\ \textbf{Linked Aptitude:} INT
\\ \textbf{What it is:} Sense is the use of psi to scan egos. Only
characters with the Psi trait (p. 147) may purchase
this skill.
What it does: See Psi, p. 220.
\\ \textbf{Specializations:} By sleight

\subsubsection{SPRAY WEAPONS}
\textbf{Type:} Active, Combat
\\ \textbf{Linked Aptitude:} COO
\\ \textbf{What it is:} The Spray Weapons skill covers the use
and maintenance of cone-effect ranged weapons (see
Spray Weapons, p. 340).
\\ \textbf{When you use it:} A player uses their Sonic Weapons
skill whenever they are attacking with a spray weapon
in ranged combat (p. 191).
\\ \textbf{Specializations:} Buzzer, Freezer, Shard, Shredder, Torch

\subsubsection{SWIMMING}
\textbf{Type:} Active, Physical
\\ \textbf{Linked Aptitude:} SOM
\\ \textbf{What it is:} Swimming is the art of moving and
not drowning within fluids. It includes floating,
surface swimming, snorkeling, diving, and related
equipment use.
\\ \textbf{When you use it:} Use Swimming whenever you
need to move and survive in water or another liquid
environment. Swimming in a non-threatening environment can be handled as a Simple Success Test.
Swimming over a long distance could be handled as
a Task Action. Diving off a cliff into a lake, preventing yourself from being swept away in a raging river
current, or making sure you’ve set a proper gas mix
for a deep-sea dive, among other things, requires a
Success Test.
\\ \textbf{Specializations:} Diving, Freestyle, Underwater Diving

\subsubsection{THROWING WEAPONS}
\textbf{Type:} Active, Combat
\\ \textbf{Linked Aptitude:} COO
\\ \textbf{What it is:} Throwing Weapons skill covers the use
and maintenance of standard throwing weapons, like
grenades (p. 340).
\\ \textbf{When you use it:} Use Throwing Weapons skill
whenever you are attacking with a throwing weapon
in ranged combat (p. 191).
\\ \textbf{Specializations:} Grenades, Knives, Rocks
%%% txt/187.txt



\subsubsection{UNARMED COMBAT}
\textbf{Type:} Active, Combat
\\ \textbf{Linked Aptitude:} SOM
\\ \textbf{What it is:} Unarmed Combat is your ability to attack and
defend without using weapons.
\\ \textbf{When you use it:} Use Unarmed Combat whenever you
want to attack someone with your fists, feet, elbows, knees,
or other body parts in melee combat (p. 191).
\\ \textbf{Specializations:} Implant Weaponry, Kick, Punch, Subdual

\begin{quotation}
\textbf{USING KNOWLEDGE SKILLS}
\\
At first glance, it may seem that Knowledge skills have
fewer in-game applications than Active skills. To some
degree this is the case. The importance of Knowledge
skills, however, should not be underestimated. While
they play a role in analyzing clues and solving mysteries, the real value of Knowledge skills is in helping the
characters—and the players—understand the world of
Eclipse Phase. In particular these skills can be used to
make plans, assess a situation, identify strengths and
weaknesses, evaluate worth, make comparisons, forecast
probable outcomes, or understand the applicable science,
socio-economic factors, or cultural or historical context.
   For example, a group of characters looking to break
into a facility could use Profession: Security Procedures
to evaluate the defenses, Academic: Architecture to
identify covert points of entry, Interests: Sports to plan
their infiltration at a time when the guards are likely to
be distracted, Interests: Triads to identify a local crime
group that can sell them breaking and entering gear,
and Art: Sculpture when picking a valuable art piece
with which to bribe an insider. When used appropriately,
these skills can be just as beneficial as the Active skills
used to break inside, if not more so because the plan
is more likely to succeed as a result of this preparation.
   It is largely up to the gamemaster to enforce how
useful Knowledge skills are in their game. The easiest
way to reinforce their relevance is to penalize characters who don’t take advantage of them. For example,
characters who didn’t use their Profession: Security
Procedures in the example above might end up being
surprised when they run across a security system they
are not prepared to deal with, forcing them to improvise or even abandon their plans.
\end{quotation}

\section{SPECIAL SKILLS}
While the preceding list represents the skills most commonly
used in Eclipse Phase, there may be certain skills called for in a
campaign that are not found in this book. In this case, the gamemaster
may work with the players to create a new skill to fill
this void. This option should be used sparingly to prevent skill
bloat, and all skills are subject to approval by the gamemaster
If you choose to create a new skill, keep in mind that it needs
to be linked to an existing aptitude and should be a skill that
is available to all characters, not just specific to one character.










%%%%OLD EXAMPLES

%\begin{quotation}
%\textbf{USING KNOWLEDGE SKILLS}
%\\
%
%\end{quotation}


%\begin{enumerate}
%\item Define Character Concept (p. 130)
%\item Choose Background (p. 131)
%\item Choose Faction (p. 132)
%\item Spend Free Points (p. 134)
%	\begin{itemize}
%	\item 105 aptitude points
%	\item 1 Moxie
%	\item 5,000 credit
%	\item 50 Rep
%	\item Native tongue
%	\end{itemize}
%\item Spend Customization Points (p. 135)
%	\begin{itemize}
%	\item 1,000 CP to spend
%		\begin{itemize}
%	     \item 15 CP = 1 Moxie
%	     \item 10 CP = 1 aptitude point
%	     \item 5 CP = 1 psi sleight
%	     \item 5 CP = 1 specialization
%	     \item 2 CP = 1 skill point (61-80)
%	     \item 1 CP = 1 skill point (up to 60)
%	     \item 1 CP = 1,000 credit
%	     \item 1 CP = 10 rep
%	    \end{itemize}
%    \item Active skill minimum: 400 skill points
%    \item Knowledge skill minimum: 300 skill points
%    \item Choose Starting Morph (pp. 136 and 139)
%    \item Choose Traits (pp. 136 and 145)
%	\end{itemize}
%\item Purchase Gear (p. 136)
%\item Choose Motivation (p. 137)
%\item Calculate Remaining Stats (p. 138)
%\item Detail the Character (p. 138)
%\end{enumerate}
%

%
% \begin{tabular}{ccccccc}
%COG & COO & INT & REF & SAV & SOM & WIL \\ 
%20 & 15 & 15 & 20 & 10 & 15 & 10 \\ 
%\end{tabular} 
%


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ep"
%%% End: 
