\chapter{Mind Hacks}
\label{cha:mind-hacks}
%%%from the 2nd printing

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ep"
%%% End: 

\section{PSI}

\begin{quotation}

$\triangleright $ Desdemona: Glad to have you back. I hope
you had a pleasant farcast from Pelion and don’t
feel too much lack. While you were out, a message
from Aeneas with a precis on psi, extracted from
the infomorph backup of psigeneticist Daborva
(StellInt, Dipole Research Station on Ganymede),
was rerouted for distribution to your Firewall node.
\end{quotation}

Coined by the biologist Bertold P. Wiesner,
“psi” was originally an umbrella term
used to describe a number of so-called
“psychic” abilities and other speculative
paranormal phenomena such as telepathy
and extra-sensory perception. While
the term was used extensively in the
field of parapsychology and pop culture
in the twentieth and early twenty-first
centuries, the study of psi was largely
considered a pseudoscience with flawed
methodologies and gradually lost funding
and support.
During the Fall, however, repeated
rumors and accounts of unexplained
phenomenon drew the attention of
scientists, military leaders, and singularity
seekers alike. Numerous nanovirii had
been unleashed upon transhumanity,
racing through populations and transforming
as they spread. Some inflicted
only minor biological or mental changes
and impairments, but many were vicious
and deadly. The most feared variants,
however, were those that Firewall has
come to label as the Exsurgent virus—a
transformative nano-plague that mutates
its victims and subverts them to its will.
The Exsurgent virus was also observed
to radically modify the subject’s neural
patterns and mental state, affecting synaptic
arrangement and even modulating
synaptic currents. These changes alter
and enhance the victim’s cognition and
seemed to endow an ability to sense and
even affect the minds of others from a
short distance—an ability dubbed “psi”
as the causal factors continue to mystify
us. The existence and nature of this phenomenon
remains carefully concealed
and under wraps in controlled habitats,
so as not to trigger widespread panic.
Among anarchist and other open communities,
knowledge of psi is more widespread,
but details are vague and reports
are generally greeted with skepticism.
The Exsurgent virus is exceptionally
mutable and adaptive, however, and two
argonaut researchers who were aware of
and studying it soon made an interesting
discovery. One variant strain of the virus
was found that endowed the subject
with exceptional mental abilities without
engaging the transformative process
of the other strains. Though infection
still has other drawbacks, Firewall and
other agencies have come to regard
this strain as “safe” in the sense that
the subject does not transmogrify into
something else and their general personality
remains intact. Intrigued that this
avenue of inquiry might lead to a way
to nullify the effects of other Exsurgent
strains, Firewall and others continue
to experiment with the strain with the
cooperation of willing test subjects (or
according to some reports, unwilling
victims in the case of certain authorities
and hypercorps).

\subsection{THE NATURE OF PSI}
Labeled the Watts-MacLeod strain
after the researchers who isolated it,
further study has gained insight into
the effect this virus has on transhuman
brains. Careful analysis of infected
subjects discovered that their altered
synapses generate a modulated brainwave
pattern that is extremely difficult
to detect. Those “in-the-know” have
come to refer to these asynchronous
brainwaves as “psi waves,” fitting with
the Greek letter designation of other
brainwaves (alpha, beta, delta, gamma,
theta). Likewise, affected individuals are
known as “asyncs.”
Exploration of the explicit causal factors
behind psi waves remains stymied.
Theories regarding extraordinary mental
processes with the ability to change
quantum states have been explored but
remain frustratingly inconclusive. Neuroimaging
and mapping have enabled
scientists to pinpoint structures within
the brain, neural activity, and perturbations
in the brain’s bioelectric field that
are associated with psi processes, but
attempts to duplicate these features
in non-infected brains have resulted in
failure or worse. Attempts to identify
asyncs by psi brainwave patterns are not
even assured of success. Numerous dead
ends have prompted many researchers to
postulate that the mechanics underlying
psi are simply too strange and too far
beyond transhumanity’s understanding
of physical sciences—perhaps reinforcing
theories that the Exsurgent virus is in
fact of alien origin.
One leading speculation is that the
changes wrought in the mind by infection
actually entangle some of the
neural sub-systems, enable some sort of
quantum field within the brain, or possibly
create Bose-Einstein condensates
within the brain, allowing for quantum
computation or perhaps hypercomputation.
This enhances the async’s mental
capabilities to the level provided by
modern implants and neuro-mods—and
sometimes beyond. This does not explain
the capabilities of other asyncs, however,
especially those used to read or affect
other biological minds. These abilities
seem to involve reading brain waves
from a short range or affecting another’s
mind via direct physical contact with the
target’s bio-electric fields. Of course I can
only speculate in accordance with what
Firewall has uncovered—it is quite possible
that certain hypercorps or other factions
have made further breakthroughs,
but are keeping the information to
themselves.
The initiation and use of psi talents
is generally understood to take place on
a subconscious level, meaning that the
async is not actively aware of the fundamental
processes that fuel the psi-waves.
Training in certain skills, however, allows
an async to focus on certain tasks and
psi abilities. These are called “sleights:”
mnemonic or cognitive algorithms of psi
use rooted in the async’s ego.
The percentage of the transhuman
population believed to have contracted
the Watts-MacLeod strain remains statistically
insignificant—less than .001\%
of the population. The vast number of
asyncs have been recruited by various
agencies, “disappeared” for study, or
simply eliminated as a potential threat.
Ten years after the Fall, Firewall and
other agencies have come to regard
Watts-MacLeod infection as comparatively
safe, though we remain quite
wary of unforeseen side effects or other
hidden dangers. Most of us engaged in
studying the phenomenon now consider
asyncs to be useful as a tool for fighting
the Exsurgent virus and other threatsdespite the protests of those who are
convinced that asyncs are not in control
of their own minds and are not to be
trusted. As of yet we have encountered
no cases of Watts-MacLeod infection
that have inflicted anything other than
psi abilities, though there seems to be
an increased risk for asyncs to succumb
to other Exsurgent strains should they
encounter them. There are other risks associated
with Watts-MacLeod infection,
such as extreme fatigue and even lethal
biofeedback resulting from extensive use
of psi sleights and a statistically likelihood
of developing mental disorders due
to the increased mental stress placed on
the async’s mind.

\newpage
\subsection{ÆTHER JABBER: ASYNCS}
\# Start Æther Jabber \# \\
\# Active Members: 2 \#

\textbf{\textit{1}} Sorry to bother you, but my muse just
alerted me to this excerpt that was
sent around to my Firewall team. Is
this for real? I’ve heard the talk about
psi before—enough to be convinced
that there’s something to it, even
if we can’t explain it—but this bit
about variant Exsurgent infection is
too much. Are we seriously going to
be working with someone who’s a
known carrier? And can you shed any
more light on how asyncs do their
mojo? I’m worried now. And since
you are connected to the Medeans, I
thought I’d take the chance and ask.

\textbf{\textit{2}} Well, as to the Medeans ... that’s history.
I am back on the freelancing market
right now. But no problem, I’ll try and
explain. I know it is not easy to grasp.

\textbf{\textit{1}} Shiny.

\textbf{\textit{2}} Yes, Srit was once infected with a
strain of the Exsurgent virus, probably
on Mars near the end of the Fall. I say
“was” because the Watts-MacLeod
strain seems to go dormant shortly
after it finishes rewiring the victim’s
brain; the plague nanobots die off and
get flushed out of the system, unlike
other Exsurgent strains, which continue
to stick around and transform
the subject. At least, that’s the dominant
theory—I’ve also seen some
speculation that async minds might
be modified so that they continue to
produce bio-nanobots that linger in
the brain, though what function these
serve remains unclear. However, the
prevailing opinion among our best
neuroscientists is that people like Srit
are safe and non-infectious once the
virus has run its course. I’ll even go
a bit further and say that prevailing
opinion is that they can be trusted,
assuming they don’t catch another
infection ... which they unfortunately
seem to be a bit prone too. Not everyone
agrees of course, but we have an
abundance of paranoia in our circles.
So far, we haven’t seen any evidence
that any of our asyncs have been
turned by that initial infection, and the
utility and usefulness of having psiactives
on our side has simply been
too important to push aside.
All right. I can’t say that I’ll trust her,
but I’ll try and give her the benefit of
the doubt. I’ll be damned if I’m going
to trust an async that’s not vouched
for by Firewall though—who knows
what the hell a hypercorp like Skinthetic
might be cooking up in their
black labs.

\textbf{\textit{2}} That seems like a wise choice.

\textbf{\textit{1}} Maybe you can put my mind at ease
by explaining to me in a bit more
detail how Watts-MacLeod infection
occurs.

\textbf{\textit{2}} Well, like the other Exsurgent strains
you are unfortunately familiar with,
the primary transmission vector is a
nanovirus, but we speculate that it
may also be transmitted as a digital
computer virus or possibly even as
a basilisk hack. The physical plague
form is spread by highly-advanced
techno-organic nanobots that infect
a biomorph and use bio-mimicry
mechanisms to pass as normal cells
and penetrate the blood-brain barrier
and central nervous system. The
nanobots are several steps beyond
anything our technology can produce,
are very difficult to detect, and can
overwhelm most defensive countermeasures.
Infected minds are essentially
rewired, and these changes will
be copied when the ego is uploaded.
Synthmorphs and infomorphs remain
immune to this nano-infection, but
they are theoretically vulnerable to
other transmission vectors.

\textbf{\textit{1}} I’ve heard that synthmorphs are effectively
invulnerable to psi as well. This
true?

\textbf{\textit{2}} Yes. As far as we can tell, async
abilities only effect biological
minds—either their own or others.
And they can only read/affect others
from a very short distance, requiring
physical contact in most cases. The
half-biological minds of pods are also
vulnerable, though to a lesser extent.
Likewise, asyncs need a biological
brain to use their abilities—they
can’t use their psi if sleeved in a synthmorph
and have difficulty in a pod.

\textbf{\textit{1}} Interesting. So, I have to ask again—
you’re sure she’s safe? I’ve heard
that some of these asyncs can be real
nut-jobs.

\textbf{\textit{2}} I’ve heard from several of these
asyncs directly. The fact is, infection
rewrites their brain, and some of
them came out the other side feeling
fundamentally altered. Either they
felt like a different person, or they
felt like there was something new
that was part of them—something
that they didn’t necessarily like. One
described it as presence, another as
a black void that whispered at them.
Yet another described it as giving a
personality to their unconscious mind,
which only made the gulf between
unconscious and conscious mind all
the more intimidating. Some of them
preferred to suicide and revert to a
pre-infection backup. While they may
be more prone to cracking up as a
result, I haven’t ever heard one talk
about their abilities as something
they couldn’t control.

1 Well, that’s fucking cheery. There’s
nothing else we have on how this psi
stuff actually works?

\textbf{\textit{2}} Unfortunately, we don’t. Even the
Prometheans haven’t been much
help. There are theories, of course, but
nothing that we’ve been able to verify
with rigorous experimentation. It
doesn’t help that the factions that are
aware of psi’s existence don’t exactly
compare notes—they’re all too busy
looking into ways to weaponize it
and use it against each other, instead
of figuring out how to use it for the
benefit of transhumanity.

\textbf{\textit{1}} Of course. The TITANs didn’t get us,
but we can still get ourselves. It worries
me that the best we’ve come up
with is nothing.

\textbf{\textit{2}} It’s important to keep perspective.
Transhumanity has come quite a
distance and made some impressive
accomplishments, but our understanding
of the universe is still in its infancy.
What we may be facing here is something
concocted by an intelligence so
far beyond our own that we are but
insignificant insects in comparison.
It likely has a grasp on the universe
that is simply beyond our ability to
understand. We shouldn’t be cocky
and think that we can decipher any
mystery thrown at us ... we should
instead be very, very afraid.

\newpage

Though neuroscience has ascended to impressive
pinnacles, allowing minds to be thoroughly scanned,
mapped, and emulated as software, the transhuman
brain remains a place that is complicated, not fully
understood, and thoroughly messy. Despite a prevalence
of neural modifications, meddling with the seat
of consciousness remains a tricky and hazardous
procedure. Nevertheless, psychosurgery—editing the
mind as software—remains common and widespread,
sometimes with unexpected results.
Likewise, even as the knowledge of neuroscientists
grows on an exponential basis, some are discovering
that minds are far more mysterious than they had ever
imagined. During the Fall, scattered reports of “anomalous
activity” by individuals infected by one of the
numerous circulating nanoplagues were discounted
as fear and paranoia, but subsequent investigations
by black budget labs has proven otherwise. Now, toplevel
confidential networks whisper that this infection
inflicts intricate changes in the victim’s neural network
that imbue them with strange and inexplicable abilities.
The exact mechanism and nature of these abilities
remains unexplained and outside the grasp of modern
transhuman science. Given the evidence of a new
brainwave type and the paranormal nature of this
phenomenon, it is loosely referred to as “psi.”

\section{PSI}
In Eclipse Phase, psi is considered a special cognitive
condition resulting from infection by the mutant—and
hopefully otherwise benign—Watts-Macleod strain
of the Exsurgent virus (p. 367). This plague modifies
the victim’s mind, conferring special abilities. These
abilities are inherent to the brain’s architecture and
are copied when the mind is uploaded, allowing the
character to retain their psi abilities when changing
from morph to morph.

\subsubsection{PREREQUISITES}

To wield psi, a character must acquire the Psi trait (p.
147) during character creation. It is theoretically also
possible to acquire the use of psi in game via infection
by the Watts-MacLeod strain; see The Exsurgent
Virus, p. 362.
Psi ability is considered an innate ability of the
ego and not a biological or genetic predisposition of
the morph. While psi researchers do not understand
how it is possible to transfer this ability via uploads,
backups, and farcasting, it has been speculated that
all components of an async’s ego are entangled on a
quantum level, or that they possess the ability to entangle
themselves or form a unique conformation or
alignment as a whole even after they have been copied,
up-, or downloaded. This speculated entanglement
process is also thought to be the origin of the impairment
that asyncs experience when adapting to a new
morph (see below).

\subsubsection{MORPHS AND PSI}

Asyncs require a biological brain to draw on their
abilities (the brains of uplifted animals count). An
async whose ego is downloaded into an infomorph or
fully computerized brain (synthmorphs) has no access
to their abilities as long they remain in that morph.
Asyncs inhabiting a pod morph may use psi, but
their abilities are restricted as pod brains are only
partly biological. Pod-morphed asyncs suffer a –30
modifier on all tests involving the use of psi sleights
and the impact from using sleights would be doubled.

\subsubsection{MORPH ACCLIMATIZATION}

Async minds undergo extra difficulty adjusting to new
morphs. For 1 day after the character has resleeved,
they will suffer the effects of a single derangement
(p. 210). The gamemaster and player should choose a
derangement appropriate to the character and story.
Minor derangements are recommended, but at the
gamemaster’s discretion moderate or major derangements
may be applied. No trauma is inflicted with
this derangement.

\subsubsection{MORPH FEVER}

Asyncs find it irritating and traumatizing to endure life
as an infomorph, pod, or synthmorph for long periods
of time. This phenomenon, known as morph fever,
might cause temporary derangements and trauma to
the asyncs’ ego, possibly even to the grade of permanent
disorders. If stored or held captive as an active
infomorph (i.e. not in virtual stasis), the async might
go insane if not psychologically aided by some sort of
anodyne program or supporting person during storage.
In game terms, asyncs take 1d10 $\div$ 2 (round up)
points of mental stress damage per month they stay
in a pod, synthmorph or infomorph form without
psychological assistance by a psychiatrist, software,
or muse.

\subsubsection{PSI DRAWBACKS}

There are several drawbacks to psi ability:

\begin{itemize}
\item The variant Exsurgent strain that endows psi ability
rewires the character’s brain. An unfortunate side
effect to this change is that asyncs acquire a vulnerability
to mental stress. Reduce the async’s Trauma
Threshold by 1.
\item The mental instability that accompanies psi infection
also tends to unhinge the character’s mind.
Asyncs acquires one Mental Disorder negative trait
(p. 150) for each level they have of the Psi trait
without receiving any bonus CP. The gamemaster
and player should agree on a disorder appropriate
to the character. This disorder may be treated over
time, according to normal rules (see Mental Healing
and Psychotherapy, p. 215).
\item Characters with the Psi trait are also vulnerable to
infection by other strains of the Exsurgent virus.
The character suffers a –20 modifier when resisting
Exsurgent infection (p. 362).
\item Critical failures when using psi tend to stress the
async’s mind. Each time a critical failure is rolled
when making a sleight-related test, the async suffers
a temporary brain seizure. They suffer a –30 modifi
er and are incapable of acting until the end of the
next Action Turn. They must also succeed in a WIL
+ COG Test or fall down.
\end{itemize}

\subsubsection{PSI SKILLS AND SLEIGHTS}

Transhuman psi users can manipulate their egos and
otherwise create effects that can often be neither
matched nor mimicked by technological means. To use
these abilities, they train their mental processes and
practice cognitive algorithms called sleights, which
they can subconsciously recall and use as necessary.
Sleights fall into two categories: psi-chi (cognitive
enhancements, p. 223) and psi-gamma (brainwave
reading and manipulation, p. 225). Psi-chi sleights are
available to anyone with the Psi trait (p. 147), but psigamma
sleights are only available to characters with
the Psi trait at Level 2. In order to use these sleights,
the async must be skilled in the Control (p. 178), Psi
Assault (p. 183), and/or Sense skills (p. 184), as appropriate
to each sleight.

\subsubsection{ROLEPLAYING ASYNCS}

Any player who chooses to play an async should keep
the origin of their abilities in mind: Watts-MacLeod
strain infection. The character may not be aware of
this source, but they undoubtedly know that they underwent
some sort of transformation and have talents
that no one else does. If unaware of the infection, they
have likely learned to keep their abilities secret lest
they be ridiculed, attacked, or whisked away to some
secret testing program. Learning the truth about their
nature could even be the starting point of a campaign
and/or their introduction to Firewall. If they know the
truth, however, the character must live with the fact
that they are the victim of a nanoplague likely spread
by the TITANs that may or may not lead to complications,
side effects, or other unexpected revelations in
their future.
Gamemasters and players should make an effort
to explore the nature of this infection and how the
character perceives it. As noted previously, asyncs are
often profoundly-changed people. The invasive and
alien aspect of their abilities should not be lost on
them. For example, an async might conceive of their
psi talents as a sort of parasitic entity, living off their
sleights, or they might feel that using these powers
puts them in touch with some sort of fundamental
substrate of the universe that is weird and terrifying.
Alternately, they could feel as if their personality was
melded with something different, something that
doesn’t belong. Each async is likely to view their situation
differently, and none of them pleasantly.

\subsection{USING PSI}

Using psi—i.e., drawing on a certain sleight to procure
some kind of effect—does not always require a
test. Each sleight description details how the power
is used.

\subsubsection{ACTIVE PSI}

Active psi sleights must be “activated” to be used.
These sleights usually require a skill test. Sleights that
target other sentient beings or life forms are always
Opposed Tests, while others are handled as Success
Tests. The level of concentration required to use these
sleights varies, and so may call for a Quick, Complex,
or Task Action. Active sleights also cause strain (p.
223) to the async. Most psi-gamma sleights fall into
this category.

\subsubsection{PASSIVE PSI}

Passive psi sleights encompasses abilities that are considered
automatically active and subconscious. They
rarely require an action to be activated and require
no effort or strain by the psi user. Passive sleights typically
add bonuses to various activities or allow access
to certain abilities rather than calling for some kind of
skill test. Most psi-chi sleights fall into this category.

\subsubsection{PSI RANGE}

Sleights have a Range of either Self, Touch, or Close.
Self: These sleights only affect the async.
Touch: Sleights with a Touch range may be used
against other biological life, but the async must have
physical contact with the target. If the target avoids
being touched, this requires a successful melee attack,
applying the touch-only +20 modifier. This attack
does not cause damage, and is considered part of the
same action as the psi use.
Close: Close sleights involve interaction with other
biological life from a short distance. The optimal distance
is within 5 meters. For each meter beyond that,
apply a –10 modifier to the test.
Psi vs. Psi: Due to the nature of psi, sleights are
more effective against other psi users. Sleights with
a range of Touch may be used from a Close range
against another async. Likewise, a sleight with a Close
range may be used at twice the normal distance (10
meters) when wielded on another async.

\subsubsection{TARGETING}

Synthmorphs, bots, and vehicles may not be targeted
by psi sleights, as they lack biological brains. Pods
- with brains that are half biological and half computer -
are less susceptible and receive a +30 modifier
when defending against psi use. Note that infomorphs
may never be targeted by psi sleights as psi is not effective
within the mesh or simulspace.
Multiple Targets: An async may target more than
one character with a sleight with the same action, as
long as each of them can be targeted via the rules
above. The psi character only rolls once, with each of
the defending characters making their Opposed Tests
against that roll. The psi character suffers strain (p.
223) for each target, however, meaning that using psi
on multiple targets can be extremely dangerous.
Animals and Less Complex Life Forms: Psi works
against any living creature with a brain and/or
nervous system. Against partially-sentient and partially-
uplifted animals, it suffers a –20 modifier and
increases strain by +1. Against non-sentient animals,
it suffers a –30 modifier and increases strain by +3. It
has no effect on or against less complex life forms like
plants, algae, bacteria, etc.
Factors and Aliens: At the gamemaster’s discretion,
psi sleights may not work on alien creatures at all, depending
on their physiology and neurology. If it does
work, it is likely to suffer at least a –20 modifier and
+1 strain.

\subsubsection{OPPOSED TESTS}

Psi that is used against another character is resisted
with an Opposed Test. Defending characters resist with
WIL x 2. Willing characters may choose not to resist.
Unconscious or sleeping characters cannot resist.
If the psi-wielding character succeeds and the defender
fails, the sleight affects the target. If the psi user
fails, the defender is unscathed. If both parties succeed
in their tests, compare their dice rolls. If the psi
user’s roll is higher, the sleight bypasses the defender’s
mental block and affects the target; otherwise, the
sleight fails to affect the defender’s ego.

\subsubsection{TARGET AWARENESS}

The target of a psi sleight is aware they are being
targeted any time they succeed on their half of the
Opposed Test (regardless on whether the async
rolls higher or not). Note that awareness does not
necessarily mean that the target understands that psi
abilities are being used on them, especially as most
people in Eclipse Phase are unaware of psi’s existence.
Instead, the target is simply likely to understand that
some outside influence is at work, or that something
strange is happening. They may suspect that they
have been drugged or are under the influence of some
strange technology.
Targets who fail their roll remain unaware.

\subsubsection{PSI FULL DEFENSE}

Like full defense in physical combat (p. 198), a defender
may spend a Complex Action to rally and concentrate
their mental defenses, gaining a +30 modifier
to their defense test against psi use until their next
Action Phase.

\subsubsection{CRITICALS}

If the defender rolls a critical success, the character
attempting to wield psi is temporarily locked out of
the target’s mind. The psi user may not target that
character with sleights until an appropriate “reset”
period has passed, determined by the gamemaster.
If the async rolls a critical failure, they suffer temporary
incapacitation as their mind dysfunctions in some
harsh and distressing ways (see Psi Drawbacks, p. 221).
If a psi user rolls a critical success against a defender,
or the defender rolls a critical failure, double
the potency of the sleight’s effect. In the case of psi attacks,
the DV can be doubled or mental armor can be
bypassed. Alternately, when using Psi Assault (p. 183),
the targeted character may be in danger of infection
by the Watts-Macleod strain (p. 362).

\subsubsection{MENTAL ARMOR}

The Psi Shield sleight (p. 228) provides mental armor,
a form of neural hardening against psi-based attacks.
Like physical armor, this mental armor reduces the
amount of damage inflicted by a psi assault.

\subsubsection{DURATION}

Psi sleights have one of four durations: constant, instant,
temporary, or sustained.
Constant: Constant sleights are always “on.”
Instant: Instant sleights take effect only in the
Action Phase in which they are used.
Temporary: Temporary sleights last for a limited
duration with no extra effort from the async. The
temporary duration is determined by the async’s WIL
$\div$ 5 (round up) and is measured in either Action Turns
or minutes, as noted. Strain for the sleight is applied
immediately when used, not at the end of the duration.
Sustained: Sustained sleights require active effort
to maintain for as long as the async wants to keep
it active. Sustaining a sleight requires concentration,
and so the async suffers a –10 modifier to all other
skill tests while the sleight is sustained. The async
must also stay within the range appropriate to the
sleight, otherwise the sleight immediately ends. More
than one sleight may be sustained at a time, with a
cumulative modifier. Strain for the sleight is applied
immediately when used, not at the end of the duration.
At the gamemaster’s discretion, sleights that are
sustained for long periods may incur additional strain.

\subsubsection{STRAIN}

The use of psi is physically (and sometimes psychologically)
draining to a psi user. This phenomenon is
known as strain, and manifests as fatigue, exhaustion,
pain, neural overload, cardiovascular stress, and
adynamia (loss of vigor). Though strain has only
rarely been known to actually kill an async, the use
of too much active psi can be life-threatening in some
circumstances.
In game terms, every active sleight has a Strain
Value of 1d10 $\div$ 2 (round up) DV. Every active sleight
lists a Strain Value Modifier that modifies this amount.
For example, a sleights with a Strain Value Modifier
of –1 inflicts (1d10 $\div$ 2) –1 DV.
If the damage points suffered from strain exceed
the character’s Wound Threshold, they may inflict a
wound just like other damage (see Wounds, p. 207).

\begin{quotation}
\textbf{EXAMPLE}
\\
Matric is investigating a disappearance, so he
decides to use his Qualia sleight to boost his
Intuition while hunting for clues. That psi-chi
sleight takes only a Quick Action to initiate and
requires no test. Matric’s WIL is 25, so the duration
of this temporary sleight is 5 Action Turns
(25 $\div$ 5 = 5). The sleight’s Strain modifier is –1,
so he is facing (1d10 $\div$ 2) –1 DV. He rolls a 1, so
he takes no strain at all!
Later on, Matric finds himself in a life-or-death
struggle with a kidnapper. Lucky for Matric,
they’re in a melee, so he’s close enough to try
and touch his opponent. On his Action Phase, he
makes an Unarmed Combat Test with a +20 modifi
er (for a touch-only attack) and succeeds. This
allows him to try and use his Psychic Stab sleight.
He rolls his Psi Assault of 57 against the target’s
WIL x 2 (32). His target is in a worker pod morph,
however, which is less susceptible to psi, so he
receives a +30 modifier (32 + 30 = 62). Matric
rolls a 32 and the worker pod a 64—Matric wins!
For damage, he rolls 1d10 + (WIL $\div$ 10). His WIL
is 25, so that’s 1d10 + 3. He rolls a score a 7 and
inflicts 10 (7 + 3) points of damage. The worker
pod screams in pain, suffering a wound from the
psychic assault.

\end{quotation}



\subsection{PSI-CHI SLEIGHTS}
Psi-chi sleights are async abilities that speed up cognitive
informatics (internal information processing) and
enhance the user’s perception and cognition.

\subsubsection{AMBIENCE SENSE}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
This sleight provides the async with an instinctive
sense about an area and any potential threats nearby.
The async receives a +10 modifier to all Investigation,
Perception, Scrounging, and Surprise Tests.

\subsubsection{COGNITIVE BOOST}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Quick \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Temp (Action Turns) \\
\textbf{STRAIN MOD:} –1 \\
The async can temporarily elevate their cognitive
performance. In game terms, Cognition is raised by 5
for the chosen duration. This boost to Cognition also
raises the rating of skills linked to that aptitude.

\subsubsection{DOWNTIME}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Task (min. 4 hours) \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Sustained \\
\textbf{STRAIN MOD:} 0 \\
This sleight provides the async with the ability
to send the mind into a fugue-state regenerative
downtime, during which the character’s psyche is
repaired. The async must enter the downtime for
at least 4 hours; every 4 hours of downtime heals
1 point of stress damage. Traumas, derangements,
and disorders are unaffected by this sleight. For
all sensory purposes, the async is catatonic during
downtime, completely oblivious to the outside
world. Only severe disturbances or physical shock
(such as being wounded or hit by a shock weapon)
will bring the async out of it.

\subsubsection{EMOTION CONTROL}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
Emotion Control gives the async tight control
over their emotional states. Unwanted emotions
can be blocked out and others embraced. This has
the benefit of protecting the async from emotional
manipulation, such as the Drive Emotion sleight or
Intimidation skill tests. The async receives a +30
modifier when defending against such tests.

\subsubsection{ENHANCED CREATIVITY}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
An async with Enhanced Creativity is more imaginative
and more inclined to think outside the box.
Apply a +20 modifier to any tests where creativity
plays a major role. This level of ingenuity can
sometimes seem strange and different, manifesting
in odd or creepy ways, especially with artwork.

\subsubsection{FILTER}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
Filter allows the async to filter out out distractions
and eliminate negative situational modifiers from
distraction, up to the gamemaster’s discretion.

\subsubsection{GROK}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Instant \\
\textbf{STRAIN MOD:} –1 \\
By using the Grok sleight, the async is able to
intuitively understand how any unfamiliar object,
vehicle, or device is used simply by looking at and
handling it. If the character succeeds in a COG x 2
Test, they achieve a basic ability to use the object,
vehicle, or device, no matter how alien or bizarre.
This sleight does not provide any understanding of
the principles or technologies involved—the psi user
simply grasps how to make it work. If a test is called
for, the psi user receives a +20 modifier to use the
device (this bonus only applies to unfamiliar devices,
and/or tests the character is defaulting on—it does not
apply to devices the character is familiar with).

\subsubsection{HIGH PAIN THRESHOLD}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
This sleight allows the async to block out, ignore, or
otherwise isolate pain. The async reduces negative
modifiers from wounds by 10.

\subsubsection{HYPERTHYMESIA}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
Hyperthymesia grants the async a superior autobiographical
memory, allowing them to remember the
most trivial of events. A hyperthymestic async can be
asked a random date and recall the day of the week
it was, the events that occurred that day, what the
weather was like, and many seemingly trivial details
that most people would not be able to recall.

\subsubsection{INSTINCT}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
Instinct bolsters the async’s subconscious ability to
gauge a situation and make a snap judgment that
is just as accurate as a careful, considered decision.
For Task Actions that involve analysis or planning
alone (typically Mental skill actions), the async may
reduce the timeframe by 90\% without suffering a
modifier. For Task Actions that involve partial analysis/
planning, they may reduce the timeframe by 30\%
without penalty.

\subsubsection{MULTITASKING}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
The async can handle vast amounts of information
without overload and can perform more than one
mental task at once. The character receives an extra
Complex Action each Action Phase that may only be
used for mental or mesh actions.

\subsubsection{PATTERN RECOGNITION}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
The character is adept at spotting patterns and correlating
the non-random elements of a jumble—related
items jump out at them. This is useful for translating
languages, breaking codes, or find clues hidden among
massive amounts of data. The character must have
a sufficiently large sample enough time to study, as
determined by the gamemaster. This might range from
a few hours of listening to a spoken transhuman language
to a few days of investigating inscriptions left
by long-dead aliens to a week or more of researching
a lengthy cipher. Languages may be comprehended by
reading or listening to them being spoken. Apply a
+20 modifier to any appropriate Language, Investigation,
Research, or cod-breaking Tests (note that this
does not apply to Infosec Tests made by software to
decrypt a code). The async may also use this ability to
more easily learn new languages, reducing the training
time by half.

\subsubsection{PREDICTIVE BOOST}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
The Bayesian probability machine features of the
async’s brain are boosted by this sleight, enhancing
their ability to estimate and predict outcomes of
events around them as they unfold in real-time and
update those predictions as information changes.
In effect, the character has a more intuitive sense
for which outcomes are most likely. This grants the
character a +10 bonus on any skill tests that involve
predicting the outcome of events. It also bolsters
the async’s decision-making in combat situations by
making the best course of action more clear, and so
provides a +10 bonus to both Initiative and Fray Tests.

\subsubsection{QUALIA}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Quick \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Temp (Action Turns) \\
\textbf{STRAIN MOD:} –1 \\
The async can temporarily increase their intuitive
grasp of things. In game terms, Intuition is raised by
5 for the chosen duration. This boost to Intuition also
raises the rating of skills linked to that aptitude.

\subsubsection{SAVANT CALCULATION}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
The character possesses an incredible facility with
intuitive mathematics. They can do everything from
calculate the odds exactly when gambling to predicting
precisely where a leaf falling from a tree will land
by observing the landscape and local wind currents.
The character specializes in calculation involving the
activity of complex chaotic systems and so can calculate
answers that even the fastest computers could not,
including things like patterns of rubble distribution
from an explosion. However, this mathematic facility
is largely intuitive, so the character does not know the
equations they are solving, they merely know the solution
to the problem.
This sleight also provides a +30 modifier to any
skill tests involving math (which the character is calculating,
not a computer).

\subsubsection{SENSORY BOOST}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Quick \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Temp (Action Turns) \\
\textbf{STRAIN MOD:} –2 \\
An async uses this sleight to increase their natural or
augmented sensory perception (sight, audio, smell,
augmented) by enhanced cerebral processing, granting
a +20 bonus modifier on sensory-based Perception
Tests.

\subsubsection{SUPERIOR KINESICS}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
The async acquires more insight into people’s emotive
signals, gestures, facial expressions, and body
language when it comes time to predict the person’s
emotional state, intent, or reactions. Apply a +10
modifier to Kinesics Skill Tests.

\subsubsection{TIME SENSE}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Temp (Action Turns) \\
\textbf{STRAIN MOD:} –1 \\
An async with this ability can slow down his perception
of time, making everything appear to move
in slow motion or at reduced speed. In game terms,
this sleight grants the async a Speed of +1. This extra
Action Phase, however, can only be spent on mental
and mesh actions.

\subsubsection{UNCONSCIOUS LEAD}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Temp (Action Turns) \\
\textbf{STRAIN MOD:} +0 \\
This sleight allows the async to override their consciousness
and let their unconscious mind take point.
While in this state, the async’s conscious mind is only
dimly aware of what is transgressing, and any memories
of this period will be hazy at best. The advantage
is that the unconscious mind acts more quickly, and
so the async’s Speed is boosted by +1. The character
remains aware and active, but is incapable of complex
communication or other mental actions and is
motivated by instinct and primitive urges more than
conscious thought. Though it is recommended that
the player retain control of their character while using
Unconscious Lead, the gamemaster should feel free to
direct the character’s actions as they see fit.


\subsection{PSI-GAMMA SLEIGHTS}
Psi-gamma sleights deal with contacting (reading
and communicating) and influencing the function of
biological minds (egos within a biomorph, but also
including animal life). Psi-gamma is only available to
characters with Level 2 of the Psi trait.
Most psi-gamma use is handled as an Opposed Test
between the async and the target (p. 222).

\subsubsection{ALIENATION}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Temp (Action Turns) \\
\textbf{STRAIN MOD:} +0 \\ 
\textbf{SKILL:} Psi Assault \\
Alienation is an offensive sleight that create a sense of
disconnection between an ego and its morph—similar
to that experienced when resleeved into a new body.
The ego finds their body cumbersome, strange, and
alien, almost like they are a prisoner within it. If the
async beats the target in an Opposed Test, treat the
test as a failed Integration Test (p. 272) for the target.
This effect lasts for the sleight’s duration.

\subsubsection{CHARISMA}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Quick \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Temp (Minutes) \\
\textbf{STRAIN MOD:} –1 \\ 
\textbf{SKILL:} Control\\
The async uses this sleight to influence the target’s
mind on a subconscious level, so that the target perceives
them to be charming, magnetic, and persuasive.
If the async beats the target in an Opposed Test, they
gain a +30 modifier on all subsequent Social Skill
Tests for the chosen duration.

\subsubsection{CLOUD MEMORY}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Temp (Minutes) \\
\textbf{STRAIN MOD:} –1 \\ 
\textbf{SKILL:} Control\\
Cloud Memory allows the async to temporarily disrupt
the target’s ability to form long-term memories. If
the async wins the Opposed Test, the target’s memorysaving
ability is negated for the duration. The target
will retain short-term memories during this time, but
will soon forget anything that occurred while this
sleight was in effect.

\subsubsection{DEEP SCAN}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Sustained \\
\textbf{STRAIN MOD:} +1 \\ 
\textbf{SKILL:} Sense\\
Deep Scan is a more intrusive version of Thought
Browse (p. 228), made to extract information from the
targeted individual. If the Opposed Test succeeds, the
async telepathically invades the target’s mind and can
probe it for information. For every 10 full points of MoS
the async achieves on their test, they retrieve one piece
of information. Each item takes one full Action Turn to
retrieve, during which the sleight must be sustained. The
target is aware of this mental probing, though they will
not know what information the async acquired.

\subsubsection{DRIVE EMOTION}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Temp (Action Turns) \\
\textbf{STRAIN MOD:} –1 \\ 
\textbf{SKILL:} Control\\
This sleight allows the async to stimulate cortical areas
of the target’s brain related to emotion. This allows
the async to induce, amplify, or tone down specific
emotions, thereby manipulating the target. If the async
beats the target in an Opposed Test, they will act in accordance
with the emotion for the duration and under
certain circumstances may suffer from certain penalties
(up to +/–30), as determined by the gamemaster. For
example, an async might receive a +30 Intimidation
Test modifier against a target imbued with fear.

\subsubsection{EGO SENSE}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Close \\ 
\textbf{DURATION:} Temp (Action Turns) \\
\textbf{STRAIN MOD:} –1 \\ 
\textbf{SKILL:} Sense\\
Ego Sense can be used to detect the presence and
location of other sentient and biological life forms
(i.e., egos) within the async’s range. To detect these
life forms, the async makes a single Sense Test, opposed
by each life form within range. The async may
suffer a modifier for detecting small animals and
insects, similar to the modifier applied for targeting
them in ranged combat (see p. 193); likewise, a
modifier for detecting larger life forms may also be
applied. If successful, the async has detected that the
life form is nearby. Every 10 full points of MoS will
ascertain another piece of information regarding the
detected life: direction from async, approximate size,
type of creature, distance from async, etc. The async
will know if the target moves, if they do so during the
sleight’s duration.

\subsubsection{EMPATHIC SCAN}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Quick \\ 
\textbf{RANGE:} Close \\ 
\textbf{DURATION:} Sustained \\
\textbf{STRAIN MOD:} –2 \\ 
\textbf{SKILL:} Sense\\
Empathic Scan enables the async to sense the target’s
base emotions. If the async wins the Opposed Test,
they intuitively feel the target’s emotional current state
for as long as the sleight is sustained. At the gamemaster’s
discretion, this knowledge may provide a modifier
(up to +30) for certain Social skill tests.

\subsubsection{IMPLANT MEMORY}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Instant \\
\textbf{STRAIN MOD:} +0 \\ 
\textbf{SKILL:} Control\\
An async using this sleight can implant a memory
of up to an hour’s length inside the target’s mind.
This memory very obviously does not belong to the
target—there is no way they will confuse it for one
of their own. The intent is not to fake memories, but
to place one of the async’s memories in the target’s
mind so that the target can access it just like any other
memory. This can be useful for “archiving” important
data with an ally, providing a literal alternate perspective,
or simply making a “data dump” for the target
to peruse. Implant Memory requires an Opposed Test
against unwilling participants. At the gamemaster’s
discretion, particularly traumatic memories might
inflict mental stress on the recipient (p. 215).

\subsubsection{IMPLANT SKILL}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Temp (Action Turns) \\
\textbf{STRAIN MOD:} +0 \\ 
\textbf{SKILL:} Control\\
Similar to Implant Memory, this sleight allows the
async to impart some of their expertise and implant it
into the target’s mind. For the duration of the sleight,
the target benefits when using that skill. If the async’s
skill is between 31 and 60, the target receives a +10
bonus. If the async’s skill is 61+, the target receives
a +20 bonus. Implant Skill requires an Opposed Test
against unwilling participants. In some cases, the
target has been known to use the skill with the async’s
flair and mannerisms.

\subsubsection{MIMIC}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Quick \\ 
\textbf{RANGE:} Close \\ 
\textbf{DURATION:} Instant \\
\textbf{STRAIN MOD:} +0 \\ 
\textbf{SKILL:} Sense\\
In a setting where changing your body and face is
not unusual, people learn to recognize habits and
personality quirks more often. The async must use
this sleight on a target and succeed in a Success Test.
If successful, the async acquires an “imprint” of the
target’s mind that they can take advantage of when
impersonating that ego. The async then receives a +30
bonus on Impersonation Tests when mimicking the
target’s behavior and social cues.

\subsubsection{MINDLINK}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Quick \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Sustained \\
\textbf{STRAIN MOD:} +1/target after first \\ 
\textbf{SKILL:} Control\\
Mindlink allows two-way mental communication
with a target. This may be used on more than one
target simultaneously, in which case the async can act
as a telepathic “server,” so that everyone mindlinked
with the async may also telepathically communicate
with each other (via the async, however, so they
overhear). Language is still a factor in mindlinked
communications, but this barrier may be overcome by
transmitting sounds, images, emotions, and other sensations.
Mindlink requires an Opposed Test against
unwilling participants.

\subsubsection{OMNI AWARENESS}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Quick \\ 
\textbf{RANGE:} Close \\ 
\textbf{DURATION:} Temp (Minutes) \\
\textbf{STRAIN MOD:} –1 \\ 
\textbf{SKILL:} Sense\\
An async with Omni Awareness is hypersensitive to
other biological life that is observing them. During
this sleight’s duration, the async makes a Sense Test
that is opposed by any life that has focused their attention
on them within the sleight’s range; if successful,
the async knows they are being watched, but not
by whom or what. It does, however, apply a +30 Perception
bonus to spot the observer. This sleight does
not register partial attention or fleeting attention, or
simple perception of the async, it only notices targets
who are actively observing (even if they are concealing
their observation). This sleight is effective in spotting a
tail, as well as finding potential mates in a bar.

\subsubsection{PENETRATION}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Quick \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Instant \\
\textbf{STRAIN MOD:} 1 per AP point \\ 
\textbf{SKILL:} Psi Assault \\
Penetration is a sleight that works in conjunction with
any offensive sleight that involves the Psi Assault skill.
It allows the async to penetrate the Psi Shield of an opponent
by concentrating their psi attack. Every point
of Armor Penetration applied to a psi attack inflicts
1 point of strain. The maximum AP that may be applied
equals the async’s Psi Assault skill divided by 10
(round down).

\subsubsection{PSI SHIELD}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
Psi Shield bolsters the async’s mind to psi attack and
manipulation. If the async is hit by a psi attack, they
receive WIL $\div$ 5 (round up) points of armor, reducing
the amount of damage inflicted. They also receive a
+10 modifier when resisting any other sleights.

\subsubsection{PSYCHIC STAB}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Instant \\
\textbf{STRAIN MOD:} +0 \\ 
\textbf{SKILL:} Psi Assault \\
Psychic Stab is an offensive sleight that seeks to inflict
physical damage on the target’s brain and nervous
system. Each successful attack inflicts 1d10 + (WIL $\div$
10, round up) damage. Increase the damage by +5 if
an Excellent Success is scored.

\subsubsection{SCRAMBLE}
\textbf{PSI TYPE:} Passive \\ 
\textbf{ACTION:} Automatic \\ 
\textbf{RANGE:} Self \\ 
\textbf{DURATION:} Constant \\
Scramble allows the async using the sleight to hide
from another async using the Ego Sense or Omni
Awareness sleights. Apply a +30 modifier to the defending
async’s Opposed Test.

\subsubsection{SENSE BLOCK}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Temp (Action Turns) \\
\textbf{STRAIN MOD:} –1 \\ 
\textbf{SKILL:} Psi Assault \\
Sense Block disables and short circuits one of the target’s
sensory cortices (chosen by the async), interfering
with and possibly negating a specific source of sensory
input for the chosen duration. If the async beats the
target in the Opposed Test, the target suffers a –30
modifier to Perception Tests with that sense equal (doubled
to –60 if the async scores an Excellent Success).


\subsubsection{SPAM}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Temp (Action Turns) \\
\textbf{STRAIN MOD:} +0 \\ 
\textbf{SKILL:} Psi Assault \\
The sleight allows the async to overload and flood one
of the target’s sensory cortices (chosen by the async),
spamming them with confusing and distracting sensory
input and thereby impairing them. If the async
wins the Opposed Test, the target suffers a –10 modifi
er to all tests the duration of the sleight (doubled to
–20 if the async scores an Excellent Success).

\subsubsection{STATIC}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Close \\ 
\textbf{DURATION:} Sustained \\
\textbf{STRAIN MOD:} +0 \\ 
\textbf{SKILL:} None\\
The async generates an anti-psi jamming field, impeding
any use of ranged sleights within their range. All
such ranged sleights suffer a –30 modifier. This sleight
has no effect on self or touch-range sleights.

\subsubsection{SUBLIMINAL}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Instant \\
\textbf{STRAIN MOD:} +2 \\ 
\textbf{SKILL:} Control\\
The Subliminal sleight allows the async to influence
the train of thought of another person by implementing
a single post-hypnotic suggestion into the mind
of the target. If the async wins the Opposed Test, the
recipient will carry out this suggestion as if it was
their own idea. Implanted suggestions must be short
and simple; as a rule of thumb, the gamemaster may
only suggestions encompassed by a short sentence
(for example: “open the airlock,” or “hand over the
weapon”). At the gamemaster’s discretion, the target
may receive a bonus for resisting suggestions that are
immediately life threatening (“jump off the bridge”)
or that violate their motivations or personal strictures.
Suggestions do not need to be carried out immediately,
they may be implanted with a short trigger condition
(“when the alarm goes off, ignore it”).

\subsubsection{THOUGHT BROWSE}
\textbf{PSI TYPE:} Active \\ 
\textbf{ACTION:} Complex \\ 
\textbf{RANGE:} Touch \\ 
\textbf{DURATION:} Sustained \\
\textbf{STRAIN MOD:} –1 \\ 
\textbf{SKILL:} Sense\\
Thought Browse is a less-intrusive form of mind
reading which scans the target’s surface thoughts for
certain “keywords” like a particular word, phrase,
sound, or image chosen by the async. Rather than digging
through the target’s ego as with the Deep Scan
sleight, Thought Browse merely verifies whether a
target has a particular person, place, event, or thing
in mind, which can be used by a savvy investigator to
draw conclusions without the need to invade the mind
directly. Thought Browse may be sustained, allowing
the async to continue scanning the target’s thoughts
over time. The async must beat the target in an Opposed
Test for each scanned item.

\section{PSYCHOSURGERY}
Given the reach of neuroscience in the time of
Eclipse Phase, it is easy to think of the mind as
programmable software, as something that can be
reverse-engineered, re-coded, upgraded, and patched.
To a large degree, this is true. Aided by nanotechnology,
genetics, and cognitive science, neuroscientists
have demolished numerous barriers to understanding
the mind’s structure and functions, and even
made great leaps in unveiling the true nature of
consciousness. Genetic tweaks, neuro-mods, and
neural implants offer an assortment of options for
improving the brain’s capabilities. The transhuman
mind has become a playground—and a battlefield.
Nanovirii unleashed during the Fall infected millions,
altering their brains in permanent ways, with
occasional outbreaks still occurring a decade later.
Cognitive virii roam the mesh, plaguing infomorphs
and AIs, reprogramming mind states. An “infectious
idea” is now a literal term.
In truth, mind editing is not an easy, safe, and
error-proof process—it is difficult, dangerous, and
often flawed. Neuroscience may be light years
ahead of where it was a century ago, but there are
many aspects of the brain and neural functions
that continue to confound and elude even the
brightest experts and AIs. Technologies like nanoneural
mapping, uploading, digital mind emulation,
and artificial intelligence are also comparatively
in their infancy, being mere decades old. Though
transhumanity has a handle on how to make these
processes work, it does not always fully understand
the underlying mechanisms.
Any neurotech will tell you that mucking around
in the mind’s muddy depths is a messy business.
Brains are organic devices, molded by millions of
years of unplanned evolutionary development. Each
is grown haphazardly, loaded with evolutionary
leftovers, and randomly modified by an unlimited
array of life events and environmental factors. Every
mind features numerous mechanisms—cells, connections,
receptors—that handle a dizzying array
of functions: memory, perception, learning, reasoning,
emotion, instinct, consciousness, and more. Its
system of organization and storage is holonomic,
diffused, and disorganized. Even the geneticallymodifi
ed and enhanced brains of transhumans are
crowded, chaotic, cross-wired places, with each
mind storing its memories, personality, and other
defining features in unique ways.
What this means is that though the general architecture
and topography of neural networks can
be scanned and deduced, the devil is in the details.
Techniques used to modify, repair, or enhance one
person’s mind are not guaranteed equal success
when applied to another’s brain. For example, the
process by which brains store knowledge, skills, and
memories results in a strange chaining process where
these memories are linked and associated with other
memories, so attempts to alter one memory can
have adverse affects on other memories. In the end,
minds are slippery and dodgy things, and attempts
to reshape them rarely go as planned.

\subsection{THE PROCESS OF PSYCHOSURGERY}
Psychosurgery is the process of selective, surgical alteration
of a transhuman mind. It is a separate field from
neural genetic modification (which alters genetic code),
neuralware implantation (adding cybernetic or biotech
inserts to the brain or nervous system), or brain hacking
(software attacks on computer brains, neural inserts, and
infomorphs), though they are sometimes combined.
Psychosurgery is almost always performed on a
digital mind-state, whether that be a real-time emulation,
a backup, or a fork. In most cases, the subject’s
mind-state is copied via the same technology and process
as uploading or forking, and run in a simulspace.
The subject need not be willing, and in these cases
the subject’s permissions are restricted. Numerous
psychosurgery simulspace environments are available,
each custom-designed for facilitating specific psychosurgical
goals and programmed with a thorough
selection of psychotherapy treatment options.
The actual process of psychosurgery breaks down
into several stages. First is diagnosis, which can
involve the use of several neuro-imaging techniques
on morphed characters, mapping synaptic connections,
and building a neurochemical model. It can
also involve complete psychological profiling and
psychometric behavioral testing, including personality
tests and simulspace scenario simulations. Digital
mind-states can be compared to records of people
with similar symptoms in order to identify related
information clusters. This analysis is used to plan
the procedure.
The actual implementation of psychosurgical alteration
can involve several methods, depending on
the desired results. Applying external modules to the
mind-state is often the best approach, as it doesn’t
meddle with complicated connections and new inputs
are readily interpreted and assimilated. For treatments,
mental health software patches compiled from
databases of healthy minds are matched, customized,
and applied. Specialized programs may be run to
stimulate certain mental processes for therapeutic
purposes. Before an alteration is even applied, it may
first be performed on a fork of the subject and run at
accelerated speeds to evaluate the outcome. Likewise,
multiple treatment choices may be applied to time-accelerated
forks this way, allowing the psychosurgeon
to test which is likely to work best.
Not all psychosurgery is performed for the subject’s
benefit, of course. Psychosurgery can be used to interrogate
or torture prisoners, erase memories, modify
behavior, or inflict crippling impairments. It is also
sometimes used for legal punishment purposes, in
an attempt to impair criminal activity. Needless to
say, such methods are often brute-forced rather than
fine-tuned, ignoring safety parameters and sometimes
resulting in detrimental side effects.


\begin{quotation}
\textbf{SOLARCHIVE SEARCH: “THE HUMAN COGNOME PROJECT” }
\\
The Human Cognome Project was an
academic research venture to reverse
engineer the human brain, paralleling
in many ways the Human Genome
Project and its success in deciphering
the human genome. The HCP was a
multidisciplinary undertaking, relevant
to biology, neuroscience, psychology,
cognitive science, artificial intelligence,
and philosophy of mind.
Funded and supported by scientific
and corporate entrepreneurs and early
transhumanist groups, the HCP developed
the fundamentals of digitizing an ego
and was a major driving force towards
the first transhumans with elevated
intelligence and brain capacity. The HCP
has also been instrumental in cataloging
transhuman minds and developing
databases of “mind patches” based on
the mind-states of healthy individuals for
treating mental diseases and damage.
Though most HCP data is available to the
public, some argonauts claim that certain
data is held hostage by some hypercorps,
potentially for the development of proprietary
mind-altering technologies.
After the Fall, the remnants of this
project were acquired by the Planetary
Consortium.
\end{quotation}

\subsection{PSYCHOSURGERY MECHANICS}
In game terms, psychosurgery is handled as a Task
Action requiring an Opposed Test. The psychosurgeon
rolls Psychosurgery skill against the target’s WIL x 3.
Apply modifiers as appropriate from the Psychosurgery
Modifiers table.
If the psychosurgeon succeeds and the subject fails,
the psychosurgery is effective and permanent. The
alteration becomes a permanent part of the subject’s
ego, and will be copied when uploaded (and sometimes
when forking).
If both sides succeed but the psychosurgeon rolls
higher, the psychosurgery is effective but temporary. It
lasts for 1 week per 10 points of MoS.
If the subject rolls higher, or if the psychosurgeon
fails their roll, the attempt does not work.
The timeframe listed for psychosurgical procedures
is according to the patient’s subjective point of view.
Since most subjects are treated in a simulspace, time
acceleration may drastically reduce the amount of
real-time such a procedure requires (see Defying Nature’s
Laws, pp. 240–241).


\subsubsection{MENTAL STRESS}
Psychosurgery is a modification to the transhuman
mind, and sometimes to the actual person that resides
in that mind. It is unsurprising then that psychosurgery
places stress on the subject’s mental state and
sometimes even inflicts mental traumas.
Each psychosurgery option lists a Stress Value
(SV) that is inflicted on the subject regardless of
the tests’ success or failure. If the psychosurgeon
achieves an Excellent Success (MoS 30+), this stress
is halved (round down). If the psychosurgeon rolls
a Severe Failure (MoF 30+), the stress is doubled.
Alternately, a Severe Failure could result in unintended
side effects, such as affecting other behaviors,
emotions, or memories.
If a critical success is rolled, no stress is applied at
all. If a critical failure is rolled, however, an automatic
trauma is applied in addition to the normal stress.
Some psychosurgery conditions may also affect the
SV, as noted on the Psychosurgery Modifiers table.


\subsection{ROLEPLAYING MIND EDITS}
Many of the changes incurred by psychosurgery are
nebulous and difficult to pin down with game mechanics.
Alterations to a character’s personality and
mind-state are often better handled as roleplaying factors
anyway. This means that players should make a
real effort to integrate any such mental modifications
into their character’s words and actions, and gamemasters
should ensure that a character’s portrayal
plays true to their mind edits. Some psychosurgical
mods can be reflected with ego traits, while others
might incur modifiers to certain tests or in certain
situations. The gamemaster should carefully weigh a
brain alteration’s effects, and apply modifiers as they
see appropriate.


\subsection{PSYCHOSURGERY PROCEDURES}
The following alterations may be accomplished with
psychosurgery. At the gamemaster’s discretion, other
mind-editing procedures may be attempted, using
these as a guideline.

\begin{quotation}
\textbf{PSYCHOSURGERY MODIFIERS}
\\
\\
\begin{tabular}{lll}
SITUATION & PSYCHOSURGERY TEST MODIFIER & SV MODIFIER \\
Improper Preparatory Diagnosis & –30 & +1 \\
Safety Protocols Ignored & +20 & x2 \\
Simulspace Time Acceleration & –20 & +2 \\
Subject is an AI, AGI, or uplift & –20 & +1 \\
\end{tabular} 
\end{quotation}

\subsubsection{BEHAVIORAL CONTROL}
\textbf{TIMEFRAME:} 1 week \\
\textbf{PM:} Limit/Boost –10; Block/Encourage –20, Expunge/Enforce –30 \\
\textbf{SV:} (1d10 $\div$ 2, round up) \\
Commonly used for criminal rehabilitation, behavioral
control attempts to limit, block, or expunge a
specific behavior from the subject’s psyche. For example,
a murderer may be conditioned against acts of aggression,
or a kleptomaniac might be restricted from
stealing. Some people seek this adjustment willingly,
such as socialite glitterati who restrict their desire to
eat, or an addict who cuts out their craving for a fix.
Behavioral control can also be applied as an unleashing
or reinforcement. A companion may desire
to eliminate their sexual inhibitions, for example, or
a hypercorp exec may boost his commitment to place
work above all else.
A character will simply feel compelled to avoid
a behavior that is limited (perhaps suffering a –10
modifier), but will find it quite difficult to pursue a behavior
that is blocked (requiring a WIL x 3 Test, and
suffering a –20 modifier). They will find themselves
completely incapable of initiating a behavior that is
expunged, and if forced into the behavior will suffer
a –30 modifier and (1d10 $\div$ 2, round up) points of
mental Stress.
Likewise, a character will feel compelled to pursue a
behavior that is boosted, and will find it hard to avoid
engaging in a behavior that is encouraged (requiring a
WIL x 3 Test to avoid). They will have no choice but
to engage in enforced behaviors, and will suffer (1d10
$\div$ 2, round up) points of mental Stress if prevented
from doing so.

\subsubsection{BEHAVIORAL MASKING}
\textbf{TIMEFRAME:} 1 week \\
\textbf{PM:} –20 \\
\textbf{SV:} 1d10 $\div$ 2, round up \\
Given the ability to switch bodies, many security and
law enforcement agencies have resorted to personality
and behavioral profiling as a means of identifying
people even when they resleeve. Though such systems
are far from perfect, someone’s unconscious habits
and quirks could potentially give them away. Characters
who wish to elude identification in this way may
undergo behavioral masking, which seeks to alter and
change the character’s unconscious habits and social
cues. Apply a +30 modifier when defending against
such identification systems and Kinesics Tests.

\subsubsection{DEEP LEARNING}
\textbf{TIMEFRAME:} Skill Learning Time $\div$ 2 \\
\textbf{PM:} +20 \\
\textbf{SV:} 1 \\
Using tutorial programs, memory reinforcement protocols,
conditioning tasks, and deep brain stimulation,
the subject’s learning ability is reinforced, allowing
them to learn new skills more quickly.

\subsubsection{EMOTIONAL CONTROL}
\textbf{TIMEFRAME:} 1 week \\
\textbf{PM:} Limit/Boost –10; Block/Encourage –20, Expunge/Enforce –30 \\
\textbf{SV:} (1d10 $\div$ 2, round up) + 2 \\
Similar to behavioral control, emotional control seeks
to modify, enhance, or restrict the subject’s emotional
responses. Some choose these modifications willingly,
such as limiting sadness in order to be happier, or
encouraging aggression in order to be more competitive.
Mercenaries and soldiers have been known
to expunge fear. Follow the same rules as given for
Behavioral Control.


\subsubsection{INTERROGATION}
\textbf{TIMEFRAME:} Variable (gamemaster discretion; 1 week default) \\
\textbf{PM:} +30 \\
\textbf{SV:} 1d10 \\
Psychosurgery can be used for interrogative purposes
via the application of mental torture and manipulation.
A successful Psychosurgery Test applies a +30
modifier to the Intimidation Test for interrogation.

\subsubsection{MEMORY EDITING}
\textbf{TIMEFRAME:} 1 week (2 weeks adding/replacing) \\
\textbf{PM:} –10 (willing) or –30 (forced) \\
\textbf{SV:} (1d10 $\div$ 2, round up) \\
By monitoring memory recall (forcibly invoked if necessary),
psychosurgeons can identify where memories
are stored in the brain and target them for removal.
Memory storage is complex and diffused, however,
and often linked to other memories, so removing one
memory may affect others (gamemaster discretion).
Adding or replacing memories is a much more complicated
operation and requires that such memories be
copied from someone who has experienced them or
manufactured with XP software. Even when successfully
implanted, fake memories may clash with other
(real) memories unless those are also erased.

\subsubsection{PERSONALITY EDITING}
\textbf{TIMEFRAME:} 1 week \\
\textbf{PM:} Minor –10; Moderate –20, Major –30 \\
\textbf{SV:} (1d10 $\div$ 2, round up) + 3 \\
Possibly the most drastic psychosurgery procedure,
personality editing involves altering the subject’s core
personality traits. The personality factors that may be
modified is almost unlimited, including traits such as
openness, conscientiousness, altruism, extroversion/
introversion, impulsiveness, curiosity, creativity, confidence, sexual orientation, and self-control, among
others. These traits may be enhanced or reduced to
varying degrees. The effect is largely reflected by roleplaying,
but the gamemaster may apply modifiers as
they see fit.

\subsubsection{PSYCHOTORTURE}
\textbf{TIMEFRAME:} Variable \\
\textbf{PM:} +30 \\
\textbf{SV:} 1d10 SV per day \\
Psychotorture is mental manipulation for the simple
intention of causing pain and anguish, reflected in
game terms as mental stress and traumas. Prolonged
torture can lead to serious mental disorders or worse.

\subsubsection{PSYCHOTHERAPY}
\textbf{TIMEFRAME:} Variable \\
\textbf{PM:} +0 \\
\textbf{SV:} 0 \\
Therapeutic psychosurgery is beneficial for characters
suffering from mental stress, traumas, and disorders. A
successful Psychosurgery Test applies a +30 modifier
to mental healing tests, as noted on p. 215.

\subsubsection{SKILL IMPRINTS}
\textbf{TIMEFRAME:} 1 week per +10 \\
\textbf{PM:} +0 \\
\textbf{SV:} 1 per +10 \\
Skill imprinting is the use of psychosurgery to insert
skill-set neural patterns in the subject’s brain, temporarily
boosting their ability. Skill imprints are artificial
boosts, however, degrading at the rate of –10 per day.
No skill may be boosted higher than 60.

\subsubsection{SKILL SUPPRESSION}
\textbf{TIMEFRAME:} 1 day per –10 \\
\textbf{PM:} –10 \\
\textbf{SV:} 1 per +10 \\
Skill suppression attempts to identify where skills are
stored in the brain and then block or remove them.
The subject’s skill is impaired and may be lost entirely.

\subsubsection{TASPING}
\textbf{TIMEFRAME:} 1 day \\
\textbf{PM:} +10 \\
\textbf{SV:} 1 \\
Tasping is the use of deep brain stimulation techniques
to tickle the mind’s pleasure centers. Though this
procedure is often used for therapeutic purposes for
patients suffering from depression or other mental
illnesses, the intent with tasping is to overload the
subject into a prolonged state of almost unendurable
bliss. Such stimulation is highly addictive, however, so
character’s exposed to it for any length of time (over
1 hour, subjective) are likely to pick up the Addiction
trait (p. 148). Some criminal organizations have been
known to use tasping addiction and rewards as a
means of controlling those under their thrall.


\subsection{THE LOST}

\begin{quotation}
$\prec$ begin excerpt $\succ$ \\
PSICLONE Project Quarterly Board Meeting \\
2nd Quarter 8 AF \\
FUTURA Project Conclusion—  \\
Executive Summary Report \\
Prepared by Dr. Amelia Sheppard	
\\
Per request, I have compiled a review of the
Futura Project and its fallout, 5 years after
whistleblowers and intense media attention
forced us to end the project and release the
remaining subjects (dubbed “the Lost” by
popular media).

Futura was a joint initiative spearheaded by
Hanto Genomics and strongly backed by Cognite,
with numerous other partners (complete
list). The project was initially proposed by my
mentor, Dr. Antonio Pascal, whose team had
proven the feasibility of Accelerated Life Experience
Training (ALET) after a series of pilot
studies with two small (N $ \prec $ 1000) samples.
While it is true that these early pilot studies
used both older subjects and a lesser amount
of time dilation, the rationale for the Futura
Project’s ambitious program was justified by a
remarked decrease in transhumanity’s population
due to the Fall, a system-wide stagnant
population growth rate (blamed on various
factors including increased longevity, available
contraception, and rising despair over troubling
times), as well as a desire to move aggressively
into a new technological sector in the hopes of
obtaining a competitive advantage.

Futura began immediately in the wake
of the Fall with an initial seed population
of *** test subjects culled from extant genetic
material and gestated to between 1 week and
6 months after birth. Of these, less than 10\%
were live births from either a surrogate or
genetic birth mother who had perished during
the Fall. The majority came from our Lunar and
Martian labs and were brought to term within
an exowomb.

After the sample was selected, all subjects
were sleeved into our fast-growth futura-brand
biomorph bodies and inducted into customized
simulspace accelerated learning environments.
The project made extensive use of emergent
technologies and techniques culled from recaptured
TITAN facilities, including neogenetic
traits for the futura morphs and time distortion
applications for captive simulspace populations.
Futura ran concurrently on three different
research stations with a combined staff of
2,211 researchers and support personnel and
45 AGIs custom-programmed for expert child
development. Project goals were to raise each
child to a subjective 18 years life experience in
3 years objective time.

Despite omnipresent observation and
real-time adjusting of the simulspace and educational
programming for optimal normality,
somewhere along the way the project suffered
a breakdown in quality assurance and parameter
monitoring that resulted in a near total
failure at empathy modeling. We first observed
this effect 11 months into the project when the
subjects had aged to approximately 6 years of
age. Incidences of animal cruelty and acting
out had spiked, though at that time they remained
within acceptable standards. Over the
next few months this trend continued and Dr.
Pascal authorized the usage of more authoritative
“parenting” to attempt to correct for the
borderline sociopathic behavior that was being
exhibited by 23.19\% of all subjects by the 18-
month mark (9 years of age).

We now know that these changes had
the unintended consequence of suppressing
overt displays of cruelty and violence and
merely taught the majority of subjects how
to conceal their psychoses. It was also at this
time that the first deaths occurred. The initial
waves were thought to be accidents and
both the victim and perpetrator were usually
backed up to a week or so of subjective time.
Post-project analysis now shows that 43.87\%
of our subjects had engaged in at least one
act of premeditated murder by the 24-month
mark (12 years of age) and the counseling
protocols were only training them how to lie
more effectively.

It was at this point that myself and Dr.
Aaron Bharani advocated pulling the plug on
the project and bringing the subjects out to
real time and intensive counseling. Dr. Pascal
vetoed our concerns without ever taking them
to the board. As the project spiraled towards
its conclusion, a fork of Dr. Bharani went public
at the 34 month mark, inciting a firestorm of
controversy. While Dr. Pascal successfully tied
up investigators, hoping to see the project
through to its conclusion, the incident at our
Legacy research station occurred. Initial findings
concluded that one or more of the subjects
had escaped the program and were in fact
responsible for the habitat’s environmental failures
and the thousands of subsequent deaths.

In the face of intense public and private
scrutiny, many of the partners involved in
the project attempted to pull out and even
eliminate all traces of their involvement. In the
resulting chaos, an estimated *** subjects
were quietly released into the system’s general
population. It was only after this occurred that
all known subjects were identified as having
been infected with the Watts-Macleod strain
of the Exsurgent virus, though when and how
this occurred remains troubling and unclear.
Though later orders resulted in all remaining
subjects being euthanized and/or backed up
into cold storage, only of *** the released
subjects were recaptured. Of the rest, *** pursued
sanctuary with sympathetic authorities,
*** went public and submitted themselves
to extensive psychotherapy, *** were killed in
incidents of violence and not resurrected, and
the rest presumably went into hiding.

$\prec$ end excerpt $\succ$ 
\end{quotation}