\chapter{The Mesh}
\label{cha:mesh}

Before the Fall, humanity interfaced with each other through the internet, interconnected networks that served as the technical backbone for the evolving world wide web. While it began as a electronic medium for retrieving information from various sources (replacing even older paper-based infosources), succeeding generations emphasized digital communities and hosted services such as networking sites, wikis, blogs, and folksonomies. These facilitated openness, collaboration, and sharing, thereby laying the groundwork for a modern, interconnected information society. Further stages emphasized wireless interaction, geolocation, and semantic web approaches and achieved quantum leaps in the realm of user interaction with the advent of brain-computer interfaces, augmented reality (AR), virtual reality (VR), and experience playback (XP).

This environment, coupled with the exponential growth of processing power and memory storage, created an evolutionary path for the development of intelligent agents—designed to augment human information processing—that then transformed into artificial intelligences (AIs) in the following decades. While these “weak” AIs did not possess the full range of humancognitive abilities, tended towards overspecialization, and were restrained by programmed limitations, the digital evolution toward artificial general intelligences (AGIs)—”strong” AIs with intelligence capabilities that equaled or exceeded human abilities—could not be halted. From this point it was but a matter of time
 before so-called seed AI would come into existence, machine minds capable of recursive self-improvement, leading to an exponential growth in intelligence. Unfortunately for humanity, the TITANs were the result.

Even before the Fall, however, the internet of old was transforming into something new. Instead of connecting via central servers, users were wirelessly linking to each other, creating a decentralized intermeshed network of handheld devices, personal computers, robots, and electronic devices. Users were online all of the time and connected with everything and everyone around them in a ubiquitous computing environment. This was especially true of those participating in humanity’s expansion into space. Disconnected from the internet due to distance and light-speed communication lags, these users were nevertheless connected with all of the people and objects in their nearby environment or habitat, creating local wireless mesh networks. Thus was the mesh born, taking the place of the old internet of earth, lost during the Fall.


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ep"
%%% End: 
