%%% 018
%%% 019
%%% 020
We humans have a special way of pulling ourselves up 
and kicking ourselves down at the same time. We'd 
achieved more progress than ever before, at the cost 
of wrecking our planet and destabilizing our own 
governments. But things were starting to look up.

With exponentially accelerating technologies, we 
reached out into the solar system, terraforming worlds 
and seeding new life. We re-forged our bodies and 
minds, casting off sickness and death. We achieved 
immortality through the digitization of our minds, 
resleeving from one biological or synthetic body to 
the next at will. We uplifted animals and AIs to be our 
equals. We acquired the means to build anything we 
desired from the molecular level up, so that no one 
need want again.

Yet our race toward extinction was not slowed, and 
in fact received a machine-assist over the precipice. 
Billions died as our technologies rapidly bloomed into 
something beyond control ... further transforming 
humanity into something else, scattering us throughout
the solar system, and reigniting vicious conflicts.
Nuclear strikes, biowarfare plagues, nanoswarms, 
mass uploads ... a thousand horrors nearly wiped 
humanity from existence.

We still survive, divided into a patchwork of restrictive
inner system hypercorp-backed oligarchies and
libertarian outer system collectivist habitats, tribal 
networks, and new experimental societal models. We 
have spread to the outer reaches of the solar system 
and even gained footholds in the galaxy beyond. But 
we are no longer solely ``human'' ... we have evolved 
into something simultaneously more and different—
something \textbf{transhuman}.

\section{Starting Out}

\textit{Eclipse Phase} is a post-apocalyptic roleplaying game 
of transhuman conspiracy and horror. Humans are 
enhanced and improved, but humanity is battered and 
bitterly divided. Technology allows the re-shaping of 
bodies and minds and liberates us from material needs, 
but also creates opportunities for oppression and puts 
the capability for mass destruction in the hands of everyone
Many threats lurk in the devastated habitats of
the Fall, dangers both familiar and alien.

\subsection{What Is A Roleplaying Game?}

Have you ever read a book or seen a movie or a television
show where a character does something really
stupid, like heading into a basement at night when the 
character knows the serial killer is around? The whole 
time, you're thinking: ``I wouldn't walk down those 
creepy stairs to the dark basement, especially without a 
flashlight. I'd do X, Y, or Z instead!'' Since you're in the 
passenger's seat for the plot you're reading or watching, 
however, you simply have to sit back and let it unfold. 

What if you could take hold of the driver's seat? 
What if you could take the plot in the direction \textit{you'd} 
choose? That is the essence of a roleplaying game.

A roleplaying game (or RPG, for short) is part improvisational
theater, part storytelling, and part game.
A single person (the gamemaster) runs the game for 
a group of players that pretend to be characters in a 
fictitious world. The world could be a mystery game 
set in the 1920s that takes you adventuring around 
the globe, a fantasy realm inhabited by dragons 
and trolls and sword-wielding barbarians, or a science
fiction setting with aliens and spaceship and
world-crushing weaponry. The players pick a setting 
that they find cool and want to play in. The players 
then craft their own characters, providing a detailed 
%%% 021
history and personality to bring each to life. These 
characters have a set of statistics (numerical values) 
that represent skills, attributes, and other abilities. 
The gamemaster then explains the situation in which 
the characters find themselves. The players, through 
their characters, interact with the storyline and each 
others' characters, acting out the plot. As the players 
roleplay through some scenarios, the gamemaster will 
probably ask a given player to roll some dice and the 
resulting numbers will determine the success or failure 
of a character's attempted action. The gamemaster 
uses the rules of the game to interpret the dice rolls 
and the outcome of the character's actions.

As a group exercise, the players control the storyline 
(the adventure), which evolves much like any movie or 
book but within the flexible plot created by the gamemaster
This gamemaster plot provides a framework
and ideas for potential courses of action and outcomes, 
but it is simply an outline of what might happen—it 
is not concrete until the players become involved. If 
you don't want to walk down those stairs, you don't. 
If you think you can talk yourself out of a situation 
in place of pulling a gun, then try and make it happen. 
The script of any roleplaying session is written by the 
players, and the story, based upon the character's actions
and their responses to the events of the plot, will
constantly change and evolve.

The best part is that there is no ``right'' or ``wrong'' 
way to play an RPG. Some games may involve more 
combat and dice rolling-related situations, where other 
games may involve more storytelling and improvised 
dialogue to resolve a situation. Each group of players 
decides for themselves the type and style of game they 
enjoy playing!
%%% 022

\subsection{But How Do You Actually Play?}

To play a game of \textit{Eclipse Phase}, you need the following:
•  A group of players and a place to meet (real life 
or online!)
•  One player to act as the gamemaster
•  The contents of this book
• Something for everyone to take notes with notepads
laptops, whatever!)
•  Two 10-sided dice per player (or a digital 
equivalent)
•  Imagination

\subsubsection{A Group Of Players And A Place To Meet}

While roleplaying games are flexible enough to allow 
any number of people, most gaming groups number 
around four to eight players. That number of people 
brings a good mix of personalities to the table and 
ensures great cooperative play.

Once a group of players have determined to play 
\textit{Eclipse Phase,} they'll need to designate someone as 
the gamemaster (see below). Then they'll need to determine
a time and place to meet.

Most roleplaying groups meet once a week at a regularly
scheduled time and place: 7:00 PM, Thursday
night, Rob's house, for example. However, each group 
determines where, how they'll play, and how often. 
One group may decide they can only get together 
once a month, while another group is so excited to 
dive into the story potential of \textit{Eclipse Phase} that 
they want to meet twice a week (they decide to rotate 
between their houses, though, so as not to overload 
a particular player). If a group is lucky enough to 
have a favorite local gaming store that supports in-store
play, the group might meet there. Other gaming
groups meet in libraries, common rooms at their 
school, bookstores that have generously-sized reading
rooms,'' quiet restaurants, and so on. Whatever
fits for your gaming group, make it work!

When getting together for a game, most RPGs 
use the phrase ``gaming session.'' The length of each 
gaming session is completely dependent upon the consensus
of the playing group, as well as the limitations
of the locale where they're playing. The particular 
story that unfolds in a given session can also impact a 
session's length. If playing in a game store, the group 
may only have a four-hour slot and the gamemaster 
and group may have determined—through several sessions
of play—that this is a perfect time frame to enjoy
the story they're participating in each week. Another 
group, however, may want an even shorter length of 
time. Yet another group may decide that while they'll 
usually do four-hour sessions, once a month they'll 
set aside an entire Saturday for a great all-day gaming 
session. Players will need to dive in and start playing
and be flexible to decide what will provide the
ultimate enjoyment for their gaming group.

While the camaraderie of a shared experience of 
playing face-to-face with a group of friends remains 
%%% 023
the strength of roleplaying games, groups need not 
confine themselves to a single mode of play. There are 
myriad options that can be used. Email, instant messages
message boards, video chats, phone/voip calls,
text messages, wikis, (micro-)blogs: any and all of 
these can be utilized to play the game without having 
warm bodies in seats directly across the table from 
one another.

Finally, when playing groups meet for the first time, 
they should generate their characters (as opposed to 
generating characters by themselves). While a gaming 
group can decide to generate characters individually, 
often it is far easier once the players are together. This 
allows those more experienced in roleplaying games to 
help those new to RPGs. Even more important, it enables
the entire group to tailor the characters so there
is not too much overlap in capabilities and style. After 
all, with the wealth of character opportunities available
you don't want to show up at the table with an
almost identical character to the player next to you.

\subsubsection{The Gamemaster}

Once a group has been organized, someone needs to 
step up and take the reins of the gamemaster. Some 
groups have a single gamemaster that runs all their 
gaming sessions month after month. Other groups 
rotate a gamemaster, with a single gamemaster running
a given portion of the unfolding story for several
sessions before handing the work off to another
player. Once again, the participants should be flexible. 
Some groups may have the perfect person who loves 
the work involved and is more than willing to run 
session after session, while other groups may decide 
that they all want to take turns both as the gamemaster
and as players.

The gamemaster controls the story. They keep track 
of what is supposed to happen when, describes events 
as they occur so that the players (as characters) can 
react to them, keep track of other characters in the 
game (referred to as non-player characters, or NPCs), 
and resolve attempts to take action using the game 
system. The game system comes into play when characters
seek to use their skills or otherwise do something
that requires a test to see whether or not they
succeed. Specific rules are presented for situations that 
involve rolling dice to determine the outcome (see 
\textit{Game Mechanics,} p. \textbf{112}).

The gamemaster describes the world as the characters
see it, functioning as their eyes, ears, and other
senses. Gamemastering is not easy, but the thrill of 
creating an adventure that engages the other players' 
imaginations, testing their gaming skills and their 
characters' skills in the game world, makes it worthwhile
Posthuman Studios and Catalyst Game Labs
will follow the publication of \textit{Eclipse Phase} with supporting
supplements and adventures to help this process
along, but experienced gamemasters can always
adapt the game universe to suit their own styles. In 
fact, since \textit{Eclipse Phase} is published under a Creative 
Commons License (see p. 5), players are encouraged 
to tailor the universe to their style of play and also 
to share that with other players. You never know 
when a specific choice you've made in the running of 
a campaign is exactly what another gamemaster and 
his group is looking for.

\subsubsection{The Contents Of This Book}

Whether you have purchased the print or electronic 
version, this book is specifically organized to present 
the information you need to know to start telling 
\textit{your} stories in the \textit{Eclipse Phase} universe. Below 
you'll find a summary of each chapter of the book.

\textbf{A Time of Eclipse:} A comprehensive history and setting
fully describes the \textit{Eclipse Phase} universe and how
humanity transitioned from here to there. See p. 30.

\textbf{Game Mechanics:} The player's desired actions 
become reality within the universe through quick and 
easy-to-use game mechanics. See p. 112.

\textbf{Character Creation and Advancement:} Creating a 
unique character can be one of the most enjoyable 
experiences of roleplaying. Even more rewarding is 
watching that character evolve and grow across numerous
gaming sessions, far beyond anything your
imagination first envisioned. See p. 128.

\textbf{Skills:} Beyond a character's innate abilities, their 
skills are what set them apart. This is what your 
character knows and what they know how to do. See 
p. 170.

\textbf{Action and Combat:} What is a dramatic story without
action and violence? When words fail, weapons
will blaze. See p. 186.

\textbf{Mind Hacks: }The unusual possibilities offered by 
psi abilities and mental reprogramming. See p. 216.

\textbf{The Mesh: }The all-pervasive nature of the mesh 
ensures that it is a key element to any story telling. 
See p. 234.

\textbf{Accelerated Future:} The wonders of advanced technologies
and how they work. See p. 266.

\textbf{Gear:} Personal enhancements, weapons, robots, and 
everything else in between. See p. 294.

\textbf{Game Information:} The quintessential set of insider 
secrets for gamemasters. See p. 350.

\subsubsection{Taking Notes}

Whether a gamemaster or player, you'll need a way to 
track information. Players will be generating characters
and making changes to those characters from session
to session. Meanwhile, the gamemaster will have
a host of information to track: notes on how the story 
is unfolding due to player character interaction that 
you'll need to fold into next week's session; changes 
to NPCs; changes to player characters that the players 
are not yet aware off (such as a character has been 
mind hacked but doesn't yet know it); and so on.

Additionally, some groups enjoy a synopsis of 
each session that can be compiled and read at a 
later time in order to enjoy and share their exploits, 
just as you might fileshare clips from your favorite 
video game to show off your skill in taking the 
bad guy down (traditionally this has been called 
%%% 024
``bluebooking''). This can be particularly useful if a 
player was unable to attend a given session, providing
a quick re-cap that they can read before attending
the next gaming session and thus avoiding
a bog-down up-front as that player tries to catch 
up on current events in the game. The session scribe 
can be a shared responsibility or assigned, all based 
upon what a given playing group finds works best 
for them. Likewise, some gaming groups audio-record
their entire game session, both for later
reference and for ``actual play'' podcasts.

The old standard of a pencil and paper still works 
wonders. A host of additional technologies, however
provide many new options for players. From
a text file on a laptop to a shared wiki, the ability 
to track large amounts of information in a quick 
and useful fashion—while simultaneously making 
appropriate information available to each player 
from session to session—significantly decreases 
how much time everyone needs to spend tracking 
information. That time can now be redirected into 
the enjoyment of participating in a great story.

\subsubsection{Dice}

As described in the \textit{Game Mechanics} section (p. \textbf{112}), 
two ten-sided dice are required to play \textit{Eclipse Phase}. 
While most players enjoy the feel of tossing dice onto 
a table, there are many other mechanisms for rolling 
two ten-sided dice to achieve a 00 to 99 result. Players 
who make heavy use of any online technologies for 
game play—such as using online chatting or video 
blogging—should find it easy to track down and 
implement a quick dice-rolling program.

\subsubsection{Imagination}

All too often, it's easy for someone looking at an 
RPG to be intimidated. So many concepts to grasp, 
so many ideas that seem overwhelming. Just as described
under \textit{What is a Roleplaying Game?,} however
how often have you read a book or watched
that movie and decided that you would have done 
it better? That's your imagination at work. Just dive 
in and you'll be amazed at how quickly you can immerse
yourself in the \textit{Eclipse Phase} universe. Soon
you'll be spinning stories with the best of them.

Also, don't forget to tap your resources. Your 
gaming group is your best resource. What's going on, 
ideas for how to handle a situation, or how to take 
on a bad guy: these are just some of the things that 
can and should be discussed by the gaming group 
in between sessions, and each is an opportunity to 
strengthen your imagination.

Another resource is simply watching TV or reading
a good book. Pay attention to how the story is
put together, how the characters are built, and how 
the plot unfolds. Push your imagination and soon 
you'll be figuring out subplots and who the bad guy 
is long before it's revealed. Knowing how a story is 
put together enables you to put together your own 
stories during each gaming session.

Finally, eclipsephase.com is the official site for 
\textit{Eclipse Phase.} If you have questions about the 
game or want to see how another group of players 
handles a given situation, post on the forums. The 
online community can be just as helpful and enjoyable
as a local gaming group.

\subsection{What Do Players Do?}

The players can take on a variety of roles in \textit{Eclipse }
\textit{Phase}. Due to advances in digital mind emulation 
technology, uploading, and downloading into new 
morphs (physical bodies, biological or synthetic), it is 
possible to literally be a new person from session to 
session. With bodies taking on the role of gear, players 
can customize their forms for the task at hand.

\subsubsection{The Default Campaign}

In the default story (also known as ``campaign setting'')
every player character is a ``sentinel,'' an
agent-on-call (or potential recruit) for a shadowy 
network known as ``Firewall.'' Firewall is dedicated 
to counteracting ``existential risks''—threats to the 
existence of transhumanity. These risks can and do 
include biowar plagues, nanotech swarm outbreaks, 
nuclear proliferation, terrorists with WMDs, net-breaking
computer attacks, rogue AIs, alien encounters
and so on. Firewall isn't content to simply
counteract these threats as they arise, of course, 
so characters may also be sent on information-gathering
missions or to put in place pre-emptive
%%% 025
or failsafe measures. Characters may be tasked to 
investigate seemingly innocuous people and places 
(who turn out not to be), make deals with shady 
criminal networks (who turn out not to be trustworthy
or travel through a Pandora's Gate wormhole
to analyze the relics of some alien ruin (and see if 
the threat that killed them is still real). Sentinels 
are recruited from every faction of transhumanity; 
those who aren't ideologically loyal to the cause are 
hired as mercenaries. These campaigns tend to mix 
a bit of mystery and investigation with fierce bouts 
of action and combat, also stirring in a nice dose of 
awe and horror.

\subsubsection{Alternate Campaigns}

When they're not saving the solar system, sentinels are 
free to pursue their own endeavors. The gamemaster 
and players can use this rulebook to generate any 
type of story they wish to tell. However, the following 
examples provide a brief look at the most obvious opportunities
for adventure in \textit{Eclipse Phase}.

After each campaign variant below, a list of archetypes
for \textit{Eclipse Phase} are provided in parenthesis
Archetypes are the names applied to the most
common character types featured in those scenarios. 
For example, in a traditional detective story, the archetypes
would be the Detective, the Damsel In Distress
the Hard-bitten Cop, and so on. In a cowboy
movie, the archetypes would be the Gunfighter, the 
Bartender, the Marshal, the Indian Brave, and so 
on. Players will note that some archetypes fit into 
multiple story settings. The character creation system 
(p. 128) allows players to create any of the suggested 
archetypes. Just as roleplaying games are designed 
for players to build their own stories, however, these 
archetypes are just suggestions and players can mix 
and match how they will.

\textbf{Salvage and Rescue/Retrieval Ops:} The Fall left 
two worlds and numerous habitats in ruins—but 
these devastated cities and stations contain untold 
riches for those who are brave and foolhardy 
enough. Potential hauls include: weapon systems; 
physical resources; lost databanks; left-behind uploads
of friends, family, or important people; new
technologies developed and lost in the brief singularity
takeoff; valued heirlooms of immortal oligarchs;
and much more. Outside of these once-inhabited 
realms, space itself is a big place and lots of people 
and things get lost out there. Some need to be saved 
and some are beyond saving. This option lets players 
explore the unknown or seek out specific targets on 
contract. (Archeologist/Scavenger/Pirate/Free Trader/
Smuggler/Black Marketeer)

\textbf{Exploration:} There are plenty of opportunities 
to be had as an explorer, colonist, or long-range 
scout—perhaps even as one of the few lucky or suicidal
individuals who explore through an untested
Pandora's Gate. Even the Kuiper Belt, on the fringe 
of our solar system, is still sparsely explored; there 
may be riches and mysteries still to be found. Many 
dangers also lurk in odd corners of the system, from 
isolationist posthuman factions to secretive criminal 
cartels, as well as pirates, aliens, and others wishing
to remain out of sight. (Explorer/Archeologist/
Scavenger/Singularity Seeker/Techie/Medic)

\textbf{Trade:} While the majority of inner system trade is 
controlled by sleek hypercorporations, many of the 
smaller or more independent stations rely on small 
traders. In the post-scarcity outer system, trade takes 
on a different form, with information, favors, and 
creativity serving as currency among those who no 
longer want for anything due to the availability of 
cornucopia machines. (Free Trader/Smuggler/Black 
Marketeer/Pirate)

\textbf{Crime:} The patchwork of city-state habitats and 
widely varying laws throughout the system create 
ample opportunity for those who would make a 
living from this situation. Black market commodities
and activities include infomorph-slave trading,
pleasure pod sex industries, data brokerage and 
theft, extracting/smuggling advanced technologies 
and scientists, political/economic espionage, assassination
drug and XP dealing, soul-trading, and
much more. Whether as an independent or part of 
an organized criminal element, there are always opportunities
for those with a thirst for adventure or
profit and questionable morals. (Criminal/Smuggler/
Pirate/Fixer/Black Marketeer/Genehacker/Hacker/
Covert Ops)

\textbf{Mercenaries:} The constant maneuvering of 
ideologically-driven factions, the squabbling over 
contested resources, and the rush to colonize new 
exoplanets beyond the Pandora Gates all spark new 
conflicts on a regular basis. Some of these simmer 
and seethe as low-intensity conflicts for years, occasionally
flaring into raids and clashes. Others break
out into all-out warfare. Women and men willing 
to bear arms for credits are always in demand for 
good wages. Players can engage in commando and 
military campaigns in habitats, between the stars, or 
in hostile planetary environments. (Merc/Security 
Specialist/Fixer/Bounty Hunter/Ex-Cop/Medic)

\textbf{Socio-Political Intrigue:} The corporations and 
political factions that span the solar system do not 
always play nice with each other, but neither is it 
wise for them to openly confront each other except 
under extreme circumstances. Many battles are 
fought with diplomacy and political maneuvering, 
using words and ideas more potent than weapons. 
Even within factions, social cliques can compete 
ruthlessly, or heated class conflicts can come to 
a boil, tearing a society apart from within. In this 
campaign, the players can start as pawns of some 
entity who rise through the ranks as they become 
more enmeshed in the intrigues of their sponsor, play 
a group of ambassadors and spies stationed in the 
opposition's capital, or can play a group of activists
and radicals fighting for social change. (Politico/
Socialite/Covert Ops/Hacker/Security Specialist/
Journalist/Memeticist)
%%% 026

\subsection{Where Does It Take Place?}

While  \textit{Eclipse Phase} is set in the not-too-distant 
future, the changes that have taken place due to the 
advancements of technology have transformed the 
Earth and its inhabitants almost beyond recognition. 
As players dive into the universe, they'll generally 
encounter one of the following settings.

\subsubsection{Humanity'S Habitats}

The Earth has been left an ecologically-devastated 
ruin, but humanity has taken to the stars. When 
Earth was abandoned, so too were the last of the 
great nation-states; transhumanity now lacks 
a single unifying governing body and is instead 
subject to the laws and regulations of whomever 
controls a given habitat.

The majority of transhumanity is confined to 
orbital habitats or satellite stations scattered 
throughout the Sol system. Some of these were 
constructed from scratch in the orbit or Lagrange 
points of planetary bodies, others have been hewn 
out of solid satellites and large asteroids. These stations
have myriad purposes from trade to warfare,
espionage to research.

Mars continues to be one of transhumanity's 
largest settlements, though it too, suffered heavily 
during the Fall. Numerous cities and settlements 
remain, however, though the planet is only partially 
terraformed. Venus, Luna, and Titan are also home 
to significant populations. Additionally, there are a 
small number of colonies that have been established 
on exoplanets (on the other side of the Pandora 
Gates) with environments that are not too hostile 
towards humanity.

Some transhumans prefer to live on large colony 
ships or linked swarms of smaller spacecraft, moving 
nomadically. Some of these rovers intentionally exile 
themselves to the far limits of the solar system, far 
from everyone else, while others actively trade from 
habitat to habitat, station to station, serving as 
mobile black markets.

\subsubsection{The Great Unknown}

The areas of the galaxy that have felt the touch of 
humanity are few and far between. Lying betwixt 
these occasional outposts of questionable civilization 
are mysteries both dangerous and wonderful. Ever 
since the discovery of the Pandora Gates, there has 
been no shortage of adventurers brave or foolhardy 
enough to strike out on their own into the unknown 
regions of space in hopes of finding more alien artifacts
or even establishing contact with one of the
other sentient races in the universe.

\subsubsection{The Mesh}

While not a ``setting'' in the traditional sense, as 
the sections describe above, the computer networks
known as the ``mesh'' are all-pervasive. This
ubiquitous computing environment is made possible
thanks to advanced computer technologies and
nanofabrication that allow unlimited data storage 
and near-instantaneous transmission capacities. With 
micro-scale, cheap-to-produce wireless transceivers so 
abundant, literally everything is wirelessly connected 
and online. Via implants or small personal computers
characters have access to archives of information
that dwarf the entire 21st-century internet and sensor 
systems that pervade every public place. People's 
entire lives are recorded and lifelogged, shared with 
others on one of numerous social networks that link 
everyone together in a web of contacts, favors, and 
reputation systems.

\subsection{Ego Vs. Morph}

The distinction between ego (your mind and personality
including memories, knowledge, and skills)
and morph (your physical body and its capabilities) 
is one of the defining characteristics of \textit{Eclipse }
\textit{Phase}. A good understanding of the concept right 
up front will allow players a glimpse at all the story 
possibilities out of the gate.

Your body is disposable. If it gets old, sick, or too 
heavily damaged, you can digitize your consciousness
and download it into a new one. The process
isn't cheap or easy, but it does guarantee you effective
immortality—as long as you remember to
back yourself up and don't go insane. The term 
\textit{morph }is used to describe any type of form your 
mind inhabits, whether a vat-grown clone sleeve, 
a synthetic robotic shell, a part-bio/part-synthetic 
``pod,'' or even the purely electronic software state 
of an infomorph.

A character's morph may die, but the character's 
ego may live on, assuming appropriate backup 
measures have been taken. Morphs are expendable, 
but your character's ego represents the ongoing, 
continuous life path of your character's mind and 
personality. This continuity may be interrupted by 
an unexpected death (depending on how recently 
the backup was made), but it represents the totality 
of the character's mental state and experiences.

Some aspects of your character—particularly skills, 
along with some stats and traits—belong to your 
character's ego and so stay with them throughout 
the character's development. Some stats and traits, 
however, are determined by morph, as noted, and so 
will change if your character leaves one body and 
takes on another. Morphs may also affect other skills 
and stats, as detailed in the morph description.

\subsection{Where To Go From Here?}

Now that you know what this game is about, we suggest
that you next read the \textit{Time of Eclipse} chapter
(p.  30), to get a feel for the game's default setting 
(which you are, of course, free to change to suit your 
whims). Then read the \textit{Game Mechanics} chapter 
(p. 112) to get a grasp of the rules. After that, you 
can move on to \textit{Character Creation and Advancement}
(p. 128) and create your first character!
%%% 027

\subsection{Terminology}

\textit{Eclipse Phase} uses a host of jargon to simply convey 
the numerous concepts covered within the pages of 
this book. While not all-inclusive, this list of terminology
will allow players to quickly acclimate themselves
for their journey into \textit{Eclipse Phase}. If you read something
and are confused, don't worry. These concepts
are fully explained in later sections of this book.
Note that several of the words on this list are 
standard scientific terms, often used in astronomy. As 
\textit{Eclipse Phase} attempts to remain as close to ``hard science
as possible—while allowing players to interact
with the great stories waiting to unfold—such terms 
are used liberally.
\textbf{ Aerostat:} A habitat designed to ﬂ oat like a balloon in a 
planet's upper atmosphere.
\textbf{ AF:} After the Fall (used for reference dating).
\textbf{ AGI:} Artificial General Intelligence. An AI that has cognitive
faculties comparable to that of a human or higher.
Also known as ``strong AI'' (differentiating from more 
specialized ``weak AI''). See also ``seed AI.''
\textbf{ AI:} Artificial Intelligence. Generally used to refer to weak 
AIs; i.e., AIs that do not encompass (or in some cases, are 
completely outside of) the full range of human cognitive 
abilities. AIs differ from AGIs in that they are usually 
specialized and/or intentionally crippled/limited.
\textbf{ Anarchist:} Someone who believes government is 
unnecessary, that power corrupts, and that people should 
control their own lives through self-organized individual 
and collective action.
\textbf{ Arachnoid:} A spider-like robotic synthmorph.
\textbf{ Argonauts:} A faction of techno-progressive scientists 
that promote responsible and ethical use of technology.
\textbf{ AR:} Augmented Reality. Information from the mesh universal
data network) that is overlaid on your real-world
senses. AR data is usually entoptic (visual), but can also 
be audio, tactile, olfactory, kinesthetic (body awareness), 
emotional, or other types of input.
\textbf{ Async:} A person with psi abilities.
\textbf{ AU:} Astronomical unit. The distance between the Earth 
and the Sun, equal to 8.3 light minutes, or about 150 
million kilometers. 
\textbf{ Autonomists:} The alliance of anarchists, Barsoomians, 
Extropians, scum, and Titanians.
\textbf{ Barsoomian:} A rural Martian, typically resentful of 
hypercorp control.
\textbf{ Basilisk Hack:} An image or other sensory input that 
affects the brain's visual cortex and pattern recognition 
abilities in such a way as to cause a glitch and possibly 
exploit it and rewrite neural code.
\textbf{ Beehive:} A microgravity habitat made from a tunneled-out
asteroid or moon.
\textbf{ BF:} Before the Fall (used for reference dating).
\textbf{ Bioconservative:} An anti-technology movement that argues 
for strict regulation of nanofabrication, AI, uploading, forking, 
cognitive enhancements, and other disruptive technologies.
\textbf{ Biomorph:} A biological body, whether a ﬂ at, splicer, 
genetically engineered transhuman, or pod.
\textbf{ Body Bank:} A service for leasing, selling, acquiring, or 
storing a morph. Aka dollhouse, morgue.
\textbf{ Bots:} Robots. AI-piloted synthetic shells.
\textbf{ Bracewell Probe:} A type of autonomous monitoring deep-space
probe meant to make contact with alien civilizations.
\textbf{ Brinkers:} Exiles who live on the fringes of the system, 
as well as other isolated and well-hidden nooks and 
crannies. Also called isolates, fringers, drifters.
\textbf{ Case:} A cheap, common, mass-produced synthetic shell.
\textbf{ Chimeric:} Transgenic, containing genetic traits from 
other species.
\textbf{ Circumjovian:} Orbiting Jupiter.
\textbf{ Circumlunar:} Orbiting the Moon.
\textbf{ Circumsolar:} Orbiting the Sun.
\textbf{ Cislunar:} Between the Earth and the Moon.
\textbf{ Clade:} A species or group of organisms with common 
features. Used to refer to transhuman subspecies and 
morph types. 
\textbf{ Cole Bubble:} A habitat made from a hollowed-out 
asteroid or moon, spun for gravity.
\textbf{ Cornucopia Machine:} A general-purpose nanofabricator.
\textbf{ Cortical Stack:} An implanted memory cell used for ego 
backup. Located where the spine meets the skull; can be 
cut out.
\textbf{ Cyberbrain:} An artificial brain, housing an ego. Used in 
both synthmorphs and pods.
\textbf{ Darkcast:} Illegal and black market farcasting and 
egocasting services.
\textbf{ Domain Rules:} The rules that govern the reality of a 
virtual reality simulspace.
\textbf{ Drone:} A robot controlled through teleoperation (rather 
than directly via onboard AI).
\textbf{ Ecto:} Personal mesh devices that are ﬂ exible, stretchable, 
self-cleaning, translucent, and solar-powered. From 
ecto-link (external link).
\textbf{ Ego:} The part of you that switches from body to body. 
Also known as ghost, soul, essence, spirit, persona.
\textbf{ Egocasting:} Term for sending egos via farcasting.
\textbf{ Entoptics:} Augmented-reality images that you ``see'' in 
your head. (``Entoptic'' means ``within the eye.'')
\textbf{ ETI:} Extraterrestial intelligence. The term Firewall uses 
to refer to the god-like post-singularity alien intelligence 
theorized to be responsible for the Exsurgent virus.
\textbf{ Exalts:} Genetically-enhanced humans (between genefixed
and transhumans). Aka genefreaks, the ascended,
the elevated.
\textbf{ Exoplanet:} A planet in another solar system.
\textbf{ Exsurgent:} Someone infected by the Exsurgent virus.
\textbf{ Exsurgent Virus:} The multi-vector virus created by 
an unknown ETI and seeded throughout the galaxy in 
Bracewell probes. The Exsurgent virus is self-morphing and 
can infect both computer systems and biological creatures.
\textbf{ Extrasolar:} Outside the solar system.
%%% 028
\textbf{ Factors:} The alien ambassadorial race that deals with 
transhumanity. Also called Brokers.
\textbf{ The Fall:} The apocalypse; the singularity and wars that 
nearly brought about the downfall of transhumanity.
\textbf{ Farcasting:} Intrasolar communication utilizing classical 
communication technologies (radio, laser, etc.) and 
quantum teleportation.
\textbf{ Farhauler:} Long distance space shipper.
\textbf{ Firewall:} The secret cross-faction conspiracy that works 
to protect transhumanity from ``existential threats'' (risks 
to transhumanity's continued existence).
\textbf{ Flatlander:} Someone born or used to living on a planet 
or moon with gravity.
\textbf{ Flats:} Baseline humans (not genetically modified). Also 
called norms.
\textbf{ Flexbot:} A shape-changing synthmorph also capable of 
joining together with other ﬂ exbots in a modular fashion 
to create larger shapes.
\textbf{ Forking:} Copying an ego. Not all forks are full copies. 
AKA backups.
\textbf{ FTL:} Faster-Than-Light.
\textbf{ Fury:} A transhuman combat morph.
\textbf{ Gatecrashers:} Explorers who take their chances using a 
Pandora gate to go somewhere previously unexplored.
\textbf{ Genehacker:} Someone who manipulates genetic code to 
create genetic modifications or even new life.
\textbf{ Ghost:} A transhuman combat morph optimized for 
stealth and infiltration.
\textbf{ Ghost-riding:} The act of carrying an infomorph in a 
special implant module inside your head.
\textbf{ Greeks:} Trojan asteroids or moons that share the same 
orbit as a larger planet or moon, but are 60 degrees 
ahead in the orbit at the L4 Lagrange point. The term 
Greeks normally refers to the asteroids orbiting around 
Jupiter's L4 point. See also ``Trojans.''
\textbf{ Habtech:} A habitat technician.
\textbf{ Heliopause:} The point where pressure from the solar wind 
balances with the interstellar medium (about 100 AU out).
\textbf{ Hibernoid:} A transhuman modified for hibernation, for 
extensive travel in space.
\textbf{ Iceteroid:} An asteroid made from mostly ice rather than 
rock or metals.
\textbf{ Iktomi:} The name given to the mysterious alien race 
whose relics have been found beyond the Pandora Gates.
\textbf{ Indentures:} Indentured servants who have contracted 
their labor to a hypercorp or other authority, usually in 
exchange for a morph.
\textbf{ Infolife:} Artificial general intelligences and seed AIs.
\textbf{ Infomorph:} A digitized ego; a virtual body. Also known 
as datamorphs, uploads, backups.
\textbf{ Infugee:} ``Infomorph refugee,'' or someone who left 
everything behind on Earth during the Fall—even their 
own body.
\textbf{ Isolates:} Those who live in isolated communities far 
outside the system (in the Kuiper Belt and Oort Cloud); 
aka outsters, fringers.
\textbf{ Jamming:} The act of ``becoming'' a teleoperated drone 
thanks to XP technology. Also sometimes applied to 
accesing the real-time XP feed from lifeloggers and others.
\textbf{ Kuiper Belt:} A region of space extending from Neptune's 
orbit out to about 55 AU, lightly populated with asteroids, 
comets, and dwarf planets.
\textbf{ Lagrange Point:} One of five areas in respect to a 
small planetary body orbiting a larger one in which the 
gravitational forces of those two bodies are neutralized. 
Lagrange points are considered stable and ideal locations 
for habitats.
\textbf{ Lifelog:} A recording of one's entire life experience, made 
possible due to near unlimited computer memory.
\textbf{ Lost Generation:} In an effort to repopulate post-Fall, a 
generation of children were reared using forced-growth 
methods. The results were disastrous: many died or went 
insane, and the rest were stigmatized.
\textbf{ Main Belt:} The main asteroid belt, a torus ring orbiting 
between Mars and Jupiter.
\textbf{ Meme:} A viral idea.
\textbf{ Mentons:} Transhumans optimized for mental and 
cognitive ability.
\textbf{ Mercurials:} The non-human sentient elements of the 
transhuman ``family,'' including AGIs and uplifted animals.
\textbf{ Mesh:} The omnipresent wireless mesh data network. 
Also used as a verb (to mesh) and adjective (meshed or 
unmeshed).
\textbf{ Mesh ID:} The unique signature attached to one's mesh 
activity.
\textbf{ Microgravity:} Zero-g or near weightless environments.
\textbf{ Mist:} The clouds of AR data that sometimes fog up your 
perception/displays.
\textbf{ Morph:} A physical body. Aka suit, jacket, sleeve, shell, form.
\textbf{ Muse:} Personal AI helper programs.
\textbf{ Nanobot:} A nano-scale machine.
\textbf{ Nano-ecology:} Pro-tech ecological movement.
\textbf{ Nanoswarm:} A mass of tiny nanobots unleashed into an 
environment.
\textbf{ Neo-Avians:} Uplifted ravens and gray parrots.
\textbf{ Neogenesis:} The creation of new life forms via genetic 
manipulation and biotechnology.
\textbf{ Neo-Hominids:} Uplifted chimpanzees, gorillas, and 
orangutans.
\textbf{ Neotenics:} Transhumans modified to retain a child-like 
form.
\textbf{ Novacrab:} A pod created from genetically-engineered 
spider crab stock.
\textbf{ Olympian:} A transhuman biomorph modified for athleticism
and endurance.
\textbf{ O'Neill Cylinder:} A soda-can shaped habitat, spun for 
gravity.
\textbf{ Oort Cloud:} The spherical ``cloud'' of comets that 
surrounds the solar system out to about one light-year 
from the sun. 
%%% 029
\textbf{ PAN:} Personal area network. The network created when 
you slave all of your minor personal electronics to your 
ecto or mesh inserts.
\textbf{ Pandora Gates:} The wormhole gateways left behind by 
the TITANs.
\textbf{ Pods:} Mixed biological-synthetic morphs. Pod clones are 
force-grown and feature computer brains. Also known as 
bio-bots, skinjobs, replicants. From ``pod people.''
\textbf{ Posthuman:} A human or transhuman individual or 
species that has been genetically or cognitively modified 
so extensively as to no longer be human (a step beyond 
transhuman). Aka parahuman.
\textbf{ Prometheans:} A group of transhuman-friendly seed AIs 
that were created by the Lifeboat Project (precursors to 
the argonauts) years before the TITANs became self-aware
and that (mostly) avoided Exsurgent infection. The
Prometheans secretly back Firewall and work to defeat 
existential threats.
\textbf{ Proxies:} Members of the Firewall internal structure.
\textbf{ Psi:} Parapsychological powers acquired due to infection 
by the Watts-MacLeod strain of Exsurgent virus.
\textbf{ Reaper:} A warbot synthmorph.
\textbf{ Reclaimers:} A transhuman faction that seeks to lift the 
interdiction and reclaim Earth.
\textbf{ Redneck:} A rural Martian. See Barsoomian. Aka Reds.
\textbf{ Re-instantiated:} Refugees from Earth who escaped 
only as bodiless infomorphs, but who have since been 
resleeved.
\textbf{ Resleeving:} Changing bodies, or being downloaded into 
a new one. Also called remorphing, reincarnation, shifting, 
rebirthing.
\textbf{ Rusters:} Biomorphs optimized for life on Mars.
\textbf{ Scorching:} Hostile programs that can damage or affect 
cyberbrains.
\textbf{ Scum:} The nomadic faction of space punks/gypsies that 
travel from station to station in heavily-modified barges 
or swarms of ships. Notorious for being a roving black 
market.
\textbf{ Seed AI:} An AGI that is capable of recursive self-improvement
allowing it to reach god-like levels of intelligence.
\textbf{ Sentinels:} Agents of Firewall.
\textbf{ Shell:} A synthetic physical morph. Aka synthmorph.
\textbf{ Simulmorph:} The avatar you use in VR simulspace 
programs.
\textbf{ Simulspace:} Full-immersion virtual reality environments.
\textbf{ Singularity:} A point of rapid, exponential, and recursive 
technological progress, beyond which the future becomes 
impossible to predict. Often used to refer to the ascension 
of seed AI to god-like levels of intelligence.
\textbf{ Singularity Seeker:} People who pursue relics and 
evidence of the TITANs or other possible avenues to 
super-intelligence, either to learn more about it or to 
become part of a super-intelligence themselves.
\textbf{ Skin:} A biological physical morph. Aka meat, ﬂ esh.
\textbf{ Skinning:} Changing your perceived environment via 
augmented reality programming.
\textbf{ Sleight:} A psi power.
\textbf{ Slitheroid:} A snake-like robotic synthmorph.
\textbf{ Smart Animals:} Partially-uplifted animal species including
dogs, cats, rats, and pigs). Some other large smart
animals (whales, elephants) are nearly extinct.
\textbf{ Spime:} Meshed, self-aware, location-aware devices.
\textbf{ Splicers:} Humans that are genetically modified to 
eliminate genetic diseases and some other traits. Also 
known as genefixed, cleangenes, tweaks.
\textbf{ Swarmanoid:} A synthetic morph composed from a 
swarm of tiny insect-sized robots.
\textbf{ Sylphs:} Transhuman biomorphs with exotic good looks.
\textbf{ Synthmorph:} Synthetic morphs. Robotic shells possessed 
by transhuman egos.
\textbf{ Synths:} A specific type of synthmorph. Synths are 
standard androids/gynoids; robots that are designed to 
look humanoid, though they are usually noticeably not 
human.
\textbf{ Teleoperation:} Remote control.
\textbf{ Titanian:} Someone from Titan, a moon of Saturn.
\textbf{ TITANs:} The human-created, recursively-improving, 
military seed AIs that underwent a hard-takeoff singularity
and prompted the Fall. Original military designation
was TITAN: Total Information Tactical Awareness Network.
\textbf{ Torus:} A donut-shaped habitat, spun for gravity.
\textbf{ Transgenic:} Containing genetic traits from other species.
\textbf{ Transhuman:} An extensively modified human.
\textbf{ Trojans:} Asteroids or moons that share the same orbit 
as a larger planet or moon, but follow about 60 degrees 
ahead or behind at the L4 and L5 Lagrange points. The 
term Trojans normally refers to the asteroids orbiting at 
Jupiter's Lagrange points, but Mars, Saturn, Neptune, and 
other bodies also have Trojans. See also ``Greeks.''
\textbf{ Uplifting:} Genetically transforming an animal species to 
sapience.
\textbf{ Vacworker:} Space laborer.
\textbf{ Vapor:} A failed mind emulation or crippled fork/infomorph
(from vaporware).
\textbf{ VPNs:} Virtual private networks. Networks that operate 
within the mesh, usually encrypted for privacy/security.
\textbf{ VR:} Virtual Reality. Imposing an artificially-constructed 
hyper-real reality over one's physical senses.
\textbf{ X-Caster:} Someone who transmits/sells XP recordings of 
their experiences.
\textbf{ Xenomorph:} Alien life form.
\textbf{ Xer:} As in ``X-er''—someone who is addicted or 
obsessed with XP. Sometime used to refer to people 
making XP as well.
\textbf{ XP:} Experience Playback. Experiencing someone else's 
sensory input (in real-time or recorded). Also called 
experia, sim, simsense, playback.
\textbf{ X-Risk:} Existential risk. Something that threatens the 
very existence of transhumanity.
\textbf{ Zeroes:} People without wireless mesh access. Common 
with some indentures.
%%% 030
%%% 031
