%%% 114
%%% 115
%%% 116
In every game, there comes a time when the gamemaster
must decide if a character succeeds or fails in an
action. This is when the players roll dice and the characters
stats and abilities come into play. This chapter
defines the core mechanics and rules that govern the 
outcomes of events in \textit{Eclipse Phase}.

\subsection{The Ultimate Rule}

One rule in \textit{Eclipse Phase} outweighs all of the others: 
have fun. This means that you should never let the 
rules get in the way of the game. If you don't like a 
rule, change it. If you can't find a rule, make one up. 
If you disagree over a rule's interpretation, flip a coin. 
Try not to let rules interfere with the game's flow and 
mood. If you're in the middle of a really good scene 
or intense roleplaying and a rule suddenly comes into 
question, don't stop the game to look it up and argue 
about it. Just wing it, make a decision quickly, and 
move on. You can always look the rule up later so 
you'll remember it next time. If there are disagreements
over a rule's interpretation, remember that the
gamemaster gets the final say.

This rule also means that you shouldn't let the story 
be guided solely by rolls of the dice. The element of 
chance that dice rolls provide lends a sense of randomness
uncertainty, and surprise to the game. Sometimes
this is exciting, like when a character makes an unexpectedly
difficult roll and saves the day. At other
times, it is brutal, such as when a lucky shot from an 
opponent takes one of the characters out for good in a 
fight. If the gamemaster wants a scenario to result in a 
pre-planned dramatic outcome and an unexpected die 
roll threatens that plan, they should feel free to ignore 
that roll and move the story in the direction they desire.

\subsection{Dice}

\textit{Eclipse Phase} uses two ten-sided dice (each noted as 
a \textit{d10}) for random rolls. In most cases, the rules will 
call for a percentile roll, noted as \textit{d100,} meaning that 
you roll two ten-sided dice, choosing one to count first, 
and then read them as a result between 0 and 99 (with 
a roll of 00 counting as zero, not 100). The first die 
counts as the tens digit, and the second die counts as 
the ones digit. For example, you roll two ten-sided 
dice, one red and one black, calling out red first. The 
red one rolls a 1 and the black die rolls a 6, for a result 
of 16. Some sets of d10s, as shown above, are specifically
marked for easier rolling and reading.

Occasionally the rules will call for individual die 
rolls, with each individual ten-sided die listed as a \textit{d10. }
If the rules call for several dice to be rolled, it will be 
noted as 2d10, 3d10, and so on. When multiple ten-sided
dice are rolled in these instances, the results are
added together. For example, a 3d10 roll of 4, 6, and 
7 counts as 17. On d10 rolls, a result of 0 is treated as 
a 10, not a zero.

Most players of \textit{Eclipse Phase} can get by with 
having two ten-sided dice, but it doesn't hurt to 
have more on hand. These dice can be purchased 
at your friendly local game store or borrowed from 
another gamer.
%%% 117

\subsection{Making Tests}

In  \textit{Eclipse Phase,} your character is bound to find 
themself in adrenalin-pumping action scenes, high-stress
social situations, lethal combats, spine-tingling
investigations, and similar situations filled with drama, 
risk, and adventure. When your character is embroiled 
in these scenarios, you determine how well they do by 
\textit{making tests}—rolling dice to determine if they succeed 
or fail, and to what degree.

You make tests in \textit{Eclipse Phase} by rolling d100 and 
comparing the result to a target number. The target 
number is typically determined by one of your character's
skills (discussed below) and ranges between 1
and 98. If you roll \textit{less than or equal to} the target 
number, you succeed. If you roll \textit{higher} than the target 
number, you fail.

A roll of 00 is \textit{always} considered a success. A roll of 
99 is \textit{always }a failure.

\subsection{Target Numbers}

As noted above, the target number for a d100 roll in 
\textit{Eclipse Phase} is usually the skill rating. Occasionally, 
however, a different figure will be used. In some cases, 
an aptitude score is used, which makes for much 
harder tests as aptitude scores are usually well below 
50 (see \textit{Aptitudes,} p. 123). In other tests, the target 
number will be an aptitude rating x 2 or x 3 or two 
aptitudes added together. In these cases, the test description
will note what rating(s) to use.

\subsection{When To Make Tests}

The gamemaster decides when a character must make 
a test. As a rule of thumb, tests are called for whenever
there is a chance that a character might fail at an
action or when success or failure may have an effect 
on the ongoing story. Tests are also called for whenever
two or more characters act in opposition to one
another (for example, if they are arm wrestling or haggling
over a price). On the other hand, routine use of
a skill by someone with at least a rating of 30 in that 
skill can be assumed to be successful with no test.

It is not necessary to make tests for everyday, run-of-the-mill
activities, such as getting dressed or checking
your email (especially in \textit{Eclipse Phase,} where so
many activities are automatically handled by the machines
around you). Even an activity such as driving
a car does not call for dice rolls as long as you have 
a small modicum of skill. A test might be necessary, 
however, if you happen to be driving while bleeding 
to death or are pursuing a gang of motorcycle-riding 
scavengers through the ruins of a devastated city.

Knowing when to call for tests and when to let the 
roleplaying flow without interruption is a skill every 
gamemaster must acquire. Sometimes it is better to 
simply make a call without rolling dice in order to 
maintain the pacing of the game. Likewise, in certain 
circumstances the gamemaster may decide to make tests 
for a character in secret, without the player noticing. If 
an enemy is trying to sneak past a character on guard, 
for example, the gamemaster will alert the player that 
something is amiss if they ask them to make a perception 
test. This means that the gamemaster should keep a copy 
of each character's record sheet on hand at all times.

\subsection{Difficulty And Modifiers}

The measure of a test's difficulty
is reflected in its \textit{modifiers.}
Modifiers are adjustments made 
to the target number (not the 
roll), either raising or lowering 
it. A test of average difficulty 
will have no modifiers, whereas 
actions that are easier will have 
positive modifiers (raising the 
target number, making success 
more likely) and harder actions
will have negative modifiers
(lowering the target number,
making success less likely). It is 
the gamemaster's job to determine
if a particular test is harder
or easier than normal and to 
what degree (as illustrated on the 
Test Difficulty table) and to then 
apply the appropriate modifier.
Other factors might also play a 
role in a test, applying additional 
modifiers aside from the test's 
general level of difficulty. These factors include the environment
equipment (or lack thereof), and the health
of the character, among other things. The character 
might be using superior tools, working in poor conditions
or even wounded, and each of these factors must
be taken into account, applying additional modifiers 
to the target number and adjusting the likelihood of 
success or failure.
For simplicity, modifiers are applied in multiples of 
10 and come in three levels of severity: Minor (+/–10), 
Moderate (+/–20), and Major (+/–30). Any number of 
modifiers may be applied, as the gamemaster deems 
appropriate, but the cumulative modifiers may not 
exceed + or – 60.

Effortless

+30

Simple

+20

Easy

+10

Average

+0

Difficult

–10

Challenging

–20

Hard

–30

Minor

+/– 10

Moderate

+/– 20

Major

+/– 30
%%% 118

\subsection{Criticals: Rolling Doubles}

Any time both dice come up with the same number—
00, 11, 22, 33, 44, etc.—you have scored a critical 
success or critical failure, depending on whether your 
roll also beats your target number. 00 is always a 
critical success, whereas 99 is always a critical failure. 
Rolling doubles means that a little something \textit{extra }
happened with the outcome of the test, either positive 
or negative. Criticals have a very specific application 
in combat tests (see p. 192), but for all other purposes 
the gamemaster decides what exactly went wrong or 
right in a specific situation. Criticals can be used to 
amplify a success or failure: you finish with a flourish
or fail so spectacularly that you remain the butt
of jokes for weeks to come. They can also result in 
some sort of unexpected secondary effect: you repair 
the device \textit{and} improve its performance; or you fail 
to shoot your enemy \textit{and} hit an innocent bystander. 
Alternately, a critical can be used to give a boost (or 
a hindrance) to a follow-up action. For example, you 
not only spot a clue, but you immediately suspect it 
to be red herring; or you not only fail to strike the 
target, but your weapon breaks, leaving you defenseless
Gamemasters are encouraged to be inventive with
their use of criticals and choose results that create 
comedy, drama, or tension.

\subsection{Defaulting: Untrained Skill Use}

Certain tests may call for a character to use a skill they 
don't have—a process called \textit{defaulting.} In this case, 
the character instead uses the rating of the aptitude 
(see p. 123) that is linked to the skill in question as the 
target number.

Not all skills may be defaulted; some of them are 
so complex or require such training than an unskilled 
character can't hope to succeed. Skills that may not be 
defaulted on are noted on the Skill List (p. 176) and in 
the skill description.

In rare cases, a gamemaster might allow a character 
to default to another skill that also relates to a test 
(see p. 173). When allowed, defaulting to another skill 
incurs a –30 modifier.
%%% 119

\subsection{Teamwork}

If two or more characters join forces to tackle a test 
together, one of the characters must be chosen as the 
primary actor. This leading character will usually (but 
not always) be the one with the highest applicable 
skill. The primary acting character is the one who 
rolls the test, though they receive a +10 modifier for 
each additional character helping them out, up to a 
maximum +30 modifier. Note that helping characters 
do not necessarily need to know the skill being used 
if the gamemaster decides that they can follow the 
primary actor's lead.

\section{Types Of Tests}

There are two types of tests in \textit{Eclipse Phase:} Success 
and Opposed.

\subsection{Success Tests}

Success Tests are called for whenever a character is 
acting without direct opposition. They are the standard
tests used to determine how well a character
exercises a particular skill or ability.

Success Tests are handled exactly as described under 
\textit{Making Tests,} p. 115. The player rolls d100 against a 
target number equal to the character's skill +/– modifiers
If they roll equal to or less than the target number,
the test succeeds, and the action is completed as desired. 
If they roll higher than the target number, the test fails.

\subsubsection{Trying Again}

If you fail at a test, you can take another shot. Each 
subsequent attempt at an action after a failure, however
incurs a cumulative –10 modifier. That means the
second try suffers –10, the third –20, the fourth –30, 
and so on, up to the maximum –60.

\subsubsection{Taking The Time}

Most skill tests are made for Automatic, Quick, or 
Complex Actions (see pp. 119–120) and so are resolved
within one Action Turn (3 seconds, see p. 119).
Tests made for Task actions (p. 120) take longer.

Players may choose to take extra time when their 
character undertakes an action, meaning that they 
choose to be especially careful when performing the 
action in order to enhance their chance of success. For 
every minute of extra time they take, they increase 
their target number by +10. Once they've modified 
their target number to over 99, they are practically assured
of success, so the gamemaster can waive the dice
roll and grant them an automatic success. Note that 
the maximum +60 modifier rule still applies, so if their 
%%% 120
skill is under 40 to start with, taking the time may still 
not guarantee a favorable outcome. You may take the 
time even when defaulting (see \textit{Defaulting, }p. 116).

Taking extra time is a solid choice when time is not 
a factor to the character, as it eliminates the chance 
that a critical failure will be rolled and allows the 
player to skip needless dice rolling. For certain tests it 
may not be appropriate, however, if the gamemaster 
decides that no amount of extra time will increase the 
likelihood of success. In that case, the gamemaster 
simply rules that taking the time has no effect.

For Task action tests (p. 120), which already take 
time to complete, the duration of the task must be 
increased by 50 percent for each +10 modifier gained 
for taking extra time.

\subsubsection{Simple Success Tests}

In some circumstances, the gamemaster may not be 
concerned that a character might fail a test, but instead 
simply wants to gauge how well the character performs
In this case, the gamemaster calls for a \textit{Simple }
\textit{Success Test,} which is handled just like a standard 
Success Test (p. 117). Rather than determining success
or failure, however, the test is assumed to succeed.
The roll determines whether the character succeeds 
\textit{strongly }(rolls equal to or less than the target number) 
or succeeds \textit{weakly }(rolls above the target number).

\subsubsection{Margin Of Success/Failure}

Sometimes it may be important that a character not 
only succeeds, but that they kick ass and take names 
while doing it. This is usually true of situations where 
the challenge is not only difficult but the action must 
be pulled off with finesse. Tests of this sort may call 
for a certain \textit{Margin of Success (MoS)}—an amount by 
which the character must roll under the target number. 
For example, a character facing a target number of 55 
and a MoS of 20 must roll equal to or less than a 35 
to succeed at the level the situation calls for.

At other times, it may be important to know how 
badly a character fails, as determined by a \textit{Margin }
\textit{of Failure (MoF)}, which is the amount by which the 
character rolled over the target number. In some cases, 
a test may note that a character who fails with a 
certain MoF may suffer additional consequences for 
failing so dismally.

\subsubsection{Excellent Successes/Severe Failures}

Excellent Successes and Severe Failures are a method 
used to benchmark successes and failures with an 
MoS or MoF of 30+. Excellent Successes are used in 
situations where an especially good roll may boost the 
intended effect, such as inflicting more damage with 
a good hit in combat. Severe Failures denote a roll 
that is particularly bad and has a worse effect than a 
simple failure. Neither Excellent Successes or Severe 
Failures are as good or bad as criticals, however.
%%% 121

\subsection{Opposed Tests}

An Opposed Test is called for whenever a character's 
action may be directly opposed by another. Regardless 
of who initiates the action, both characters must make 
a test against each other, with the outcome favoring 
the winner.

To make an Opposed Test, each character rolls 
d100 against a target number equal to the relevant 
skill(s) along with any appropriate modifiers. If only 
one of the characters succeeds (rolls equal to or less 
than their target number), that character has won. 
If both succeed, the character who gets the highest 
dice roll wins. If both characters fail, or they both 
succeed and roll the same number, then a deadlock 
occurs—the characters remain pitted against each 
other, neither gaining ground, until one of them takes 
another action and either breaks away or makes another
Opposed Test.

Note that critical successes trump high rolls in an 
Opposed Test—if both characters succeed and one rolls 
54 while the other rolls 44, the critical roll of 44 wins.

Care must be taken when applying modifiers in an 
Opposed Test. Some modifiers will affect both participants
equally, and should be applied to both tests. If
a modifier arises from one character's advantage in 
relation to the other, however, that modifier should 
only be applied to benefit the favored character; it 
should not also be applied as a negative modifier to 
the disadvantaged character.

\subsubsection{Opposed Tests And Margin Of Success/Failure}

In some cases, it may also be important to note a character's
Margin of Success or Failure in an Opposed
Test, as with a Success Test above. In this case, the 
MoS/MoF is still determined by the difference between 
the character's roll and their target number—it is not 
calculated by the difference between the character's 
roll and the opposing character's roll.

\subsubsection{Variable Opposed Test}

In some cases, the rules will call for a \textit{Variable Op-}
\textit{posed Test,} which allows for slightly more outcomes 
than a standard Opposed Test. If both characters 
succeed in a Variable Opposed Test, then an outcome 
is obtained which is different from just one character 
winning over the other. The exact outcomes are noted 
with each specific Variable Opposed Test.

\section{Time And Actions}

Though the gamemaster is responsible for managing 
the speed at which events unfold, there are times 
when it is important to know exactly who is acting 
when, especially if some people are acting before or 
after other people. In these circumstances, gameplay 
in \textit{Eclipse Phase} is broken down into intervals called 
\textit{Action Turns.}

\subsection{Action Turns}

Each Action Turn is three seconds long, meaning 
there are twenty Action Turns per minute. The order 
in which characters act during a turn is determined 
by an Initiative Test (see \textit{Initiative,} p. 121). Action 
Turns are further subdivided into \textit{Action Phases.} Each 
character's Speed stat (p. 121) determines the amount 
of actions they can take in a turn, represented by how 
many Action Phases they may take.

\subsection{Types Of Actions}

The types of actions a character may take in an Action 
Turn are broken down to: \textit{Automatic, Quick, Com-}
\textit{plex, }and \textit{Task }actions\textit{.}

\subsubsection{Automatic Actions}

Automatic actions are ``always on'' and require no 
effort from the character, assuming they are conscious.
\textbf{Examples:} basic perception, certain psi sleights

\subsubsection{Quick Actions}

Quick actions are simple, so they can be done fast and 
can be multi-tasked. The gamemaster determines how 
many Quick actions a character may take in a turn.
\textbf{Examples:} talking, switching a safety, activating an 
implant, standing up

d

w
%%% 122

\subsubsection{Complex Actions}

Complex actions require concentration or 
effort. The number of Complex actions a 
character may take per turn is determined 
by their Speed stat (see p. 121).

\textbf{Examples:} attacking, shooting, acrobatics, 
disarming a bomb, detailed examination

\subsubsection{Task Actions}

Task actions are any actions that require 
longer than one Action Turn to complete. 
Each Task action has a \textit{timeframe,} usually 
listed in the task description or otherwise 
determined by the gamemaster. The timeframe
determines how long the task takes
to complete, though this may be reduced by 
10 percent for every 10 full points of MoS 
the character scores on the test (see \textit{Margin }
\textit{of Success/Failure,} p. 118). If a character 
fails on a Task action test roll, they work 
on the task for a minimum period equal to 
10 percent of the timeframe for each 10 full 
points of MoF before realizing it's a failure.

For Task actions with timeframes of one 
day or longer, it is assumed that the character
only works eight hours per day. A character
that works more hours per day may
reduce the time accordingly. Characters 
working on Task actions may also interrupt
their work to do something else and
then pick up where they left off, unless the 
gamemaster rules that the action requires 
continuous and uninterrupted attention.

Similar to taking the time (p. 117), a 
character may \textit{rush the job} on a Task action, 
taking a penalty on the test in order to decrease
the timeframe. The character must
declare they are rushing the job before they 
roll the test. For every 10 percent they wish 
to reduce the timeframe, they incur a –10 
modifier on the test (to a maximum reduction
of 60 percent with a maximum modifier
of –60).

\section{Defining Your Character}

In order to gauge and quantify what your 
character is merely good at and what they 
excel in—or what they are clueless about 
and suck at—\textit{Eclipse Phase} uses a number 
of measurement factors: \textit{stats, skills, traits,} 
and  \textit{morphs.} Each of these characteristics 
is recorded and tracked on your character's 
\textit{record sheet} (p. 399).

\subsection{Concept}

Your character concept defines who you are 
in the \textit{Eclipse Phase }universe. You're not just 
a run-of-the-mill plebeian with a boring and 
mundane life, you're a participant in a post-apocalyptic
transhuman future who gets
caught up in intrigue, terrible danger, unspeakable
horrors, and scrambling for survival
Much like a character in an adventure,
drama, or horror story, you are a person 
to whom interesting things happen—or if 
not, you make them happen. This means 
your character needs a distinct personality 
and sense of identity. At the very least, you 
should be able to sum up your character 
concept in a single sentence, such as cantankerous
neotenic renegade archaeologist with
anger management issues'' or ``thrill-seeking 
social animal who is dangerously obsessed 
with conspiracy theories and mysteries.'' If 
it helps, you can always borrow ideas from 
characters you've seen in movies or books, 
modifying them to fit your tastes.

Your character's concept will likely be 
influenced by two important factors: \textit{back-}
\textit{ground} and \textit{faction.} Your background denotes
the circumstances under which your
character was raised, while your faction 
indicates the post-Fall grouping to which 
you most recently held ties and allegiances. 
Both of these play a role in character creation
(p. 128).

\subsection{Motivations}

The clash of ideologies and memes is a core 
component of \textit{Eclipse Phase,} and so every 
character has three \textit{motivations}—personal 
memes that dominate the character's interests 
and pursuits. These memes may be as abstract
as ideologies the character adheres to
or supports—for example, social anarchism, 
Islamic jihad, or bioconservatism—or they 
may be as concrete as specific outcomes the 
character desires, such as revealing a certain 
hypercorp's corruption, obtaining massive 
personal wealth, or winning victories for 
uplifted rights. A motivation may also be 
framed in opposition to something; for example
anti-capitalism or anti-pod-citizenship,
or staying out of jail. In essence, these are 
ideas that motivate the character to do the 
things they do. Motivation is best noted as a 
term or short phrase on the character sheet, 
marked with a + (in favor of) or – (opposed 
to). Players are encouraged to develop their 
own distinct motivations for their characters, 
in cooperation with the gamemaster. Some 
examples are provided on p. 138.

In game terms, motivation is used to help 
define the character's personality and influence
their actions for roleplaying purposes.
It also serves to regain Moxie points (p. 122) 
and earn Rez Points for character advancement
(p. 152).
%%% 123

Motivational goals may be short-term or long-term, 
and may in fact change for a character over time. 
Short-term goals are more immediately obtainable 
objectives or short-lived interests, and these goals are 
likely to change once achieved. Even so, they should 
reflect intentions that will take more than one game 
session to reach, possibly covering weeks or months. 
These short-term goals may in fact tie directly into the 
gamemaster's current storyline. Examples include conducting
a full analysis of an alien artifact, completing
a research project, or living life as an uplifted dog for a 
while. Long-term goals reflect deeply rooted beliefs or 
tasks that require major efforts and time (possibly lifelong
to achieve. For example, finding the lost backup
of a sibling missing since the Fall, overthrowing an 
autocratic regime, or making first contact with a new 
alien species. For purposes of awarding Moxie or Rez 
Points, long-term goals are best broken down into 
obtainable chunks. Someone whose goal is to track 
down the murderer who killed their parents when they 
were a child, for example, can be considered to achieve 
that goal every time they discover some evidence that 
brings them a little closer to solving the puzzle.

\subsection{Ego Vs. Morph}

\textit{Eclipse Phase}'s setting dictates that a distinction must 
be made between a character's \textit{ego} (their ingrained 
self, their personality, and inherent traits that perpetuate
in continuity) and their \textit{morph} (their ephemeral
physical—and sometimes virtual—form). A character's
morph may die while the character's ego lives on
(assuming appropriate backup measures have been 
taken), transplanted into a new morph. Morphs are 
expendable, but your character's ego represents the 
ongoing, continuous life path of your character's mind, 
personality, memories, knowledge, and so forth. This 
continuity may be interrupted by an unexpected death 
(depending on how recent the backup was made), or 
by forking (see p. 273), but it represents the totality of 
the character's mental state and experiences.

Some aspects of your character—particularly skills, 
along with some stats and traits—belong to your character's
ego, which means they stay with them throughout
the character's development. Some stats and traits,
however, are determined by morph, as noted, and so 
will change if your character leaves one body and 
takes on another. Morphs may also affect other skills 
and stats, as detailed in the morph description.

It is important that you keep ego- and morph-derived
characteristics straight, especially when updating
your character's record sheet.

\subsection{Character Stats}

Your character's stats measure several characteristics 
that are important to game play: \textit{Initiative, Speed, }
\textit{Durability, Wound Threshold, Lucidity, Trauma }
\textit{Threshold,} and \textit{Moxie.} Some of these stats are inherent
to your character's ego, others are influenced or
determined by morph.

\subsection{Ego Stats}


\subsection{Morph Stats}


\subsubsection{Initiative (Init)}

Your character's Initiative stat helps determine when 
they act in relation to other characters during the 
Action Turn (see \textit{Initiative,} p. 188). Your Initiative stat 
is equal to your character's Intuition + Reflexes aptitudes
(see \textit{Aptitudes,} p. 123) multiplied by 2. Certain
implants and other factors may modify this score.

\subsubsection{Speed (Spd)}

The Speed stat determines how often your character 
gets to act in an Action Turn (see \textit{Initiative,} p. 188). 
All characters start with a Speed stat of 1, meaning 
they act once per turn. Certain implants and other 
advantages may boost this up to a maximum of 4.

\subsubsection{Durability (Dur)}

Durability is your morph's physical health (or structural
integrity in the case of synthetic shells, or system
integrity in the case of infomorphs). It determines the 
amount of damage your morph can take before you are 
incapacitated or killed (see \textit{Physical Health,} p. 206).
Durability is unlimited, though the range for baseline
(unmodified) humans tends to fall between 20 and
60. Your Durability stat is determined by your morph.

\subsubsection{Wound Threshold (Wt)}

A Wound Threshold is used to determine if you receive a 
wound each time you take physical damage (see \textit{Physi-}
\textit{cal Health,} p. 206). The higher the Wound Threshold, 
the more resistant to serious injury you are.
Wound Threshold is calculated by dividing Durability
by 5 (rounding up).

\subsubsection{Death Rating (Dr)}

Death Rating is the total amount of damage your 
morph can take before it is killed or destroyed 
beyond repair. Death Rating is equal to DUR x 1.5 for 
biomorphs and DUR x 2 for synthmorphs.

il
124

\subsubsection{Lucidity (Luc)}

Lucidity is similar to Durability, except that it measures 
mental health and state of mind rather than physical 
well-being. Your Lucidity determines how much stress 
(mental damage) you can take before you are incapacitated
or driven insane (see \textit{Mental Health,} p. 209).

Lucidity is unlimited, but generally ranges from 
20 to 60 for baseline unmodified humans. Lucidity is 
determined by your Willpower aptitude x 2.

\subsubsection{Trauma Threshold (Tt)}

The Trauma Threshold determines if you suffer a 
trauma (mental wound) each time you take stress (see 
\textit{Mental Health,} p. 209). A higher Trauma Threshold 
means that your mental state is more resilient against 
experiences that might inflict psychiatric disorders or 
other serious mental instabilities.

Trauma Threshold is calculated by dividing Lucidity 
by 5 (rounding up).

\subsubsection{Insanity Rating (Ir)}

Your Insanity Rating is the total amount of stress your 
mind can take before you go permanently insane and 
are lost for good. Insanity Rating equals LUC x 2.

\subsubsection{Moxie}

Moxie represents your character's inherent talent at 
facing down challenges and overcoming obstacles 
with spirited fervor. More than just luck, Moxie is 
your character's ability to run the edge and do what 
it takes, no matter the odds. Some people consider it 
the evolutionary trait that spurred humankind to pick 
up tools, expand our brains, and face the future head 
on, leaving other mammals in the dust. When the sky 
is falling, death is imminent, and no one can help you, 
Moxie is what saves the day.

The Moxie stat is rated between 1 and 10, as purchased
during character creation (and perhaps raised
later). In game play, Moxie is used to influence the 
odds in your favor. Every game session, your character 
begins with a number of Moxie points equal to their 
Moxie stat. Moxie points may be spent for any of the 
following effects:
•  The character may ignore all modifiers that apply 

to a test. The Moxie point must be spent \textit{before }

dice are rolled.
•  The character may flip-flop a d100 roll result. For 

example, an 83 would become a 38.
•  The character may upgrade a success, making it a 

critical success, as if they rolled doubles. The char-

acter must succeed in the test before they spend 

the Moxie point.
•  The character may ignore a critical failure, treat-

ing it as a regular failure instead.
•  The character may go first in an Action Phase (p. 189).
Only 1 point of Moxie may be spent on a single roll. 
Moxie points will fluctuate during gameplay, as they 
are spent and sometimes regained.
\textbf{Regaining Moxie:} At the gamemaster's discretion, 
Moxie points may be refreshed up to the character's 
full Moxie stat any time the character rests for a significant
period. Moxie points may also be regained if
the character achieves a personal goal, as determined 
by their Motivation (see p. 121). The gamemaster determines
how much Moxie is regained in proportion
to the goal achieved.

\subsubsection{Damage Bonus}

The Damage Bonus stat quantifies how much extra 
oomph your character is able to give their melee and 
thrown weapons attacks. Damage Bonus is determined
by dividing your Somatics aptitude (see below)
by 10 and rounding down.

\subsection{Character Skills}

Skills represent your character's talents. Skills are 
broken down into \textit{aptitudes} (ingrained abilities that 
everyone has) and \textit{learned skills} (abilities and knowledge
picked up over time). Skills determine the target
number used for tests (see \textit{Making Tests,} p. 115).

\subsubsection{Aptitudes}

Aptitudes are the core skills that every character has 
by default. They are the foundation on which learned 
skills are built. Aptitudes are purchased during character
creation and rate between 1 and 30, with 10 being
average for a baseline unmodified human. They represent
the ingrained characteristics and talents that your
character has developed from birth and stick with you 
even when you change morphs—though some morphs 
may modify your aptitude ratings.
%%% 125
Each learned skill is linked to an aptitude. If a 
character doesn't have the skill necessary for a test, 
they may default to the aptitude instead (see \textit{Default-}
\textit{ing,} p. 116).
There are 7 aptitudes in \textit{Eclipse Phase:}
•  \textbf{Cognition (COG)} is your aptitude for problem-

solving, logical analysis, and understanding. It 

also includes memory and recall.
•  \textbf{Coordination (COO)} is your skill at integrat-

ing the actions of different parts of your morph 

to produce smooth, successful movements. It 

includes manual dexterity, fine motor control, 

nimbleness, and balance.
•  \textbf{Intuition} \textbf{(INT)} is your skill at following your 

gut instincts and evaluating on the fly. It includes 

physical awareness, cleverness, and cunning.
•  \textbf{Reflexes (REF)} is your skill at acting quickly. This 

encompasses your reaction time, your gut-level 

response, and your ability to think fast.
•  \textbf{Savvy (SAV)} is your mental adaptability, social in-

tuition, and proficiency for interacting with others. 

It includes social awareness and manipulation.
•  \textbf{Somatics (SOM)} is your skill at pushing your 

morph to the best of its physical ability, includ-

ing the fundamental utilization of the morph's 

strength, endurance, and sustained positioning 

and motion.
•  \textbf{Willpower (WIL)} is your skill for self-control, 

your ability to command your own destiny.

\subsubsection{Learned Skills}

Learned skills encompass a wide range of specialties 
and education, from combat training to negotiating 
to astrophysics (for a complete skill list, see p. 176). 
Learned skills range in rating from 1 to 99, with an 
average proficiency being 50. Each learned skill is 
linked to an aptitude, which represents the underlying 
competency in which the skill is based. When a learned 
skill is purchased (either during character generation 
or advancement), it is bought starting at the rating of 
the linked aptitude and then raised from there. If the 
linked aptitude is raised or modified, all skills built off 
it are modified appropriately as well.
Depending on your background and faction, you 
may receive some starting skills for free during 
character creation. Like aptitudes, learned skills stay 
with the character even when they change morphs, 
though certain morphs, implants, and other factors 
may sometimes modify your skill rating. If you lack 
a particular skill called for by a test, in most cases 
you can default to the linked aptitude for the test (see 
\textit{Defaulting,} p. 116).

\subsubsection{Specializations}

Specializations represent an area of concentration and 
focus in a particular learned skill. A character who 
learns a specialization is one who not only grasps the 
basic tenets of that skill, but they have trained hard 
to excel in one particular aspect of that skill's field. 
Specializations apply a +10 modifier when the character
utilizes that skill in the area of specialization.
Specializations may be purchased during character 
creation or advancement for any existing skill the 
character possesses with a rating of 30 or more. Only 
one specialization may be purchased for each skill. 
Specific possible specializations are noted under individual
the skill descriptions (see \textit{Skills,} p. 170).

\subsection{Character Traits}

Traits include a range of inherent qualities and features
that help define your character. Some traits are
positive, in that they give your character a bonus to 
certain stats, skills, or tests, or otherwise give them 
an edge in certain situations. Others are negative, in 
that they impair your abilities or occasionally create 
a glitch in your plans. Some traits apply to a character's
ego, staying with them from body to body, while
others only apply to a character's morph. 
Traits are purchased during character generation. 
Positive traits cost customization points (CP), while 
negative traits give you extra CP to spend on other 
things (see \textit{Traits,} p. 145). The maximum number of 
CP you may spend on traits is 50, while the maximum
you may gain from negative traits is 50. In rare
circumstances—and only with gamemaster approval—traits
may be purchased, bought off, or inflicted
during gameplay (see p. 153).

\subsection{Character Morph}

In \textit{Eclipse Phase,} your body is disposable. If your body 
gets old, sick, or too heavily damaged, you can digitize 
your consciousness and download it into a new body. 
The process isn't cheap or easy, but it offers effective 
immortality—as long as you remember to back yourself
up and don't go insane. The term \textit{morph }is used to
describe any type of form your mind inhabits, whether 
it be a vat-grown clone sleeve, a synthetic robotic shell, 
a part-bio/part-flesh pod, or even the purely electronic 
software state of an infomorph.
You purchase your starting morph during character 
creation (see p. 128). This is likely the morph you 
were born with (assuming you were born), though it 
may simply be another morph you've moved onto.
Physical looks aside, your morph has a large impact 
on your characteristics. Your morph determines certain 
physical stats, such as Durability and Wound Threshold, 
and it may also influence Initiative and Speed. Morphs 
may also modify some of your aptitudes and learned 
skills. Some morphs come pre-loaded with specific traits 
and implants, representing how it was crafted, and you 
can always bling yourself out with more implants if 
you choose (see \textit{Implants,} p. 126). All of these factors 

ou

it

ct 

n 

, 
%%% 126
are noted in the individual morph descriptions (see 
p. 139).

If you plan on switching your current morph to 
another during gameplay, you may first want to back 
yourself up (see \textit{Backups and Uploads,} p. 268). Backing
up regularly is always a smart option in case you
suffer an accidental or untimely death. Acquiring a 
new morph is not always easy, especially if you want 
it pre-loaded according to certain specifications. The 
full process is detailed under \textit{Resleeving,} p. 271.

\subsubsection{Aptitude Maximums}

Every morph has an aptitude maximum, sometimes 
modified by traits. This maximum represents the highest
value at which the character may use that aptitude
while inhabiting that morph, reflecting an inherent 
limitation in some morphs. If a character's aptitude 
exceeds the aptitude maximum of their morph, they 
must use it at the maximum value for the duration 
of the time they remain in that morph. This may also 
affect the skills linked to that aptitude, which must be 
modified appropriately.

Some implants, gear, psi, and other factors may 
modify a character's natural aptitudes. These augmented
values may exceed a morph's aptitude maximums
as they represent external factors boosting the
morph's ability. No aptitude, however, augmented or 
not, may ever exceed a value of 40. Innate ability only 
takes a person so far—after that point, actual skill is 
what counts.

\section{Things Characters Use}

In the advanced technological setting of \textit{Eclipse Phase,} 
characters don't get by on their wits and morphs 
alone; they take advantage of their credit and reputation
to acquire gear and implants and use their social
networks to gather information. Some characters also 
have the capability to use mental powers known as psi.

\subsection{Identity}

In an age of ubiquitous computing and omnipresent 
surveillance, privacy is a thing of the past—who you 
are and what you do is easily accessed online. Characters
in \textit{Eclipse Phase,} however, are often involved
in secretive or less-than-legal activities, so the way to 
keep the bloggers, news, paparazzi, and law off your 
back is to make extensive use of fake IDs. While Firewall
often provides covers for its sentinel agents, it
doesn't hurt to keep a few extra personas in reserve, in 
case matters ever go out the airlock in a hurry. Thankfully
the patchwork allegiances of city-state habitats
and faction stations means that identities aren't too 
difficult to fake, and the ability to switch morphs 
makes it even easier. On the other hand, anyone with a 
copy of your biometrics or geneprint is going to have 
an edge tracking you down or finding any forensic 
traces you leave behind (for more on ID, see p. 279).

\subsection{Social Networks}

Social networks represent people the character knows 
and social groups with which they interact. These 
contacts, friends, and acquaintances are not just maintained
in person, but also through heavy Mesh contact.
Social software allows people to stay updated on what 
the people they know are doing, where they are, and 
what they are interested in, right up to the minute. 
Social networks also incorporate the online projects 
%%% 127
of individual members, whether it's a mesh-site loaded 
with a band member's songs, a personal archive of 
stored media, a decade of blog entries reviewing the 
best places to score cheap electronics, or a depository 
of research papers and studies someone has worked 
on or finds interesting.

In game play, social networks are quite useful 
to characters. Their friends list is an essential 
resource—a pool of people you can actively poll for 
ideas, troll for news, listen to for the latest rumors, 
buy or sell gear from, hit up for expert advice, and 
even ask for favors.

While a character's social networks are nebulous 
and constantly shifting, the use of them is not. A 
character takes advantage of their social networks via 
the Networking (Field) skill (p. 182). The exact use 
of this skill is covered under \textit{Reputation and Social }
\textit{Networks,} p. 285.

\subsection{Cred}

The Fall devastated the global economies and currencies
of the past. In the years of reconsolidation
that followed, the hypercorps and governments 
inaugurated a new system-wide electronic monetary 
system. Called \textit{credit,} this currency is backed by all 
of the large capitalist-oriented factions and is used to 
trade for goods and services as well for other financial 
transactions. Credit is mainly transferred electronically
though certified credit chips are also common
(and favored for their anonymity). Hardcopy bills are 
even used in some habitats.

Depending on your background or faction, your 
character may be given an amount of credit at the 
start of the game. During game play, your character 
must earn credit the old-fashioned way: by earning 
or stealing it.

\subsection{Rep}

Capitalism is no longer the only economy in town. 
The development of nanofabricators allowed for the 
existence of post-scarcity economies, a fact eagerly 
taken advantage of by anarchist factions and others. 
When anyone can make anything, concepts like 
property and wealth become irrelevant. The advent 
of functional gift and communist economies, among 
other alternative economic models, means that in 
such systems you can acquire any goods or services 
you need via free exchange, reciprocity, or barter—
presuming you are a contributing member of such 
a system and respected by your peers. Likewise, art, 
creativity, innovation, and various forms of cultural 
expression have a much higher worth than they do in 
capitalist economies.

In alternative economies, money is often meaningless
but \textit{reputation} matters. Your reputation score
represents your social capital—how esteemed you 
are to your peers. Rep can be increased by positively 
influencing, contributing to, or helping individuals or 
groups, and it can be decreased through antisocial 
behavior. In anarchist habitats, your likelihood of 
obtaining things that you need is entirely based on 
how you are viewed by others.
Reputation is easily measured by one of several 
online social networks. Your actions are rewarded 
or punished by those with whom you interact, who 
can ping your Rep score with positive or negative 
feedback. These networks are used by all of the factions
as reputation can affect your social activities in
capitalist economies as well. The primary reputation 
networks include:
• \textbf{The @-list:} the Circle-A list for anarchists, Barsoomians
Extropians, scum, and Titanians, noted
as @-rep.
• \textbf{CivicNet:} used by the Jovian Republic, Lunar-Lagrange
Alliance, Morningstar Constellation,
Planetary Consortium, and many hypercorps, 
referred to as c-rep.
• \textbf{EcoWave:} used by nano-ecologists, preservationists
and reclaimers, referred to as e-rep.
•  \textbf{Fame:} the seen-and-be-seen network used by 
socialites, artists, glitterati, and media, referred 
to as f-rep.
• \textbf{Guanxi:} used by the triads and numerous criminal
entities, referred to as g-rep.
•  \textbf{The Eye:} used by Firewall, noted as i-rep.
• \textbf{RNA:} Research Network Affiliation, used by argonauts
technologists, scientists, and researchers,
referred to as r-rep.
Reputation is rated from 0-99. Depending on 
your background and faction, you may start with 
a Rep score in one or more networks. This can be 
bolstered through spending customization points 
during character creation. During game play, your 
Rep scores will depend entirely on your character's 
actions. For more information, see \textit{Reputation and }
\textit{Social Networks,} p. 285.
Note that each Rep score is tied to a particular 
identity.

\subsection{Gear}

Gear is all of the equipment your character owns and 
keeps on their person, from weapons and armor to 
clothing and electronics. You buy gear for your character
with customization points during character creation
(see p. 136) and in the game with Credit or Rep.
Certain restricted, illegal, or hard-to-find items may 
require special efforts to obtain (see \textit{Acquiring Gear,} 
p. 298). If you have access to a nanofabricator, you 
may be able to simply build gear, presuming you have 
the proper blueprints (see \textit{Nanofabrication,} p. 284). 
For a complete listing of equipment options, see the 
\textit{Gear} chapter, p. 296.
Even among the remaining capitalist economies, 
prices can vary drastically. To represent this, all 
gear falls into a cost category. Each category defines 
a range of costs, so the gamemaster can adjust the 
prices of individual items as appropriate to the situation
Each category also lists an average price for that

\textit{,}

ou

ve 

). 

e
%%% 128
category, which is used during character generation 
and any time the gamemaster wants to keep costs 
simple. See the Gear Costs table on p. 137.

\subsection{Implants}

Implants include cybernetic, bionic, genetech, and 
nanoware enhancements to your character's morph 
(or mechanical enhancements in the case of a synthetic 
shell). These implants may give your character special 
abilities or modify their stats, skills, or traits. Some 
morphs come pre-equipped with implants, as noted in 
their descriptions (see p. 139). You may also special-order
morphs with specific implants (see \textit{Morph Ac-}
\textit{quisition,} p. 277). If you want to upgrade a morph 
you are already in, you can undergo surgery or other 
treatments to have an enhancement installed (see 
\textit{Healing Vats,} p. 326. For a complete list of available 
implant/enhancement options, see pp. 300-311, \textit{Gear.}

\subsection{Psi}

Psi is a rare and anomalous set of mental abilities that 
are acquired due to infection by a strange nanovirus 
released during the Fall. Psi abilities are not completely 
understood, but they give characters certain advantages—as
well as some disadvantages. A character
requires the Psi trait (p. 147) to use psi abilities, which 
are called \textit{sleights.} Psi users are called \textit{asyncs.} A full 
explanation of psi and details on the various sleights 
can be found in the \textit{Mind Hacks }chapter, p. 216.
%%% 129
