%%% 172
%%% 173
%%% 174
In a setting where physical looks and capabilities are 
easily changed at the push of a button, who you are and 
what you know is more important than any inborn ability
Skills represent the knowledge your character has,
the accumulated set of experience, education, and inherent
know-how possessed by each and every sentient
transhuman in \textit{Eclipse Phase}. They are what allow you 
to sneak into a hypercorp station, disable the security 
systems, hack the mesh hub, and then impersonate security
personnel to make your escape. Your skills represent
the one thing you have no matter what you look like or 
where you find yourself. When your characters explore 
what they can do, their skills, or lack thereof, often determine
the margin between success and failure.

Having a well-rounded set of skills is vital to survival 
and success in \textit{Eclipse Phase}. The skills below encompass 
a wide selection of talents, enough so that each character 
can be unique in their abilities and knowledge.

\section{Skills Overview}

Skills are divided into \textit{aptitudes} and \textit{learned skills }(see 
\textit{Character Skills,} p. 123). Most (but not all) learned 
skills are built on and linked to an aptitude. If a 
character lacks the specific skill needed in a situation, 
they may default to the linked aptitude. You may also 
choose to specialize in certain skills (see \textit{Specializa-}
\textit{tions,} p. 123), reflecting an enhanced knowledge of a 
particular aspect of a certain skill.

\subsection{Core Skills: Aptitudes}

Aptitudes represent inherent skills and abilities acquired 
at birth or during the course of growing up. Aptitudes 
are sometimes used for tests, but their primary use is 
determining the starting point at which learned skills 
are developed. Aptitudes determine the starting value 
of their linked skills. For example, a character with 
Somatics aptitude 10 who wishes to purchase points 
in the Freerunning skill (which is linked to Somatics) 
would start with a Freerunning rank of 10 and then 
buy additionally points in that skill.

Aptitudes are also used when a character doesn't 
posses knowledge of a needed skill (see \textit{Defaulting,} p. 
116). Aptitudes represent the basic knowledge that a 
character has acquired regarding rudimentary use of that 
skill. They may not have ever received any formal training
with the skill, but they can still attempt to use it.

Aptitudes range in value from 1 to 30, with 10 being 
the unaugmented human average and 15 representing 
the average of most genetically modified transhumans\textit{.}
Since aptitudes represent untrained ability, they are 
capped at a maximum rating of 30.

There are seven different aptitudes that all players 
possess. These aptitudes are purchased during character
creation (p. 128), but depending on the morph the
character is currently inhabiting, they may find their aptitudes
capped by the quality of the morph (see p. 124).

\subsection{Learned Skills}

A player's learned skills are the most important part of 
their character, representing the acquired knowledge 
they carry with them from morph to morph, knowledge
that plays a fundamental role in helping define
the person's ego. Learned skills encompass nearly any 
skill that you might need to use in \textit{Eclipse Phase,} and 
they range in value from 0 to 99.

All learned skills have a linked aptitude that is 
used to calculate their initial value, and which is 
also defaulted to if the player does not have that 
particular skill.

\subsubsection{Skill Categories}

Each learned skill is classified as either an \textit{Active skill}
or a \textit{Knowledge skill.} Active skills represent skills 
that typically require physical actions and are used in 
action scenes within game play. Knowledge skills are 
more knowledge-based and intellectual, representing 
ideas and facts. Knowledge skills may play a less dramatic
role in certain action-oriented game play moments
but they flesh out the character's background
and interests and are integral to roleplaying interactions
Active and Knowledge skills are purchased
separately during character creation.

Active skills are further divided into Combat, 
Mental, Physical, Psi, Social, Technical, and Vehicle 
skills. Certain traits and abilities may apply to specific 
categories.

\subsubsection{Field Skills}

Some learned skills are \textit{field skills,} meaning that when 
this skill is chosen a particular field of emphasis must 
also be selected. For example, the skill of Academics 
requires the character to specify a specific academic 
discipline in which they are knowledgeable, such as 
Biology, Chemistry, or Xenosociology. Field skills are 
written as ``[skill]: [field];'' for example: ``Art: Painting
Field skills can be taken multiple times, choosing
a different area of emphasis each time, reflecting skills 
in different fields; that is to say, each field is a separate 
skill. Several suggested fields are listed for each field 
skill, but gamemasters and players may also cooperate 
to create others that fit their games.

Field skills may also have specializations; for example
Professional: Accounting (Money Laundering).

\subsubsection{Psi Skills}

Psi refers to the ability to perceive and manipulate 
biological minds via psi waves and/or other inexplicable
phenomena. Due to the uniqueness of this ability
characters that wish to wield psi must acquire the
%%% 175
Psi trait (p. 147). Psi use also requires a number of 
specialized skills (Control, Psi Assault, and Sense) that 
reflect special training characters acquire to tap into 
their psi powers. Psi skills may not be defaulted on; 
the only way to use a psi skill is to possess the trait 
along with training in that skill. For more details, see 
\textit{Psi,} p. 220.

\subsection{Specializations}

Any character may opt to specialize in a given skill 
(see \textit{Specializations,} p. 123). This specialization reflects 
increased knowledge in one particular aspect of the 
skill. Many of the skills offered below include sample 
specializations. Gamemasters and players are encouraged
to develop other specialization ideas together for
their campaigns.

Specialization provides a +10 modifier when using 
that skill in a situation appropriate to that specialization.

\section{Using Skills}

Whenever a character wants to do something using 
a skill, they must succeed at a skill test (see \textit{Making }
\textit{Tests,} p. 115). The difficulty of the action is applied 
as a modifier, as are any other extenuating circumstances
that may affect the test (see \textit{Difficulty and }
\textit{Modifiers,} p. 115). As with other types of tests, all skill 
tests are successful when the character rolls less than 
or equal to the test's target number after any modifiers
have been applied. In the case of skill tests, the
target number is the character's skill rating with that 
particular skill. Modifiers representing difficulty and 
other factors are applied directly to the target number 
(see \textit{Difficulty and Modifiers,} p. 115). A roll of a 00 is 
always a success, regardless of modifiers, and a result 
of 99 is always a failure, again despite any modifiers 
that may increase a character's target number over 
100. Standard critical success and failure rules apply 
to skill tests (see \textit{Criticals: Rolling Doubles,} p. 116), 
so any time a character rolls a double (i.e. 00, 11, 22, 
33, etc.) they score a critical success or failure.

\subsection{Defaulting}

Sometimes you lack the skill needed in a certain situation
In these instances, characters may default their
skill test to the linked aptitude. This reflects the fact 
that most learned skills are developed from some sort 
of baseline physical ability. Even though you may not 
know how to do something, you've likely seen how 
it's done at some point or have some idea of how to 
do it, or can at least take a shot at it. Naturally, you're 
not as good as someone who has training in that skill, 
but it still allows you to make an attempt.

Not all skills can be defaulted. Some skills are 
simply too complex or obscure, or demand special 
knowledge or ability, for someone to attempt their 
use untrained. For example, brain surgery or most psi 
skills are simply beyond anyone who doesn't have that 
ability or the knowledge of what they're attempting.

\subsubsection{Defaulting To Field Skills}

In some cases, a character may not possess the particular
field skill that a test calls for, but they may be
skilled in another related field. For example, a test to 
conduct an alien autopsy might call for an Academics: 
Xenobiology roll, but a character who doesn't have 
that skill may be allowed to default to Academics: 
Biology instead. The gamemaster decides if and when 
to allow this, perhaps applying a modifier to the test 
based on the difference between fields.

\subsubsection{Defaulting To Related Skills}

If the gamemaster allows it, characters may default to 
a related skill that also has some relevance to the test 
at hand. For example, a character skilled in Kinetic 
Weapons might not be trained in the use of a laser, but 
they know enough to point at the target and pull the 
trigger. Likewise, a character might not be skilled in 
Investigation, but the gamemaster could still allow 
them to use their Perception skill instead in order to 
realize that a body had been moved from the place 
where it had been shot. In situations like this, when the 
gamemaster allows defaulting to a related skill, a –30 
modifier should be applied to the test.

\subsection{Complementary Skills}

Sometimes more than one skill may apply to a particular
test, or knowledge in one area can aid your skill
in another. In this case, the gamemaster may apply a 
modifier to the skill test based on the strength of the 
complementing skill, as noted on the Complementary 
Skill Bonus table.

01–30

+10

31–60

+20

61+

+30
%%% 176

\subsection{Skill Ranges}

What is the difference between being a clumsy neophyte
wobbling in zero gravity and being a veteran
gliding effortlessly through space as though you were 
dancing? The answer is training and skill. The greater 
your skill, the more likely you are to not only succeed 
at what you want to do, but succeed well.
Aptitudes in \textit{Eclipse Phase} range from 1 to 30, 
while learned skills range from 0 to 99. These numbers
are an abstraction of the range of transhuman
abilities and traits. The Aptitude Range table provides 
a breakdown of different aptitude levels and how they 
relate to each other. Likewise, the Learned Skill Range 
table provides an interpretation for the capabilities at 
different skill levels.

\section{Aptitudes}

There are 7 aptitudes in \textit{Eclipse Phase, }described on p. 
123\textit{.} Each character has these aptitudes at a minimum 
rating of 1.

\subsection{Aptitude-Only Tests}

In rare cases, a test may call for using an aptitude 
only, rather than a learned skill. This should only 
occur when no learned skills are appropriate to the 
test, and these circumstances are usually noted in 
the rules.
Aptitude-only tests must be handled carefully, as 
the range of aptitude ratings (1–30) is typically much 
smaller than the rating of learned skills (0–99). For 
this reason, most aptitude tests should use a target 
number equal to the aptitude x 3. In rare cases where 
1–5
child average

inept
6–10
adult average

weak
11–15
transhuman average

fit

coordinated
16–20
enhanced

enhanced
21–25
superhuman

gifted
26-30
posthuman

elite
the test is more difficult, the gamemaster may simply 
use an aptitude x 2, or just the straight aptitude rating. 
In some cases, more than one aptitude may be relevant
to the test, and so they may be added together
to derive the target number.
What follows are a few examples where an aptitude-only
test might be appropriate. Gamemasters may call
for similar tests in other situations, but learned skills 
should be used whenever possible.
clumsy
slow
limited

aware

awkward

distracted
able
paced
intelligent

perceptive

personable

controlled
swift
bright

sharp

charismatic

focused
agile
fast
learned

uncanny

dazzling

resolute
nimble
lightning
brilliant

prescient

mesmerizing

unwavering
unerring
synaptic
genius

near omniscient

hypnotic

unshakable
00
No exposure or familiarity, completely unskilled
10
Very rudimentary knowledge
20
Basic operator's proficiency (driver's license, gun permit, 
high school diploma)
30
Hands-on experience, some professional training
40
Basic professional certification (police driving, army riﬂ e 
certified, college diploma)
50
Experience from professional-level work, some 
advanced training
60
Expert competence (competitive driver, marksman, PhD)
70
Experience from expert-level work, has had unique innovations
or insights
80
Worthy of being a system-renowned authority on 
the subject
90
Nobel/Olympic/grandmaster
99
Pinnacle of current understanding and innovation
%%% 177

\subsubsection{Brute Strength}

Any test that involves simple brute strength can be 
handled as an SOM x 3 Test. Use this when smashing 
down a door, breaking an item in half, engaging in a 
tug-of-war, or lifting and carrying a heavy item.

\subsubsection{Catching Thrown Objects}

Use REF + (COO x 2) any time you need to catch a 
thrown or dropped object, such as catching a baseball, 
saving a priceless vase from shattering, or throwing 
back a grenade (see p. 200).

\subsubsection{Composure And Resolve}

Various game situations may frighten your character, 
turn their stomach, horrify them, or rattle them to the 
core of their being. Use WIL x 3 to determine if your 
character can hold their ground, keep it down, and 
pull themselves together.

\subsubsection{Escape Artist}

If a character wants to slip free of physical bonds (such 
as ropes or handcuffs) or otherwise contort themselves 
(such as wriggling out from under a collapsed wall or 
an overturned vehicle), an Escape Artist Test may be 
called for using the character's COO + SOM. Apply 
modifiers appropriate to the difficulty of the situation. 
At the gamemaster's discretion, escaping from some 
restraining situations may be considered a Task Action 
with an appropriate timeframe.

\subsubsection{Having An Idea}

Sometimes the players miss the obvious, or their personal
mindset or biases cause them to misinterpret a
situation or understand events in a way different from 
how the actual character would. In cases like this, the 
gamemaster can call for an INT x 3 or COG x 3 roll 
(whichever is more appropriate) to determine if the 
character gets an idea that will help them along. This 
test should be used sparingly, and only for assessing 
the character's interpretation of obvious and known 
facts and details.

\subsubsection{Memorizing And Remembering}

Memories are what egos use to maintain continuity of 
self from morph to morph, but humans are notorious 
for remembering things incorrectly. Whenever characters
attempt to recall a memory or memorize some
piece of information, use COG x 3 to determine how 
well they succeed. Note that characters with eidetic 
memory (p. 146 or 301) or mnemonic augmentation 
(p. 307) have perfect memory, so no test is required.
%%% 178

\section{Complete Skill List}

This section details all of the learned skills available
in \textit{Eclipse Phase}. Gamemasters and players
may, of course, agree to add additional skills to 
this list as appropriate to their campaign.

\subsection{Academics: [Field]}

\textbf{Type:} Field, Knowledge
\textbf{Linked Aptitude:} COG

\textbf{What it is:} Academics covers any sort of specialized
non-applied knowledge you can only get
through intensive education. Most theoretical and 
applied sciences, social sciences, transhumanities, 
etc. are covered by this skill. Most of the other 
skills listed in this chapter could also be taken as 
an Academics field, reflecting a working theoretical 
knowledge of the skill—for example, Academics: 
Armorer or Academics: Interrogation.

\textbf{When you use it:} Academics is used when a 
character wishes to call upon a specific body of 
knowledge. For example, Academics: Chemistry 
could be used to identify a particular substance, 
understand an unusual chemical reaction, or determine
what elements are needed to nanofabricate
something that requires exotic materials. At the 
gamemaster's discretion, some Academics-related 
tests might not be defaultable, given that only 
someone who has been educated in that subject is 
likely to be able to tackle it.
\textbf{Sample Fields:} Archeology, Astrobiology, As-

tronomy, Astrophysics, Astrosociology, Bio-

chemistry, Biology, Botany, Computer Science, 
Academics: [Field]

COG

Knowledge
Animal Handling

SAV

Active, Social
Art: [Field]

INT

Knowledge
Beam Weapons

COO

Active, Combat
Blades

SOM

Active, Combat
Climbing

SOM

Active, Physical
Clubs

SOM

Active, Combat
Control
WIL (no defaulting)

Active, Mental, Psi
Deception

SAV

Active, Social
Demolitions
COG (no defaulting)

Active, Technical
Disguise

INT

Active, Physical
Exotic Melee Weapon: [Field]

SOM

Active, Combat
Exotic Ranged Weapon: [Field]

COO

Active, Combat
Flight

SOM

Active, Physical
Fray

REF

Active, Combat
Free Fall

REF

Active, Physical
Freerunning

SOM

Active, Physical
Gunnery

INT

Active, Combat
Hardware: [Field]

COG

Active, Technical
Impersonation

SAV

Active, Social
Infiltration

COO

Active, Physical
Infosec
COG (no defaulting)

Active, Technical
Interest: [Field]

COG

Knowledge
Interfacing

COG

Active, Technical
Intimidation

SAV

Active, Social
Investigation

INT

Active, Mental
Kinesics

SAV

Active, Social
Kinetic Weapons

COO

Active, Combat
Language: [Field]

INT

Knowledge
Medicine: [Field]

COG

Active, Technical
Navigation

INT

Active, Mental
Networking: [Field]

SAV

Active, Social
Palming

COO

Active, Physical
Perception

INT

Active, Mental
Persuasion

SAV

Active, Social
Pilot: [Field]

REF

Active, Vehicle
Profession: [Field]

COG

Knowledge
Programming
COG (no defaulting)

Active, Technical
Protocol

SAV

Active, Social
Psi Assault
WIL (no defaulting)

Active, Mental, Psi
Psychosurgery

INT

Active, Technical
Research

COG

Active, Technical
Scrounging

INT

Active, Mental
Seeker Weapons

COO

Active, Combat
Sense
INT (no defaulting)

Active, Mental, Psi
Spray Weapons

COO

Active, Combat
Swimming

SOM

Active, Physical
Throwing Weapons

COO

Active, Combat
Unarmed Combat

SOM

Active, Combat
%%% 179

Cryptography, Economics, Engineering, Genetics, 

Geology, Linguistics, Mathematics, Memetics, 

Nanotechnology, Old Earth History, Physics, Politi-

cal Science, Psychology, Sociology, Xeno-archeology, 

Xenolinguistics, Zoology
\textbf{Specializations:} As appropriate to the field

\subsection{Animal Handling}

\textbf{Type:} Active, Social
\textbf{Linked Aptitude:} SAV

\textbf{What it is:} Skilled animal handlers are able to 
train and control a wide variety of natural and transgenic
animals, including partial uplifts. Though many
animal species went extinct during the Fall, a few 
``ark'' and zoo habitats keep some species alive, and 
many others can be resurrected from genetic samples. 
Exotic animals are considered a sign of prestige 
among the hypercorp elites, and guard animals are 
occasionally used to protect high-security installations. 
Likewise, many habitats and settlements employ small 
armies of partially uplifted, genetically modified, and 
behavior-controlled creatures for sanitation or other 
purposes. Many new and strange breeds of animal are 
created daily to serve a variety of roles.

\textbf{When you use it:} Animal Handling is used whenever 
you are trying to manipulate an animal, whether your 
intent is to calm it down, keep it from attacking, intimidate
it, acquire its trust, or goad it into attacking.
Your Margin of Success determines how effective you 
are at convincing the creature. At the gamemaster's 
discretion, modifiers may be applied to the test. Likewise
winning an animal over may sometimes take
time, and so could be handled as a Task Action with a 
timeframe of five minutes or more.
\textbf{Specializations:} Per animal species (dogs, horses, smart 

rats, etc.)

\subsubsection{Training Animals}

Training animals is a time-consuming task requiring 
repeated efforts and rewards to reinforce the trained 
behavior. Treat this as a Task Action with a timeframe 
of one day to one month, depending on the complexity
of the action. Apply modifiers to this test based on
the relative intelligence of the animal being trained, 
how domestic it is, and the complexity of the task.

Once an animal has been trained, commanding 
it is treated as a Simple Success Test (p. 118) except 
for unusual or stressful situations, in which case the 
trainer receives a +30 modifier on their Animal Handling
Tests when convincing the animal to complete
the trained action.

\subsection{Art: [Field]}

\textbf{Type: }Field, Knowledge
\textbf{Linked Aptitude:} INT

\textbf{What it is:} Art confers the ability to create and evaluate
artistic endeavors. This is a particularly useful
skill in \textit{Eclipse Phase,} especially in the post-scarcity 
economies where creativity and vision can be a key 
component to a character's reputation.
\textbf{When you use it:} The Art skill can be used to either 
create a new work of art or to duplicate an existing 
piece of art in the hopes of passing it off as your 
own. The skill can also determine the approximate 
value of a piece of art either on the open market, for 
monetary exchange systems, or in terms of reputation
for the artist.
\textbf{Sample Fields:} Architecture, Criticism, Dance, Drama, 
Drawing, Painting, Performance, Sculpture, Simulspace
Design, Singing, Speech, Writing
\textbf{Specializations:} As appropriate to the field

\subsection{Beam Weapons}

\textbf{Type:} Active, Combat
\textbf{Linked Aptitude:} COO
\textbf{What it is:} The Beam Weapons skill covers the usage 
and maintenance of standard coherent beam energy 
weapons such as lasers, particle beam weapons, 
plasma rifles, and microwave weapons (p. 338).
\textbf{When you use it:} A player uses their Beam Weapons
skill whenever attacking with a beam weapon in
combat (p. 191). Beam Weapons may also be used for 
tests involving maintenance of the weapon, but not 
for repairing or modifying the weapon (that would be 
Hardware: Armorer skill).
\textbf{Specializations:} Lasers, Microwave Weapons, Particle 
Beam Weapons, Plasma Rifles

\subsection{Blades}

\textbf{Type:} Active, Combat
\textbf{Linked Aptitude:} SOM
\textbf{What it is:} The Blades skill covers the usage and 
maintenance of standard bladed weapons (p. 334).
\textbf{When you use it:} A player uses their Blades skill 
whenever attacking with a blade weapon in melee 
combat (p. 191). Blades may also be used for tests 
involving maintenance of the weapon, but not for 
repairing or modifying the weapon (that would be 
Hardware: Armorer skill). This skill is used for blade 
weapons implanted in the body at the end of an appendage
(hands, forearms, feet, octomorph arms, etc.),
but the Exotic Melee Weapon skill is used for blades 
implanted in other parts of the body.
\textbf{Specializations:} Axes, Implant Blades, Knives, Swords

\subsection{Climbing}

\textbf{Type:} Active, Physical
\textbf{Linked Aptitude:} SOM
\textbf{What it is:} Climbing is the skill of ascending and 
descending sheer surfaces with or without the aid of 
specialized equipment.
\textbf{When you use it:} This skill is used whenever a character
wishes to scale a climbable surface. For heights
greater than one story, climbing is handled as a Task 
Action with a timeframe equivalent to one meter per 
Action Phase. For rappelling, the timeframe for descent
is 50 meters per Action Turn. Climbing gear (p.
332-333) provides appropriate modifiers.
\textbf{Specializations:} Assisted, Freehand, Rappelling

nd

of

r-

s
%%% 180

\subsection{Clubs}

\textbf{Type:} Active, Combat
\textbf{Linked Aptitude:} SOM

\textbf{What it is:} The Clubs skill covers the usage and 
maintenance of standard blunt melee weapons such 
as batons or sticks (see p. 334).

\textbf{When you use it:} Players use their Clubs skill whenever
they want to attack with a blunt weapon in melee
combat (p. 191). The Clubs skill may also be used for 
tests involving maintenance of the weapon, but not 
for repairing or modifying the weapon (that would be 
Hardware: Armorer skill).
\textbf{Specializations:} Batons, Hammers, Staffs

\subsection{Control}

\textbf{Type:} Active, Mental, Psi
\textbf{Linked Aptitude:} WIL (no defaulting)

\textbf{What it is:} Control is the use of psi to manipulate 
individuals or actively penetrate their mental processes
This skill is only available to characters with
the Psi trait (p. 147).

\textbf{When you use it:} Use Control when taking a psionic 
tour through a foreign ego—messing around included. 
See \textit{Psi,} p. 220.
\textbf{Specializations:} By sleight

\subsection{Deception}

\textbf{Type:} Active, Social
\textbf{Linked Aptitude:} SAV

\textbf{What it is:} Deception is your ability to act, bluff, 
con, fast talk, lie, misrepresent, and pretend. Accomplished
users of deception are able to convince anyone
of nearly anything. This skill does not include using a 
physical disguise to appear to be another person (the 
Impersonate skill covers that area).

\textbf{When you use it:} Use this skill whenever you want 
to deceive someone with words or gestures. A successful
skill test means that you have passed off your deception
convincingly. At the gamemaster's discretion,
someone who is actively alert for signs of deception 
may make an Opposed Test using the Kinesics skill.
\textbf{Specializations:} Acting, Bluffing, Fast Talk

\subsection{Demolitions}

\textbf{Type:} Active, Technical
\textbf{Linked Aptitude:} COG (no defaulting)

\textbf{What it is:} Demolitions covers the use of controlled 
explosives.

\textbf{When you use it:} Use it when making, placing, and 
disarming explosives and explosive devices. See \textit{De-}
\textit{molitions,} p. 197.
\textbf{Specializations:} Commercial Explosives, Disarming, 

Improvised Explosives

\subsection{Disguise}


\textbf{Type:} Active, Physical

\textbf{Linked Aptitude:} INT

\textbf{What it is: }Disguise is the art of physically altering 
your appearance so that you look like someone else. 
This includes both the use of props (wigs, contacts, 
skin pigments) and the altering of subtle physical 
characteristics (gait, posture, poise).

\textbf{When you use it:} Use Disguise to fool someone into 
thinking you're someone you're not. This can be used 
to hide your identity or to make yourself look like 
someone in particular. When used against someone 
who knows your true look or the appearance of the 
person you are imitating, this is handled as an Opposed
Test against Perception or Investigation.
\textbf{Specializations:} Cosmetic, Theatrical

\subsection{Exotic Melee Weapon: [Field]}

\textbf{Type:} Field, Active, Combat
\textbf{Linked Aptitude:} SOM

\textbf{What it is:} The Exotic Melee Weapon skill covers 
the use and maintenance of all melee weapons not 
covered by the Clubs or Blades skills (see p. 334).

\textbf{When you use it:} Use the Exotic Melee Weapon skill 
when attacking someone with an exotic melee weapon 
in melee combat (p. 191).
\textbf{Sample Fields:} Morning Star, Spear, Whip
\textbf{Specializations:} N/A

\subsection{Exotic Ranged Weapon: [Field]}

\textbf{Type:} Field, Active, Combat
\textbf{Linked Aptitude:} COO

\textbf{What it is:} Exotic Ranged Weapon skill includes the 
use and maintenance of all ranged weapons not covered
by the Beam, Kinetic, Seeker, Spray, or Throwing
Weapons skills.

\textbf{When you use it:} Use this skill whenever attacking with 
an exotic ranged weapon in ranged combat (p. 191).
\textbf{Sample Fields:} Blowgun, Crossbow, Flamethrower, 

Slingshot
\textbf{Specializations:} N/A

\subsection{Flight}

\textbf{Type:} Active, Physical
\textbf{Linked Aptitude:} SOM

\textbf{What it is:} Flight is the skill of using your body to 
fly. This skill is used when sleeved in or jamming a 
winged or otherwise flight-capable morph (manual and 
remote-control flight are handled using Pilot skill).

\textbf{When you use it:} Use this skill whenever you need to 
make an aerial maneuver, land in difficult conditions, 
maintain your course in steep winds, or otherwise 
keep from crashing or falling.
\textbf{Specializations:} Diving, Landing, Takeoff, specific 

maneuvers

\subsection{Fray}

\textbf{Type:} Active, Combat
\textbf{Linked Aptitude:} REF

\textbf{What it is:} Fray is the ability to get out of the way 
of incoming attacks, debris, or inconvenient passers-by. 
Characters that have a high Fray score are able to react 
quicker than others when dodging or maneuvering.

\textbf{When you use it:} Whenever a character is physically 
attacked by an opponent in melee combat, roll Fray to 
avoid getting hit (see p. 191). Fray may also be used 
%%% 181
to dodge other events that may harm the character, 
such as avoiding a charging vehicle or jumping out of 
the way of a collapsing stack of crates.
\textbf{Specializations:} Blades, Clubs, Full Defense, Unarmed

\subsection{Free Fall}

\textbf{Type:} Active, Physical
\textbf{Linked Aptitude:} REF

\textbf{What it is:} Free Fall is about moving in free-fall and 
microgravity environments.

\textbf{When you use it:} Use whenever you need to maneuver
in a zero-g situation, such as propelling yourself
across a large open space or making sure you don't 
accidentally send yourself spinning off into space. 
Free Fall is also used when moving with spacesuit 
maneuvering jets and when parachuting.
\textbf{Specializations:} Microgravity, Parachuting, Vacsuits

\subsection{Freerunning}

\textbf{Type:} Active, Physical
\textbf{Linked Aptitude:} SOM

\textbf{What it is:} Freerunning is part running, part gymnastics
It is about moving fast, maneuvering over/under/
around/through obstacles, and placing your body 
where it needs to go. Freerunning/parkour is a popular 
pastime in habitats where open space is limited.

\textbf{When you use it:} Use Freerunning whenever you 
need to overcome an obstacle via movement, such 
as hurdling a railing, rolling across the hood of a 
car, jumping across a pit, or swinging around a pole. 
Freerunning is also used for sprinting (p. 191) and full 
defense against attacks (p. 198).
\textbf{Specializations:} Balance, Gymnastics, Jumping, Running

\subsection{Gunnery}

\textbf{Type:} Active, Combat
\textbf{Linked Aptitude:} INT

\textbf{What it is:} Gunnery skill covers the use and maintenance
of large, vehicular, or non-portable weapons
systems. Firing these weapons is more like playing a 
video game than firing a gun.

\textbf{When you use it:} Use Gunnery when attacking with 
a vehicle-mounted weapon or weapon emplacement 
in ranged combat (p. 191).
\textbf{Specializations:} Artillery, Missiles

\subsection{Hardware: [Field]}

\textbf{Type:} Field, Active, Technical
\textbf{Linked Aptitude:} COG

\textbf{What it is:} This skill encompasses the ability to 
build, repair, physically hack, and upgrade equipment 
of a specific type.

\textbf{When you use it:} Hardware is primarily used to repair 
devices, vehicles, habitat systems, or synthetic morphs. 
See \textit{Building, Repairing,} and \textit{Modifying} below.
\textbf{Sample Fields:} Aerospace (all air and space vehicles), 

Armorer (armor and weapons), Electronics (all 

computerized devices), Groundcraft, Implants, In-

dustrial (habitat, factory, and life support systems), 

Nautical (watercraft and submarines), Robotics 

(synthetic morphs)
\textbf{Specializations:} As appropriate to the field

\subsubsection{Building}

Creating an item from scratch is handled as a Task 
Action with a timeframe determined by the gamemaster
The timeframe should be set according to the complexity
of the object and could range from an hour
(constructing a set of shelves) to days (assembling a 
robot from spare parts) to even months (building a 
house). Numerous factors may apply modifiers  to 
the test, such as the use of entoptic blueprints/help 
manuals (+20) or poor working conditions (–10 to 
–30). Tools are also a factor, perhaps making the job 
easier (superior tools +10 to +30), more difficult (poor 
or inadequate tools, –10 to –30), or even impossible 
(lack of required tools).

\subsubsection{Repairing}

Damaged items may be repaired in a similar manner. See 
the rules for \textit{Synthmorph and Object Repair,} p. 209.

\subsubsection{Modifying}

Altering an object's design and function follows the 
same basic rules as build and repair, above. The timeframe
is determined by the gamemaster as appropriate
to the modification.
%%% 182

\subsection{Impersonation}

\textbf{Type:} Active, Social
\textbf{Linked Aptitude:} SAV

\textbf{What it is:} Impersonation is the skill of trying to 
pass yourself off as someone else in social situations, 
including virtual ones. This includes copying mannerisms
and speech patterns and using accumulated information
to convince others that you are that person.
In a universe where appearance is highly variable, the 
question of identity is largely one of both trust and 
picking up on behavioral quirks and verbal cues to 
recognize a given individual.

\textbf{When you use it:} Sometimes it's fun to pretend 
you're someone else, and sometimes it's profitable 
or lifesaving. Use this skill whenever you attempt 
to convince someone that you are actually someone 
else through some sort of social or online interaction. 
Forks use this skill when passing themselves off as 
their alpha ego. Impersonate is handled as an Opposed
Test against the Kinesics skill.
\textbf{Specializations:} Avatar, Face-to-Face, Verbal

\subsection{Infiltration}

\textbf{Type:} Active, Physical
\textbf{Linked Aptitude:} COO

\textbf{What it is:} Infiltration is the art of escaping 
detection.

\textbf{When you use it:} Use Infiltration whenever you 
need to physically hide or move with stealth to avoid 
someone sensing you, whether you are hiding behind 
a tree, sneaking past a guard, or blending into a 
crowd. Infiltration can also be used to follow people 
(shadowing) without them detecting you. Infiltration 
is an Opposed Test against the Perception of whomever
you are hiding from. The gamemaster may wish
to roll such tests in secret so the player does not know 
whether they have succeeded or failed.
\textbf{Specializations:} Blending In, Hiding, Shadowing, 

Sneaking

\subsection{Infosec}

\textbf{Type:} Active, Technical
\textbf{Linked Aptitude:} COG (no defaulting)

\textbf{What it is:} Infosec is short for ``information security
It encompasses training in electronic intrusion
and counterintrusion techniques, as well as encryption 
and decryption.

\textbf{When you use it:} Infosec is used both for hacking 
into electronic devices and mesh networks and for 
protecting them. See the \textit{Mesh} chapter, p. 234, for 
more details.
\textbf{Specializations:} Brute-Force Hacking, Decryption, 

Probing, Security, Sniffing, Spoofing

\subsection{Interest: [Field]}

\textbf{Type:} Field, Knowledge
\textbf{Linked Aptitude:} COG

\textbf{What it is:} Interest includes just about any topic 
that captures your attention that isn't covered by 
another skill. This includes hobbies, obsessions, 
causes, pastimes, and other recreational pursuits.

\textbf{When you use it:} Use the Interest skill whenever 
you need to recall or use knowledge related to the 
particular interest in question.

\textbf{Field Examples:} Ancient Sports, Celebrity Gossip, 
Conspiracies, Factor Trivia, Gambling, Hypercorp 
Politics, Lunar Habitats, Martian Beers, Old Earth 
Nation-States, Reclaimer Blogs, Science Fiction, Scum 
Drug Dealers, Spaceship Models, Triad Economics, 
Underground XP
\textbf{Specializations:} As appropriate to the field

\subsection{Interfacing}

\textbf{Type:} Active, Technical
\textbf{Linked Aptitude:} COG

\textbf{What it is:} Interfacing is about using computerized 
electronic devices and software.

\textbf{When you use it:} Use Interfacing to understand an 
electronic device you are not familiar with, use a program
according to its normal operating parameters,
manipulate electronic files of various types (including 
images, video, XP, and audio files), scan for wireless 
devices, and otherwise interact with and command 
your ecto, muse, and other computerized devices. 
Some Interfacing actions may be Task Actions, with 
a timeframe determined by the gamemaster. For more 
detail, see the \textit{Mesh} chapter, p. 234.
\textbf{Specializations:} Forgery, Scanning, Stealthing, by 

program

\subsection{Intimidation}

\textbf{Type:} Active, Social
\textbf{Linked Aptitude:} SAV

\textbf{What it is:} Intimidation is convincing someone to 
do what you want based on direct threats (implied or 
actual) or sheer force of personality.

\textbf{When you use it:} Use Intimidation to scare someone 
into submission, browbeat them into getting your way, 
command them to follow your orders, or berate them 
into giving up information. Influence is handled as an 
Opposed Test, pitted against the target's WIL + WIL 
+ SAV.
\textbf{Specializations:} Interrogation, Physical, Verbal

\subsection{Investigation}

\textbf{Type:} Active, Mental
\textbf{Linked Aptitude:} INT

\textbf{What it is:} Investigation is the art of analyzing 
evidence, piecing together clues, solving mysteries, 
and making logical deductions from groups of facts. 
Investigation differs from Perception in that it is the 
conscious search for clues or pieces of a puzzle.

\textbf{When you use it:} Use Investigation to draw conclusions
from assorted details. For example, Investigation
could be used to determine the likely sequence
of events at a crime scene, determine a possible social 
connection between two people, or deduce how an 
enemy made their escape. Investigation is a great way 
%%% 183
to provide clues to players, especially when the subject 
matter is something their character might know well 
but the player does not.
\textbf{Specializations:} Evidence Analysis, Logical Deductions, 

Physical Investigation, Physical Tracking

\subsection{Kinesics}

\textbf{Type:} Active, Social
\textbf{Linked Aptitude:} SAV

\textbf{What it is:} Kinesics is the art of empathy and nonvocal
communication.

\textbf{When you use it:} Use Kinesics to read body language, 
tells, social cues, and other subconscious indicators. It 
can also be used to emote more effectively. Kinesics 
is used defensively whenever someone is trying to deceive
you; make an Opposed Test against that person's
Deception or Impersonation skill.

Though synthetic morphs are also designed to 
emote, reading them is not as easy. Apply a –30 modifier
when judging a synthetic morph inhabited by a
character or AGI. Likewise, standard AIs are also difficult
to read; apply a –60 modifier when judging a
synthetic morph or pod operated by an AI.
\textbf{Specializations:} Judge Intent, Nonvocal Communication

\subsubsection{Judging Emotions And Intentions}

Transhumans are empathic beings, and so you can 
attempt to gauge the demeanor and/or intent of 
someone you are dealing with by rolling a Kinesics 
Test. This attempt to read someone is far from exact, 
however, and it is easy to misjudge. The gamemaster 
should make this test in secret and only allow a hint 
if successful—it is not possible to read someone with 
absolute certainty. If the person being judged is intentionally
trying to deceive the character, this should be
an Opposed Test against their Deception skill.

\subsubsection{Nonvocal Communication}

Experts in Kinesics can effectively communicate 
with each other simply by posture, stances, gestures, 
demeanors, and looks. Such communication is necessarily
limited in the amount of information it can
convey, but feelings, attitudes, affirmation/negation, 
and simple concepts may be passed. To effectively 
communicate complex concepts, the gamemaster may 
require successful Kinesics Tests from both parties, applying
modifiers as appropriate.

\subsection{Kinetic Weapons}

\textbf{Type:} Active, Combat
\textbf{Linked Aptitude:} COO

\textbf{What it is:} Kinetic Weapons covers the use and 
maintenance of standard kinetic projectile weapons 
like firearms and railguns (p. 335).

\textbf{When you use it:} Use this skill whenever attacking 
with a kinetic weapon in ranged combat (p. 191).
\textbf{Specializations:} Assault Rifles, Machine Guns, Pistols, 

Sniper Rifles, Submachine Guns

\subsection{Language: [Field]}

\textbf{Type:} Field, Knowledge
\textbf{Linked Aptitude:} INT
\textbf{What it is:} Language covers the speaking and 
reading of languages other than the player's native 
tongue. A speaker is considered fluent at a skill level 
of 50; anything above this indicates further refinement
in technical vocabulary, accents, and knowledge
of dialects.
\textbf{When you use it:} Use the Language skill whenever 
you want to speak, understand, or read something in 
a language at which you are skilled. Most speaking 
and reading comprehension tests can be considered 
Simple Success Tests if your skill is over 50, unless the 
gamemaster rules the subject is sufficiently complex 
that a non-native speaker would have trouble understanding
it.
\textbf{Sample Fields:} Arabic, Cantonese, English, French, 
Hindi, Japanese, Mandarin, Portuguese, Russian, 
Spanish
\textbf{Specializations:} As appropriate to the field, representing
dialects, technical jargon, and subcultural slang

the

ex

er-

h, 

,
%%% 184

\subsection{Medicine: [Field]}

\textbf{Type:} Field, Active, Technical
\textbf{Linked Aptitude:} COG

\textbf{What it is:} Medicine is the applied care and maintenance
of biological beings and life.

\textbf{When you use it:} Use Medicine whenever you need 
to apply medical care beyond the immediate help 
provided by first responders. This includes conducting
physical exams, diagnosing ailments, treating
problems and illnesses, surgery, using biotech and 
nanotech medical tools, and long-term care. See \textit{Heal-}
\textit{ing and Repair,} p. 208.
\textbf{Sample Fields:} Biosculpting, Exotic Biomorphs, 

Gene Therapy, General Practice, Implant Surgery, 

Nanomedicine, Mercurials (by type), Paramedic, 

Pods, Psychiatry, Remote Surgery, Trauma Surgery, 

Veterinary
\textbf{Specializations:} As appropriate to the field

\subsection{Navigation}

\textbf{Type:} Active, Mental
\textbf{Linked Aptitude:} INT

\textbf{What it is:} Navigation is the art of finding your way, 
whether using AR maps, a compass, the stars, or an 
astrogation AI.

\textbf{When you use it:} Use Navigation whenever you 
need to plot out a course, determine a direction, or 
otherwise keep from getting lost.
\textbf{Specializations:} Astrogation, Map Making, Map 

Reading

\subsection{Networking: [Field]}

\textbf{Type:} Active, Social
\textbf{Linked Aptitude:} SAV

\textbf{What it is:} Networking is your skill at working your 
contacts, trading favors, and keeping your finger on 
the pulse of a particular faction or cultural grouping.

\textbf{When you use it:} Use Networking to gather information
or call on services using your Reputation (see
\textit{Reputation and Social Networks,} p. 285).
\textbf{Sample Fields:} Autonomists (@-rep), Criminals (g-rep), 

Ecologists (e-rep), Firewall (i-rep), Hypercorps (c-

rep), Media (f-rep), Scientists (r-rep). At the game-

master's discretion, this list can be expanded to 

other (sub)cultural groupings.
\textbf{Specializations:} As appropriate to each field

\subsection{Palming}

\textbf{Type:} Active, Physical
\textbf{Linked Aptitude:} COO

\textbf{What it is:} Palming is the skill of handling items quickly
and nimbly without others noticing. Palming is not
only about dexterous manipulation of objects but also 
relies heavily on obfuscation, timing, and misdirection.

\textbf{When you use it:} Use Palming any time you are 
trying to conceal an item on your person, shoplift, 
pick a pocket, surreptitiously discard something, or 
perform a magic trick. Palming is an Opposed Test 
against the Perception of any onlookers. The gamemaster
may wish to make this roll secretly.
\textbf{Specializations:} Pickpocketing, Shoplifting, Tricks

\subsection{Perception}

\textbf{Type:} Active, Mental
\textbf{Linked Aptitude:} INT

\textbf{What it is:} Perception is the use of your physical 
senses (including cybernetic) and awareness of the 
physical world around you. Perception differs from 
Investigation in that it is noticing things by chance, 
rather than actively searching for something.

\textbf{When you use it:} Use Perception whenever you 
wanted to take a detailed account of your surroundings
(see \textit{Detailed Perception,} p. 190). Perception
can also be considered an Automatic Action (see 
\textit{Basic Perception, }p. 190) and so the gamemaster may 
call for a Perception Test to determine if you notice 
something; it is recommended that such tests be rolled 
secretly by the gamemaster. Perception is also used as 
an Opposed Test whenever someone around you is 
trying to be sneaky with Infiltration or Palming.
\textbf{Specializations:} Aural, Olfactory, Tactile, Taste, Visual
%%% 185

\subsection{Persuasion}

\textbf{Type:} Active, Social
\textbf{Linked Aptitude:} SAV

\textbf{What it is:} Persuasion is the art of convincing 
someone to do what you want through the use of 
words and gestures. This does not include persuasion 
through threats or force (that is covered by Intimidation
or by lying (covered by Deception).

\textbf{When you use it:} Use Persuasion any time you are 
trying to bargain with, convince, or manipulate someone
This can include motivating your subordinates or
peers to take action, seducing a companion, winning 
a political debate, or negotiating a contract, among 
other things. Persuasion is handled as an Opposed 
Test against the target's WIL + WIL + SAV when one 
person is simply trying to win over another. If both 
parties are trying to convince each other, make it an 
Opposed Test between Persuasion skills.
\textbf{Specializations:} Diplomacy, Morale Boosting, Negoti-

ating, Seduction

\subsection{Pilot: [Field]}

\textbf{Type: }Field, Active, Vehicle
\textbf{Linked Aptitude:} REF

\textbf{What it is:} Pilot is your skill at driving/flying a vehicle
of a particular type.

\textbf{When you use it:} You use Pilot skill whenever you 
need to maneuver, control, or avoid crashing a vehicle, 
whether you are in the pilot's seat, remote controlling 
a robot, or directly jamming a vehicle with VR. Each 
vehicle has a Handling modifier that applies to this 
test, along with other situational modifiers (see \textit{Bots, }
\textit{Synthmorphs, and Vehicles,} p. 195).
\textbf{Sample Fields:} Aircraft, Anthroform (walkers), Exotic 

Vehicle, Groundcraft (wheeled or tracked), Space-

craft, Watercraft
\textbf{Specializations:} As appropriate to the field

\subsection{Profession: [Field]}

\textbf{Type:} Field, Knowledge
\textbf{Linked Aptitude:} COG

\textbf{What it is:} Profession skills indicate training in a 
profession practiced in \textit{Eclipse Phase.} This can indicate
either formal training or informal, on-the-job type
training, and includes both legal and extralegal trades.

\textbf{When you use it:} Use Profession to perform work-related
tasks for a specific trade (i.e. mining, balancing
accounts, designing a security system, etc.) or to reference
specialized knowledge that someone trained in
that profession might have.
\textbf{Sample Fields:} Accounting, Appraisal, Asteroid Pros-

pecting, Banking, Cool Hunting, Con Schemes, 

Distribution, Forensics, Lab Technician, Mining, 

Police Procedures, Psychotherapy, Security Ops, 

Smuggling Tricks, Social Engineering, Squad Tactics, 

Viral Marketing, XP Production
\textbf{Specializations:} As appropriate to the field

\subsection{Programming}

\textbf{Type:} Active, Technical
\textbf{Linked Aptitude:} COG (no defaulting)
\textbf{What it is:} Programming is your talent at writing 
and modifying software code.
\textbf{When you use it:} Use Programming to write new 
programs, modify or patch existing software, break 
copy protection, find or introduce exploitable flaws, 
write virii or worms, design virtual settings, and so 
on. See the \textit{Mesh} chapter, p. 234. Programming is also 
applied when using nanofabrication devices.
\textbf{Specializations:} AI Code, Malware, Nanofabrication, 
Piracy, Simulspace Code

\subsubsection{Nanofabrication}

Nanofabrication is use of Programming skill to create 
objects using a cornucopia machine, fabber, or maker 
(p. 327). If you have appropriate blueprints and raw 
materials, most uses of a nanofabricator can be treated 
as a Simple Success Test (p. 118). If you wish to create 
an item for which you do not have blueprints or the 
proper raw materials, however, or you wish to alter an 
item's design, then a Nanofabrication Test is called for. 
See \textit{Nanofabrication,} p. 284.
\textbf{Specializations:} Art, Clothing, Electronics, Food, Forgery
Weapons

\subsection{Protocol}

\textbf{Type:} Active, Social
\textbf{Linked Aptitude:} SAV
\textbf{What it is:} Protocol is the art of making a good 
impression in social settings. This includes keeping 
up with the latest memes, trends, gossip, interests and 
habits of various (sub)cultural group.
\textbf{When you use it:} Use Protocol whenever you need 
to choose your words carefully, determine who is the 
appropriate person to speak to, impress someone with 
your grasp of customs, or otherwise fit into a specific 
social/cultural grouping. Part etiquette, part streetwise, 
Protocol allows you to navigate treacherous social 
waters and put people at ease. If the character is dealing
with a suspicious or hostile audience, make this an
Opposed Test against the target's WIL + WIL + SAV.
\textbf{Specializations:} Anarchist, Brinker, Criminal, Factor, 
Hypercorp, Infomorph, Mercurial, Reclaimer, Preservationist
Scum, Ultimate

\subsubsection{Negating Social Gaffes}

Sometimes a player will make a mistake that their 
character never would, whether that's failing to stand 
in the presence of hypercorp royalty, confusing a gang 
leader for a peon, or accidentally insulting someone's 
heritage. In cases like this, the player may make a Protocol
Test for the appropriate field in order to negate
the gaffe. If successful, the character never actually 
screwed up, or at least managed to cover their tracks 
without ruffling any feathers.

eir

nd

ng

's

o
186

\subsection{Psi Assault}

\textbf{Type:} Active, Mental, Psi
\textbf{Linked Aptitude:} WIL (no defaulting)

\textbf{What it is:} Psi Assault is the skill of damaging another
ego's mind. It can only be purchased by characters
with the Psi trait (p. 147).

\textbf{What it does:} Use Psi Assault when attacking another
ego's mind in psi combat.
\textbf{Specializations:} By sleight

\subsection{Psychosurgery}

\textbf{Type:} Active, Technical
\textbf{Linked Aptitude:} INT

\textbf{What it is:} Psychosurgery is the use of machine-aided
psychological techniques to repair, damage, or
manipulate the psyche.

\textbf{When you use it:} Use Psychosurgery to attempt the 
tricky process of editing someone's mind (see \textit{Psycho-}
\textit{surgery,} p. 229). Psychosurgery can be used beneficially
to help patients who remember their deaths, feel
disconnected after remorphing, or have experienced 
other sorts of mental traumas. This skill may also be 
used to interrogate, torture, or otherwise mess with 
captive minds in a VR environment.
\textbf{Specializations:} Memory Manipulation, Personality 

Editing, Psychotherapy

\subsection{Research}

\textbf{Type:} Active, Technical
\textbf{Linked Aptitude:} COG

\textbf{What it is:} Research is the skill for looking up information
on the Mesh: searching, sifting, mining, and
interpreting data. This includes knowing where to look, 
what links to follow, and how to optimize your queries.

\textbf{When you use it:} Use the Research skill whenever 
you need to look up the answer to a question, find 
databases, search archives, or track down anything 
online. Research is typically a Task Action with the 
timeframe and difficulty modifier determined by the 
gamemaster. See the \textit{Online Research,} p. 249.
\textbf{Specializations:} Tracking, by information type

\subsection{Scrounging}

\textbf{Type:} Active, Mental
\textbf{Linked Aptitude:} INT

\textbf{What it is:} Scrounging is your ability to find things, 
particularly things of use or value that are concealed, 
buried, or hard to find. This includes knowing where 
to look and what to look for. Scrounging differs from 
both Perception and Investigation in that it is about 
finding items hidden among others, and in most cases 
about finding something in particular (food, valuables
etc.).

\textbf{When you use it:} Use Scrounging to dumpster-dive a 
meal, search ruins for relics, find bargains at a bazaar, 
forage berries in the forest, locate a spacesuit in an 
abandoned ship, etc. Scrounging is typically handled 
as a Task Action with a timeframe and difficulty 
modifier determined by the gamemaster.
\textbf{Specializations:} Bazaars, Forests, Habitats, Ruins

\subsection{Seeker Weapons}

\textbf{Type:} Active, Combat
\textbf{Linked Aptitude:} COO

\textbf{What it is:} Seeker Weapons covers the use and 
maintenance of seeker launchers (p. 339) and seeker 
missiles (p. 340).

\textbf{When you use it:} Use this skill when attacking with 
a seeker in ranged combat (p. 191).
\textbf{Specializations:} Armband, Pistol, Rifle, Underbarrel

\subsection{Sense}

\textbf{Type:} Active, Mental, Psi
\textbf{Linked Aptitude:} INT (no defaulting)

\textbf{What it is:} Sense is the use of psi to scan egos. Only 
characters with the Psi trait (p. 147) may purchase 
this skill.

\textbf{What it does:} See \textit{Psi,} p. 220.
\textbf{Specializations:} By sleight

\subsection{Spray Weapons}

\textbf{Type:} Active, Combat
\textbf{Linked Aptitude:} COO

\textbf{What it is:} The Spray Weapons skill covers the use 
and maintenance of cone-effect ranged weapons (see 
\textit{Spray Weapons,} p. 340).

\textbf{When you use it:} A player uses their Sonic Weapons 
skill whenever they are attacking with a spray weapon 
in ranged combat (p. 191).
\textbf{Specializations:} Buzzer, Freezer, Shard, Shredder, Torch

\subsection{Swimming}

\textbf{Type:} Active, Physical
\textbf{Linked Aptitude:} SOM

\textbf{What it is:} Swimming is the art of moving and 
not drowning within fluids. It includes floating, 
surface swimming, snorkeling, diving, and related 
equipment use.

\textbf{When you use it:} Use Swimming whenever you 
need to move and survive in water or another liquid 
environment. Swimming in a non-threatening environment
can be handled as a Simple Success Test.
Swimming over a long distance could be handled as 
a Task Action. Diving off a cliff into a lake, preventing
yourself from being swept away in a raging river
current, or making sure you've set a proper gas mix 
for a deep-sea dive, among other things, requires a 
Success Test.
\textbf{Specializations:} Diving, Freestyle, Underwater Diving

\subsection{Throwing Weapons}

\textbf{Type:} Active, Combat
\textbf{Linked Aptitude:} COO

\textbf{What it is:} Throwing Weapons skill covers the use 
and maintenance of standard throwing weapons, like 
grenades (p. 340).

\textbf{When you use it:} Use Throwing Weapons skill 
whenever you are attacking with a throwing weapon 
in ranged combat (p. 191).
\textbf{Specializations:} Grenades, Knives, Rocks
%%% 187

\subsection{Unarmed Combat}

\textbf{Type:} Active, Combat
\textbf{Linked Aptitude:} SOM

\textbf{What it is:} Unarmed Combat is your ability to attack and 
defend without using weapons.

\textbf{When you use it:} Use Unarmed Combat whenever you 
want to attack someone with your fists, feet, elbows, knees, 
or other body parts in melee combat (p. 191).
\textbf{Specializations:} Implant Weaponry, Kick, Punch, Subdual

\section{Special Skills}


While the preceding list represents the skills most commonly 
used in \textit{Eclipse Phase,} there may be certain skills called for in a 
campaign that are not found in this book. In this case, the gamemaster
may work with the players to create a new skill to fill
this void. This option should be used sparingly to prevent skill 
bloat, and all skills are subject to approval by the gamemaster

If you choose to create a new skill, keep in mind that it needs 
to be linked to an existing aptitude and should be a skill that 
is available to all characters, not just specific to one character.
ly 
a 

188
 
