%%% 236
%%% 237
%%% 238
Before the Fall, humanity interfaced with each other 
through the internet, interconnected networks that 
served as the technical backbone for the evolving
world wide web. While it began as a electronic
medium for retrieving information from various 
sources (replacing even older paper-based info-sources
succeeding generations emphasized digital
communities and hosted services such as networking 
sites, wikis, blogs, and folksonomies. These facilitated
openness, collaboration, and sharing, thereby
laying the groundwork for a modern, interconnected 
information society. Further stages emphasized wireless
interaction, geolocation, and semantic web approaches
and achieved quantum leaps in the realm
of user interaction with the advent of brain-computer 
interfaces, augmented reality (AR), virtual reality 
(VR), and experience playback (XP).

This environment, coupled with the exponential 
growth of processing power and memory storage, created
an evolutionary path for the development of intelligent
agents—designed to augment human information
processing—that then transformed into artificial
intelligences (AIs) in the following decades. While these 
``weak'' AIs did not possess the full range of human 
cognitive abilities, tended towards overspecialization, 
and were restrained by programmed limitations, the 
digital evolution toward artificial general intelligences 
(AGIs)—''strong'' AIs with intelligence capabilities 
that equaled or exceeded human abilities—could not 
be halted. From this point it was but a matter of time 
before so-called seed AI would come into existence, 
machine minds capable of recursive self-improvement, 
leading to an exponential growth in intelligence. Unfortunately
for humanity, the TITANs were the result.

Even before the Fall, however, the internet of old 
was transforming into something new. Instead of connecting
via central servers, users were wirelessly linking
to each other, creating a decentralized intermeshed
network of handheld devices, personal computers, 
robots, and electronic devices. Users were online all of 
the time and connected with everything and everyone 
around them in a ubiquitous computing environment. 
This was especially true of those participating in 
humanity's expansion into space. Disconnected from 
the internet due to distance and light-speed communication
lags, these users were nevertheless connected
with all of the people and objects in their nearby 
environment or habitat, creating local wireless mesh 
networks. Thus was the mesh born, taking the place 
of the old internet of earth, lost during the Fall.

\section{Mesh Capabilities}

The mesh, as it exists in \textit{Eclipse Phase,} is only possible 
thanks to major developments made in computer and 
communication technologies and nanofabrication. 
Wireless radio transmitters and receivers are so unobtrusively
tiny that they can literally be factored into
anything. As a result, everything is computerized and 
connected, or at least tagged with a radio frequency 
ID (RFID) chip. Even food is tagged with edible chips, 
complete with expiration date and nutritional content. 
Other communications mediums, such as laser and 
microwave links, add to the information flow.

Data storage technology has advanced to such high 
levels that even an individual user's \textit{surplus }storage 
capacity can maintain an amount of information 
easily surpassing the entire 20th-century internet. 
Lifeloggers can literally record every moment of their 
life and never fear about running out of room. The 
amount of data that people carry around in the mesh 
inserts in their head or in portable ecto personal computers
is staggering.

Processing capabilities also exist at hyper-efficient 
levels. Even massive supercomputers are a thing of the 
past when modest handheld devices can fulfill almost 
all of your needs, even while simultaneously running a 
personal AI assistant, downloading media, uploading 
porn, and scanning thousands of newsfeeds. Within 
the mesh network, devices that near their processing 
limits simply share the burden with devices around 
them, creating a massively distributed framework that 
in some ways is like an entire supercomputer to itself, 
shared by everyone.

Similarly, transmission capacity now far exceeds 
most citizens' definition of need. Anyone born within 
the last several generations has always lived in a world 
in which hyper-realistic, multi-sensory media of nearly 
any length is available for instantaneous download 
or upload from anywhere. Massive databases and 
archives are copied back and forth with ease. Bandwidth
is such a non-issue that most people forget it ever
was. In fact, given the sheer amount of data available, 
\textit{finding }the information or media you're looking for 
takes considerably longer than downloading it. The 
mesh is also \textit{never }down. As a decentralized network, 
if any one device is taken offline, connections merely 
route around it, finding a path via the thousands if not 
millions of available nodes. Similarly, the entire mesh 
behaves like a peer-to-peer network, so that large transfers
are broken into manageable chunks that take independent
routes. In fact, most users maintain personal
torrent archives that are publicly accessible and shared.

Private networks still exist, of course. Some are 
physically walled away behind closed-access wired networks
or even wireless-inhibiting infrastructure that
keep a network isolated and contained. Most, however, 
operate on top of the public mesh, using encrypted 
tunneling protocols that provide private and secure 
communications over unsecured networks. In other 
words, these private networks are part of the mesh 
along with everything else, but only the participants 
%%% 239
can interact with them thanks to encryption, user authentication
and message integrity checking.

With the factionalization of transhumanity, attempts
to unify software into standard formats have
still failed. However, different operating systems 
or protocols are rarely an obstacle anymore due to 
easily accessible conversion tools and AI-aided compatibility
oversight.

\subsection{Meshing Technologies}

Almost all biomorphs in the solar system are equipped 
with basic mesh inserts (p. 300)—implanted personal 
computers. These implants are grown in the brain via 
non-intrusive nanosurgery. The processor, wireless 
transceiver, storage devices, and other components 
are directly wired to the user's cerebral neuronal cells 
and cortical centers responsible for language, speech, 
and visual perception among others. Thought-to-communication
emulations (so called \textit{transducing})
enables the user to control the implant just by thinking
and to communicate without vocalizing. Input
from the mesh inserts is transmitted directly into the 
brain and sometimes perceived as augmented reality, 
overlaid on the user's physical senses. In a similar 
vein, the mesh inserts installed in synthmorphs and 
pods are directly integrated with their cyberbrains 
(creating a potential security concern as cyberbrains 
are vulnerable to hacking).

External devices called ectos (p. 325) are also 
used to access the mesh, though these are growing 
increasingly rare given the prevalence of mesh inserts. 
Ecto interface options include haptic interfaces like 
touch-display controls, bracelets or gloves that detect 
arm, hand, or finger movements (virtual mouse and 
keyboards), eye tracking and blink control, body 
scanning grids (body axis control or all-limb controls 
for non-humanoids), voice controls, and more. Sensory
information is handled via lenses, glasses, earplugs
(subdermal bone-vibrating speakers), bodysuits,
gloves, nose plugs, tongue dams, and other devices 
that are wirelessly linked to (or physically plugged 
into) the ecto.

\subsection{Information Overload}

The mesh contains massive amounts of personal and 
public information shared by users, a digital commons 
%%% 240
of news, media, discourse, knowledge, environmental 
data, business, and culture. Transhumans embrace 
the mesh as a tool for exchange, communication, 
and participation with other users, both local and far 
away. As such, the mesh is an up-to-date, authoritative 
source on all transhuman knowledge and activities.

Not everything online is available for free, of course, 
except perhaps in the autonomist zones. Quite a bit of 
proprietary data is kept off the grid in secure storage 
or sequestered away in private networks. Some of this 
is for sale, and heavily encumbered with digital restrictions—software
media, nanofabrication blueprints,
skillsofts, etc. A thriving open source movement offers 
free and open source alternatives to much proprietary 
data, however, and numerous digital piracy groups 
deal out cracked versions of proprietary material, 
despite pressure from some authorities. Other data is 
simply secured from competitive interests (hypercorp 
research projects) or is an extremely private affair, 
such as ego backups.

\subsubsection{Spimes}

Along with the accumulated data of transhuman 
affairs, the mesh is also cluttered with information 
derived from untold numbers of wireless-capable sensor-enabled
devices that continuously update the mesh
with their location, sensor recordings, and other data. 
Colloquially called ``spimes,'' these location-aware, 
environment-aware, self-logging, self-documenting 
objects broadcast their data to anyone who cares to 
listen. Since visual, auditory, and other sensors are 
absurdly tiny and inexpensive, they are ubiquitously 
incorporated into nearly every object or product a 
person might wear, apply, use, or internalize. This 
allows almost any user to reach out through the mesh 
and gather environmental data and ambient sensor 
recordings from a specific location (or at least public 
locales—private areas typically block such signals or 
slave them to a local AI that filters their output).

\subsubsection{Surveillance, Privacy, And Sousveillance}

While spimes are easily trackable, they also contribute
to an environment of universal surveillance.
Between spimes, microsensors, ubiquitous security 
systems, and the recording capabilities of the mesh 
inserts used by almost everyone, just about everything
is recorded. Factor in the availability of mesh-tracking
facial recognition, rep/social networks, and
other data mining software, and it rapidly becomes 
clear that privacy is an outdated concept. Special 
considerations must be embraced by anyone that 
seeks to mask their identity or cover their movements. 
Alternately, off-the-shelf looks common with some 
morphs (especially synthmorphs and pods) allow a 
user to blend in with the masses.

Though in part this may seem an Orwellian surveillance
nightmare, the abundance of recording
tech actually works as ``sousveillance'' (watching 
from below), serving a role in making everything 
transparent and putting checks on abuses of power. 
%%% 241
Authoritarian regimes tread carefully, as they are also 
universally monitored, despite their attempts to control
information flow. Many people also willing participate
in this open ``participatory panopticon.'' With
nearly infinite storage capacity, dedicated lifeloggers 
record every moment of their lives and share it for 
others to experience.

\section{Interfacing: Ar, Vr, And Xp}

Mesh media is accessed using one of three protocols: 
augmented reality (AR), virtual reality (VR), or experience
playback (XP).

\subsection{Augmented Reality}

Most users perceive data from the mesh as augmented
reality—information overlaid on the user's
physical senses. For example, computer-generated 
graphics will appear as translucent images, icons, 
or text in the user's field of vision. While visual AR 
data—called  \textit{entoptic} data—is the most common, 
other senses may also be used. AR input includes 
acoustic sounds and voices, odors, tastes, and 
even tactile sensations. This sensory data is high-resolution
and seemingly ``real,'' though it is usually
presented as something ghostly or otherwise
artificial so as not to be confused with real-world 
interactions (and also to meet safety regulations).

User interfaces are customized to the user's preferences
and needs, both graphically and content-wise.
Filters allow users to access the information they 
are interested in without needing to worry about 
extraneous data. While AR data is typically placed 
in the user's normal field of vision, entoptics are not 
actually limited by this and may be viewed in the 
``mind's eye.'' Nevertheless, icons, windows and other 
interaction prompts can be layered, stacked, toggled, 
hidden, or shifted out of the way if necessary to interact
with the physical world.

\subsubsection{Avatars}

Every mesh represents themselves online via a digital
avatar. Many people use digital representations
of themselves, whereas other prefer more iconic 
designs. This may be an off-the-shelf look or a 
customized icon. Libraries of avatars may also be 
employed, enabling a user to switch their representation
according to mood. Avatars are what other
users see when they deal with you online—i.e., how 
you are represented in AR. Most avatars are animated
and programmed to reflect the user's actual
mood and speech, so that the avatar seems to speak 
and have emotions.

\subsubsection{E-Tags}

Entoptic tags are a way for people to ``tag'' a physical 
person, place, or object with a piece of virtual data. 
These e-tags are stored in networks local to the tagged 
item, and move with the item if it changes location. 
%%% 242
E-tags are viewable in AR, and can hold almost any 
type of data, though short notes and pictures are the 
most common. E-tags are often linked to particular 
social networks or circles within that network, so that 
people can leave notes, reviews, memorabilia media, 
and similar things for friends and colleagues.

\subsubsection{Skinning}

Since reality can be overlaid with entoptics of hyper-real 
quality, modern users can ``skin'' their reality by modifying
their perceptual input. Environments around them
may be modified to fit their particular tastes or mood. 
Need your spirits boosted? Pull up a skin that makes it 
seems like you're outdoors, with the sun shining down, 
the sounds of gentle surf in the background, and butterflies
drifting lazily overhead. Pissed off? Be comforted as
flames engulf the walls and thunder grumbles ominously 
in the distance. It is not uncommon for people to go 
about their day, accompanied by their own personal 
soundtrack that only they can hear. Even olfactory and 
taste receptors can be artificially stimulated to experience
sensations like the smell of roses, fresh air, or freshly-baked
pastries. While originally developed to make
``space food'' less distasteful and as a method to counter 
space-induced cabin fever for those that weren't born in 
space, vast archives of aromas, tastes, and environments 
are available for download.

Skins do not need to be kept private, they may 
also be shared with others via the mesh. Tired of 
your cramped habitat cubicle? Decorate it with a 

\subsection{Æther Jabber}

custom-themed skin and share it with visitors to make 
them feel more comfortable. Found a new music track 
that livens up your day? Share it with others around 
you, so they can nod to the same beat.

Skinning can also be used for the opposite effect. 
Any undesired content of reality can be edited out, 
veiled, or censored by modern software programs or 
muses that engage in real-time editing. Tired of looking
at someone's face? Add them to your killfile, and
you'll never have to acknowledge their presence again. 
AR censorware is also common in some communities 
with strict religious or moral convictions.

\subsection{Virtual Reality}

Virtual reality overrides the user's physical senses and 
places them inside an entirely computer-generated 
environment called a \textit{simulspace.} While AR is used 
for all common day activities and interactions, VR is 
used mainly for recreation (gaming, virtual tourism, 
escapism), socializing, meeting (when face-to-face 
meetings are not possible), and training. Dedicated 
networks with high-capacity information processing 
are required to render and run large and complex 
hyper-real simulspaces with many users, and these 
are often hard-wired for additional stability. Smaller 
simulspaces capable of hosting a smaller amount of 
users can be run on a smaller distributed network of 
linked devices. Many infomorphs and AIs effectively 
reside within simulspaces, and some transhumans 
have sworn off the physical world altogether.
%%% 243

\subsubsection{Defying Nature'S Laws}

A plethora of simulspace environments are available, 
ranging from simulations of real places to historical 
recreations to fantastic worlds representing almost 
every genre imaginable. All of these simulations are 
bolstered by the fact that possible scenarios are not 
bound by the laws of nature. The fundamental forces 
of reality and nature, like gravitation, electromagnetism
atmosphere, temperature, etc., are programmable
in VR, allowing for environments that are
completely unnatural, such as escheresque simulspace 
where gravity is relative to position. These \textit{domain }
\textit{rules} may be altered and manipulated according to 
the whim of the designer.

Time itself is an adjustable constant in VR, though 
deviation from true time has its limits. So far, transhuman
designers have achieved time dilation up to
60 times faster or slower than real time (roughly one 
minute equaling either one hour or one second). Time 
slowdown is far more commonly used, granting more 
time for simulspace recreational activities (more time, 
more fun!), learning, or work (economically effective). 
Time acceleration, on the other hand, is extremely 
useful for making long distance travel through space 
more tolerable.

\subsubsection{Accessing Simulspaces}

Most simulspaces can be accessed through the mesh 
just like any other node. Since VR takes over the 
user's sensorium, however, and sometimes involves 
time perception dilation, users are cut off from other 
mesh-delivered sensory input and interacting directly 
with other nodes. Instead, outside mesh interactions 
are routed through the simulspace's interface meaning
that a character may browse the mesh, communicate
with others, etc. from inside a simulspace, if the
domain rules allow it).

Since physical senses are overridden when a user 
accesses VR, most people prefer to rest their body 
in a safe and comfortable environment while in the 
simulspace. Body-fitting cushions and couches help 
users relax and keep them from cramping up or injuring
themselves if they happen to thrash around.
In case of long-term virtual sojourns (for instance, 
during space travel), morphs are normally retained 
in tanks that sustain them in terms of nutrition and 
oxygen. Many VR entertainment and game networks 
offer dedicated and hardwired physical VR cafes 
with private pods. Visitors rent a pod and physically 
jack in, using either access jacks or an ultrasonic 
trode net that reads and transmits brain patterns 
when placed on the head.

When accessing a simulspace, the user first enters 
an electronic buffer ``holding space'' known as a 
white room. Here the user chooses a customizable 
avatar-like persona to represent them in the simul-space
called a simulmorph. From this point, the user
immerses themself in the virtual reality environment, 
effectively becoming their simulmorph.

\subsection{Experience Playback}

Every morph with mesh inserts has the capability 
to transmit or record their experiences, a form of 
technology called experience playback, or XP. Since 
the first programs were developed that provide a 
simple interface to ``snapshot'' one's experiences, 
it has become extremely popular to share XP with 
friends and social networks, or with the online public 
at large.
The level of experiences depends on how much of 
the recorded sensory perception is kept when the clip 
is made. Full XP includes exteroceptive, interoceptive, 
and emotive tracks. Exteroceptive tracks include the 
traditional senses of sight, smell, hearing, touch, and 
taste that process the outside world. Interoceptive 
tracks include senses originating within the body, such 
as balance, a sense of motion, pain, hunger and thirst, 
and a general sense of the location of one's own body 
parts. Emotive tracks include the whole spectrum of 
emotions which can be aroused in a transhumans. 
Due to the biological requirements (neuronal and 
endocrine systems) of expressing emotions, hardcore 
XP aficionados deem only the experience in and from 
biomorphs as the real deal.

\section{Mesh Uses}

There are many reasons people use the mesh. The 
foremost is communication: voice and video calls 
(typically displaying avatars rather than actual 
video), electronic messaging (e-mail, instant messaging
microblogging), and file and data transfers.
Socializing is also key, handled via social and reputation
networks, personal profiles, lifelogging, chats
and conferences (both AR and VR), and discussion 
groups and forums. Information gathering is also 
at the top, whether its browsing the popular Solarchive
or other databases and directories, tapping
the latest newsfeeds, browsing mesh sites, tracking 
your friends, taking lessons in VR, or looking up just 
about anything conceivable. Recreation rounds out 
the pack, covering everything from gaming (AR and 
VR) to experiencing other people's lives (XP) to VR 
tourism and club-hopping.

\subsection{Personal Area Networks}

Since everything a person carries is meshed, most 
people maintain personal area networks that route 
all of these devices through their mesh inserts or ecto, 
which acts as a hub. This is both a security measure, 
ensuring they maintain control over their own accessories
and a convenience factor, as it focuses all of the
controls in one place.

\subsection{Virtual Private Networks}

Virtual private networks (VPNs) are communications
networks tunneled through the mesh, which
are dedicated for a specific group of people. The 
primary use of VPNs is to create privacy and security
for its users, and so they typically use security

cto,

re, 

es-

e
%%% 244
features such as ego authentication and public key 
encryption. VPNs are regularly used to mesh mobile 
offices into a corporate network or mesh people together
who work on or contribute a certain project.
Other VPNs—particularly social networks and rep 
networks—operate with minimal security features, 
simply serving as a network of specific users within 
the mesh and making it easier to keep in touch, 
transfer information, make updates, and so on. Most 
VPNs come as specialized software suites that run 
custom environmental software that integrates into 
the user's normal mesh interface and AR.

\subsection{Social Networks}

Social networks are the fabric of the mesh, weaving 
people together. They are the means by which most 
people keep in contact with their friends, colleagues, 
and allies, as well as current events, the latest trends, 
new memes, and other developments in shared 
interests. They are an exceptionally useful tool for 
online research, getting favors, and meeting new 
people. In some cases, they are useful for reaching 
or mobilizing masses of people (as often illustrated 
by anarchists and pranksters). There are thousands 
of social networks, each serving different cultural 
and professional interests and niches. Most social 
networks allow users to feature a public profile to 
the entire mesh and a private profile that only those 
close to them can access.

Reputation plays a vital part in social networks, 
serving as a measure of each person's social capital. 
Each person's reputation score is available for lookup, 
along with any commentary posted by people who 
favored or disfavored them and rebuttals by the user. 
Many people automate their reputation interactions, 
instructing their muse to automatically ping someone
with a good review after a positive action and
to likewise provide negative feedback to people with 
whom the interaction went poorly.

\subsection{Mobile Offices}

Due to the lack of office space and the wireless accessibility
of most information, most businesses
now operate virtually, with few or no fixed offices 
or even assets. Instead, individuals have become their 
own mobile office. Bit-pushers and bureaucrats like 
hypercorp executives, clerical workers, accountants, 
and researchers—as well as innovators like artists, 
writers, engineers, and designers—work wherever 
they want to.

The most prominent example of this phenomenon 
are the bankers of the Solaris hypercorp. Each employee
acts as a mobile one-person banking office,
managing transactions via Solaris's robust VPN.

 On rare occasions, office environments are run 
in simulspace with time dilation to maximize efficiency
Since this requires the workers to access a
centralized wired network and leave their bodies 
unattended while accessing simulspace, however, it 
requires an extra level of physical security that is 
typical only of some governmental installations and 
corporate habitats.

\section{Islands In The Net}

In the time of \textit{Eclipse Phase,} information can become 
outdated quite fast, and the accessibility of new information
depends on your location. It's easy to keep up-to-date
on your local habitat/city or planetary body,
but keeping current on events elsewhere is typically 
reliant on the speed of light.

If you happen to be in a station in the Kuiper Belt, 
on the edge of the solar system 50 astronomical units 
from the terrestrial inner planets, waiting on a message
from Mars, the signal carrying the message will
be roughly seven hours old when it reaches you. Of 
course it will only reach you that fast if you are using 
quantum farcast, which is only limited by the speed 
of light (not to mention rare and expensive in most 
habitats). If you are not using a quantum farcaster, 
the signal may take even longer and is prone to interference
and noise, deteriorating the quality and
possibly losing some of the content, especially over 
major distances. Whenever you start dealing with 
communication between habitats, you have to factor 
in the light-speed lag, the amount of time it takes 
even the fastest transmission to reach you. This lag 
works both ways, so trying to hold a conversation 
with someone just 5 light-seconds away means that 
you're waiting at least 10 seconds to get the reply 
to whatever you just said. For this reason, AR and 
VR communications are almost always conducted locally
while standard messaging is used for nonlocal
communications. For detailed discussions, it is often 
simpler to send a fork of yourself (p. 273) to have the 
conversation and then return.

Quantum-entanglement communicators (p. 314) 
are one solution to this light-speed lag, although a 
burdensome and expensive one. QE comms allow 
for faster-than-light communication to an entangled 
communicator, though each transmission uses up a 
precious amount of quantum-entangled bits, which 
are in limited supply.

Transmissions made between habitats almost 
always occur via each station's massive data relays, 
where they are then distributed into the local mesh. 
This bottleneck is often used by authoritarian habitats 
to monitor data transmissions and even filter or censor 
certain public non-encrypted content. Some messages 
are also prioritized over others, potentially meaning 
further delays.
%%% 245

The method of transmission between habitats also 
sometimes matters. Radio and neutrino broadcasts 
can be intercepted by anyone, whereas tight-beam 
laser or microwave links are specifically used as a 
point-to-point method that minimizes interception 
and eavesdropping. The use of quantum farcasting 
using neutrino systems is completely secure, however, 
and is the most frequently-used intra-habitat link.

What these lags, bottlenecks, and prioritizations 
mean is that some news and data takes a particularly 
long-time to trickle from one local mesh network to 
another, passing slowly from habitat to habitat. This 
means that there are always gradients of information 
available to different local mesh networks, typically 
depending on proximity and the importance of the 
information. Some data even gets lost along the way, 
never making it further than a habitat or two before 
it is lost in the noise. The only way to retrieve such 
information is to track it down to its source.

\subsection{Darkcasts}

``Darkcasts'' are ranged communications that go outside
of legal and approved channels. Since certain habitats
have strict regulations on transmission content,
forking, egocasting, infomorphs, muse abilities, and 
AGI code, underworld groups like the ID crew profit 
by offering illegal data transmission services. Primarily 
used for censored data and banned content (like illegal 
XPs or malware), local organized crime factions also 
often offer egocasting services complete with resleeving
and leasable morphs, allowing egos that prefer
discretion to enter or leave a habitat without drawing 
attention. Though such authorities hunt down these 
darkcast networks whenever they get a chance, many 
habitats have a sophisticated darkcast infrastructure 
that makes use of decoys, temporary communications 
lines, relays, and regular transmitter relocation—not 
to mention judicious bribing and blackmailing.

\section{Mesh Abuses}

As with all things, the mesh has its darker side. At 
the basic level, this amounts to flamewar-starting 
trolls, stalkers, or griefers whose intent is to mess 
with others for a laugh. At the more organized level, 
it expands to illicit or criminal enterprises that utilize 
the mesh, such as selling black/snuff/porn XPs, illegal
software, pirated media, or even egos. The most
infamous threats—thanks both to the Fall and to 
the continuous sensationalism applied by media and 
stern authorities—are, of course, malware and hackers
Given the capabilities of modern hackers and the
vulnerability of many habitats—where damage to life 
support systems could kill thousands—the threat may 
not be over-exaggerated.

\subsection{Hackers}

Whether individuals who are genuinely interested 
in exploring new technologies and seeking ways to 
break them in order to make them better, hacktivists
who utilize the mesh in order to undermine the
power of authorities, or ``black hats'' who seek to 
circumvent network security for malicious or criminal
intent, hackers are a permanent fixture of the
mesh. Unauthorized network break-ins, infiltration 
of VPNs, muse subversion, cyberbrain hijacking, data 
theft, cyber-extortion, identity fraud, denial of service 
attacks, electronic warfare, spime hijacking, entoptic 
vandalism—these are all common occurrences on the 
mesh. Thanks to smart and adaptive exploit programs 
and assisting muses, even a moderately skilled hacker 
can be a threat.

In order to counter hacking attempts, most people, 
devices, and networks are protected by a mix of 
access control routines, automated software intrusion
prevention systems, encryption, and layered
firewalls, typically overseen by the user's muse who 
plays the role of active defender. Extremely sensitive 
%%% 246
systems—such as space traffic control, life support, 
power systems, and hypercorporate research facilities—are
usually limited to isolated, tightly-controlled,
heavily-monitored, hard-wired networks to minimize 
the risk of intrusion from snoopers and saboteurs. 
Various countermeasures may be applied against 
such intruders, ranging from locking them out of the 
system to tracking them back and counterhacking.

\subsection{Malware}

The number of worms, virii, and other malware programs
that ripped through computer systems during
the Fall was staggering. Many of these were part of 
the netwar systems prepared by old Earth nation-states 
and corporations and unleashed on their enemies. 
Others were products of the TITANs, subversive programs
that even the best defenses had trouble stopping.
Even 10 years later, many of these are still reappearing, 
brought back to life by the accessing of some long-forgotten
data cache or the accidental infection of a
scavenger mucking through old ruins. New ones pop 
every day, of course, many of them programmed by 
criminal hacker outfits, while others that enter circulation
are modifications and variations of suspected
TITAN designs, perhaps implying that certain parties 
are intentionally tinkering with this code and releasing 
it into the wild. Rumors and whispers circulate that 
some of these TITAN worms are even more potent and 
frightening than previously hinted at, with stunning 
adaptive capabilities and intelligence. These rumors 
are steadfastly denied by authority figures and security 
experts ... who then quietly turn around and do their 
best to ensure that their own networks remain safe.

\section{Ai}


\section{ And Infolife}

Self-aware helper programs were originally designed 
and realized to augment transhuman cognitive abilities
These specialized-focus AIs were then developed
into the more complete, independent digital consciousnesses
known as AGIs. The further evolution of these
digital life forms into seed AIs unfortunately led to the 
emergence of the TITANs and then the Fall. This created
a rift in transhuman society as fear and prejudice
turned popular opinion against unrestricted AGIs, an 
attitude of mistrust that still lingers to this day.

\subsection{Ai}

The term AI is used to refer to narrow, limited-focus 
AIs. These digital minds are expert programs with processing
capabilities equal to or even exceeding that of
a transhuman mind. Though they have a personality 
matrix with individual identities and character, and 
though they are (usually) conscious and self-aware, 
their overall complexity and capabilities are limited. 
The programmed skills and abilities of AIs are typically
very specific in scope and aligned towards a particular
function, such as piloting a vehicle, facilitating
mesh searches, or coordinating the functions of some 
habitat sub-system. Some AIs, in fact, can only barely 
be considered sapient, and their emotional programming
is usually narrow or non-existent.

AIs have a number of built-in safety features and 
programmed limitations. They must serve and obey the 
instructions of authorized users within their normal 
functioning parameters and (in the inner system at 
least) must also obey the law. They lack self-interest 
and self-initiative, though they have limited empathy 
and may be programmed to anticipate the needs and 
desires of users and pre-emptively take action on 
their behalf. Perhaps most importantly, however, is 
that their psychological programming is specifically 
based on universal human modes of thought and an 
understanding and support of transhuman goals 
and interests. This is part of an initiative to engineer 
so-called ``friendly AIs,'' who are programmed with 
sympathy towards transhumanity and all life and seek 
out their best interests.

In most societies, basic AIs are considered ``things'' 
or property rather than people and are accorded no 
special rights.

\subsection{Muses}

Muses are a specific type of AI designed to function 
as a personal aide and companion. Most people in 
\textit{Eclipse Phase} have grown up with a muse at their 
virtual side. Muses tend to have a bit more personality 
and psychological programming than standard AIs and 
over time they build up an extensive database of their 
user's preferences, likes and dislikes, and personality 
quirks so that they may more effectively be of service 
and anticipate needs. Muses generally have names 
and reside within the character's mesh inserts or ecto, 
where they can manage the character's personal area 
network, communications, data queries, and so on.

\subsection{Agi}

AGIs are complete and fully operational digital consciousnesses
self-aware and capable of intelligent
action at the same level as any transhuman. Most have 
full autonomy and the capacity for self-improvement 
by a processing similar to learning—a slow optimization
and expansion of their code that features
programmed limitations to prevent it from achieving
the self-upgrading capabilities of seed AIs. They
have more fully-rounded personalities and emotional/
empathic abilities than standard AIs, due in part to 
a development process where they are literally raised 
within a VR simulation analogous to the rearing of 
transhuman children, and so are more fully socialized. 
As a result, they have a fairly human outlook, though 
some deviation is to be expected, and sometimes is 
apparent in great degrees. Despite this attempt to humanize
AGIs, they do not have the same evolutionary
and biological origins that transhumans have, and so 
their social responses, behavior, and goals are sometimes
off-mark or decidedly different.

AGIs bear the social stigma of their non-biological 
origin and are often met with bias and mistrust. Some 
habitats have even outlawed AGIs or subject them to 
%%% 247
strict restrictions, forcing such infolife to hide their 
true natures or illegally darkcast to enter habitats or 
stations. AGI mind programming emulates transhuman
brain patterns sufficiently well that they can be
sleeved into biomorphs if they choose.

\subsection{Seed Ai}

Due to the capability for unlimited self-upgrading, 
seed AIs have the capacity to grow into god-like 
digital entities far beyond the level of transhumans 
or AGIs. They require massive processing power and 
are always increasing in complexity due to a continual 
metamorphosis of their code. Seed AIs are too complex
to be downloaded into a physical morph, even
a synthetic one. Even their forks require impressive 
processing environments, so doing so is rare. In fact, 
most seed AIs require the capacities of hard-wired 
networks to survive.

The only seed AIs known to the public are the 
infamous TITANs who are widely regarded as being 
responsible for the Fall. In truth, the TITANs were 
not the first seed AIs and will probably not be the last. 
There are no publicly known TITANs (or other seed 
AIs) currently residing in the solar system, despite 
circulating rumors of damaged TITANs who were left 
behind on Earth, speculated TITAN activity under the 
clouds of Venus, or whispers of new seed AIs hidden 
away in secret networks on the edges of the system.

\subsection{Transhuman Infomorphs}

For thousands of infugees, embodying a digital form 
is their only choice. Some of these are locked away 
in mesh-separated virtual holding areas or even inactive
storage, locked up by habitats who didn't have
enough resources to handle them. Others are imprisoned
inside simulspaces, killing time in whatever way
they choose until an opportunity to resleeve comes 
their way. Quite a few are free to roam the mesh, interacting
with physically-sleeved transhumans, keeping
up with current events, and sometimes even forming
activist political blocs to campaign for infomorph
rights or interests. Still others find or are forced into 
virtual careers, slaving away in the digital sweatshops 
of hypercorps or criminal syndicates. A few find companions
who are willing to bring them along in their
ghostrider module and become an integral part of 
their lives, much like a muse.
Some transhumans willingly choose the infomorph 
lifestyle, either for hedonism (custom simulspace and 
VR games until the end of time), escapism (loss of a 
loved ones leads them to write off physical concerns 
for awhile), freedom (go anywhere the mesh takes 
you—some have even beamed copies of themselves to 
far distant solar systems, hoping someone or something 
will receive their signal when they arrive), experimentation
(forking and merging, running simulations, and
weirder things), or because it is ensured immortality.

\section{Everyday Mesh Mechanics}

Everyone (and everything) is meshed in \textit{Eclipse Phase. }
The following rules and concerns apply to standard 
mesh use. Note that various mesh-related terms are 
explained, along with other \textit{Eclipse Phase }concepts, 
under \textit{Terminology,} p. 25.

\subsection{Mesh Interface}

Characters have a choice of which interface to use, the 
entoptic interface of basic mesh inserts or the haptic 
interface of an ecto.
The basic mesh inserts used by most users allows 
them to interact with AR, VR, XP, and the mesh at the 

\textit{ase.}

rd 

re 

s, 
%%% 248
speed of thought. This is the default method of mesh 
use and suffers no modifiers. They are, however, more 
prone to visual and operative impairments (virtual illusions
denial-of-service effects) when hacked.

Characters who use the haptic interface of an ecto, 
however, suffer a slight delay on their mesh activities 
due to manual toggling, physical controls, and physical
interaction with virtual controls. In game terms,
the use of haptics imposes a –10 skill modifier  to 
all mesh tests where timing is rushed (particularly 
combat and or any sort of mesh use under pressure). 
Additionally, increase the timeframe for mesh-based 
Task Actions by +25% when interfacing via haptics. 
On the positive side, ectos can be easily removed and 
discarded if compromised—for this reason, many 
hackers and security-conscious users use an ecto in 
addition to their mesh inserts, routing all high-risk 
traffic through the ecto as an extra line of defense.

\subsection{Mesh Id}

Every mesh user (and, in fact, every device) has a 
unique code called their \textit{mesh ID.} This ID distinguishes
them from all other users and devices, and is
the mechanism by which others can find them online, 
like a combination phone number, email address, and 
screen name. Mesh IDs are used in almost all online interactions
which are often logged, meaning that your
activities online leave a datatrail that can be tracked 
(p. 251). Fortunately for Firewall sentinels and others 
who value their privacy, there are ways around this 
(see \textit{Privacy and Anonymity,} p. 252). AIs, AGIs, and 
infomorphs also each have their own unique mesh ID.

\subsection{Accounts And Access Privileges}

Devices, networks (such as PANs, VPNs, and hardwired
networks), and services require that every user
that accesses them does so with an account. The 
account serves to identify that particular user, is 
linked to their mesh ID, and determines what access 
privileges they have on that system. There are four 
types of accounts: \textit{public, user, security,} and \textit{admin.}

\subsubsection{Public Accounts}

Public accounts are used for systems that allow 
access (or access to parts of their system) to anyone 
%%% 249
on the mesh. Public accounts do not require any sort 
of authentication or login process, the user's mesh ID 
is enough. These accounts are used to provide access 
to any sort of data that is considered public: mesh 
sites, forums, public archives, open databases, social 
network profiles, etc. Public accounts usually have 
the ability to read and download data, and sometimes
to write data (forum comments, for example),
but little else.

\subsubsection{User Accounts}

User accounts are the most common accounts. User 
accounts require some form of authentication (p. 253) 
to access the device, network, or service. Each user account
has specific access privileges allotted to it, which
are tasks the user is allowed to perform on that system. 
For example, most users are allowed to upload and 
download data, change basic content, and use the 
standard features of the system in question. They are 
not, however, usually allowed to alter security features, 
add new accounts, or do anything that might impact 
the security or functioning of the system. As some systems
are more restrictive than others, the gamemaster
decides what privileges each user account provides.

\subsubsection{Security Accounts}

Security accounts are intended for users that need 
greater rights and privileges than standard users, but 
who don't need control over the entire system, such 
as security hackers and muses. Security access rights 
usually allow for reading logs, commanding security 
features, adding/deleting accounts, altering the data of 
other users, and so on.

\subsubsection{Admin Accounts}

Admin accounts provide complete control over the 
system. Characters with admin rights can do everything
security accounts can, plus they can shut down/
reboot the system, alter access rights of other users, 
view and edit all log files and statistics, and stop or 
start any software available on the system.

\subsection{Mesh Gear Quality}

Not all gear is created equal, and this is especially true 
of computers and software, where new innovations 
are made on a daily basis. Keeping up-to-date with 
the last specs isn't too difficult, but on occasion the 
–30

Bashed-up devices, no-longer-supported software, relics from Earth or the early expansion into space
–20

Malfunctioning/inferior devices, buggy software, pre-Fall technology
–10

Outdated and low quality systems

0

Standard ectos, mesh inserts, and software
+10

High-quality goods, standard security-grade products
+20

Next-generation devices, advanced software
+30

Newly-developed, state-of-the-art, top-of-the-line technology
>+30

TITANs and/or alien technology
characters will get their hands on some old relic or 
find themselves in secluded or decrepit places with 
local systems and gear that aren't up to date. Likewise, 
they may acquire some cutting-edge gear straight 
from the labs or may run up against a state-of-the-art 
installation with next-generation defenses. To reflect 
this, mesh tests can be modified according to the state 
of the hardware or software being used, as noted on 
the Mesh Gear Modifiers table
.

\subsection{Computer Capabilities}

Computerized electronics can be broken down into 
three simple categories: \textit{peripherals, personal comput-}
\textit{ers,  }and  \textit{servers.} In game terms, all are collectively 
referred to as \textit{devices.}

\subsubsection{Peripherals}

Peripherals are micro-computerized devices that don't 
need the full processing power and storage capacity of 
a personal computer, but benefit from online networking
and other computerized functions. Peripherals may
run software, but the gamemaster may decide that too 
many programs (10+) will degrade the system's performance
AIs and infomorphs are incapable of running
on peripherals, though they may access them. Peripherals
only have public and user accounts (users accounts
include security and admin functions; see p. 247).

Common peripherals include: spimes, appliances, most 
cybernetic implants, individual sensors, and weapons.

\subsubsection{Personal Computers}

Personal computers account for a wide range of computer
types, but essentially account for anything that
has the capabilities evolved from generations and generations
of personal computers to meet an everyday
user's needs. Most personal computers are portable 
and tailored for use by multiple users at a time. Personal
computers may run one AI or infomorph at a
time. They may not run simulspace programs.

Common personal computers include: mesh inserts, 
ectos, and vehicles.

\subsubsection{Servers}

Servers have much greater processing power and 
data management capabilities than personal computers
They are capable of handling hundreds of users,
multiple AIs and infomorphs, and they may run 
%%% 250
simulspace programs. Though few are portable, some 
of them come close.

\subsection{Software}

A wide manner of software is available for mesh users, 
from firewalls and AIs to hacking and encryption tools 
to tacnets and skillsofts. These programs are listed on 
p. 331 of the \textit{Gear }chapter. Like other gear, software 
may enable a character to perform a task they could 
not otherwise. The quality of the software may also be 
a factor, applying a modifier as appropriate (see \textit{Mesh }
\textit{Gear Quality,} p. 247).

Some software is equipped with digital restrictions 
to prevent it from being copied and shared with others. 
These restrictions may be defeated, but it is a time-consuming
task, requiring a Task Action Programming Test
with a timeframe of 2 months. Thanks to the efforts 
of the open source movement and numerous individual 
software pirates, however, quite a bit of software is 
available free or online. The availability of pirated software
or freeware will depend on the local habitat and
legalities. Finding it may be a matter of a simple search, 
or it may require some use of reputation to find someone
who has it. Usually there is at least one local crime
syndicate that is willing to help you out—for a price.

\subsubsection{Software Compatibility}

In most instances, software compatibility is not going 
to be an issue for characters. Gamemasters who like 
it as a plot device, however, can introduce compatibility
problems in certain cases, whether this is done
to increase drama, slow the characters down, or create 
obstacles that they must overcome. Such incompatibilities
are more likely to arise when dealing with
outdated systems or devices, or at least ones that are 
unlikely to have the latest patches and software updates
Incompatibilities can also be used as a drawback
to acquiring software from untrustworthy sources.

Conflicting software issues are going to have one of 
two effects. Either the software will simply not work 
with certain devices, or it will inflict a –10 to –30 
modifier due to instabilities and glitches. If the gamemaster
allows it, a character may reduce this penalty
by patching up the software, requiring a successful 
Programming Task Action (1 day). For every 10 points 
of MoS, reduce the incompatibility modifier by 10.

\subsection{Traffic Filters And Mist}

Mesh networks and AR are overrun with yottabytes of 
information. Though mesh inserts and ectos can deal 
with a lot of data traffic in terms of bandwidth and 
processing power, using filters to weed out unwanted 
traffic is quite simply a necessity. This is especially true 
of AR, where unwanted entoptics can clutter your 
vision and distract you. Nevertheless, entoptic spam of 
many flavors—advertisements, political screeds, porn, 
scams—do their best to bypass these filters, and in 
many areas the amount of unfiltered entoptics can be 
overwhelming—a phenomenon colloquially referred 
to as ``mist.''

At the gamemaster's discretion, mist can interfere 
with a user's sensory perceptions. This modifier can 
range from –10 to –30, and in some cases might be 
so distracting as to affect all of a character's actions. 
To lift the data fog, a character or muse must adjust 
their filter settings by succeeding in an Interfacing Test 
modified by the mist modifier. Alternately, the character
can completely disable AR input, but this is likely
to impede them in other ways.

\subsection{Slaving Devices}


For ease of use, as well as for privacy and security 
purposes, one or more devices may be slaved to each 
other. One device (usually the character's mesh inserts 
or ecto) takes the role of \textit{master,} while the other 
device(s) takes the role of \textit{slave.} All traffic to and 
from slaved devices is routed through the master. This 
allows a slaved device to rely on the master's security 
features and active monitoring. Anyone that wants to 
connect to or hacked into a slaved device is rerouted 
to the master for authentication and security scrutiny.
Slaved devices automatically accept commands from 
their master device. This means that a hacker who 
penetrates a master system can legitimately access 
and issue commands to a slaved device, assuming their 
access privileges allow for it.

PANs are typically formed by slaving all of a character's
devices to their ecto or mesh inserts. Similarly,
individual components of a security system (doors, 
sensors, etc.) are usually slaved to a central security 
node that serves as a chokepoint for anyone hoping 
to hack the system. The same is often true for other 
networks and facilities.

\subsection{Issuing Commands}

Characters may issue commands to any slaved device 
or teleoperated bot (see \textit{Shell Remote Control,} p. 196) 
with a Quick Action. Each command counts separately, 
unless the character is issuing the same command to 
multiple devices/drones.

\subsection{Distance Lag}

Anytime you extend your communications over great 
distances, you run into the risk of time lags. Most 
communications are restricted to ``local'' for this 
reason, which generally means your local habitat 
and any others within 50,000 kilometers. On planetary
bodies like Mars, Venus, Luna, or Titan, ``local''
encompasses all of the habitats and linked mesh networks
on that planetary body.

If a character is searching the mesh beyond their 
local area, the most efficient way is to transmit a search 
AI (usually a copy of your muse) or a fork to the non-local
area, which will then run its search and return
completed results. This process does, however, add to 
the time of transmission to the timeframe (i.e., searching
the mesh of a station 10 light-minutes away adds
20 minutes to the search as the search is transmitted 
over and the results are transmitted back). Since long-distance
communications are sometimes interfered with
%%% 251
or bumped for higher-priority items, the gamemaster 
can increase this time at their discretion. If the research 
involves correlation and fine-tuning the search parameters
based on data accumulated from different locals,
the timeframe may be exponentially increased due to 
the need for back-and-forth interaction.

If the character is simply communicating with or 
accessing non-local networks, an appropriate time lag 
must be introduced between communications and actions
The effects of this lag are largely up to the gamemaster
as fitting current distances and other factors.

\subsection{Accessing Multiple Devices}

Meshed characters may connect to and interact with 
numerous devices, networks, and services simultaneously
There is no penalty for doing this, but the
character may only focus on one system at a time. In 
other words, you may only interact with one system 
at a time, though you may also switch between them 
freely, even within the same Action Phase. You could, 
for example, spend several Quick Actions to send a 
message with your ecto, tell your spime oven at home 
to start cooking dinner, and look up a friend's updated 
profile on a social network. You may not, however, 
hack into two separate systems simultaneously.

Note that you may send the same command to 
multiple slaved devices or teleoperated drones with 
the same Quick Action, as noted above.

\section{Online Research}

The Research skill (p. 184) represents a character's 
ability to track down information in the mesh. Such 
information includes any type of digitized data: text, 
pictures, vids, XP, sensor feeds, raw data, software, etc. 
This data is culled from all manner of sources: blogs, 
archives, databases, directories, social networks, rep 
networks, online services, forums, chat rooms, torrent 
caches, and regular mesh sites of all kinds. Research 
is conducted using various public and private search 
engines, both general and specialized, as well as data 
indices and search AIs.

Research has other uses as well. Hackers use it 
when looking for specific information on a network or 
device on which they have intruded. Likewise, since 
everyone inevitably uses and interacts with the mesh, 
Research skill is also a way to identify, backtrack, and/
or gather information on people as long as they have 
not hidden their identity, worked anonymously, or 
covered their identity with a shroud of disinformation.

\subsection{Search Challenges}

Due to the sheer amount of data available, finding
what you're looking for may sometimes seem a
daunting task. Thankfully, information is fairly well 
organized, thanks to the hard work of ``spider'' AIs 
that cruise the mesh and constantly update data and 
search indices. Additionally, information on the mesh 
is tagged with semantics, meaning that it's presented 
with code that allows a machine to understand the 
\textit{context }of that information as well as a human reader 
would. This helps AIs and search software correlate 
data more efficiently. So finding the data is usually not 
as difficult as analyzing it and understanding it. Finding
specialized or hidden info or correlating data from
multiple sources is usually the real challenge.

Perhaps a larger issue is the amount of incorrect 
data and misinformation online. Some data is simply 
wrong (mistakes happen) or outdated, but the nature 
of the mesh means that such things can linger on for 
years and even propagate far and wide as they are 
circulated without fact-checking. Likewise, given 
%%% 252
the amount of transparency in modern society, some 
entities actively engage in disinformation spreading 
in order to clutter the mesh with enough falsehoods 
that the truth is hidden. Two factors help to combat 
this, the first being that data sources themselves have 
their own reputation scores, so that untrustworthy 
or disreputable sources can be identified and ranked 
lower in search results. Second, many archives take 
advantage of crowdsourcing—that is, harnessing the 
collaborative power of mesh users (and their muses) 
everywhere—to verify data integrity so that these sites 
are dynamic and self-correcting.

\subsection{Handling Searches}

Online research is often a crucial element of \textit{Eclipse }
\textit{Phase} scenarios, as characters take to the mesh to 
research backgrounds and uncover clues. The following
suggestions present a method of handling research
that does not rely solely on dice rolls and integrates it 
with the flow of the plot.

First, common and inconsequential information 
should be immediately available without requiring 
a roll at all. Most characters rely on their muses to 
handle searches for them, passing on the results while 
the character focuses on other things.

For searches that are more detailed, difficult, or 
central to the plot, a Research Test should be called 
for (made either by the character or their muse). 
This test indicates the process of finding links to 
and/or accumulating all data that may in fact be 
relevant to the search topic. This test should be 
modified as appropriate to the obscurity of the topic, 
ranging from +30 for common and public topics 
to –30 for obscure or unusual intel. This initial 
search has a timeframe of 1 minute. If successful, it 
turns up enough data to give the character a basic 
overview, perhaps with cursory details. The gamemaster
should use the MoS to determine the depth
of the data provided on this initial excursion, with 
an Excellent Success providing some bonus details. 
Similarly, a Severe Failure (MoF 30+) may result in 
the character working with data that is incorrect or 
intentionally misleading.

The next step is not so much accumulating links 
and data as it is analyzing and understanding the 
information acquired. This requires another Research
Test, again modified by the obscurity of the
topic. If the gamemaster allows it, complementary 
skills (p. 173) may apply to this test, providing bonus 
modifiers (for example, an understanding of Academics
Chemistry would help research the effects of a
strange drug). Muses may also perform this task, 
though their skills are typically inferior. As above, 
success determines the quality and depth of the 
analysis, with an Excellent Success providing the full 
story and potential related issues and a Severe Failure
marking completely incorrect assumptions. The
timeframe for this phase of research largely depends 
on two factors: the amount of data being analyzed 
and the importance to the storyline. Gamemasters 
%%% 253
need to carefully measure out their distribution of 
intel and clues to players. Give them too much too 
soon, and they may spoil the plot. Fail to give them 
enough, and they may get frustrated or pursue dead 
ends. Timing is everything.

\subsection{Real-Time Searches}

Characters may also set up ongoing mesh scans that will 
alert them if any relevant information comes up new or 
updated, or is somehow changed. This is a task usually 
assigned to muses for continuous oversight. Whenever 
such data arises, the gamemaster secretly makes a Research
Test, modified by the obscurity of the topic. If
successful, the update is noted. If not, it is missed, though 
the gamemaster may allow another test if and when the 
topic reaches a wider range of circulation or interest.

\subsection{Hidden Data}

It is important to remember that not everything can 
be found online. Some data may only be acquired (or 
may be more easily gotten) by asking the right people 
(see \textit{Networking,} p. 286). Information that is considered
private, secret, or proprietary will likely be stored
away behind VPN firewalls, in off-mesh hardwired 
networks, or in private and commercial archives. This 
would require the character to gain access to such 
networks in order to get the data they need (assuming 
they even know where to look).

It's worth noting that some entities send out AIs 
into the mesh with the intent of finding and erasing 
data they'd rather hide, even if this requires hacking 
into systems to alter such information.

\section{Scanning, Tracking, }


\section{And Monitoring}

Most users leave traces of their physical and digital 
presence all throughout the mesh. Accounts they access, 
devices with which they interact, services they use, entoptics
they perceive—all of these keep logs of the event,
and some of these records are public. Simply passing 
nearby some devices is enough to leave a trail, as near-field
radio interactions are often logged. This electronic
datatrail can be used to track a user, both to ascertain 
their physical location or to note their online activities.

\subsection{Wireless Scanning}

To interface with a wireless device or network, whether 
to establish a connection or for other purposes, the 
target device/network must be located first. To locate 
an active node, it's wireless radio transmissions must be 
detected. Most wireless devices automatically scan for 
other devices in range (see \textit{Radio and Sensor Ranges,} p. 
299) as a matter of course, so no test is required. This 
means that it's trivial for any character to pull up a 
list of the wireless devices and networks around them, 
along with associated mesh IDs. Likewise, a muse or 
device can be instructed to alert the user when a new 
signal (or a specific user) comes into range.
Detecting stealth signals (p. 252), however, is a bit 
more challenging. To detect a stealthed signal, the scanning
party must actively search for such signals, taking
a Complex Action and making an Interfacing Test with 
a –30 modifier. If successful, they detect the hidden emissions
If the character aiming for stealth engages in active
countermeasures, also requiring a Complex Action, then 
an Opposed Interfacing Test is called for (with the –30 
modifier still applying to the scanning party).
For covert devices that are only transmitting in 
short bursts, wireless detection is only possible during 
the short period the burst transmission is being made.

\subsection{Physical Tracking}

Many users willingly allow themselves to be physically 
tracked via the mesh. To them, this is a useful feature—
it allows their friends to find them, their loved ones to 
know where they are, and for authorities to come to 
their aid in the event of some emergency. Finding their 
location is simply a matter of looking them up in the 
local directory, no test required (assuming you know 
who they are). Mesh positioning is accurate to within 
5 meters. Once located, the position of the target can 
be monitored as they move as long as they maintain 
an active wireless connection to the mesh.

\subsubsection{Tracking By Mesh Id}

An unknown user's physical location can also be 
tracked via their online mesh activity—or more specifically
by their mesh ID (p. 246). Network security
will often trace intruders this way and then dispatch 
security squads to bring them in. To track an unknown 
user by their mesh ID alone requires a Research Test. 
If successful, they have been tracked to their current 
physical location (if still online) or last point of interaction
with the mesh. If the character is in privacy
mode (p. 252), a –30 modifier applies.

\subsubsection{Tracking By Biometrics}

Given the existence of so many spimes and public 
cameras and sensors, people may also be tracked by 
their facial profile alone using facial recognition software
This software scans accessible video feeds and
attempts to match it to a photo of the target. Given 
the sheer volume of cameras, however, and the typical 
range of false-positives and false-negatives, finding 
the target often boils down to luck. Priority can be 
given to cameras monitoring major thoroughfares, to 
narrow the search, but this risks missing the target if 
they avoid heavy traffic areas. The success of searches 
of this nature is best left to gamemaster fiat, but a 
Research Test can also be called for, modified appropriately
by the range of the area being watched,
whenever there is a chance the target may be spotted.
Other biometric signatures may also be used for 
tracking this way, though these are usually less available
than cameras: thermal signatures (requires infrared
cameras), walking gait, scent (requires olfactory
sensors), DNA (requires DNA scanners), etc. Each 
biometric scan requires a separate type of software.

t if 

hes

a 

p-

d, 
%%% 254

\subsection{Digital Activity Tracking}

Tracking someone's online activities (meshbrowsing, 
entoptic interactions, use of services, messaging, etc.) 
is slightly more difficult, depending on what exactly 
you're after. Gathering information on a user's public 
mesh activities—social network profiles, public 
forums posts, public lifelogging, etc.—is handled just 
like standard online research (p. 249).

\subsubsection{Tracking By Mesh Id}

A more investigate search can attempt to use the 
target's mesh ID (p. 246), using it as a sort of digital 
fingerprint to look up where else they've been online. 
This primarily involves checking access/transaction 
logs, which are not always publicly accessible. This 
sort of search requires a Research Test, handled as a 
Task Action with a timeframe of 1 hour.

\subsection{Sniffing}

Wireless radio traffic is broadcast through the air (or 
space), meaning that it can be intercepted by other wireless
devices. ``Sniffing'' involves the capture and analysis
of data traffic flowing through the wireless mesh.

To eavesdrop on wireless communications, you need 
a sniffer program (p. 331) and you must be within 
radio range (p. 299) of the target (alternately, you can 
access a device that is within radio range of the target, 
and sniff from that location). To capture the information
you must succeed in an Infosec Test. If successful,
you capture data traffic from any targeted devices in 
range. Note that sniffing does not work on encrypted 
traffic (including VPNs and anything else using public 
key cryptography) as the results are gibberish. Quantum
encrypted communications cannot be sniffed.

Once you have the data, finding the information 
you're looking for can be a challenge. Handle this as 
a standard Research Test (p. 245).

\subsubsection{Remote Sniffing Via Mesh Id}

Finally, a mesh ID may also be actively monitored 
to see what mesh activity it engages in. This requires 
special sniffer software (p. 331) and a Research Test. 
If successful, the monitoring will provide information 
on that user's public mesh activities (how much is 
determined by the gamemaster and the MoS), such 
as which sites they access, who they message, etc. It 
will not, however, uncover anything that is encrypted 
(unless the encryption is broken) or anything that 
takes place on a VPN (unless the VPN is hacked first), 
though it will show that encrypted communications 
and/or VPN use are taking place.

\section{Privacy And Anonymization}

Given how easily mesh activities are monitored, many 
users pursue privacy and anonymization options.

\subsection{Privacy Mode}

Characters who go into privacy mode hide their online 
presence and activities from others to a limited degree. 
The exact settings are adjustable, but typically involve 
masking their social profiles and presence to other 
users in the immediate vicinity, like having an unlisted 
phone number. Privacy mode can also be used to limit 
the use of mesh IDs and other data in access and 
transaction logs, applying a –30 modifier to attempts 
to research or track them by their online activity.

\subsubsection{Stealthed Signals}

Another tactic that can be taken for privacy is to stealth 
the wireless radio signals you emit. This method uses 
a combination of spread-spectrum signals, frequency 
hopping, and modulation to make your radio transmissions
harder to detect with scanning (p. 251). Stealthing
your signals is either a passive activity (Automatic
Action, –30 modifier on Interface Tests to locate the 
signal) or an active one (Complex Action, requires an 
Opposed Test to locate).

\subsection{Anonymization}

Anonymization takes the issue of privacy a bit further. 
The user does not just hide their mesh ID, but they 
actively use false mesh IDs and take other measures to 
reroute and obfuscate their datatrail. Anonymization 
is a necessity both for clandestine operatives and those 
engaging in illicit mesh activities.

\subsubsection{False Mesh Ids}

The easiest method of making mesh activities anonymous
is to set your muse to supply false mesh IDs in
online transactions. Though illegal in many jurisdictions
this is an easy task for any character or muse to
do. Multiple false IDs are used, making it extremely difficult
for anyone to tie all of the user's activities together.

This method makes it extremely difficult for anyone 
to track the user's online actions. Someone attempting 
to track the character via these false mesh IDs must 
beat them in an Opposed Test, pitting their Research 
skill with a –30 modifier against the character's (or 
more likely, their muse's) Infosec skill. This is a Task 
Action with a base timeframe of 1 hour, adjusted 
higher according to the amount of activity they hope 
to track. If successful, the tracker manages to dig 
together enough correlating evidence and records of 
false IDs to get a picture of the character's activities 
(how thorough this picture is depends on their MoS). 
If the fail, the anonymous character has effectively 
camouflaged themselves in the mesh.

Actively monitoring a character who is fluctuating 
their mesh ID with a sniffer program, or physically 
tracking them via the mesh, is next to impossible, as 
the continual shifting of IDs and intentional decoys 
make it too difficult to keep up.

\subsubsection{Anonymous Account Services}

A number of people—not just criminals, hackers, and 
secret agents—have an interest in keeping some of 
their affairs anonymous. To meet this demand, various 
online service vendors offer anonymous accounts for 
messaging and credit transfers. Some of these vendors 
%%% 255
are legit business (in places where it is legal), some 
are criminals operating illegally, others are hacktivists 
promoting the privacy meme, and still others are hypercorps
or other organizations offering such services
internally to their own staff/membership.

The interaction between the vendor and user is 
encrypted and anonymous, with no logs kept, so even 
if the vendor's servers are hacked, an intruder will 
not find any leads. While some anonymous accounts 
are established for regular use, the truly paranoid use 
(multiple) one-time accounts for maximum security. 
One-time accounts are used for a single message incoming
or outgoing) or credit transaction, and then
are securely erased.

Tracking an anonymous account is a practical 
impossibility, and something that only an extremely 
resourceful organization employing a systematic and 
expensive effort could attempt.

\subsubsection{Disposable Ectos}

Another option for those seeking privacy and security
is to simply use disposable ectos. Using this
method, all activity is routed through a specific ecto 
(using its mesh ID), the ecto is used for a limited 
period (until it gets hot), and then it is simply discarded
or destroyed.

\section{Mesh Security}

Given the lessons of the Fall and the very real risk still 
posed by hackers, virii, and similar threats, network 
security is taken extremely seriously in \textit{Eclipse Phase. }
Four methods are typically used: authentication, firewalls
active monitoring, and encryption.

\subsection{Authentication}

Most devices, networks (PANs, VPNs, etc.), and services
require some kind of authentication (a process
by which a system determines whether the claimed 
identity of a user is genuine) before they grant an account
and access privileges (p. 246) to a user. There
are several different ways for a system to authenticate 
a user. Some are more reliable and secure than others, 
but for the most part, the more secure the method, the 
higher the operational expenses.

\textbf{Account:} If you have access to an account on one 
system, this may give you automatic access to related 
systems or subsystems. This is typical of slaved devices 
(p. 248), where access to the master automatically 
grants you access to slaves.

\textbf{Mesh ID:} Some systems accept mesh IDs as authentication
This is extremely common with most public
systems, which merely log the mesh ID of any user 
that wishes access. Other systems will only allow 
access to specific mesh IDs, but these are vulnerable 
to spoofing (p. 255).

\textbf{Passcode:} This is a simple string of alphanumeric 
characters or logographic symbols, submitted in an 
encrypted format. Anyone with the passcode can 
access the account.
\textbf{Biometric Scan:} This calls for a scan of one or more 
of the user's biometric signatures (fingerprint, palm 
print, retinal scan, DNA sample, etc.). Popular before 
the Fall, such systems have fallen out of use as they 
are impractical with synthmorphs or users that frequently
resleeve.
\textbf{Passkey:} Passkey systems call for some of encrypted 
code that is either hardwired into a physical device 
(that is either implanted or physically jacked into 
an ecto) or extracted from specialized software. Advanced
passkeys combine hardwired encryption with
physical nanotech etching to create a unique key. To 
access such systems, the passkey must either be acquired
or somehow spoofed.
\textbf{Ego Scan:} This system authenticates the user's ego 
ID (p. 279).
\textbf{Quantum Key:} Quantum key systems rely on the unbreakable
encryption of quantum cryptography (p. 254).

\subsection{Firewalls}

Firewalls are software programs (sometimes hardwired
into a device) that intercept and inspect all traffic
to and from a protected network or device. Traffic
that meets specified criteria that designates it as safe 
is passed through, whereas all other traffic is blocked. 
In \textit{Eclipse Phase,} every network and device can be 
assumed to have a firewall by default. Firewalls are 
the main obstacle that an intruder must overcome, as 
discussed under \textit{Intrusion Tests,} p. 255.
Like other gear, firewalls come in varying quality 
levels and so may apply modifiers to certain tests.

\subsection{Active Monitoring}

Instead of relying on authentication and firewalls 
alone, secure systems are actively monitored by a security
hacker or a muse. These digital security guards
inspect network traffic using a number of software 
tools and applications that flag conspicuous events. 
Active surveillance makes intrusions more difficult, 
since the interloper must beat the monitoring hacker/
AI in an Opposed Test (see \textit{Intrusion,} p. 254). Active 
monitoring also includes monitoring any devices 
slaved to the monitored system.
Characters may actively monitor their own PANs if 
they so choose, though this requires a moderate level 
of attention (count as a Quick Action). It is far more 
common for a muse to actively guard a user's PAN.

\subsection{Encryption}

Encryption is an exceptionally effective extra layer of 
security. There are two types of encryption commonly 
used in \textit{Eclipse Phase:} public key cryptosystems and 
quantum cryptography.

\subsubsection{Public Key Crypto}

In public key cryptosystems, two keys are generated 
by the user, a public key and a secret key. The public 
key is used to encrypt messages to that user, and is 
made freely available. When messages are encrypted 
using that public key, only the secret key—controlled 

of 

ly

d
%%% 256
by the user—can decrypt them. Public key crypto is 
widely used both for encrypting data traffic between 
two users/networks/devices and for encrypting 
files. Due to the strength of the public key system 
algorithms, such crypto is essentially unbreakable 
without a quantum computer (see \textit{Quantum Code-}
\textit{breaking,} p. 254).

\subsubsection{Quantum Cryptography}

Quantum key distribution systems use quantum mechanics
to enable secure communications between two
parties by generating a quantum key. The major advantage
of transmitting information in quantum states is
that the system itself instantly detects eavesdropping 
attempts as quantum systems are disturbed by any sort 
of external interference. In practical terms, this means 
that quantum encrypted data transfers are unbreakable 
and attempts to intercept automatically fail. Note that 
quantum crypto doesn't work for basic file encryption, 
its only use is in protecting communication channels.

While quantum key systems have an advantage over 
public key systems, they are both more expensive and 
less practical. In order to generate a quantum key, 
the two communications devices must be entangled 
together on a quantum level, in the same location, 
and then separated. So quantum key encrypted 
communications channels require some setup effort, 
especially if long distances are involved. Since the 
implementation of quantum cryptographic protocols 
is an extraordinary expense, it is usually only adopted 
for major high-security communications links.

\subsubsection{Breaking Encryption}

What this means is that encrypted communications 
lines and files are very safe if using public key systems, 
and that data transfers are absolutely safe if using 
quantum crypto. Gamemasters should take note, however
while this may be useful to player characters, it
may also hinder them. If the characters need to get at 
something that is encrypted, they're going to need to 
figure out some way to get the secret key's passcode. 
Common methods include the old standbys of bribery, 
blackmail, threats, and torture. Other options involve 
espionage or social engineering to somehow acquire 
the passcode. Hackers could also find some other 
method to compromise the system and gain inside 
access, bypassing the encryption entirely.

\subsubsection{Quantum Codebreaking}

As noted above, quantum computers can also be used 
to break public key encryption. This requires an Infosec 
Task Action Test with a +30 modifier and a timeframe 
of 1 week (once started, the quantum computer finishes 
the job on its own; the user does not need to provide 
constant oversight). Gamemasters should feel free to 
modify this timeframe as fits the needs of their game. 
Note that quantum computers cannot break quantum-encrypted
communications, only encrypted files.

\section{Intrusion}

The art of intrusion lies in penetrating a device's security
The best methods involve infiltrating a system quietly
without catching a watchdog's attention, by using
exploits—code glitches, flawed security protocols—to 
create a path circumventing the target's defenses. When 
called for, however, a hacker can toss aside pretenses 
and attempt to brute-force their way in.

\subsection{Preconditions}

In order to hack a device, the hacker needs to establish 
a direct connection to the target computer system. If 
the hacker is making a direct wireless connection to 
the target, the target system must be wireless-capable 
and within range (p. 299), and the hacker must know 
the target is there (see \textit{Wireless Scanning,} p. 251). If 
the system is hard-wired, the hacker must physically 
jack in by using a regular jacking port or somehow 
tapping into a cable that carries the network's data 
traffic. If the hacker is accessing the target through the 
%%% 257
mesh, the target system must be online and the hacker 
must know it's mesh ID (p. 246) or otherwise be able 
to track it down (p. 251).

\subsection{Circumventing Authentication}

Rather than hacking in, an intruder can try to subvert
the authentication system used to vet legitimate
users. The easiest manner of doing this is to somehow 
acquire the passcode, passkey, or whatever authentication
method the target uses (p. 253). With this in hand,
no test is necessary to access the system; the hacker 
simply logs in just like a legitimate user and has all of 
the normal access privileges of that user.

Lacking a passcode, the hacker can try to subvert 
the authentication system in one of two other ways: 
spoofing or forgery.

\subsubsection{Spoofing Authentication}

Using this method, the hacker attempts to disguise 
their signals as coming from the legitimate, authenticated
user, rather than from themself. If successful,
the system is fooled by this masquerade, accepting the 
hacker's commands and activity as if they came from 
a legitimate user. Spoofing is more difficult to pull off, 
but is very effective when it works.

To spoof a legitimate user, the hacker must be using 
both sniffer and spoofing software (p. 331). The 
hacker must then monitor a connection between the 
legitimate user and the target system, and succeed 
in an Infosec Test to sniff the traffic between them 
(p. 252). Apply a –20 modifier if the user has security
account privileges, –30 if they have admin rights
(p. 247). If the connection is encrypted, this will fail 
unless the hacker has the encryption key.

Armed with this data, the hacker then uses it to 
disguise their signals. This requires an Infosec Test, 
modified by the quality of the system's firewall and 
the hacker's spoofing program. If successful, communications
sent by the hacker are treated as coming
from the legitimate user.

\subsubsection{Forging Authentication}

Biometric and passkey systems used for authentication 
(p. 253) can potentially be forged hackers who are 
able to get a look at the originals. The means and techniques
for doing so differ, and are beyond the scope of
this book, but successfully forging such systems would 
allow a hacker to log in as the legitimate user.

\subsection{Intrusion Tests}

Hacking into a node is a time-consuming task. The 
target system must be carefully analyzed and probed 
for weaknesses, without alerting its defenses. Depending
on the type of security in place, more than one test
may be called for.

Hackers require special exploit software (p. 331) to 
take advantage of security holes, but software does 
not a hacker make. What really counts is Infosec skill 
(p. 180), which is the ability to use, modify, and improvise
exploits to their full advantage.

\subsubsection{Defeating The Firewall}

Lacking a passcode, the hacker must break in the old-fashioned
way: discreetly scanning the target, look for
weaknesses, and take advantage of them. In this case 
the hacker takes their exploit software and makes an 
Infosec Test. This is handled as a Task Action with a 
timeframe of 10 minutes. Various modifiers may apply, 
such as the quality of the exploit software, the quality 
of the Firewall, or the alertness of the target system. The 
gamemaster may also modify the timeframe, shortening 
it to reflect systems that are cookie-cutter common with 
known security flaws or raising it as fitting for a top-of-the-line
system with still-unreleased defenses.

By default, a hacker trying to break in this way is 
pursuing standard user access rights (p. 247). If the 
hacker wishes to obtain security or admin privileges on 
the system, apply a –20 or –30 modifier, respectively.

If the Infosec Test succeeds, the intruder has invaded 
the system without triggering any alarms. If the system is 
actively monitored (p. 253), they must now avoid detection
by that watchdog (see below). If there is no active
monitor, the intruder gains the status of \textit{Covert} (see \textit{In-}
\textit{truder Status,} p. 256). If the intruder scored an Excellent 
Success, however, their status is \textit{Hidden }(p. 256).

\textbf{Probing:} Players may choose to take the time (p. 
116) when probing the target for weakness and exploits
In fact, this is a common procedure when a
hacker wants to ensure success.

\subsubsection{Bypassing Active Security}

If a system is also actively monitored (p. 253), the 
hacker must avoid detection. Treat this as a Variable 
Opposed Infosec Test between the intruder and the 
monitor. The outcome depends on both rolls:

If only the intruder succeeds, the hacker has accessed
the node without the monitor or the system
%%% 258
noticing. The hacker has acquired \textit{Covert} status (p. 
256). If the hacker scored an Excellent Success, their 
status is \textit{Hidden }(p. 256).

If only the monitor succeeds, the hacking attempt 
is spotted and the monitor may immediately lock the 
hacker out of the system before they manage to fully 
break in. The intruder may try again, but the monitor 
will be vigilant for further intrusions.

If both succeed, the intruder has gained access but 
the monitor is aware that something strange is going 
on. The hacker acquires \textit{Spotted} status.

If both fail, continue to make the same test on each 
of the hacker's Action Phases, until one or both succeed.

\subsection{Intruder Status}

\textit{Intruder status} is a simple way of measuring an 
invader's situation when they are intruding upon 
a system. This status has an impact on whether the 
hacker has caught any attention or if they managed 
to remain unobtrusive. Status is first determined when 
the intruder access the system, though it may change 
according to events.

Note that intruder status is a separate matter from 
account access privileges (p. 246). The latter represents 
what a user can legally do on a system. The former 
indicates how aware the system is of the hacker's true 
nature as an intruder.

\subsubsection{Hidden}

An intruder with Hidden status has managed to silently
sneak into the system without anyone noticing. The
system's security is totally unaware of their presence 
and may not act against them. In this case, the hacker 
is not using an account so much as they are exploiting
a flaw in the system that grants them a nebulous,
behind-the-scenes sort of presence in the system. The 
hacker effectively has admin access rights, but does 
not show up as an admin-level user in logs or other 
statistics. Hidden characters receive a +30 modifier on 
any efforts to subvert the system.

\subsubsection{Covert}

An intruder with Covert status has accessed the 
system in a manner that doesn't attract any unusual 
attention. For all intents and purposes, they appear 
to be a legitimate user with whatever access rights 
they sought. Only extensive checking will turn up any 
abnormalities. The system is aware of them, but does 
not consider them a threat.

\subsubsection{Spotted}

Spotted status indicates that the system is aware of an 
anomaly or intrusion but hasn't zeroed in on the intruder
yet. The hacker appears to be a legitimate user
with whatever access rights they sought, but this will 
not hold up under close scrutiny. The system goes on 
passive alert (inflicting a –10 modifier to the hacker's 
activities on that system) and may engage the hacker 
with passive countermeasures (p. 257).

\subsubsection{Locked}

Locked status means that the intruder—including their 
datatrail—has been pinned down by system security. 
The hacker has access and account privileges, but they 
have been flagged as an interloper. The system is on 
active alert (inflicting a –20 modifier on the hacker's actions
and may launch active countermeasures (p. 257)
against the intruder.

\subsection{Changing Status}

An intruder's status is subject to change according to 
their actions and the actions of the system.

\subsubsection{Upgrading Status}

A hacker can attempt to improve their status in order 
to better protect themself. This requires a Complex 
Action and an Infosec Test. If the hacker has Spotted 
status, this is an Opposed Test between monitor and 
intruder. If the hacker wins and scores an Excellent 
Success (MoS of 30+), they have upgraded their status 
by one level (for example, from Covert to Hidden). 
Intruders with Locked status may not upgrade.

\subsubsection{Zeroing In}

A security hacker or muse that is actively monitoring 
a system can take a Complex Action and attempt to 
hone in on a Spotted intruder. An Opposed Infosec Test 
is made between both parties. If the system's defender 
wins, the hacker is downgraded to Locked status.

\subsubsection{Failing Tests}

Any time an intruder scores a Severe Failure (MoF 
30+) on a test involving manipulating the system, they 
are automatically downgraded one status level (from 
Covert to Spotted, for example). If a critical failure 
is rolled, they immediately give themselves away and 
achieve Locked status.

\subsection{Brute-Force Hacking}

Sometimes a character simply doesn't have time to do 
the job right, and they need to hack in now or never. 
In this case the hacker engages the target system immediately
head on, without taking any time to prepare
an attack. The hacker simply brings all of their
software exploit tools to bear, throwing them at the 
target and hoping that one works. This is handled as 
an Infosec Test, but as a Task Action with a timeframe 
of 1 minute (20 Action Turns). The hacker receives 
a +30 modifier on this test. Many hackers choose to 
rush the job (see \textit{Task Actions,} p. 120), in order to cut 
this time even shorter.

The drawback to brute-force hacking is that it immediately
triggers an alarm. If the system is actively monitored
the hacker must beat the monitor in an Opposed
Infosec Test or be immediately locked out as soon as 
they break in. Even if they succeed, the hacker has 
\textit{Locked} status and is subject to active countermeasures.
%%% 259

\section{Intrusion Countermeasures}

If an intruding hacker fails to penetrate a system's 
defenses (i.e., they are Spotted or Locked, see p. 256), 
then the system goes on alert and activates certain 
defenses. The nature of the applied countermeasures 
depends on the capabilities of the system, the abilities 
of its security defender(s), and the policy of its owner/
admins. While some nodes will simply seek to kick 
the intruder out and keep them shut out, others will 
actively counterattack, seeking to track the intruder 
and potentially hack the intruder's own PAN.

\subsection{Security Alerts}

Security alerts come in two flavors: passive and active.

\subsubsection{Passive Alert}

Passive alerts are triggered when an intruder hits Spotted
status. The system immediately flags a visual or
acoustic cue to anyone actively monitoring the system 
and possibly the owner or admins. It immediately 
launches one or more passive countermeasures (see 
below). Depending on the system, extra security hackers
or AIs may be brought in to help investigate. If the
intruder is not encountered again or located within a 
set time period (usually about 10 minutes), the alarm 
is deactivated and the event is logged as an anomaly. 
Depending on the security level of the system, someone
may analyze the logs at some point and try to
ascertain what happened—and prevent it from happening
again.

All intruders suffer a –10 modifier for tests involving
a system that is on passive alert.

\subsubsection{Active Alert}

An active alert is triggered when an intruder hits 
Locked status. The system immediately alerts the 
owners, admins, and monitoring security agents. 
Additional security assets (hackers and AIs) may be 
called in. The system also launches active countermeasure
against the intruder (see below). Active alerts are
maintained for as long as the intruder is present, and 
sometimes for a lengthy period afterwards just in case 
the hacker returns.

\subsection{Passive Countermeasures}

Passive countermeasures are launched as a precaution 
whenever an intruder acquires Spotted status.

\subsubsection{Locate Intruder}

A security hacker or AI monitoring a system may attempt
to track down the source of the passive alert.
See \textit{Zeroing In,} p. 256.

\subsubsection{Re-Authenticate}

When a passive alert is triggered, a firewall can be set 
to re-authenticate all active users, starting with the 
most recent. At the beginning of the next Action Turn, 
everyone on the system must take an action to log 
back in. For intruders, this means making an Infosec 
Test, modified by –10 for the passive alert, to satisfy 
the system that they are a legitimate user.

\subsubsection{Reduce Privileges}

As a protective measure, some systems will immediately
reduce access privileges available to standard
users, and sometimes security users as well. One 
common tactic is to protect all logs, backing them up 
and making sure no one has rights to delete them.

\subsection{Active Countermeasures}

Active countermeasures can only be launched if the 
intruder has acquired Locked status.

\subsubsection{Counterintrusion}

A security hacker or guardian AI can proactively 
defend a system by attacking the intruder's source. For 
this to occur, the intruder must first be successfully 
traced (p. 251). Once this occurs, the security forces 
can then launch their own intrusion on the hacker's 
home ecto/mesh inserts and/or PAN.

\subsubsection{Lockout}

A system that has locked onto an intruder may also 
attempt to lock them out. Lockout is an attempt to 
remove the compromised account, sever the connection
between the two, and dump the hacker from
the system.
Lockout must be initiated by someone with security
or admin privileges. An Opposed Infosec Test is
made, with the intruder suffering a –20 modifier for 
being Locked. If the character defending the system 
succeeds, the intruder is immediately ejected from 
the system and the account they used will be placed 
on quarantine or deleted. That account will not be 
usable again until a security audit approves it and 
replaces the authentication. Any attempt to access 
the system from the same mesh ID as the intruder 
automatically fails.

\subsubsection{Reboot/Shutdown}

Perhaps the most drastic option for dealing with an 
interloper is to simply shut down the system. In this 
case, the system closes all wireless connections (if it 
has any), logs off any users, terminates all processes, 
and shut itself down—thereby locking out the intruder. 
The disadvantage, of course, is that the system must 
interrupt its activities. For example, shutting down 
your mesh inserts or ecto means losing all communication
with teammates, access to augmented reality, and
control over slaved/linked devices.
Initiating a reboot/shutdown is only a Complex 
Action, but the actual process of shutdown takes 
anywhere from 1 Action Turn (personal devices) to 1 
minute (large hardwired networks with multiple users), 
determined by the gamemaster. Rebooting a system 
takes the same amount of time to get started again.

wn

ca-

nd 

x
%%% 260

\subsubsection{Trace}

For high-security systems, a popular countermeasure 
is to track the infiltrator's physical location via their 
mesh ID (see \textit{Physical Tracking,} p. 251). In most cases, 
habitat physical security is subsequently alerted and 
forwarded the position to take care of the criminal.

\subsubsection{Wireless Termination}

An alternative to shutdown or rebooting is simply to 
sever all wireless connections by shutting down the 
wireless capabilities of the system. The system will 
lose all active connections, but any intruders will be 
dumped. Wireless termination is a Complex Action to 
initiate and completes at the end of that Action Turn. 
Re-starting wireless connectivity takes 1 Action Turn.
%%% 261

\section{Subversion}

Once an intruder has successfully invaded a device 
or network, they can pursue whatever tasks they are 
interested in, as fitting that particular system. Depending
on the type of account the intruder hacked, they
may or may not have access privileges to do what 
they want to do. If their access rights allow it, the 
activity is handled like that of a legitimate user and 
no test is called for (unless the activity itself calls for 
some kind of test, such as Research). For example, a 
hacker who infiltrates a habitat's security system with 
a security account can monitor cameras, deactivate 
sensors, review recorded surveillance footage, and so 
on, as any legitimate user with security right would be 
allowed to do.
Engaging in any sort of activity for which you 
don't have access rights is more difficult and requires 
hacking the system. This typically requires an Infosec 
Success Test, modified by the difficulty of the action 
as noted on the Subversion Difficulties table. In most 
cases this in not an Opposed Test even if the system is 
actively monitored, unless specifically stated otherwise. 
Failing such tests, however, will result in a change of 
the hacker's intruder status (see \textit{Failing Tests,} p. 256).
Examples for different types of system subversion 
are given in the \textit{Subversion Examples} sidebar. This 
is not an exhaustive list, however, and gamemasters 
and players are encouraged to improvise game effects 
in case an action has not been explicitly described.

\subsection{Augmented Reality Illusions}

A hacker who has infiltrated an ecto, mesh inserts, or 
some other device with an AR interface may inject 
different kinds of visual, auditory, tactile, and even 
emotional illusions into the augmented reality of 
the device's user, depending on the type of interface 
used. How the hacked user will respond to the illusion
depends on a number of factors, such as whether
they are aware of the intruder (i.e., the hacker has 
Spotted or Locked status), what type of interface they 
are using (entoptic or haptic), and how realistic the 
illusion is.
The best illusions are, of course, crafted in advance, 
using the best image and sensory manipulation tools 
available. Such illusions are hyper-realistic. Anyone 
making a Perception Test to identify them as fake 

–0

Execute commands, view restricted information, run restricted software, open/close connections to other systems, 

read/write/copy/delete files, access sensor feeds, access slaved devices

–10

Change system settings, alter logs/restricted files

–20

Interfere with system operations, alter sensor/AR input

–30

Shut system down, lockout user/muse, launch countermeasures at others
–0

Give orders to drones
–10

Alter sensor system parameters, disable sensors or weapon systems
–20

Alter smartlink input, send false data to AI or teleoperator
–30

Lockout AI or teleoperator, seize control via puppet sock
–0

Interact with entoptics, befriend everyone in range, make online 

purchases using user's credit, intercept communications, log activity
–10

Alter social network profile/status, adjust AR filters, tweak sensory 

interface, change AR skin, change avatar, access VPN
–20

Block or shufﬂ e senses, inject AR illusions, spoof commands to 

drones/slaved devices
–30

Boot user out of AR
–0

 Open/close doors, stop/start elevators, operate intercom
–10

Adjust temperature/lighting, disable safety warnings, 

replace entoptic skin, lock doors, switch traffic timers
–20

Disable subsystems (plumbing, recycling, etc.), disable wireless links, 

dispatch repair crews
–30

Override safety cutoffs
–0

Move/manipulate cameras/sensors, locate security systems/guards/bots
–10

Adjust patterns of sensor sweeps, view security logs, 

disable weapon systems
–20

Delete security logs, dispatch security teams
–30

Disable alerts
–0

View current status of simulspace, simulmorphs, and accessing egos
–10

Change domain rules, add cheats, alter parameters of story, 

alter simulmorphs, change time dilation
–20

Eject simulmorph, alter/erase character AIs
–30

Abort simulation
–0

Get status report, use device functions
–10

Adjust AI/voice personality settings, adjust timed operation schedule
–20

Disable sensors, disable device functions
%%% 262
suffers a –10 to –30 modifier (gamemaster's discretion
An eclectic collection of software programs
offer a diverse range of AR illusions.

Hackers may also improvise illusions on the 
fly, usually by patching in sensor data from others 
sources, though this is more difficult and more easily 
spotted (typically adding a +10 to +30 modifier to 
Perception Tests). The advantage is that the hacker 
can modify the illusion in reaction to the user's actions
or environmental factors on the fly. AR illusion
software, however, also offers some template illusions 
that can be modified and controlled in real-time via a 
connected interface.

Whenever a user is bombarded with AR illusions, the 
gamemaster should make a secret Perception Test to see 
if they spot the deception. Even if they do, however, the 
character may still react to them. Almost anyone will 
duck when they see an object suddenly flying at their 
face, as their body reacts on its own before the brain 
comprehends that it's an illusion and not a threat.

Aside from their deceptive value, illusions can be 
used to distract users or otherwise impair their perceptive
faculties. For example, dark illusory clouds can
obscure vision, ear-wrenching high-volume noises can 
make people cringe, and a persistent tickling sensation 
might drive anyone crazy. Such effects can apply a –10 
to –30 modifier to Perception Tests and other actions, 
but the user can also adjust their filters and/or turn 
their AR off if necessary.

\subsection{Backdoors}

A backdoor is a method of bypassing a system's normal 
authentication and security features. It enables a hacker 
to sneak into a system by exploiting a flaw (which can 
take the form of an installed program or modification 
to an existing program or hardware device) that was 
integrated into the system previously, either by themself 
or another hacker (who shared the backdoor).

To install a backdoor, the hacker must successfully 
infiltrate the system and succeed in both a Programming
and an Infosec Test (or an Opposed Infosec Test
if the system is actively monitored). The Programming 
Test determines how well the backdoor is crafted and 
hidden within system processes, while the Infosec Test 
represents incorporating it into the system without security
noticing. Modify the Programming Test by –20
if the hacker wants to have security privileges when 
using the backdoor, –30 for admin.

Once installed, using a backdoor requires no test to 
access the system—the hacker simply logs on as if they 
were a legitimate user, gaining Covert status. Anyone 
who is aware of a backdoor's details may use it.

How long the backdoor lasts depends on many factors
and is largely up to the gamemaster. Backdoors
are only likely to be spotted during complete security 
audits, so more paranoid systems are likely to detect 
them earlier. Security audits may also occur when an 
intruder is Spotted but never Locked. Security audits 
are a Task Action with a timeframe of 24 hours. The 
character conducting the audit makes an Infosec 
Test to spot the back door. If the backdoor's installer 
scored an Excellent Success on their Programming 
Test, this Infosec Test suffers a –30 modifier.

\subsection{Crashing Software}

Intruders can attempt to crash software programs 
by killing the processes that run them. This requires a 
Complex Action and an Infosec Test. Note that some 
software is set to immediately respawn, but this can take 
from 1 Action turn to 1 minute, depending on the system.

Hackers may crash AIs, AGIs, and even infomorphs 
this way, but the process is more difficult. In this case, 
an Opposed Infosec Test is made against the target, 
who is immediately aware they are under attack. Two 
consecutive tests must succeed in order to crash an AI, 
three in order crash an AGI or infomorph. If successful, 
the AI/infomorph immediately reboots, which generally 
takes 3 Action Turns, longer if the gamemaster chooses.

\subsection{Eliminating Intrusion Traces}

Hackers who have avoided being Locked may attempt 
to clean up all traces of their intrusion before they 
exit a system. This involves erasing incriminating data 
in the access and security logs, and otherwise hiding 
any evidence of system tampering. This requires a 
Complex Action and an Infosec Test, or an Opposed 
Infosec Test if the system is actively monitored. If successful
the intruder has wiped anything that might be
used to track them down later, such as mesh ID, etc.

\subsection{Hacking Vpn}

Virtual private networks (VPNs) are more challenging
to hack than standard devices. Because they exist
as an encrypted network within the mesh, accessing 
channels of communication within a VPN is all but 
impossible without the encryption key. This means 
any attempt to sniff the VPN traffic is also impossible
without the key.

The only way to hack a VPN is to hack into a 
device that is part of the VPN and running the VPN 
software. Once an intruder has access to such a device, 
they can attempt to access the VPN. The account the 
hacker has compromised may have VPN privileges, in 
which case they are in. If not, they must hack access, 
requiring an Infosec Test with a Minor modifier (–10).

Once access to the VPN is acquired, the hacker may 
treat the VPN like any other network. They may hack 
other devices on the VPN, sniff VPN traffic,  track 
other users on the VPN, research data hidden away 
on the VPN, and so on.

\subsection{Scripting}

A script is a simple program—a batch of instructions—
that a hacker can embed in a system to be executed at 
a later pre-scheduled time or upon a certain trigger 
event, even without the hacker being present. When 
activated, the script will undertake a number of system 
operations limited by the abilities of the operating 
system and the access rights the hacker had when 
they implemented the script in the system. Scripts are 
%%% 263
a great way for a hacker to subvert a system without 
necessarily being in danger when they do it.

Scripts can be programmed on the fly or preprogrammed
When composing the script, the character
must detail what system operations the set will call 
for, in what order and at what times (or at what trigger
steps). The script cannot contain more steps/tasks
than the character's Programming skill ÷ 10 (round 
down). To program a script, the character must succeed
in a Programming Test with a timeframe determined
by the gamemaster.

To load the script, the character must have successfully
intruded in the system and must succeed
in an Infosec Test (or an Opposed Infosec Test if the 
system is actively monitored). If successful, the script 
is loaded into the system and will run as programmed.

Once the script is activated, it carries out the preprogrammed
sequence of actions. The programmer's
Infosec skill is used for any tests those actions call for.

Inactivated scripts may be detected in a security 
audit, just like backdoors (p. 260).

\section{Cyberbrain Hacking}

Pods and synthmorphs (including some bots and vehicles
are equipped with cybernetic brains. While this
technology allows a transhuman ego to sleeve into and 
control these forms, they carry the disadvantage of being 
vulnerable to hacking, like any other electronic device.

Cyberbrains are not wireless-enabled for security 
reasons, but they do have access jacks (p. 306) and 
are directly linked to mesh inserts. This means that 
in order to hack a cyberbrain, the hacker either must 
have direct physical access to the morph's body in 
order to jack in, or they must first hack into the mesh 
inserts and then break into the cyberbrain from there.

Due to their importance, cyberbrains are equipped 
with numerous hard-coded security features that make 
intrusion very difficult. Apply a –30 modifier to all 
attempts to hack into and subvert a cyberbrain. (Note 
that the –30 modifier for hacking an admin account 
does not apply to cyberbrains.)
Cyberbrains are treated just like other systems for 
intrusion and subversion purposes, but since they 
house the morph's controlling ego they present several 
unique hacking opportunities.

\subsection{Entrapment}

An intruding character can attempt to \textit{lock in} an ego, 
preventing it from evacuating the cyberbrain. The 
hacker (with the –30 modifier noted above) must 
beat the defending character or muse in an Opposed 
Infosec Test. If successful, the ego is prevented from 
transferring itself to another system.
To fully pen the ego in, the ego character and its 
protecting muse must also be locked out (p. 257) from 
controlling the cyberbrain's system, otherwise the ego 
could potentially be freed.
 Trapped egos are quite vulnerable. They could, for 
example, be subject to enforced uploading, enforced 
forking, or psychosurgery.

\subsection{Memory Hacking}

All cyberbrains incorporate mnemonic augmentation 
(p. 307), or digitally recorded memories. A hacker 
who has accessed the cyberbrain can read, alter, or 
delete these memories with a successful Research or 
Interfacing Test (the –30 modifier applies).

\subsection{Puppeteering}

Most cyberbrains also incorporate a puppet sock (p. 307), 
enabling remote users to take over the pod or synthmorph
body and control it via teleoperation or jamming
(p. 196). This allows a hacker to seize control of the body 
and manipulate it remotely. To do so, the hacker must 
take a Complex Action and beat the defending character 
or muse in an Opposed Infosec Test; the hacker suffers 
the –30 modifier noted above.
A defender who is not locked out may continue 
to fight for control of the morph, using a Complex 
Action. In this case, another Opposed Infosec Test 
is called for. This can result in a situation where the 
morph repeatedly slips control from the hacker to the 
defender, or perhaps slips into a catatonic state as the 
two sides battle it out.

\subsection{Scorching}

Having direct access to a cyberbrain opens the possibility
for certain kinds of attacks that are normally infeasible
due to the strict content filtering that occurs on
the link between the cyberbrain and mesh inserts. One 
of these possibilities is \textit{scorching}—the use of damaging 
neurofeedback algorithms to harm the victim's mind.
In order to make a scorching attack, the cyberbrain 
intruder must deploy a scorch program. To utilize a 
scorch program, the intruder must beat the defending 
Ego in an Opposed Infosec Test. The –30 modifier for 
cyberbrain hacking applies to the attacker.
Several types of scorch programs exist, with different
effects: cauterizers (damage), bedlams (stress),

in-

on

ne 

g
%%% 264
spasms (pain), nightmares (fear), and shutters (sensory 
deprivation). These are described on p. 332 of \textit{Gear.}

\subsection{Shutdown}

If a cyberbrain is shut down (p. 257), the morph immediately
ceases activity, perhaps collapsing or rolling
to a stop. Pods will appear to be in a coma. The ego, 
however, will be rebooted along with the cyberbrain.

\subsection{Terminate Cortical Stack Feed}

The cyberbrain feeds backup data to the cortical stack. 
This is a one-way connection, so the cortical stack may 
not be hacked, but the transfer of data may be cut off. 
This termination action requires an Opposed Infosec 
Test between the hacker (with the –30 modifier) and 
the defender. The ego's backup will not be updated for 
as long as the connection remains off.

\section{Radio Jamming}

Radio jamming is a method of transmitting radio 
signals that deliberate interfere with other radio 
signals in order to disrupt communications. In the 
highly-networked world of \textit{Eclipse Phase,} intentional 
jamming is often illegal, not to mention rude.

Radio jamming does not require any special equipment
other than a standard wireless device, such as
an ecto or mesh inserts. Jamming can be \textit{selective }
or \textit{universal.} Selective jamming targets a particular 
device or set of devices. In order to selectively jam, 
the character must have scanned the target device(s) 
(p. 251). Universal devices target all radio-equipped 
devices indiscriminately.

Jamming simply requires a Complex Action and an 
Interfacing Test. If successful, affected devices within 
range have their radio communications disrupted—
they are cut off from the mesh and may not communicate
wirelessly. Wired devices are unaffected.

Devices equipped by AIs will automatically attempt 
to overcome jamming, which requires a Complex 
Action (transhuman users may also do the same). In 
this case, a Variable Opposed Test is made between 
the jammer and defender. If the jammer wins, all communications
are blocked; if the defender wins, they
are unaffected. If both parties succeed, the defender's 
communications are impacted but not completely cut 
off. The gamemaster decides how much information 
the defender can get through, and how this situation 
affects mesh use.

\subsection{Jamming Radar}

Jamming can also be used to interfere with radar. In 
this case, if the jammer makes an Interface Test. If 
successful, the radar suffers interference, imposing 
a –30 modifier on all sensor-related tests. The entity 
operating the radar may attempt to overcome this 
interference by beating the jammer in an Opposed 
Interface Test.

\section{Simulspaces}

Simulspaces are virtual reality environments where 
the resolution advances beyond realistic high definition
and into the hyper-real. The environments they
create are comprehensive and authentic illusions, from 
aspects like lighting, day or lunar cycles, and weather 
down to minute details and sensations. Jacking into 
a simulspace is much like crossing over into a alternate
world or reality, which is why simulspaces have
become increasingly popular in entertainment.

While simulspaces usually cannot harm characters 
immersed in them as the sensory algorithms are not 
intended to be offensive programs or routines, experiences
in simulspaces can have a strong psychological
impact on an ego, as the simulation is as close to 
reality as you can get. A character who is ``physically'' 
tortured within a simulspace will not be physically 
harmed, but the mental stress of the experience might 
still be sufficient to cause permanent traumas.

\subsection{Simulmorphs}

Characters access simulspace using an avatar-like persona
called a simulmorph. This simulmorph is created by
the simulspace, based on the domain rules of the simulation
and certain characteristics of the morph or ego
accessing the simulation. Depending on the simulation, 
this simulmorph may be customizable to varying degrees.

While interacting with the simulation, treat simulmorphs
as basic infomorphs for all rules purposes,
even if the egos are still possessing another morph 
body in reality.

When accessing a simulspace, muses are usually not 
transferred into the simulation, though they can potentially
come along if domain rules permit it. In this
case, muses are treated as separate characters within 
the simulspace with their own simulmorph body.

Depending on the role a simulspace is intended to play 
in the story, the gamemaster may want to invent physical
stats'' for the simulmorph bodies, especially if the
characters are likely to spend a lot of time in the simulation
These statistics can literally be made up—it is a virtual
reality after all, and anything goes. Alternately, the
gamemaster can simply wing it and invent any necessary 
statistics on the fly as the need for them comes up.

\subsection{Immersion}

When a character immerses themselves in a simulspace, 
they ``become'' the simulmorph. The character's physical
body, typically secluded and protected in a vat or
couch, slumps inertly. While immersed, they suffer –60 
on all Perception Tests or attempts to take action with 
their physical morph. Characters can enter and leave 
the simulspace as will, but toggling in or out takes a 
Complex Action.

If the simulspace crashes or the character is otherwise
dumped from it, they immediately resume control
of their own morph as normal. VR dumpshock is
extremely jarring, and the character suffers 1d10 ÷ 2 
mental stress.
%%% 265

\subsection{External Mesh Interaction}

A character accessing a simulspace may still interact 
with the mesh (and through it, the outside world) 
assuming the domain rules allow for it. Any outside 
interactions are subject to time dilation issues, however
For example, in a simulspace running faster
than real time, holding a chat with someone in outside
meatspace is excruciatingly slow, as real-world
seconds translate into minutes in VR. If a character 
wishes to directly access other mesh nodes, they must 
toggle or log out of the simulspace.

\subsection{Simulspace Rules}

Since a simulspace is an alternate world whose realism 
matches reality, characters use their physical skills and 
aptitudes as if they were acting in the real world with 
few exceptions:
•  Though intrusion and hacking can be represented 

as another layer of the simulation, there is no 

actual hacking within the simulspace (see \textit{Hack-}

\textit{ing Simulspaces}).
•  Asyncs cannot use their psi abilities in simulspace, 

though such abilities can be simulated.
•  Any ``physical'' damage taken in the simulspace is 

treated as ``virtual'' damage. While virtual injuries 

and wounds use the same mechanics, characters 

that die in a simulspace are usually simply ejected 

from the simulation. In some cases ``dead'' char-

acters are brought into a white room and can 

re-enter or just watch the simulation, depending 

on the domain rules.
•  Mental stress or trauma inflicted during a simu-

lation carries over to the ego as real Lucidity 

damage. At the gamemaster's discretion, some 

mental stress may be reduced if the character is 

aware that they are in a simulation.

\subsection{Domain Rules}

Anything goes in a simulspace, as dictated by 
the domain rules. A simulspace may range from 
approximating reality very closely to differing drastically
Gravity might fluctuate, the visual light spectrum
might not exist, characters might heal virtual
damage effortlessly, simulmorphs may be capable of 
transmogrifying into other creatures, everything might 
be underwater—the possibilities are endless, limited 
only by imagination. In game terms, this allows the 
gamemaster to make up rules on the fly.

\subsection{Cheating}

As with any good game, simulspaces provide ways 
to cheat. Cheats are either built into the simulspace 
software or (externally) programmed in by a hacker. 
Cheats allow for a character to break the domain rules 
in some way. This may be a special power, a way to 
alter some environmental factor (like flying), altering 
the time dilation, some sort of power-up ability, a 
way to get info on other simulmorphs, or a short-cut 
through part of the simulation. In game terms, cheats 
might provide bonus modifiers to certain skill or stat 
tests made by a simulmorph. Cheating is usually forbidden
Players who cheat in a simulspace game and
who get caught may face eviction from the simulspace.

\subsection{Hacking Simulspaces}

Since simulspaces are complex virtual environments 
and often run on time dilation, hackers cannot hack 
them in a normal manner when they participate in the 
simulation. There are ways to affect and influence the 
simulation from within, but the degree of subversion 
that is achievable is limited. For this reason, hackers 
rarely enter into VR to hack. Hacking into the external
system running a simulspace is just like breaking
into any other system. Use all of the standard rules for 
intrusion and subversion.

\subsubsection{Meddling From The Inside}

Within a simulspace, a hacker's only choice for interacting
with the VR controls is through the standard
interface that any simulmorph can pull up. Typically 
used for standard user features like adjusting your 
simulmorph or chatting with or checking the status 

ers

ter-

ng 

or
%%% 266
of other users, a clever hacker might find some ways 
to subvert the system. Such options are usually limited
however, as a number of system controls and
processes cannot be accessed and manipulated from 
the inside.
Most of the hacker's options are going to involve 
meddling with the simulation and its specific domain 
rules or possibly gaining access to cheats. To make a 
change requires a successful Interface Test. Ultimately 
the gamemaster decides what the hacker can and 
cannot get away with, based on the limitations of that 
particular simulspace.
Most simulspaces are monitored to prevent cheating
and abuse, though the monitors are typically preoccupied
with maintaining the simulspace as a whole,
dealing with other users, etc. At the gamemaster's 
discretion, such a monitor might get to make an Interface
Test (possibly with a modifier for distraction) to
notice the hacker's efforts.

\section{Ai}


\section{ And Muses}

AIs are sentient but specialized programs. Like other 
software, they must be run on a computerized system. 
Most AIs are run on bots, vehicles, and other computerized
devices where they can assist transhuman
users or operate the machine themselves. They are also 
commonly used to actively monitor computer systems 
against intrusion attempts. \textit{Muses} are AIs that specialize
as personal companions, always at a character's
virtual side every since they were a child.
Sample AIs and muses can be found on p. 331 of \textit{Gear.}

\subsection{Ai Limitations}

AIs feature a number of built-in restrictions and 
limitations. To start with, they can be loaded in the 
cyberbrains of pods and synthmorphs, but they may 
not be downloaded into biomorph brains. As software, 
they use the same rules as other software and may be 
shut down, restarted, copied, erased, stored as inert 
data, infected with virii, and reprogrammed. Due to 
their size and complexity, only one AI (or infomorph) 
may be run on a personal computer at a time (see 
\textit{Computer Capabilities,} p. 247), and they may not run 
on peripheral devices.
While they possess cognition and intelligence, they 
are incapable of self-improvement and cannot expand 
their programming and skills on their own. Although 
–0

Analyze simulation parameters, view domain rules, shape appearance of simulmorph, switch simulmorph character or morph type
–10

Change probability of test outcomes, become invisible (``out-game'') to others
–20

Interfere with simulation (e.g. make it rain, generate earthquakes), generate items, ignore domain rules, 

kill or lockout other simulmorphs
–30

Go into god mode, command simulated characters, take over the simulation
they are not able to learn they do possess memory 
storage that grants them the ability to remember and 
a limited form of adaptation. AIs do not earn Rez 
Points, nor do they have Moxie.

AIs have aptitudes no greater than 20 but are 
incapable of defaulting. If they don't possess a skill, 
they don't know how to do it. (At the gamemaster's 
discretion, they may default to field skills or similar 
skills as noted on p. 173 with a –10 to –30 modifier). 
They can use skills like any character in \textit{Eclipse Phase}, 
however they may not possess any Active skill at a 
rating higher than 40 or Knowledge skill higher than 
90—the maximum amount of expertise that their skill 
software allows.

While AIs are programmed with personality templates
and empathy, they are generally less emotive
and difficult to read (apply a –30 modifier to Kinesics 
Test made against them, when in pod bodies). When 
combined with non-expressive synthetic morphs, they 
are even more difficult (–60 modifier). Some AIs lack 
emotive capabilities altogether and are impossible to 
read with Kinesics skill.

AIs do have a Lucidity and Trauma Threshold stat, 
and are capable of suffering mental stress and traumas.

\subsection{Commanding Ai}

AIs and muses are programmed to accept commands 
from authorized users. In some circumstances, they 
may also be programmed to follow the law or some 
ethical code. Programming is never perfect, however, 
and AIs can be quite clever in how they interpret 
commands and act on them. In most cases, an AI will 
rarely refuse to follow a request or obey a command. 
Given that they also usually have a duty to protect the 
person commanding them, the AI may be reluctant to 
follow commands that could be construed as dangerous
or having a negative impact on the user. Under
certain circumstances, preprogrammed imperatives 
can force an AI to ignore or disobey their owner's 
commands (gamemaster's discretion).

\section{Agi}


\section{ And Infomorphs}

The term ``infomorph'' is used to refer to any ego in 
digital body, whether that be an AGI or the digital 
emulation of a biological mind (including backups 
and forks). The following rules apply to infomorph 
and AGI characters.
%%% 267

\subsection{Software Minds}

At their core, infomorphs are just programs and so 
they are treated like other software in terms of rules. 
They must be run on a specific personal computer or 
server (see \textit{Computer Capabilities,} p. 247). If that 
device is shut down, the infomorph also shuts down 
into a state of unconsciousness, restarting along with 
the device (infomorphs may also shut themselves 
down, though it is rare that they do so). If the device is 
destroyed, the infomorph is killed along with it (unless 
their data can somehow be extracted from any surviving
components, perhaps resulting in a \textit{vapor,} p. 274).
Infomorphs may copy themselves, though in some 
places this is illegal and in most places is frowned 
upon as it raises numerous ethical and legal questions. 
For this reason many infomorphs that copy and transfer
themself to run on a new device will thoroughly
erase themselves off the old one.

As digital beings, infomorphs have no physical 
mind, but it is a simple matter for them to possess 
an uninhabited synthmorph, taking up residence in 
the cyberbrain (see \textit{Resleeving Synthmorphs,} p. 271). 
They may also download into biomorph bodies according
to standard resleeving rules (p. 271). Even
when disembodied, they may interact with the 
physical world via the mesh, viewing through sensors, 
streaming XP feeds, communicating with characters, 
commanding slaved devices, and teleoperating/jamming
drones.
Infomorphs have a Speed of 3, reflecting their digital
nature and their ability to act at electronic speeds.
If an infomorph sleeves into a body, however, it takes 
on the Speed of that morph.

\subsection{Agi Characters}

Though AGIs were not born in a biological body, their 
programming encompasses the full spectrum of human 
personality, outlook, emotions, and mental states. AGIs 
are in fact raised in a manner similar to human children
so that they are socialized much like humans are.
Nevertheless, on a fundamental level they are non-humans
programmed to act human. There are inevitably
points where the programming does not mask or alter 
the fact that AGIs often possess or develop personality 
traits and idiosyncrasies that are quite different from 
human norms and often outright alien.
Unlike standard crippled AIs, AGIs are capable of 
full-fledged creativity, learning, and self-improvement 
(at a slow but steady pace equivalent to humans). Just 
like other characters, they earn Rez points and may 
improve their skills and capabilities. AGIs suffer none 
of the skill limitations placed on weak AIs, using skills 
just like any other character.
On an emotional level, AGIs run emotional subroutines
that are comparable to biological human emotions
AGIs are, in fact, programmed to have empathy
and share an interest in human affairs and prosperity, 
and to place significant relevance on life of all kinds. 
In game terms, AGIs emote like humans (and so Kinesics
may be used against them) and are vulnerable to
emotionally manipulative effects, fear, etc.

ne 

lls 

u-

o-
