%%% 296
%%% 297
%%% 298
The accelerated technological levels of \textit{Eclipse Phase}
enable a number of devices for personal enhancement, 
survival, and other uses.

\section{Equipment Rules}

The following rules apply to all technological items in 
\textit{Eclipse Phase.}

\subsection{Acquiring Gear}

During character creation, players purchase gear for 
their characters using the credits they have during 
the character creation process. Once play begins, 
however, characters must obtain any equipment they 
need the usual way: by buying, borrowing, making, 
or stealing it.

In the inner system, hypercorp, and Jovian Republic 
settlements—and other places where capitalism still 
reigns—gear acquisition is simply a matter of finding 
a seller and buying it. Each item has a listed cost, from 
Trivial to Expensive, as noted on the Gear Costs table. 
Due to local availability of resources, supply and 
demand, and legalities, these listed costs are meant 
to be approximations. When no other factors apply, 
the listed Average Cost for that category can be used. 
Otherwise the gamemaster should modify the item's 
worth as they see fit, according to local economic factors
while still keeping it within that cost category
range. The Cost Modifiers table lists out some suggested
changes to an item's cost, but these are simply
recommendations, and can be ignored or followed as 
the gamemaster deems fit. The exact local conditions 
are largely up to the gamemaster to determine, as best 
fits their game.

In some circumstances, characters may attempt to 
haggle over gear prices. This is best handled as roleplaying
but the gamemaster may also call for an
Opposed Persuasion Test (or possibly an Intimidation 
Test). The character who wins may increase or reduce 
the price by 10% per 10 points of MoS.

In the outer system, anarchist, Titanian, scum, and 
other habitats that use the reputation economy, characters
must rely on their rep scores to acquire the goods
and services they need. The mechanics for this are covered
under \textit{Reputation and Social Networks,} p. 285.

Characters are of course free to get their hands 
on equipment by any other means they devise—con 
schemes, borrowing from friends, and outright 
theft, with all of the appropriate tests and consequences
In some cases, acquiring gear may be an
adventure unto itself.

\subsubsection{Fabricating Gear}

Thanks to nanofabrication technology, characters may 
also create their own equipment using cornucopia 
machines and similar nanofab devices (p. 327). The 
character must have the appropriate blueprints to do 
so, whether they come with the fabber, are bought 
legitimately or on the black market, acquired with rep, 
or found online. Characters may also code their own 
blueprint desires, using the Programming: Nanofabrication
skill.

\subsection{Gear Modifiers}

In the technological future, gear is a necessity. In many 
cases, use of equipment provides no bonuses, it simply 
allows a character to perform a task they would otherwise
be unable to do. For example, it is impossible
to pick a mechanical lock without lockpick or some 
sort of tool.
In other cases, however, gear provides a bonus to 
the task at hand. Climbing a wall may be possible 
without tools, but if you happen to have gecko gloves 
or other climbing gear, it's going to be a lot easier. The 
specific modifier applied is usually noted in the gear 
item's description, typically ranging from +10 to +30.

\subsubsection{Gear Quality}

In both of the situations above, it is possible to have 
items that are of either exceptional or inferior quality, 
with corresponding positive or negative modifiers. The 
gear may be well-crafted, state-of-the-art, cutting-edge 
experimental, or simply top-of-the-line, applying an 
additional +10 to +30. Or it may be outdated, shoddy, 
or in disrepair, inflicting a –10 to –30 modifier  (in 
some cases canceling out the basic gear bonus).
Trivial

1–99

50
Low

100–499

250
Moderate

500–1,499

1,000
High

1,500–9,999

5,000
Expensive

10,000+

20,000
Item Stolen

–50%
Item Used

–25%
Item Restricted

+25%
Item Illegal

+50%
Item Scarce

+25%
Item Extremely Rare

+50%
Item Common

–25%
%%% 299

\subsubsection{Gear Sizes}

On occasion, you'll need to know how small or large 
a certain piece of equipment is. Though this is largely 
something the gamemaster can wing on the fly using 
common sense, we've listed sizes for many gear items 
that are unusual or so futuristic that the average player 
may not have a feel for what dimensions the tech 
likely is. These size categories are listed on the Gear 
Sizes table (p. 297). These sizes should be considered 
approximations, as depending on the manufacturer 
and process, some items may be smaller or larger than 
similar items. It is also important to keep in mind that 
as technology advances, the size and components of 
various equipment items shrink, so when in doubt, go 
with smaller.

\subsubsection{Mass And Encumbrance}

A character who is carrying too much gear should 
be slowed down, suffering negative modifiers  both 
to their movement rates and their skill tests. Rather 
than micromanaging the weights of individual pieces 
of equipment, however, this matter is largely left to 
the gamemaster's discretion, using common sense. If a 
character loads up beyond reason, apply modifiers as 
seem appropriate. The gamemaster should, however, 
keep in mind that many of the manufacturing materials
used in \textit{Eclipse Phase} allow for items that are
much lighter than current standards without any loss 
of durability or function (see \textit{Future Materials,} p. 298). 
Likewise, characters in low or microgravity environments
can carry much larger loads.

\subsubsection{Concealing Gear}

Characters may attempt to conceal items on their 
person, hoping at least to hide them from casual 
notice if not an intensive search. To determine how 
Nano

So small that the item cannot be seen without 

the aid of a microscope or nanoscopic vision 

(p. 311), and may not be manipulated without 

fractal digits (p. 311) or similar tools.
Micro

Anything ranging from the size of a barely 

visible small dot to an average insect.
Mini

Mini items may be concealed within some-

one's palm or small pockets.
Small

Small items may be held in one hand and 

concealed in normal pockets.
Medium

Medium size items are cumbersome to hold 

with one hand, ranging from the size of a 2-li-

ter bottle to the size of a medium dog. They do 

not fit in pockets, but they may be concealed 

by larger coverings.
Large

Roughly human-sized.
Huge

Vehicles and other more massive objects.
effectively the character conceals the equipment, make 
a Palming Test and note the MoS (the gamemaster 
may wish to roll this secretly). Whenever another 
character has a chance to notice the concealed item, 
they must succeed in a Perception Test and achieve 
a higher MoS than was scored on the Palming Test. 
The gamemaster should apply modifiers to both tests 
as appropriate. For example, concealing a large item 
like a sword would be difficult (–30), whereas wearing 
concealing clothing like a longcoat or multi-pocketed 
jumpsuit would help (+20). Likewise, a character 
who is not actively looking is less likely to notice the 
hidden gear (–30), whereas someone who conducts a 
physical search (+30) or who has enhanced vision to 
pierce protective layers will fare better.

\subsection{Design And Fashion}

Many objects in \textit{Eclipse Phase} closely resemble their 
early 21st century equivalents—a bottle of soda is still 
a transparent container holding a brightly colored 
liquid, clothing is obviously something you wear, and 
a knife still consists of a blade and a handle. The materials
processes, and mindsets that go into making
them, however, are quite different. To start, very few 
items look have a uniform, mass-produced look, even 
if they were. The procedures of minifacturing and 
nanofabrication allow every individual item to be 
manufactured with a unique (or at least different) 
look. In areas with anarchist/reputation economies, in 
fact, where personal possessions have very little intrinsic
value, expression and creativity are favored and so
many items are artistically personalized (and actual 
hand-crafted items are rare and prized). Likewise, 
almost all equipment is designed with ergonomics 
and ease-of-use prioritized, so gear with soft curves, 
pleasing colors, and form-fitting shapes are common. 
Many items of personal technology, such as flashlights 
or small tools, are made in the form of ovoids that 
fit comfortably in the user's hand or in similar forms 
that can be easily worn or attached to clothing.

To 
someone from the 20th century, many common 
devices look like oddly colored rocks or decorative 
pieces of plastic or ceramic (in fact, many such items 
are referred to as ``blobjects'' by older transhumans).
The materials used to create everyday items are 
also advanced, ranging from aerogel and graphene 
to smart materials (p. 298) and exotic metamaterials
with unusual physical properties. In practice, this
means that most items are light, durable (with both 
tensile strength and/or flexibility, as needed), waterproof
dirt-repellent, and self-cleaning. Most gear is
also designed with zero-G or microgravity functionality
in mind, and can easily be clipped, tethered, or
stuck to a surface with grip pads.
Almost all gear available in \textit{Eclipse Phase} is also 
available in forms that are wearable/usable by uplifted
animals and non-humanoid morphs, such
as novacrabs, slitheroids, and so on. Even if such 
customized gear is not immediately available, it is 
%%% 300
usually not difficult to nanofabricate. Smart materials 
(p. 298) also make interoperability between different 
morphs easy.

\subsubsection{Interface}

It is not uncommon for everyday devices to have no 
visible controls as they are designed to be operated via 
radio broadcasts from the user's ecto or mesh inserts. 
Any items crafted for use in emergency, combat, survival
or exploration situations, however, will feature
basic physical controls, just in case. Physical interfaces 
are typically controlled by touch pads that are nothing
more than colored spots on the device's surface,
though some may also project a holographic interface 
display. Most equipment of this sort can can also be 
voice-activated and controlled.

Almost all devices are loaded with a complete set 
of help files and tutorials. Most electronics are also 
mesh-capable and equipped with specialized AIs (see 
\textit{Meshed Gear}, next page).

\subsubsection{Smart Materials}

Many common items of technology are made from 
so-called smart materials. These devices contain—or 
sometimes consist entirely of—many small nanomachines
that can both move and reshape themselves
to alter the object's shape, color, and texture. For 
example, smart clothing can transform from a suit of 
specialized cold weather clothing suitable for the Martian
poles in winter to a fashionable suit in the latest
style due to hundreds of thousands of tiny nanomachines
in the clothing that shift and move to reshape
the garment. Similarly, a tool made of smart materials 
can switch from a powered screwdriver to a wrench 
or a hammer, as the nanomachines move around and 
completely reshape the tool. Smart materials all contain
specialized advanced nanomachine generators (p.
328) that keep them in perfect repair as long as they 
are regularly recharged.
%%% 301

\section{Meshed Gear}

Almost all technology in \textit{Eclipse Phase} is designed 
to be operated via radio signals from the user's basic 
implant, although models usable by characters without
basic implants are also available. In addition all
devices contain a nearly microscopic computer and 
radio link (known as a ``voice'') that allows the user to 
easily locate the object and that reports on the condition
of the object or device, how to properly use and
care for it, as well as telling the user when it needs 
to be repaired and how. Most are discrete and highly 
useful, but cheaply made goods sometimes have overly 
annoying voices.
This means that almost all devices can be accessed 
via the mesh or directly if within radio range. This 
makes them vulnerable to hacking and intrusion attempts
(p. 254) as well as radio jamming (p. 262).
Many devices are, however, publicly accessible (see 
\textit{Spimes,} p. 238). Meshed gear may also be tracked 
through the mesh (p. 251). For privacy and security, 
these devices are often slaved to other systems (see 
\textit{Slaving Devices,} p. 248); devices worn/carried by 
Nano

20 meters
Micro

50 meters
Mini

1 kilometer
Small

5 kilometers
Medium

25 kilometers
Large

500 kilometers
characters are usually made part of the personal area 
network and slaved to the character's mesh inserts/
ecto. For more info on meshed devices, see the \textit{Mesh}
chapter, p. 234.

Many devices come equipped with AIs, who are 
equipped with skillsofts that enable them to operate
the device on their own, as according to voiced
instructions or commands issued through the net. AIs 
are described on p. 264 and p. 331.

\subsection{Radio And Sensor Ranges}

In \textit{Eclipse Phase,} almost all devices are equipped with 
small radios so that they may be meshed. Likewise, 
many pieces of gear are equipped with sensors such as 
cameras, microphones, or other detectors. The Radio 
and Sensor Ranges table notes what range these devices
operate at.

\subsection{Power}

All of the powered devices in \textit{Eclipse Phase} require 
electricity to function. With rare exceptions, most of 
them rely on either solar cells or powerful batteries. 
These batteries are high-density, room-temperature 
100 meters

Smart Dust, Nanobot/Microbot Swarms
500 meters

Microbugs
20 kilometers

Mesh Inserts
50 kilometers

Ectos, Miniature Radio Farcasters, Portable Sensors
250 kilometers

Radio Boosters, Vehicle Sensors
5,000 kilometers 
%%% 302
superconductors with 25 times the capacity of the 
best batteries in common use in the early 21st century. 
Such batteries may also be constructed so that they 
are flexible, printed on devices, or woven into fabric. 
They are good for 100–500 hours of use, and will alert 
the user when they start running low. More powerful
radio-isotope nuclear batteries are also available,
heavily shielded so they do not emit radiation and 
good for 3 years or more of use.

In short, power should rarely be an issue in \textit{Eclipse }
\textit{Phase} games, unless it happens to fit the plot. Power 
failure could also result from a critical failure roll.

\section{Personal Augmentation}

Almost all citizens of the solar system, whether human, 
AI, or uplifted animal, use various forms of biological, 
cybernetic, or nanotechnological augmentation. The 
following is a list of the most common types.

Unless otherwise noted, any bonuses from personal 
augmentations are both compatible and cumulative 
with bonuses from other enhancements.

\subsection{Standard Augmentations}

Most morphs produced in the solar system include the 
following augmentations.

\subsubsection{Basic Biomods}

Almost universal in biomorphs, many habitats will not 
allow individuals to visit/immigrate if their biomorph 
does not possess these biomods in order to preserve 
public health. Basic biomods consists of a series of 
genetic tweaks, tailored virii, and bacteria that speed 
healing, greatly increase disease resistance, and impede 
aging. A morph with basic biomods heals twice as fast 
as an early 21st century human, gradually regrows lost 
body parts, is immune to all normal diseases (from 
cancer to the flu), and is largely immune to aging. In 
addition, the morph requires no more than 3-4 hours 
of sleep per night, is immune to ill-effects from long-term
exposure to low or zero gravity, and does not
naturally suffer from biological problems like depression
shock reactions after being injured, or allergies.
\textbf{[Moderate, but included for free in most biomorphs]}

\subsubsection{Basic Mesh Inserts}

Mesh inserts are ubiquitous among modern morphs. 
This network of cybernetic brain implants is essential 
equipment for anyone who wants to stay connected 
and make full use of the wireless mesh. The interconnected
components of this system include:
•  \textbf{Cranial Computer:} This computer serves as the 

hub for the character's personal area network and 

is home to their muse (p. 264). It has all of the 

functions of a smartphone and PDA, acting as a 

media player, meshbrowser, alarm clock/calendar, 

positioning and map system, address book, ad-

vanced calculator, file storage system, search engine, 

social networking client, messaging program, and 

note pad. It manages the user's augmented real-

ity input and can run any software the character 

desires (see \textit{Software,} p. 331). It also processes XP 

data, allowing the user to experience other people's 

recorded memories, and also allowing the user to 

share their own XP sensory input with others in 

real-time. Facial/image recognition and encryption 

software (p. 331) are included by default.
•  \textbf{Radio Transceiver:} This transceiver connects the 

user to the mesh and other characters/devices 

within range. It has an effective range of 20 kilome-

ters in deep space or other locations far from radio 

interference and 1 kilometer in crowded habitats.
•  \textbf{Medical Sensors:} This array of implants monitors 

the user's medical status, including heart rate, 

respiration, blood pressure, temperature, neural 

activity, and much more. A sophisticated medical 

diagnostic system interprets the data and warns 

the user of any concerns or dangers.
Using any of these functions is as easy as thinking. 
\textbf{[Moderate, but included for free in most morphs]}

\subsubsection{Cortical Stack}

A cortical stack is a tiny cyberware data storage unit 
protected within a synthdiamond case the size of a 
grape, implanted at the base of the skull where the 
brain stem and spinal cord connect. It contains a digital
backup of that character's ego. Part nanoware, the
implant maintains a network of nanobots that monitor
synaptic connections and brain architecture, noting
any changes and updating the ego backup in real time, 
right up to the moment of death. If the character dies, 
the cortical stack can be recovered and they may be 
restored from the backup (see \textit{Resleeving,} p. 271). 
Cortical stacks do not have external or wireless access 
(for security), they must be surgically removed (see \textit{Re-}
\textit{trieving a Cortical Stack,} p. 268). Cortical stacks are 
extremely durable, requiring special effort to damage 
or destroy. They are commonly recovered from bodies 
that have otherwise been pulped or mangled. Cortical 
stacks are intentionally isolated from mesh inserts and 
other implants, as a security measure to prevent hacking
or external tampering. \textbf{[Moderate, but included for }
\textbf{free with most morphs]}

\subsubsection{Cyberbrain}

Cybernetic brains are where the ego (or controlling 
AI) resides in synthmorphs and pods. Modeled on biological
brains, cyberbrains have a holistic architecture
and serve as the command node and central processing
point for sensory input and decision-making. Only
one ego or AI may ``inhabit'' a cyberbrain at a time; to 
accommodate extras, mesh inserts (p. 300) or a ghost-rider
module (p. 307) must be used. Since cyberbrains
store memories digitally, they have the equivalent of 
mnemonic augmentation (p. 307). They also have 
a built-in puppet sock (p. 307) may be remote-controlled
though this option may be removed by those
who value their security. Cyberbrains are vulnerable 
%%% 303
to brainhacking (p. 261) and other forms of electronic 
infiltration/attack. Cyberbrains come equipped with 
two or more pairs of external access jacks (p. 306), 
usually located at the base of the skull, which allow 
for direct wired connections. \textbf{[Moderate, but included }
\textbf{for free in all synthetic morphs and pods]}

\subsection{Bioware}

Bioware augmentations can be acquired either as a genemod
when the morph is designed and grown or as a
later modification to an existing morph, either by using 
nanomachines to modify the morph's tissue or by externally
growing the organ and implanting it. Bioware
may be used to enhance biomorphs (including pods and 
uplifts), but not synthmorphs. Bioware may be used to 
enhance biomorphs (including pods and uplifts), but not 
synthmorphs (see Synthmorphs and Bioware, p. 306).

\subsubsection{Enhanced Senses}

The following are a list of the most common enhanced 
senses. Each is also available as a cybernetic implant, 
but bioware is much more common.

\textbf{Direction Sense:} The character has an innate sense 
of direction and distance using advanced inertial 
navigation. The character can arbitrarily define any 
point as ``north'' and keep track of which direction 
that is, as well as knowing approximately how far 
they have come. Characters with this augmentation 
can always retrace any route they have taken, only 
experiencing difficulty with three-dimensional routes 
lacking navigational markers (such as deep space or 
undersea; apply a –30 modifier). Since positioning 
inside habitats by anyone with basic mesh inserts is an 
automatic affair, only characters venturing to remote 
locations require this augmentation. \textbf{[Low]}

\textbf{Echolocation:} The character possesses sonar similar 
to that of a bat or dolphin. The character bounces 
brief ultrasonic pulses off their surroundings and uses 
them to form an image of these surroundings through 
the pattern of reflections of these pulses received by 
the character's ears. For more details, see \textit{Using En-}
\textit{hanced Senses,} p. 302. This augmentation works in 
both air and water and has a range of 20 meters in air 
and 100 meters in water. \textbf{[Low]}

\textbf{Enhanced Hearing:} The morph's ears are enhanced 
to hear both higher and lower frequency sounds—the 
range of sounds they can hear is twice that of normal 
human ears (see \textit{Using Enhanced Senses,} p. 302). In 
addition, their hearing is considerably more sensitive, 
allowing them to hear sounds as if they were five 
times closer than they are. A character with this augmentation
can easily overhear even a softly spoken
conversation at another table in a small restaurant. 
This augmentation provides a +20 modifier to all 
Perception Tests involving hearing. \textbf{[Low]}

\textbf{Enhanced Smell:} The morph's sense of smell is 
equal to that of a bloodhound. The user can identify
both chemicals and individuals by smell, and can
track people and chemically reactive objects by smell 
as long as the trail was made within the last several 
hours and has not been obscured. The character can 
also gain a general sense of the emotions and health 
of any character within 5 meters (+20 to Perception or 
Kinesics Tests to do so). \textbf{[Low]}

\textbf{Enhanced Vision:} The morph's eyes have tetrachromatic
vision capable of exceptional color differentiation
These eyes can also see the electromagnetic
spectrum from terahertz wave frequencies to gamma 
rays, enabling them to see a total of several dozen 
colors, instead of the seven ordinary human eyes can 
perceive. In addition, these eyes have a variable focus 
equivalent to 5 power magnifiers or binoculars. This 
augmentation provides a +20 modifier to all Perception
Tests involving vision. For further applications,
see \textit{Using Enhances Senses,} p. 302. \textbf{[Low]}

\subsubsection{Mental Augmentations}

Mental augmentations are extremely common.

\textbf{Eidetic Memory:} The character can remember everything
that ever happened to them, in detail, with
no long term memory loss. For example, they can 
recite a page they read in a book a month ago, recall 
a string of 200 random characters they viewed a year 
ago, or even tell you what they had for breakfast on 
a particular date a decade ago. However, they can 
only remember things they paid attention to. The 
character will not remember the contents of a note 
on someone's desk if they merely glanced at it; they 
must specifically have read it. No effort is required to 
use this augmentation, the character merely needs to 
attempt to remember a specific fact. \textbf{[Low]}

\textbf{Hyper Linguist: }The morph's brain maintains the 
linguistic flexibility of a small child, allowing the 
character to learn languages with great ease. This 
functions as the Hyper Linguist trait, p. 146. \textbf{[Low]}

\textbf{Math Boost:} This implants functions as the Math 
Wiz trait, p. 146. \textbf{[Low]}

\textbf{Multiple Personalities}: The character's brain is intentionally
partitioned to accommodate an extra personality
This multiplicity is not viewed as a disorder,
but as a cognitive tool to help people deal with their 
hypercomplex environments. This extra personality 
can be an NPC run by the gamemaster, a separate 
character (in ego form only) made by the player, or 
the downloaded fork of another character. For all intents
and purposes, the extra personality is treated as
a separate ego (i.e., it may fork separately), except that 
both personalities are backed up in the same cortical 
stack and if downloaded they must be placed in separate
morphs or in another morph with this implant.

Only one ego may be in control of the morph at a 
time. The other resides in the background, still active, 
but not on a surface level. Each ego is completely 
aware of what the other is doing, thinking, etc. If for 
some reason the subsumed personality wants to come 
to the fore, but the other personality won't relinquish 
control, make an Opposed WIL x 3 Test. Each ego has 
its own Lucidity and Trauma Threshold, and they track 
stress and trauma separately. Any psi attacks or social/
mental influences only affect the personality at the fore.
%%% 304

Having an extra ego in your head, working in the 
background, is helpful for multitasking. The character 
receives an extra Complex Action each turn that may 
only be used for mental or mesh actions. \textbf{[High]}

\subsubsection{Physical Augmentations}

Most physical bioware augmentations are derived 
from the capabilities of animals.

\textbf{Adrenal Boost:} This adrenal gland enhancement 
supercharges the character's adrenal response to 
situations that invoke stress, pain, or strong emotions 
(fear, anger, lust, hate). When activated, the concentrated
burst of norepinephrine accelerates heart rate
and blood flow and burns carbohydrates. In game 
terms, this allows the character to ignore the –10 
modifier from 1 wound and temporarily increases 
REF by +10 (also boosting REF-linked skills and 
Initiative). These modifiers apply until the character 
has calmed down (if the character also has endocrine 
control, p. 304, then adrenal boosts can be activated 
and deactivated at will, and the negated wounds are 
cumulative). \textbf{[High]}

\textbf{Bioweave Armor (Light):} Bioweave armor involves 
lacing the morph's skin with artificial spider silk biological
fibers. This provides an Armor rating of 2/3
without changing the appearance, texture, or sensitivity
of the morph's skin. This armor is cumulative with
worn armor. \textbf{[Low]}
%%% 305

\textbf{Bioweave Armor (Heavy):} Heavy bioweave armor 
involves lacing the morph's skin with a denser and 
thicker network of the same fibers. The  morph's 
skin becomes thicker and somewhat less flexible 
except at the joints. The morph's skin also has an 
unusually smooth look, and a distinctively smooth 
and tough-feeling texture. This provides an Armor 
rating of 3/4 without decreasing the morph's mobility
The character's sense of touch, however, is
significantly reduced (–20 modifier) except on their 
hands, feet, and face. This armor is cumulative with 
worn armor. \textbf{[Moderate]}

\textbf{Carapace Armor:} Carapace armor combines 
bioweave armor with hard but flexible plates of 
a chitin-ceramic hybrid material modeled on the 
microscopic structure and texture of arthropod exoskeletons
This armor is obvious and has a somewhat
crocodilian or insectoid appearance (character's 
choice). The morph is completely hairless as well. This 
provides an Armor rating of 11/11. This armor is not 
cumulative with worn armor. \textbf{[Moderate]}

\textbf{Chameleon Skin: }The morph's skin is augmented 
with complex chromatophores so that it changes color 
like the skin of a chameleon or an octopus. The morph 
can match the appearance of almost any color and 
most patterns. This provides a +20 modifier to Infiltration
Tests to avoid being seen or noticed, as long as
the character is stationary or not moving faster than 
%%% 306
a slow walk. The character must be nude or wearing 
smart clothing (p. 325) of the same color/pattern. If 
incompletely camouflaged, or if moving faster, reduce 
the modifier to +10. In addition to blending in, the 
character can also consciously change the color and 
pattern of their skin to deliberately stand out (+20 on 
Perception Tests to notice) or simply to produce attractive
or interesting colors or patterns. \textbf{[Low]}

\textbf{Circadian Regulation:} The morph only requires 
2 hours of sleep to maintain health and function at 
peak mental capacity. The character dreams constantly 
while asleep and can both fall asleep and wake up 
almost instantly. In addition, the character can easily 
and with no ill-effects shift to a 2-day cycle, where they 
are awake for 44 hours and sleep for 4. \textbf{[Moderate]}

\textbf{Claws: }The morph has retractable claws like those 
of a cat. These claws do not interfere with the character's
manual dexterity and are razor sharp. However
they are relatively small and only do 1d10 + 1
+ (SOM ÷ 10) damage, with an AP of –1. As a result, 
they are legal in almost all habitats and are considered 
tools as much as weapons. \textbf{[Low]}

\textbf{Clean Metabolism: }The morph's symbiotic bacteria, 
gut flora, and glands have been genetically engineered 
to keep the morph ``clean.'' The morph also produces 
smart antibiotics that prevent the growth of any 
bacteria or yeasts in it or on its skin. As a result, the 
morph is completely immune to infections, dental cavities
and bad breath, its sweat has no scent, and the
morph's efficient digestion produces somewhat less 
solid waste and less odorous chemicals. \textbf{[Moderate]}

\textbf{Drug Glands: }The morph has specially-tailored 
glands designed to produce specific hormones or 
chemicals and release them in the body. The character
has control over these glands and can release the
chemicals at will. Each type of drug gland is considered
a separate enhancement. For potential drugs and
chemicals, see p. 317. \textbf{[One Cost Category Higher }
\textbf{Than Drug Cost]}

\textbf{Eelware:} Derived from electric eel genetics, a character
can have eelware implanted so that it connects
to a network of bioconductors in the hands and feet 
(or other limbs), allowing the character to generate 
stunning shocks with a touch. Eelware inflicts shock 
damage (p. 204) exactly like a pair of shock gloves. 
Eelware can also be used to power implants and specially
designed handheld devices by touch. \textbf{[Low]}

\textbf{Emotional Dampers: }This low-cost alternative to 
endocrine control (p. 304) allows the user to voluntarily
damp their morph's emotional responses
and various non-verbal cues like pupil dilation, eye 
movement, or vocal tone. Using this augmentation 
allows the user to lie and conceal their emotions in 
such as way as oo fool the keenest observer; apply a 
+30 modifier to Deception and Impersonation Tests. 
This modification does not affect methods of detecting 
lies and emotions that involve reading the character's 
neural state, including psi-gamma sleights. However, 
this augmentation damps out all emotional responses 
and so causes the character to be less persuasive in real-time
personal interactions, imposing a –10 modifier
to other Social skill tests like Persuasion. Characters 
can turn this augmentation on or off at will. \textbf{[Low]}

\textbf{Endocrine Control:} This augmentation modifies the 
morph's endocrine system, giving the character fine 
control over their hormone output. This allows the 
character to completely control their appetite and 
emotions and to regulate pain. They receive a +30 
modifier against the effects of hunger, fear, and any 
forms of emotional manipulation, such as the Drive 
Emotion sleight. This augmentation also allows character
to lie with perfect conviction and to completely
fool all methods of lie detection that do not rely 
on the target's neural output; apply a +20 modifier 
to Deception Tests. It also allows the character to 
remain awake for 48 hours without penalty, but after 
this time the character begins experiencing normal 
fatigue. Finally, the ability to regulate pain reception 
%%% 307
allows the character to ignore the –10 modifier from 
1 wound. \textbf{[High]}

\textbf{Enhanced Pheromones:} The morph's biochemistry 
has been altered so that it produces enhanced pheromonal
signals that subconsciously affect the behavior
of other humans in the vicinity. These pheromones 
make the character more attractive and trustworthy to 
the target; apply a +10 modifier to appropriate Social 
skill tests, such as Persuasion. This augmentation only 
affects characters who can smell the pheromones, and 
it does not affect uplifts or xenomorphs. \textbf{[Low]}

\textbf{Enhanced Respiration:} By boosting both lung efficiency
and the blood's oxygen-carrying capacity, the
character can live comfortably in both high and low 
pressure environments, from 0.2 atmospheres to 5 
atmospheres, with no dizziness or need for gradual 
decompression. In addition, the character can hold 
their breath for up to 30 minutes when performing 
minimal activity or for up to 10 minutes while performing
highly strenuous activity. \textbf{[Low]}

\textbf{Gills: }The morph's lung tissue has been adapted to 
function as gills, allowing the morph to breathe both 
air and water, as long as the water is not toxic or too 
stagnant. Characters with this augmentation breathe 
in water and then expel the water through slits just 
underneath their lowest pair of ribs that seal when the 
character is not underwater. \textbf{[Low]}

\textbf{Grip Pads: }The morph possesses specialized pads 
on its palms, lower arms, shins, and the bottoms of 
its feet. Designed to emulate the pads on gecko feet, 
characters can support themselves on a wall or ceiling 
by placing any two of these pads against any surface 
not made from a material specially designed to resist 
this augmentation. Characters can climb any surface 
and move easily across ceilings that can support their 
weight. Apply a +30 modifier to Climbing Tests. The 
pads must be free to touch the surface the character is 
climbing (no gloves). The nature of these pads is obvious
to anyone looking at them, but they do not impair
the character's sense of touch or manual dexterity. If 
combined with the vacuum sealing augmentation, the 
character can even stick to surfaces in the vacuum of 
space. \textbf{[Low]}

\textbf{Hibernation:} The character can voluntarily reduce 
the morph's metabolism to the point that the morph 
requires only 5% of the normal amount of food, 
water, and air. The character appears to sink into a 
deep sleep, but can maintain a dim awareness of both 
touch and sound and so can be easily awakened. Entering
or leaving this state requires 3 minutes where
the character is relatively helpless. With sufficient 
air, characters can safely hibernate for up to 40 days 
without food or water. \textbf{[Low]}

\textbf{Muscle Augmentation:} The morph's muscle mass 
has been enhanced and toned and myofibers strengthened
Apply a +5 modifier to SOM. \textbf{[High]}

\textbf{Neurachem:} This bioware modification enhances 
the character's chemical synapses and juices their 
neurotransmitters, drastically speeding up neural 
connections. Neurachem can be mentally activated or 
triggered by charged emotions. Level 1 neurachem increases
the character's Speed stat by +1, with no side
effect. Level 2 raises the Speed stat by +2, but each 
time it is used the character suffers a nervous system 
fatigue hangover for 1 hour after the boost wears off 
(apply a –20 modifier to all actions). \textbf{[High (Level 1), }
\textbf{Expensive (Level 2)]}

\textbf{Poison Gland:} Similar to the drug gland, this 
morph has special glands that produce poisons, 
like the venom glands of a snake. The morph has 
poison glands in its fingers and mouth, so that it can 
deliver either poison by scratching someone with a 
fingernail, biting them hard enough to draw blood, 
or even by sharing a beverage with someone or spitting
into their drink. The morph is immune to the
poisons it produces. These glands may not produce 
nanotoxins. \textbf{[Low]}

\textbf{Prehensile Feet: }The morph's feet and leg joints are 
altered so that its toes are longer and more dexterous
and the big toe is transformed into an opposable
thumb. Physically, the morph's feet resemble a longer 
narrower hand or a human foot with finger  (and 
thumb)-like toes. The character can walk normally 
but must wear specially designed shoes. However, this 
morph runs somewhat slower than a morph with unmodified
feet (–1 meter per Action Turn). In addition,
the morph's hips are slightly modified to allow greater 
mobility. In a properly constructed chair, or when 
floating in zero-G, the character can use both their 
hands and their feet to manipulate the same object. 
Most morphs used by characters who live in zero-G 
possess this augmentation. \textbf{[Low]}

\textbf{Prehensile Tail:} A long (1.5 meters) prehensile 
tail is added to the morph's backside, extending out 
from the tailbone. This tail is prehensile and may 
be used to grab, hold, and even manipulate objects. 
The character can control the tail's movements with 
concentration, but it otherwise tends to move on its 
own. The tail also improves the character's balance; 
apply a +10 to any Physical skill tests where balance 
is a factor. \textbf{[Low]}

\textbf{Sex Switch:} A complex suite of alterations allows 
the character to switch their physical sex to male, 
female, hermaphrodite, or neuter. This change is 
mentally triggered but takes approximately 1 week to 
complete. \textbf{[Moderate]}

\textbf{Skin Pocket: }The morph has a pocket within its skin 
layer, capable of holding and providing concealment 
(+30) for small items. \textbf{[Trivial]}

\textbf{Temperature Tolerance: }The morph's temperature 
regulation and circulation are both substantially enhanced
allowing the character to survive in temperatures
as low as –30 degrees Celsius and as high as 60
degrees Celsius without discomfort or ill effects. \textbf{[Low]}

\textbf{Toxin Filters: }The morph gains an improved liver 
and kidneys and biological filters in its lungs. Characters
with this augmentation are immune to all chemical
and biological toxins, including everything from
recreational chemicals to nerve agents to spoiled food. 
In addition, the character can safely and comfortably 
%%% 308
breathe smoke and drink salt water. Unlike medichines
toxin immunity prevents the character from experiencing
even brief harm or discomfort from a toxin
(medichines merely rapidly repair damage caused by 
the toxin and then remove it from the morph). This 
augmentation provides no resistance to concentrated 
acid, nanotechnological attacks, or similar destructive
agents. Some characters with this augmentation
learn to enjoy the taste of various chemical toxins like 
cyanide or arsenic. \textbf{[Moderate]}

\textbf{Vacuum Sealing:} To possess this augmentation, the 
character must also possess some form of bioware 
armor or carapace armor. The morph has been specially
designed to survive the effects of vacuum. The
character's skin resists vacuum as well as protecting 
the wearer from temperatures from –75 to 100 C. 
In addition, the character's mouth, nose, and other 
orifices can seal sufficiently well to resist vacuum, 
and the morph possesses a special membrane that 
extends over their eyes, allowing the character to 
see in vacuum without risking any eye damage. This 
augmentation is usually combined with either the enhanced
respiration or oxygen storage augmentation,
or both together. \textbf{[High]}

\subsection{Cyberware}

Very little cyberware is physically implanted. Instead, 
the morph is placed in a healing vat (p. 326) and the 
vat's nanobots construct the cyberware inside the biomorph's
body. Cyberware is rarely used for anything
that can be accomplished using bioware.

Synthmorphs and bots may also also use cyberware.

\subsubsection{Enhanced Senses}

In addition to being able to duplicate the affects of all 
bioware enhanced senses, there are a few enhanced 
senses that can only be produced using cyberware.

\textbf{Anti-Glare:} This visual mod eliminates penalties for 
glare. \textbf{[Low]}

\textbf{Electrical Sense:} The character can sense electric 
fields. Within 5 meters, the character can instantly 
tell if an electrical device is on or off and can see the 
precise location of electrical wiring behind a wall or 
inside a device. This sense gives the character a +10 
modifier on any test involving analyzing, repairing, or 
modifying electrical equipment. \textbf{[Low]}

\textbf{Radiation sense:} The character can sense the presence
and approximate source of all forms of dangerous
radiation, including neutrons, charged particles,
and cosmic rays. \textbf{[Low]}

\textbf{T-Ray Emitter:} Mounted under the skin of the 
user's forehead, this implant generates low-powered 
beams of terahertz radiation (T-rays) that allow the 
character to see using reflected T-rays. As discussed in 
\textit{Using Enhanced Senses, }p. 302, this implant combined 
with the enhanced vision enhancement (or a terahertz 
sensor) allows the user to effectively see through cloth, 
plastic, wood, masonry, composites, and ceramics as 
well as being able to determine the composition of 
various materials. This implant allows the user to see 
using reflected T-rays for 20 meters in a normal atmosphere
and for 100 meters in vacuum. \textbf{[Low]}

\subsubsection{Mental Augmentation}

These cybernetic augmentations enhance the brain 
and mental functions.

\textbf{Access Jacks:} Usually located in the base of the skull 
or neck, this implant is an external socket with a direct 
neural interface. It allows the character to establish a 
direct wired connection using a fiberoptic cable to 
external devices or other characters, which can be 
useful in places where wireless links are unreliable or 
complete privacy is required. Two characters linked 
via access jack can ``speak'' mind-to-mind and transfer 
information between their mesh inserts and other implants
All synthmorphs have these by default. \textbf{[Low]}

\textbf{Dead Switch: }This cortical stack (p. 300) accessory is 
designed to keep the stack from falling into the wrong 
hands. If the morph is killed, the dead switch wipes 
and melts the cortical stack completely, so that the ego 
cannot be recovered. This option is generally only used 
by covert operatives with recent backups. \textbf{[Low]}

\textbf{Emergency Farcaster: }Only characters with cortical
stacks can possess this augmentation. The morph
has an implanted quantum farcaster (p. 314) linked 
to a highly secure storage facility. The high cost of 
this implant also covers the cost of this storage. Using 
standard radio and quantum encryption, the farcaster 
broadcasts full backups of the character's ego (pulled 
from the cortical stack) once every 48 hours. At the 
gamemaster's discretion, the backup interval may be 
scheduled more or less frequently, keeping in mind 
that ego broadcasts are generally limited for security 
%%% 309
purposes and because they hog bandwidth. These 
broadcasts only work when the character is in radio 
contact with the storage facility and is typically only 
used inside a habitat to broadcast backups back to a 
nearby space ship. If the radio broadcasts are blocked 
or jammed, this device cannot make backups.

In the event of a farcaster failure, this augmentation 
also includes a single-use emergency neutrino broadcaster
(p. 314) as well. This broadcaster contains approximately
10 nanograms of antimatter stored in an
orange-sized triply-redundant magnetic containment 
vessel. If the character is dying or urgently wishes to 
depart the morph, this tiny amount of antimatter is 
brought into contact with a similarly tiny amount of 
matter in a controlled fashion that generates a single 
brief and carefully coded neutrino pulse of the ego's 
most recent backup. However, the heat generated by 
this process literally cooks the entire morph, killing it 
and destroying all implants and electronics in or on it.

This entire process takes less than 0.1 second and 
the broadcast can be received as long as the neutrino
receiver is within 100 astronomical units of
the character. Within the solar system, this implant 
effectively guarantees the character's backup. It is less 
useful on exoplanets where the character is out of 
neutrino range of their backup facility. The amount 
of antimatter carried by this implant is sufficiently 
small enough that it does not produce an explosion 
and will not damage any surrounding objects. Most 
habitats carefully scan all visitors to determine if they 
have this implant and if the amounts of antimatter 
involved are sufficiently low as not to pose a danger 
to the habitat and its inhabitants, and some ban this 
implant entirely. \textbf{[Expensive]}

\textbf{Ghostrider Module:} This implant allows the 
character to carry another infomorph inside their 
head. This infomorph could be another muse, an AI, 
a backed-up ego, or a fork. The module is linked to 
the character's mesh inserts, so the ghost-rider can 
access the mesh. The character may limit the ghost-rider's
access, or may allow them direct access to their
sensory information, thoughts, communications, and 
other implants. \textbf{[Low]}

\textbf{Mnemonic Augmentation:} A character with this 
augmentation and a cortical stack can access digital 
recordings of all of the sensory data they have experienced
in XP format (and they may share these
recordings with others). Mnemonic augmentation 
differs from the eidetic memory bioware because it 
allows characters to digitally share all of their sensory 
data with others. It also allows them to closely examine
sensory data they did not initially look at. For
example, If the character glanced at a note but did not 
read it, they can later use image enhancement software
to enhance this image and in most cases actually
read what the note said. Mnemonic augmentation 
allows the character to clearly hear all background 
noises, like a conversation at a nearby table that the 
character only initially heard a few words of. Using 
mnemonic augmentation to retrieve a specific piece of 
information is quite easy, but usually requires between 
2 and 20 minutes of concentration. \textbf{[Low]}

\textbf{Multi-Tasking:} Only characters with cortical stacks 
can possess this augmentation. The character has an 
advanced computer installed in their brain that uses 
the data in the cortical stack to create several simultaneous
short-term forks to handle various mental tasks.
By design, this computer automatically reintegrates 
all of these forks into the character's core personality 
after a maximum of 4 hours, earlier if desired. This 
augmentation allows the character to both plan a 
speech and engage in intensive mesh-browsing while 
simultaneously fighting a gun battle or running from 
pursuit, since each of the forks operates independently. 
However, these forks can only perform purely mental 
or on-line interactions. This augmentation can produce
a maximum of two forks at a time, giving the
character an extra two Complex Actions on every 
Action Phase for mental or on-line actions. This implant
cannot be used simultaneously with any other
augmentation that allows for extra actions, or with 
the mental speed augmentation (p. 308). \textbf{[High]}

\textbf{Puppet Sock:} This implanted computer allows the 
biomorph's body (the ``puppet'') to be controlled by 
another character (the ``puppeteer''). While active, the 
puppet has no control over their body and is simply 
along for the ride (at the gamemaster's discretion, 
puppets who are tormented by repeated or extensive
loss of control may suffer mental stress). The
puppeteer may directly ``jam'' the puppet or remote 
control it in the same way that robots and pods are 
teleoperated (p. 196). The puppeteer must either be 
ghost-riding the puppet (see the \textit{Ghostrider Module,} p. 
307) or have a direct communications link (via mesh, 
radio, laser, etc.). \textbf{[Moderate]}

\subsubsection{Physical Augmentations}

This implants enhance the morph's physical body.

\textbf{Cyberclaws: }The bones on the back of the morph's 
hand are bonded to smart material claws. These claws 
can extend through concealed ports in the morph's 
skin and extend 6 inches past the morph's knuckles. 
These razor-sharp weapons inflict 1d10 + 3 + (SOM 
÷ 10) damage and have an AP of –2. If combined 
with eelware (p. 304), they can also inflict  electric 
shocks. Likewise, cyberclaws can also deliver poison 
or nanotoxins secreted from a poison gland (p. 305) 
or implanted nanotoxins. \textbf{[Low]}

\textbf{Cyberlimb:} In an age when arms and legs can 
easily be regrown, many people consider cybernetic 
prostheses to be vulgar and distasteful. The Scum 
and others, however, treat them as iconic symbols 
of self-expression. Standard replacement cyberlimbs 
function the same as their biological equivalents, 
though that particular limb receives a +3/+3 Armor 
bonus when targeted specifically (this bonus does not 
apply to synthmorphs). Cyberlimbs may be masked 
to look real (see \textit{Synthetic Mask,} p. 311), and may 
also feature small compartments for hiding/storing 
small objects. \textbf{[Moderate]}
%%% 310

\textbf{Cyberlimb Plus:} More extravagant cyberlimb 
models are also available, though they require more 
severe body alteration to accommodate. These limbs 
apply a +5 SOM bonus per limb (maximum +10). 
They may be replacement limbs or ``extra'' limbs anchored
in the body's skeletal frame. These cyberlimbs
may not be masked. \textbf{[High]}

\textbf{Hand Laser: }The morph has a weapon-grade laser 
implanted in its forearm, with a flexible waveguide 
leading to a lens located between the first two knuckles
on the morph's dominant hand. The laser fires
from this waveguide, inflicting 2d10 damage with 0 
AP. The laser is powered by a small nuclear battery 
located in the morph's torso, good for 50 shots before 
it must be recharged like other beam weapon batteries 
(p. 338). \textbf{[Moderate]}

\textbf{Hardened Skeleton: }The morph's skeleton has been 
laced with strengthening materials. Apply a +5 DUR 
and +5 SOM bonus. \textbf{[High]}

\textbf{Oxygen Reserve: }The morph has a miniature 
oxygen tank and rebreather installed in its torso. This 
implant provides the equivalent of the life support 
system in a light vacsuit (p. 333), allowing the character
to breathe comfortably for up to 3 hours. It feeds
oxygen directly to the morph's blood stream, avoiding 
problems with pressure changes. Implanted sensors 
automatically cause the character to use the stored 
oxygen if they detect poisonous or insufficient atmosphere
Without vacuum sealing, the character can
only survive in vacuum for 5 minutes, but remains 
conscious and active for the entire time, giving them 
far more time to find shelter or a vacsuit than characters
without this implant. For every hour the character
is in a breathable atmosphere, this implant recovers 
one hour of oxygen storage. The implant can be fully 
recharged within 15 minutes if the character is in a 
high-pressure mostly oxygen atmosphere. \textbf{[Low]}

\textbf{Reflex Boosters:} The morph's spinal column and 
nervous system is rewired with superconducting 
materials, boosting transmission speed. This raises 
the character's REF by +10 and improves Speed by 
+1. \textbf{[Expensive]}

\subsection{Nanoware}

All augmentation nanoware is advanced nanotechnology
(p. 328), consisting of a grape-sized nanobot
generator that produces specialized nanomachines. 
Nanoware is available for synthmorphs and bots in 
addition to biomorphs.

\textbf{Implanted Nanotoxins:} The morph has an implanted
nanobot hive that produces nanotoxins (p.
324). This implant is designed so that the character 
can deploy these nanobots instantly via a scratch with 
claws, spraying with saliva, or simply making continuous
bare-skin contact. Characters can choose whether
or not to deploy these nanobots. Each nanotoxin 
generator only produces a single variety of nanobots, 
with the most common types being ones designed to 
kill or incapacitate almost any living target or ones 
designed to destroy delicate machinery. Characters 
are immune to their own nanotoxins. Nanotoxins 
are highly restricted and many habitats will not allow 
anyone with this implant on board. \textbf{[Moderate]}

\textbf{Medichines:  }This is the most common form of 
nanoware. These nanobots monitor the user's body at 
a cellular level and fix any problems that arise. 

Medichines eliminate most diseases, drugs, and 
toxins (but not nanodrugs or nanotoxins) before 
they can do more than minor harm to the host (see 
\textit{Drug Effects,} p. 318). If desired, the user can temporarily
override this protection to permit intoxication
or other effects, but unless the character activates a 
second specially labeled override, medichines prevent 
the toxins from accumulating to lethal or permanently 
harmful levels. In this case, they can also be activated 
at a later point to reduce a drug or toxin's remaining 
duration by half. 

Medichines allow the character to ignore the effects 
of 1 wound. They also speed normal healing as noted 
under \textit{Biomorph Healing,} p. 208. If the user suffers 
5 or more wounds at once, or more than 6 wounds 
in an hour, the damage has exceeded the medichines' 
ability to repair. In this case, the medichines place the 
character into a medical stasis, where their mind and 
body are perfectly preserved, but where the character 
cannot act in any way. Under these circumstances the 
medichines also send out a priority call for emergency 
services via the character's mesh inserts.

Medichines for synthmorphs and bots consist 
of nanobots that monitor and repair the shell's 
integrity and internal system functions. Note that 
the synthmorph version of medichines allows the 
synthmorph to self-repair in the same manner by 
which a biomorph with medichines would naturally 
heal (p. 208). \textbf{[Low]}

\textbf{Mental Speed: }With this nanoware system, nanobots
alter the character's neural architecture and
augment the functioning of their neurons. The character
can deliberately speed up their mind to think
and also receive and process sensory information far 
faster than ordinary humans. Time seems to subjectively
slow down for the character, allowing them
to carefully plan their next action, even if they only 
have a split second to do so. With this system active, 
the character can discern things occurring too fast 
for a normal human to perceive, such as the individual
frames of an old analog film or understanding
sounds that were accelerated to many times their 
normal speed. The character can also read 10 times 
faster than normal and can track the paths of bullets 
and similar fast-moving objects with a successful 
Perception Test.

When using this augmentation, the character gains 
two extra Complex Actions during each Action Phase 
that may only be used for mental actions. The character
also receives a +30 Initiative bonus. The character
thinks at normal speed whenever this nanoware is inactive
This nanoware is incompatible with any other
augmentation that provides any form of extra actions, 
such as multi-tasking. This augmentation can be used 
%%% 311
as often as desired, but actively using it renders ordinary
conversation and social interactions difficult and
requires concentration to maintain. \textbf{[High]}

\textbf{Nanophages:} These nanobots patrol the body, alert 
for signs of intrusive nanodrugs or -toxins and destroying
them before they have more than a minor
effect. Nanophages provide automatic immunity 
against nanodrugs and nanotoxins unless they are 
specifically commanded to stand down by the user, via 
their mesh implants. \textbf{[Moderate]}

\textbf{Oracles: }These neural macrosensing nanobots pay 
attention to the sensory input on which the character
is not focusing, alerting them about important
things they might otherwise overlook. Oracles also 
act as a sort of memory buffer and search aid, extending
short term memory, helping the character
recall memories and details, and crosschecking 
them with other memories. Oracles negate Perception
modifiers for distraction, apply a +10 modifier
to Investigation Tests, and add a +30 bonus to
memory-related tests. \textbf{[Moderate]}

\textbf{Respirocytes:} These nanobots act as highly-efficient 
artificial red blood cells, increasing the ability to 
transfer oxygen and carbon dioxide. This increases 
the morph's ability to hold their breath to 4 hours 
and increases DUR by +5. \textbf{[Moderate]}

\textbf{Skillware: }The morph's brain is laced with a network
of artificial neurons that may be formatted with
downloaded information. This allows the user to 
download skillsofts (p. 332) into their brains, gaining 
the use of those programmed skills until the skillsoft 
is erased or replaced. Skillware systems are only 
capable of handling 100 total skill points worth of 
skillsofts at a time. \textbf{[High]}

\textbf{Skinflex: }This disguise implant allows the user to 
restructure their facial features and musculature and 
alter skin tone and hair color. The entire process takes 
a mere 20 minutes. Skinflex adds +30 to Disguise 
Tests. \textbf{[Moderate]}

\textbf{Skinlink:} Skinlink nanobots live on the morph's 
external skin or shell, automatically swarming over 
and creating a physical connection with any electronics
the user touches. They also take advantage of the
electrical field in a biomorph's skin for communication
They allow the user to communicate and mesh
with any devices merely by touching them. This 
is considered a wired link, and so is not subject to 
wireless interception or interference. Two skinlinked 
characters can also communicate and mesh simply by 
touching. \textbf{[Moderate]}

\textbf{Wrist-Mounted Tools: }The morph has a 6 centimeter-wide
metal band containing nanobot generators
implanted around each wrist. These nanobots link together
to duplicate the function of a utilitool (p. 326),
creating narrow, highly flexible arms that each ends in 
a specialized tool. These nanobots can also produce 
tiny fiber optics to allow the character to see through 
small openings, as well as being able to create small 
weapons equal to bioware claws. The fact that these 
tool are mentally controlled gives the character a +20 
modifier to skills involving repairing or modifying devices
with mechanical parts, opening locks or disarming
alarm systems, or performing first aid. \textbf{[Moderate]}

\subsection{Cosmetic Mods}

In an age of universal beauty, artistic cosmetic modification
of your body is commonly pursued by many
transhumans. Body mods once considered dangerous 
or edgy are now safe and commonplace, especially 
among factions like the anarchists, scum, or brinkers.

\textbf{Bodysculpting:} If your morph's enhanced physique 
isn't enough, you can take it further with custom 
bodysculpting such as as elongated ears or fingers, 
nose alteration, hair addition/removal, feathers, exotic 
%%% 312
eyes, snakeskin, endowed genitalia, and more unusual 
physical alterations. \textbf{[Low]}

\textbf{Nanotats:} Tattoos created with nanobots can 
move around the body, change shape/color/brightness
texture, alternate text and images, and/or even
create minor holographic effects on the skin's surface
all controllable via mesh inserts. \textbf{[Low]}

\textbf{Piercings:} Name any part of the body and someone's
figured out a way to pierce it, probably multiple
times. Hoops, barbells, plugs, and chains are
extremely common, often made of shapechanging 
smart materials. \textbf{[Trivial]}

\textbf{Scarification:} Given modern medical abilities, scars 
of any sort are purely an affectation. \textbf{[Trivial]}

\textbf{Scent Alteration:} Minor changes to a body's biochemistry
can alter a character's natural smell or
constantly perfume them. \textbf{[Low]}

\textbf{Skindyes:} Dye jobs are available in all conceivable 
colors and patterns. \textbf{[Trivial]}

\textbf{Subdermal Implants:} Adding small implants under 
the skin can create bumps, ridges, piercing anchors, 
and similar textures and alterations. \textbf{[Trivial]}

\subsection{Robotic Enhancements}

The following modifications are only available to 
synthmorphs/robots.

\subsubsection{Armor}

These armor modifications replace the synthmorph's 
built-in Armor rating.

\textbf{Heavy Combat Armor:} The synthmorph's frame is 
loaded with armor that offers protection from heavy 
weapons for serious combat operations. This modification
is bulky and noticeable; the bot frame is encased
in a heavy-duty carapace. It increases the bot's
built-in Armor to 16/16. The shell's mobility systems 
and power output are also enhanced to deal with the 
extra load. \textbf{[High]}

\textbf{Industrial Armor:} The shell is equipped with protection
against collisions, extreme weather, industrial accidents
and similar wear-and-tear. Increase the bot's
built-in Armor rating to 10/10. \textbf{[Moderate]}

\textbf{Light Combat Armor:} The synthmorph's frame is 
protected by armor designed for policing and security
duties. This increases the bot's built-in Armor to
14/12. \textbf{[Moderate]}

\subsubsection{Mobility Systems}

Shells are designed with a wide-range of propulsion 
systems, and are sometimes built for a specific environment/gravity
Some synthmorphs may have multiple
mobility systems. Many such systems are retractable, 
meaning they can be folded away into the shell's frame.

\textbf{Hopper:} Hoppers have two or more legs designed 
to propel the morph forward or up, much like a frog 
or grasshopper. \textbf{[Moderate]}

\textbf{Hovercraft:} The shell uses an impeller to blast a 
cushion of high-pressure air off the surface below, 
repelling the frame off the ground (modern hovercraft 
do not use rubber skirts). Most hovercraft travel a 
meter or so above the ground, but can temporarily 
levitate themselves higher for short periods. \textbf{[Low]}

\textbf{Ionic:} The shell uses principles of magnetohydrodynamics
to levitate and fly, by ionizing surrounding
air into plasma to create lift and momentum. The 
shell is also spun for stability. This system does not 
work in vacuum, but an underwater version uses the 
same mechanics for propulsion in liquid environments
\textbf{[High]}

\textbf{Microlight:} Popular in low-grav and microgravity 
environments, microlights encompass several types 
of ultralight or lighter-than-air systems, such as powered
paragliders, autogyros, balloons, aerostats, and
blimps. These systems do not work in vacuum. \textbf{[Low]}

\textbf{Roller:} Only for circular shells, this system allows 
the synthmorph to roll like a ball. The shell rolls 
around an interior axle, propelled by a motor-driven 
pendulum. \textbf{[Moderate]}

\textbf{Rotorcraft:} Rotating blades create lift, allowing 
the shell to move and hover like a helicopter. Most 
models use tilt-rotors or tilt-wings so that the rotorblades
may be moved forward (for faster propeller-like
propulsion) and for better maneuverability in
zero-G. This system does not work in vacuum. \textbf{[Low]}

\textbf{Snake:} Commonly used by slitheroids, these shells 
use lateral undulation, flexing their body from left to 
right and waving their frame forward. Such shells may 
also use sidewinding or a concertina motion straightening
forward, then retracting the rear) to move. They
also featured gyroscope stabilization so that they may 
circle into a hoop and roll like a wheel. \textbf{[Moderate]}

\textbf{Submarine:} Designed for undersea mobility, submarine
shells use propellers or pumpjets to push through
water. \textbf{[Moderate]}

\textbf{Tracked:} Tracked shells use smart rotating treads to 
work their way across surfaces that would bog down 
other ground vehicles. They can prop themselves up 
in order to overcome taller obstacles or to lay themselves
down to bridge across a ditch or crevice. \textbf{[Low]}

\textbf{Thrust Vector:} These shells use either turbofans or 
turbojets to create atmospheric lift with a set of wings. 
The engines may be maneuvered to point and generate 
thrust in different directions for vertical takeoffs/landings
and better maneuverability in zero-G. \textbf{[Moderate]}

\textbf{Walker:} Walkers use two or more limbs to walk or 
crawl across a surface. Many use grip pads (p. 305) or 
magnetic systems (p. 310) to stick to surfaces. \textbf{[Low]}

\textbf{Wheeled:} Most wheeled shells feature smart 
spokes that allow the wheels to conform their shape 
to obstacles and even climb stairs. Some low-grav 
shells feature puncture-resistant and self-repairing 
compressed-gas tires. \textbf{[Low]}

\textbf{Winged:} Primarily used by smaller shells, this system 
of four independently-controlled wings allows the 
shell to hover or move rapidly in any direction. \textbf{[Low]}

\subsubsection{Physical Modifications}

These mods are applied to the shell's physical frame.

\textbf{Extra Limbs:} The shell is equipped with one or 
more extra limbs. A character using these limbs 
%%% 313
suffers an off-hand modifier (p. 193). These limbs 
may be arms (with hand/grippers/etc.), legs, tentacle-like
or otherwise articulated and/or prehensile. Some
shells have rotational frames that allow them to move 
limbs around their body. \textbf{[Low]}

\textbf{Fractal Digits:} The synthmorph has ``bush robot'' 
digits that are capable of splitting into smaller digits, 
and those smaller digits into micro digits, and so on 
down to the micrometer scale, allowing for ultra-fine 
manipulation. Apply a +20 COO modifier where such 
fine manipulation is a factor (such as detailed repair 
work). The bot must have functioning nanoscopic 
vision (p. 311) to get this bonus. \textbf{[Moderate]}

\textbf{Hidden Compartment:} The shell has a concealed 
aperture for a shielded interior compartment, ideal for 
storing valuables or smuggling contraband. Apply a 
–30 modifier to detect this compartment either manually
or with sensor scans. \textbf{[Low]}

\textbf{Magnetic System:} A magnetic system allows the 
shell to cling to most ferrous materials. This enables 
the character to walk in zero-G situations by magnetically
adhering surfaces, hang upside down, and hold
onto devices without letting them drop or drift away. 
The shell receives a +30 modifier whenever maintaining
a magnetic hold on something. \textbf{[Low]}

\textbf{Modular Design:} This shell is designed to lock together
with similar modular morphs in different architectural
patterns to create larger gestalt forms. When
united with other modules, the group is treated as a 
single unit/morph, with shared capabilities. If damaged
and then separated, damage and wounds are distributed
evenly between modules; uneven amounts are
allocated randomly. The exact capabilities of different 
shapes depends on the composition, and is largely left 
in the gamemaster's hands. \textbf{[High]}

\textbf{Pneumatic Limbs:} The limbs are equipped with 
pneumatic cylinder systems that can generate up to 
1,500 pounds of thrust. This allows the shell to push 
off and make impressive jumps (a synth of human 
size/weight can leap over 2 meters up). Apply a +20 to 
Freerunning Tests. A pneumatic limb used to strike an 
opponent in unarmed combat inflicts an extra 1d10 
damage. \textbf{[Low]}

\textbf{Retracting/Telescoping Limbs:} The shell's limbs can 
either be retracted completely inside it's frame and/or 
extended for extra length (usually up to 1 or 2 meters 
extra). Telescoping limbs may give the shell a reach 
advantage in melee combat (p. 204). \textbf{[Low]}

\textbf{Shape Adjusting:} This shell is made from smart 
materials that allow it to alter its shape, altering its 
height, width, circumference, and external features, 
while retaining the same mass. This modification is 
typically employed to reshape the morph into special 
configurations adapted to specific tasks (for example, 
lengthening to crawl through a tunnel, widening its 
base for stability, expanding to reach out and attach 
to multiple access point simultaneously, and so on). 
This mod also allows the morph to change its features 
for disguise purposes; apply a +30 modifier to Disguise
Tests. \textbf{[High]}

\textbf{Structural Enhancement:} This modification bolsters 
the shell's structural integrity, boosting its ability to 
take damage. Increase Durability by 10 and Wound 
Threshold by 2. \textbf{[Moderate]}

\textbf{Swarm Composition:} The shell is not a single unit 
but a swarm of hundreds of insect-sized robotic microdrones
Each individual ``bug'' is capable of crawling
rolling, hopping several meters, or using nano-copter
fan blades for airlift. The cyberbrain, sensor
systems, and implants are distributed throughout the 
swarm. Though the swarm can ``meld'' together into 
a roughly child-sized shape, the swarm is incapable of 
tackling physical tasks like grabbing, lifting, or holding
as a unit. Individual bugs, however, are quite capable
of interfacing with electronics. Swarms cannot
carry most gear or wear armor, and may not make 
strength-based SOM-linked skill tests. For combat 
purposes, use the same rules as given for nanoswarms, 
p. 328. Damage and wounds are reflected as damaged/massacred
bugs. The swarm may be ``healed'' by
manufacturing more bugs.\textbf{[High]}

\textbf{Synthetic Mask:} The synthmorph is equipped 
with a realistic outer casing of faux-skin and carefully
sculpted to pass as a biomorph (perhaps even a
particular person). The morph can cry, spit, have sex, 
and will even bleed if cut. Only a detailed physical 
examination or a radar, terahertz, or x-ray scan will 
detect the synthmorph's true nature, and even then 
such exams/scans suffer a –30 modifier. \textbf{[Moderate]}

\textbf{Weapon Mount:} The shell carries a built-in (or built-on
weapon. This weapon mount may be either internal
(concealed, only weapons small in relation to the shell 
may fit, –30 to Perception Tests to detect) or external 
(visible). It may be fixed (one direction only), swiveling 
(limited field of fire), or on an articulated mount (all 
directions). \textbf{[Low; Moderate for concealed/articulated]}

\subsubsection{Sensors}

\textbf{360° Vision: }The shell's visual sensors are situated for 
a 360-degree field of vision. \textbf{[Low]}

\textbf{Chemical Sniffer:} This sensor detects molecules in the 
air and analyzes their chemical composition. It enables 
Chemistry Tests to determine the presence of gases, including
toxins and other fumes. It can also detect the
presence of explosives and firearms. \textbf{[Moderate]}

\textbf{Lidar:} This sensor emits laser light and measures 
the reflections to judge range, speed, and image the 
target. See \textit{Using Enhanced Senses,} p. 302. \textbf{[Low]}

\textbf{Nanoscopic Vision:} The shell's visual sensors can 
focus like a microscope, using advanced superlens 
techniques to beat the optical diffraction limit and 
image objects as small as a nanometer. This allows 
the character to view and analyze objects as small as 
blood cells and even individual nanobots. The synthmorph
must stay relatively steady to view objects at
this scale. \textbf{[Moderate]}

\textbf{Radar:} This sensor system bounces radio or microwaves
off targets and measures the reflected waves
to judge size, composition, and motion. See \textit{Using }
\textit{Enhanced Senses,} p. 302. \textbf{[Low]}
%%% 314

\subsection{Armor}

Modern personal armor systems have advanced from 
the high modulus polyethylene thermoplastics and 
aramid fabrics of the early 21st century. Armor in 
\textit{Eclipse Phase} is derived from biotech, in the form of 
organoweave fibers and crystalline-grown plates, and 
nanotech, in the form of shock-absorbing fullerene (p. 
298) materials. Occasionally other materials are used, 
such as metallic glass plates or shear-resistant fluids 
that harden against impacts. Such armor protects 
against (armor-piercing) bullets and kinetic impacts 
as well as bladed weapons and piercing sharp objects. 
They also insulate against both the explosive heating 
of energy weapons and electrical shocks. While such 
armor protects against bullets, the layers of material
catch the bullet and redistribute its kinetic energy
across the body, which can still result in severe blunt 
force trauma.

Rules for armor in combat can be found on p. 194. 
Armored exoskeletons are listed on p. 343.

\textbf{Armor Clothing:} The extra-resilient organoweave 
fibers and fullerene materials that offer basic protection
against kinetic and energy weapons can be woven
in with normal smart materials to create a wide range 
of discreet armor clothing that provides a subtle level 
of security. Such protective garments are indistinguishable
from regular clothing and come in all styles
and designs. Armor clothing provides an Armor Value 
of 3/4. \textbf{[Trivial]}

\textbf{Armor Vest:} Armor vests provide more thorough 
protection to a body's vital areas, covering the abdomen
and torso completely, protecting the neck with a
rigid collar, and even providing wrap-under protection
for the groin. Though armor vests are not bulky,
they are obvious as armor. Armor vests may be worn 
with armor clothing without penalty. Armor vests 
provide an Armor Value of 6/6. \textbf{[Low]}

\textbf{Body Armor (Light):} These high performance armor 
outfits protect the wearer from head to toe. An integrated
armor vest is supplemented with increased protection
on the limbs and joints, while still managing to
be flexible and non-restrictive. Body armor is typically 
worn by security and police forces, and supplemented 
with a helmet. It provides an Armor Value of 10/10. 
\textbf{[Low]}

\textbf{Body Armor (Heavy):} Similar to light body armor, 
but with extra protective layers, often ergonomically 
manufactured to conform to a specific  character's 
body, and an environmental seal with climate control 
to protect the wearer from hostile environments. It 
provides an Armor Value of 13/13. \textbf{[Moderate]}

\textbf{Crash Suit:} Designed for both industrial worksite 
safety and protection from accidental zero-G collisions, 
crash suits are also favored by sports enthusiasts and 
explorers. The basic jumpsuit offers comfortable protection
equal to that of armor clothing. When activated
with an electronic signal, however, elastic polymers
within the suit stiffen and form rigid impact protection 
for vital areas. Crash suits provide an Armor Value of 
3/4 when inactive and 4/6 when activated. \textbf{[Low]}
\textbf{Helmet:} This armor accessory is usually worn with 
body armor or a battle suit. Light helmets are open, 
whereas full helmets latch on and provide an environmental
seal with a 12 hour supply of air. Light helmets
provide an Armor Value bonus of +2/+2, whereas full 
helmets add +3/+3. Helmets are often equipped with 
an ecto (p. 325), a radio booster (p. 313), and sensors 
equal to specs (see p. 325). \textbf{[Trivial]}
\textbf{Riot Shield:} Used for mob suppression, riots shields 
are light-weight, tough, and may be set to electrify on 
command, stunning anyone who comes into contact 
with the outer surface (treat as shock glove effects, p. 
334). Riot shields provide an Armor Value bonus of 
+3/+2. \textbf{[Low]}
\textbf{Second Skin:} This lightweight bodysuit, woven 
from spider silks and fullerenes, is typically worn 
as an underlayer, though some athletes use it as a 
uniform. It provides minimal protection, but may be 
worn with other armor without penalty. It provides an 
Armor Value of 1/3. \textbf{[Low]}
\textbf{Smart Skin:} Smart skin is an advanced nanofluid
that covers the wearer's skin. It resembles liquid
mercury but retains the texture and flexibility  of 
normal skin until activated, at which point the material
becomes rigid enough to protect the wearer and
Armor Clothing 

3

4

311
Armor Vest 

6

6

312
Battle Suit Powered Exoskeleton 

18

18

344
Bioweave Armor (Light) 

2

3

302
Bioweave Armor (Heavy) 

3

4

302
Body Armor (Light) 

10

10

312
Body Armor (Heavy) 

13

13

312
Carapace Armor  

11

11

303
Crash Suit (Inactive) 

3

4

312
Crash Suit (Active) 

4

6

312
Exowalker 2

4

344
Hard Suit 

15

15

334
Helmet (Light) 

+2

+2

312
Helmet (Full) 

+3

+3

312
Hyperdense Exoskeleton 

6

12

344
Riot Shield 

+3

+2

312
Second Skin 

1

3

312
Smart Skin 

3

2

312
Smart Vac Clothing

2

4

325
Spray Armor 

2

2

312
Synthmorph Industrial Armor  

10

10

310
Synthmorph Combat Armor (Light) 

14

12

310
Synthmorph Combat Armor (Heavy) 

16

16

310
Transporter Exoskeleton 

2

4

344
Trike Exoskeleton 

2

4

344
Vacsuit (Light)  

5

5

333
Vacsuit (Standard)  

7

7

333
%%% 315
distribute the kinetic energy (though still flexible 
enough at the joints not to impede movement). A specialized
hive, worn by the character, replenishes the
nanobots and stores them when not in use. Deploying 
the nanobots across the body takes a full Action Turn. 
Smart skin has an Armor Value of 3/2, and may be 
worn with other armor without penalty. \textbf{[Low]}

\textbf{Spray Armor:} This fast armor application comes in 
a spray can and disperses a smart chemical polymer 
that sticks to bare flesh (but does not adhere to hair 
and eyes). The polymer solidifies into a form fitting 
body armor fabric when exposed to body temperature 
with the look and feel of a latex suit. Spray armor 
does not work on synthetic morphs or on clothing or 
other armor. The color and feel of the armor can be 
adjusted with electric currents and additional polymers
making it popular among some socialite and
nightlife scenes. The spray-on armor does not wash 
off, but degrades 1 point of armor (both energy and 
kinetic) every 12 hours. It may be removed with a 
special nanotech solvent. Spray armor has an Armor 
Value of 2/2. \textbf{[Low]}

\subsection{Armor Mods}

Armor modifications add extra materials or coatings 
that either enhance the armor's resistance to certain 
dangers or provide other effects. Armor mods may be 
easily added or removed with the appropriate nanobot
applicators.

\textbf{Ablative Patches:} These thin and light slap-on 
patches of stick to armor and are designed to absorb 
heat and energy from beams and explosions, safely 
vaporizing and blowing hot gas away. Ablative 
patches increases the Armor Value by +4/+2, but each 
hit reduces both the energy and kinetic value of the 
ablative armor by 1. \textbf{[Trivial]}

\textbf{Chameleon Coating:} This provides the armor with 
the same effect as the chameleon cloak (p. 315). \textbf{[Trivial]}

\textbf{Fireproofing: }Fireproofing includes the addition of 
heat-resistant ceramic or fire-resistant layers, both 
capable of withstanding extremely high temperatures. 
Fireproofing increases the Armor Value by +2/+0, and 
provides an additional 10 points of armor against 
heat or fire specifically. \textbf{[Trivial]}

\textbf{Immunogenic System:} The immunogenic mod 
adds an active nanobot swarm, maintained by a specialized
hive, that coats the outer layer of armor and
also the non-armored parts of the wearer's morph. 
It acts as an outer immune system designed to neutralize
toxic agents and nanotoxins with which it
comes into contact. This provides immunity to drugs, 
toxins, and nanotoxins applied dermally, such as 
with a slap patch or splash grenade. It has no effect 
on inhaled, oral, or injected drugs (including coated 
weapons). \textbf{[Low]}

\textbf{Lotus Coating:} The armor has been impregnated 
with a superhydrophic coating (contact angle of 
around 170°) that repels all water-like liquids. If the 
armor is splashed by liquid toxins or chemicals, the 
effect is reduced since the liquids starts to roll off the 
armor. Apply a +30 modifier when defending against 
liquid-based attacks. \textbf{[Trivial]}

\textbf{Offensive Armor: }When activated, the outer layer of 
this armor is rigged to shock anyone or anything that 
contacts it with electricity. Treat its DV and effect as a 
shock baton (p. 334). \textbf{[Low]}

\textbf{Reactive Coating:} A thick layer of advanced 
nanotech is applied to the armor, protecting it with 
a colony of nanobots designed to sense incoming 
attacks. When an attack strikes the coating, it detonates
to disrupt the attack. Bursts and full autofire
are treated as a single attack. A reactive coating 
increases the Armor Value by +5/+5, but each detonation
automatically inflicts 1 point of damage on
the wearer. Reactive armor also works against melee 
attacks, but the attacker also suffers 1d10 ÷ 2 (round 
up) points of damage per attack (armor protects) 
from the microexplosion. Reactive coating only 
works against 5 attacks, after which the specialized 
nanobot hive replenishes the coating at the rate of 1 
use per hour. \textbf{[Moderate]}

\textbf{Refractive Glazing:} A combination of reflectors, 
refractive metamaterials, and an energy transfer 
system with heat radiators provides extra protection 
against energy weapons. Increase the Armor Value 
by +3/+0. \textbf{[Low]}

\textbf{Self-Healing:} The armor is equipped with a nanohive
that acts like repair spray (p. 333). \textbf{[Moderate]}

\textbf{Shock Proof:} Shock proof armor is electronically 
insulated to discharge and reduce the effect of shock 
weapons. Apply an additional +10 modifier when resisting
the DV and effects of shock weapons (p. 204). \textbf{[Low]}

\textbf{Thermal Dampening:} Thermal dampening obfuscates
heat signatures by converting body heat into
electric energy. It makes the target more difficult to 
spot with thermal sensors; apply a –30 modifier for 
Perception Tests. \textbf{[Moderate]}

\subsection{Communications}

The oldest and most widespread communications 
technology still in regular use is radio. Every habitat 
and world inhabited by transhumanity is awash in 
radio traffic, with humans, machines, and uplifts all 
constantly communicating with one another. The 
smallest radios are no larger than a spec of dust and 
have a range of no more than 20 meters, while the 
largest are the size of a truck and have a range of 
many thousands of miles. Radios large and small are 
ubiquitous and almost all devices contain at least 
short-range radios so they may interact with the 
mesh. Most morphs are equipped with basic mesh 
inserts (p. 300) that include an implanted radio. For 
radio ranges, see p. 296.

\textbf{Fiberoptic Cable:} Fiberoptic cables are used to 
establish wired connections between two devices. 
Given the ubiquity of radios and the tangled mess 
wires cause, they are typically only used for privacy 
(unlike radio communication, fiberoptic signals may 
not be intercepted) or in areas with heavy radio 
interference. \textbf{[Trivial]}
%%% 316

\textbf{Laser/Microwave Link:} These portable devices are 
used to establish a tight-beam, line-of-sight communications
channel with another laser or microwave
link. The range of these transceivers varies widely 
with environmental factors, but approximates 50 
kilometers in atmosphere and 500 kilometers in 
space (though horizon limits must be kept in mind, 
being 5 kilometers at ground level on Earth and less 
on smaller bodies). Lasers are subject to interference 
from fog, dirt, smoke, and similar visual chaff, while 
microwaves may be hindered by metallic obstructions. 
These links may only be intercepted by getting directly 
in between the beams. Some teams carry a micro version
of this system, worn on their person, allowing
line of sight intra-team communications that cannot 
be intercepted like radio. \textbf{[Moderate]}

\textbf{Radio Booster:} This device boosts the range and 
sensitivity of short-range radios, like those from implants
ectos, or microbugs. The booster must be with
the shorter-ranged device's range (or directly linked 
via fiberoptic cable). It will repeat any transmissions 
received from that device, but at its extended range of 
25 kilometers in urban areas (250 kilometers remote 
areas). Broadcasts from a radio booster are easy to 
receive by anyone looking for broadcasts (see \textit{Wire-}
\textit{less Scanning,} p. 251), though transmissions may be 
stealthed (p. 252). Boosters are commonly used by 
characters traveling far from habitats or other civilized
regions. \textbf{[Low]}

\subsection{Neutrino Communicators}

Neutrinos are particles that can pass through any 
solid matter with ease and are impossible to block. As 
a result, they make an ideal medium for communications
Unfortunately, they are also easy to intercept.
Even a tight beam of neutrinos sent between two locations
can be intercepted simply by placing another
receiver behind the location the broadcaster is sending 
to. Neutrino communicators require a large power 
plant to power the high energy particle interactions 
required to generate the neutrino broadcast. Neutrino 
receivers are also relatively large, with the smallest 
occupying 100 cubic meters. In most cases, neutrino 
communicators are designed to broadcast neutrinos 
in all directions, though tight-beam transmissions are 
also possible. Quite often neutrino communications 
take advantage of quantum farcasting for security.

\textbf{Neutrino Transceiver:} This transceiver is capable of 
generating and receiving neutrino signals at a range of 
at least 100 astronomical units. It is large, with a size 
of 8 cubic meters (in a cube 2 meters on a side), but 
they can be loaded onto large vehicles. To function, it 
must be connected to a large power plant, such as one 
found in habitats or large spacecraft. The cost and 
size of this device includes the computer necessary for 
quantum farcasting. \textbf{[Expensive]}

\subsection{Quantum Farcasters}

Quantum farcasters are special computers designed to 
protect a communications channel (such as fiberoptic, 
radio, laser/microwave, or neutrino) with unbreakable
encryption. To function, two or more quantum
farcaster computers must first be entangled together 
(on a quantum level) in the same physical location. 
The farcasters may then be separated, at which point 
they may continue to exchange encrypted data via 
quantum teleportation. This data exchange requires 
a standard communications link (fiberoptic,  radio, 
laser/microwave, or neutrino), and so is limited by 
the speed of light, but it is a high bandwidth form 
of communications. The quantum encryption used 
by these entangled farcasters is unbreakable, and any 
attempted interception is immediately detected and 
neutralized. A quantum farcaster may not be used to 
securely communicate with any farcasters other than 
the ones it is entangled with.
%%% 317

Because it is exceptionally safe and secure, quantum 
farcasting via neutrino communications is the primary 
means of both long-distance communication between 
habitats and egocasting (p. 276). The neutrino signal 
cannot be blocked and it can only be decrypted if a 
character has access to the computer that is sending 
or receiving the signal.

\textbf{Miniature Radio Farcaster: }Miniature farcasters 
communicate with each other using standard radio 
transceivers. As noted above, they may only securely 
communicate with the other farcasters with which they 
are entangled. Most miniature farcasters are worn as 
jewelry or fitted into clothing or other equipment. \textbf{[Low]}

\subsection{Quantum Entangled Communication}

The rarest form of communications is quantum entangled
(QE) communication. QE communication is
instantaneous and works over any distance, but is 
also very limited. QE communication requires pairs 
of entangled particles known as qubits. To use QE, 
large number of pairs of qubits are created and then 
separated from each other. Millions of these separated
pairs of particles are stored in special containers
known as qubit reservoirs. If two QE communicators 
each have a qubit reservoir containing qubits that are 
each entangled with qubits in the other communicator's
qubit reservoir, then characters can use the two
QE communicators to commutate with one another 
instantaneously. Characters can use QE to instantly 
communicate between any two locations, even if one 
character is in the solar system and the other has 
passed through a Pandora gate and is standing on a 
planet 500 light years away.

Each bit of data transmitted between these two QE 
comms uses up one qubit. Once all of the qubits are 
used up, the two QE comms can no longer communicate
with each other until they each get a new batch
of entangled qubits. Qubits are expensive to produce, 
contain, and transport, making this an exceedingly 
expensive form of communication. As a result, extremely
high bandwidth communications like full
sensory AR and egocasting cannot be performed using 
QE communication.

\textbf{Portable QE Comm: }This is a handheld FTL communications
device. The actual communications unit
can be made as small as desired, but must be large 
enough to connect to or hold a qubit reservoir. Because
qubit reservoirs are relatively large and must be
replaced, they are rarely implanted. Some miniature 
farcasters are designed so that users can also attach 
qubit reservoirs to enable them to be used for both 
light speed and FTL communication. \textbf{[Low]}

\textbf{Low-Capacity Qubit Reservoir:} Low-capacity qubit 
reservoirs can be used for 10 hours of high-resolution 
video conferencing or meshbrowsing and 100 hours 
of voice or text only communications. \textbf{[High]}

\textbf{High-Capacity Qubit Reservoir:} High-capacity qubit 
reservoirs can be used for 100 hours of high-resolution 
video conferencing or meshbrowsing and 1,000 hours 
of voice or text only communications. \textbf{[Expensive]}

\subsection{Covert And Espionage Technologies}

These technologies allow characters to acquire protected
information and to gain access to places that
others try to keep them out of. Many of these devices 
are mesh-capable and equipped with radios, see p. 296 
for radio ranges.

\textbf{Chameleon Cloak: }This loose, poncho-like cloak 
contains a network of sensors that perceive wavelengths
from microwave to ultra-violet. A similar network
of miniature emitters precisely replicate the information
its sensors receive, making the wearer seem
transparent to those wavelengths. A chameleon cloak 
allows a character to effectively become invisible as 
long as they are stationary or not moving faster than 
a slow walk. When worn by someone moving faster, 
the cloak still provides a +30 modifier to Infiltration 
Tests to avoid being seen or noticed.

Chameleon cloaks are not effective against radar, 
x-ray, or gamma-ray sensors. They do hide the character
from thermal infrared, however, by absorbing
the character's body heat into its heat sink. The cloak 
can only absorb a character's body heat for one hour 
before it must emit this heat. Heat emission also requires
one hour, during which time the character is
easily visible in the thermal infrared spectrum. \textbf{[Low]}

\textbf{Covert Operations Tool (COT): }This handheld 
device is the ultimate in infiltration technology. It contains
both smart matter micromanipulators, cutting
%%% 318
tools, and an advanced nanotechnology generator 
capable of producing nanobots that can bore or cut 
through almost any material and disable or open 
almost any electronic lock.

Cutting out a lock or boring a 1-millimeter hole in 
a wall with a COT requires ((Durability + Armor) ÷ 
10) seconds. Cutting out a 1-meter diameter hole in 
a wall requires ((Durability + Armor) ÷ 10) minutes. 
These same nanobots can later be used to repair this 
damage so that it is invisible to any but the most careful
and detailed examination.

A COT can easily open any old-fashioned mechanical
lock simply by analyzing it and shaping an appropriate
key, though this takes a full Action Turn.
It can also open electronic locks by infiltrating them 
with nanobots that influence the lock's electronics, 
no matter what authentication system the lock uses. 
Opening electronic locks takes a full Action Turn, but 
success is practically guaranteed. Opening an electronic
lock in this manner will, however, trigger an
alarm and/or be logged as an event. For more details, 
see \textit{Electronic Locks,} p. 291. \textbf{[High]}

\textbf{Cuffband:} This smart plastic loop restricts around 
a prisoner's limbs when activated. If the prisoner 
struggles, it will tighten more. Cuffbands will inform 
the user if they are cut or loosened and are electronically-controlled
so the user can release the prisoner
remotely. Some cuffband variants including a shock 
system (treat as a shock baton, p. 334) to zap and 
restrain unruly prisoners. \textbf{[Low]}

\textbf{Dazzler:} The dazzler is a tiny laser system set on 
a rotating ball. When activated, it consistently spins 
and emits laser pulses in all directions. These laser 
pulses are not dangerous, but they detect the lenses 
of camera systems (including specs, viewers, and bot/
synthmorph sensors) and repeatedly zap them with 
laser pulses of varying strength to overload and dazzle 
them. For as long as a dazzler is active, any camera 
system (visual, infrared, and ultraviolet) within line 
of sight and within 200 meters is blinded. \textbf{[Moderate]}

\textbf{Disabler:} This handy device emits an overloading 
surge that completely incapacitates and disables a 
synthetic morph or pod (anything with a cyberbrain) 
when it is plugged into an access jack and activated. 
The affected cyberbrain will be unable to function until 
the signal is deactivated, effectively shutting down the 
ego (or AI). In order to plug a disabler into an unwilling
target, the target must first be grappled or a called
shot must be successfully made in melee combat. This 
device does not work on larger synthetic morphs (like 
vehicles) or on cyberbrainless robots.\textbf{[High]}

\textbf{Fiber Eye:} This is a flexible and electronically-controllable
length of fiberoptic cable and viewer, which
can be worked through cracks, under doors, and 
around corners to peep unobtrusively. \textbf{[Low]}

\textbf{Invisibility Cloak:} This cloak is made of metamaterials
with a negative refractive index, so that light
actually bends around it, making it and anything it 
covers invisible. This invisibility works from the 
microwave to ultraviolet spectrums, but not against 
radar or x-rays. The drawback is that anything concealed
within the cloak can't see out. This is easily
overcome by using external sensor feeds (if available) 
and entoptics to navigate. Alternately, a small piece of 
anti-cloak, which cancels the cloak's invisibility properties
when touched together, can be used to create a
small window to peep out of, though this increases 
the chance of being spotted. Noticing such a window 
requires a Perception Test with a –30 modifier. \textbf{[High]}

\textbf{Microbug: }This device is a tiny camera and microphone
1 millimeter across. It has the visual capabilities
of a set of specs (p. 325). It can hear everything
within 20 meters and see everything within the same 
range that is in its line of sight. A microbug can record 
up to 100 hours of information. Microbugs can be 
set to broadcast continuously, at set intervals, or only 
when they receive a special signal. If desired, they 
can also be set to only record if there is movement 
or voices in the room they are in. Microbugs have 
adhesive backs and can stick to almost any surface. 
Microbugs can also establish their location via mesh 
positioning or GPS, and so double as tracking devices. 
To avoid being detected by their radio transmissions, 
some microbugs are attached to miniature quantum 
farcasters (p. 314). These microbugs are much larger 
(1 centimeter) and easy to see, but their transmissions 
cannot be detected or blocked. \textbf{[Trivial, Low for quan-}
\textbf{tum farcaster bugs]}

\textbf{Prisoner Mask:} This hood tightens around the 
head of a prisoner, blocks all vision frequencies, and 
engages in low-level jamming in order to prevent any 
wireless communication via mesh inserts. \textbf{[Medium]}

\textbf{Psi Jammer:} This device jams frequencies used 
by brainwaves within a 20-meter radius. This has 
no effect on brain functions, but it does prevent 
any ranged used of psi sleights within this area of 
effect. \textbf{[Moderate]}

\textbf{Quantum Computer:} These advanced devices 
make use of quantum computation, allowing them 
to handle extremely large numbers with ease. This 
makes them especially useful for codebreaking, as 
noted on p. 254. \textbf{[Expensive]}

\textbf{Smart Dust:} This device is a walnut-sized specialized
nanobot generator that creates tiny sensor nanobots
each one of which is a tiny sphere the diameter
of a human hair. A packet of smart dust nanobots is 
sufficient to perform detailed surveillance on a large 
room like an auditorium has a volume of 1 cubic 
centimeter and contains 3 million nanobots. Each 
nanobot contains tiny cameras, microphones, a tiny 
computer, a radio, and chemical sensors, as well as 
short legs that allow them to walk and climb at a rate 
of 5 cm per second.

When a character dumps a packet of smart dust in 
a room, it will cover every surface in the room within 
20 minutes, including all furniture and the insides 
of every drawer and other space that is not airtight. 
At this point, the smart dust has recorded all data 
about the room that can be obtained by exceedingly 
detailed observation, including the DNA of everyone 
%%% 319
who has visited the room in the last week or two. The 
smart dust can then either broadcast a brief, highly 
compressed signal, or it can send all of its information
to a few hundred nanobots that then walk to a
pre-arranged destination for pickup and downloading 
by their user. The user need only find a single nanobot
with a nanodetector to acquire the information
obtained by the smart dust. If ordered to do so, the 
remaining nanobots can either power down and await 
further orders or self-destruct in a fashion that turns 
them into a tiny amount of dust made mostly of metal 
and silicon. \textbf{[Moderate]}

\textbf{Traction Pads:} This set of specialized fingerless 
gloves, shoes, and kneepads is designed to emulate the 
pads on geckos' feet. Characters can support themselves
on a wall or ceiling by placing any two of these
pads against any surface not made from a material 
specially designed to resist such devices. Characters 
can climb any surface and move easily across walls 
and ceilings that can support their weight (+30 to 
Climbing Tests). In addition to climbing, these devices 
are also very popular in zero-g environments. Wearing 
this item does not impair the user's agility or manual 
dexterity. \textbf{[Low]}

\textbf{White Noise Machine:} This small and wearable 
device generates masking sounds that protect a conversation
from being audibly recorded or overheard
by anyone not in the immediate vicinity. \textbf{[Trivial]}

\textbf{X-Ray Emitter:} This device is designed to be used 
with either the enhanced vision augmentation (p. 
301) or specs (p. 325). It emits a focused beam of 
low-powered x-rays that allows the user of either 
device to both see and see through most objects using 
backscatter x-ray radiation (p. 303). This allows the 
character to literally see through walls and into containers
including ones made of metal. \textbf{[Low]}

\section{Drugs, Chemicals, And Toxins}

In \textit{Eclipse Phase,} the transhuman desire to enhance 
the body and mind—especially with chemicals—
merges right into humanity's popular pastime of 
recreational substance abuse. Drugs of all kinds, 
whether they be chemical, nano-based, or electronic, 
are not only popular but widespread. While advances 
in biotechnology have eliminated many of the side 
effects that once plagued drug users, transhuman 
bodies remain complicated environments, and so side 
effects (especially with long-term use) are still a factor. 
Additionally, addiction is always a consideration for 
anyone who gets comfortable with popping the same 
pills too often, though there are also drugs for addiction
of course.

Drug descriptions include benefits, side effects, noticeable
signs that a person is using the drug, addictiveness
and effects from long-term use). Descriptions also
include the drug's Duration and its Addiction Modifier 
(see \textit{Addiction and Substance Abuse}).

\subsection{Substance Rules}

These rules explain how to handle drugs and toxins.

\subsubsection{Classification Of Substances}

Substances fall into four categories:

\textbf{Chemicals}: These are pharmacologically-active 
small chemical compounds (toxins, pharmaceuticals, 
chemical drugs) that have been produced by chemical 
synthesis, nanotech fabrication, or enzymatic biosynthesis
in (transgenic) organisms. They include naturally-occurring
drugs from known species of (exo-)flora
and fauna, endotoxins produced by biological organisms
enhancements of endogenic substances (designer
drugs), and de novo developments designed for a 
specific medical or recreational application. Chemical 
drugs affect only biological morphs and pods.

\textbf{Biologicals}: These include peptides, hormones, and 
biologically-based substances like biotoxins, bacteria, 
and viral organisms—drugs devised or based on 
naturally-occurring endogenic biological substances. 
This category also includes infectious biological organisms
that can produce drug-like effects, like virii
and bacteria. Biologicals affect biomorphs and pods 
but not synthetic morphs or infomorphs.

\textbf{Nanodrugs: }These are temporary nanobot colonies 
programmed to create a certain effect. While nanobots
are generally able to target or infect all morph
types except infomorphs, exactly which morphs are 
affected usually depends on the pre-programmed 
effect (i.e., whether it targets a biological or mechanical
mechanism).

\textbf{Electronic: }Electronic drugs include software and 
technology that affect the brain directly, such as manipulative
XP programs or retro-tech like transcranial
magnetic stimulation or cranial electrotherapy. It also 
includes narcoalgorithms—programs that reproduce 
drug-like effects for AIs, infomorphs, and egos residing
in cyberbrains.

\subsubsection{Application Methods}

There are number of vectors by which a substance 
may be applied to a morph.

\textbf{Dermal (D): }This drug or toxin is absorbed via 
the skin (or exterior hull with some nanotoxins) as 
either a gas, liquid, or solid (e.g., paste). Slap patches 
and slap bands are commonly used, loaded with the 
chemical DMSO, which transfers the drug through 
the skin.

\textbf{Inhalation (INH):} This is a gas that is breathed into 
the lungs or snorted nasally. Used for inhalers, aerosols
powders, and gas grenades/seekers.

\textbf{Injected (INJ): }This liquid is applied via either 
an intramuscular or intravenous injection. Used for 
needles and piercing weapons.

\textbf{Oral (O):} This is a liquid or solid that is absorbed 
through the stomach or oral cavity (eating or drinking
Used with pills and liquids.

\subsubsection{Drug Effects}

%%% 320
If a character is exposed to a drug via its method of 
application—for example, they pop a pill, slap on 
a dermal patch, are soaked with a splash grenade, 
breathe in gas, or get stabbed with a coated weapon—
then they are subject to the drug's effects. The onset 
time determines how long these effects take to kick in, 
and the duration determines how long they last.

While there is no resistance test to ignore a drug or 
toxin's effects once exposed, in some cases (especially 
toxins) a test might be called for to determine the 
\textit{severity} of the effects. 

Unless otherwise noted or specifically overridden, 
medichines (p. 308) will protect a character from 
drug/toxin effects (but not nanodrugs/nanotoxins). 
Enhancements like toxin filters (p. 305) may also 
impede a drug's effect or provide complete resistance. 
If an antidote is taken in advance or before the effects 
kick in, the drug will not work.

\subsubsection{Addiction And Substance Abuse}

Some drugs are addictive, either physically affecting
the morph) or mentally (affecting the ego)—and
sometimes both. Every time a character uses the drug 
(or after an appropriate amount of use, as determined 
by the gamemaster), they must make a WIL x 3 Test to 
avoid addiction. Each drug has an Addiction Modifier 
that will modify this test.

Failure indicates that the character has become 
addicted—they immediately acquire the Addiction 
negative trait (p. 148). Addiction is measured in three 
levels: Minor, Moderate, and Major. The severity determines
how often an addicted character needs the
drug and what the negative effects of not using the 
drug are.

An addicted character must continue to make WIL 
x 3 Tests as they use the drug, as determined by the 
gamemaster. Failure indicates the character's addiction
severity increases.

The negative effects from not using a drug end 
whenever the character does the drug again. Durability
and Lucidity penalties are not damage, but temporary
decreases to the character's maximum values; the
character immediately regains the lost Durability or 
Lucidity when they do the drug again.

Addiction is of indefinite duration. To clean up, 
the character must stay off the drug for 1 week for 
each level of addiction. Resisting this craving is difficult
and should at least require another WIL x 3
Test, modified by the drug's Addiction modifier. Players
and gamemasters are encouraged to roleplay an
attempt to kick a habit. Each week the character is 
off the drug, the addiction drops by one level. When 
it reaches 0, the character is clean ... though there is 
always danger of a relapse.

Physical addictions do not carry over to a new 
morph if the character resleeves, but mental addictions
do. If the character uploads and resleeves, the
mental addictions persist, and the morph the character
leaves behind remains physically addicted. This
means that poor or unlucky characters may occasionally
find themselves resleeved into a morph that has a
physical addiction. In this case, the character is subject
to the physical addictiveness of the drug but not
the mental addiction, although if they break down 
and indulge in the drug, they may themself become 
physically addicted.

Characters who resleeve as infomorphs can remain 
mentally addicted to a substance despite no longer 
having a body. The market is always happy to provide, 
though; a wide variety of narcoalgorithms mirroring 
the effects of most of the drugs described below are 
available for infomorphs and AIs. For the infomorph-ported
narcoalgorithm version of any physically-only
addictive drug described below, consider the
Addictiveness to be effectively physical. The character 
remains addicted as long as they are an infomorph, 
but they do not remain addicted if they sleeve into a 
physical morph.

\subsection{Drugs}

The drugs described here are usually (but not always 
beneficial), and are typically taken intentionally. Drugs 
and chemicals used offensively are described under 
\textit{Chemicals }and\textit{ Toxins,} both on p. 323.

Note that the drugs here are just a representative 
sampling. There are thousands if not millions of 
drugs in circulation in \textit{Eclipse Phase}—gamemasters 
are encouraged to introduce their own, using these 
as guidelines.

\subsubsection{Cognitive Drugs}

Nootropics and similar drugs are intended to boost 
the user's mental faculties.

\textbf{Drive:} This nootropic speeds up left-right brain 
hemisphere communication, stimulates idea production
and improves concentration, with no usual side
effects. Users receive a +5 bonus to COG while the 
drug lasts. \textbf{[Low]}

\textbf{Klar:} Klar boosts alertness and enhances clarity and 
perception. Users report a feeling of being ``elevated'' 
to a higher level. They receive +5 INT while the drug 
lasts. \textbf{[Low]}

\textbf{Neem:} Neem is a mnemonic drug that works by 
``tagging'' experiences and mental input with a set of 
unique sensations that contribute to the formation 
of state-based memories. Neem gummy chews come 
in a variety of fruit flavors shaped like extinct old 
Earth animals. Neem gives characters a +20 bonus 
on COG Tests to recall information they learned 
while on Neem (see \textit{Memorizing and Remembering,}
p. 176). The drawback to Neem is that memories they 
accumulate while under the drug's influence have no 
emotional association. For example, a character who 
witnessed something horrible happening to a friend 
or who had a fight with a romantic partner while on 
Neem would feel no emotional connection whatsoever 
to what happened. \textbf{[Moderate]}

\subsubsection{Combat Drugs}

%%% 321
Combat drugs are an easy way of evening the odds 
in a fight.

\textbf{BringIt:} In some respects more a social than a 
combat drug, BringIt stimulates massive bursts of aggression
pheromones designed to make the user the
center of attention in a fight. In combat, opponents 
within 3 meters of the character not already in unarmed
or melee combat with another character must
pass a WIL x 3 Test or attack the character using BringIt
The nature of airborne pheromones is imprecise,
however, so if the character using BringIt is within 
1 meter of another character hostile to the character 
affected, the affected character may opt to attack the 
proximate character instead of the BringIt user. Characters
using this drug suffer a –20 modifier on social
skill tests. \textbf{[Low]}

\textbf{Grin:} Grin is an effective opiate and pain suppressant
Users may ignore the –10 modifiers from
2 wounds (not cumulative with similar effects), and 
in fact may not even be aware they are injured. Grin 
users suffer from tunnel vision, however, and so suffer 
a –10 modifier on Perception Tests. \textbf{[Low]}

\textbf{Kick:} Kick is a strong stimulant that increases the 
user's response time and puts them on edge. The character
gains +10 REF and +1 Speed for the duration of
the drug. Characters under the influence of Kick are 
twitchy, however, reacting in a jumpy, cat-like fashion 
to sudden or unexpected stimuli. At the gamemaster's 
discretion, they must make a WIL x 2 Test or react 
without thinking towards unexpected noises or other 
surprises. Long-term users suffer –5 COO. \textbf{[Moderate]}

\textbf{MRDR:  }MRDR is a straightforward and brutal 
combat drug. It increases pain tolerance, speed, and 
strength. The character receives +10 SOM, +1 Speed, 
+10 Durability, and may ignore the –10 modifier of 
one wound. Any damage incurred while under the 
effects of the drug is taken from the bonus Durability
first. MRDR users are easily identifiable by the
broken blood vessels in their eyes, tense posture, and 
visible tension in the muscles of the face, arms, and 
legs. Long-term users suffer –5 SOM. \textbf{[Low]}

\textbf{Phlo:} Phlo increases alertness and coordination, 
making the user more graceful and nimble in a fray. 
The character gains +5 COO and +10 on Perception 
Tests for the duration of the drug. Everything feels 
possible to a character on Phlo, and so they are vulnerable
to being goaded into actions that might be
foolish or dangerous (apply a –10 modifier to appropriate
Social Skill Tests). \textbf{[Moderate]}

\subsubsection{Health Drugs}

Pharma-foods that boost the consumer's health and 
physical state are common.

\textbf{Bananas Furiosas:} This drug reverses some of the 
effects of de-ionizing radiation on the cells of the 
body. Although a pill form is available, it most commonly
comes in large bunches of bright orange-red
bananas. Bananas reduce the severity of a radiation 
dosage (gamemaster determines effect). \textbf{[Low]}

\textbf{Comfurt:} This tasty yogurt treat blocks stress 
%%% 322
hormones, stabilizes mood, and relieves anxiety, allowing
them to ignore the effect of 1 trauma and temporarily
boosting Lucidity by +5. Any stress suffered
while the drug is in effect is taken from the bonus Lucidity
first. Comfurt also provides a +10 bonus when
resisting attempts to manipulate the user's emotions. 
Excessive use of Comfurt can lead to chronic itchiness 
caused by histamine release. \textbf{[Low]}

\subsubsection{Recreational Drugs}

These drugs compete with petals (p. 321) and black 
market XP for wasting people's time and lives away.
\textbf{Buzz:} This gene-modified variant of BZ is an odorless
invisible, extremely powerful hallucinogen. Users
or affected characters will undergo extremely realistic 
hallucinations for the duration, and may even ``share'' 
hallucinations with other affected characters. Characters
will suffer a –30 modifier to any tests to remember
what occurred while under the influence. \textbf{[Moderate]}
\textbf{Mono No Aware:} Taken from the Japanese term 
for sadness at the ephemerality of worldly things, this 
drug, typically ingested as a tea, is a depressant that 
induces a meditative state. Mono No Aware gives the 
character a +10 bonus on Art and Sense Tests. With 
frequent use, Mono No Aware reacts with pigments 
in the skin to create a pallor with a slight bluish tinge, 
even in darker-skinned morphs. \textbf{[Low]}
\textbf{Orbital Hash:} Good ol' reefer—but grown in space 
using powerful lighting and post-singularity hydroponics
Because space is at a premium in habitats and scum
barges, blocks of hashish are the preferred mode of 
transport and delivery. However, for the wealthy and 
\textbf{Cognitive Drugs}
Drive

Chem
Klar

Chem
Neem

Chem
\textbf{Combat Drugs}
BringIt

Bio

Inh, Inj, O
Grin

Chem

Inh, Inj, O
Kick

Chem

Inh, Inj, O
MRDR

Chem
Phlo

Chem
\textbf{Health Drugs}
Bananas Furiosas

Chem
Comfurt

Bio
\textbf{Recreational Drugs}
Buzz

Chem
Mono No Aware

Chem
Orbital Hash

Chem
\textbf{Social Drugs}
Alpha

Bio
Hither

Bio
Juice

Chem
on planets, buds in leaf form are not uncommon. Hash 
allows the character to ignore the effects of 1 trauma, 
but inflicts a –10 penalty on all memory-related tests 
and Knowledge Skill Tests. Hash users exhibit bloodshot
eyes, lethargic behaviors, and the munchies. \textbf{[Low]}

\subsubsection{Social Drugs}

These social lubricants affect the user's interactions 
with others.
\textbf{Alpha:} Alpha is a more subtle version of BringIt, 
popular with hypercorp execs, street thugs, and 
anyone else who wants to come across as a domineering
asshole. The pharm designer who invented it had
a retro sensibility (and maybe a sick sense of humor); 
Alpha is typically synthesized as a sparkling white 
powder designed to be snorted. Alpha stimulates 
production of threat pheromones, but less bluntly 
than BringIt. Alpha imparts confidence, a feeling of 
power, and alertness. Users can function without sleep 
for 4 days, after which point they need to catch up 
with at least 4 hours of sleep (remember morphs with 
basic biomods require less sleep). Dosed characters 
receive a +20 modifier on Intimidation Tests and +10 
on Persuasion and Networking Tests where attitude is 
a factor (gamemaster discretion). These bonuses only 
apply to characters within 2 meters of the Alpha user.
On the downside, alpha users are impatient, unfocused
assholes. At the gamemaster's discretion, Social
skill modifiers may be reversed to penalties with 
certain types of people. Additionally, Alpha users 
suffer –10 on all COG skill tests related to memory 
and coherent or logical thinking. Long-term users may 
O
20 minutes
8 hours

—

Mental
O
20 minutes
8 hours

—

Mental
O
20 minutes
12 hours

—

Mental
1 minute
15 minutes

+10

Physical
3 Action Turns
3 hours

–10

Physical
3 Action Turns
2 hours

–10

Physical
O
20 minutes
1 hour

–10

Physical
O
20 minutes
1 hour

–10

Physical
O
20 minutes
1 day

—

—
O
20 minutes
12 hours

–10

Mental
Inh, O
1 hour
36 hours

—

Mental
O
20 minutes
8 hours

–10

Mental
Inh
3 minutes
3 hours

—

Mental
Inh
1 minute
2 hours

–10

Mental
D
1 minute
6 hours

–10

Physical
O, Inh
20 minutes
8 hours

—

Mental
%%% 323
suffer the COG penalty even when not on the drug; 
on it, they may be worse. \textbf{[High]}
\textbf{Hither:} Want to ooze sexy like a pleasure morph 
on a hot tin roof? For those desiring that slinky je-ne-sais-quoi
Hither is the tool. Hither is a clear, slippery
gel, sometimes with a faint, musky, floral scent. Hither 
is applied to parts of the body with large concentrations
of sweat glands, where the skin quickly absorbs
it. Hither is a mild euphoriant, imparting a feeling of 
confidence and you-know-you-want-it-ness to the user. 
It also stimulates abundant production of lust pheromones
The character gains a +10 bonus on Persuasion
Tests against targets who are possible to seduce. At the 
gamemaster's discretion, this extends to Deception, 
Impersonate, and Networking Tests. \textbf{[Low]}
\textbf{Juice:} This potent anti-depressant makes it almost 
impossible to have bad feelings or negative thoughts. 
The character is unnaturally happy—often irritatingly 
or strangely so. The character receives a +30 bonus 
against fear or attempts to manipulate their emotions 
in a negative direction, but is also likely to act inappropriately
like giggling over the massive amount of
spilled blood or cheerfully changing the subject to 
inane topics when someone else is freaking out. \textbf{[Low]}

\subsection{Nanodrugs}

Nanodrugs are temporary nanobot infestations that 
apply a specific effect.
\textbf{Frequency:} Frequency (or Freeq) is a nanodrug designed
as a tool for scientific visualization. It releases
a small swarm of nanobots into the character's bloodstream
that settle in the epidermis, where they act as
sensors of electromagnetic radiation. This sensory 
input is then injected into the character's visual and 
tactile sensoria, hitting the user with a sequence of 
novel stimuli, typically a light show or weird tactile 
sensations. Aside from its recreational uses, Frequency 
is good at picking up on localized field radiation with 
a standard Perception Test. A character can take advantage
of this to spot sensors and hidden electronics.
Similar to now-obsolete 20th-century hallucinogens 
like LSD and psilocybin, however, a Frequency trip 
can be disorienting and upsetting (the gamemaster 
should apply any modifiers, mental stress, or even 
trauma as they feel appropriate). Characters typically 
experience a period about 1/3 of the way through 
their trip in which sensory input is extremely intense; 
during this period, which usually lasts about 2 hours, 
they are unable to read. \textbf{[Moderate]}
\textbf{Gravy:} Gravy assists characters in acclimating to 
Frequency

Nano
Gravy

Nano
Petals

Nano
Schizo

Nano
high gravity environments. It comes in a variety of 
flavors and is often added as a sauce to food. For 
Gravy to be 100% effective, the character must begin 
using it in advance. Reduce penalties for high-gravity 
acclimation by 20. \textbf{[Low]}
\textbf{Schizo:} Schizo is a nanodrug that mirrors the effects
of paranoid schizophrenia. It is popular in
some hyperelite social circles as a truly daring and 
intriguing experience. A dose of schizo looks like a 
disposable antique razor blade. Making an incision 
in the skin releases a swarm of nanobots that travel 
to the central nervous system and induce the effects 
of the drug. While in effect, the character is severely 
paranoid and hears voices. How this plays out is at 
the discretion of the gamemaster, but should include 
irrational fears, unusual compulsions based on the 
instructions of the voice or voices, and a strong possibility
that the character will behave in a violent or
destructive fashion. The character may make WIL x 3 
Tests to avoid violent acts against objects or strangers. 
Friends and trusted acquaintances are probably less 
likely to be targets of violence (+30 modifier to avoid 
hurting people the character cares about or destroying
important possessions). Note that the character's
muse is unaffected by Schizo and can make efforts 
to babysit the character. Characters who take Schizo 
suffer 1d10 mental stress. \textbf{[Low]}

\subsubsection{Petals}

Inj, O
8 hours

–10

Mental
Inj, O
special

—

—
O
2 hours–1 day

+10 to –20

Mental
Inj
1 day

—

Mental
%%% 324
Petals is a term for a type of narrative hallucinogen, 
a nanodrug that hijacks the senses and takes the user 
on a game-like, highly immersive trip. Known by a 
myriad of intriguing names—Forgotten Hand, Darkly 
Selving, Inquisitive Green, to name a few—Petals are 
post-Fall society's heroin—the drug of choice for the 
desperate and fucked. Petals almost always appear as 
nanopharmaceutical flowers, potted or with a nutrient 
pack attached to the stem. Plucking and swallowing 
the petals from the flower triggers the effects immediately
Flowers have 5-10 petals. Multiple users may
share the experience if they take the Petals within 1 
minute of the first one being plucked; after this all 
petals remaining on the flower fade to translucent 
white and become inert.

Petal experiences are like entire scenarios in and of 
themselves. Some take place entirely in the user's mesh 
inserts (the user must cede control of their implants 
voluntarily; if they do not, the drug has no effect other 
than producing very low-intensity LSD-like visual hallucinations
taking control of the character's entoptic
displays, linking to secretive mesh servers and other 
trippers, and invading the character's sensorium with 
AR ``hallucinations.'' Others put the character into a 
near-comatose state during which they go on a head 
trip. Normally there is some kind of well-developed 
theme or plot to a Petal experience, although in some 
cases they just experience a stream of images.

Though most societies seek to suppress Petals, new 
ones appear constantly, fueled by a persistent subculture
of crafters and users. Petalcrafters view their
work as an art form (or at least as really good entertainment
and the better Petals are lovingly crafted,
hauntingly beautiful experiences—even if they're also 
terrifying. The subculture of Petal use ranges from 
casual users who occasionally do an easy, short-duration
flower to hardcore addicts who spend much of
their time not on Petals trying to hunt down the most 
%%% 325
intense and esoteric varieties. From this subculture 
comes a lot of information on what various Petals 
look like and their effects. Because Petals combine 
custom nanobots with tailored chemical payloads and 
sometimes connections to mesh servers, duplicating 
them using fabricators is impossible, leading to an 
active market of crafters, dealers, and traders.

Petals sometimes contain easter eggs and rewards, 
called ``sweets'' by petal users. Getting the sweets usually
requires fulfilling certain conditions within the
trip, such as correctly answering questions or fulfilling 
goals. Typical sweets include skillsofts, new clothing 
or product designs, and custom infomorph sleeves.

On the negative side, some Petal trips go bad, inflicting
1d10 mental stress or more on the user. Perhaps
worse, some Petals are loaded with malware that
takes over the user's mesh inserts and worse—some 
sentinels even whisper of Petals carrying strains of the 
Exsurgent virus. \textbf{[Trivial to High]}

\subsection{Narcoalgorithms}

Narcoalgorithms are software programs that simulate 
the effects of drugs on biological bodies. Almost all 
bio, chemical, and nano drugs can be replicated as 
narcoalgorithms, with corresponding effect gamemaster
discretion). Narcoalgorithms may be run by
infomorphs, egos encased in cyberbrains (pods and 
synthmorphs), simulmorphs, and even AIs.

\textbf{DDR:} Originally crafted by prankster hackers 
and distributed as a virus, DDR (for ``Dance Dance 
Robot'') triggers impulses in the target's motor control
circuits. Primary targeting robot AIs, the effect is
that targets ``dance'' in jerky, automated movements. 
Pleasure receptors are also activated so that dancing—
and movement of any kind—feels good. Different 
software variants invoke different motions and styles. 
The target suffers a –20 modifier on other actions 
while dancing, but the dancing may be overridden 
with a WIL x 3 Test. \textbf{[Low]}

\textbf{Linkstate:} This software actually connects the user 
to a peer-to-peer network, where it randomly connects
to other linkstate users and samples a bit of their
XP feed and randomly accessed memories—typically 
just enough to provide context, but not enough to 
acquire private personal details. These inputs are 
spliced together, their emotional inputs amplified, and 
then the entire package is spiked with some hormonal 
circuit triggers and artificial synaesthesia. The effect 
is a mind-blowing mixed sampling of people's lives, 
mashed together in a sensory soup, that hits the mind 
with a euphoric rush. Linkstate users are catatonic 
while under the effects (typical sessions run 3-4 hours), 
but afterwards they often report that they have flashbacks
of events in other people's lives. \textbf{[Low]}

\subsection{Chemicals}

\textbf{Atropine:} Though poisonous in large doses, atropine 
is an effective antidote against nerve agents like BTX
and Nervex. Easily synthesized in a maker, atropine 
will avert the effect whether taken soon before or after 
dosage by a nerve agent. \textbf{[Trivial]}

\textbf{DMSO:} This chemical acts as a carrier, allowing
other chemicals to be absorbed through the
skin. It allows any chemical agent to be applied 
dermally. \textbf{[Trivial]}

\textbf{Liquid Thermite:} Similar to scrapper's gel, liquid 
thermite comes in a gel form that is easily applied 
under all environmental conditions (by the nature 
of its chemical reaction, thermite is oxygenated and 
will burn underwater or in space). It is ignited with 
an electric charge, burning at temperatures exceeding 
2,500 degrees Celsius and melting through whatever 
it is touching. Liquid thermite inflicts 3d10 + 5 DV per 
Action turn to whatever it is touching. Armor will also 
be burnt through, offering no protection once the full 
Armor rating has been reached. \textbf{[Moderate]}

\textbf{NotWater:} NotWater is an effective liquid fire retardant
that does not get objects wet, no matter how
absorbent they are—it simply beads up and slides 
right off. \textbf{[Trivial]}

\textbf{Scrapper's Gel:} This goo turns into a potent acid 
when given an electrical charge. It comes in a gel-like 
state and may be smeared like jelly, and may even be 
used in space. In acid form, scrapper's gel does 1d10 
+ 5 DV per Action Turn to anything it touches, unless 
the material has been treated against acid. Armor will 
protect against this acid at first, but the acid will eat 
through the armor, so that it will no longer protect 
after its full armor value has been reached. \textbf{[Low]}

\textbf{Slip:} This liquid is almost entirely frictionless. When 
spread around an area (commonly used in splash 
grenades), anyone attempting to walk or run on the 
affected surface must make a COO Test or fall down. 
Likewise, any coated surface becomes extremely 
hard to grip onto, requiring a SOM Test to hang on. 
Anyone attempting to grapple a slip-soaked character 
suffers a –30 modifier. \textbf{[Low]}

\textbf{Tracker Dye:} This liquid is colorless at normal light 
but becomes recognizable under pre-specified different 
wavelengths (such as infrared or ultraviolet). \textbf{[Trivial]}

\subsection{Toxins}

Chemical warfare involves using the toxic properties 
of biological and chemical substances to kill, injure, 
or incapacitate an enemy. Note that an antidote can 
be constructed for most toxins if a sample is acquired 
and an appropriate Medicine or Academics Test is 
made. This is considered a Task Action with a timeframe
of 1 hour. These toxins only affect biomorphs;
synthmorphs are immune.

\textbf{BTX}

\textbf{:} BTX-squared (also called Frog Bite) is a 
genetically-enhanced variant of the extremely potent 
cardiotoxic and neurotoxic batrachotoxin. It leads to 
fast paralysis and cardiac arrest that usually kills the 
target within a few Action Turns. Affected characters 
suffer 2d10 + 10 damage a turn for 3 Action Turns; 
medichines reduce this damage by half. They must 
also make a SOM x 2 Test (+30 with medichines) or 
be paralyzed for 1 hour. \textbf{[High]}

\textbf{CR Gas:} This potent incapacitating agent causes 
%%% 326
eye twitching and temporary blindness, severe coughing
and breathing difficulty, skin irritation, and panic.
Affected characters suffer 1d10 ÷ 2 damage, a –30 
modifier to sight-based Perception Tests, and a –20 
modifier to all other actions for 20 minutes (5 minutes
if the character has medichines). \textbf{[Low]}

\textbf{Flight:} This drug is derived from human pheromones 
released due to fear, and is intended to instill alarm or 
even terror in the character. Affected characters must 
make a WIL x 3 Test (+30 with medichines) or suffer 
a panic attack, inflicting 1d10 stress. Dosed characters 
also suffer a –30 modifier for resisting intimidation or 
fear-based emotional manipulations. Flight affects last 
for 1 hour (5 minutes with medichines). \textbf{[Low]}

\textbf{Nervex:} Derived from deadly nerve agents like cyclosarin
VX, and novichok, this genetically-modified
toxin is deployed as a colorless, odorless gas that 
turns safely inert 10 minutes after deployment. It 
causes involuntary contraction of the muscles, seizures
and death by respiratory failure. One minute
after exposure, the character must make a SOM Test 
or be incapacitated by seizures, paralysis, or nausea 
and vomiting; unaffected characters still suffer a –20 
modifier to all actions. After 10 minutes, the character 
will die unless an antidote (such as atropine, p. 323) is 
applied. Characters with medichines suffer the initial 
effects, but recover after 5 minutes. \textbf{[High]}

\textbf{Oxytocin-A:} A genetically-improved variant of oxytocin
this drug induces trust in the recipient. Drugged
characters suffer a –30 modifier on all WIL and Kinesics
Tests where trust is a factor. Medichines provide
immunity. \textbf{[Low]}

\textbf{Twitch:} Twitch is a convulsive agent, a nonlethal 
nerve gas. Affected characters must succeed in a SOM 
Test (+30 with medichines) or become incapacitated 
with severe muscle tremors. Unaffected characters still 
suffer a –20 on all actions. The effects of Twitch last for 
10 minutes, 5 if the character has medichines. \textbf{[Low]}

\subsection{Nanotoxins}

\textbf{Disruption:} This nanotoxin attacks the myelin sheath 
on nerves, disrupting nerve impulses and inflicting 
symptoms of multiple sclerosis. Every hour the morph 
suffers a –5 modifier to COO, REF, and COG. If any 
aptitudes are reduced to zero,the morph is effectively 
paralyzed and catatonic. \textbf{[Moderate]}

\textbf{Necrosis:} Necrosis nanobots attack the walls of cells 
inside the body, killing tissue. This nanotoxin inflicts 
1d10 ÷ 2 damage per Action Turn for one minute, after 
which the nanobots disable and flush from the body. 
Necrosis only affects biomorphs. \textbf{[Moderate]}

\textbf{Neuropath:} These nanobots are designed to stimulate
the pain receptors of a morph on a systemic level
to cause agony and impairment. While most neuropaths
target biological receptors, variants are available
that induce comparable (phantom) pain stimulations
in the cyberbrains of synthmorphs to create an
equivalent effect. The affected character must succeed 
in a WIL x 3 Test or become incapacitated. Even if 
they succeed, they suffer –30 from the inflicted agony. 
Any form of pain resistance that allows a character 
to ignore wound modifiers will negate the neuropath 
pain modifier by an appropriate amount. \textbf{[Moderate]}

\textbf{Nutcracker:} Nutcrackers are nanobots designed to 
locate, migrate, and decompose the synthdiamond 
case of a cortical stack within a morph by attacking 
its crystal lattice. This process takes approximately 
6 hours, after which the cortical stack is destroyed. 
These nanobots also attack the cortical stack's connections
to the (cyber)brain and brain-mapping nanobots.
After 1 hour, the victim will be aware that their cortical
stack is threatened. After 3 hours, all connections
will be severed and the cortical stack will no longer be 
able to back up the character. \textbf{[High]}

\subsection{Pathogens}

A pathogen is an infectious biological agent that 
causes disease or illness to its host. While natural 
pathogens rarely strive to kill their hosts, germ warfare
programs revived during the Fall—or instigated
by the TITANs—sought to modify and use pathogens 
as a weapon of war. The ideal characteristics of lethal 
biological agents are high infectivity, high potency, 
availability of vaccines, and delivery as an aerosol. 
Most biomorphs are immune to standard pathogens 
thanks to their basic bio-mods, and medichines will 
protect against most others. However, even these 
defenses may not protect against diseases left by 
the TITANs or a new terrorist cell's biowar bug. It 
is largely recommended that pathogens be handled 
as a plot device, rather than an active threat to the 
characters. Pathogens have no effect on synthmorphs.

\textbf{Degen:} Characters exposed to this degenerative 
neurological disease must make a DUR x 2 Test or 
become infected. Medichines will defeat the disease, 
but others will not show signs of infection for 1 week, 
when the symptoms of a rapidly progressing dementia 
will become clear: memory loss, personality changes, 
and hallucinations. If untreated, Degen will progress 
for another week with more serious symptoms, including
speech impediments, jerky movements, loss
of balance and coordination, and even seizures. This 
is reflected by a 5 point loss in all aptitudes per day 
(after the first week). When any aptitude reaches 0, 
the character dies. Degen is notorious for its effect 
in corrupting cortical stack backups before infection 
symptoms manifest. \textbf{[Expensive]}

\textbf{Trigger:} Trigger is a designer virus that selectively 
targets and infects mast cells to trigger a hyper-allergic 
reaction. The resulting anaphylactic shock due to systemic
vasodilatation (associated with a sudden drop
in blood pressure) and bronchial swelling (resulting in 
constriction and difficulty breathing) usually leads to 
death in a matter of minutes after onset, if not treated. 
Infected characters must succeed in a DUR Test (using 
their current Durability score minus damage) or die 
quickly. Even medichines have difficulty reacting in 
time against this virus; characters with medichines 
must make a DUR x 2 Test to survive. \textbf{[Expensive]}

\subsection{Psi Drugs}

%%% 327
Research into the Watts-MacLeod strain has resulted in 
several exceptional breakthroughs involving the creation 
of psi-impacting drugs. Each of these drugs is in the experimental
stage, but they are already finding some use
among Firewall and similar secretive groupings.

\textbf{Inhibitor:} Inhibitor is a cocktail of neurochemicals 
that block some brain receptor and transmitter functions
in an attempt to reduce psi-waves and block or
impair sleights. This drug is commonly used to restrain 
async prisoners from using their abilities. A drugged 
character must make a WIL x 2 Test. If they fail, they 
lose all psi abilities for the drug's duration. If they succeed
they suffer a –30 impairment on Psi skills and
all strain is doubled. Inhibitor has an unfortunate side 
effect of doping the character down, however; apply 
a –10 modifier to their COG. Inhibitor-influenced 
characters tend to have a glazed, dopey expression and 
have difficulty getting excited or emotional. \textbf{[High]}

\textbf{Psi-Opener:} Psi-opener drugs are variants of the 
Watts-MacLeod strain with a temporary effect and 
which do not permanently alter the user's brain. Psi-opener
temporarily imbues the user with the ability
to use one particular sleight, regardless of whether or 
not they have the Psi trait. Each type of Psi-opener is 
customized for a particular sleight. While primarily 
intended for non-asyncs, non-asyncs may not possess 
Psi skills, so they must default to WIL. For this reason, 
Psi-Opener is often doubled up with Psike-out.

Using Psi-opener is a mind-wrenching experience. 
Users are occasionally subject to hallucinations 
(gamemaster discretion). When the drug wears off, it 
inflicts 1d10 points of mental stress, +2 if the drug 
imbues a psi-gamma sleight. \textbf{[Expensive]}

\textbf{Psike-Out:} Psike-out bolsters an async's psi abilities. 
Apply a +20 modifier to the async's Psi skill tests for the 
drug's duration. However, also apply +2 to all strain 
DVs for the drug's duration. Psike-out is mentally addictive
with an Addiction Modifier of –10. \textbf{[Expensive]}

\section{Everyday Technology}

The following devices are all exceptionally common 
and can be acquired in almost any habitat. Almost 
everyone in \textit{Eclipse Phase} either owns these devices or 
knows several people who do.

\textbf{Ecto:} Ectos are the external version of basic mesh 
inserts (p. 300), minus the medical sensors. These colorful
devices serve as a wearable mesh terminal, PDA,
locator, and camera-phone. The devices are flexible 
(often worn as bracelets), dirt-resistant, self-cleaning, 
and may be stretched out to increase screen size. They 
may project holographic displays and are typically 
equipped with wireless-enabled glasses or contact 
lenses and decorative earpieces or earrings so that the 
user may access augmented reality. Given the ubiquity 
of mesh inserts, ectos are growing less common, but 
they are still used by bioconservatives, others without 
implants, and those who prefer to access the mesh via 
an external device for security concerns. \textbf{[Low]}

\textbf{Holographic Projectors: }These devices are capable 
of projecting high-definition, ultra-realistic three-dimensional
images and movies. From a distance (20+
meters), such holograms can be difficult to distinguish 
as fake, but up close they are easier to see for what they 
are (+20 Perception Test modifier). Holograms do not 
appear wavelengths other than visual light, and so are 
easily identified by anyone with enhanced vision. \textbf{[Low]}
\textbf{Micrograv Shoes:} These shoes are equipped with 
velcro and/or a magnetic system, allowing the wearer 
to walk normally on appropriate surfaces in micrograv
and zero-G environments, rather than floating or
bouncing. \textbf{[Trivial]}
\textbf{Portable Sensor:} This is a small portable (possibly 
even wearable) sensor system. The type of sensor must 
be chosen (for example: infrared, lidar, radar, x-ray). 
Combined sensor systems are also available, at a cumulative
cost. See \textit{Radio and Sensor Ranges,} p. 299.
and \textit{Using Enhanced Senses,} p. 302. \textbf{[Moderate]}
\textbf{Smart Clothing: }Smart clothing can change its 
color, texture, and even its cut, taking only a minute 
or two to transform from a solid color jumpsuit to a 
plaid party dress or a replica of a pinstriped, late 20th 
century business suit. It can also camouflage the wearer, 
providing a +20 bonus to Infiltration Tests to avoid 
being seen, as long as the wearer is stationary or not 
moving faster than a slow walk, and as long as the 
wearer is completely covered or also using chameleon 
skin (p. 303) of the same color/pattern. If incompletely 
camouflaged, or if moving faster, reduce the modifier 
to +10. Smart clothing also keeps the character warm 
or cool, allowing the character to exist comfortably in 
environments from –40 to 70 C. \textbf{[Low]}
\textbf{Smart Vac Clothing:} Just like regular smart clothing
this outfit can also transform into a light vacsuit
(p. 333). It also functions as armor with a rating of 
2/4. \textbf{[Moderate]}
\textbf{Specs:  }Specs are vision-enhancing glasses. They 
\textbf{Chemical Toxins}
BTX

Chem

D, Inj, O

1 Action Turn 3 Action Turns/1 hour
CR Gas

Chem

D, Inh

1 Action Turn

20 minutes
Flight

Bio

Inh

3 Action Turns

1 hour
Nervex

Chem

D, Inh, Inj, O

1 minute

death
Oxytocin-A

Bio

Inh, Inj

3 minutes

2 hours
Twitch

Chem

D, Inh, Inj, O 3 Action Turns

10 minutes
\textbf{Nanotoxins}
Degeneration

Nano

Inj, O

Immediate

8 hours
Necrosis

Nano

Inj, O

3 Action Turns

1 minute
Neuropath

Nano

D, Inj, O

3 Action Turns

8 hours
Nutcracker

Nano

Inj, O

Immediate

6 hours
\textbf{Psi Drugs}
Inhibitor

Chem

Inj, O

3 Action Turns

6 hours
Psi-Opener

Bio

Inj, O

20 minutes

1 hour
Psike-Out

Chem

Inj, O

1 minute

1 hour
%%% 328
deliver sensory data directly into the wearer's visual 
cortex by connecting with their basic mesh inserts 
(p. 300), though visual displays are available for bioconservatives
and other characters without implants.
Specs extend the range of the wearer's vision from 
terahertz waves to gamma rays (p. 302). Specs include
a t-ray emitter (p. 306), however, using x-rays,
or gamma rays for visual purposes requires a separate
emitter, since neither of these sorts of radiation
are common inside habitats, or in any safe environments
Specs have a variable focus equivalent to 5
power magnifiers and provide the wearer with a +10 
bonus to all Perception Tests involving vision. \textbf{[Low]}

\textbf{Tools: }Tools come in kits (portable), shops (can fit 
into a large vehicle), and facilities (large, non-mobile). 
Each set of tools applies to a particular skill, such 
as Hardware: Electronics or Hardware: Groundraft. 
\textbf{[Low (Kit), High (Shop), Expensive (Facility)]}

\textbf{Utilitool:} This hand tool includes a specialized 
small nanobot generator. In its basic form, a utilitool 
is the size and shape of a large fountain pen. It can 
transform into almost any tool, however, from a 
wrench, knife, or powered screwdriver to a rotary 
grinder or pair of pliers. Some inexpensive utilitools 
are optimized for specialized tasks, like cooking 
or wilderness survival, but more expensive models 
become almost any imaginable hand tool. Utilitools 
are normally mentally controlled using the character's 
basic mesh inserts. Characters without such implants 
can control the tool via voice commands and touch 
controls. Characters using a utilitool gain a +10 
modifier to skills involving repairing or modifying devices
with mechanical parts, opening locks, disarming
alarm systems, or performing first aid. \textbf{[Low]}

\textbf{Viewers: }These small and highly advanced binoculars 
possess all the visual enhancement of specs (p. 325), 
but also provide 50x magnification. They also include 
a directional microphone that magnifies sound from 
the direction the viewers are pointed by a factor of 50. 
Viewers provide the user with a +30 bonus to all Perception
Tests involving vision or hearing for the target they
are aimed at. This bonus is not cumulative with bonuses 
from any other device or augmentation. \textbf{[Low]}

\section{Nanotechnology}

Nanotechnology is the precise manipulation of matter 
at the atomic level, typically using millions of micro-scale
nanomachines. Nanotechnology transformed
manufacturing, enabling new techniques and materials
The advent of nanofabrication—building objects
from the molecular level up—transformed economies, 
allowing people to simply manufacture whatever they 
needed from raw materials. Nanotechnology is still a 
growing field, however, and has its limitations. While 
the TITANs unleashed self-replicating nanoswarms 
with the ability to transform or destroy anything 
through the power of geometric growth, such technology
remains far beyond transhumanity's grasp.

\subsection{Basic Nanotechnology}

Basic nanotechnology is exceedingly widespread 
and used throughout the solar system, serving as the 
primary method of manufacturing for decades. The 
nanobots of basic nanotech are confined to delicate 
and specially-maintained environments like the 
insides of cornucopia machines or healing vats and 
cannot operate elsewhere.

\subsubsection{Healing Vats}

Healing vats were the first type of nanotech medicine 
developed and remain the most powerful medical 
devices in common use. With the exception of a few 
exceptionally deadly nanoplagues, a healing vat can 
cure any disease and heal any injury. As long as the 
patient is alive when they are place in the healing 
vat, they will not only survive, but emerge without a 
scratch. A healing vat can even take a severed head (as 
long as it has been stabilized by medichines or nanotech
first aid) and regrow an entire body based on the
head's genetics. If the patient's body or medical records 
contain information about their implants, bioware, or 
%%% 329
advanced nanotechnology, all of these modifications 
are also fully restored.

Few people suffer injuries serious enough to require 
a healing vat. Most are used as a safe and easy way 
to perform bodysculpting or to install implants or 
bioware. Healing vats use specialized nanomachines 
to either alter the patient's body or integrate implants 
or bioware. One advantage of using a healing vat is 
that no additional healing time is needed, the patient 
leaves the vat fully recovered from the augmentation 
and ready to go. Every hospital, clinic, bodyshop, and 
augmentation parlor has several healing vats. The 
time required by a healing vat varies with the severity
of the damage it is healing or the extent of the
modification being made, as noted on the Healing Vat 
table, p. 327. \textbf{[High]}

\subsubsection{Nanodetectors}

Nanodetectors are small devices that suck in air and 
micro debris in order to scan for and detect nanobots
Given that nanobots are so small, the density of
nanobots in the area has a large impact on its success. 
The nanodetector has a base skill of 30 for detecting 
nanobots, modified by +30 if an active nanoswarm or 
hive is present, +0 if a nanoswarm or hive was active 
recently, and –10 for the presence of nanobots outside 
of a swarm or hive. Once a nanobot is detected it may 
be analyzed either by the user or the nanodetector's AI, 
using Academics: Nanotechnology 30 skill. Nanodetectors
are often worn and left on, set to alert the user
if a hostile nanoswarm is detected. \textbf{[Low]}

\subsubsection{Nanofabricators}

Nanofabrication machines are universal assemblers 
that perform almost all of the manufacturing in the 
solar system. The user loads in raw materials and electronic
plans and it can produce literally any manufactured
good, from a weapon to an ultralight plane to a
Healing normal damage to a character who has taken 3 or fewer wounds. 
Restoring major lost body parts like arms or legs, or healing dying or nearly dead character 
who has taken 4 wounds. 
Restoring recently dead character who was placed in medical stasis to avoid death, but 
who is mostly intact.
Restoring recently dead character who is placed in medical stasis to avoid death,
and who is missing most of their body.
\textbf{Augmentation}
Minor implants and bioware, minor cosmetic changes like alterations in skin color, eye 
color or shape, or hair color, texture or distribution, minor alterations to face shape or 
body fat distribution.
Major brain and neural implants, nanoware or bioware, sex changes, changing height by 
no more than 5% or weight by no more than 20%.
Major physical modifications like adding limbs or radical changes to height and weight.
hot and delicious dinner. Many nanofabricators come 
equipped with a library of common-use blueprints 
(basic foods, standard clothing, common tools, etc.). 
Other blueprints must either be purchased online, self-programmed
or acquired through some other method
(see \textit{Nanofabrication,} p. 284). The largest nanofabrication
units are more than 10 meters on a side and are
used to produce small consumer goods in bulk as well 
as building large devices like orbital transfer vehicles.
The availability and legality of nanofabricators 
varies widely throughout the system. In the inner 
system and Jovian Republic, cornucopia machines 
are commonly restricted and sometimes illegal, with 
licenses only available to hypercorps, military units, 
and other officials and elites. In these habitats, only 
more limited fabbers are available to the general 
populace. Additionally, blueprints are licensed and 
protected by copyright laws, and many nanofabricators
feature pre-programmed restrictions that prevent
them from using unlicensed blueprints as well as 
from manufacturing weapons, explosives, or other 
restricted items. Among the autonomists of the outer 
system, however, nanofabricators are commonly accessible
shared by everyone, and unrestricted.
For rules on creating goods in a nanofabricator, see 
\textit{Nanofabrication,} p. 284.
\textbf{Desktop Cornucopia Machine:} Cornucopia machines
(CMs) are general-purpose nanofabricators.
The smallest CMs are desk-sized cubes approximately 
half a meter on a side with a volume of at least 40 
liters. They can produce any small object, from tools 
to well-folded suits of clothing to handguns or a bowl 
of cereal. It is sometimes possible to assemble larger 
items, but they must be manufactured in smaller 
pieces and then assembled (likely requiring an appropriate
Hardware Test).
While users can purchase bulk raw materials, CMs 
also come equipped with a disassembler. The user 

2 hours per wound 

(min. 1 hour for 0 wounds)

12 hours per wound

1 day per wound

3 days per wound

1 hour

12 hours

3 days
%%% 330
loads garbage and other objects into the disassembler 
so that they can be turned into raw materials for the 
CM. All legally-available disassemblers only deconstruct
non-living material. \textbf{[Expensive]}

\textbf{Fabber:} Fabbers are specialized nanofabricators, 
portable and considerably smaller than CMs. There 
are a wide variety of portable fabbers, including 
ones that can make any hand tool or small piece of 
personal electronics, ones that can turn any organic 
material into food and drink, and ones that can create 
any drug or medicine as well as bandages and specialized
dressings. The most common fabbers have a
volume of 4 liters. Larger hand tools and devices are 
produced as 2 or 3 separate parts that must be fitted 
together. Like CMs, fabbers also contain miniature 
disposal units. \textbf{[Moderate]}

\textbf{Maker:} Makers are specially-designed to produce 
food and drink for the user. Raw materials can be provided
by the addition of any water-containing liquid
and collected biomass like leftover food, grass, dirt, 
dead animals, or transhuman waste. Some models are 
built into standard vacsuits. Makers can produce water 
and various flavored beverages, as well as ration bars 
or thick pudding-like edible gels. With adequate raw 
material, a maker can indefinitely provide food and 
drink for up to three transhumans. Most units, however
have a very limited range of flavors and textures
that are widely considered to be fairly bad. Models 
with a wider and better range of flavors and textures 
are more expensive, but produce food that is considered
adequate or occasionally good. \textbf{[Low to Moderate]}

\textbf{Blueprints:} If you want a nanofabricator to make 
something, you need to instruct the device how to 
create it from the molecular level up. Such blueprints 
are available for almost every conceivable item out 
there. The cost of such blueprints typically exceeds 
the cost of purchasing the item, though factors like 
legality and quality may affect the cost as usual (see 
\textit{Acquiring Gear,} p. 296). \textbf{[One Cost Category Higher }
\textbf{Than Item Cost]}

\subsection{Advanced Nanotechnology}

Advanced nanotechnology includes more recent 
developments. Like basic nanotech, advanced nanotechnology
cannot self-replicate but the nanobots
can function normally in most environments and are 
highly resistant to bacterial attacks and other environmental
problems. Typical advanced nanotech consists
of a generator—known as a ``hive''—that produces 
nanobots as long as it is supplied with raw materials. 
Every such hive also includes a miniature disassembly 
unit and/or specialized nanomachines that collect 
raw materials for the generator. These hives produce 
nanobot swarms that are set loose to perform some 
function in the world.

Examples of advanced nanotech include COTs (p. 
315), medichines (p. 308), smart dust (p. 316), and 
utilitools (p. 326), among others.

\textbf{General Hive:} General hives are capable of producing
any conceivable type of nanobot with the right
blueprints and/or programming. Even at their smallest 
size they are not really portable, with a minimum size 
being cubes 30 centimeters on a side and a volume of 
25 liters. \textbf{[Expensive]}

\textbf{Specialized Hive:} Specialized hives are far more 
common than general hives, though they can produce 
only one type of nanomachines (i.e., choose one type 
of nanoswarm per hive). The smallest specialized 
hives are approximately the size of a 12-gauge shotgun
shell or a large cherry tomato. \textbf{[Moderate, plus }
\textbf{Cost of Programmed Nanoswarm]}

\subsubsection{Ego Bridges}

Ego bridges are vat devices used for uploading and 
downloading minds. See \textit{Backups and Uploading,} p. 268, 
and \textit{Resleeving}, p. 271. \textbf{[Expensive]}

\subsubsection{Nanoswarms And Microswarms}

Swarms are colonies of nanobots or larger microbots 
created in a hive, programmed with specific instructions
and then set free to perform a set task. Each
swarm is composed of hundreds or thousands of 
nanobots or microbots, ranging in size from a microbe 
to a small insect. Nanobots are typically invisible to 
the naked eye, though they can be detected with a 
nanodetector (p. 326) or nanoscopic vision (p. 311). 
Microbots are more noticeable but still quite small, 
usually the size of a grain of sand or a dust mote, 
or occasionally as big as a flea. Individual bots in a 
swarm are directed by nanocomputers, with behavioral
routines modeled on biological insect and animal
swarms. These swarms stick together and work as a 
whole, communicating with nanoradios, nanolasers, 
or chemical cues, and sharing information between 
each bot in the swarm. Note that nanoswarms don't 
invade inside living bodies (though they may attack 
externally)—internal nano is handled by nanoware (p. 
308), nanodrugs (p. 321), and nanotoxins (p. 324).

Nanobots and microbots may be designed with 
all manner of miniaturized propulsion systems (see 
\textit{Mobility Systems,} p. 310), with the exception of ionic 
drives. They are powered by tiny batteries or solar 
cells. Their tiny sensors are very effective at allowing 
them to identify materials and objects, and so to target 
discriminatingly. Nanobots or microbots could, for example
be programmed to ignore metal objects, certain
types of plants, specific morphs, females, or specific individuals
Swarms may either be released directly from
a hive or from pre-packaged programmable canisters.

Swarms must be programmed before they are released
The programming first determines how long
the swarm is active. This timeframe is open-ended, 
though most swarms deteriorate into ineffectiveness 
after 2 weeks unless they are replenished by a hive. 
The programming then sets what area the swarm is 
to occupy. This is also open to interpretation and 
can vary from ``coat this person'' to ``spread out to a 
diameter of 20 meters'' to ``find the nearest chemical 
traces and track them to their source.'' Finally, programming
sets any other parameters for the swarm's
%%% 331
mission—for example, if it should ignore certain materials
if it should send a report at a predetermined
time, or if it should self-destruct into harmless dust 
when it has completed a certain task.

Programming is generally handled as a Simple 
Success Test using Programming (Nanoswarm) skill. 
Failure simply infers that the programming is imperfect
and so the swarm may not operate completely as
planned. An actual Programming (Nanoswarm) Success
Test is only called for if the swarm's programming
is substantially complex or if the character seeks to 
have the swarm act outside of its usual set functions. 
The bots in each swarm are specially equipped for the 
task they are designed for, however, so attempting 
to drastically repurpose a swarm may be difficult or 
pointless at the gamemaster's discretion.

Swarms may also be teleoperated, controlled, and/
or (re)programmed once they are released, via radio 
or laser link.

Swarms are treated as a whole. The standard 
swarm size is enough to cover a 10 x 10 x 10 meter 
cube, and this is the standard ``unit'' of swarm released
by a canister or hive. Swarms may be larger,
but they are treated as individual swarm units. Each 
swarm has a Durability of 50 and is immune to 
wounds. Most attacks against a swarm simply inflict 
1 point of damage. Area-effect weapons, plasma rifles, 
and fire inflict 1d10 damage, plasma grenades do full 
damage. EMP weapons (p. 340) are very effective 
against swarms, inflicting 2d10 + 5 damage and a -10 
modifier to all tests due to their damaging effects on 
the swarm's communication abilities until repaired. 
Swarms are not affected by vacuum.

\textbf{Cleaners:} This nanoswarm cleans, polishes, and 
removes dirt and stains. It may be used on an area, 
specific objects, or people. Some facilities employ 
permanent cleaner swarms to keep their area spotless. 
Cleaners may also be programmed to remove specific 
toxins, chemicals, or other hazardous substances in 
order to decontaminate an area. Covert operatives 
and criminals sometimes use cleaners to eliminate 
any evidence they may have left at a scene usable for 
forensics purposes, such as blood, hair, or anything 
that could be DNA-typed. \textbf{[Low]}

\textbf{Disassemblers:} Also known as smart corrosives, 
these nanobots break down any matter. Their advantage
over common acids is that not only are
they able to break down any material by using 
energy to disrupt chemical bonds, but that they 
can be programmed to take apart certain components
while ignoring others, leaving them intact.
Disassemblers are a common weapon used against 
synthmorphs, eating away their components without
having to worry about accidentally splashing
biomorphs. Upon contact, these nanobots inflict 
1d10 ÷ 2 damage (round up) per Action Turn. Accumulated
damage counts as a wound when the
Wound Threshold is reached. Both Energy and Kinetic
armor protect against this damage, but these
armors are eaten away as well, so the Armor Value 
is reduced by the soaked DV. \textbf{[High]}

\textbf{Engineers:} Engineer microswarms are used for 
various construction purposes: erecting walls, digging 
tunnels, sealing holes, reinforcing foundations, and so 
on. \textbf{[Moderate]}

\textbf{Fixers:} This is the nanoswarm version of repair 
spray (p. 333). \textbf{[Moderate]}

\textbf{Injectors:} Injector microswarms are equipped with 
tiny needles and a drug payload. A biological target 
affected by an injector swarm suffers 1 point of 
damage and the effects of the carried drug, chemical, 
or toxin. \textbf{[Moderate]}

\textbf{Gardeners:} This microswarm is useful for a number 
of agricultural purposes: killing weeds, planting seeds, 
trimming plants, pollinating, and even harvesting 
small items. It may also be programmed to simply 
defoliate an area. \textbf{[Moderate]}

\textbf{Guardians:} Guardians watch for and attack other 
unauthorized swarms. Guardians inflict 1d10 ÷ 2 
damage (round up) on other swarms they come into 
contact with per Action Turn. \textbf{[Moderate]}

\textbf{Proteans:} This nanoswarm is designed to disassemble 
other materials and objects and to create a single specific, 
pre-programmed device from the components (much 
like a specialized nanofabricator). The proteans must be 
able to scavenge appropriate raw materials (for example, 
to create a metallic device the nanobots must transform 
something else made of metal). The construction time 
takes 1 hour per cost category of the item (1 hour for a 
Trivial cost item, 2 hours for Low, etc.). \textbf{[High]}

\textbf{Saboteurs:} Sab nanobots are designed to infiltrate 
electronics or machinery and sabotage them in small 
but difficult to discern ways: severing connections, 
disabling components, gumming up moving parts, 
etc. Saboteurs inflict damage on devices similar to 
disassemblers, but the target is not destroyed and 
such damage is not immediately obvious. They inflict 
1d10 ÷ 2 points of damage to synthmorphs, bots, and 
other devices every Action Turn. Armor has no effect, 
but accumulated damage counts as a wound when the 
Wound Threshold is reached. \textbf{[High]}

\textbf{Scouts:} A scout nanoswarm will systematically
map and explore an area, collecting samples
of all materials and substances it encounters. The 
samples are carried back to the hive or canister 
and chemically analyzed. Scouts can also be used 
for forensic purposes, collecting DNA samples, 
analyzing chemical residues, and examining other 
evidence. \textbf{[High]}

\textbf{Taggants:} Taggants seek to lodge themselves onto 
everything in their area of dispersal. Each carries 
a unique identifier, so that if it is found later, the 
tagged person or object can be linked back to 
the point they were tagged. Taggants can be programmed
to remain silent, only responding to query
broadcasts made with the proper crypto codes, or 
they can be programmed to broadcast their location 
back to the deployer via the mesh. \textbf{[Low]}

\subsection{Pets}

These partially-uplifted and bio-engineered animals 
%%% 332
have rudimentary intelligence and limited communication
skills. They make for fine companions and helpers.
\textbf{Fur Coat:} A so-called ``fur coat'' is outerwear made 
from a living primitive organism. The creature's skin, 
fur, or scales are real. The organism is cultivated from 
transgenic stocks and grown around molds into clothing
shapes, often with actual usefulness: polar bear
parkas, seal diving suits, porcupine coats, etc. Fur 
coats are modified with wireless controls and haptic 
systems, so they can be made to move, shiver, massage, 
or prickle up on command. \textbf{[Low]}
\textbf{Smart Dogs:} Commonly used as discriminatory 
guardians, smart dogs are sometimes enhanced with 
combative bioware or cybernetics. \textbf{[Moderate]}
\textbf{Smart Monkey:} Commonly used by criminal groups 
for minor larceny such as pickpocketing, smart monkeys
can be useful and intelligent aides. \textbf{[Moderate]}
\textbf{Smart Rats:} These upgrades of the common Norwegian
rat are clever and dexterous, and they easily fit
into a pocket or hood. \textbf{[Low]}
\textbf{Space Roach:} Grown to the size of a small dog, 
these insects are often biosculpted for bright colors 
and patterns. They are useful for minor janitorial 
duties. \textbf{[Low]}

\subsection{Scavenger Tech}

This technology is often employed by gatecrashers, 
space scavengers, and Firewall teams during missions.
\textbf{Disassembly Tools:} These tools are useful for salvage
ops, breaking down wrecks, or dissembling anything
from a habitat room to a vehicle or synthmorph.
They include plasma torches, laser cutters, pneumatic 
jaws, and smart tools like spanners and wrenches that 
can be adapted to a wide array of connections and 
fittings. \textbf{[High]}
\textbf{Mobile Lab: }The mobile lab is a handheld device 
that contains all different types of sensors to investigate
organic and inorganic liquid, gaseous, and solid
components (from soil to tissue samples) and compositions
It performs material analysis using different
methods of spectrometry and biochemical testing, 
Fur Coat

1

1

1

5
Smart Dog

5

10 15 15
Smart Monkey

5

15 15 15
Smart Rat

5

15 15 15
Space Roach

1

10 10 15

comparing results to a built-in database of element 

and compound spectra. Its built in AI comes equipped 

with Academic: Chemistry 30. \textbf{[Moderate]}

\textbf{Specimen Container:} This capsule container is 

designed to hold samples of any sort (chemical, bio-

logical, etc.) in near stasis. It can be programmed to 

reproduce whatever conditions the user specifies, from 

cryogenic freezing to extreme heat, or even vacuum or 

high-pressure atmosphere. \textbf{[Low]}

\textbf{Superthermite Charges:} These powerful and highly 

stable demolition charges are made from a combina-

tion of nanometals and metal oxides. A single charge 

can be used to create an explosive blast inflicting 

2d10+5 damage. This charge can be shaped with a 

successful Demolitions Test, focusing the blast in a 

particular 90-degree direction (for example, to blow 

through a door). This triples the damage of the blast 

in the focused direction; in all other directions, the 

damage is reduced to 1/3rd (round down). Multiple 

charges apply a cumulative effect. \textbf{[Moderate]}

\subsection{Services}


\textbf{Anonymous Accounts:} These accounts are crucial 

for anyone who wants to be discreet with their 

online transactions. See \textit{Anonymous Account Ser-}

\textit{vices,} p. 252. \textbf{[Moderate]}

\textbf{Backup:} A single, one-time backup without insur-

ance is sometimes all the poor can afford, hoping 

that they can buy backup insurance later or that 

someone that cared about them will see to a re-

sleeving. \textbf{[Moderate]}

\textbf{Backup Insurance:} In the event of verifiable 

death, or after a set period of being missing, backup 

insurance will arrange for your cortical stack to be 

retrieved and your ego downloaded into another 

morph. If the cortical stack cannot be retrieved, your 

most recent backup is used. Most policies require 

that the holder provide a backup to be uploaded 

into secure storage at least twice a year. This industry 

works in a manner similar to insurance underwriting 

in terms of cost and individuals engaged in high risk 
1
5
1
12
1
15

3

2

1 —
5
15 10 60
1
25

5

20

4

Fray 30, Freerunning 30, Intimidation 30, Percep-

tion 30, Scrounging 30, Unarmed Combat 40
5
10 10 60
1
20

4

20

4

Climbing 50, Fray 30, Freerunning 30, Infiltra-

tion 30, Perception 30, Scrounging 30, Unarmed 

Combat 30
5
5
10 60
2
5

1

20

4

Climbing 40, Fray 40, Freerunning 30, Infiltration 

50, Perception 20, Scrounging 50
5
5
5
50
1
5

1

10

2

Fray 30, Free Fall 30, Infiltration 50, Perception 20, 

Scrounging 50
%%% 333
professions can expect to pay a premium for the service
Additionally, attempts to retrieve a cortical stack
are minimal unless one wants to pay for some extra 
effort (a thriving industry of paramilitary ego-repo 
operatives exists for this purpose). \textbf{[Low to Moderate }
\textbf{per month]}

\textbf{Body Bank:} People who are egocasting to another 
station but whom hope to download back into the 
same body they have before when they return may 
put the morph on ice for the duration of their absence. 
\textbf{[Moderate per month]}

\textbf{Bot/Pod Rental:} When you need a helping hand or 
a personal companion for a day or two, renting a bot 
or pod is often the way to go. \textbf{[Moderate per day]}

\textbf{Egocasting:} This is the use of a farcaster to transmit
an ego/infomorph. Farcasting is not cheap, and
the cost is impacted by factors such as distance to 
receiver station and priority service (paying extra to 
get bumped ahead in line). \textbf{[Expensive]}

\textbf{Fake Ego ID:} This forged ID will pass in most inner 
system and Jovian Republic habitats, and sometimes 
others. \textbf{[High]}

\textbf{Morph Brokerage:} Acquiring a new morph is not 
always easy and is affected by factors such as the type 
of morph, sought-after enhancements/customizations, 
and local availability. Numerous brokerage services 
exist to find you what you need, or close to it. With 
enough lead-time, it may be possible to grow a pod 
that closely imitates your morph of choice. A willingness
to accept used/traded-in morphs helps to reduce
costs. For more details, see \textit{Morph Brokerage,} p. 276.

\textbf{Psychosurgery:} A character can purchase time in an 
immersive high-fidelity simulspace with expert care 
from psychosurgeons and AIs in order to cope with 
derangements and disorders that build up as a result 
of existing in a transhuman universe. For an additional
price the procedure can be time shifted to speed
up the relative time within the simulspace. For more 
details, see \textit{Mental Healing and Psychotherapy,} p. 215, 
and \textit{Psychosurgery,} p. 229. \textbf{[Moderate per month]}

\textbf{Simulspace Subscription: }This will by you access to 
the simulspace of your choice, whether you want it for 
a private meeting/vacation or to play the latest and 
hottest VR game. \textbf{[Low (single use/1 day) to Moderate }
\textbf{(monthly subscription)]}

\textbf{Space Travel:} Space transport cost depends on a 
number of factors like distance, quality of lodgings, 
and how much cargo you're bringing with. At the low 
end, an intra-habitat shuttle trip within the same cluster
or a trip to or from a planetary body's surface and
orbit, is not cheap but affordable \textbf{[High]}. Just about 
anything else is progressively more costly. \textbf{[Expensive]}

\subsection{Software}

For information on using software, see the \textit{Mesh }
chapter, p. 234.

\subsubsection{Programs}

These programs can be run on any computerized device.

\textbf{AR Illusions:} These databases of AR clips can be used 
to create realistic illusions in someone's entoptic display. 
See \textit{Augmented Reality Illusions,} p. 259. \textbf{[Moderate]}

\textbf{Exploit:} Exploits are hacker tools that take advantage
of known vulnerabilities in other software. They
are required for intrusion attempts (p. 254). \textbf{[High]}

\textbf{Facial/Image Recognition:} This program can be 
used to take an image and run a pattern-matching 
search among public archives. Similar version of this 
program exist for other biometrics: gait recognition, 
vocal recognition, etc. \textbf{[Low]}

\textbf{Firewall:} This program protects a device from 
hostile intrusion. Every system comes with a standard 
version of this software by default. \textbf{[Low]}

\textbf{Sniffer:} Sniffer programs collect all of the transmission
that pass to, from, or through the device they are
running on. See \textit{Sniffing,} p. 252. \textbf{[Moderate]}

\textbf{Spoof:} Spoof is a hacker tool used to fake commands
and transmissions, making them seem as if
they came from another source. See \textit{Spoofing Authen-}
\textit{tication,} p. 255. \textbf{[Moderate]}

\textbf{Tactical Networks:} These programs allow people in 
the same squad to share tactical data in real-time. See 
\textit{Tactical Networks,} p. 205. \textbf{[Moderate]}

\textbf{Tracking:} This software is used to track people by 
their presence online. See \textit{Scanning, Tracking, and }
\textit{Monitoring,} p. 251. \textbf{[Moderate]}

\textbf{XP:} Experience playback recordings are clips of 
someone else's experiences. Depending on the content, 
some XP (porn, snuff, crime, etc.) may be restricted in 
certain jurisdictions. Some XP clips are intentionally 
modified so that their emotive tracks are more intense, 
giving the viewer a greater thrill. \textbf{[Low to High]}

\subsubsection{Ai}


S

\subsubsection{ And Muses}

Every character starts with a personal muse for free. 
Many devices also come with pre-installed AIs, capable
of helping the user, responding to commands, or
even operating the device on their own. Rules for AIs 
can be found on p. 264.

Below are some commonly available AI programs. 
Unless otherwise noted, these AIs have aptitudes of 10. 
These AIs may also be equipped with skillsofts (p. 332).

\textbf{Bot/Vehicle AI:} These AIs are designed to be capable
of piloting the robot/vehicle without transhuman
assistance. REF 20. Skills: Hardware: Electronics 20, 
Infosec 20, Interests: [Bot/Vehicle] Specs 80, Interface 
40, Research 20, Perception 40, Pilot: [appropriate 
field] 40. \textbf{[High]}

\textbf{Device AI:} These AIs are designed to operate a particular
device without transhuman assistance. Skills:
Infosec 20, Interests: [Device] Specs 80, Interface 30 
(Device Specialization), Programming 20, Research 20, 
Perception 20. \textbf{[Moderate]}

\textbf{Kaos AI:} Kaos AIs are used by hackers and covert 
ops teams to create distractions and sabotage systems. 
%%% 334
REF 20. Skills: Hardware: Electronics 40, Infosec 40, 
Interface 40, Professional: Security System 80, Programming
40, Research 20, Perception 30 plus one
weapon skill at 40. \textbf{[Expensive]}

\textbf{Security AI:} Security AIs provide overwatch for 
electronic systems. Skills: Hardware: Electronics 30, 
Infosec 40, Interface 40, Professional: Security Systems
80, Programming 40, Research 20, Perception
30, plus one weapon skill at 40. \textbf{[High]}

\textbf{Standard Muse:} Muses are digital entities that have 
been designed as personal assistants and lifelong 
companions for transhumans (see \textit{AIs and Muses,}
p. 264). INT 20. Skills: Academics: Psychology 60, 
Hardware: Electronics 30, Infosec 30, Interface 40, 
Professional: Accounting 60, Programming 20, Research
30, Perception 30, plus three other Knowledge
skills at 40. \textbf{[High]}

\subsubsection{Scorchers}

Scorchers are damaging neurofeedback programs used 
to torment hacked cyberbrains (p. 261).

\textbf{Bedlam:} Bedlam programs assault the ego with 
traumatic mental input, inflicting mental stress. 
Victims are overwhelmed with horrific,  monstrous, 
sanity-ripping sensory and emotional input. Each 
attack inflicts 1d10 SV. \textbf{[High]}

\textbf{Cauterizer:} This scorch program rips into the ego 
with destructive neurofeedback routines. Each attack 
with a cauterizer inflicts 1d10 + 5 DV on the target 
ego. This damage is reflected as digitized neurological
damage. \textbf{[High]}

\textbf{Nightmare:} Nightmare programs trigger anxiety 
and panic attacks within the victim by stimulating the 
neural circuitry representing the amygdala and hippocampus
The target ego must make a WIL x 2 Test.
If they succeed, they are shaken but otherwise unaffected
suffering a –10 modifier to all actions until the
end of the next Action Turn. If they fail, they suffers 
1d10 ÷ 2 stress damage and are overcome with panic. 
This causes them either to blindly flee, have a nervous 
breakdown, or cower in frozen shock (gamemaster's 
discretion). This panic episode lasts for 1 Action Turn 
per 10 points of MoF. \textbf{[High]}

\textbf{Shutter:} Shutters target the victim's sensory cortices, 
inflicting a –30 modifier to one chosen sense. Double 
this modifier if the attacking hacker scored an Excellent
Success. This modifier reduces at the rate of 10
points per Action Turn. \textbf{[High]}

\textbf{Spasm:} Spasm programs are design to incapacitate 
the ego with excruciating pain. Affected targets must 
immediately make a WIL x 2 Test. If they fail, they 
immediately convulse, are disabled, and writhe in 
agony for 1 Action Turns per 10 full points of MoF. 
If they succeed, they still suffer a –30 modifier to all 
actions, which reduces at the rate of 10 points per 
Action Turn. Due to the nature of the delivery, pain 
tolerance of any sort has no effect. \textbf{[High]}

\subsubsection{Skillsofts}

Skillsofts are used with skillware implants (p. 309).

\textbf{Standard Skillsoft:} These programs provide the 
character with a rating of up to 40 in a single Active 
skill. \textbf{[High]}

\subsection{Survival Gear}

The following gear is often critical to the survival of 
soldiers, spies, criminals, gatecrashers, emergency service
personnel, and others who regularly venture into
unsafe or unfamiliar regions.

\textbf{Breadcrumb Positioning System: }This worn device 
leaves micro ``breadcrumbs'' behind as the character 
moves. These devices interact with mesh inserts (or 
ectos) as long as they are within range (50 meters), 
allowing the user to map their position in relation to 
the breadcrumb trail. This is useful in derelict habitats, 
wilderness, and other areas where there is no local 
functioning mesh, and is helpful both for mapping 
and for finding one's way back. \textbf{[Low]}

\textbf{Electrogravitics Net:} Also called a safety net, this failsafe
system uses electric fields to counter gravity when
falling. While the system is not able to actually levitate 
heavy objects, it will slow down a fall enough that the 
user can land safely if the gravitational force is not too 
high (the fall height is not greater than 50 meters in 
1G). Generating these electric fields consumes a lot of 
energy, so the net is only charged for one use only and 
needs to be recharged afterwards. \textbf{[Moderate]}

\textbf{Electronic Rope:} The fibers in this rope can be 
controlled electronically, making it move in a snakelike
fashion, stiffen up, and
even wrap around objects. 
Typically comes in a 50-meter
length capable of
supporting 250 kg. \textbf{[Low]}

\textbf{Emergency Bubble:} Commonly
used as a last resort
``life raft'' on spaceships, 
an emergency bubble is 
made of advanced smart 
materials and comes in 
a portable package that 
can be quickly inflated 
(1 Action Turn) around 
the user, usually inside an 
airlock. The bubble has a 
5-meter diameter and can 
comfortably accommodate 
4 people. It maintains 1 atmosphere
of pressure in a vacuum, protect the inhabitants
from temperatures ranging from –175 to 140 C,
and provide light, breathable air and water and food 
recycling for up to four human-sized inhabitants, using 
its built in maker (p. 327). It features a simple airlock, 
carries an emergency distress beacon (below), and can 
be transparent, opaque, or polarized. It is powered by 
a small nuclear battery and also includes comfortable 
inflatable furniture. \textbf{[Moderate]}

\textbf{Emergency Distress Beacon:} This small but powerful
transmitter is powered by a nuclear battery and
%%% 335
will broadcast any programmed distress call for years. 
Though portable and medium-sized, this beacon has 
a range of 500 km in urban areas and 5,000 km elsewhere
\textbf{[Moderate]}

\textbf{Flashlight:} These handheld, wearable, or portable 
lights can display light in the normal visual spectrum, 
infrared, or ultraviolet, as desired. \textbf{[Trivial]}

\textbf{Nanobandage: }Characters without medichines must 
rely on external sources of healing. The most common 
option is the nanobandage—a plum-sized advanced 
nanotechnology generator built into a reusable, self-sterilizing
bandage. It can treat all forms of injury and
illness, from poisoning to burns to trauma. Characters 
simply apply the bandage to the wound and let the 
nanobots do the work. It removes pain and discomfort
and speeds healing (see \textit{Biomorph Healing,} p.
208). For especially severe injuries, physical first aid 
such as setting bones and removing projectiles may 
be necessary (gamemaster's choice). If the wounds are 
too severe (the patient has suffered more than five 
wounds), the unit places the patient in medical stasis 
and radios for emergency services. \textbf{[Trivial]}

\textbf{Repair Spray: }This nanobot generator creates 
nanobots designed to repair synthmorphs, vehicles, 
and other common objects. Repair spray contains 
the specifications and plans for almost all commonly 
used synthmorphs and devices and is a ubiquitous 
household item. If it does not contain the specifications
for something it is being used to repair, it must
query the object's voice for these details, otherwise 
it cannot repair it. Simply touch it to the damaged 
area, push the button on top, and it sprays out a 
number of nanobots sufficient to make repairs. 
These nanobots repair 1d10 points of damage per 
2 hours. Once all damage is restored, the nanobots 
repair wounds at the rate of 1 per day. Repair spray 
also cleans and polishes items and returns them to a 
pristine and new state. Repair spray is not effective 
on any object with more than 3 wounds, but it provides
a +30 to all repair rolls on anything too badly
damaged for it to fully repair (see \textit{Synthmorph and }
\textit{Object Repair,} p. 208). \textbf{[Low]}

\textbf{Shelter Dome:} A variant of the emergency bubble, 
this package unfolds into a dome with a 2.5-meter 
ceiling and a floor 4 meters across. To safely use this 
shelter, it must be staked down to the surface it is 
placed on. \textbf{[Moderate]}

\textbf{Spindle: }A spindle is an advanced nanotechnology 
generator that produces a super-strong cable. It can 
produce up to 2 kilometers of 0.2 millimeter diameter 
line than can support up to 250 kilograms before it 
needs more raw materials. The spindle can produce 
up to 20 meters of cable every second. It can produce 
line in a continuous length or cut the cable it produces 
to any length. Spindles can also reabsorb their cable, 
retracting it at a rate of 5 m per second. As long as 
it is recharged and has small amounts of additional 
material added every 1,000 hours of use, a spindle 
can keep producing and retracting cable indefinitely. 
By setting the maximum production speed at 10 m/
second a character with a spindle can safely jump 
off a building and land safely, using the cable to slow 
their descent. \textbf{[Moderate]}

\textbf{Spindle Climber:} This device attaches to a spindle 
and transforms it into a highly effective climbing 
device. The spindle climber has two functions. First, it 
attaches hardened tips to the spindle's cable and fires 
it at high speed, up to 50 meters, with sufficient force 
to imbed the tip into almost any sufficiently durable 
surface. Second, the spindle climber can pull itself and 
up to 250 kg up the cable at a speed of up to 2 m/sec. 
A spindle climber has enough power to shoot and pull 
up the cable 50 times before it must be recharged. A 
spindle fits inside a spindle climber. \textbf{[Low]}

\subsubsection{Vacuum Suits}

Most vacuum suits are skin-tight garments that use 
the pressure of their advanced smartfabrics on the 
wearer's body to resist vacuum. When the wearer is in 
a breathable atmosphere, the smartfabric also loosens 
the suits to serve as ordinary clothing or be easily 
put on or taken off. In all cases, the suits can become 
skin-tight within 3 Action Turns. All vacsuits contain 
advanced rebreather units capable of maintaining a 
breathable atmosphere for several hours or days.

\textbf{Light Vacsuit:} Everyone living in a sealed habitat 
owns at least one of these suits. They come in a variety
of forms. Inexpensive versions are typically lightweight
jumpsuits made of simple smart fabric that
adjusts to fit and folds up small enough to fit into a 
coat pocket. The best models include suits of high-end 
smart clothing that can transform into a vacsuit and 
an advanced nanotech generator the size of a large 
orange that deploy nanobots that cover the user and 
fit together into a vacuum suit. Both can transform 
into a vacsuit in 2 full Action Turns and do so either 
on command or if their sensors reveal that life support
is needed.

All models include a lightweight belt or torc 
containing a miniature oxygen tank and advanced 
rebreather unit that provides 3 hours of air. However
the suits contain no food or water recycling. All
models include an ecto (p. 325) and a headlight, but 
typically little else beyond atmosphere sensors to let 
the wearer know when it is safe to take off the suit. 
They protect the wearer from temperatures from –75 
to 100 C. These vacuum suits also provide an Armor 
rating of 5/5 and instantly self-seal breaches unless 
more than 20 points of damage are inflicted at once. 
\textbf{[Low, Moderate for smartfabric suits]}

\textbf{Standard Vacsuit:} These suits resemble light vacsuits
made from thicker and more durable materials
that resist tearing and provides the wearer with light 
armor. They are fitted with more substantial life support
belts that includes a maker (p. 327) capable of
recycling all wastes and producing air for up to 48 
hours and food and water indefinitely. The best suits 
are made of smart materials that can transform from 
standard clothing to vacuum suits in a single Action 
Turn, and will do so automatically if life support is 
%%% 336
needed. Each suit also contains an ecto (p. 325), a 
radio booster (p. 313), and sensors equal to specs (see 
p. 325). These suits have an Armor rating of 7/7 and 
protect the wearer from temperatures from –175 to 
140 C. They can almost instantly seal any hole unless 
more than 30 points of damage are inflicted at once. 
\textbf{[Moderate, High for smartfabric suits]}

\textbf{Hard Suit}: This heavy-duty suit can almost be 
considered a miniature space ship. Hard Suits look 
like large metallic ovals with jointed arms and legs. 
They are quite heavy, but the user can move relatively 
easily by using servo assist motors in all the major 
joints of the arms and legs. Unlike other vacsuits, they 
are solid and can resist both vacuum and up to 100 
atmospheres of external pressure. Characters wearing 
hard suits can safely explore the upper atmosphere 
of a gas giant. They are well armored against punctures
and radiation and possess miniature plasma
thrusters capable of delivering 0.01G for 10 hours. A 
built-in high quality maker produces sufficient food, 
air, and water that a user can remain in a hard suit 
indefinitely. Explorers have used them continuously 
for up to 2 months. Their gloves incorporate smart 
materials that allow each hand to use the equivalent 
of a utilitool (p. 326). Hard suits also contain radios 
and sensors equivalent to those on standard vacsuits. 
These suits have an Armor rating of 15/15, are maintained
by a fixer nanohive (p. 329), and are instantly
self-sealing of any breach unless more than 30 points 
of damage are inflicted at once. They protect the 
wearer from temperatures of –200 to 180 C. \textbf{[High]}

\section{Weapons}

A wide range of weapons are available in \textit{Eclipse }
\textit{Phase,} from the primitive to the technologically 
advanced.

\subsection{Melee Weapons}

Melee weapons are those wielded by hand (or foot) in 
melee combat. They are divided by the skill be which 
they are used.

\subsubsection{Blades}

These weapons are wielded with Blades skill.

\textbf{Diamond Axe:} Commonly found on many habitats 
for fire and emergency purposes, axes require two 
hands to wield. Their blades are diamond-coated for 
superior cutting ability. \textbf{[Low]}

\textbf{Flex Cutter:} The blade of this machete-like weapon 
is made of a memory polymer. When deactivated, the 
blade is limp and flexible, and may even be rolled up 
or otherwise easily concealed. When activated, however
the blade stiffens and sharpens into a vicious
slashing weapon. \textbf{[Low]}

\textbf{Knife:} A standard cutting implement, still carried 
by many. \textbf{[Trivial]}

\textbf{Monofilament  Sword:} Though swords are rather 
archaic in the time of \textit{Eclipse Phase,} a few eccentrics 
take advantage of modern versions with a self-sharpening
near-monomolecular edge, easily capable
of slicing through metal or limbs. \textbf{[Low]}

\textbf{Vibroblade:} These buzzing electronic blades vibrate 
at a high frequency for extra cutting ability. This 
has little extra effect when stabbing or slashing, but 
provides an extra –3 AP and +2d10 damage when 
carefully sawing through something. \textbf{[Low]}

\textbf{Wasp Knife:} Wasp knives are equipped with a canister
in their handle. The common use is to fill these
canisters with pressured air, which inflates  inside 
the target. This is potentially lethal in vacuum or 
pressurized environments (like underwater), as the 
gas bursts out of the body cavity to escape (+2d10 
damage in such situations). Wasp knives may also 
be loaded with chemicals, drugs, or nanobots. The 
target must be damaged for the canister's contents to 
affect them. \textbf{[Low]}

\subsubsection{Clubs}

Characters use Clubs skill when using these weapons.

\textbf{Club:} Clubs encompasses a wide range of one-handed
blunt objects, from saps to sticks to pipes. \textbf{[Trivial]}

\textbf{Extendable Baton:} This hardened composite baton 
retracts into its handle for easy carrying, storage, or 
concealment. Extending it simply requires a flick or 
an electronic signal. \textbf{[Trivial]}

\textbf{Shock Baton:} Shock batons are standard clubs 
used for policing duties, but when activated they also 
deliver an electric shock to struck targets (see \textit{Shock }
\textit{Attacks,} p. 204). \textbf{[Low]}

\subsubsection{Exotic Melee Weapons}

Unusual weapons requires a specific Exotic Melee 
field skill to use.

\textbf{Monowire Garrote:} This assassin's weapon features
a dangerous monomolecular wire wrapped
around a contained spool with two handles. One 
handle grips the spool, while the other extends the 
wire so that it may be used to wrap around targets 
(typically necks or limbs) and slice through them 
when pulled. Monofilament tensile strength is weak, 
however, usually breaking after one use. \textbf{[Moderate]}

\subsubsection{Unarmed}

These weapons are wielded using Unarmed Combat skill.

\textbf{Densiplast Gloves:} These gloves extra-harden when 
activated, for extra punch. \textbf{[Trivial]}

\textbf{Shock Gloves:} When activated, these gloves deliver 
an incapacitating shock along with every punch or 
grab. Note that the effect is the same whether wearing 
one glove or two. \textbf{[Low]}

\subsection{Kinetic Weapons}

Kinetic weapons damage the target by firing a hard 
%%% 337
impact projectile at high-velocities. Slugthrowers 
have evolved from the mechanical firearms of the 
early 21st century, however, and now fall into two 
categories: chemical firearms and railguns. Though 
their mechanisms for firing are different, they are 
roughly similar in effect. Railguns have a higher 
penetration and inflict more damage, which is offset 
by more limited ammunition choices. While modern 
beam weapons have their uses, they rarely match the 
punch of kinetic weapons, therefore slugthrowers 
are still perceived as the most versatile and effective 
weapon system.

Kinetic weapons are constructed from lightweight, 
reinforced plastoceramic materials, which are easily 
produced even without nanofabrication. By default, 
modern kinetic weapons are ambidextrous but more 
importantly feature safety and smartlink systems (p. 
342) that automatically connect to the wielder's mesh 
inserts for firing assistance, target recognition, and 
tactical networking.

The wielder of a firearm or railgun uses Kinetic 
Weapons skill. For information on firing modes, see 
p. 198. For different ammunition types, see p. 336. 
Ranges are listed on p. 203.

\subsubsection{Firearms}

Modern chemical firearms use caseless ammunition 
that is auto-loaded from a magazine. They are effectively
recoilless (thanks to rheological smart fluid
mechanisms) and electronically fired (an electric 
Diamond Ax
Flex Cutter
Knife
Monofilament Sword
Vibroblade
Wasp Knife
Club
Extendable Baton
Shock Baton
Monowire Garrote
Bioware Claws (p. 304)
Cyberclaws (p. 307)
Densiplast Gloves
Eelware (p. 304)
Shock Gloves
Unarmed
charge vaporizes the propellant, using the expanding 
steam and plasma to eject and accelerate the projectile).
Note that older, pre-Fall firearms still exist and are 
traded by black marketeers, though they use outdated 
system such as liquid propellants or cased ammunition
At the gamemaster's discretion, these relics may
suffer shorter ranges, less penetration, fewer firing 
modes, or reduced damage.
\textbf{Pistols:} Pistols are small-sized (p. 297) and designed
for one-hand use. Light pistols sacrifice penetrating
ability for concealability. Heavy pistols focus
on stopping power, with medium pistols occupying 
a middle ground. All versions fire in semi-automatic, 
burst-fire, and full-auto modes. \textbf{[Low]}
\textbf{Submachine Guns:} SMGs use pistol ammunition
but are medium-sized (p. 297) and may fire in
semi-auto, burst fire, or full auto modes. They typically
are designed in a bullpup configuration for
close quarters operations and are ideal for tactical 
and strike teams. \textbf{[Moderate]}
\textbf{Automatic Rifles:} Automatic rifles use rifle ammunition
and have greater range and penetration than
SMGs. They fire in semi-auto, burst fire, or full auto 
modes. They are two-handed weapons. \textbf{[Moderate]}
\textbf{Sniper Rifle:} Sniper rifles are optimized for range, 
accuracy, penetration, and stopping power. They fire 
in semi-auto, burst fire, or full auto modes, and are 
two-handed weapons. \textbf{[High]}
\textbf{Machine Gun:} Machine guns are heavy weapons
typically mounted, and intended to provide
–3
2d10 + 3 + (SOM ÷ 10) 

14 + (SOM ÷ 10)
–1
1d10 + 3 + (SOM ÷ 10) 

8 + (SOM ÷ 10)
–1
1d10 + 2 + (SOM ÷ 10) 

7 + (SOM ÷ 10)
–4
2d10 + 2 + (SOM ÷ 10)

13 + (SOM ÷ 10)
–2
2d10 + (SOM ÷ 10) 

11 + (SOM ÷ 10)
–1
1d10 + 2 + (SOM ÷ 10) 

7 + (SOM ÷ 10)
—
1d10 + 2 + (SOM ÷ 10) 

7 + (SOM ÷ 10)
—
1d10 + 2 + (SOM ÷ 10) 

7 + (SOM ÷ 10)
—
1d10 + 2 + (SOM ÷ 10) + shock (p. 204)

7 + (SOM ÷ 10)
–8

3d10

16
–1
1d10 + 1 + (SOM ÷ 10) 

6 + (SOM ÷ 10)
–2
1d10 + 3 + (SOM ÷ 10) 

8 + (SOM ÷ 10)
—
1d10 + 2 + (SOM ÷ 10) 

7 + (SOM ÷ 10)
—

shock (p. 204)

—
—
1d10 + (SOM ÷ 10) + shock (p. 204)

5 + (SOM ÷ 10)
—
1d10 + (SOM ÷ 10)

5 + (SOM ÷ 10)
%%% 338
continuous fire for support or suppressive purposes. 
They fire in burst fire or full auto modes, and are two-handed
weapons. \textbf{[High]}

\subsubsection{Railguns}

Railguns use a pair of electromagnetic rails to slide 
and accelerate a non-explosive conductive projectile at 
extremely high velocities (Mach 6+) to create an overwhelming
penetrating attack. The kinetic energy of
the projectile exceeds that of an explosive-filled shell 
of greater mass and creates shock and heat waves 
Light Pistol

–3
Medium Pistol

–5
Heavy Pistol

–7
Submachine Gun

–5
Automatic Riﬂ e

–9
Sniper Riﬂ e

–15
Machine Gun

–9
Light Pistol

—
Medium Pistol

–2
Heavy Pistol

–4
Submachine Gun

–2
Automatic Riﬂ e

–6
Sniper Riﬂ e

–12
Machine Gun

–6
upon impact that shatter and incinerate the target, or 
portions of it. While railguns are more potent than 
firearms, the ammunition choices are limited as the 
projectile must be conductive and able to survive 
both acceleration and heat created in the process due 
to friction. Nanofabrication allows railguns to be 
manufactured on the personal weapons scale while 
high-energy portable batteries provide the power to 
fire them. Railgun operation is silent except for the 
supersonic crack of the projectile.
Railguns are available in the same models as firearms
(pistols through machine guns), with the following
modifications:
•  Increase AP by –3
•  Increase damage by +2
•  Increase the maximum for each range category 
by x1.5
•  Increase Cost category by one
• Railguns may only use regular and armor-piercing
ammunition
•  Railguns also require battery power for each shot. 
Standard railgun batteries hold enough power for 
200 shots, after which they must be recharged at 
the rate of 20 points per hour.

\subsubsection{Kinetic Ammunition}

Ammunition is defined by its various types (standard, 
gel, APDS, etc.) and by the class of gun (light pistol, 
heavy pistol, SMG, etc.). For simplicity, each gun 
can trade ammunition with another gun of its class, 
though ammunition for firearms and railguns is not 
exchangeable. For example, all railgun SMGs can 
share ammo.
2d10 + 2
13

SA, BF, FA

10
2d10 + 4
15

SA, BF, FA

12
2d10 + 6
17

SA, BF, FA

16
2d10 + 5
16

SA, BF, FA

20
2d10 + 8
19

SA, BF, FA

30
2d10 + 12
23

SA

40
2d10 + 8
19

BF, FA

50
2d10
11

SA, BF, FA

10
2d10 + 2
13

SA, BF, FA

12
2d10 + 4
15

SA, BF, FA

16
2d10 + 3
14

SA, BF, FA

20
2d10 + 6
17

SA, BF, FA

30
2d10 + 10
21

SA

40
2d10 + 6
17

BF, FA

50
%%% 339

The ammunition's Damage Value and Armor Penetration
modifiers are added to the weapon's base
DV and AP. With the exception of regular and armor-piercing
rounds, none of this ammunition may be
used with railguns. Listed costs are per 100 rounds 
of ammunition.

\textbf{Armor-Piercing: }This tungsten-carbide ammunition 
penetrates armor effectively. \textbf{[Low]}

\textbf{Bug: }Bug rounds are equipped with a microbug and 
medical sensor nanobots. They attempt to gather information
on the target's location (via standard mesh
tracking), health (querying the target's medichines), 
and surroundings (typically hindered by being inside 
the target's body). They will transmit status reports 
in a pre-programmed manner via the mesh or a pre-chosen
frequency band either continuously or in preset
intervals. \textbf{[Low]}

\textbf{Capsule:} Capsule ammo carries a payload (drug, 
toxin, nanobots) that is released inside the target after 
the round penetrates. \textbf{[Trivial plus payload cost]}

\textbf{Flux:} Flux ammo is made from rheological materials
that allow each bullet to be ``programmed'' so that
they may change from regular rounds to less-lethal soft 
plastic-like rounds. This allows the firer to choose the 
type of round (regular or plastic) made with each shot 
or burst, and then change with the next one. \textbf{[Low]}

\textbf{Hollow-Point:} Hollow-point bullets are designed to 
deform and widen once they penetrate a target, thus 
Armor-Piercing

–5

–2
Bug

+1

–1d10
Capsule

+1

–half
Flux
as ammo type

as ammo type
Hollow-Point

+2

+1d10
Jammer

—

no damage
Plastic
(AV doubled)

–half
Reactive

–2

+2
Reactive Armor-Piercing


–6

–1
Regular

—

—
Splash

—

no damage
Zap

+2

–half + shock (p. 204)
 
Accushot

—

—
Biter

—

+1d10
Flayer

—

+2
Homing

—

—
Laser-Guided

—

—
Proximity

–1

+2
Zero

—

—
%%% 340
inflicting more damage. \textbf{[Trivial]}

\textbf{Jammer:} Jammers stick to the target and pulse out 
jamming electromagnetic signals, jamming the target's 
wireless communications. If an Opposed Test is called 
for, these devices have an Interface of 30. See \textit{Radio }
\textit{Jamming,} p. 262. \textbf{[Low]}

\textbf{Plastic: }Plastic ammo is designed to hurt but not 
wound targets, and is commonly used for crowd control
purposes. \textbf{[Trivial]}

\textbf{Reactive: }The casing on these projectiles is made 
of reactive materials that release a large amount of 
energy when subjected to a sudden shock or impact—
such as striking a target. In other words, they explode 
or superheat when they hit. \textbf{[Low]}

\textbf{Reactive Armor-Piercing (RAP):} This is a tungsten-carbide
armor-piercing round with a reactive
casing, allowing the ammunition to penetrate even 
further. \textbf{[Moderate]}

\textbf{Regular Ammo: }This standard metal projectile is 
designed to put holes into morphs. \textbf{[Trivial]}

\textbf{Splash: }Splash rounds carry a payload like capsule 
ammo, but are designed to break upon impact rather 
than penetrating, splashing their contents on the 
target's exterior. Splash rounds are typically loaded 
with paint, taggant nanobots, tracker dye, and similar 
substances. \textbf{[Trivial plus payload cost]}

\textbf{Zap:} Zap rounds are rubber or gel bullets that 
create an electric charge upon firing in a piezoelectric 
like manner to stun the target effectively with both 
the bullet and the electric shock. \textbf{[Trivial]}

\subsubsection{Smart Ammo}

Smart ammunition takes advantage of nanotechnology
to produce bullets that can alter their flight path,
home in the target, and correct aim. Smart ammo may 
not be used with railguns. With the exception of biter, 
flayer, and proximity rounds, smart ammo may be 
combined with other ammo types (accushot armor-piercing
for example).

\textbf{Accushot:} Accushot bullets change shape within 
flight to keep dead on course, countering the effects 
of wind, drag, and gravity over distance. Attacks 
made with accushot bullets ignore all range modifiers
\textbf{[Low]}

\textbf{Biter:} Biters are specially-designed to fragment in 
opposite proportion to the hardness of the target they 
strike. For hard targets (synthmorphs), they fragment
very little, blasting a big hole. For soft targets
(biomorphs), they fragment and tumble in multiple 
directions within the body. \textbf{[Low]}

\textbf{Flayer:} Flayers have nanosensors to detect an oncoming
impact, shooting out monomolecular barbs
as they are about to strike a target. These monowires 
cut through the target along with the bullet, inflicting 
additional damage. \textbf{[Low]}

\textbf{Homing:} When fired with a smartlink system, the 
bullet identifies the target and uses nanosensors to 
lock on, correcting the bullet's trajectory with surface 
alterations and tiny vectored nozzles. Apply a +10 
modifier to the Attack Test, cumulative with aiming 
and smartlink modifiers. Homing bullets may also be 
used for indirect fire (p. 195). \textbf{[Low]}

\textbf{Laser-Guided:} These bullets function like homing 
smart rounds (apply the +10 attack modifier), except 
rather than requiring a smartlink system, they lock 
onto the reflection of the laser sight used to paint 
the target. Laser-guided bullets may also be used for 
indirect fire (p. 195). \textbf{[Low]}

\textbf{Proximity:} Proximity is an explosive ammunition 
that identifies the target when fired via smartlink. If 
the round determines that it will miss the target, it will 
still explode if it reaches the close proximity of the 
target. If the attack misses with an MoF of 10 or less, 
the round explodes 1d10 meters away from the target 
and inflicts 1d10 area effect damage (see \textit{Blast Effect,}
p. 193) in the proximity of the target. \textbf{[Moderate]}

\textbf{Zero:} Similar to homing smart rounds, zero bullets 
identify the target when fired via smartlink. Whether 
the round hits or misses, however, it sends telemetry
data back to the next zero bullet, allowing it to
course-correct and ``zero in'' to hit the target (or hit 
more accurately). Apply a +10 modifier to each shot 
(or burst) fired after the first against the same target in 
the same Action Turn. \textbf{[Low]}

\subsection{Beam Weapons}

Beam weapons is a broad category for a number of electromagnetic
weapons with a wide range of effects. With
a few exceptions, energy weapons are primarily used for 
less-than-lethal purposes, designed to impair the target 
rather than kill it. Their poor performance against armor, 
lesser ability to damage targets, and high power requirements
make them less versatile than kinetic weapons.
The wielders of such weapons use Beam Weapons skill.

Beam weapons are powered by nuclear batteries. 
This battery is good for a list number of shots before 
it is depleted. Batteries may be recharged at the rate of 
20 shots per hour; they have a Cost of \textbf{[Low]}

\textbf{Laser Pulsers: }Laser weapons use focused beams of 
light to inflict damage on the target by burning into it 
and causing it's outer surface to vaporize and expand, 
creating an explosive effect. The laser beam is pulsed 
in order to bite into the target before the beam is diffused
Pulsers are vulnerable to atmospheric effects
like dust, mist, smoke, or rain, however—the gamemaster
should reduce their effective range categories
as appropriate. Note that laser pulses are invisible in 
the normal visual spectrum (but are visible to characters
with enhanced vision). Pulsers are medium-sized
(p. 297) and fire in semi-auto mode. \textbf{[Moderate]}

One advantage to the pulser is that it can be placed 
in less-lethal mode. In this case, it first fires a pulse at 
the target to create a ball of plasma, quickly fired by 
a second pulse that strikes the plasma and creates a 
flash-bang shockwave to stun and disorient the target. 
This blast has an area of effect with a 1-meter radius. 
Anyone caught in the blast must make a SOM x 2 
Test (SOM x 3 for synthmorphs or biomorphs with 
any form of pain tolerance). Failure means the target 
%%% 341
is temporarily stunned and disoriented and loses 
their next action. A critical failure means the target 
is knocked down and paralyzed for 1 Action Turn per 
10 points of MoF. In this stun setting, the pulser fires 
only in single-shot mode.

\textbf{Microwave Agonizer: }The agonizer fires millimeter-wave
beams that create an unpleasant burning sensation
in skin (even through armor) and to metals.
Agonizers have two settings. The first is an active 
denial setting that causes extreme burning pain in the 
target, inflicting –20 to the target's actions and forcing
them to move away from the beam on their next
action unless they succeed in a WIL Test (targets with 
Level 1 Pain Tolerance or the equivalent only suffer 
a –10 modifier and roll WIL x 2). Synthetic morphs 
and biomorphs with Level 2 Pain Tolerance (or the 
equivalent) are immune to this weapon. The second 
setting (colloquially known as the ``roast'' setting) has 
the same effect of the first, but also actually burns the 
target, inflicting the listed damage. Originally developed
for crowd control, the agonizer is also useful for
repelling animals. The agonizer is small-sized (p. 297) 
and fires in semi-auto mode. \textbf{[Moderate]}

\textbf{Particle Beam Bolter:} This weapon shoots a bolt of 
accelerated particles at near light speed that transfer 
massive amounts of kinetic energy to the target, superheating
and creating an explosion when striking.
The bolter's beam is not diffused by the cloud that 
occurs when it strikes, and so it has greater penetration
than the laser pulser. Likewise, the bolter is not
affected by smoke, fog, or rain. The bolter's beam is 
invisible. Note that bolters are designed for either 
atmospheric or exoatmospheric (vacuum) operation, 
and will not function in the opposite environment 
(though bulkier dual models, combining both models, 
are also available). Bolters fire in semi-auto mode and 
are rifle-sized two-handed weapons. \textbf{[High]}

\textbf{Plasma Rifle:} This bulky, heavy, two-handed 
weapon blasts a stream of nova-hot plasma at the 
target, inflicting severe burns and thermal damage, 
possibly melting or evaporating the target entirely. 
Plasma rifles are perhaps the deadliest man-portable 
weapons in use. Plasma guns suffer from dangerous 
overheating, however, and so require 1 full Action 
Turn of cool-down time after every 2 shots. Plasma 
Cybernetic Hand Laser (p. 308)
Laser Pulser
   Stun Mode
Microwave Agonizer
   Roast Mode
Particle Beam Bolter
Plasma Riﬂ e
Stunner
rifles only fire in single shot mode. \textbf{[Expensive]}
\textbf{Stunner:} The stunner is an electrolaser that creates
an electrically-conductive plasma channel to the
target, down which it transmits a powerful electric 
current, shocking the target. Stunners do not work in 
vacuum. Stunners fire in semi-auto mode. \textbf{[Moderate]}

\subsection{Seekers}

Seekers are a combination of automatic grenade 
launcher, micromissile, coilgun, and smart munitions 
technology. Unlike traditional launchers of the past, 
miniaturization allows the manufacture of seeker micromissile
launchers in personal weapon sizes. Seeker
rounds are fired at high-velocity via rings of magnetic 
coils, after which the micromissile or minimissile uses 
scramjet technology to propel itself and maintain high 
velocities over great distances. Seekers are wielded 
using Seeker Weapon skill.
Seeker missiles are detailed on p. 340. Like grenades
seekers may be programmed for a variety of
trigger events (see \textit{Grenades and Seekers,} p. 199). All 
seeker weapons are smartlink-equipped (p. 342).
\textbf{Disposable Launcher (Standard Missile):} This 
launcher is pre-packed with one standard missile. 
\textbf{[Moderate (includes missile)]}
\textbf{Seeker Armband (Micromissile):} This weapons unit 
is worn on the arm, allowing the user to point and 
fire using an entoptic smartlink system. Though highly 
portable, the armband's micromissile supply is low. It 
fires in single-shot mode. \textbf{[Moderate]}
\textbf{Seeker Pistol (Micromissile):} This pistol-sized 
seeker launcher fires micromissiles in semi-auto 
mode. \textbf{[Moderate]}
\textbf{Seeker Rifle (Micromissile/Minimissile):} The seeker 
rifle comes in a bullpup configuration and fires either 
—
2d10

11

SA

50
—
2d10

11

SA

100
—
1d10

5

SS

—
—
pain (see description)

—

SA

100
–5
2d10

11

SA

50
–2
2d10 + 4

15

SA

50
–8
3d10 + 12

28

SS

10
—
(1d10 ÷ 2) + shock (p. 204)

—

SA

200
Disposable Launcher

SS

1
Seeker Armband

SS

4
Seeker Pistol

SA

8
Seeker Riﬂ e

SA

12 micromissile/6 minimissile
Underbarrel Seeker

SA

6
%%% 342
micromissiles or minimissiles in semi-auto mode. It is 
a two-handed weapons. \textbf{[High]}
\textbf{Underbarrel Seeker (Micromissile):} This seeker 
micromissile launcher is commonly attached to the 
underbarrel of SMGs or assault rifles. It fires in semi-auto
mode. \textbf{[Moderate]}

\subsection{Spray Weapons}

Spray weapons blast their ammunition outwards in a 
widening cone, allowing them to strike several targets 
at once. These weapons are wielded with Spray Weapons
skill. Spray weapon ammunition has a flat cost
of Low per 100 shots (with the exception of buzzers, 
which use nanoswarms).
\textbf{Buzzer:} Equipped with a specialized nanobot hive, 
Buzzers are used to spray a nanoswarm (p. 328) on a 
target or area. They have a limited capacity of swarms, 
though the nanohive can construct one new swarm 
each hour. This weapon is two-handed. \textbf{[Moderate]}
\textbf{Freezer:} Freezers spew out a fast-hardening foam 
that immediately begins to harden. They are primarily
used as a non-lethal method of immobilizing or
securing a target. Struck characters must immediately 
make a REF x 3 Test or become trapped. Apply a –30 
modifier to this test if the attacker scored an Excellent
Success (MoS 30+). The foam allows characters
to breath even if their mouth and nose are covered, 
but it may impede sight. Freezer foam can be spiked 
with contact toxins or drugs to additionally sedate 
the target. It can also be used to construct temporary 
barricades or cover. Hardened foam has an Armor of 
10 and Durability of 20. It slowly breaks down and 
degrades over a 12 hour period. Freezers are two-handed
\textbf{[Moderate]}
\textbf{Shard Pistol:} The shard pistol is a flechette weapon, 
firing a stream of of diamondoid monomolecular 
shards at high velocities. These micro flechettes are 
very good at penetrating armor, but they do not disperse
kinetic energy well and so do not cause as much
tissue damage as kinetic weapons. Shard ammunition 
is often coated with drugs or toxins for extra efficiency
\textbf{[Low]}
\textbf{Shredder:} A heavier version of the shard pistol, 
the shredder fires a larger cloud of lethal flechettes, 
enough to shred a portion of the target into a fine 
mist. \textbf{[Moderate]}
\textbf{Sprayer:} This is a general-purpose two-handed 
squirtgun, loaded with tanks filled with the chemical 
Buzzer

—
Freezer

—
Shard

–10
Shredder

–10
Sprayer

as chemical/drug
Torch

–4
or drug of the wielder's choice. \textbf{[Low]}
\textbf{Torch:} This modern flamethrower uses condensed 
ammunition capsules rather than fuel tanks, scorching
targets and setting them on fire. Any hit that is
an Excellent Success (MoS 30+) sets the target on 
fire, where they will continue to take 2d10 damage 
per Action Turn. These chemical fires are particularly 
difficult to put out unless they are deprived of oxygen. 
Torches are two-handed. \textbf{[Moderate]}

\subsection{Grenades And Seekers}

Grenades and seeker missiles come in similar munitions
packages and with similar trigger mechanisms,
though their packaging, physical form, and methods 
of application differ. Seeker missiles are fired from a 
seeker launcher (p. 339) using Seeker Weapons skill. 
Grenades are thrown at targets using Throwing Weapons
skill. If a grenade or seeker misses, use the rules
for scatter (p. 204).
Grenades are available in standard form or as 
microgrenades. Similarly, missiles are available in 
standard, minimissile, or micromissile sizes. Standard 
grenades and minimissiles are the baseline standard 
for listed effects. All are area effect weapons (p. 193). 
Minigrenades and micromissiles inflict –1d10 damage 
(or have another decreased effect as noted). Standard 
missiles double the listed DV. For weapons with a 
unfirom blast effect or other static blast area, divide 
the base listed radius in half for minigrenades and micromissiles
and double it for standard missiles. Listed
costs are for 10 grenades/missiles.
Each seeker has one smart ammo option (p. 338) 
other than biter or flayer.
\textbf{Concussion:} These devices emit a concussive blast 
designed to knock opponents off their feet and stun 
them. Any character caught within a base blast radius 
of 10 meters must make a SOM x 2 Test. If they fail, 
they are knocked down. If their MoF is 30+, they are 
additionally stunned until the end of the next Action 
Turn. Anyone caught in the blast radius suffers a 
–10 action modifier for the rest of that Action Turn. 
\textbf{[Moderate]}
\textbf{EMP:} EMP munitions fire off a strong electromagnetic
pulse when they ``detonate.'' Since most electronics
in \textit{Eclipse Phase} are built with optical technology,
and power supplies and sensitive microcircuits are 
shielded and surge-protected, this has no major 
damaging effect. Antennas, however, are vulnerable, 
nanoswarm

—

SS

3
incapacitation

—

SA

20
1d10 + 6

11

SA, BF, FA

100
2d10 + 5

16

SA, BF, FA

100
as chemical/drug

as chemical/drug

SA

20
3d10

16

SS

20
%%% 343
especially finer wires like those used with mesh inserts. 
As a result, the primary effect of EMP is to disable 
radio communications—every radio within range of 
the blast is reduced to 1/10th the normal range. The 
base blast radius for EMP is 50 meters. \textbf{[High]}

\textbf{Frag:} Fragmentation explosives spread a cloud of 
lethal flechettes over the area of effect. They are resisted
with kinetic armor. \textbf{[Moderate]}

\textbf{Gas/Smoke:} Gas/smoke munitions emit a cloud of 
their contained substance. Smoke impedes sight by 
releasing thick fumes upon ignition of the seeker. The 
smoke can be of any color and is often heated (called 
thermal smoke) to obfuscate heat signatures moving 
through the smoke as cover. Note that gases dissipate 
much more quickly under certain environmental conditions
(wind, rain, etc.) \textbf{[Low]}

\textbf{High-Explosive: }High-explosive seekers and grenades
are designed to create a very destructive shock
and heat wave. This damage is resisted with energy 
armor. \textbf{[Moderate]}

\textbf{High-Explosive Armor-Piercing (HEAP):} A design 
only available for seekers (not grenades), HEAP 
warheads use high explosives to blast a path for 
a penetrating round. HEAPs lose –4 damage per 
meter distance from the blast, as opposed to the 
usual –2. \textbf{[Moderate]}

\textbf{Overload:} Overload grenades and seekers launch an 
all-out assault on the target's sensory spectrum. This 
attack includes blinding by intense flashing light, a 
deafening thunderclap followed by intense ultrasonic 
screaming, nausea-inducing malodorants, and infrasonic
frequencies that can trigger unpleasant emotional
responses (anxiety, uneasiness, extreme sorrow, nervous
feelings of revulsion or fear). For an extra kick,
overloads are also packed with ``stingballs''—rubber 
pellets that inflict pain when detonated near an under-armored
target. Anyone caught in the base 10-meter
blast radius must make a SOM + WIL Test. If they fail, 
they must immediately leave the area of effect. If they 
fail with an MoF of 30+, they are incapacitated for 
3 Action Turns with disorientation and/or vomiting, 
after which they must roll again. Overload munitions 
remain in effect for 1 full minute. Anyone in the area 
of effect suffers a –30 action modifier, which reduces 
Concussion
Frag
EMP
Gas/Smoke
High-Explosive
HEAP
Overload
Plasmaburst
Splash
Thermobaric
by 10 per Action Turn after they leave the area. Additionally
anyone facing the direction of the overload
round suffers a –10 glare modifier (neutralized by 
anti-glare systems). \textbf{[Moderate]}
\textbf{Plasmaburst:} Also called ``hellballs,'' these munitions
release a burst of plasma upon detonation that
causes searing heat and fire damage across the area of 
effect without the devastating shockwaves of explosions
that might rebound in an enclosed environment
and/or breach a habitat's infrastructure. \textbf{[High]}
\textbf{Splash:} Splash rounds spread a contained substance 
(drug, chemical, nanoswarm, paint) over a base 10-meter
blast radius when they detonate. \textbf{[Low plus }
\textbf{payload cost]}
\textbf{Thermobaric:} Thermobaric grenades and seekers 
utilize a more deadly form of explosion. When they 
detonate, they disperse a cloud of aerosol explosive 
over an area and then ignite, literally setting the air 
on fire, generating a devastating pressure wave, and 
sucking the oxygen out of the area. Thermobarics use 
the rules for uniform blast (p. 194). \textbf{[High]}

\subsubsection{Sticky Grenades}

Sticky grenades have a special coating that when triggered
becomes a sticky adhesive, allowing the grenade
to be stuck to almost any surface. Sticky grenades can 
even be wielded in melee combat, smacking them on 
an opponent to be detonated later. \textbf{[Trivial]}

\subsection{Exotic Ranged Weapons}

These weapons are either rare or distinctly separate 
from other weapons types. These weapons are wielded 
with an Exotic Ranged Weapon skill of the appropriate
field.
\textbf{Vortex Ring Gun:} This less-lethal two-handed 
weapon detonates a blank cartridge and accelerates 
the explosive pressure down a widening barrel so that 
it develops into a high-speed vortex ring—a spinning, 
donut-shaped blast vortex. This concussive blast is 
used to knock down and incapacitate close-range 
targets. Struck targets suffer a –10 action modifier for 
the rest of that Action Turn and must must succeed in 
a SOM x 2 Test or fall down. If their MoF is 30+, they 
are additionally stunned and unable to act until the 
—
1d10 ÷ 2

5

E
–4
3d10 + 6

22

K
—
—

—

—
—
—

—

—
—
3d10 + 10 

26

E
–8
3d10 + 12

28

K
(AV x 2)
1d10 ÷ 2 

5

K
–6
3d10 + 10

26

E
—
—

—

—
–10
3d10 + 5

21

E
%%% 344
end of the next Action Turn. Drugs, chemicals, and 
similar agents may be loaded into the charge as well. 
\textbf{[Moderate]}

\subsection{Weapon Accessories}

The following accessories are available for various 
weapons.

\textbf{Arm Slide:} This slide-mount can hold a pistol-sized
weapon under a character's sleeve, pushing the
weapon into the character's hand with an electronic 
signal or specific sequence of arm movements. \textbf{[Low]}

\textbf{Extended Magazine:} This ammunition case has 
an increased capacity. Increase the weapon's ammo 
capacity by +50%. Only available for firearms and 
seekers. \textbf{[Low]}

\textbf{Gyromount:} This weapon harness features a gyro-stabilized
weapon mount that keeps the weapon steady.
Negates all modifiers from movement. \textbf{[Moderate]}

\textbf{Imaging Scope:} Imaging scopes attach to the top of 
the weapon and act like specs (p. 325). Scopes may 
also bend like a periscope, along a character to point 
the weapon and target around corners without leaving
cover. \textbf{[Low]}

\textbf{Flash Suppressor:} This device obscures the muzzle 
flash on firearms, applying a –10 modifier on Perception
Tests to locate a firing weapon by its flash. \textbf{[Low]}

\textbf{Laser Sight:} This underbarrel laser emits a beam 
that places a glowing red dot on the target to assist 
targeting. Apply a +10 modifier to Attack Tests (not 
cumulative with a smartlink modifier). Laser sights 
may also be used to paint a target for laser-guided 
smart ammo or seekers. Infrared and ultraviolet lasers 
are also available, so that the dot is only visible to 
characters able to see in those spectrums. \textbf{[Low]}

\textbf{Safety System: }A biometric (palmprint or voiceprint
or ego ID (p. 279) sensor is embedded in the
weapon, disabling it if anyone other than an authorized
user attempts to fire it. \textbf{[Low]}

\textbf{Shock Safety:} Just like a safety system, except that 
an unauthorized user is zapped with an electric shock. 
Treat as a shock baton (p. 334). \textbf{[Moderate]}

\textbf{Silencer/Sound Suppressor:} This barrel-mounted 
accessory reduces the sound of a firearm's discharge. 
Apply a –10 modifier on hearing-based Perception 
Tests to hear or locate the gun's firing. \textbf{[Moderate]}

\textbf{Smartlink:} A smartlink system connects the weapon 
to the user's mesh inserts, placing a targeting bracket 
in the character's field of vision and providing range 
and targeting information. Apply a +10 modifier to 
the Attack Test. Smartlinks also incorporate a micro-camera
that allows the user to see what the weapon
is pointed at, fire around corners, etc. Smartlinks also 
allow certain other types of weapon system control, 
such as changing flux ammo (p. 337) or programming 
seeker trigger conditions (p. 199). \textbf{[Moderate]}

\textbf{Smart Magazine:} A smart magazine allows the 
character to pick and choose what ammo round will 
be fired with each shot. This system leaves less room 
for bullets, however, so reduce the weapon's ammunition
capacity by half (round up). Smart magazines
may be combined with extended magazines, in which 
case ammo capacity is normal. \textbf{[Moderate]}

\section{Robots And Vehicles}

%%% 345
The following is a small selection of the many vehicles 
in use in the solar system. Almost all of the vehicles 
in current use, including all of the vehicles listed here, 
have built-in AIs capable of piloting the vehicle under 
almost all circumstances. In most cases, passengers 
simply state their destination and the vehicle takes 
them there. Manual piloting is primarily used in 
emergencies or by people who prefer the exotic thrill 
of controlling their own vehicle.
Rules for handling robots and vehicles are detailed 
on p. 195. Any of these shells may be modified for use 
as a synthetic morph by adding a cyberbrain system 
(p. 300). Each of the shells listed here comes with a 
puppet sock (p. 307) for remote-control operation.

\subsection{Aircraft}

On Mars, Venus, and within large open-space habitats 
like O'Neil cylinders, aircraft of various kinds see regular 
use. This includes modern version of rotorcraft helicopters
autogyros, tilt-rotors), fixed-wing planes, and zeppelins
and other lighter-than-air craft. These are typically
propelled by turbofan or jet engines, rotors, or vectored 
thrust. These vehicles are piloted with Pilot: Aircraft skill.
\textbf{Microlight:} This ultra-light personal aircraft is 
not much more than a strut-based wing, an airframe, 
and an electric propeller engine. They are ideal for 
getting around inside large habitats with enclosed 
airspace. \textbf{[Low]}
\textbf{Portable Plane}: Powered by superconducting batteries
and with an exceedingly small but powerful
electric motor, this light but durable propeller plane 
is made of smart materials that allow it to be swiftly 
folded up into a small portable package. Different versions
are designed for flight on Mars, Titan, or Venus,
each taking 10 minutes to assemble or disassemble. 
The Martian version unpacks into an airplane with a 
wingspan of 11 m with a top speed of 250 kph and a 
cruising speed of 220 kph and a range of 1,300 km. 
The Venusian version has a wingspan of 9 m, a top 
speed of 200 kph and a range of 1,000 km. The version
designed for use on Titan has a wingspan of 8
m and has a top speed of 200 kph and a range of 
2,000 km. In all versions, the two occupants ride in 
an inflatable and insulated pressurized bubble with a 
life support system capable of providing clean air and 
comfortable temperatures for 20 hours on Mars and 
Venus, and 15 hours on Titan. \textbf{[High]}
\textbf{Rocket Buggy}: This vehicle is the most common 
form of medium to long distance personal transport 
Microlight 1
Portable Plane

2
Rocket Buggy

4
Small Jet

6
on Luna, and is in common on most other moons 
and large asteroids. On these airless worlds, a rocket 
buggy can reach orbit and return or take a parabolic 
path to any destination on that moon in less than 
an hour. This vehicle is also regularly used to travel 
between habitats that are less than 30,000 km apart. 
The vehicle is pressurized, but is designed for short 
duration travel only. The seats are relatively small 
and the life support system contains no provisions 
for recycling food or water and can only support the 
passengers for an absolute maximum of 50 uncomfortable
hours. Rockets buggies come equipped with
headlights, radio boosters, and radar with a range of 
up to 250 km.
A version of this vehicle is also used on both Mars 
and Titan, but here the frame has been modified to 
act as a lifting body, and it has a top speed in the thin 
Martian atmosphere of 2,500 km/hour and a range 
of 8,000 km on Mars. On Titan is has a top speed 
of 3,000 kph in the atmosphere, but it can also reach 
orbit. \textbf{[Expensive]}
\textbf{Small Jet}: Methane-powered jet planes are one of 
the most common forms of fast transport on Mars 
and Venus. Similar planes are used on Titan, except 
that they carry both liquid methane and liquid oxygen. 
These jets range in size from huge vehicles the size of 
late 20th-century airliners to small planes designed to 
carry half a dozen passengers. All jets are made using 
smart materials, so that their wings and frames can 
adapt to a wide range of speeds and altitudes. One 
common small jet has similar versions in use on Venus, 
Mars, and Titan, has a single jet engine and has a life 
support system capable of providing air for up to 100 
hours. The Venusian and Martian versions both have 
a top speed of 900 kph, a wingspan of 11 m, and a 
maximum range of 5,000 km. The version designed 
for Titan has a wingspan of 8 m, a top speed of 650 
kph, and a range of 4,000 km. Jets are equipped with 
headlights, radio boosters, and radar with a range of 
up to 250 km. \textbf{[Expensive]}

\subsection{Exoskeletons}

Exoskeletons are powered mechatronic skeleton 
frameworks worn by a person. Servo-hydraulic joints 
allow the exoskeleton to be maneuvered by mimicking
the wearer's own movements, as well as enhancing
their strength. Exoskeletons may also be piloted electronically
A character wearing an exoskeleton (other
than the trike or transporter) maneuvers as normal, 
+20
8/40
100

—

30

10
+10
—
200–250

10/6

50

10
–10
8/32
2,500–3,000

24/16

100

20
+20
—
650-900

30/20

200

30
%%% 346
because the exoskeleton is like an extension of their 
own body. A character jamming an exoskeleton remotely
uses Pilot: Walker skill (except for the trike
and transporter).
\textbf{Battle Suit: }The battle suit powered exoskeleton
features a military-grade fullerene armor shell
with flexible aerogel for thermal insulation and a 
diamond-hardened exterior designed to resist even 
potent ballistic and energy-based weapons. The suit 
also enhances the wearer's strength and mobility, applying
a +10 bonus to strength-based tests, inflicting
an extra +1d10 damage and AP of –2 on melee attacks
and doubling the distance by which the character
may jump. Battlesuits are completely sealed to
protect the wearer from environmental factors, and 
fitted with life support features and a maker (p. 327) 
capable of recycling all wastes and producing air 
for up to 48 hours and food and water indefinitely. 
Battle suits are equipped with each an ecto (p. 325), 
a radio booster (p. 313), and sensors equal to specs 
(see p. 325). These suits have an Armor Value of 
18/18 (not cumulative with any other armor) and 
protect the wearer from temperatures from –175 to 
140 C. \textbf{[Expensive]}
\textbf{Exowalker:} Exowalkers are minimal framework 
exoskeletons, primarily designed to bolster the 
wearer's strength and movement. They provide a an 
Armor Value of 2/4, a +10 modifier to strength-based 
tests, and double the distance by which the character 
may jump. \textbf{[Moderate]}
\textbf{Hyperdense Exoskeleton}: These powered exoskeletons
are larger (roughly twice human-sized) and built
for heavy-use industrial purposes, such as handling 
heavy/large objects. The wearer is partially encapsulated 
to protect them from debris and industrial accidents. 
Hyperdense exoskeletons provide no movement bonus, 
but provide a +30 bonus to strength-based tests and 
inflict an extra +3d10 damage and –5 AP on physical 
attacks. They have an Armor Value of 6/12. \textbf{[Expensive]}
\textbf{Transporter:} This exoskeleton framework includes 
a pair of vector-thrust turbofan engines, giving the 
user flight capabilities in gravity and increased maneuverability
in zero-G. It provides partial protection to
the wearer with an Armor Value of 2/4. Piloted with 
Pilot: Aircraft skill. \textbf{[High]}
\textbf{Trike:} The trike exoskeleton is a three-wheeled 
personal motorcycle design, rather than a walker. It 
Battlesuit 1
Exowalker

1
Hyperdense 
Exoskeleton

1
Transporter

1
Trike

1
provides partial protection to the wearer with an 
Armor Value of 2/4. Piloted with Pilot: Groundrcraft 
skill. \textbf{[Moderate]}

\subsection{Groundcraft}

In \textit{Eclipse Phase,} trains and bicycles remain the most 
common form of ground transportation, especially on 
habitats. In larger habitats and on moons and planets, 
cycles and cars are used as well.
\textbf{Cycle}: Because of the high cost of enclosing a habitat
and providing life support, space is at a premium
in all cities except some of the newest cities on Mars. 
As a result, there is rarely room for large roads or the 
cars that once carpeted the roads of Earthly cities. 
Instead, the ubiquitous modern vehicle is the cycle, 
which is designed to drive down narrow streets only a 
little wider than sidewalks in Earth cities.
There are many different varieties of cycle. Some 
have only a single wheel and are gyro-stabilized, but 
most have two wheels and resemble old Earth motorcycles
In some, the driver and passenger are enclosed
by a streamlined pod. These vehicles are powered 
by superconducting batteries, have a range of 600 
km and a top speed of 120 kph, but must usually 
drive more slowly in crowded streets. Cycles are all 
equipped with radio boosters, headlights, and a portable
radar sensor. Tires are solid state (not inflated),
or in some cases smart spokes capable of handling 
stairs. Some luxury versions have limited life-support 
in the small cabin, capable of providing air for the 
passengers for up to 10 hours. \textbf{[Moderate]}
\textbf{Mars Buggy}: One of the most ubiquitous vehicles 
on Mars is the so-called Mars buggy, a four-wheeled 
vehicle with large balloon tires that is designed for 
use both on roads and on almost any terrain. Mars 
buggies can travel at speed of up to 110 kph on roads, 
90 kph over relatively flat terrain, and up to 40 kph 
on jagged and rocky terrain. They can maintain these 
speeds because smart materials in both the suspension 
and the tires reshape themselves to adapt to uneven 
conditions and their nuclear batteries give them an 
effectively unlimited range. Most Mars buggies are 
enclosed but unpressurized. Similar vehicles are used 
on Luna and Titan, however, though the passenger 
compartments of these vehicles includes life support 
gear that provides the occupants with air for at least 
100 hours. Buggies are powered by nuclear batteries 
—
8/32
30

18/18

60

12
—
8/40
40

2/4

30

6
—
8/20
30

6/12

100

20
+10
8/40
200

2/4

50

10
+10
8/40
120

2/4

50

10
%%% 347
and come in a variety of sizes, from small two-person 
buggies to large trucks. Mars buggies come equipped 
with headlights, radio boosters, and a vehicle radar 
system. \textbf{[High]}

\subsection{Personal Vehicles}

These one-person movement aids primarily are used in 
space, but do not count as spacecraft per se.
\textbf{EVA Sled:} This small sled uses air impellers to maneuver
in zero-G. It is commonly used to carry attached gear,
but may also pull along 1 human-sized morph. \textbf{[Low]}
\textbf{Rocket Pack}: This is a miniature metallic hydrogen 
rocket that the wearer straps to their back, with two 
rocket exhausts extending out to either side, away 
from the wearer's body or legs. Biomorphs and pod 
morphs can only safely use this vehicle when wearing
a vacuum suit or some garment that is similarly
heat resistant. Also, to prevent harm to the wearer, the 
thrust must be kept sufficiently low that it can only 
take off on Mars or moons with even lower gravity. A 
rocket pack can keep the wearer airborne for up to 15 
minutes in Mars gravity, or 30 minutes on Luna, Titan, 
or any of the four large Jovian moons. On Mars, it 
has a maximum speed of 700 kph. It can be used to 
reach orbit and land again on Luna, Titan, and other 
similarly small bodies like the Jovian moons. Rocket 
packs are equipped with radio boosters but no other 
sensors or communication devices. \textbf{[Low]}
\textbf{Thruster Pack:} Worn for EVA duties, this thruster 
pack uses vectored thrust nozzles, allowing a character
to maneuver in open space. This is not a jetpack
and does not produce enough thrust for atmospheric 
movement. \textbf{[Low]}

\subsection{Robots}

Robots are a common sight and accepted fact of daily 
life within \textit{Eclipse Phase.} Numerous varieties exist, 
from robo-pets to mechanical workers to warbots. 
If a job can be done more cheaply (and sometimes 
safely) by a bot, it usually is. The robots listed here 
are not generally used as synthetic shells by transhuman
egos, often for cultural reasons (sleeving a case
is bad enough, sleeving a creepy is just ... wrong), and 
they are not equipped to be sleeved into (though the 
may be jammed; see p. 196). Any of these bots may 
Cycle

1–3
Mars Buggy

2–6
EVA Sled

1
Rocket Pack

1
Thruster Pack

1
be modified for use as a synthetic morph, however, by 
adding a cyberbrain system (p. 300).
\textbf{Automech: }Automechs are general purpose repair 
drones, found just about everywhere. Each particular 
automech tends to specialize in a particular type of 
repair work and so carries the appropriate tools and 
AI skills, whether it be habitat waste recyclers, outer 
hull integrity, or servitor systems. Standard automechs 
are wheeled cubes with articulated limbs, though they 
are also equipped with vectored-thrust drives for 
zero-G work. \textbf{[Moderate]}
\textbf{Creepy: }Creepies are small crawler bots that come 
in an eclectic variety of shapes and forms, from robo-squirrels
to insectoids to bizarre and artsy mechanical
creatures. Creepies were originally designed as a sort of 
robotic pet, but they are commonly used as a general 
purpose household minion, like a more beloved servitor
Many people in fact wear a creepy on their person,
dropping it to handle small tasks for them and letting it 
crawl up and down and over their body. \textbf{[Low]}
\textbf{Dr. Bot:} These wheeled medical robots are designed 
to tend to and transport injured or sick people. They 
carry a healing vat (p. 326), a specialized pharmaceuticals
maker, miscellaneous medical gear, and articulated
arms for conducting remote surgery. \textbf{[High]}
\textbf{Dwarf:} These large industrial bots are named not 
just for their primary use—mining, excavation, tunneling
and construction—but because the default
AIs they shipped with had a programmed tendency 
to happily whistle as they worked. Dwarfs are quadrapedal
walkers, equipped with massive modular
industrial tools like boring drills, shovels, hydraulic 
jacks, jackhammers, scooping arms, acid sprays, and 
so on. \textbf{[Expensive]}
\textbf{Gnat:} Gnats are small rotorcraft camera/surveillance
drones. Many people use gnats for personal
lifelogging, while socialites and media use them to 
capture the glamor or hottest news. \textbf{[Low]}
\textbf{Guardian Angel:} Similar to gnats, guardian angel rotorcraft
hover around their charges, keeping a watchful
eye out to protect them from threats. \textbf{[Moderate]}
\textbf{Saucer:} These disc-shaped drones are lightweight 
and quiet. They are typically launched by throwing 
them like a frisbee, after which they propel themselves 
with an ionic drive (p. 310). Saucers make excellent 
+20
4/40
120

12/10

50

10
+10
8/32
40/90/110

30/20

150

30
–30
4/16
16

5

40

8
–20
—
700

+5/+5

40

8
–10
4/20
40

+4/+4

30

6
%%% 348
``eye in the sky'' monitors and scouts. \textbf{[Low]}
\textbf{Servitor:  }Servitors are the most common robot, 
acting as cooks, janitors, universal helpers, movers, 
and personal aides. Every home has one, if not several. 
Servitors are intentionally built in non-humanoid 
forms so as not to confuse them with common 
synthmorphs and in order to defuse bad feelings at 
ordering them around. However, they all have some 
form of ``face'' to interact with, so as not to be too 
machine-like. \textbf{[Low]}
\textbf{Speck:} Specks are tiny insectoid spy drones, 2.5 mm 
long and 2 mm wide, about the size of a small fruit fly. 
They fly with tiny wings, carry a microbug, and are 
excellent for surveillance purposes or otherwise being 
a ``speck on a wall.'' Specks are difficult to notice (–30 
Perception modifier) and almost impossible to distinguish
from an actual insect. \textbf{[Low]}

\subsection{Spacecraft}

Though egocasting is a common method of personal 
transport and it's often easier to simply transmit the 
specifications for various goods and to allow nano-factories
to create duplicates, spacecraft play an
important role in the solar system, carrying both passengers
and valuable cargo. Both in terms of materials
and propulsion, spacecraft in the post-Fall era are far 
superior to the primitive vessels used in the 20th and 
early 21st centuries, but they are still based on the 
same principles.
Spacecraft have few stats in \textit{Eclipse Phase,} as they 
are primarily handled as setting rather than vehicles. 
Automech

4/8
Enhancements: Access Jacks, Electrical Sense, Extra Limbs (4), Headlights, Magnetic System, Radiation sense, Utilitool, misc. tools
Creepy 

4/12
Enhancements:\textbf{ }+5 COO, Access Jacks, Chameleon Skin, Extra Limbs (2-8), Grip Pads
Dr. Bot

4/16
Enhancements:\textbf{ }Access Jacks, Enhanced Smell, Fabber, Fractal Digits, Healing Vat, Nanoscopic Vision
Dwarf

4/12
Enhancements:\textbf{ }+10 SOM, Access Jacks, Extra Limbs (4), Industrial Armor, Radar, Sonar, misc. tools
Gnat

8/40
Enhancements:\textbf{ }360-Degree Vision, Access Jacks, Enhanced Hearing, Enhanced Vision, Radar
Guardian Angel

8/40
Enhancements: +5 REF, 360-Degree Vision, Access Jacks, Chameleon Skin, Eelware, Enhanced Hearing, Enhanced Smell, Enhanced Vision, 

Lidar, Light Combat Armor, Neurachem, T-Ray Emitter
Saucer

8/40
Enhancements:\textbf{ }360-Degree Vision, Access Jacks, Chameleon Skin, Enhanced Hearing, Enhanced Vision, Radar
Servitor

4/20
Enhancements:\textbf{ }Access Jacks, Extra Limbs (2-6)
Speck

1/5
Enhancements: +5 REF, +5 COO, –10 SOM, Access Jacks, Grip Pads, Enhanced Hearing, Enhanced Vision, Synthetic Mask
Note also that no stats are given for spacecraft weaponry
It is highly recommended that space combat be
handled as a plot device rather than a combat scene, 
given the extreme lethality and danger involved. 
If you absolutely must know the DV of a spacecraft 
weapon, treat it as a a standard weapon with a DV 
8
4/4
30

6

Wheeled/Vector-Thrust
12
2/2
25

5

Walker or Hopper
16
—
40

8

Wheeled
20
16/12
150

30

Walker
60
2/2
25

5

Rotor
80
14/12
40

8

Rotor
200
2/2
25

5

Ionic
20
4/4
30

6

Walker or Wheeled
5
—
5

1

Winged/Hopper
%%% 349
multiplier of x3 for small craft (fighters and shuttles), 
x5 for medium craft, and x10 for larger craft.

\subsubsection{Spacecraft Propulsion}

The most important part of any spacecraft is its engine, 
and the most important features of any engine are 
the exhaust velocity, which determines how much 
fuel the rocket requires to reach a given speed, and 
the engine's thrust, which determines how high the 
acceleration can be. Any rocket that has a thrust of 
less than approximately twice the gravity of a planet 
or moon cannot take off from that planet or moon. 
Sample thrusts and gravities are listed on the \textit{Escaping }
\textit{Gravity Wells} table, p. 346.

\textbf{Hydrogen-Oxygen Rocket (HO):} Though optimized
with improved engine design and light-weight
materials, these are essentially the same primitive 
rockets that humanity used to first reach the moon 
in the 20th century. These are rarely used and only 
common with groups too poor or primitive to safely 
manufacture metallic hydrogen.

\textbf{Metallic Hydrogen Rocket (MH):} Metallic hydrogen
is a solid form of hydrogen created using exceedingly
high pressures. Although naturally unstable, it
can be stabilized with carefully controlled electrical 
and magnetic fields, and these field generators are 
an integral part of every metallic hydrogen fuel tank. 
By selectively reducing these fields near the exhaust 
nozzle, small amounts of metallic hydrogen can be 
made to swiftly and explosively revert to conventional 
hydrogen gas, propelling the rocket with great force 
in an easily controlled fashion. Metallic hydrogen 
engines are used in most planetary landers and short 
Hydrogen-Oxygen Rocket
Metallic Hydrogen
Plasma Rocket
Fusion Rocket
Anti-Matter
Rocket Buggy
\textbf{Planets, Moons, Etc.}
Earth
Europa
Jupiter
Luna
Mars
Mercury
Neptune
Pluto
Saturn
Titan
Uranus
Venus

range vehicles.

\textbf{Plasma Rocket (P):} This drive heats hydrogen into 

plasma and accelerates it using a powerful electrical 

field. This type of rocket was very common in the 

mid 21st century, but has been superseded by fusion 

rockets and is only used in older and more primitive 

spacecraft, notably scum barges.

\textbf{Fusion Rocket (F):} Similar to a plasma rockets, 

fusion rockets require significantly higher tempera-

tures and pressures, and the rocket also produces 

large amounts of power for the spacecraft. Fusion 

rockets are now the most common form of propul-

sion for spacecraft designed for long-distance voyages.

\textbf{Anti-Matter Rocket (AM):} Anti-matter rockets 

work mixing small amounts of anti-matter into the 

hydrogen fuel, producing enormous amounts of 

energy and an exceptionally fast and powerful ex-

haust. These rockets typically carry a heavily shielded 

magnetically contained anti-matter storage vessel car-

rying a mass of anti-matter equal to 1% of the mass 

of the hydrogen fuel used by the rocket. The magnetic 

containment vessels needed to safely contain anti-

matter usually weight at least 10 times the mass of 

the antimatter used.

Though anti-matter storage is exceptionally safe, 

the vast energy release possible if there was an ac-

cident means that anti-matter rockets are forbid-

den from coming closer than 25,000 km from any 

inhabited planet or moon. Also, very few habitats 

will allow an anti-matter rocket to dock with them, 

and instead require the spacecraft to remain at least 

10,000 km away and for all cargo and passengers to 

be transferred using a small craft like a small LOTV. 

Anti-matter is exceedingly expensive to produce and 

so anti-matter rockets are only used in military vessels 

and in fast couriers designed to carry critical cargoes 

across the solar system in short periods of time.

\subsubsection{Sample Spacecraft}


The following is a representative sample of the most 

common type of spacecraft used in the solar system today.

\textbf{Bulk Carrier: }This vessel is simply a standard 

transport refitted to carry large amounts of cargo in 

external cargo grapples. Used for carrying refined ores, 

ice, and similar forms of large, useful, but low priority 

cargos, bulk carriers transport large cargos at rela-

tively low velocities. They also offer an inexpensive, 

reliable, and slow method for passengers to travel 

from one habitat to another and are not infrequently 

used by individuals who wish to disappear for a while. 

Unlike the standard transport, the bulk carrier lacks 

the rotating habitat pods.

\textbf{Courier:} In a standard transport, a typical journey 

from Luna to Mars requires approximately three 

weeks, while a journey from Mars to Jupiter requires 

approximately four months. This is sufficient for most 

purposes, but occasionally characters need to take 

themselves or sufficiently valuable cargoes across the 

solar system in a matter of days or weeks, instead of 

weeks or months.
4+
3
0.01
0.05
0.2
0.5
\textbf{Gravity}
1
0.13
2.53
0.17
0.38
0.38
1.14
0.06
0.91
0.14
0.89
0.9
%%% 350
Anti-matter drive fast couriers are vessels designed
for this specific purpose. This vessel can
travel from Venus to Mars in a week and from 
Mars to Jupiter in a month. The fast courier is the 
swiftest vessels currently made and is able to reach 
at much as one half of one percent of the speed 
of light. To manage this, this spacecraft must also 
carry 6 tons of antimatter in a 100 ton magnetic 
containment vessel. In an emergency, this containment
facility can be jettisoned.
\textbf{Destroyer:} One of the largest military spacecraft 
in common use, destroyers use an antimatter drive 
holding 150 tons of antimatter in a 2,000-ton magnetic
containment vessel. This antimatter can also
be used to provide the spacecraft's missiles with 
anti-matter for devastatingly powerful anti-matter 
warheads. This spacecraft is also armed with 
railguns, nuclear and high explosive missiles, and 
point defense lasers.  In addition, all destroyers 
carry a contingent of 20 fighters.
\textbf{Fighter: }This small, short range military vessel is designed
to be crewed by an infomorph or AI. If needed,
however, it can hold a single synthmorph or vaccum-adapted
biomorph as a pilot. It carries 3 lasers and 2
railguns mounted on small pods placed around the 
middle of ship that can fire in any direction. A single 
missile launcher is located in the nose of the fighter 
and typically holds 6 small high explosive missiles or 
tactical nuclear missiles (or even anti-matter missiles 
if facing high-threat targets).
\textbf{General Exploration Vehicle (GEV)}: A GEV is one 
of the standard vehicles used for exploration beyond 
the Pandora Gates. It is specifically designed to handle 
almost any environment. It is a boxy vehicle, 6 meters 
long, 2.2 meters wide, and 2 meters high. It makes 
extensive use of smart matter in the lower part of the 
chassis, and can create wheels or short legs primarily
useful for exceedingly rough terrain). It can even
produce limited hull streamlining and propulsion 
suitable for travel both on and underwater. In addition
it contains a small metallic hydrogen engine that
Bulk Carrier
Courier
Destroyer
Fighter
GEV
LLOTV (HO)

20 (high-velocity)/100 (low-velocity)
LLOTV (MH)

250 (high-velocity)/350 (low-velocity) 
Scum Barge

20,000
SLOTV (HO)

3 (high-velocity)/30 (low-velocity) 
SLOTV (MH)

70 (high-velocity)/100 (low-velocity) 
Standard Transport
allow it to maneuver in space with an acceleration of 
up to 0.1 G. GEVs have a Maximum Velocity of 200 
(wheeled)/40 (walker)/60 (sea)/40 (submerged).
The GEV also has a closed cycle life support system 
that can support up to 6 (fairly cramped) living 
occupants for up to one month and limited electromagnetic
shielding against charged particle radiation.
All models are fitted with advanced AI piloting and 
navigation as well as limited self-repair capacity. In 
addition, GEV's have an extensible airlock, a single 
healing vat, several desktop CMs, and a variety of 
sensors, including both radar and telescopic full spectrum
cameras.
\textbf{Large Lander and Orbit Transfer Vehicle (LLOTV):}
This common vehicle is used for transporting passengers
and cargo between a planet or moon and orbit
and for short distance transfers between habitats 
less than 100,000 km apart. This conical vehicle has 
a curved heat shield on the base and smart material 
landing legs and grapples so that it can rest securely 
on any stable terrain and link up with all forms of 
docking clamps. It comes in variants designed to use 
either a hydrogen-oxygen chemical rocket or a metallic
hydrogen rocket. The use of light-weight smart
materials allows the interior to be easily and rapidly 
reconfigured to accommodate different amounts of 
fuel, passenger seats, and cargo space. LLOTVs that 
are not designed for planetary landing or which are 
designed only to land on airless moons are unstreamlined
and look considerably blockier.
LLOTVs come in two configurations: high or low 
velocity. High velocity configuration allows the vehicle 
to land and take off again on Venus or Earth without 
refueling and for rapid transport between nearby 
habitats. Low velocity configuration is designed to 
land and take off again on Mars or various large 
moons without refueling and for slower and more 
fuel efficient transport between nearby habitats. The 
extensive use of smart materials in this vehicle means 
that LLOTVs that use metallic hydrogen engines can 
be easily converted between the high and low velocity 
110
—

20

750

150
13
—

15

500

100
90
—

30

2,000

500
1
+30

20

240

60
6
–10

15

200

40
–10

20

800

160
–10

20

800

160
—

20

1,500

150
–10

20

400

80
–10

20

400

80
200
—

20

750

150
%%% 351
configurations, requiring less than a day in a well-equipped
maintenance facility. However, vessels using
hydrogen oxygen engines cannot be converted. Since 
metallic hydrogen is a much more efficient propellant, 
landers using it always include significant amounts of 
extra propellant for emergencies.

\textbf{Scum Barge:} These huge craft were originally designed
for use during the first stages of the evacuation
of Earth. They were built to carry up to 20,000 people 
and to allow them to survive for months or even years, 
in relatively cramped conditions, until more suitable 
habitats could be constructed. A number of these vessels
are still in service, primarily used as mobile habitats
by various anarchic subcultures. The best have
had their plasma rockets replaced by modern fusion 
rockets and carry 5-10,000 in relative comfort. The 
worst use aging plasma rockets and stretch their life 
support systems and living spaces to the limit, carrying
up to 25,000 poor and desperate residents.

\textbf{Small Lander and Orbit Transfer Vehicle (SLOTV): }
This vehicle is identical in use and design to the 
LLOTV, except that it is one third the total mass and 
correspondingly less expensive to build and refuel. 
Some exceptionally wealthy individuals own private 
small LOTVs. Using a small LOTV with a hydrogen-oxygen
engine to take off and land on Venus or for
other high velocity uses is exceptionally cramped 
and allows for absolutely no room for error. Like the 
LLOTV, this vehicle can be easily converted between 
low and high velocity configurations and is made in 
both streamlined and non-streamlined versions.

\textbf{Standard Transport}: This vessel is one of the most 
common freighter and passenger vessel in the solar 
system. While egocasting is by far the most common 
form of inter-habitat transport, some people prefer to 
travel by ship and others do not wish to leave their 
current morph behind. In addition, some goods are 
easier or cheaper to physically transport rather than 
duplicating their templates. As a result, standard 
transports regularly travel to and from every large 
habitat and inhabited planet and moon in the solar 
system. These are modern fusion-drive ships that offer 
fast and comfortable travel for passengers as well as 
relatively swift transport for small cargoes.

One of the additional benefits of the standard 
transport is the fact that it contains four separate passenger
compartments, each of which is mounted on
a 90 meter-long booms that can extend and rotate 
to simulate gravity. When rotating at a comfortable 2 
rpm, passengers experience Mars level gravity. Typically
the gravity maintained in these pods starts at
the local gravity (or Mars gravity, if the local gravity 
is higher) and over the course of the journey gradually 
increases or decreases to the gravity of the destination. 
However, these pods cannot rotate to produce gravity 
higher than that found on Mars.
