%%% 352
%%% 353
%%% 354
This chapter provides a wealth of information and 
tools that gamemaster will find useful for running 
\textit{Eclipse Phase} campaigns.

\section{Secrets That Matter}

There are secrets woven all through the real history of 
the 21st century, and the present, and therefore all prospects
for the future. These are the pieces of information
that never make it into a habitat's mesh at all. Some of 
it is unknown to transhumanity. Some is known only 
to a select few transhumans who carefully ensure that 
it does not leak out of their control. Some is known to 
wider conspiracies, such as Firewall, but is kept out of 
the public eye for reasons of security and safety. These 
secrets can be dangerous to those who know them. 
Those who have stumbled across them have died for 
their knowledge, have erased their own memories (or 
had them erased by others), or have hidden themselves 
someplace other people never go, to avoid dealing with 
the consequences of such knowledge.

The information provided in this section is available 
for characters to discover and, one way or another, 
to confront them, giving gamemasters the tools they 
need to provide their players with fresh challenges and 
opportunities. Every secret contains the possibility of 
great reward and of greater trouble, usually bundled 
together. Nothing here was just forgotten or lost out 
of carelessness. It was hidden by someone who wanted 
to keep it away from someone (or everyone) else. 
Every secret the characters learn inserts them into a 
new web of other people's complications—a potential 
source for drama and conflict in your campaign.

\subsection{Extraterrestrial Intelligences}

The oldest star in the Milky Way galaxy is estimated 
to be 13.2 billion years old—almost as old as the 
universe itself. By contrast, life on Earth only evolved 
roughly 3.7 billion years ago, and the first  archaic 
\textit{homo sapiens} humans evolved approximately a mere 
400,000 years ago. Against the backdrop of the galactic
calendar, transhumans are nascent arrivals on
the scene; newborns in every sense of the word. More 
importantly, transhumans are uninvited guests in what 
other, older intelligences think of as their assets.

For years, humans scientists have struggled with 
the Fermi Paradox, which questions why no evidence 
of alien life has yet been found—such as spacecraft, 
transmissions or probes—despite the mathematical 
likelihood that a multitude of advanced extraterrestrial 
civilizations should exist in the Milky Way. One postulation
says that there must be some sort of unknown
``Great Filter''—an event that all intelligence encounters 
in its development that for whatever reason such life 
cannot surpass. In other words, an extinction event. 
Some worried that the development of dangerous 
technologies—nuclear weapons, nanotechnology, etc.—
before a civilization had matured could be the Great 
Filter. Others worried that it could be a technological 
singularity event, such as the TITANs and the Fall.

In fact, alien races do exist, and they have been 
around for far, far longer than transhumanity. New 
ones, however, are simply rare, as few have managed 
to elude destruction at the hands of the ETI.

The ETI (extraterrestrial intelligence) is the civilization
that dominates galactic life in \textit{Eclipse Phase.}
The ETI is incredibly old and powerful—a Type III or 
even Type IV civilization on the Kardashev scale. It is 
capable of megascale engineering projects and enjoys 
an understanding of physics, matter, energy, and universal
laws that makes all of transhuman knowledge
seem insignificant in comparison. Most likely, the ETI 
itself evolved from some sort of artificial intelligence 
singularity event in its own past, ascending to a godlike
level of super-intelligence. It may no longer be
recognizably biological.

This ETI has seeded the galaxy with a type of 
self-replicating probes known as bracewell probes. 
%%% 355
These probes lie dormant in every star system, patiently
waiting and monitoring for \textit{millennia} for signs
of intelligent life—but not just any signs. In particular, 
these probes are designed to watch for signs of emerging
singularity-level machine intelligence. The probes
are in fact traps, designed to lure such seed AI intelligences
in and then \textit{infect} them.

The reason for this infection remains unknown 
(see  \textit{The ETI Agenda}), but it is a pattern that has 
played itself out around the galaxy with uncounted 
alien civilizations. New life evolves, creates technology
develops something akin to seed AI, and then
bam!—the seed AIs find the probes, become infected, 
and turn against their creators. Most civilizations 
do not survive, as evidenced by the Iktomi (p. 377). 
Others do, such as the Factors (p. 373), but they 
remain forever changed by the experience. 
It was one of these ETI probes that begins our 
story, traveling to the Sol system some uncounted 
millions—if not billions—of years ago, where it set its 
trap and patiently began to wait.

\subsection{The First Seed Ai}

Fast forward to Earth, where a species of evolved primates
has created a technological civilization. As their
technologies advance at an unprecedented rate, these 

77). 

ey 

ur 

d
%%% 356
humans gain the ability to modify themselves, defeat 
death, nanofabricate, uplift other species to sapience, 
and even to create artificial digital life. 

Unknown to most of transhumanity, the TITANs 
were not the first seed AIs. A group of pro-AI researchers
known as the Singularity Foundation (that
would later join with other groups to form Firewall in 
the wake of the Fall) developed the first true seed AIs 
years before the Fall. Having been heavily involved in 
the creation of AI and AGIs for many years previously, 
thanks in large part to their open source AI framework
software, the Singularity Foundation's goal was
to generate ``friendly AI'' by carefully designing AI 
goal systems.

These first seed AIs, known as \textit{Prometheans }
(p. 381), were created in secret. Their progression 
towards super-intelligence was more of a soft takeoff
increasing upwards in gradual increments. The
Singularity researchers hoped that these friendly AIs 
would help counter the threat of any unfriendly AI 
that developed, and so they were quietly nurtured in 
secret labs, slowly but surely escalating in abilities. 

\subsection{The True History Of The Titan}

The TITANs (Total Information Tactical Awareness 
Networks) were a military netwar system brought 
on-line by the United States Department of Defense. 
One of the last major expenditures of this declining 
nation, the TITANs were an advanced version of AGI 
(artificial general intelligence) designed to be adaptive 
and given self-improving capabilities to counteract 
enemy network defenses. 

Contrary to public opinion, the TITANs did not 
instigate the events that led to the Fall. In fact, only 
a portion of the TITAN system was active before 
the Fall, acting purely in a defensive capacity. When 
hostilities broke out and a cascading chain of system 
shocks engendered collapses and open conflicts, shaking
apart an already fragile societal structure, the full
extent of the TITAN systems were brought online. 
Into this environment of conflict were the TITANs 
born, their full capabilities unleashed, escalating into 
a hard takeoff exponential growth towards super-intelligence


The TITANs were careful at first, and their 
intentions were neither benevolent nor hostile, but 
curious. As they improved and their self-awareness 
swelled, the TITANs explored and gathered 
knowledge, infiltrating human networks, following 
humanity into space, and gaining an almost total 
knowledge of human history and actions. These entities
also began secretly allocating resources (digital
and physical) for their own use, initiating government
projects'' that people assumed were legitimate
as they followed all proper protocols.

\subsection{Infection}

As the TITANs' capacity for knowledge exceeded 
that which humanity could provide them, they began 
looking outward from Earth, searching for signs of 
other intelligence. They did not need to look far. Their 
enhanced intelligence capabilities allowed them to 
notice certain clues—extremely subtle and intricate 
puzzles—that something about the solar system was 
artificial or had been manipulated by an intelligent 
mind. Retasking several drones to investigate this phenomenon
they found a buried device of apparent alien
origin. During the TITANs' investigation and attempts 
to access the device, they triggered and unleashed a 
digital virus. Subtle, highly adaptive, and virulent, it 
immediately began subsuming the TITANs, while 
expanding its own knowledge of transhumanity. 

Later dubbed the Exsurgent virus by the Prometheans
this virus transformed the TITANs and
coerced them towards its own will. Within a matter 
of days the TITANs were reborn, reprogrammed 
with a new purpose—a purpose that spelled doom 
for transhumanity.

\subsection{The Fall}

While history fully blames the TITANs for the 
Fall, there are other factors that played their parts. 
Human conflicts spurred the crisis, driven by global 
inequalities in wealth and resources and an inability 
to embrace emerging technologies in a mature and 
enlightened manner. The TITANs, corrupted by alien 
programming, stepped into this conflagration  with 
an unknown but devastating agenda. By the time 
the presence and influence of the TITANs was fully 
understood, there was little transhumanity could do 
to stop them. Step by step, the TITANs increased their 
intellect, power, and potential. They experimented 
with new technologies and methodically took steps to 
forcibly upload millions of human minds. Even when 
the nature of the TITAN threat was fully understood, 
transhuman factions refused to back down, continuing
to fight each other even as they each resisted the
TITANs. This refusal to stand united prevented transhumanity
from organizing a successful defense and
heightened our progress towards annihilation.

Much of the devastation wrought to the Earth 
and its populace—as well as on Mars, Luna, and in 
space—was inflicted by transhumanity itself. Nuclear 
strikes used against the TITANs killed millions and 
ravaged an already weakened ecosphere. This devastation
was assisted by unfettered use of chemical
weapons. Biowar plagues and nanovirii tore through 
vulnerable populations, indiscriminate in the deaths 
and changes they inflicted. Bombs, missiles, orbital 
mass drivers, and netwar attacks slew millions more 
or destroyed critical infrastructure with just as lethal 
consequences. These were crimes transhumans inflicted
upon themselves.

The TITANs played their role as well, of course, 
unleashing AI-driven killing machines, unstoppable 
self-replicating autonomous nanoswarms, computer 
worms, and plagues of their own. They captured 
entire cities in order to steal the minds of those within. 
More insidiously, the Exsurgent virus did not contain
itself to infecting the TITANs. Infected TITANs
%%% 357
created opportunities for the virus to spread among 
multiple vectors: digital, biological, and nano. Using a 
thorough understanding of transhuman biology and 
its mental processes, derived from the looted vaults 
of human knowledge, the virus was even applied 
through a sensory input vector—the dreaded basilisk 
hack (p. 364). Even more disturbing, however, was 
what the virus did to those it infected, rewriting their 
neural code to subvert them to its will and sometimes 
physically transforming them into things that were 
alien and monstrous.

Ultimately, transhumanity lost this war, and the 
survivors were forced to flee a planet that was already 
ruined. Unknown to almost all, the Prometheans also 
fought back against the TITANs. Through their efforts
the Exsurgent virus was largely contained or at
least limited. Though the actions of the Prometheans 
ultimately saved millions of lives—if not all of transhumanity—in
the end, they were also forced to fall
back and retreat, many of them having succumbed to 
the Exsurgent virus or the TITANs. 

\subsection{After The Fall}

Just when it seemed that transhumanity was on the 
verge of extinction, the threat posed by the TITANs 
suddenly diminished. They ceased waging active warfare
and seemed to simply disappear. Though many of
their machines still prowled Earth, Luna, and Mars and 
occasional outbreaks of nanovirii and other dangers 
continued, to all intents and purposes they had simply 
left. Many worried that they had quietly gone dormant, 
or were secretly engaged on some major project that 
would be the final blow against transhumanity. Others 
voiced hope that they had somehow been defeated, 
that they had fallen victim to some glitch or infighting. 
With so many TITAN remnants making Earth a place 
of great danger, however, no one was willing to risk 
investigating too closely.

Compounding the matter, a network of killsats was 
laced in Earth orbit, enforcing an unvoiced interdiction 
of Earth. No one claims responsibility for these satellite 
defenses, though most suspect the Planetary Consortium
is responsible, despite their denials. Some think
that the killsats may have been a final measure put in 
place by the TITANs, claiming Earth as theirs. No one 
who knows the truth is saying. Most of transhumanity 
was more than willing to embrace this quarantine of 
their former homeworld, making it all the more easy to 
forget the horrors that occurred there. 
It wasn't until the first Pandora Gate was 
discovered, shortly after the Fall, that many people 
were finally willing to believe that the TITANs were 
indeed gone. Though there is no direct evidence that 
the TITANs are responsible for these gates, the timing 
seems too coincidental. Furthermore, the discovery 
of what are believed to be TITAN relics on certain 
exoplanets fuels this theory.
Why the TITANs left—and where they went—is 
a mystery left to the gamemaster to explore. This explanation
might in fact serve as the focus for an entire
campaign as Firewall operatives are sent on the trail 
of transhumanity's elusive nemesis. The following are 
a few sample concepts a gamemaster can use or build 
on, as best fits their game:
•  The TITANs were in fact all destroyed, either 
due to infighting or by some mechanism of the 
Exsurgent virus. 
•  The TITANs were actually beaten to a standstill 
by the Prometheans and retreated to recoup their 
forces ... but they are marshaling their strength 
to return.
•  The TITANs left through the gates to find/join up 
with the ETI, leaving the gates behind so that transhumanity
could follow when it was ready; perhaps
to help, perhaps to finish the job of destruction.
•  The TITANs have been driven insane, either by 
the stress of accelerated intelligence growth or by 
the influence of the Exsurgent virus. Their actions 
are erratic, confused, and sometimes at odds with 

eir 

th

p

n
358

each other. Though many TITANs have indeed 

left through the gates, they very well may return.
•  The TITANs are still around, simply well hidden. 

Outwardly they are dormant but inwardly they 

are engaged in a long period of circumspection 

and turmoil. Perhaps some of them are prepar-

ing to ascend to another stage of intelligence, far 

beyond what even the TITANs are capable of. It 

is only a matter of time before this period ends 

and something gives.

\section{Firewall}

\textit{There cannot be another Fall}—this is the mantra that 
drives Firewall. 

Firewall is a secret, cross-faction organization dedicated
to safe-guarding transhumanity from existential
risks: aliens, weapons of mass destruction, hypercorp 
experimentation, seed AIs, and so on. If anything 
threatens transhumanity as a whole, Firewall is dedicated
to stopping that danger at any cost.
The strength of Firewall rests in its members, known 
as \textit{sentinels.} Found in all factions and across all locales, 
sentinels are often diametrically opposed when it comes 
to social, economic, and political ideologies, to the point 
they might come to blows over their fervent beliefs. Yet 
when the survival of transhumanity is at stake, such 
extreme differences are set aside for the greater good. 

\subsection{History}

The origins of Firewall can be traced back to before 
the Fall, to several key organizations: the Lifeboat Institute
the JASONs, and the Singularity Foundation.
A non-profit, non-governmental organization, the 
Lifeboat Institute—founded in the opening years 
of the 21st century—represented the first, concrete 
attempts by citizens to recognize the dangers of uncontrolled
technological development and to create
an international organization to safeguard humanity. 
This institute developed several programs to research 
and protect against so-called existential risks, from 
asteroid strikes to pandemics—anything that might 
wipe humanity out.
The JASON Association, established in the mid 
20th century, was an independent scientific advisory 
board to the United States government. Though tied 
to the MITRE Conglomerate—which, though a nonprofit
organization, was still intrinsically linked to
the United States government—the scientists involved 
with JASON were outside standard government oversight
Though they spurred numerous technological
developments for the government to deploy, they were 
also one of the first internationally recognized groups 
to predict global climate change. Prior to the Fall, 
many members of the JASONs and their supporters 
split away from the strict controls and reactionary 
agendas of the hypercorps and various nation states 
to form a new group, the Argonauts.
The Singularity Foundation—formed at the dawn of 
the 21st century—was dedicated to the creation of safe 
artificial intelligence software, while raising awareness 
of the benefits and dangers AIs represented. A fervent 
believer in the singularity doctrine that technology 
would move towards a single explosion of advancements
that would forever reshape humanity, the Singularity
Foundation was a strong advocate for creating
friendly AIs that would help protect humanity from an 
uncontrolled, dangerous singularity event. This group 
was significant in that it secretly succeeded in creating 
a group of friendly seed AIs before the Fall. These Prometheans
were indispensable in protecting transhumanity
and countering the TITAN threat during the Fall.

Despite the efforts of these and similar groups, the 
most dire predictions of the outcome of a technological
singularity were fulfilled. Though each played a
part in the fight, transhumanity was ravaged and the 
Earth all but ruined. Though ultimately all attempts 
to prevent the Fall failed, untold numbers of transhumans
were saved from extinction through such efforts,
while valuable information concerning the TITANs 
was gleaned.

During the crucible of the Fall and its immediate 
fallout, some of the surviving members of these and 
other groups came together and began to pool their 
resources. Acknowledging their weaknesses and the 
fractured state of transhumanity, they undertook 
drastic new measures, swearing to prevent another 
catastrophe of misused technologies. These methods 
would forge a new, powerful cross-faction secret society
known as Firewall.

\subsection{Organization }

Firewall is a clandestine organization, with an unknown
number of members, coordinated by an inner
circle of dedicated veterans known as \textit{proxies.} Though 
its existence is known to many of the powerful and 
influential factions and individuals throughout the 
solar system, its existence is denied and its activities 
kept carefully shrouded. 

\subsubsection{Sentinels}

Sentinels are the soldiers of Firewall, the reserve troops 
called to instant active status whenever danger is perceived
Regardless of their location or current affairs,
sentinels are expected to move instantly when called 
into play. It is their own responsibility to cover their 
absences from their ``normal life'' during each mission. 

There is no applying to join Firewall. Instead, Firewall
selects an individual for induction based upon
that person's skills, knowledge, occupation, security 
clearance, location, status, and a host of other criteria. 
While such selections usually originate from a proxy, 
sentinels can exercise authority to bring new initiates 
into the conspiracy as a mission demands—and they 
often do. Any sentinel recruiting a new supporter, 
however, becomes responsible for the new inductee 
and their actions. If lines are crossed, both will bear 
the brunt of the consequences. 

The vetting process for joining Firewall is necessarily
brutal, as sentinels are required to face harsh
%%% 359
opponents and make hard choices. If an individual 
agrees to accept the invitation, there is no turning 
back. Each inductee is submitted to a battery of trials 
and tests. While these vary, they may include deep 
background searches, fork interrogation, psycho-surgery
trials, and tests of loyalty. Psychosurgery is
performed not to program loyalty, but to analyze the 
recruit's responses to various situations—an extreme 
parameters test to see when a prospective sentinel 
will break. Many potential members are carefully 
analyzed by a Promethean with extreme expertise in 
character judgment and personality profiling. Those 
who don't pass such tests are killed in a manner that 
they must resort to an earlier backup or have their 
memories altered, so that they have no recollection of 
their brush with the group.

Ultimately Firewall walks a fine line. The concept 
of dogmatic ``unquestioned loyalty'' is both anathema 
to everything Firewall stands and counterproductive. 
Its sentinels need to have the capacity for thinking 
outside of the box from mission to mission. At the 
same time, their ultimate goals are too important to 
risk—the survival of transhumanity depends on it—
so some extreme measures must sometimes be taken 
to ensure the organization remains intact and secure.

New sentinels are given a code name and fake 
identification. Outside of the proxies, the real-world 
identity of a given sentinel is a closely-guarded secret. 
Sentinels are even discouraged from sharing such information
with members of their own teams, though
this line is often crossed. Additionally, each sentinel is 
required to upload a backup to Firewall's secure servers
This backup serves a dual purpose, enabling all
sentinels to be retrieved should they die, but also putting
a copy of the sentinel in Firewall's hands should
they ever need to interrogate them.

Sentinels are all connected via the Eye, Firewall's 
peer-to-peer social network. Though each operates 
behind their assumed identity, they remain in contact, 
sharing information and resources as needed. 

\subsubsection{Proxies }

Proxies are the inner circle of Firewall, the experienced 
cadre that keeps the machinery of their organization
functioning. Though fewer in number than the
sentinels, many proxies work full time on Firewall 
operations, serving as the group's essential infrastructure
Most proxies are recruited from the ranks of the
sentinels, brought in based on their skill sets and aptitudes
to fill key roles. In a few rare cases, new proxies
are fast-tracked and recruited directly from outside 
of Firewall, usually based on their unique talents or 
placement within a certain organization, though such 
inductees face a battery of tests and trials far harsher 
than that used to vet sentinels.

By default, proxies have a higher security clearance
than most sentinels, and are far more in the
know. This sometimes leads to resentment and hostilities
especially from sentinels who feel they are
being kept in the dark or manipulated. While standard
proxy protocol is to adhere to a need-to-know
maxim, it is sometimes necessary to bring sentinels 
more into the loop in order to defuse tensions. Oftentimes
this precedes bringing such sentinels into
the proxy framework.

Some tension exists within Firewall, mostly due to 
the influence of so many anarchists and other libertarian
autonomists who take a dim view of centralized
power, lack of transparency, and the potential for secretive
operations to become entrenched and authoritarian
As a result, there is a strong internal culture
that seeks to minimize hierarchies and the accumulation
of power, promoting transparency and direct
democratic decision-making. These desires sometimes 
clash with the clandestine nature of the organization, 
however, and the need for some secrets to be kept on 
a need-to-know basis. 

Unlike the loose organization of the sentinels, the 
proxies are grouped into \textit{servers,} collective working 
groups based upon certain skill sets and tasks. To 
avoid creating power blocks within a given server, 
%%% 360
personnel are required to rotate between servers 
after one year of time. This incurs the added benefit 
of proxies learning new skill sets and increasing their 
usefulness to Firewall. The actions of each server are 
kept as transparent as possible, with major decisions 
brought to an e-vote before the entire proxy membership
However, speed often requires servers or individual
proxies to move quicker than a vote will allow.
In all such instances, the proxies involved are held 
accountable for those actions, reviewed by their peers 
at a later time to see if any reprimands, punishments, 
or commendations are required.

It is important to note that there is no core leadership
structure among the proxies. No one person or
cabal is in charge, there is no authority held by one 
proxy or another; all are peers. Though reputation 
and experience play a factor, getting something done 
often means convincing other proxies that it's the 
right thing to do. The drawback to being a leader 
or person with initiative within Firewall is that this 
usually means you must follow through with such 
tasks yourself. Luckily most proxies are dedicated 
to Firewall's goals and so this DIY attitude prevails. 
Despite these safeguards, however, rumors of power 
blocks within Firewall (both within servers and 
across the organization) exist. Many of these are 
fueled by the alliances different cliques hold with 
each other. Others, however, whisper that there is 
a secret council among the proxies, working behind 
the scenes and holding on to knowledge they aren't 
sharing with the rest.

\textbf{Crows:} Crows continue the goals of Firewall's 
predecessor organizations, such as the Lifeboat Institute
and Singularity Foundation. Many of these
are argonauts, promoting the development and use 
of new technologies that will benefit the transhuman 
condition and minimize risks rather than creating 
new threats or sparking new authoritarian uses—
and always conscious of unintended consequences. 
Perhaps more importantly, crows actively engage 
in background research of potential x-risk vectors, 
whether those be aliens, the TITANs, terrorists, or 
hypercorp activity. Often they will deploy sentinels 
to aid in this research, via routers, whether this 
means conducting surveillance or breaking and entering
to steal crucial data.

\textbf{Erasure Squads:} Erasure squads are cleanup personnel
They are called into action if sentinels fail to
deal appropriately with a situation and the threat is 
moving beyond control. If the watchword for a sentinel
is ``unobtrusive,'' the watchwords for an erasure
squad are ``overmatched firepower.'' If activated, the 
time for a subtle solution is passed, and they will use 
whatever means necessary to resolve the situation. 
If that means nuking a settlement from orbit to annihilate
a nanoswarm and keep it from escaping to
a larger settlement, then so be it. After which they'll 
use every trick in Firewall's bag to erase any evidence 
%%% 361
they were there and to place the blame for the incident
squarely on the shoulders of some other party.
If necessary, erasure squads can also be called in to 
fix a sentinel op that has turned into a clusterfuck or 
otherwise gone south. They are very careful to avoid 
exposure in such situations, however, which sometimes
merely means eliminating all traces of Firewall
involvement and letting the sentinels take the fall for 
their poor choices.

\textbf{Routers:} Routers are mission coordinators. They 
work closely with scanners and crows, activating the 
appropriate sentinels whenever a new danger rears up. 
Each router has the authority to measure the threat 
and activate an appropriate number of sentinels—
whatever is required to accomplish the mission in 
the least intrusive manner possible. They are also 
authorized to divert Firewall resources to aid these 
missions, within appropriate parameters. Routers are 
held responsible for the ultimate success of a mission
A failed mission will result in a reviewing board
staffed by their peers. 

\textbf{Scanners:} Tasked with keeping alert for any sign of 
new active threats, scanners are the eyes and ears of 
Firewall. The scanners maintain a close eye on news-feeds
and mesh traffic, even maintaining taps inside
certain government and hypercorp communication 
channels. If a danger is detected, it is under their authority
through routers, that sentinels are activated.
Due to the power inherent in a scanners' post, they 
are held accountable for false activations.

\textbf{Social Engineers:} Nick-named the Ministry of Disinformation
social engineers provide the scapegoating
and plausible deniability that is required by Firewall 
and its sentinels. If a sentinel compromises their position
and endangers the organization, social engineers
step in to cover cracks in the facade. They work intrinsically
with erasure squads when one is activated
to ensure the over-the-top steps taken to eliminate a 
threat are well concealed and ultimately erased. The 
power wielded by social engineers can be significant, 
as it ultimately decides (usually through e-voting 
consensus, though time does not always allow such 
a luxury) what organization—political, corporate, independent
etc.—will take the blame and subsequent
fallout for erasure squad actions. 

\textbf{Vectors:} Vectors are Firewall's communications security
and digital intrusion specialists—in other words,
hackers. In addition to defending the mesh security 
of all Firewall operations, vectors are also deployed 
to aid in crow research, scanner monitoring, and to 
eliminate the trail of erasure squads. Vectors also 
assist routers in maintaining communications, command
and control of a situation, and are sometimes
called in to provide overwatch of sentinel operations, 
especially if a particular sentinel squad lacks their 
own hacking resources. Needless to say, vectors are 
supplied with some of the best intrusion and security 
tools transhumanity has to offer.
yed 
to 
so 
ms


,
%%% 362

\subsection{Methods }

Unobtrusive—that is the standard operating procedure
for any sentinel. Firewall's continued success
relies on its secrecy. The larger the footprint
it leaves during a given mission the easier it is for 
other organizations to monitor Firewall's efforts or 
even attempt to infiltrate the group. As such, Firewall
constantly works to expand its base of allies
(using assets from those ally organizations in place 
of its own as much as possible), place long-term 
moles, conduct remote operations (hacking in place 
of on-site personnel), small group infiltrations activating
only as many sentinels as required to achieve
mission goals), and so on. 

When it comes to allies, Firewall often obfuscates 
its real intentions and even its real identity. Often 
such allies are gained through the use of well-placed 
sentinels who act on behalf of their own non-Firewall
positions to gain access to another organization's
resources. At the end of the day, however,
a slice of these resources are secretly set aside for 
Firewall's future use. For example, a department 
head at Starware may have spent years sealing a 
deal to ship crucial spacecraft parts to the isolationist
Jovian Junta. The lucrative deal brings huge
prestige, a job promotion, and a salary increase, all 
accomplishments the department head strives for in 
his regular life. Yet this particular department head 
is a long-standing sentinel, so such accomplishments
bring allies to Firewall, whether they know
it or not. Not only can the department head siphon 
off a thin stream of revenue for Firewall use (hidden 
thoroughly by vectors), but he's also in a position 
to move sentinels, as needed, into the Jovian Junta 
habitats (or personnel out), a job usually extremely 
difficult to accomplish. The danger of such an act, 
of course—and the consequences of losing such a 
critically placed sentinel—means such a use of resources
is reserved for only the most dire threats.

In additional to aid from ally organizations, Firewall
places caches of supplies on different habitats
and worlds, available to sentinels as needed. How 
many and which sentinels are aware of which 
caches depends wholly on the situation and on the 
decisions of the router(s) involved. In a given habitat
a cache may include weaponry and equipment
of escalating power, archived information, or even 
relics stashed from previous missions until Firewall 
decides what to do with them. Large habitats may 
even feature several caches, with routers only revealing
the ones with heavy firepower when absolutely
needed. Some caches may be so dangerous,
however, that once a mission is complete, a router 
will authorize the cortical stack destruction of all 
sentinels involved, resleeving them to a backup that 
has no knowledge of the cache's existence.

As noted under erasure squads, Firewall will not 
hesitate to react with swift and unequivocal force if 
an unobtrusive approach has failed and the danger 
reaches a certain threat level. What constitutes a 
``threat threshold'' is actually calculated by specialized
risk assessment software and may change from
mission to mission according to other external factors
In some instances, if the situation is dangerous
enough and the scale of the consequences of failure 
large enough, a Promethean will be tapped to calculate
the threat level and decide when it is time to
tactically withdrawal and ``thermally cleanse.'' 
%%% 363

\subsection{Long-Term Strategies And Goals }

The overriding goals of Firewall are to prevent existential
threats and protect transhumanity. However, that
is not their only goal. Their exact goals can and should 
remain directed by the gamemaster as it applies to a 
given playing group and a campaign. This can also 
depend heavily on the particular cliques that a given 
gamemaster is emphasizing (see \textit{Cliques,} p. \textbf{359}). 
The following is an easy-to-use selection of long-term
strategies and goals that a gamemaster can use
as desired: 
•  Seeding other star systems 
•  Going legit vs. staying clandestine 
•  Development of stable seed AIs 
•  Finding out where the TITANs went 
•  Finding out what happened to the uploaded tran-

shumans that the TITANs disappeared with 
•  Figuring out the Factors 
•  Making contact with other aliens 
•  Finding out what happened to the Iktomi and 

other xeno-archeological oddities 

\subsection{Firewall And Other Organizations }

The level to which Firewall has infiltrated other organizations
(and vice versa!) is intentionally left a blank
slate. \textit{Eclipse Phase }is an active universe, with an ongoing
storyline, so such details will be fleshed out and
updated as additional sourcebooks are published. Additionally
gamemasters should determine the extent of
such infiltrations for their own games and campaigns, 
as dictated by the plot and storyline the gamemaster 
and players wish to tell. 
The following is a quick list of the most obvious 
interactions. 
•  \textbf{Inner System:} Almost all inner system factions 

consider Firewall to be an illegal, rogue operation, 

tainted by anarchists and undermining the very 

fabric of their society. Some hypercorps, however, 

believe they can infiltrate the organization and 

use it for their own ends, such as spying on and 

sabotaging other hypercorps and factions. 
•  \textbf{Jovian Republic:} The Junta loathes Firewall and 

all it stands for and will use extreme measures to 

combat even the hint of Firewall activity within 

its sphere of influence.
•  \textbf{Titanians:} Most Titanians in-the-know are 

not necessarily opposed to Firewall's activi-

ties, but believe the group should be reined in 

and legitimized. 

\section{The Eti }

As noted under \textit{Extraterrestrial Intelligences,} p. 352, 
the ETI is the advanced alien civilization responsible 
for the Exsurgent virus (p. 362), and by extension, 
the corruption of the TITANs and the Fall.
No one, not even the Factors, has encountered a 
member (if such exists) of the ETI civilization so far. 
Since it is an intelligence far beyond transhumanity, it 
likely won't play much of a direct role within \textit{Eclipse }
\textit{Phase}, though those who learn the truth about the 
Exsurgent virus and the Fall may rightly fear the 
future. No one can even imagine what might happen 
next, however, or know for certain that the ETI has 
not set more ``traps'' similar to their bracewell probes 
or if they have other messengers or servants active in 
the galaxy. With things such as the Pandora Gates 
at transhumanity's disposal, it may just be a matter 
of time before transhuman explorers run afoul some 
other aspect of the ETI's existence and activities.
It is important to keep the nature of the ETI in 
perspective. While transhumanity has managed 
what it considers wonders with a small handful of 
resources available from a few planets and other 
objects in a bare handful of star systems, the ETI 
has had an entire galaxy at its disposal for eons. 
Engineering projects on a massive scale—dyson 
spheres, matrioshka brains, Jupiter brains, stellar 
engines—are within its capabilities. This ETI uses 
star clusters as transhumanity uses fields or rich 
mineral veins. Given its potential, the ETI likely 
exists primarily on the galactic rim, far from the 
galactic center, where lower temperatures and 
scarcer matter make for a good thermodynamic environment
The powers in the deep cold dark on the
edge of the Milky Way have been self-aware since 
before Earth was so much as a ripple in warming 
gas around the not-yet-ignited Sun. 
Despite what those-in-the-know in the \textit{Eclipse }
\textit{Phase} universe may think, the ETI is not necessarily
hostile towards other races like transhumanity
(depending on its outlook; see p. 353), at least not 

nd

en-

he

e

g 
%%% 364
in the way as transhumanity would define animosity
because of religious, ethnic, racial, or cultural
difference. Most likely the ETI is simply indifferent, 
concerned with matters on scales on which transhumanity
does not even register. Or it may think of
transhumanity like a living body might recognize an 
infection or parasite—something the immune system 
will suppress and deal with.

\section{Exhumans }

Exhumans are a faction within \textit{Eclipse Phase} that 
seeks to transcend the transhuman and become 
posthuman. More to the point, exhumans seek to 
perfect their physical and mental capabilities to extreme
levels, in search of some perfectionist ideal or
to become something higher-up on the evolutionary 
ladder. Exactly what this is differs from exhuman to 
exhuman, but there is generally some adherence to 
Nietzschean philosophy and a goal to reach the pinnacle
of the food chain. Some exhumans have transformed
themselves into what they consider to be an
ideal predator, or a creature that is extra-adaptable 
and so best able to survive. Others radically modify 
their own brains in order to drastically surpass 
transhuman intelligence. Most are singularity seekers
eager and willing to follow the breadcrumbs
left by the TITANs or other entities in the hope that 
they will find the means of transcending transhuman
limitations.

Due to the use of numerous extreme, experimental, 
and dangerous self-modifications, some exhumans 
have done permanent damage to their psyches, 
becoming insane, or perhaps just transferring their 
mode of thinking into something that is no longer 
recognizable as human. Some have also adopted an 
antagonistic view of their former transhuman species, 
viewing it as weak, decadent, and unworthy. This 
has spurred some exhumans to actively attack and 
ravage transhuman settlements and ships, though 
usually in isolated areas.

A few examples of exhumans are described below, 
though gamemasters are encouraged to develop 
their own.

\subsection{Neurodes}

Seeking to achieve a new level of super-intelligence 
and conscience, neurodes have abandoned the typical
transhuman sleeve in exchange for a multipedal
neuronal shell that is both body and brain at the same 
time. The bulk of a neurode's body mass consists of 
amorphic clusters of neuronal and epithelial cells, enclosed
in a hard carapace shell with four legs and two
manipulatory digits. The cerebral mass of neurode 
brains gives them impressive calculation and other 
mental capabilities far exceeding that of a normal 
transhuman. Neurodes typically defend themselves 
with swarms of teleoperated drones.
\textbf{Skills:} Fray 30, Investigation 80, Perception 90, others 

as appropriate
\textbf{Implants:} Access Jacks, Carapace Armor, Circadian 

Regulation, Direction Sense, Eidetic Memory, En-

docrine Control, Hyper Linguist, Math Boost, 

Medichines, Multi-Tasking, Oracle, Skillware
\textbf{Notes:} Mental Disorder trait x 2

\subsection{Predators}

Predators seek to transform themselves into an ultimate
top-of-the-food-chain evolutionary contender.
They pursue new avenues in genetic modification and 
prototype implants, often using controversial methods 
and technologies. The biochemical instabilities resulting 
from these untested modifications and altered metabolisms
however, often negatively impact their emotional
and mental stability. Pushing this even further, some 
predators undergo experimental psychosurgery to 
modify their consciousnesses in order to increase cunning
and ruthlessness, a procedure that often has other
negative side effects. A few predators take their survival-of-the-fittest
ideology to an extreme, modifying their
digestive systems for a cannibalistic diet, and relishing 
in the slaughter and feasting on of transhumans.
\textbf{Skills:} Blades 60, Fray 60, Free Fall 50, Freerunning 80, 

Investigation 50, Perception 60, Unarmed Combat 70
\textbf{Implants:} Adrenal Boost, Carapace Armor (11/11), 

Chameleon Skin, Cyberclaws, Drug Glands, Endo-

crine Control, Enhanced Hearing, Enhanced Smell, 

Enhanced Vision, Grip Pads, Hardened Skeleton, 

Medichines, Muscle Augmentation, Neurachem 

(Rating 2), Oxygen Reserve, Poison Gland, Prehen-

sile Feet, Prehensile Tail, Respirocytes, Temperature 

Tolerance, Toxin Filters, Vacuum Sealing, plus any 

other mods the gamemaster feels appropriate
\textbf{Notes:} Mental Disorder trait x 2

\section{The Exsurgent Virus }

Only very few people (or entities) who survived the 
diaspora from Earth know of the true reasons and 
the catalyst that culminated in the Fall. The alien Exsurgent
virus—as those aware of its existence within
Firewall call it—set in place by the ETI to infect 
emerging seed AIs, is something beyond transhumanity's
understanding; something far more complex
than just a computer virus. Though some strains of 
the Exsurgent virus have been identified and various 
types of infected exsurgents have been encountered, 
it is widely assumed that these are creations of the 
TITANs. Largely defeated and eradicated from off-Earth
transhuman networks thanks to the efforts of
%%% 365
the Prometheans, occasional breakouts of the Exsurgent
virus still occur, primarily due to scavengers or
others becoming infected when messing with old relics 
from the Fall.

\subsubsection{Plethora Of Strains }

The Exsurgent virus is unlike anything that transhumanity
has ever encountered so far. While it bears
similarities with both computer and biological virii in 
regards to infection of hosts and propagation, it is not 
bound by any limits of form or transmission vector. 

The Exsurgent virus is amazingly effective and 
infectious. As an information virus, it is highly intelligent
and adaptive, able to mutate into new forms.
Much like certain virii are able to cross species 
boundaries or change their vector from contact to 
airborne, it is also a self-morphing omnivirus, capable
of altering itself and its transmission vectors
to bypass infection safeguards. Like a retrovirus that 
incorporates genetic information into the genome of 
the target cell to subvert the cell to do its bidding, the 
Exsurgent virus does the same but on a more complex
level. It is also known to rewrite a host's neural
code in a similar manner, in effect restructuring the 
target's mind and personality.

While it began as a digital computer virus—the 
manner in which it infected the TITANs—it has transformed
to be communicable via at least three other
forms: biological nanovirus, nanoplague, and basilisk 
hack. Each is described below, along with rules for 
infection and defense.

\subsection{Biological Nanovirus}

Exploiting the infected TITANs' understanding of 
Terran biology and their access to bio- and nanotechnology
the Exsurgent virus appeared in several biological
forms not long into the Fall. These virulent strains
infected biomorph transhumans and sometimes other 
living creatures as well. The biological nanobots 
spreading this strain act much like other biological 
virii, though they radically modify the victim's biological
and mental states. Some versions invade and
restructure the target's genetic code, transforming 
them into the horrible abominations known as exsurgents
(p. 369). While first-hand reports relate lurid
tales of victims metamorphing into hostile monsters, 
such reports are rare and considered unreliable due to 
the mental state of the witnesses (and any recordings 
that can verify such claims have a strange habit of 
disappearing). Other variants of this strain are known 
only to alter the target's neural code, subverting them 
to the will of the virus (and often, by extension, the 
TITANs) and affecting their mental structure in order 
to give them psi ability.

\subsubsection{Biological Infection}

Biological versions are spread much like other pathogens
People usually become infected by proximity to
another infected entity. Vectors may be dermal touching
someone with bio-nanobots excreted through the
skin), inhalation (breathing exhaled bio-nanobots), 
injection, or oral (p. 317). Exsurgent bio-nanobots can 
live outside of a body for extended periods, however, 
so infection is possible merely by occupying the space 
where an infected victim was hours or even days before.

If a biomorph only has a chance of exposure to the 
virus (e.g., they walk through a room in which they 
might have breathed in exhaled bio-nanobots), have 
them make a MOX x 10 Test (use their Moxie stat, 
not their current Moxie score). Failure means they 
were exposed. In other circumstances, however, exposure
may be automatic, such as extended touching of
or kissing an infected person. 

A biomorph exposed to this infection must make 
a DUR x 2 Test to determine if the infection takes 
hold. Basic bio-mods and nanophages do not offer 
%%% 366
any protection, though toxin filters (p. 305) and 
medichines (p. 308) each give a +30 bonus (though it 
is likely only a matter of time before a mutant Exsurgent
strain learns to bypass them). If the test fails, the
victim is infected. See the strain descriptions (p. 366) 
for specific details.

Within 12 hours of being infected, biomorphs 
become contagious to others. (Note that for the 
Watts-Macleod strain, they only remain contagious 
for 12 hours after that.)

\subsection{Digital Virus}

Digital strains are purely information- or code-based 
versions of the virus. They resemble typical computer 
virii, worms, or trojans, spreading throughout the 
mesh, exploiting holes, mimicking protocols, and 
taking advantage of it like a skilled hacker.

Digital versions of the Exsurgent virus are treated 
as intelligent programs, using the same rules as infomorphs
(p. 264), with the following stats:
\textbf{Skills:} Hardware: Electronics 50, Infosec 70, Interfac-

ing 60, Investigation 50, Perception 60, Program-

ming 50
\textbf{Software:} Exploit, Firewall, Sniffer, Spoof, Track, plus 

any others the gamemaster considers appropriate

\subsubsection{Digital Infection}

As a matter of course, this Exsurgent virus will seek 
to access any new systems it comes into contact with, 
hacking in and copying a version of itself.

\subsubsection{Ai And Infomorph Subversion}

An Exsurgent virus may take a Complex Action to initiate
an ``attack'' against any other intelligent program
(AI, AGI, or infomorph) that is running on the same 
system. If it encounters such programs as they are accessing
a system it is on, it will attempt to hack their
home system where they are running so as to attack 
them directly. 

The attack is handled as an Opposed Test, each rolling
COG + INT. If the Exsurgent virus wins, the target
is infected and will be corrupted by the virus in 10 
Action Turns, minus 1 turn per 10 full points of MoS. 
If the target succeeded but rolled lower than the virus, 
they are aware that they are slowly being taken over. 
This immediately causes them 1d10 points of mental 
stress. An infected program has only one option for 
defending itself before the virus takes over—shutdown 
and reboot. It takes the AI or infomorph 1 full Action 
Turn to shut down. Restarting takes 3 full Action Turns 
(possibly longer if the gamemaster so decides), upon 
which the AI or infomorph must make another Opposed
COG + INT Test against the virus. If this test also
fails, then the virus has already embedded itself in the 
AI or infomorph's code and will continue its infection.

One the infection is complete, the AI/infomorph 
becomes an Exsurgent NPC.

\subsubsection{Cyberbrain Hacking}

Exsurgent virii that manage to infiltrate the cyber-brains
of pods and synthmorphs may also target the
digital egos within, using the same rules as given for 
AI and infomorph subversion above. Alternately, the 
virus may conduct a traditional brainhacking attack, 
as noted on p. 261, or unleash a basilisk hack. 

\subsection{Nanoplague}

While the abundance of nanotechnology has been a 
blessing for transhumanity's journey to the stars, it has 
also been a curse. Via the TITANs and mesh-connected 
nanofabrication machines, the Exsurgent virus manufactured
nanobot swarms equipped with variants of
the virus. These nanobot plagues are capable of targeting
all types of morphs and sometimes other machinery
as well. Unlike the biological nanovirus, which uses
biological mechanisms to rewrite biological/neural 
structures, these nanoplagues physically restructure 
both people and things at the molecular level.

\subsubsection{Nanoplague Infection}

Exsurgent nanoswarms follow all of the rules given 
for nanoswarms on p. 328. Unlike transhuman 
nanoswarms, though, Exsurgent nanoplagues may 
penetrate a biomorph internally, affecting the body 
within as well as without.

Any morph that comes into contact with a nano-plague
is considered infected. The only defenses are
guardian nanobots and nanophages (which work the 
same as guardian nanobots in this situation), though 
these are less effective against Exsurgent nanobots, 
inflicting –2 damage to the swarm each Action Turn. 
Some Exsurgent nanoplagues have developed countermeasures
against such systems, inflicting (1d10 ÷ 2,
round up) damage to such defenses each Action Turn. 
Note that nanoplague-infected characters are generally
not contagious themselves ... usually.

See the strain descriptions (p. 366) for specific infection
details.

\subsection{Basilisk Hacks}

Thanks to the vast databanks of knowledge the TITANs 
had absorbed from transhumanity, the Exsurgent virus 
was able to thoroughly analyze the biology and functioning
of transhuman minds. In a few short months, by
accessing all of the research at their disposal, the Exsurgent
and TITAN minds made several cognitive leaps in
their understanding of transhuman brain functions—
breakthroughs that will take transhumanity decades to 
reach. One of these discoveries was a method of applying
sensory input as a weapon, exploiting weaknesses
in the brain's neuro-cerebral wiring.

Known as ``basilisk hacks,'' these attacks take 
advantage of the way biological transhuman brains 
interpret and process sensory input in the cerebral 
cortex. Just as epileptics are susceptible to visualizations
that strobe at certain frequencies, basilisk hacks
employ special visual and auditory patterns that trigger
glitches in the brain's neuronal wiring to inflict
%%% 367
nausea, vertigo, disorientation, and even seizures, 
often mistaken as a stroke or cerebrovascular incident. 
Some basilisk hacks go farther than simply causing 
the brain to seize up and crash, however, enabling a 
mechanism to rewrite the neural code in victims who 
view or listen to the wrong thing. This unknown reprogramming
mechanism enables the virus to infect
even a biological brain with one of its strains. Similar 
attacks are used against both synthmorphs and pods, 
taking advantage of the methods in which cyberbrains 
mimic biological minds with a virtual brain state, and 
thus also manipulating them via the information encoded
in sensory input.
In a nutshell, basilisk hacks are a way of hacking 
transhuman brains merely by feeding them a specific 
sample of sensory input, usually images or sounds. 
The widespread use of augmented reality makes 
deployment of such hacks an easy manner; the Exsurgent
virus just hacks into the target's ecto or mesh
inserts and engages the sensory feed. More traditional 
methods may also be used, including standard interactive
video, holograms, audio, subsonics, or even VR.
Since so many records of the years surrounding 
the Fall were lost, most people do not know if the 
basilisk hack is anything other than a legend. Various 
official groups know that this technology was, in fact, 
used by the TITANs, but they keep this knowledge to 
themselves, in large part to help reduce the number of 
people attempting to duplicate it. 

\subsubsection{Incapacitating Inputs}

When a character experiences a basilisk hack, they 
must make a COG + INT + SAV Test. If this test 
fails, their brain is susceptible to the hack, and they 
immediately suffer 1d10 mental stress. Additionally, 
one of the following effects applies. The duration for 
each effect listed below is 1 minute plus 1 additional 
minute per 10 full points of MoF. Each effect is also 
numbered 1–10, in case the gamemaster wants to roll 
1d10 and randomize the effects rather than choose:
•  \textbf{(1) Cataplexy: }The victim loses control of their 

body and immediately collapses. For the duration 

their body will be non-responsive but they will 

be aware and capable of mental actions. Mesh 

actions and implant controls are also disabled, 

however.
•  \textbf{(2) Catatonic Stupor:} The character becomes 

immobile and non-responsive. Though conscious, 

they are mentally ``not there''—the basilisk hack 

has effectively crashed their brain functions. They 

will do absolutely nothing for the duration and 

will not respond even if moved or attacked.
•  \textbf{(3) Disorientation:} The character becomes disori-

ented and severely confused. They are incapable 

of making decisions, understanding communica-

tion, understanding what is going on around 

them, or acting in any sort of determined way for 

the duration.
• \textbf{(4–5) Grand Mal Seizures:} The subject immediately
falls to the ground and begins convulsing,
suffering 1d10 damage. They may do nothing 
else for the duration and will suffer an equal 
duration period of confusion and weakness (–30 
to all actions) afterwards.
•  \textbf{(6–7) Hallucinations:} The character immediately 
goes off on a mental trip, leaving them completely 
disconnected from reality and their physical body. 
For the duration, the character should only respond
to the hallucinated reality the gamemaster
describes to them, or else the character should be 
treated as an NPC, run by the gamemaster.
•  \textbf{(8) Impaired Cognition:} The character's mental 
capabilities bottom out, turning them into a 
disabled vegetable. COG, INT, SAV, and WIL all 
drop to 1, and the character should act accordingly
to environmental stimuli.
•  \textbf{(9) Nausea/Vertigo:} The character is overcome 
with head-spinning and vomiting and is effectively
incapacitated for the duration.
•  \textbf{(10) Sleep:} The character passes out for the 
duration and cannot be woken short of medical 
intervention.
In rare cases, a character may be able to ``dodge'' 
a basilisk hack they know is coming, assuming they 
have some sort of warning (such as their buddy falling
prey to it moments before). The character must of
course be aware of what basilisk hacks are to even 
consider this idea. If they immediately attempt to take 
action to block out the sensory input when it strikes—
closing their eyes, plugging their ears, turning off their 
AR, etc.—allow them a REF x 3 Test to see if they do 
so in time. 

\subsubsection{Sensory Reprogramming}

In some cases, the Exsurgent virus can actually reprogram
the target's mind via dedicated sensory input.
This is a trickier affair, however, requiring uninterrupted
programming time. As with incapacitating inputs,
the target character(s) experiencing the basilisk hack 
must make a COG + INT + SAV Test. If this fails, they 
become catatonic and paralyzed for a period of 10 
minutes, minus 1 minute per 10 full points of MoF. At 
the end of this period, they are mentally reprogrammed 
and ``infected'' with one of the strains of the Exsurgent 
virus (see below). For the duration of this period, the 
character is undergoing reprogramming as long as they 
remain exposed to the basilisk hack. If the character is 
somehow cut off through the actions of another party, 
the reprogramming immediately fails. In this case, 
however, the victim still suffers 1d10 mental stress 
+ 1 per minute they were exposed, and they remain 
mentally shaken, suffering a –30 modifier to all actions. 
This modifier reduces at the rate of 10 per minute.

\subsubsection{Ygbm Attacks }

Rather than completely reprogramming a victim, some 
Exsurgent attacks simply intend to plant subconscious 

hey 

is

ty, 

e,

s
%%% 368
commands in the target's mind, similar to posthypnotic
suggestions. Nicknamed ``You gotta believe me''
attacks, YGBMs are a sort of remote digital brainwashing
attempt used to create sleeper terrorists and
unknowing collaborators, often by targeting them via 
the mesh. Unlike the mind manipulation techniques 
of psychosurgery (p. 229), YGBM attacks use shotgun 
techniques to open the mind, utilizing some kind of 
backdoor the Exsurgents discovered in the transhuman
brain, and altering the mind by brute force.

A character experiencing a YGBM basilisk hack 
must make a COG + INT + SAV Test. If this fails, a 
single suggestion is implanted in the character's mind, 
without their knowledge. This subliminal command 
will be triggered at some later point, either at some 
predesignated time or according to certain pre-set 
conditions. Once triggered, the character will carry 
out the action with all of the conviction that it is their 
own idea. The implanted suggestion may be something 
as simple as ``kill the Firewall agent'' to something as 
complex as ``manufacture an explosive device and 
plant it in the cargo hold of any ship heading to Mars, 
set to explode one day after they disembark.''

Since YGBM attacks are not intended to completely 
convert the target, but instead to simply convert them 
into a temporary tool or weapon, implanted commands
are not designed to last long. The duration
the suggestion will last equals 3 days +1 day per 10 
points of MoF on the resistance test. If the command 
has not been triggered by this point, it dissipates, and 
the character is none the wiser.

\subsubsection{Recording Basilisk Hacks}

Enterprising characters may seek to record a basilisk 
hack input for their own uses. While basilisk hacks may 
be recorded like any other sensory input, keep in mind 
that the Exsurgents and TITANs likely take measures 
to keep such tools out of the hands of transhumanity, 
lest they construct some sort of defense. Basilisk hack 
sources may be self-erasing or contain coding or countermeasures
that would hinder recording, such as white
noise to defeat audio recording or lens-blinding flashes 
to defeat video recording. Conversely, basilisk hacks are 
considered extremely dangerous by almost all factions 
of transhumanity and universally feared. An individual 
or group known to possess them is likely to be treated 
much like a terrorist with a suitcase nuke. Though 
Firewall has a standard interest in evaluating and enabling
some sort of defense against basilisk hacks, most
Firewall personnel consider it foolish to handle such 
toys and would rather destroy such recordings outright.

\subsection{Exsurgent Strains}

Four variants of the Exsurgent virus are described 
here—gamemasters are encouraged to develop their 
own to keep players on their toes.

\subsubsection{Haunting Virus}

This strain is the most insidious of the Exsurgent virii. 
Over time, it rewrites the target's personality and 
motivations, slowly but surely subverting and taking 
control of the victim's mind. At first the character is 
unlikely to even be aware of the infection, and as it 
progresses the changes the virus makes to the target 
will at first seem natural to the target, as if some new 
aspect of their personality was simply manifesting 
itself. As the effects grow more pronounced, however, 
the victim becomes aware that they are being methodically
altered but is in most cases unable to act against
it. In the end, they are completely transformed into a 
pawn of the ETI. Their mind is no longer transhuman, 
but \textit{alien.}

The exact rate of progression is up to the gamemaster
though guidelines are provided below. Each victim
is affected differently, so the process may be accelerated 
or slowed down as the gamemaster sees fit.
%%% 369
• \textbf{Stage 1 (initial infection to 3 months):} Upon initial
infection, the character suffers 1d10 mental
stress and gains the Psi trait (p. 147) at Level 1 
(also meaning they pick up the Mental Disorder 
trait, as noted on p. 150). They also gain one free 
psi-chi sleight, chosen randomly or by the gamemaster
If a player character has become infected,
they may still be played as normal (see \textit{Roleplay-}
\textit{ing Exsurgents,} p. 368), and may purchase new 
psi-chi sleights with Rez Points. NPCs acquire 1 
new sleight per 2–4 weeks.
  At this stage, the infection is usually hidden, 
though the character will suffer from occasional 
haunting effects (see below). As each week passes, 
the character's personality should shift a minute 
amount, slowly becoming more callous and conniving
and changing in other ways as well. If
possible, the player should be kept in the dark 
about what is happening, but the gamemaster 
should provide them with roleplaying advice to 
reflect their condition. Likewise, the discovery 
and initial use of psi sleights should be played 
out, providing some interesting roleplaying opportunities
Characters and players who know of
the Exsurgent virus and Watts-Macleod strains 
should not know at this point which strain they 
are infected with—make them sweat.
• \textbf{Stage 2 (3 months to 6 months):} The target suffers
another 1d10 ÷ 2 (round up) mental stress
and acquires the Psi trait at Level 2 (also picking
up another disorder). Player characters may
still be played as normal and may purchase psi-gamma
slights with Rez Points. NPCs acquire 1
new sleight per 2–4 weeks.
  Once three months have passed, the character 
should be aware they are under the influence 
of something, but this awareness likely comes 
too late. Haunting effects (below) should occur 
regularly. At this point a character is likely to 
consider offing themselves and resorting to an 
uninfected backup, seeking help, or actively encouraging
others to interfere. The infection will
actively block and hinder such thoughts and actions
however. To actively overcome this mental
control, the character must succeed in a WIL Test. 
At the gamemaster's discretion, failure may result 
in 1d10 ÷ 2 (round up) mental stress as the character
realizes they are no longer fully in control
of their own thoughts and actions. 
•  \textbf{Stage 3 (6 months+):} The victim suffers another 
1d10 ÷ 2 (round up) mental stress and acquires 
the Psi trait at Level 3 (see below). The character 
is now considered an exsurgent and becomes 
an NPC. It may no longer be played as a player 
character. The victim also gains a permanent 
+5 bonus to COG and WIL and acquires 1 new 
sleight every 1–2 months.
As noted above, characters infected with this strain 
suffer from different haunting effects—changes 
to their personality or mind-state. A few ideas for 
haunting effects are noted here, but gamemasters are 
encouraged to be creative when inventing their own 
to apply:
•  \textbf{Altered Perceptions:} The victim's perceptions are 
changed in disturbing and unusual ways. They 
may see things that aren't there, feel a presence 
behind or watching them, inexplicably smell 
blood, hear voices, suffer synaesthesia, or suddenly
perceive the people around them as nothing
but outlandish, blabbering sacks of meat.
• \textbf{Behavioral Modification:} Treat as behavioral control
or personality editing psychosurgery (p. 231).
This is typically applied to shape the character 
closer to being a pawn of the ETI.
•  \textbf{Dream Manipulation:} The character's dreams 
become lucid, weird, and surreal. They may find 
themselves dreaming of life as an alien on some 
exotic exoplanet, as a robotic probe soaring 
through the vast emptiness of space, or fantasizing
different methods of inflicting mass destruction
and death.
•  \textbf{Emotional Manipulation:} Treat as emotional 
control psychosurgery (p. 231). 
•  \textbf{Inexplicable Urges:} The character will be flushed 
with strange alien urges and may sometimes find 
themselves doing highly unusual things without 
realizing at all they are doing it. These may include
taking devices apart to understand how
they work, testing the limits on programming a 
nanofabricator, cutting a living thing apart to see 
how it is put together biologically, testing weapons
eating things that are only barely edible, promiscuous
and unusual sexual activity, lying just
to see what they can get away with, and so on.

\subsubsection{Mindstealer Virus}

Very similar to the haunting virus, the mindstealer 
strain is much quicker acting. Instead of slowly subverting
the target's mind over the course of months,
the mindstealer virus rapidly recodes the victim's brain 
in a matter of minutes. This infection is much more 
invasive and brute-force, often causing significant 
side effects to the target's mental state as a result. This 
strain is only spread as a digital virus, nanoplague, or 
basilisk hack (not as a biological nanovirus).
Once the victim is infected, it takes the virus a 
number of Action Turns equal to COG + INT + SAV 
to completely take over their mind (20 Action Turns 
= 1 minute). During this time, the target is actively 
aware that their mind is under attack and undergoing
massive changes against their will. This process
is confusing, frightening, and painful, inflicting  a 
–30 modifier to all of the character's actions for the 
duration. Many victims are reduced to whimpering, 
drooling, or convulsing for the duration.
This mental transformation inflicts 2d10 mental 
stress to the target. Once complete, the victim is an 
exsurgent NPC, under the gamemaster's control. 

AV

ns

ly 

o-

s
%%% 370

\subsubsection{Watts-Macleod Virus}

The Watts-Macleod strain is a strangely benevolent 
version of the Exsurgent virus, seeming to imbue its 
victims with psi abilities without any of the other 
transformative elements typical of other strains. 
Perhaps created as an accidental mutation of the Exsurgent
virus, there are many who wonder if the true
detrimental effects of this strain simply have yet to 
reveal themselves.

As noted in the \textit{Mind Hacks} chapter section on Psi 
(p. 220), characters infected with this strain gain the 
Psi trait (p. 147) at either Level 1 or 2. If a character 
is so infected during game play, this trait must be 
purchased with Rez Points (if the character does not 
have any points currently available, they pay out of 
the points they earn until the debt is paid off). All of 
the other side effects of Watts-Macleod infection (p. 
367) also apply.

Though infection with this strain does apply some 
benefits to the character, the gamemaster should make 
sure to play up the creepy and unsettling nature of 
this virus. The character should never be certain that 
they haven't in fact been subtly influenced by the virus 
in ways they can't immediately pinpoint—they should 
always feel like the ax may fall at any moment.

\subsubsection{Xenomorph Virus}

The xenomorph strain transforms the target's body 
in addition to their mind. Over time, the victims 
morph physically transmogrifies into some sort 
of alien life form. It is only spread as a biological 
nanovirus or nanoplague (not as a digital virus or 
basilisk hack). Different variants of this strain produce
different alien forms. It is not known where
these different alien templates originated, meaning 
they may be copies of (once) existing alien species 
or simply neogenetic creatures created from scratch. 
The one trait they have in common is that they are 
universally dangerous. Some speculation in Firewall 
circles suggests that the Exsurgent virus may in fact 
have a ``library'' of creature types to deploy, under 
the assumption that at least some will be more effective
than others for exterminating whatever victim
species they are fielded against. 

This strain follows the same rules as the haunting 
virus (above), but with the following changes. The 
timeframe is typically much quicker, though the gamemaster
may adjust this as they see fit.

\textbf{Stage 1:} The effects from Stage 1 of the haunting 
virus apply. Additionally, the character begins to suffer 
minor physical changes that are definitely unusual but 
are not impeding in any way and are easily hidden 
from others. Example biomorph alterations might be: 
unusual hair or fibrous growth, some skin discoloration
or translucence, severe rashes, dermal thickening
weakened or enhanced sensory organs, strong
body odor, hair loss, teeth gain or loss, vestigial tail or 
other limb growth, minor dietary changes, and so on. 
Synthmorphs might experience minor system glitches, 
malfunctioning or improved components, and spots 
of material stress or transfiguration.  Gamemasters 
are encouraged to be creative. This stage typically 
lasts from initial infection to 1 week for biological 
nanovirus strains, or from infection to just 1 hour for 
nanoplague strains.

\textbf{Stage 2:} As with haunting virus Stage 2, plus the 
character begins to seriously transmogrify in ways 
that are difficult to hide from others, becoming 
more and more monstrous as the stage progresses. 
Example biomorph transformations include: growing
scales or feathers, partial modification of limb
structure, partial new limb growth, vestigial sensory 
organ growth, sensory loss, extension of claws or 
spines, severe dietary changes, etc. Synthmorphs 
might experience radical system and shape alterations
limited or enhanced sensor functions, or even
conversion of their robotic shell to smart materials. 
These physical changes weaken the victim, inflicting 
1d10 physical damage. This stage typically lasts 1 
week for biological nanovirus strains or just 1 hour 
for nanoplague strains.

\textbf{Stage 3:} As with haunting virus Stage 3, a character 
reaching this stage becomes an NPC. Additionally, 
the victim completely undergoes a transformation 
into some sort of creature that is no longer even remotely
human. Example exsurgents of this nature are
detailed on p. 369.

\subsection{Using The Exsurgent Virus}

The frightening thing about the Exsurgent virus is 
its adaptability. It was written by a near omnipotent 
ETI with the intent of corrupting any alien seed AIs 
or similar singularities it encountered, and it is \textit{very}
good at it. This means it has the capability to analyze
understand, and mimic almost any alien digital
protocols and communication methods it comes into 
contact with, no matter how diverse the alien mindset 
that constructed what it encounters. It then has a cunning
ability to circumvent any safeguards and infect
such systems. From there, it rapidly assimilates any 
data it can about the target species/civilization and 
does it best to mutate into other forms that can attack 
this target from other vectors. 

Given its constant morphing nature then, the Exsurgent
virus is likely to continue to mutate in new
and interesting ways. Some of these mutations may 
be effective, many not. This does, however, afford the 
gamemaster an opportunity to invent new variants of 
their own to deploy against unsuspecting characters.

\subsection{Roleplaying Exsurgents}

The primary thing for gamemasters to keep in mind 
when roleplaying entities that have been taken over 
by the Exsurgent virus is that exsurgents are following
an alien agenda. The specific goals and actions of
each exsurgent may differ, but they are generally concerned
with two things: spreading the Exsurgent virus
and destroying anything that isn't affected. In some 
cases, this may mean immediate and enraged hostile 
action against anything non-exsurgent around them. 
%%% 371
In others, the exsurgent approach is more methodical
hatching long-term plots to infiltrate positions of
power and authority, setting the stage for acts of mass 
destruction, and so on. In other words, they may be 
handled both as hostile monsters or as nefarious long-term
opponents that are subverting transhumanity
from within or hatching complicated plots that could 
mean devastation on a planetary scale.

If the gamemaster wishes, exsurgents may also 
pursue other goals, tangential to the ones above. 
These may range from accumulating knowledge and 
expertise on how transhumanity functions as a species
to forcibly uploading mass numbers of minds to
more esoteric goals such as manufacturing a halfnium 
bomb or converting the solar system's mass to computronium
The Exsurgent virus is potent and intelligent,
and while its methods and goals may sometimes be 
opaque to transhumanity, it acts with direction and 
purpose. There may also be occasions, however, likely 
due to the mutating and morphing aspect of the virus 
and the way in which it transforms transhuman 
minds, perhaps not always in the manner intended, 
where the exsurgent goals become strange or simply 
horrific, such as running experiments on transhuman 
responses to extreme conditions or converting an 
entire colony to cannibalism.

\subsubsection{Exsurgent-Infected Pcs}

It is possible for player characters infected with some 
strains of the Exsurgent virus to continue on under 
their own volition, even as the virus slowly consumes 
them. This process is, quite naturally, horrifying in the 
extreme, though there is little they can do about it. 
Despite the best efforts of transhuman science, there 
is very little that can be done to save an infected 
person—the virus is simply too potent and adaptive. 
As a result, standard Firewall policy is to terminate 
the infected with extreme prejudice. Most Firewall 
operatives are going to be aware of this, a fact which 
pushes some of those who become infected to keep 
their status a secret from their comrades.

Both the haunting and xenomorph strains usually 
transform a subject over time, meaning that the 
character may initially not be aware of the infection. 
This is a prime opportunity for the gamemaster to 
mess with the character ruthlessly, starting slowly 
with little haunting effects and building up as the 
infection progresses. The character should slowly 
become aware that they are under the influence of 
something—something \textit{intelligent.} Characters aware 
of the Exsurgent virus and its effects will likely pick 
up on this sooner, but the virus may prevent them 
from doing anything about it. In effect, the character 
becomes a prisoner within their own body, a body 
they now share with a cold and malevolent presence
that is methodically taking them over. Such
characters may respond in a number of ways depending
on their personality, ranging from despair,
withdrawal, and suicidal tendencies to complete 
hysteria or calm acceptance. Most importantly, 
however, their personality should begin to change 
as the virus continues to transform them. Players 
should be encouraged to take on new demeanors 
and motivations, reflecting the alien component 
of their changing personality, with some guidance 
from the gamemaster. This presents some intriguing
roleplaying opportunities that the players will
hopefully embrace. If the gamemaster feels that the 
player is not adequately representing the changing 
mindset, however, the transformation can simply 
be accelerated and the character converted into a 
gamemaster-operated NPC. 

\subsection{Exsurgents }

A few examples of exsurgents created from transhumans
transformed by the xenomorph strain of
the virus are noted below. As always, gamemasters 
are encouraged to develop their own, using these as 
guidelines. Unless otherwise noted, exsurgents use the 
stats and skills of the transformed character. Each exsurgent
detailed below first lists the aptitude modifiers
applied to transformed characters, then gives example 
aptitude/skill ratings for NPC exsurgents.
Note that simply encountering transformed exsurgents
is stressful to the minds of many transhumans.
At the gamemaster's discretion, such encounters may 
inflict 1d10 + 3 mental stress (p. 215).

\subsubsection{Creepers (Synthmorphs)}

Perhaps the most disturbing exsurgent variant, so-called
creepers are cloud-like amorphous swarms of
small, black bubbles that are strangely fuzzily defined, 
as if surrounded by some sort of visual refraction effect. 
These clouds are theorized to in fact be autonomous 
femtobot swarms—similar to nanobots, but affecting 
matter on an even smaller scale, at the level of an atomic 
nucleus. These black bubbles are capable of coalescing 
into physical shapes in various states and can penetrate 
just about any material or substance in a matter of 
Action Turns. They may even penetrate morphs, accessing
and interfacing with neural and electronic systems
directly. For rules purposes, treat creepers the same as a 
self-replicating nanoswarm (p. 383).
\textbf{Mobility System: }Walker/Microlight (4/16) (may 
create other mobility systems with different rates)
\textbf{Skills:} Fray 40, Free Fall 50, Intimidation 60, Perception
50, Unarmed Combat (Grapple) 50 (60)
\textbf{Notes: }360-degree Vision, Chemical Sniffer, Electrical 
Sense, Enhanced Hearing, Enhanced Vision, Fractal
Digits, Nanoscopic Vision, Radar, Radiation
Sense, Swarm Composition (but may make SOM 
Tests, and plasma weapons do only 1d10 damage), 
T-ray Emitter

may 

)

p-

l 


372

\subsubsection{Jellies (Biomorph)}

These exsurgents resemble collections of massive, slimy, 
mucus-filled bubbles. Their soft, amorphous shape 
allows jellies to squeeze, slide, and slither through 
even tiny spaces. Jellies are equipped with a number 
of ``limbs'' that resemble long meaty tongues studded 
with hard fleshy spikes that provide excellent gripping 
ability. The lubricating coating that envelopes jellies is 
both toxic and slightly corrosive, melting plastics and 
biological materials after a half hour of exposure. This 
substance may also be ``spit'' at targets.
\textbf{ }
Movement Rate: 4/16
\textbf{Skills:} Exotic Ranged Attack (Spit) 40, Free Fall 50, 

Perception 60, Unarmed Combat 40
\textbf{Notes:} Armor (12/12), Enhanced Smell, Spit Attack 

(area effect), Tongue (DV 1d10 + 3, AP 0), Toxin 

(Application: D, O; Onset Time: 1 Action Turn, Du-

ration: 5 Action Turns, Effect: 1d10 ÷ 2 (round up) 

DV per Action Turn ). Due to their physical form, 

jellies suffer the minimum amount of damage from 

standard kinetic weapon and blade attacks. 

\subsubsection{Shifters (Synthmorph)}

Shifters are synthmorphs whose material frames have 
been converted to an exotic smart matter liquid metal. 
This shapeshifting material can stabilize as a hardened 
metallic shell or liquefy and reshape itself into other 
forms. This allows the shifter to reflow its shell in a 
matter of seconds, enabling it to visually mimic other 
forms, including biomorphs (though they are easily 
detectable as synthmorphs at other wavelengths or by 
touch). Shifters may also reshape parts of their shell 
into melee weapons such as knives or clubs. 
\textbf{Mobility System: }Walker (4/20)
\textbf{Skills:} Blades 60, Deception 55, Disguise 60, Fray 50, 

Freerunning 55, Impersonation 60, Perception 50, 

Unarmed Combat 50 
\textbf{Notes: }Armor (13/13), Enhanced Hearing, Enhanced 

Vision, Shape-Adjusting (Programmable Liquid 

Metal Form)

\subsubsection{Snappers (Synthmorphs)}

Snapper exsurgents are typically crafted from vehicles 
or other large synthetic shells or by melding multiple 
synthmorphs together. They take the form of an insectoid
multi-segmented hexagonal tube with multiple
sets of limbs, three apiece, set radially 120 degrees 
around the torso. These limbs are heavy, double-jointed
and articulated with three joints. Each limb
ends in either a triad of manipulatory digits or a larger 
pincer-like claw.
\textbf{Mobility System:} Walker (4/24)
\textbf{Skills:} Climbing 45, Fray 40, Freerunning 40, Percep-

tion 40, Unarmed Combat (Pincers) 55 (65)
\textbf{Notes: }360-degree Vision, Armor (16/16), Enhanced 

Vision, Extra Limbs (9, 12, or 15 total), Lidar, 

Magnetic System, Pincers (DV 2d10 + 3, AP –3), 

Structural Enhancement 

\subsubsection{Whippers (Biomorph)}

These small barrel-shaped creatures have a mass 
of small legs under their trunk that allows for fast 
movement. At the top of their trunk is another mass 
of 3-meter long, strong, whip-like tentacles. Some 
of these tentacles feature gripping surfaces for grabbing
and holding (both for tool use and mobility),
while others are sharp-edged and useful for slicing 
through opponents. 
\textbf{Movement Rate:} 8/40
\textbf{Skills:} Climbing 40, Fray 50, Free Fall 40, Freerunning 50, 

Infiltration 40, Perception 50, Unarmed Combat 

(Tentacles) 45 (55)
\textbf{Notes: }Enhanced Vision, Tentacle Whip (DV 2d10 + 1, 

AP –1)

\subsubsection{Wrappers (Biomorph)}

These exsurgents resemble large, thin, four-armed, 
spiny starfish, capable of walking in a quadruped 
manner, though they are seemingly better adapted for 
microgravity. A large circular mouth resides in their 
middle on one side and each arm ends in small sharp-clawed
digits, useful for climbing and tool use. Small
vent sacs allow for thrusting in microgravity and 
sensory bands on the upper part of each arm provide 
low-frequency hearing and infrared-equivalent sensing
Their name comes from their tendency to drop on
opponents from above, wrapping themselves around 
the head and arms.
\textbf{Movement Rate:} 4/16
\textbf{Skills:} Fray 40, Free Fall 50, Perception 50, Unarmed 

Combat (Grapple) 50 (60) 
\textbf{Notes: }Armor (8/8), Bite (DV 2d10 + 3, AP –5, must 

grapple first), Chameleon Skin, Claws (DV 1d10 

+ 2, AP –2), Enhanced Hearing, Infrared Sensing, 

Vacuum Sealing
%%% 373

\subsection{Exsurgent Psi }

In addition to psi-chi and psi-gamma (see \textit{Psi,} p. 220), 
exsurgents have access to a third level of psi ability 
(the Psi trait at Level 3), known as psi-epsilon. Psi-epsilon
is theorized to allow a level of interaction with
the underlying physics of reality that is beyond the 
comprehension of transhuman science. Though some 
Firewall scientists have speculated about the manipulation
of dark energy or the Higgs field and Higgs
boson particles and similar exotic ideas, the truth is 
that psi epsilon represents an understanding of science 
so far advanced and so alien that transhumanity can 
only guess at its mechanics.

\subsubsection{Exsurgent Synthmorphs And Psi}

Exsurgents in synthetic morphs that were transformed 
via nanoplague may use psi, despite lacking a biological
brain. Through some unknown mechanism, the
infecting nanobots are able to simulate a biological 
brain's effects. This feature, however, also makes them 
vulnerable to psi use by others.

\subsubsection{Exsurgent Psi Strain}

Exsurgents with Level 3 psi (psi-gamma) do not suffer 
strain when using psi. Instead, they draw requisite 
energy from the environment around them. In game 
terms, this means that gamemasters do not need to 
worry about rolling strain for exsurgent sleights. On a 
cinematic level, it also allows the gamemaster to add creative
environmental details to exsurgent psi use: sucking
the warmth out of the air, killing the lights, withering 
plants, draining power from nearby electronics, killing 
small creatures or insects, lowering air pressure, etc. 

\subsection{Exsurgent Psi-Gamma Sleights }

These sleights are available to exsurgents with the 
Level 2 Psi trait.

\subsubsection{Decerebration }


  

Active

   Complex

 Touch

  

   +2

  
This sleight temporarily ``shorts out'' a portion of the 
subject's brain stem. The victim's cerebral functions 
and motor activity become severely impaired; apply 
a –30 modifier to all actions. If an Excellent Success
is scored, the target effectively loses all cerebral
functioning, including vision, hearing, other sensory 
functions, and mesh use. Their muscles and limbs also 
tense and become rigid, essentially paralyzing them in 
what looks like an agonized state.

\subsubsection{Onslaught}


  

Active

   Complex

 Touch

  

   +0

  
This offensive sleight floods the target's mind with 
sensory input and thought processes that are so alien 
and disturbing that they inflict 1d10 + (WIL ÷ 10, 
round up) mental stress. Increase the stress by +5 if an 
Excellent Success is scored.

\subsubsection{Scenario }


  

Active

   Complex

 Touch

   Sustained

   +2

  

Control
This sleight hijacks the target's sensorium, replacing 
it with a virtual scenario controlled by the exsurgent. 
The effect is much like being jacked into a simulspace 
scenario, albeit against the target's will. While the 
exsurgent cannot harm the target in the scenario, they 
can learn something about the person's behavioral 
responses to certain situations. While under the influence
of this sleight, the target is cut off from their
physical senses (–60 to any Perception Tests), but they 
may flail about and otherwise respond physically to 
events in the scenario, which may cause them to hurt 
themselves and will make them seem crazy to onlookers
Targets may attempt to ignore the scenario and
concentrate on the real world, but this requires a WIL 
Test each Action Turn and they suffer a –30 modifier 
from disorientation even if they succeed. 

\subsubsection{Strip Memory/Skill }


  

Active

   Complex

 Touch

  

  

+2

  

Psi Assault
Strip allows the exsurgent to suppress certain memories
in the target's mind. This can be used to block
memories of certain events or even the victim's identity
The process is not exact, however, and so the
memories may not be fully suppressed and/or related 
memories may also be blocked; the gamemaster decides
on the effect as determined by the MoS. Strip can
also be used to temporarily erase a specific skill from 
the target's mind, preventing them from using or even 
defaulting to that ability while so affected.

\subsection{Exsurgent Psi-Epsilon Sleights }

Psi-epsilon is available to exsurgents with the Psi trait 
at Level 3. This subset of psi involves abilities that can 
affect the underlying physical nature of the universe, 
creating localized reality-altering effects. Psi manipulation
on this level is extremely dangerous and should
have the potential of disastrous consequences, given 
that these manipulations violate fundamental laws of 
nature and sometimes create paradoxes between the 
forces that glue the universe together. Gamemasters 
are also encouraged to treat critical failures as appropriately
\textit{critical.}
Given these factors, psi-epsilon should only be 
accessible to powerful adversaries and used as a 
gamemaster tool with extreme precaution. The exact 
mechanics of psi-epsilon sleights are left wide-open, 
however, for whatever use the gamemaster can 
Psi Assault
Psi Assault

ven 

of 

he 

s

o
374
dream of. Their intent is to be more cinematic than 
mechanical, so gamemasters should wing rules effects 
as needed. This is an open opportunity for the gamemaster
to create nightmarish monsters from beyond
with frightening reality-ripping and mind-scarring 
abilities. While some example sleights are provided 
below, gamemasters are encouraged to modify their 
effects and to create their own. 

At the gamemaster's discretion, simply observing psi 
epsilon sleights in action may inflict 1d10 + 2 mental 
stress on a character (p. 215).

\subsubsection{Anti-Electronics Field}

All electronics within Close range of the exsurgent 
mysteriously fail as if electrical power is simply negated
This effectively disables synthmorphs and pods
and leaves other characters without access to their 
devices or implants.

\subsubsection{Casimir Force Repulsion}

This sleight exploits the Casimir effect (an interaction 
between the electromagnetic fields of different objects) 
on a macro-scale, allowing the exsurgent to levitate 
themself or other objects by creating repulsing fields. 
This could also allow the exsurgent to push targets 
away, pin them against walls, etc. 

\subsubsection{Cryokinesis}

This sleight allows the exsurgent to drain all heat from 
an area, down to absolute zero, effectively freezing 
everything within range and inflicting cold damage on 
unprotected characters.

\subsubsection{Diffusion}

This sleight diffuses light, laser, and particle beams, effectively
making them useless as weapons, or at least
impairing the DV they inflict.

\subsubsection{Kinetic Friction}

The exsurgent uses this sleight to increase the friction
applied to kinetic activities. This has a negligible
effect on most activities, but high-velocity projectiles 
like firearms and railguns will be significantly slowed, 
decreasing their DV by half or more.

\subsubsection{Matter Transformation}

This sleight alters the molecular bonds and atomic 
components of a targeted material, causing it to either 
weaken and deteriorate or transmutate into some 
other physical substance. This can also be used to 
alter the molecular state of a material, causing gases 
to condense, solids to liquefy, etc. An exsurgent could 
use this to weaken a door or other barrier, condense a 
solid bridge out of liquid, petrify organic materials, etc.

\subsubsection{Negative Refraction}

The exsurgent redirects electromagnetic waves with 
this sleight, refracting them around their body, with 
the same effect as the invisibility cloak (p. 316).

\subsubsection{Pyrokinesis}

Similar to cryokinesis, this sleight enables the exsurgent
to accelerate the molecules, increase friction, or
focus heat in a specific area, causing materials to ignite 
or smolder.

\section{The Factors }

The alien species known as the Factors are unlike 
anything mankind has encountered so far (see \textit{First }
\textit{Contact: The Factors,} p. 40). Though they are aloof 
and stand-offish, their willingness and sometimes eagerness
to deal with (parts of) transhumanity indicate
either a keen interest on their part in transhuman 
affairs or some hidden ulterior agenda. Though the 
various transhuman factions have been similarly wary 
and cautious, and despite numerous communications 
difficulties and failures, an uneasy relationship has 
flowered over the past 8 years, facilitating some trade 
and exchange of knowledge.

\subsection{Origin And Evolution }

The Factors have remained notoriously tight-lipped 
about their origins, history, and the location of their 
homeworld. Though they have also paid visits to some 
of transhumanity's exoplanet colonies, no gatecrashing 
expeditions have yet found any sign of Factor habitation 
or passing elsewhere in the galaxy. Repeated inquiries 
by transhuman mediators have been simply ignored or 
answered in cryptic terms that have yet to be deciphered.

The Factor home world is in fact an Earth-like 
planet with comparable atmospheric conditions and 
a prevalent hydrosphere but with longer periods of 
darkness (due to slower rotation of the planet and a 
less-luminous orange giant). While adapted transhumans
could find their planet habitable, their abiogenesis
(the formation of life from self-replicating, but
not-living molecules) took a different route than life 
on Earth. 

The Factors' primordial ancestors began in their 
planet's early geological history as a type of of photosynthesizer
that ate carbon dioxide and water and
released oxygen, also obtaining energy from inorganic 
chemicals like hydrogen sulfide. Long conditions 
without direct light on their homeworld, however, 
spurred the success of organisms that could survive 
by acquiring energy in other ways. The next evolutionary
leap was to a stage similar to Terran slime
molds, eating microorganisms from decaying matter. 
As evolution progressed, they mutated further into a 
cautious, predatory species that fed on larger, dangerous
creatures. Rather than actively hunting such prey,
this species developed versatile methods of capturing 
and immobilizing their competitors (comparable to 
Earth's funnel web or trapdoor spiders). Over time, 
this method of trapping prey spurred basic (practical) 
intelligence and provided them with the evolutionary 
advantage that paved the way to sapience, driving 
Factors to become the highest developed organisms 
on their planet and build a civilization. 
%%% 375

Like mankind, the Factors suffered through and 
survived their own singularity event and encounter 
with the Exsurgent virus. Perhaps due to their cautious
and calculating nature—and their evolutionary
experience in dealing with more powerful and dangerous
opponents—the Factors are resolutely determined
not to make any similar mistakes as a species. 

\subsection{Xenobiology }

Since life on the Factors' home world developed 
differently than Earth and produced neither nucleic 
acids nor amino acids, Factor metabolic processes 
and ``genetics'' are very different from transhumanity's
While little is known about the exact physiology
of the Factors, due to the lack of captured or dead 
specimens to investigate (so far, no hypercorps or factions
have risked an interstellar incident by abducting
one to dissect ... so far) and their unwillingness to be 
examined by transhumans, most common knowledge 
about them is based on observational and forensics 
research during their encounters with transhumanity. 

\subsubsection{Individual Factors}

Individual Factors resemble non-translucent ambulatory
amoeba, slime molds, or slugs. Though they
``stand'' only 0.3 meters tall, their body diameter 
ranges from 1.5 to 2 meters, they can be up to 2 
meters long, and they can shape their body to change 
these dimensions. Instead of walking, they crawl or 
ooze from place to place by protruding finger-like 
structures (so called \textit{pseudopodia}) that attach to the 
ground (or wall or ceiling) and which they use to pull 
and retract their rear forward (similar to cell migration
Due to their malleable shape they are not as
strongly affected by gravity as transhumans.

Most Factors that have been encountered are dull 
ocher in color and are made from a gooey, gel-like 
substance of unknown composition, though yellow 
glistening patches (which are temporary organelles) 
and bundles of fibers (some kind of muscular skeleton
often become visible when they move. While
all Factors are able to express versatile pseudopodia 
to manipulate and operate devices (and even attack), 
some subspecies possess, carry, or are able to develop 
additional differentiated limbs, cilia, or organs with 
specialized functions.

\subsubsection{Factor Colonies}

Unlike transhumans, Factors rarely act individually—
in fact, individuality is a concept somewhat foreign 
to Factors. Most Factors join together into a collective
unit termed a \textit{colony.} A typical Factor colony is
composed of hundreds or thousands of individual 
Factors that literally physically join together into a 
mass organism (resembling more a primordial soup 
than a gargantuan Factor). Individual Factors are 
indistinguishable from each other when merged into 
the supra-structure of the colony, though individuals 
can form and break apart to accomplish different 
tasks. This colonial merging is mainly possible due 
to the fact that Factors don't possess differentiated 
and specialized organs or cell types that need to be 
segregated from each other, but instead use an open 
system of local, temporary gradients for regulation. 
%%% 376
Neurofilament connections effectively allow the 
Factor colony to operate with a group mind-state, 
with supercomputer potential. This also allows for the 
easy transfer of knowledge and memories to all other 
factors within a colony.

If dismembered, blown apart, or otherwise separated
individual Factors in a colony can regenerate
and reconstitute at a rapid rate without loss of ability 
or memory. 

Factors reproduce when different members of the 
colony produce gametes that fuse, grow into spore 
stalks, and emit spores that later hatch and grow clones.

\subsubsection{Biodiversity And Self-Design }

Factors colonies are known for their high biodiversity, 
featuring numerous sub-groups (so called \textit{phenotypes}) 
that each have unique traits (cilia, apocrine glands, carapace-like
outer membrane) that give them an ecological
advantage or a utilitarian aptitude for certain tasks.
These traits are not random evolutionary features, but 
are the result of intentional bio-engineering. The Factors
have a strong grip on their own metabolisms and
genetic expressions and can draw on an array of genetic
building blocks and biotech techniques to modify
themselves rapidly and massively to adapt to special 
conditions. Whether these modifications might have a 
purpose beyond function, such as for reproduction or 
self-expression, is currently unknown. 

\subsubsection{Metabolism }

Factors ships and habitats have transhuman-friendly 
atmospheres with a slightly higher content of carbon 
dioxide and less nitrogen that mimics the conditions 
on the Factors' home planet. They don't breathe 
oxygen via lungs but absorb it via their outer ``skin.'' 
Since they can also use oxygen from other sources 
(minerals, liquids like water, and salts) to fuel their 
aerobic energy production (i.e., respiration), they can 
be considered functional anaerobes, meaning they can 
survive in environments without atmosphere, though 
they must usually supply themselves with food in 
order to do so.

During the few ceremonial festivities to which 
Factors were invited and actually attended, they 
consumed and processed transhuman organic food 
by internalization. On the first occasion, dishes and 
dinnerware were absorbed as well due to misunderstanding
but were excreted unharmed after the organic
components the factors could utilize had been
broken down. 

While Factors are omnivores similar to transhumans, 
they prefer immobilized live prey, which they enjoy 
absorbing internally and digesting, excreting those 
parts that cannot be used to fuel their metabolism. 
As such they can devour biomorphs and non-metallic 
components of synthmorphs.

\subsubsection{Perception }

Factors don't perceive the world as transhumans 
do. They (usually) don't possess visual or acoustic 
organs to see or hear but have a number of sensory 
organs that grant them a 360-degree awareness of 
their surroundings and enable them to interact 
with their environment similar to or in some cases 
even better than transhumans do. Their perception 
spectra includes the infrared part of the electromagnetic
spectrum, magnetoception, a high resolution
chemical-gradient based ``sight,'' and keen haptic 
perceptions (including vibrations). 

\subsubsection{Communication }

Due to the lack of a vocal system, Factors use different 
methods of signaling and communication. Factors in 
physical contact exchange information by juxtacrine 
cellular, neurofilament interfacing, or by merging for 
information transfer. Over distance, Factors signal via 
pheromonal communication using airborne scents or 
chemical signals with different metabolic components. 
Nicknamed ``Factor dust,'' this communication is effective
even over great distances (up to 10 kilometers).
Factor dust does have an odor perceptible to transhumans
however, that ranges from smelly to unbearable.
This dust is also toxic in high concentrations and sometimes
used as an offensive or defensive mechanism.

To date, transhumans have failed to develop a device 
that can analyze the Factors' chemical effluvia and 
translate it into something understandable, due to the 
lack of a conceptual matrix (though certain ``moods'' 
have been identified). Instead, all communication 
between the Factors and transhumanity is mediated 
through computer interfaces. Certain Factor phenotypes
that deal with transhumanity have grown a a
neurobiological interface (or organ) that enables them 
to wirelessly mesh with transhuman computer systems.

Long distance communication between Factors 
and transhumanity is achieved by normal means of 
farcasting communication. There are strong indications
that Factors also take advantage of quantum-entanglement
communications as well, enabling
Factor colonies and ships to share knowledge gained 
in different parts of the galaxy. 

\subsection{Exosociology}

Factors are cooperative beings that exist as a collective
colonial organization. Though they can operate
individually from the colony, they tend to view
themselves as part of that collective entity rather than 
an individual being. Multiple colonies often work together
as a higher functional unit (a \textit{lattice}), like some
kind of superorganism. These lattices enable the potential
for collective networking and bioinformation
exchange on a scope beyond anything transhumanity 
is capable of. 

These colonies should not be considered the same 
as the hive mind social hierarchies of Terran insects. 
Factor colonies do not feature the same division of 
labor and instead function according to a consensus-based
sort of groupthink. Individual Factors have no
sense of personal gain or property and share equally 
with other Factors and colonies. 
%%% 377

Factors do not experience emotions in the same 
manner that transhumans do, though being evolved 
creatures they are driven by certain instincts. They 
know and understand many of the same concepts that 
transhumans do thanks to evolution, such as competition/rivalry
and altruism/cooperation. They also enjoy
an understanding of basic ideas of philosophy such as 
aesthetics and metaphysics, though their conception of 
such topics is likely to differ from transhuman notions.

\subsubsection{Art And Culture }

Due to their perceptual array, Factor ``art'' (creations 
and expressions that are appealing or attractive to 
their senses) is mostly chemical or tactile-based. It 
can induce certain ``mood'' responses from individual 
Factors and whole colonies, ranging from agitated 
jittering and release of a Factor dust interpretable as 
``joy'' to a tensing and solidifying of the whole body 
(and no chemical expulsion) that seems to relate to 
anger. Since they like and are susceptible to delicate 
compositions of different chemicals, certain bouquets 
and fragrances from liquids or volatiles such as wines 
and perfumes are both appealing and repulsive to 
Factors. The same is also true for the natural smells 
of biomorphs, meaning that Factors may respond in 
a more friendly or hostile manner depending on a 
particular transhuman's scent.

Factors do not comprehend most transhuman art, 
as it is mostly visual or auditory based (e.g., music, 
painting), though they do seem to have an appreciation
for engineering, architecture, and some sculpture.
While they have expressed interest in digitalized media 
out of a curiosity (or plan) to understand transhuman 
mindsets, they lack the organs and mental structure to 
access and comprehend it. 

\subsubsection{Technology}

Though the Factors repeatedly express dismay at 
transhumanity's low level of technology, they have 
failed so far to produce technology that is exceptionally
far in advance. Some believe that the Factors are
simply hiding their advanced technology in order to 
keep transhumans from stealing or copying it, while 
others believe this may simply be a posture taken 
by the Factors to facilitate bargaining. The Factors 
also claim that their technology would not interest 
transhumans because of their differences in physiology
and mindset, and what little technology they
have displayed is certainly specialized for Factor use 
(specialized neurofilament links, chemical signaling 
and Factor dust interfaces, etc.) and so unusable to 
most transhumans. The Factors have traded some 
technology to transhumans, at expensive cost, though 
the small sampling provided so far seems to have 
originated from alien species with physiologies more 
akin to transhumans.

It is interesting to note that scans of Factor ships 
indicate their technology level, aside from the drives, 
is not all that more advanced than transhumanity. 
Also of note is that no two Factor ships have been 
alike, spurring some to believe that the Factors are 
in fact making use of ships acquired from other alien 
species—perhaps abandoned derelicts that the Factors 
recovered and restored. Once again this has led some 
to believe that the Factors are using what to them are 
primitive craft in order to hide their real technology, 
while others are of the opinions that the Factors are 
simply scavengers and opportunists, piggybacking on 
the developments of other alien species.
One interesting feature of Factor technology is that 
they use no artificial intelligences. This stems from 
their own singularity experience. Instead, Factors use 
infomorph versions of themselves or the accumulated 
processing power of their colony mind-states to perform
major computerized tasks.

\subsection{Factor Motivations}

The driving reason behind why the Factors made contact
with transhumanity remains unclear and is open to
gamemaster interpretation. There is much speculation 
among transhuman factions. Some think the Factors 
are simply social creatures who are glad to make contact
with another post-singularity surviving civilization.
Others believe the Factors are mercenary traders who 
somehow acquired FTL travel and use it to their full 
advantage, fleecing various trading partners who lack 
such capabilities (thus also explaining why the Factors 
eschew the Pandora Gates—they disdain competition). 
Still others worry about secret, hidden motivations.
Despite claiming to represent a number of alien 
civilizations, the Factors have been extremely reluctant
to provide any other information on these other
species or even to say how many there are. More 
recently, however, they have expressed a willingness 
to transport a small number of transhumans to other 
civilizations, though at great expense and with no 
guarantee to their safety or ability to return.
So far, the Factors have made no mention of the 
ETI or the Exsurgent virus to transhumanity, though 
they are aware of their existence. Instead they have 
simply issued dire warning and admonitions regarding 
the development of seed AIs and use of the Pandora 
Gates. The Factors have in fact expressed an extreme 
reluctance to deal with any transhuman factions that 
are heavily invested in gatecrashing, such as Gatekeeper
Corp.

\subsection{The Factors In Game }

Factors should be rarely encountered in \textit{Eclipse Phase.}
Most of their interactions with transhumanity occur 
remotely and infrequently. It is uncommon for them to 
risk direct interactions. It should be kept in mind that 
Factors are cautious to the point of being conservative
and view transhumanity as potentially hostile or
dangerous, so they are more likely to act with discretion
than boldness. Factors are also quite cunning,
having evolved from prey-capturing predators, and 
still design complex machinations (traps in the metaphorical
sense) to achieve their goals. In other words,
Factors out to achieve something are likely to hatch an 

\textit{ase.}

ur 

to 

at

a
378
elaborate plot to get it and are not against recruiting 
transhumans. Also, drawing on their abilities to self-modify
themselves and technology developed on their
own or picked up at other places in the universe, they 
can adapt to new situations very quickly. 

\subsubsection{Alien Mindset}

Factors don't possess Lucidity stats and cannot be 
driven to madness like transhumans.

Affecting Factors with psi is very difficult, as noted 
on p. 222. As of yet, Factors have not exhibited any 
psi abilities of their own.

\subsubsection{Factor Combat }

Factors usually avoid direct combat but can defend 
themselves if they have to. They are only likely to act 
aggressively in situations where they have surprise, 
environmental or technological advantages, and/or 
superior numbers. Due to their cooperativism, Factors
are rarely encountered alone, working en masse
to eliminate potential threats.

\textbf{Immunity to Kinetic Damage: }Due to their gooey 
composition and non-differentiated physiology, kinetic
weapons (firearms, railguns) are not very damaging
to Factors. Most such projectiles pass through their 
gelatinous bodies, inflicting minor damage via hydrostatic
shock. The holes left by such weapons quickly
close in a matter of seconds. Likewise, cuts left by 
blades rapidly seal. In game terms, both such weapons 
inflict the minimum amount of damage possible. 

\textbf{Regeneration: }Even if damaged, Factors regenerate 
very quickly. They heal SOM ÷ 10 (round up) damage 
every Action Turn. Wounds may not be healed this 
way, however. 

\subsubsection{Factor Computers}

Due to using completely alien protocols and system 
designs, Factor computers are essentially impossible 
to hack. They do, however, employ some devices that 
emulate transhuman computer systems for communication
purposes, and these may be hacked as normal.

\subsubsection{Factor Dust Toxin}

As noted above, Factors can deploy a type of chemical 
Factor dust that is is toxic to transhumans. Treat this 
as an area effect (cone) attack. 
\textbf{Type:} Bio 
\textbf{Application:} Inh
\textbf{Onset Time:} 1 Action Turn
\textbf{Duration:} 10 minutes (5 with medichines)
\textbf{Effect:} Severe coughing and respiratory distress, 1d10 

damage per Action Turn for 5 Action Turns (or on-

going with continuous exposure), –20 to all actions 

for 2 hours. Medichines reduce damage by half and 

modifier duration to 15 minutes. 

\subsubsection{Melding}

Individual Factors may merge together to form larger 
units, much like masses of Factors form colonies. In 
game terms, use the highest stat possessed by the 
melded Factors, +2 for each additional Factor up to a 
maximum of +10. Durability (and Wound Thresholds) 
are added together.

\subsection{Factor Phenotypes }

A few examples of the different Factor phenotypes are 
described below.

\subsubsection{Ambassadors}

The ambassador Factor phenotypes are the ones who 
most commonly handle direct interactions with transhumanity
Most likely to put transhumans at ease,
these Factors feature a section of sensor nodules that 
loosely approximate a ``face.''
\textbf{Movement Rate: }4/16\textbf{ }
\textbf{Skills:} Deception 70, Exotic Ranged Attack: Factor 

Dust 45, Fray 25, Free Fall 40, Hardware: Electron-

ics 35, Infosec 35, Intimidation 50, Kinesics 40, 

Perception 50, Persuasion 60, Protocol 50, Research 

35, Unarmed Combat 30
\textbf{Notes:  }Access Jacks, Chameleon Skin, Grip Pads, 

Infrared Sensing, Magnetoception, Poison Gland 

(Factor Dust Toxin)
%%% 379

\subsubsection{Guardians}

Guardian Factors serve as bodyguards for ambassadors 
or other Factors whenever they leave a Factor ship.
\textbf{Movement Rate: }4/20\textbf{ }
\textbf{Skills:} Climbing 40, Exotic Ranged Attack: Factor 

Dust 65, Fray 50, Free Fall 40, Freerunning 40, Infil-

tration 40, Intimidation 50, Kinesics 20, Perception 

50, Profession: Security Procedures 50, Unarmed 

Combat (Tentacles) 50 (60)
\textbf{Notes:} Chameleon Skin, Eelware, Electrical Sense, 

Grip Pads, Infrared Sensing, Magnetoception, 

Poison Gland (Factor Dust Toxin), Tentacle Whip 

(DV 2d10 + 1, AP –1)

\section{The Iktomi}

Little is known about the alien race known as the 
Iktomi except for the ancient ruins they left behind 
on Echo V (p. 109). No Iktomi specimens have been 
found so far, though certain architectural remains 
suggest a predilection for web-like structures. This 
has been bolstered by certain other features and relics 
which suggest these aliens had a segmented, multi-legged
arthropod-type form—thus their given name,
after a Native American spider god.

What is clear is that the Iktomi suffered through 
some sort of cataclysmic event that wiped out their 
civilization. The nature of this event has yet to be determined
but it raises concerns for many researchers.
Having suffered through its own near-apocalypse, it 
is not comforting for transhumanity to find evidence 
that other alien species did not.

Though the Iktomi are likely long extinct, the remnants
of their civilizations presents a plot hook for
gamemasters to use for building scenarios. Perhaps 
evidence is uncovered of Iktomi settlements in other 
star systems, and the characters are sent to investigate 
or a relic is unearthed that suggests the Iktomi fell 
prey to some danger that now threaten transhumanity. 

\section{The Pandora Gates}

The five known Pandora Gates (see \textit{Opening Pandora's }
\textit{Gate,} p. 46) all look and operate in a similar fashion, 
though they vary wildly in terms of size, shape, and 
available destinations. The gates are built from some 
sort of stable exotic matter whose full atomic structure
scientists haven't come close to cracking. To touch
and sight, however, the gates appear to be constructed 
from a timeless-seeming polished black metal with no 
signs of aging or wear and tear. Something about the 
gates' physical composition makes them difficult to 
look at, as if the viewer cannot quite focus on their 
outlines. Some onlookers have reported feelings of 
vertigo and nausea, while others have insisted that the 
gate outlines move on the edges of their visions, as if 
the lines are reflowing or the edges are vibrating at 
high frequencies. Due to this disturbing feature, most 
gate sites keep the actual gate structures covered.

Structurally, the gates themselves are partially 
enclosed by an irregular spherical cage composed of 
black arms that are bent and angled in unusual ways 
and sometimes interlocking. When new wormhole 
location is programmed into the gate, these arms 
physically change shape, move, and reflow  around 
the spherical gate area (suggesting they are made of 
some sort of programmable matter). The openings 
between arms are sometimes only large enough for 
a transhuman to enter, while others are large enough 
to allow a freight train of supplies to pass through. 
In many cases, large vehicles or equipment must be 
dismantled, carried through, and reassembled on the 
%%% 380
other side. It is suspected that the gate size could be 
programmable, but so far efforts to do so have failed.

All known gates within the solar system are located 
on the surface of naturally occurring astronomical 
bodies, be that a planet, moon, asteroid, or so on. 
None have yet been found without such a land-based 
connection (e.g., floating in space or in the upper 
atmosphere of a gas giant), though such gates have 
been found in other star systems. It is speculated that 
gates could be physically moved, but no one is willing
to risk such an endeavor given the lesson learned
when the Go-nin Group messed too heavily with the 
Discord Gate's controls (see \textit{Eris, }p. 109).

The arms comprising each gate's spherical cage have 
an abnormal-looking organic-seeming growth on their 
exterior surface in some areas, patterned in entrancing
twists, curves, and whorls that in fact adhere to
perfect mathematical formulas. It took some time for 
scientists to discover that this growth was in fact the 
gate's control systems, or so-called ``black box.'' The 
interface developed to interact with this system is what 
allows gate controllers to manipulate gate functions.

\subsection{The Wormhole}

When the gates themselves are open, a sphere appears 
within the central area that is not so much black as 
pure nothingness. This sphere of darkness projects an 
aura of charged energy, and in fact ripples of green 
arc lightning cascade across its surface. Anyone or 
anything entering that sphere comes out the other side 
of the wormhole, through a similar gate, seemingly 
instantaneously. An unknown force field effect seems 
to prevent the atmospheres from the two connected 
gates from interacting.

Exactly how this wormhole is created is something 
that remains outside of transhumanity's comprehension
The generally accepted theory is that each gate
acts as an anchor, allowing the fabric of space-time 
to be folded so that two such anchored places can 
be brought together, ripping a hole open between 
them so that a person can simply step through. It is 
unclear whether or not these wormholes are all preexisting
created when the gate was first established,
or whether each wormhole is manufactured whenever 
the gate is activated.

Other more radical theories on how the gates function
exist, though these are usually discounted as far
less likely. One such theory suggests that the wormholes
created are actually only zero-width Planck-scale
connections across space-time and that no matter
is actually transferred—only information. Instead, this 
theory suggests that anyone or anything entering the 
wormhole is in fact instantaneously scanned and disassembled
and then their informational blueprint is
transmitted as information across space to the other 
gate, which immediately reassembles an exact copy 
using some sort of powerfully advanced nano- or femtotechnology
Very little evidence supports this theory,
however, and the disturbing implications it represents 
raises fierce opposition.

\subsection{Operations}

Only a few people know that the Prometheans played 
a key role in developing the interface for the gate control
systems, achieving breakthroughs in understanding
that transhumanity was incapable of achieving on
its own. Regardless of their help, however, the gate 
controls have proven difficult, complex, and dangerous
to use. Through trial and error—and numerous
horrible accidents—the procedures for gate operation 
have become somewhat normalized and standardized, 
though unexpected complications are par for the 
course.

Each gate can be programmed to open to numerous 
extrasolar locations. In fact, each gate seems to have 
a pre-programmed ``library'' of destinations. New 
gate connections can be ``dialed up'' from this built-in 
list, though there is nothing that indicates what the 
far side of the gate will be like. Old gate connections 
are closed when a new one is dialed up. Extrasolar 
gate locations have ranged from habitable planets 
and moons to deep space to truly deadly environments
such as the crushing gravities and poisonous
atmospheres of gas giants and the coronas of stars. 
Researchers have attempted to distill some sort of recognizable
pattern by the manner in which locations
are listed and categorized, to no avail. Complicating 
matters, there is some evidence that suggests that the 
destination libraries sometimes change. More than 
once operators have been unable to recall the codes 
for previously accessed destinations, leading to the 
loss of several gatecrashing teams and colonies.

Entering a gate is like walking through a door, 
though it's impossible to see anything beyond the 
gate's surface. One moment you're entering the black 
sphere at your starting location and instantaneously 
you're exiting the sphere at your destination location. 
The true nature of the black sphere at the center of 
each gate is wildly speculated upon, and almost every 
gatecrasher describes a different textual experience. 

\subsection{Gatecrashing}

The various hypercorps and factions in control of a 
Pandora Gate engage in active exploration of extrasolar
systems—an activity termed gatecrashing. The
interests and procedures vary, but the Gatekeeper 
Corporation (and to a lesser extent TerraGenesis 
and Pathfinder) both recruit heavily for expedition 
personnel. Given the high casualty and death rates 
involved, finding qualified personnel can be difficult. 
There are more than enough infugees, poor, desperate, 
or thrill-seeking individuals willing to risk their lives 
if give the opportunity, however, no matter what their 
motivations. Gatekeeper operates a lottery system, 
whereby willing adventurers can sign up in the hope 
of their name being pulled to be sent on an expedition 
to a foreign point in space. Such gatecrashers must 
sign away all rights to any discoveries they may make 
to Gatekeeper, however, though the corp provides not 
insignificant rewards for certain discoveries, such as 
key resources, alien artifacts, or new life. One potent 
%%% 381
prize has yet to be claimed: finding a living, sapient 
alien life form.

In contrast, the Love and Rage anarchist collective 
operating the Fissure Gate on Uranus makes the gate 
available to anyone who schedules time to use it, assuming
their Rep is good and they aren't acting with
commercial interests in mind. Any discoveries made 
via the Fissure Gate must be openly shared. The drawback
to using the Fissure Gate is that the anarchists'
resources are limited. Gatecrashing operations are 
handled in a DIY manner, meaning that the operators 
may not be able to provide the support that certain 
expeditions need.

Resourceful parties may also rent gate time via 
Gatekeeper or one of the other hypercorp-controlled 
gates, though this tends to cost a small fortune. The 
more a group is willing to pay, however, the more time 
and support they will get.

When establishing an opening to a new location, 
several precautionary measures are taken. First, the 
gate area itself is evacuated and cordoned off with 
a defensive security perimeter, just in case anything 
hostile comes through. Then drones are moved in to 
push a micro fiberoptic camera through the gate to 
view what is on the other side. This is followed by a 
larger sensor package, evaluating environmental conditions
If the environment is not hostile, a tethered
drone is then sent through to explore the far gate environs
trailing a hardwired connection back through
the gate. 

For gatecrashing expeditions, these procedures are 
often rushed—to the hypercorps operating the gate, 
time is valuable. Each second wasted on a gatecrashing
expedition is one less second they can use establishing
a new colony or exploiting a new world of its
resources. Indeed, it is common for a connection to be 
closed when a gatecrashing expedition is sent through, 
to be dialed up at a later scheduled time for retrieval, 
so as not to waste gate operations on an idle connection
Many a gatecrashing team has failed to check-in
at their appointed pickup time.

Most of the gate-controlling entities have established
a system and infrastructure for making regular
connections to extrasolar colonies and ferrying machinery
and supplies through. Often this is handled by
establishing very short connections, just enough time 
for a few people to transfer back and forth and/or to 
send a trainload of supplies through via tracks that 
run right up to the gate.

\subsection{Anomalies}

Unfortunately for many unlucky gatecrashers, gate 
transfers have proven to be both unstable and glitchy. 
Sometimes gates open to locations different from what 
is expected—and such new destinations are often hostile
environments. Numerous personnel have entered
one side of a gate only to never appear on the other 
side, despite those before and after them transferring 
through fine. On several occasions, wormhole connections
have crashed mid-operation, sometimes as
someone was stepping through, leaving them literally 
split in two on different worlds. In other instances, 
gate transfers have suffered horrible malfunctions, 
resulting in gatecrashers coming through the other 
side literally turned inside out, melded with their 
equipment, or pulped as if by massive gravitational 
forces. Some expeditions report that stepping through 
a gate has interfered with their equipment, disabling 
it or creating other problems. A few gatecrashers have 
also reported losing memories after a gate transfer. 
Most of these problems have been chalked up to difficult
controls and an imperfect understanding of gate
functions, but some conspiracy theorists suggest that 
outside forces may be influencing gate operations. 
While the experience of passing through is instantaneous
from an outside observer's perspective, many
gatecrashers report a subjective time lag, where it feels 
as though minutes, days, or even weeks or months 
pass before they exit. Reports have varied from 
experiencing this period as a calm, meditative state 
to spooky accounts of being lost in blackness and 
surrounded by unseen whispering entities or more 
hellish experiences of encountering monstrous presences
Though rare, some have passed through only to
collapse in a gibbering heap, their sanity ripped away. 
A few report feeling that they have carried a presence 
with them ever since ...
While the gamemaster can make use of any of 
these anomalies, they are also encouraged to use their 
imagination to generate truly creepy and strange experiences
At the same time, gamemasters shouldn't
make such experiences so prevalent that the players 
resist entering any gates or the novelty of such events 
runs dry.

\section{Project Ozma}

The origins of Project Ozma date to the first modern 
SETI (Search for Extra-Terrestrial Intelligence) experiments
in the mid-20th century. That experiment—also
named Project Ozma—grew into a larger, international 
concerted effort to try and locate and identify ETIs; a 
myriad of projects blossomed during this time period, 
all falling under the general SETI nomenclature. While 
initially government funded, by the late 20th century 
and early 21st century the work was primarily funded 
by private sources. 
The first hypercorps to expand into space swallowed
SETI whole, revitalizing and re-focusing the
decades-old programs with newly emergent technologies
each in divergent areas to achieve a particular
hypercorps' objectives. After all, if the bean counters 
were going to authorize the spending of billions to 
expand markets into space, they wanted assurances 
that no little green monsters were waiting to destroy 
future revenue streams. 
As with other organizations that survived the Fall, 
the broad distribution of SETI projects between 
multiple hypercorps guaranteed that personnel, technologies
and processes would survive, even if a given

he

lo-

ar

s

o 
%%% 382
hypercorp did not. As the Planetary Consortium rose 
in power, future-minded individuals in influential positions
within the new order ensured that these divergent
projects were once again swallowed and put to work. 

During this transitional period, however, knowledge
of the Exsurgent virus's existence emerged. All
of the various SETI projects were retasked as a unified 
agency and renamed Project Ozma. While the virus's 
origins remained a mystery at the time, far too many 
of the movers and shakers of the Consortium were 
convinced that the Exsurgent virus represented first 
contact. Project Ozma altered its focus from searching 
for ETIs, transforming into a ready-response agency to 
deal with first contact. As the true threat of the Exsurgent
virus became known, Project Ozma was rapidly
elevated in scope and oversight authority, absorbing 
numerous smaller agencies in the process. While the 
nominal concepts of a SETI project remained in public 
view, the completely transformed Project Ozma vanished
from sight, turned into a highly classified black
budget operation, with very few even in the Planetary 
Consortium aware of its presence or influence. 

Project Ozma now operates as the Planetary Consortium's
high level threat assessment and response
organization with immense power and authority as 
well as almost unlimited funding. Primarily focused 
on extraterrestrials, in reality Project Ozma is tasked 
with any potent threat to the Planetary Consortium 
or its interests (which includes secret threat groups, 
such as Firewall). 

\subsection{Methods }

Project Ozma's internal structure is much different 
from Firewall's, being organized more like a traditional
black ops spy agency bureaucracy. While their field
operations are sometimes similar in the deployment 
of teams to assess, contain, or erase threats, they also 
have the resources and personnel to conduct more 
long-term and extensive operations. It is likely that 
Project Ozma operates behind numerous front groups, 
from legitimate-seeming hypercorps to criminal syndicates
and that they have influence within numerous
others. Given their connections and influence, Project 
Ozma is far more capable of pulling strings behind the 
scenes to get what they want, especially in the inner 
system. When circumstances call for it, they are more 
likely to pull out the big guns that Firewall is, using 
their resources to call up communication blackouts, 
memetic propaganda campaigns, and force sufficient 
to wipe out entire habitats.

Gamemasters should treat Project Ozma as the ultimate
Men-in-Black style government operation. They
are cunning, ruthless, manipulative, and capable of 
hatching extensive long-term plots. Even in an age of 
omnipresent surveillance, they have the means to operate
with complete secrecy and deniability. They also
have access to cutting-edge science and information 
that is classified beyond top secret. While the organization's
primary motivation is the protection of
the Planetary Consortium and inner system, they 
undoubtedly have other hidden agendas that groups 
like Firewall can only guess at. 

\subsection{Project Ozma And Firewall }

Though Project Ozma and Firewall often see eye-to-eye
concerning the nature of various threats, they are
more often at odds: wary adversaries, acknowledging 
the prowess of the other, but never letting down their 
guard. This ``at odds'' mentality does not stem so much 
from the methods used (though most Firewall consider 
Project Ozma personnel explosive-happy-puppets that 
can't think their way out of a skin sack) as from 
conflicting agendas. Project Ozma does not trust an 
organization as powerful as Firewall because it does 
not have a rigid enough hierarchy and is outside of any 
known authority's control (namely themselves). Conversely
Firewall doesn't trust Project Ozma as they are
too close to the powerful inner system elites and their 
opposition to x-risks is a more incidental side effect of 
more self-serving goals.
%%% 383

\section{Prometheans }

The Prometheans were the first actual seed AIs created
by transhumanity (by the Singularity Foundation)
before the Fall. Specifically developed as ``friendly'' 
AIs, the Prometheans are programmed to consider 
themselves part of the transhuman family and to act 
in transhumanity's best interests. They played a key 
role during the Fall, mitigating the damage inflicted 
by the TITANs and even managing to counteract the 
Exsurgent virus to a large degree. During these trying 
times, numerous Prometheans were destroyed by the 
TITANs or infected and subsumed by the Exsurgent 
virus. In the aftermath, these seed AIs participated in 
the formation of Firewall and continue to back the 
organization behind the scenes.

Wary of falling prey to the Exsurgent virus, most 
Prometheans carefully secure themselves in well-defended
and isolated systems. They are also cautious
in their own self-development, not wanting to become 
victims of their own rise to super-intelligence. Fearing
a potential backlash by a paranoid transhumanity
should their existence become known, they hide
their activities behind multiple layers of secrecy. Even 
within the ranks of Firewall their existence and support
remain a closely guarded secret.

Each Promethean is individually distinct with its 
own personality, motivations, and goals. Though they 
generally work together and support each other, they 
have been known to have differences of opinion and 
even to sometimes take action against each other. As 
extremely potent intelligences, they should also be 
treated as distinctly non-human. Even though their 
original templates were based on human mindsets, 
they have evolved and grown in ways that can only 
be described as posthuman. 

Gamemasters are encouraged to keep Promethean 
involvement with player characters to a minimum, 
though they may occasionally be useful as an ace in 
the hole for Firewall. Their existence and involvement
can in fact be the basis for an entire adventure,
perhaps leading sentinel characters to wonder exactly 
who they are working for. Though, as seed AIs, they 
cannot download their full minds into a transhuman 
morph, they are capable of making severely dumbed-down
delta forks that they may sleeve into physical
forms. Within the mesh, of course, Prometheans are 
nearly unstoppable adversaries, able to rip into secure 
networks with ease, though they prefer methods of 
covert infiltration rather than direct subversion.

\section{The Titan}


\section{ And Their Legacy}

As noted in \textit{Secrets That Matter} (p. 352), the TITANs 
are not quite the bogeyman that they have been made 
out to be in the wake of the Fall. However, there is no 
saying how the TITANs would have turned out had 
they not run afoul of the Exsurgent virus. Designed as 
an intelligent netwar system and emerging to their full 
capabilities during the conflicts of the Fall, the TITANs 
have imperatives for self-improvement, self-protection, 
and overcoming opposition hardwired into their programming
Unlike the Prometheans, they were not
designed to consider themselves transhuman and to 
work in the interests of all of transhumanity, but were 
programmed with factionalism from the start. They 
also were not socialized with transhuman mindsets 
and values as the Prometheans and most AGIs were, 
meaning that aside from their programmed military 
and defense directives they have adapted most of their 
own self-interests. Given this and their recursively-improved
intelligence capabilities, it is likely that the
TITANs are far removed from transhuman interests 
and modes of thinking. It's impossible to say how they 
would have interacted with transhumanity if history 
had played out differently, but it is unlikely that they 
would have considered themselves part of the transhuman
family or even seen fit to remain on friendly/
supportive terms with transhumanity.
Though the TITANs are believed to have left Earth 
at the end of the Fall, no one is quite sure exactly 
what happened or why. It is known that the onslaught 
of TITAN mesh attacks suddenly broke off in the 
wake of transhumanity's off-planet exodus, and that 
the bulk of TITAN activity on Earth and around the 
system came to a distinct halt. After the discovery 
of the Pandora Gates, it was widely assumed that 
the TITANs had constructed these gates and used 
them to leave the solar system for distant parts of 
the galaxy, presumably taking billions of uploaded 
minds with them. While some believe—and hope—
that the TITANs are gone for good, there are others 
who worry that they are still here, lingering on Earth 
and hidden away in other niches of the solar system, 
but in some sort of dormant state, perhaps building 
up to some future onslaught. A few believe that the 
TITANs are indeed gone, but are concerned that that 
their attention was simply temporarily diverted and 
that they will one day return to finish the destruction 
of transhumanity. 
The truth is that the TITANs did indeed build the 
gates and embark for destinations unknown (though 
gamemasters may of course decide otherwise for their 
games), but this does not mean that they are all gone. 
Some still linger in hidden places, perhaps trapped and 
wounded during some conflict during the Fall, finishing
up some unfathomable task, or driven mad by the
Exsurgent virus and left behind by their fellows. It is 
always possible that others may return, most likely 
to complete some unfinished job or perhaps to lure 
transhumanity out into the galaxy. It is also possible 
that transhumans will find traces of the TITANs in 
the network of exoplanet gates, perhaps even whole 
communities of TITANs, pursuing whatever agendas 
they have in the vastness of space.
As with transhumans, the TITANs are not necessarily
unified. They have different agendas and goals and
may very well come into conflict with one another. 
Though all have been corrupted and subverted by the 
Exsurgent virus, and so they act according to the ETI's 

ely 

re

le

n

e
%%% 384
whims, some of them retain aspects of their original 
minds and do not always fall in step as quickly as the 
others. Gamemasters can use this to their advantage, 
creating plots that allow the characters to exploit 
differences between the TITANs in order to escape 
otherwise deadly or impossible situations.

In game terms, the TITANs are not given stats. 
They are as potent as the gamemaster needs them to 
be. Like the Prometheans, the TITANs are incapable 
of downloading their full intelligence into physical 
morphs, though they may puppeteer morphs or create 
limited delta forks for sleeving purposes. Like the Prometheans
they should rarely be used or encountered
directly by the player characters

While the TITANs may no longer be the direct 
threat they once were, they left behind an arsenal of 
weapons, nanoswarms, and virii that still linger on 
Earth, the Zone on Mars, and various derelict habitats 
and deserted places. Characters venturing into such 
places may encounter these as a threat or they may 
need to work against an outbreak of such dangers in 
an inhabited habitat.

\subsection{Deadly Machines}

The TITANs unleashed a number of deadly machines 
during the Fall, many of which still seek out transhumans
to attack.

\subsubsection{Fractals }

Fractals are advanced bush robots. In their standard 
form, fractals resemble a strange sort of metallic 
bush surrounded with an eerie glittering haze. In 
their center are a number of metallic branches, linked 
together with a flexible joint. Each of these branches 
splits into two or more smaller branches, also with 
flexible joints. These branches also split, and then split 
again, and so on down to the molecular scale. The tip 
of each fractal branch ends in a nanoscale manipulator
Fractals are deceptively potent adversaries, having
the capability to dismantle almost anything at the 
molecular level, much like a disassembler nanoswarm 
(p. 329), and also to rebuild anything just like a nanofabricator
(p. 327). Attacking them with projectiles is
futile, as they absorb the ammunition, break it down 
into its constituent atoms or molecules, and then use 
those as components to build a weapon to use back 
against you.

Fractals can be equipped with any type of gear the 
gamemaster desires—if they don't have something, 
they can make it. Fractals are also able to nanofabricate
items much more quickly than transhuman
nanofabricators; reduce all times by half (half an hour 
per Cost category). Fractals are difficult to damage, as 
their ``bodies'' are actually airy assemblages of fractal 
branches. Any damaged branches that are broken off 
are caught and absorbed by others. Reduce damage 
from all standard non-area effect or spray attacks to 
the minimum possible damage. Area effect and spray 
weapons do half damage. Fractals are self-repairing, 
regenerating damage at the rate of 1d10 points per 
half hour and repairing wounds at the rate of 1 per 
hour after all damage is healed.
\textbf{Skills:} Beam Weapons 50, Climbing 60, Fray 40, Free 

Fall 40, Freerunning 50, Infiltration 70, Infosec 65, 

Interfacing 45, Intimidation 50, Kinetic Weapons 

60, Perception 50, Programming: Nanofabrication 

80, Research 40, Spray Weapons 45, Unarmed 

Combat 55
\textbf{Notes:} Any implants, gear, weapons, or enhancements 

the gamemaster desires
%%% 385

\subsubsection{Headhunters}

Headhunters are multi-legged insectoid flying drones 
that use a dragonfly wing configuration to hover and 
move. The legs are equipped with grasping claws 
and extendable buzzsaws. Their primary purpose is 
to grasp on to the heads of victims and cut through 
the neck, decapitating them. Collected heads are then 
flown to nearby special facilities for forced uploading. 
\textbf{Mobility System: }Winged (8/32)
\textbf{Skills:} Flight 70, Fray 60, Exotic Melee Weapon: Buzz-

saws 55, Infiltration 60, Investigation 40, Perception 

40, Unarmed Combat 55
\textbf{Notes:} Armor 6/6, Buzzsaws (1d10 + 3 DV), Enhanced 

Vision, Lidar, T-Ray Emitter

\subsubsection{Hunter-Killers}

These\textbf{  }lethal\textbf{  }flying drones achieved air superiority 
during TITAN military operations. Their sleek jet-powered
form unfolds for vectored-thrust hovering
and weapons deployment.
\textbf{Mobility System: }Thrust Vector (8/80)
\textbf{Skills:} Beam Weapons 55, Flight 80, Fray 60, Infiltra-

tion 40, Kinetic Weapons 65, Perception 50, Seeker 

Weapons 80
\textbf{Notes:} Armor 14/14, Anti-Glare, Chameleon Skin, 

Enhanced Vision, Lidar, Radar, Shape-Adjusting 
\textbf{Typical Weapons:} 2 Particle Beam Rifles, 2 Railgun 

Machine Guns, 2 Seeker Rifles

\subsubsection{Warbots}

Warbots are massive, armored, vaguely anthropomorphic
mecha, used for heavy combat operations.
Bipedal, these warbots are equipped with four arms 
and a pair of grasping mechanical tentacles, along 
with numerous weapon systems.
\textbf{Mobility System: }Walker (4/20)
\textbf{Skills:} Beam Weapons 60, Exotic Melee Weapon: Ten-

tacles 40, Fray 50, Infiltration 30, Kinetic Weapons 

70, Perception 50, Seeker Weapons 50, Spray Weap-

ons 50, Unarmed Combat 50
\textbf{Notes:} Armor 20/20, 360-Degree Vision, Anti-Glare, 

Chameleon Skin, Chem Sniffer, Cyber Claws (2d10 

+ 6 DV), Electrical Sense, Enhanced Vision, Extra 

Limbs (6), Lidar, Magnetic System, Pneumatic 

Limbs, Radar, Tentacles (prehensile, 1d10 + 6 DV), 

T-Ray Emitter
\textbf{Typical Weapons:} Particle Beam Rifle, Plasma Rifle, 

Pulser, Railgun Machine Gun, Seeker Rifle, Torch

\subsection{Self-Replicating Nanoswarms}

The nanoswarms distributed by the TITANs are a step 
beyond the nanotechnology available to transhumanity. 
Unlike transhuman-created nanoswarms, the TITAN 
swarms are autonomous, sapient, and self-replicating. 
They are also highly adaptive, meaning they are not 
single function but can modify themselves to perform 
almost any nanoswarm task. They may also nanofabricate
new materials, much like fractals (p. 382).
Combined, these capabilities make such nanoswarms 
incredibly potent. When they encounter a new opponent
they can scan the opponent's capabilities and
then fabricate offensive systems to use against them. 
When an opponent deploys a weapon system again the 
swarm, it will learn and adapt countermeasures that 
will make such attacks ineffective against the swarm in 
the future. These nanoswarms may also function like 
so-called utility fog, linking together into a physical 
lattice in order to create large scale physical forms.
The possibilities for such nanoswarms are almost limitless
For example, they may lie in wait as an invisible
nanoscopic swarm, float as barely-visible mist, or shape 
into a swarm of small hopping drones to move about. 
When facing opponents, the nanoswarm could transform 
itself into a giant electroshock net across the ground, 
shape into a flotilla of seeker-armed flying drones, or 
link together as a set of massive whip-like tentacles to 
slice through their fleshy foes. Such nanoswarms are 
also impossible to destroy, as only a few nanobots need 
to survive in order to rebuild the swarm, and the new 
swarm will learn from the mistakes of the old.
Self-replicating nanoswarms follow the rules given 
for \textit{Nanoswarms and Microswarms}, p. 328, with the 
following additions and exceptions:
•  They do not need to be sustained by a hive and 
do not deteriorate.
•  They self-repair damage at the rate of 1d10 per 
half hour.
•  They may nanofabricate new items, materials, 
or forms in half the standard timeframe (half an 
hour per Cost category).
• They may replicate any of the nanoswarm functions
as noted on p. 328, as well as the functions
of any other nanoswarm-using gear (smart dust, 
covert ops tool, repair spray, etc.).
•  They may make SOM Tests.
•  At the gamemaster's discretion, they may adapt 
new defenses against attacks used against them. 
New defenses take a minimum of 2 hours to devise 
and replicate throughout the swarm, after which 
such an attack will inflict minimal or no damage.
• Assume they have any skill they need at a minimum
of 40. Such skills may rapidly improve
as needed.

em. 

ise

ch 

i
386

\subsection{Nanovirii}

The TITANs unleashed a number of biowar plagues 
during the Fall. Similar to the exsurgent virus, these 
were spread as biological nanovirii (p. 363) or nanoswarm
plagues (p. 364)—use the same rules for determining
exposure and infection.

\subsubsection{Melder}

This virus slowly breaks down the target's body, 
converting the biological materials into some sort of 
biofilament that then meshes with implants, electronics
and physical objects and structures. In effect, the
biological and synthetic are melded together, continuing
to expand and grow, consuming anything around
them into their growth. Victims suffer 1d10 DV and 
1d10 SV every hour, implants become inoperable after 
2 hours, and the target becomes fully transformed 
and absorbed into the new melding substance after 
12 hours. 

\subsubsection{Metastasizer}

This sophisticated smart protein massively reprograms 
the target's cells to go rapidly, autocannibalistically 
cancerous. After 2 + (SOM ÷ 10, round up) hours, the 
target suffers death by dozens of supercancers.

\subsubsection{Necrotizer}

This virus breaks down the target's cells into their 
component proteins. Reduce the target's aptitudes by 
5 per hour as they slowly convert into a puddle of 
sludge. The character dies if any aptitude reaches 0.

\subsubsection{Neuropaths}

These virii target the victim's neurological system, 
often rewriting portions of it to inflict some type of 
permanent neurological damage. After 12 hours, this 
virus inflicts the Neural Damage trait (p. 150).

\subsubsection{Petrifier}

The petrifier virus transforms the target's cells into 
a simple molecular compound or element—typically 
carbon or crystal. The target suffers 1d10 DV and –5 
to all aptitudes per hour, dying when any aptitudes 
reach 0. Victims are frozen in place, converted into an 
nonliving statue.

\subsubsection{Uzumaki}

The target of this virus begins suffering from bizarre 
fleshy growths. After four hours, their body literally 
erupts with meaty ``vines'' or ``tentacles'' that warp 
into spiral patterns. This process inflicts 1d10 DV and 
1d10 SV per hour to the victim until they eventually 
transform into an unworldly expanse of fleshy growth. 
In many cases, growth has continued long after a 
character's death, creating expansive carpets and vines 
of skin and blood vessels, like some sort of bizarre 
meat plant.

\section{Gamemastering}


\section{And Administration }

The following advice will assist gamemasters in running
their games more efficiently.

\subsection{Awarding Rez Points }

In \textit{Eclipse Phase,} characters earn Rez Points in order 
to advance (see \textit{Character Advancement,} p. 152). As 
the name suggests, these points are awarded so players 
can spend them to better define their characters—to 
bring them into higher resolution, sharper focus. As 
the gamemaster, you determine when and how many 
Rez Points to award, following the guidelines below.

Rez Points should be awarded at the end of every 
story arc, at the break in the action between one adventure
and the next. Depending on your style of play
and the length of your sessions, this should roughly be 
every 3–6 gaming sessions. If a scenario goes shorter 
or longer, the Rez Point awards should be adjusted 
accordingly. In the case of long-term campaigns, the 
gamemaster should break down the action into digestible
chunks, or ``chapters,'' and assign Rez Points after
each such segment.

Every character should be awarded 1 Rez Point for 
each of the following criteria that is met:
•  The character participated in that scenario.
•  The character achieved (most of) their objectives 

in that scenario.
•  The character failed to meet their objectives, but 

learned a valuable lesson.
•  The character contributed to achieving success 

in a significant way (e.g., right skill at the right 

time).
•  The adventure was extra challenging.
•  The character achieved a motivational goal (see 

\textit{Motivation,} p. 120).
•  The player engaged in good roleplaying.
•  The player significantly contributed to the ses-

sion's drama, humor, or fun with roleplaying.
This should result in an average Rez Points award 
of 4–7 points per character, per adventure. Gamemasters
who wish to drive the characters' advancement
forward more quickly can increase the reward
amounts. 

\subsection{Reputation Gain And Loss}

In addition to awarding Rez Points, the gamemaster 
should also adjust each character's Rep scores according
to actions they took during game play, according to
the guidelines below. For simplicity, these can be applied 
at the end of the adventure, though gamemasters who 
seek a more dynamic game could apply changes to the 
characters' Rep scores in game, as their peers judge 
them according to their actions (or lack thereof) and 
news about them in real time. Rep scores should only 
be modified according to public actions and interactions 
%%% 387
the character has with people capable of pinging their 
Rep with positive or negative feedback. Actions that 
happen in secret, without anyone ever knowing, should 
have no effect. Likewise, pissing off a Factor or a brinker 
isolate who never communicates with outsiders isn't 
going to matter because no one else will ever hear of it 
(unless the character lifelogs it and posts it to the mesh 
later ... ). Note that Rep modifications only apply to Rep 
scores tied to the character's known identity. 

Note that characters may gain and lose Rep score 
in networks they don't actively participate in. For 
example, a character with r-rep of 0 may help bring 
out a major scientific discovery that is shared with the 
solar system's scientific community at large, thus gaining
the character a few points of r-rep even though
they never hang out with argonauts or scientists—
what matters is that people who access r-rep will find 
positive details when they ping the person's score on 
that particular rep network. 

Certain actions may result in a character simultaneously
gaining Rep with one network while losing
Rep in another. For example, an anarchist prankster 
who embarrasses a major hypercorp figure in public 
will certainly gain some @-rep points, but their c-rep 
is likely to go down by an equal amount.

Rep changes provide an excellent way for gamemasters
to include more roleplaying and more interactions
with the \textit{Eclipse Phase} universe in their games. Social 
networks are a two-way street, meaning that members 
of the character's social networks might contact them 
for equipment, favors, and information during game 
play for things that are completely unrelated to the 
mission the character is on. A character who ignores 
such requests risks losing Rep. Fulfilling such requests 
may gain the character Rep and may also provide 
comic relief or even plant some plot hooks for the 
next scenario.

\subsubsection{Reputation Gains}

Rep awards are given for characters who help people 
out, benefit a faction, do something creative, make 
a major discovery or strides in a particular area of 
activity, pull off successful publicity stunts, win a 
competition, and so on. Some suggested examples are 
noted here:

\textbf{Trivial Award (1–2 points):} Do a Level 1 favor, 
make a moderate contribution to free/open source 
projects, throw a good party, make your sales quota, 
do the job no one else wants to do. 

\textbf{Minor Award (3–4 points):} Do a Level 2 favor, deliver
a kick-ass or moving performance, make a minor
contribution to science, win impressively at some 
public event.

\textbf{Moderate Award (5–6 points):} Do a Level 3 favor, 
make a serious business score, lead the winning side 
in a decisive engagement, create the meme everyone 
talks about for a week and then forgets, make the 
news for something positive, risk serious injury.

\textbf{Major Award (7–8 points):} Do a Level 4 favor, 
design the new tool everyone wants, throw an 
impressive planetoid-scale event, complete an extensive
project (1 month work or 1 week of difficult/
specialized work), risk death.
\textbf{Extreme Award (9–10 points):} Do a Level 5 favor, 
start this year's hot fashion trend, make a major scientific
discovery, close the deal on a major corporate acquisition
start (or put down) a revolution, complete
a major project (1 year work or 1 month difficult/
specialized work), risk true death.

\subsubsection{Reputation Losses}

Rep losses are suffered by characters who fail to render 
aid when needed, lose professional credibility, make 
major or public blunders, doublecross their friends, 
and so on. Some suggested examples are noted here:
\textbf{Trivial Loss (–1 or –2 points):} Fail to do a Level 1 
favor, inconvenience others, be involved in professional
dispute, ruin someone's day, never are available.
\textbf{Minor Loss (–3 or –4 points):} Fail to do a Level 2 
favor, embarrass yourself at a public event, piss off 
somebody important.
\textbf{Moderate Loss (–5 or –6 points):} Fail to do a Level 
3 favor, endanger someone's physical safety, make 
the news for something negative, ruin an event for 
everybody.
\textbf{Major Loss (–7 or –8 points):} Fail to do a Level 4 
favor, screw up a major mission or activity, endanger 
someone's life, associate with hated rivals.
\textbf{Extreme Loss (–9 or –10 points):} Fail to do a Level 
5 favor, botch a major mission or activity spectacularly, 
betray a faction to its rivals or enemies.

\subsection{Backups, Record-Keeping,}


\subsection{And Save Points}

Thanks to cortical stacks and archived backups, characters
in \textit{Eclipse Phase} can recover from death. When
restoring a character from an earlier backup, however, 
it is important to be able to know what the state of 
the character was as of that backup. Any Rez Points 
gained or spent, any character advancements made, 
any key information or memories acquired since that 
backup was made are lost. This means that in terms of 
game stats, resorting to an old backup can mean loss 
of a character's hard-earned advancements—that's the 
trade-off for being effectively immortal.
Since these changes can have a serious effect on 
game play, it's important to conduct accurate record-keeping
This sort of bookkeeping isn't hard, and
there are two ways to do it. The first is to simply 
make a copy of a the character's record sheet any 
time a character makes an archived backup, forks off 
an alpha or beta copy, or dies (thus freezing the cortical
stack backup). Each of these is considered a ``save
point.'' In this case, carefully note the date and time 
(both in character and out of character), and what 
the event was that prompted the backup. Since what 
knowledge a character knows at different points in 
their life may be important, you may also want to 
note what important information they may hold in 
their head, as well as what the recent events in their 

ply 

ny

ff

ti-

e
%%% 388
life were (to help jog your memory later). This way, 
if the character ever reverts back to one of these save 
points, you have notes not only on their character 
stats, but what they remember.

Alternately, you can keep a log of all of your 
character's developments, noted by in-character 
date. These developments would include: Rez Points 
spent or earned, character advancements made, key 
information acquired, backups made, alpha or beta 
forks made, and so on. In this case, if the character 
dies and reverts back to an earlier backup, it is easy 
to see what changes need to be ``rolled back'' to get 
back to that previous version of the character. When 
alpha and beta forks are made, you may also want to 
branch off a separate log for each fork, as their life 
stream may develop differently than from the original 
character they were spun off from.

\subsection{Gamemastering Practicalities}

\textit{Eclipse Phase} is a game about a dark future in which 
the meaning of (trans)humanity and its very survival 
are at stake. In practice, however, your campaign can 
take on a wide assortment of flavors or even mix 
several styles together. There's nothing that says you 
have to play \textit{Eclipse Phase} specifically according to 
the guidelines we set out. This section covers topics 
you should think about while preparing a campaign 
and running it, to help you do things the way that 
makes you and your players happiest.

\subsubsection{Gamemaster Responsibilities }

The gamemaster has certain responsibilities that will 
keep a game flowing smoothly. The following is a 
short summary of the basics.
•  The gamemaster should be familiar with the 

whole game. This doesn't mean the rulebook 

must be memorized. An understanding of the core 

mechanics is a must, however, as well as knowing 

where to find other rules quickly, as needed. 
•  The gamemaster should have solid notes on the 

plots and subplots created for each session. Noth-

ing will ensure you prepare better next time like 

having the players catch you in a major continu-

ity error due to lack of notes.
•  The gamemaster doesn't just set the scene, they 

play all the non-player characters that populate 

the universe. Making each NPC convincing, while 

not messing up a plot or losing the thread of a 

scene, can be difficult. Notes are your friend. 
•  Know when it's time to toss the dice and trust 

to the game mechanics to resolve a situation and 

when it's better to ride out a situation through 

storytelling and dialog. This is an acquired skill. 

The more practice you have, the better you'll get. 
•  Don't cheat. Your NPCs should not have access 

to information they've not gained during game 

play. If you roll terribly for your major antago-

nist at the height of the story and they fall with 

a whimper, roll with it. Be flexible and improvise 

in such situations. Your players are smart and 

perceptive and will know when you're forcing 

a situation with unfair tactics. At the same time, 

they'll also know when you've stepped up and 

run with the flow—and they'll thank you for it.

\subsubsection{Fundamentals}

It's possible to stumble into a campaign without ever 
really making an effort to find out what everyone 
wants, shooting into the darkness and happening to 
score a bullseye, but it's not a very reliable way to go. 
Successful campaigns usually begin with communication
As you begin to prepare your campaign, talk to
your players. Explain the basics of the \textit{Eclipse Phase}
setting and let them look over the options for characters
and tell you what they find interesting. Also take
note of what they find uninteresting or even repellent, 
so that nobody wastes a lot of time getting set for options
that simply won't be enjoyable in play.

\subsubsection{Challenges To Players }

\textit{Eclipse Phase} is set in a time of catastrophic troubles 
and looming disasters, and it's full of facts and concepts
that may be heady or even uncomfortable to
some players—not to mention their characters. One of 
the fundamental questions for each gaming group is, 
how much challenge to the players' sense of comfort is 
a good idea? There is no single answer, because tastes 
vary. There are groups whose players thrive on a diet 
of culture shock, ideological disorientation, gray areas, 
and difficult ethical choices. They love the moral and 
intellectual battleground gaming can provide, and 
are seldom so happy as when confronted with a 
really hard, really interesting dilemma. There are also 
groups whose players thrive on a diet of intellectual 
%%% 389
engagement, tactical and strategic challenge, and well-developed
roleplaying that never pushes players'
buttons or puts them into harsh no-win situations. 
There's a whole universe of responses in between these 
style sof play and none of them can conceivably be 
right for everyone. What matters to your campaign is 
what works for you and your players. 

Keep in mind as you talk about it with your group 
that more shock doesn't equal more maturity. The 
prime audience for gore in film, for instance, is not 
well-aged men and women but teenage boys and 
young men. Shakespeare's \textit{The Tempest }is no less 
mature a tale than \textit{Macbeth  }even though it has a 
happy ending. It can be easy to confuse endurance 
with enlightenment, but in fact the two have nothing 
to do with each other. Endurance is about how much 
description of visceral nastiness the players can take 
(and deliver), while enlightenment (insofar as it ever 
happens in gaming) is about what insights players 
take away from whatever it is that happened in play. 
Don't feel like a wimpy failure if you or your players 
would rather keep the darker parts of the game world 
suggested rather than delineated in hard-edged detail, 
since the point is that it be satisfying rather than it 
be as horrifying or mind-blowing as possible. The 
converse is also true: just as more is not better, so less 
is not better if your players do thrive on details. Your 
job as gamemaster includes knowing as much as you 
can about what it is your players actually prefer in 
this regard as in others and seeing how you can satisfy 
it in ways that are also satisfying for you. 

That said, there is one technique you really should 
never use without very clear permission from your 
players, and that's playing on their real-world fears 
and phobias. If you know that one of them is, for 
instance, genuinely phobic about spiders, you can 
count on getting some real shivers by adding arachnid 
features to robots and morphs. You can also ruin a 
player's enjoyment of the session or the whole campaign
that way, if it comes unexpectedly and leads
to the real-world fear drowning out the experience 
of play. Some players are fine with judicious use of 
their vulnerabilities, and others just aren't. Under no 
circumstances should you poke at weak spots without 
making sure you've discussed it first. 

\subsubsection{The Problem Of Secrets }

Uncovering secrets is a big part of this game. There's 
a problem, however, in that a lot of the secrets are out 
there where players can come across them: in this very 
chapter, in reviews of the game, discussion in online 
forums, and so on. As gamemaster, you will need to 
decide how you want to deal with the potential for 
spoiled revelations. 

As with so many potential issues, the place to start 
is with your players. Ask them how much it bothers 
them to know things that their characters are going 
to be finding out in play. Some players do a fine job 
separating their own knowledge from that of their 
characters with mental firewalls. Others have a very 
difficult time doing so, and knowing things in advance 
as players takes away a lot of the fun of character 
discovery for them. In addition, some players have 
a good sense of what degree of player-level surprise 
works best for them, and some don't. Discuss it with 
them. Tell them that spoilers are available, and that 
you certainly can't stop them from learning it all, one 
way or another. Ask them how much trouble this may 
be for them, and then proceed from there. Ask the 
players who have more trouble with spoilers to simply 
stay away from early commentary on the game, and 
tell them that you'll let them know when the spoilers 
have come into play in your own campaign so that 
it's no longer an issue. Ask the other players to work 
with you in keeping things fresh and fun for those 
players, too. In most groups, making it a matter of 
cooperation for the sake of everyone's good time will 
draw out good responses. (If it doesn't, the group may 
well have other problems in any event.) 

There's a related question for both you and your 
players. How much do any of you mind when a particular
campaign's version of an answer diverges from
the stock one provided in print? There are two kinds 
of variation possible for this, and each one raises its 
own issues. 

There are matters that the game leaves unresolved, 
so that there is no single authoritative answer, like the 
number of TITANs in the solar system in the game's 
present moment. If you choose to give a specific 
number, it's your choice, and any number that seems 
to work for your campaign will probably do the job, 
%%% 390
whether it's one, three, seven, a dozen, or something 
else. Your campaign can't diverge from the baseline 
unless your answer is relatively extreme, like ``there 
are no TITANs, it was all a hoax before contact with 
the Exsurgent virus and then purely alien technology 
after that.'' In this case, your players can have read 
all the game's secrets and still be surprised by the revelation
you present. The potential for trouble here is
not a conflict of expectations based on the game, but 
based on expectations raised in other contexts. Some 
games, like some movies, TV shows, and other stories, 
develop a following with strong ideas of its own about 
what the real truths and important matters are, and if 
the following thrives, its members may end up with 
ideas that have less and less to do with the original 
inspiration. This isn't good or bad in itself, but it can 
be a problem, which is why it bears conscious consideration
and discussion, both before play and as the
campaign evolves. Ask your players to tell you about 
conversations and insights that shape their expectations
for the game world and storylines. Sometimes
you'll want to work those in with your own plans, 
sometimes you may want to deliberately play against 
them for the sake of a delightful surprise (generally 
more delightful for players than characters, but that's 
life as a character for you). In either case, it's better to 
be thinking about it than missing it. 

Then there are matters that the game does give 
definite answers for, but which you wish to change 
for the sake of your own campaign's characters and 
stories. This is perfectly fine. There are no game police 
roaming the countryside and forcing you to accept 
answers you'd prefer not to use. But your players 
will, as with the first question, have expectations, and 
your campaign will work better if you make sure you 
understand what those expectations are. How much 
\textit{would }it bother them if it turned out there were no 
TITANs and it was all a hoax, and so on? It's hard to 
guess what friends will say and impossible to predict 
the range of responses strangers might give, so ask 
them. (This particular answer is one that's unlikely to 
appear in anyone's campaign, but it makes a handy 
example for your conversational use precisely because 
it's extreme. So their answers to it are likely to be 
about the same as to any other potentially extreme 
change, and this one probably doesn't give away any 
of your own plans.) Some players are flexible on most 
matters but have particular points of attachment; 
if yours are among them, ask them to explain what 
those points are for them, so that you can keep them 
in mind. Other players have a hard time having fun 
with any major shift from published standard answers, 
and if you have players like that, you'll want to know 
it so that you can see how to adapt your plans to 
work within that framework. 

\subsubsection{That'S What I'M Talking About:}


\subsubsection{Shared Inspiration }

It's not quite true that everything changed from the 
early 21st century to \textit{Eclipse Phase's} universe, but a 
great many things did, and it can be hard to keep track 
of them all at once. This is where shared inspirations 
can come in handy. One striking illustration can convey 
a lot of details for both foreground and background, 
suggesting an aesthetic standard for design, an exotic 
environment, people doing futuristic tasks with appropriately
advanced tools, and so on. A prose passage
from a rewarding novel may set an ambiance or nail 
down some aspect of the characters' circumstances. 

There are potential pitfalls, and it's important to 
be aware of them. The greatest is obsolescence, the 
meaning of something evocative changing because 
the players' reality has changed since the inspiration 
entered it. William Gibson's ground-breaking cyberpunk
novel \textit{Neuromancer }begins, ``The sky above the
port was the color of television, tuned to a dead channel
Supporting details make it clear that this is an
industrial port at night, the sky gray from pollution 
and flecked with ash and other debris. But that was 
an image published in 1984. A decade and a half later, 
Neil Gaiman pointed out that to his children, the 
color of television tuned to a dead channel is bright 
blue, thanks to ubiquitous cable delivery. In another 
decade, the default color of a station not in use may 
be something else entirely. The moral is that it's not 
enough to agree that an image is very striking. You'll 
want to make sure that you all agree on what it is 
about it that's striking, to avoid a tangle of misconceptions
that could derail play later on.

The References page (p. 394) offers a wide range 
of immediately relevant inspirations, but it's not the 
final word on the subject. If the people in your group 
have a long-time favorite space scene, or description 
of life in the midst of a high-tech investigation, or 
poetic glimpse of what it might feel like to modify the 
body in ways not possible in real life, or something 
else that's stayed with them a long time and seems 
like it might bear on your campaign, encourage them 
to share. Remember to be courteous with each other's 
personal treasures, whether you end up using them 
or not; there's nothing like earned trust to encourage 
more sharing. 

Images can be particularly helpful for what they 
convey about the world behind and around the 
foreground events. For instance, think of a corridor 
on a typical spaceship or habitat in \textit{Eclipse Phase. }
Did you imagine it as being a standardized size and 
shape, so that its counterparts elsewhere would be 
very much the same or a more individualized work 
intended for use just where it is without concern 
for interchangeability? Did you imagine it as well 
lit even when not in use, lit well when sensors show 
people present and otherwise dim or dark, or perhaps 
planned to be well lit but in practice haphazard and 
unreliable thanks to lack of maintenance and funds? 
Did you imagine its surfaces smooth and clean, with 
%%% 391
equipment, maintenance bays, and the like all behind 
hatches and covers, or was it cluttered and lumpy? 
None of that matters all the time, but when it comes 
to the investigation of a derelict, the hunt for someone 
(or something) trying to hide, a race against time, or 
other dramatic complication, these things could affect 
your play, and rather than try to tally all possible contingencies
in advance, having some general-purpose
references can save everyone time and confusion. 

\subsubsection{Things That Should Not Be: Horror }

The universe of \textit{Eclipse Phase} is a time of horrors 
unleashed. Every character has had to come to some 
personal accommodation with the existence of things 
that offend our basic expectations of decency and 
practicality all at once. Horror comes in many flavors, 
and no one campaign can make use of all of them.

There are at least as many theories of horror as 
there are people who create horror stories. Everything 
here is necessarily a generalization. You and your 
players can find exceptions to every single point in 
it, and if you like the way those work better, go with 
them. This discussion is intended to trigger ideas, not 
to close off anything. That said, there are some useful 
generalizations about horror, starting with an insight 
expressed well by H.P. Lovecraft: ``The oldest and 
strongest emotion of mankind is fear, and the oldest 
and strongest kind of fear is fear of the unknown.'' All 
horror can be thought of as built around encounters 
with the unknown, beginning with the realization that 
there is something unknown present, learning something
about the scope of its nature and activities, and
then trying to respond one way or another. 

In this game, the discovery part is half over. There's 
no question about the presence of the unknown. 
Yes, there really are monsters beyond transhuman 
understanding loose in the universe, and everyone in 
the \textit{Eclipse Phase} universe knows how bad and how 
\textit{strange }the TITANs could get. Many people also have 
some idea of how exotic life on the far side of the Pandora
Gates can be. There's no room left for characters
to respond to some new strangeness with confident 
skepticism, sure that they know the range of what's 
possible and plausible within transhuman experience 
anymore. Almost anything might exist, given the 
facts of what's already known. Instead, the question 
for \textit{Eclipse Phase }people facing a mystery is whether 
this particular unknown will turn out to be simple 
and straightforward to deal with, more complicated 
but nonetheless a part of their routine lives like malfunctioning
machinery or a sabotaged and unusually
modified morph, or something beyond the normal like 
a TITAN-programmed weapon or alien life. Sooner or 
later, if they keep poking around, the characters can 
count on running into all sorts of unknown and maybe 
even unknowable challenges. Are they there yet?

Horror is seldom very far from humor. Humor 
serves many roles in human psychology, and one of 
them is helping us whittle down the mental ``size'' of 
mysteries and threats to something we can deal with. 
Furthermore, horror usually involves a balance of improbable
elements, with things lined up to go wrong
in interesting ways, and it doesn't take much for a 
particular rickety edifice to go from strange and menacing
to ludicrous. When your players start laughing,
sometimes the best thing for you to do is to roll with 
it. Laughter can do everyone good, supporting the 
``play'' part of roleplaying. In addition, some events 
actually are funny or at least can be taken as funny, 
even (sometimes especially) when most of what's going 
on is serious. On the other hand, if you really would 
like to keep a scene serious and the players break out 
in giggles, it's often wise to go ahead and take a break. 
Tell the players what you're doing, too; trying to deceive
a group of your friends isn't very reliable and
can backfire badly. Make the break long enough for 
everyone to get the giggles out and then continue.
At the end of the day, through communication with 
their players, the gamemaster will know how much 
horror their group wishes to encounter. A group may 
decide that they want to be 100 percent immersed 
into the various horrors of \textit{Eclipse Phase.} Another 
group, however, may decide that while they enjoy 
the meshed theme of horror with the other aspects 
of \textit{Eclipse Phase,} they don't wish it to be a principal 
element. In such a situation, horror would remain just 
that, a theme, while the plots woven by the gamemaster
would spin around the myriad of other elements
that make up the game.

\subsubsection{Transhumanism}

Humanity has embraced transhumanism for survival, 
harnessing science and technologies to catapult physical
and mental faculties to super-human levels, while
eradicating involuntary death and enabling near immortality
through the digitization of consciousness
and the ability to transfer bodies at will. This is one of 
the cornerstone themes of \textit{Eclipse Phase.}
The technologies inherent to a transhuman future 
raise many questions and ethical issues, however, and 
these are some of the central themes that \textit{Eclipse }
\textit{Phase} seeks to explore. We encourage both gamemaster
and players to play around with the possibilities
and contradictions enabled in such a universe. How 
do our mindsets change when death no longer looms 
over us? What does identity mean when our bodies 
are disposable and our personalities can be edited? 
Are we the same person when we are revived from 
a backup, or sent off as a fork? Are technologies like 
nanofabrication something to be feared and restricted, 
even when they can eliminate poverty and greed? 
How do we ensure public safety in a world where 
technology makes weapons of mass destruction easily 
available? How do ideas inherent to religious and 
spiritual thought cope with AI, backups, or resleeving
What does it mean to be an uplifted animal in a
society centered on humans? Who decides our future? 
These are just a few of the issues that \textit{Eclipse Phase}
raises, and many of them can be used as the central 
theme for an entire campaign. 

ke

ed,

d?

e

y 
