%%% 392
1. Define Character Concept (p. 130)
2. Choose Background (p. 131)
3. Choose Faction (p. 132)
4. Spend Free Points (p. 134)
a) 105 aptitude points
b) 1 Moxie
c) 5,000 credit
d) 50 Rep
e) Native tongue
5. Spend Customization Points (p. 135)
a) 1,000 CP to spend
 

15 CP = 1 Moxie
 

10 CP = 1 aptitude point
 

5 CP = 1 psi sleight
 

5 CP = 1 specialization
 

2 CP = 1 skill point (61-80)
 

1 CP = 1 skill point (up to 60)
 

1 CP = 1,000 credit
 

1 CP = 10 rep
b) Active skill minimum: 400 CP
c) Knowledge skill minimum: 300 CP
d) Choose Starting Morph (pp. 136 and 139)
e) Choose Traits (pp. 136 and 145)
6. Purchase Gear (p. 136)
7. Choose Motivation (p. 137) 
8. Calculate Remaining Stats (p. 138)
9. Detail the Character (p. 138)
%%% 393
%%% 394
•  Combat is handled as an Opposed Test.
•  Attacker rolls attack skill +/– modifiers.
•  Melee: Defender rolls Fray or melee skill 
+/– modifiers.
•  Ranged: Defender rolls (Fray skill ÷ 2, 
round down) +/– modifiers.
•  If attacker succeeds and rolls higher than 
the defender, the attack hits.
•  Critical hits are armor-defeating (armor 
does not apply).
•  Armor is reduced by the attack's Armor 
Penetration value (AP).
•  The weapon's damage is reduced by the 
target's modified Armor rating (unless 
the attack is armor-defeating).
•  If the damage exceeds the target's 
Wound Threshold, a wound is also scored. 
(If the damage exceeds the Wound 
Threshold by multiple factors, multiple 
wounds are inﬂ icted.)
Step 1: Roll Initiative ((INT + REF) x 2) + 1d100
Step 2: Begin First Action Phase (Speed 1)
Step 3: Declare and Resolve Actions
Step 4: Rotate and Repeat (Speed 2-4)
%%% 395
\textbf{TASKS RESULTS}
1. Defeat the Firewall 
2. Bypass Active Security 
  a. Hacker Wins with Excellent Success, Defender Fails 
  b. Hacker Succeeds, Defender Fails 
  c. Both Succeed 
  d. Defender Succeeds, Hacker Fails 
1. Common data = automatic acquisition
2. Uncommon data:
  a. Research Task Test (timeframe: 1 minute) modified by 

data obscurity to accumulate data
  b. Measure of Success determines depth of data found
3. Analyzing data:
  a. Research Task Test (timeframe: GM call) using 

complementary skill to understand data
Infosec Task Action (10 minutes)
Opposed Infosec Test
Hidden status/admin privileges/+30 all Subversion Tests (p. 256)
Covert Status (p. 256)
Spotted status/passive alert/-10 all Subversion Tests (p. 256)
Locked status/active alert/-20 all Subversion Tests (p. 256) 
